---
title: "iNEXT_EBS_Richness"
output: html_notebook
---
I will plot how richness changes over time but will use iNEXT to calculate richness

```{r setup}
library(data.table)
library(iNEXT)
```

Abundance Data
```{r abundance data}
ebs_data <- readRDS("ebs_data_gridded.RData")

head(ebs_data)
```
First, we will plot richness for the entire region (ignoring cells)

Format as data frame with species as rows and years as columns , "wtcpue - numeric - the weight in kilograms captured per unit effort"

```{r sum across years}
ebs_data_year <- ebs_data[,wtcpue_sum := sum(wtcpue, na.rm = T), .(matched_name2, year)][,wtcpue_sum_round := round(wtcpue_sum, 0)]

ebs_data_year.r <- unique(ebs_data_year[,.(matched_name2, year, wtcpue_sum_round)])
```
I want to calculate richness for each individual year (site)
```{r all years x abundances}

#list of years 
ebs_years <- unique(ebs_data_year.r[,year])
sort(ebs_years)

#long to wide
ebs_wide <- dcast(ebs_data_year.r, matched_name2 ~ year, value.var = "wtcpue_sum_round", fill = 0)

```

Apply iNEXT. iNEXT will literally only take data frames, so, I will make it into a data frame first.
```{r applying iNEXT to overall richness}
ebs_wide.df <- data.frame(ebs_wide)

ebs_overall_biomass_iNEXT_richness <- iNEXT(ebs_wide.df[,2:length(ebs_wide.df)], q = 0, datatype = "abundance")
saveRDS(ebs_overall_biomass_iNEXT_richness, "ebs_overall_biomass_iNEXT_richness.RData")

ebs_overall_biomass_iNEXT_shannon <- iNEXT(ebs_wide.df[,2:length(ebs_wide.df)], q = 1, datatype = "abundance") #how to deal with the fact abundance should be a round number
saveRDS(ebs_overall_biomass_iNEXT_shannon, "ebs_overall_biomass_iNEXT_shannon.RData")


ebs_overall_biomass_iNEXT_simpson <- iNEXT(ebs_wide.df[,2:length(ebs_wide.df)], q = 2, datatype = "abundance")
saveRDS(ebs_overall_biomass_iNEXT_simpson, "ebs_overall_biomass_iNEXT_simpson.RData")
# to assess sample completeness (this takes a bit), iNEXT focuses on three measures of Hill numbers of order q: species richness (q = 0), Shannon diversity (q = 1, the exponential of Shannon entropy) and Simpson diversity (q = 2, the inverse of Simpson concentration), I should look at all three

ebs_overall_iNEXT_min_coverage_richness <- min(ebs_overall_biomass_iNEXT_richness$DataInfo$SC)
ebs_overall_iNEXT_min_coverage_shannon <- min(ebs_overall_biomass_iNEXT_shannon$DataInfo$SC)
ebs_overall_iNEXT_min_coverage_simpson <- min(ebs_overall_biomass_iNEXT_simpson$DataInfo$SC)

#minimum sample coverage across sites is 0.9998, so this will be my SC value

#this step below takes a long time as well... keep in mind

ebs_overall_diversity_richness <- estimateD(ebs_wide.df[,2:length(ebs_wide.df)], datatype = "abundance", base = "coverage", level = ebs_overall_iNEXT_min_coverage_richness) # compare point diversities for specified completeness

saveRDS(ebs_overall_diversity_richness, file = "ebs_overall_diversity_richness.RData")

ebs_overall_diversity_shannon <- estimateD(ebs_wide.df[,2:length(ebs_wide.df)], datatype = "abundance", base = "coverage", level = ebs_overall_iNEXT_min_coverage_shannon) # compare point diversities for specified completeness

saveRDS(ebs_overall_diversity_shannon, file = "ebs_overall_diversity_shannon.RData")

ebs_overall_diversity_simpson <- estimateD(ebs_wide.df[,2:length(ebs_wide.df)], datatype = "abundance", base = "coverage", level = ebs_overall_iNEXT_min_coverage_simpson) # compare point diversities for specified completeness

saveRDS(ebs_overall_diversity_simpson, file = "ebs_overall_diversity_simpson.RData")
#only thing that would change above is coverage estimates (q = 0-2)

#convert to data table to make easier to manipulate
ebs_overall_diversity_richness.dt <- data.table(ebs_overall_diversity_richness)
ebs_overall_diversity_shannon.dt <- data.table(ebs_overall_diversity_shannon)
ebs_overall_diversity_simpson.dt <- data.table(ebs_overall_diversity_simpson)

ebs_overall_diversity_richness.dt[,year := as.numeric(substr(site, 2, 5))][,metric_type := "richness"]
ebs_overall_diversity_shannon.dt[,year := as.numeric(substr(site, 2, 5))][,metric_type := "shannon"]
ebs_overall_diversity_simpson.dt[,year := as.numeric(substr(site, 2, 5))][,metric_type := "simpson"]

#... turns out it calculates all three metrics (hill numbers) for you anyway... no need to do three times, cool bro

#rarefied richness overtime plot
ggplot(aes(x = year, y = qD), data = ebs_overall_diversity_richness.dt[order == 0,]) +
  geom_smooth(aes(ymin = qD.LCL, ymax = qD.UCL), stat = "identity", color = "black") +
  labs(x= "Year", y = "Estimated Richness") +
  theme_classic() +
  theme(text = element_text(size=20))

#simpson and shannon diversity plot
ggplot(aes(x = year, y = qD, linetype = as.factor(order)), data = ebs_overall_diversity_richness.dt[order != 0,]) +
  geom_smooth(aes(ymin = qD.LCL, ymax = qD.UCL), stat = "identity", color = "black", alpha = 3) +
  labs(x= "Year", y = "Diversity Metric") +
  theme_classic() +
  theme(text = element_text(size=20))


```