---
title: "1. Setup NE Trawl Data"
output: html_notebook
---

Instructions from Malin:

Is anyone up for an exploratory analysis with the trawl data? It would be very cool to look at changes in spatial turnover, similar to [Magurran, Dornelas, Moyes, Gotelli, McGill et al. 2015](http://www.nature.com/ncomms/2015/150924/ncomms9405/full/ncomms9405.html) but across a much wider set of regions. An idea from conversations at [HIFMB](https://hifmb.de/en/) yesterday. 

From Magurran et al. *"Here, we analyse an exceptionally comprehensive 29-year time series of North Atlantic groundfish assemblages monitored over 5° latitude to the west of Scotland. These fish assemblages show no systematic change in species richness through time, but steady change in species composition, leading to an increase in spatial homogenization: the species identity of colder northern localities increasingly resembles that of warmer southern localities."*

Methods from Magurran et al. 2015

* First assign the rectangles to nine 30' latitudinal bands
* Compile community time series for each latitudinal band
* Sample rarefaction ensures equal sampling effort across bands and is used in the calculation of temporal $\alpha$ diversity and temporal $\beta$ diversity. 
* First calculate dissimilarity in relation to the start of the survey
* Next, for each year, we compute pairwise compositional dissimilarity of these latitudinal bands
* We also construct distance-decay plots for each year

```{r setup}
library(data.table)
library(vegan)
library(ggpubr)
library(betapart)
```

## Download data from Ocean Adapt
* [NEUS Strata](https://github.com/pinskylab/OceanAdapt/blob/master/data_raw/neus_strata.csv)
* [NEUS SVPP](https://github.com/pinskylab/OceanAdapt/blob/master/data_raw/neus_SVSPP.RData)
* [NEUS Survey Data](https://github.com/pinskylab/OceanAdapt/blob/master/data_raw/neus_Survdat.RData)

```{r import data for NEUS from Ocean Adapt on GitHub}

neus_strata <- fread('https://raw.githubusercontent.com/pinskylab/OceanAdapt/master/data_raw/neus_strata.csv')
setnames(neus_strata, c('stratum', 'stratum_name'), c('STRATUM', 'STRATUM_NAME'))
load(url('https://raw.githubusercontent.com/pinskylab/OceanAdapt/master/data_raw/neus_Survdat.RData')) # loads survdat
load(url('https://raw.githubusercontent.com/pinskylab/OceanAdapt/master/data_raw/neus_SVSPP.RData')) # load spp

```

### Link data tables together
```{r link data tables together}
# merge
stratamerge <- merge(survdat[, .(CRUISE6, STATION, STRATUM, TOW, SVSPP, YEAR, SEASON, LAT, LON, BIOMASS)], 
                     neus_strata[, .(STRATUM, STRATUM_NAME)], by = 'STRATUM') #merge strata with survey data
neus_full <- merge(stratamerge, spp[, .(SVSPP, SCINAME, COMNAME)], by = "SVSPP") #merge species with strata and survey data

```

### Summarize and clean up data
Make sure every observation is of an identifiable fish to species
Sum across multiple rows of the same species (e.g., different length classes or sexes in NEFSC surveys)
```{r clean up}
#only want fish that ID to species, therefore must a string with a space
#first remove any floating white space before or after words
neus_full <- neus_full[,SCINAME := trimws(SCINAME)]
#keep only rows with space in SCINAME column (signifying genus and species are identified)
neus_full.r <- neus_full[grep(" ", SCINAME),]
#get rid of any rows with TRASH SPECIES IN CATCH or UNIDENTIFIED FISH as SCINAME
neus_full.ID <- neus_full.r[SCINAME != "UNIDENTIFIED FISH" & SCINAME != "TRASH SPECIES IN CATCH",]
#sum biomass across cruise, station, stratum, tow, species, year, season, and location
neus_full.ID <- neus_full.ID[ , .(BIOMASS = sum(BIOMASS)), by = .(CRUISE6, STATION, STRATUM, STRATUM_NAME,TOW, SVSPP, SCINAME, YEAR, SEASON, LAT, LON)]

```

### Save out
```{r save}
save(neus_full.ID, file = "data/NE_trawl_data_full.Rdata")
```

