---
title: "3. Spatial Temporal Diversity By Strata"
output: html_notebook
---
Here, we will do a similar analysis but by strata instead of subregion
```{r setup}
library(data.table)
library(ggplot2)
library(geosphere)
library(betapart)
```

Data
```{r data}
load("NE_trawl_data_full.Rdata")
```

#Distance Decay with Stratum instead of region
Jaccard Dissimilarities Distance Decay all in one plot by STRATA
```{r stratum distance matrix}
neus_full.ID.uniquestrata <- unique(neus_full.ID, by = "STRATUM") #unique strata

neus_full.ID.uniquestrata.latlon <- neus_full.ID.uniquestrata[,c("LAT", "LON", "STRATUM")] #only lat lon and stratum

stratum_distances <- distm(neus_full.ID.uniquestrata.latlon[,c(2,1)]) #distance matrix, consistent across years
rownames(stratum_distances) <- neus_full.ID.uniquestrata.latlon$STRATUM
colnames(stratum_distances) <- neus_full.ID.uniquestrata.latlon$STRATUM

#reorient to long form 
stratum_distance.l <- melt(as.matrix(stratum_distances), varnames = c("strat1", "strat2")) #matrix to data frame, 
stratum_distance.l <- data.table(stratum_distance.l) #and then to data table
stratum_distance.l <- stratum_distance.l[strat1 >= strat2,] # to get rid of repetitions
colnames(stratum_distance.l)[colnames(stratum_distance.l)=="value"] <- "distance(m)" #change column name to distance

```
Above is distance info, now I need to build community matrix
```{r distance decay by strata}

#one PRESENT value for each YEAR, SEASON, STRATUM and SVSPP combination
neus_full.ID.stratum.unique <- na.omit(unique(neus_full.ID, by = c("YEAR", "STRATUM", "SVSPP")), cols = c("YEAR", "STRATUM", "SVSPP")) #unique STRATUM, year, species combinations, and get rid of any rows where any are NA

#reduce data table to just be svspp, year, sciname, comname, STRATUM

neus_full.ID.stratum.unique.r <- data.table((neus_full.ID.stratum.unique[, c("SVSPP", "SCINAME", "COMNAME", "YEAR", "STRATUM")]))

#add presence column 
neus_full.ID.stratum.unique.r[,"present" := 1]

#now I need to populate a comparison data table with every possible year, region, and species combination to add 0s
stratum <- as.integer(unique(neus_full.ID.stratum.unique.r$STRATUM)) 



full_set_stratum<- data.table(expand.grid("SVSPP" = spp.list, "YEAR" = years, "STRATUM" = stratum))

```
Bring in presences to expansion for presence by STRATUM
```{r bring in actual presences by STRATUM to expansion}
merge <- neus_full.ID.stratum.unique.r[full_set_stratum, on = c("YEAR", "STRATUM", "SVSPP")]
merge[,"present" := ifelse(is.na(SCINAME), 0,1)][,"SCINAME" := NULL][,"COMNAME" := NULL] #if scientific name and common name are empty, that is an absence, if populated, presence, also, we don't need these columns, as we'll bring them in below

#simplify spp key
#bring in spp key
spp_key

neus_pres_abs_STRATUM <- merge[spp_key, on = c("SVSPP")] #also don't actually need all columns

```

Make distance decay model for each year by loop for all STRATUM, YEAR, SVSPP combos. I've commented out lines that attempt to make a distance decay model because of error. Instead, I will just use this loop to populate a data table with distance and dissimilarities between all strata. I attempt to use decay.model() function from betapart package. 
--> I can't get this to work
Error: *cannot find valid starting values: please specify some*


```{r jaccard disdissimilarities between STRATUM in multiple years}
distances_dissimilarities_allyears_STRATUM <- data.table("strat1" = integer(), "strat2" = integer(), "distance(m)" = numeric(),"jaccard_dissimilarity" = numeric(), YEAR = integer())

#Now loop through all years starting in 1967
for (i in 1:length(years)) {
  reduced_year <- neus_pres_abs_STRATUM[YEAR == years[i],]
  reduced_year_wide <- dcast(reduced_year, YEAR + STRATUM ~ SCINAME, value.var = "present") #long to wide data for community matrix, column names are SVSPP
  
  communitymatrix <- cbind(reduced_year_wide[,c(3:819)], reduced_year_wide[,c(1:2)]) #community matrix
  
#there are some STRATA with no presences or observations
#rowSums(communitymatrix.r[,c(1:817)]) #I guess I will get rid of those

communitymatrix.r <- communitymatrix[as.logical(rowSums(communitymatrix[,c(1:817)] != 0)), ] #reduce to only rows with atleast one species present
stratum.r <- communitymatrix.r$STRATUM

  
  dissimilarities <- vegdist(communitymatrix.r[,c(1:817)], method = "jaccard", binary = T) #dissimilarity 
  

#make into matrix
dissimilarities.m <- as.matrix(dissimilarities, labels=TRUE)
colnames(dissimilarities.m) <- rownames(dissimilarities.m) <- stratum.r

  #reshape dissimilarities
  dissimilarities.l <- melt(dissimilarities.m, varnames = c("strat1", "strat2"))
  dissimilarities.l <- data.table(dissimilarities.l) #and then to data table
  dissimilarities.l <- dissimilarities.l[strat1 >= strat2,] # to get rid of repetitions
  colnames(dissimilarities.l)[colnames(dissimilarities.l)=="value"] <- "jaccard_dissimilarity" #change column name to distance


#add year for these values
  dissimilarities.l[, "YEAR" := years[i]]

#merge distance with jaccard_dissimilarity for this year
  dissimilarities.l.full <- stratum_distance.l[dissimilarities.l, on = c("strat1", "strat2")]

distances_dissimilarities_allyears_STRATUM <- rbind(distances_dissimilarities_allyears_STRATUM, dissimilarities.l.full) #add to data table

#extract info for distance decay exponential model for each year
#reduce stratum_distances to stratum.r
stratum_distances.r <- stratum_distances[as.character(stratum.r), as.character(stratum.r)]

#modname <- paste0(years[i], "_decay_mod_exponential")

#assign(modname, decay.model(dissimilarities.m, stratum_distances.r, model.type = "exponential", y.type = "dissimilarities", perm = 100))

#jpeg(file=paste0("jaccardplot_", years[i], ".jpg"))
    #plot(x=dissimilarities.l.full$`distance(m)`, y=dissimilarities.l.full$jaccard_dissimilarity, xlab = "Distance(m)", #ylab="Jaccard dissimilarity", ylim = c(0:1), pch = 19)
 #dev.off()
 
 #PCA
#vare.pca <- rda(communitymatrix[,c(1:817)])
#jpeg(file=paste0("jaccardpca_", years[i], ".jpg"))
#plot(vare.pca, display = "sites", xlim = c(-4,4), ylim = c(-4,4))
#dev.off()
  
}

summary(distances_dissimilarities_allyears_STRATUM) #here we have jaccard, distance, and 

save(distances_dissimilarities_allyears, file = "jaccard_similarities_distances_4reg.Rdata")
save(distances_dissimilarities_allyears_STRATUM, file = "jaccard_similarities_distances_stratum.Rdata")

```

Instead, I will look at distance decay through time visually

```{r distance decay all years}
#order data table by year
setorder(distances_dissimilarities_allyears_STRATUM, by = YEAR)

distance_decay_stratum <- ggplot(data = distances_dissimilarities_allyears_STRATUM, aes(x = `distance(m)`, y = jaccard_dissimilarity, color = YEAR)) +
  geom_point() +
  theme_classic() +
  labs(x = "Distance (m)", y = "Jaccard Dissimilarity") +
  theme(text=element_text(size = 20))
distance_decay_stratum


ggsave(distance_decay_stratum, path = "plots/", file = "distance_decay_stratum.eps")


```

Extract coefficient (slope) for each year for simple exponential, how does it change through time? I did this because I couldn't get betapart decay.model to work.
```{r each year build linear model}
#I'm interested in the rate of decrease
#make new column for 1-jaccard dissimilarity
distances_dissimilarities_allyears_STRATUM[,jaccard_similarity := 1-jaccard_dissimilarity]

#reduce to one year
expmod_coefs <- data.table("year" = years, "beta.0" = 0)
for (i in 1:length(years)) {
  reduced <- distances_dissimilarities_allyears_STRATUM[YEAR == years[i],]
  
  # Select an approximate $\theta$, since theta must be lower than min(y), and greater than zero
theta.0 <- min(reduced[jaccard_similarity != 0,]$jaccard_similarity) * 0.5

# Estimate the rest parameters using a linear model
model.0 <- lm(log(jaccard_similarity - theta.0) ~ `distance(m)`, data=reduced)  
alpha.0 <- exp(coef(model.0)[1])
beta.0 <- coef(model.0)[[2]]

expmod_coefs[i,2] <- -(beta.0)

}

plot(expmod_coefs$year, expmod_coefs$beta.0)
abline(lm(expmod_coefs$beta.0~expmod_coefs$year))

#decreasing in rate of decay through time

```


Below, I continue to attempt to make distance decay from betapart package work, no luck yet. Some years work, others don't

Fitting distance decay models to pair-wise assemblage similarity: fits a negative exponential or power law function(GLM) describing decay of assemblage similarity or increase of dissimilarity with distance.


```{r distance decay model}

#haven't figured out error here, same error that appears above

#reduce stratum_distances to stratum.r
stratum_distances.r <- stratum_distances[as.character(stratum.r), as.character(stratum.r)]

mod_1968 <- decay.model(dissimilarities.m, stratum_distances.r, model.type = "exponential", y.type = "dissimilarities", perm = 1000)


```