---
title: "Sample and Coverage Rarefaction"
output: html_notebook
---

Sample and coverage based rarefaction (will look more into later)
```{r setup}
library(data.table)
library(iNEXT)
library(vegan)
library(ggplot2)
library(tidyverse)
library(vegan)
library(betaC) #devtools::install_github("T-Engel/betaC")
library(cowplot)
theme_set(theme_cowplot())

neus_data <- readRDS("neus_data_gridded.RData")
neus_distances.l <- readRDS("neus_reg_distances.l.RData")
```

gamma diversity scale: all grid cells
alpha diversity scale: one grid cell
beta diversity: between grid cells


Using betaC from https://github.com/T-Engel/betaC

"To achieve this, we use a combination of individual-based and coverage-based rarefaction (coverage is a measure of sample completeness). The main idea of our approach is that within a species pool alpha and gamma scale diversity estimates are standardized to a common number of individuals, while across species pools we allow the sample size to vary in order to keep a constant gamma-scale sample sample coverage (C) instead."


smallest C_target() output of all the communities

Setup site by species matrix for 1982
```{r site by species NEUS}
neus_summed <- neus_data[,.(wtcpue_sum = sum(wtcpue)), .(cell, year, matched_name2)]

#year by year in a list
setorder(neus_summed, year)
years <- unique(neus_summed$year)

neus_summed_spread_by_year <- list()

for (i in 1:length(years)) {
  this_year <- years[i]
  subset <- neus_summed[year == this_year,]
  spreaded_sites_spp <- spread(data = subset, key = matched_name2, value = wtcpue_sum, fill = 0)
  neus_summed_spread_by_year[[i]] <- spreaded_sites_spp
}


```
Calculate the recommended maximum coverage value for the computation of beta_C from a site by species matrix
```{r coverage value}
neus_C_targets <-c()
for (i in 1:length(neus_summed_spread_by_year)) {
  
neus_C_targets[i] <- C_target(neus_summed_spread_by_year[[i]]) 
  
}

neus_min_C_target <- min(neus_C_targets)
```

Coverage-based rarefaction to quantify non-random component in beta-diversity
```{r non-random beta}
neus_beta_C <-c()
for (i in 1:length(neus_summed_spread_by_year)) {
  
  neus_beta_C[i] <- beta_C(neus_summed_spread_by_year[[i]], C = neus_min_C_target)
  
}

neus_beta_C.dt <- data.table(year = years, beta_C = neus_beta_C)


```
Methods from Prelimns
```{r jaccard dissimilarity}
distances_dissimilarities_allyears_neusgridcell <- data.table("cell1" = integer(), "cell2" = integer(), "distance(m)" = numeric(),"jaccard_dissimilarity" = numeric(), year = integer())

#Now loop through all years starting in 1982
for (i in 1:length(years)) {
  reduced_year <- neus_summed[year == years[i],]
  reduced_year_wide <- dcast(reduced_year, year + cell ~ matched_name2, value.var = "ABUNDANCE_sum", fill = 0) #long to wide data for community matrix, column names are year, cell, then scientific 
  
  ncols <- ncol(reduced_year_wide)
  communitymatrix <- cbind(reduced_year_wide[,3:ncols], reduced_year_wide[,1:2]) #community matrix with year and STRATUM on far right
  
  #there are some STRATA with no presences or observations
  #rowSums(communitymatrix.r[,c(1:817)]) #I guess I will get rid of those

  communitymatrix.r <- communitymatrix[as.logical(rowSums(communitymatrix[,1:(ncols - 2)] != 0)), ] #reduce to only rows with atleast one species present
  cell.r <- communitymatrix.r$cell

  dissimilarities <- vegdist(communitymatrix.r[,1:(ncols-2)], method = "jaccard", binary = T) #dissimilarity 
  #?vegdist

  #make into matrix
  dissimilarities.m <- as.matrix(dissimilarities, labels=TRUE)
  colnames(dissimilarities.m) <- rownames(dissimilarities.m) <- cell.r

  #reshape dissimilarities
  dissimilarities.l <- reshape2::melt(dissimilarities.m, varnames = c("cell1", "cell2"), value.name = "jaccard_dissimilarity")
  dissimilarities.l <- data.table(dissimilarities.l) #and then to data table
  dissimilarities.l <- dissimilarities.l[cell1 > cell2,] # to get rid of repetitions and self-comparisons


  #add year for these values
  dissimilarities.l[, "year" := years[i]]

  #merge distance with jaccard_dissimilarity for this year
  dissimilarities.l.full <- neus_reg_distances.l[dissimilarities.l, on = c("cell1", "cell2")]

  #add to data table
  distances_dissimilarities_allyears_neusgridcell <- rbind(distances_dissimilarities_allyears_neusgridcell, dissimilarities.l.full)

  #extract info for distance decay exponential model for each year
  #reduce stratum_distances to stratum.r
  #stratum_distances.r <- stratum_distances[as.character(stratum.r), as.character(stratum.r)]

  #modname <- paste0(years[i], "_decay_mod_exponential")

  #assign(modname, decay.model(dissimilarities.m, stratum_distances.r, model.type = "exponential", y.type = "dissimilarities", perm = 100))

  #jpeg(file=paste0("jaccardplot_", years[i], ".jpg"))
  #plot(x=dissimilarities.l.full$`distance(m)`, y=dissimilarities.l.full$jaccard_dissimilarity, xlab = "Distance(m)", #ylab="Jaccard dissimilarity", ylim = c(0:1), pch = 19)
  #dev.off()
 
  #PCA
  #vare.pca <- rda(communitymatrix[,c(1:817)])
  #jpeg(file=paste0("jaccardpca_", years[i], ".jpg"))
  #plot(vare.pca, display = "sites", xlim = c(-4,4), ylim = c(-4,4))
  #dev.off()
  
}
```


Extract coefficient (slope) for each year: how does it change through time? I did this because I couldn't get betapart decay.model to work.
```{r coef per year to look at distance decay}
mod_coefs <- data.table("year" = years, "beta" = 0)
for (i in 1:length(years)) {
  # Estimate the parameters using a linear model for this year
  mod <- lm(I(1 - jaccard_dissimilarity) ~ `distance(m)`, data = distances_dissimilarities_allyears_neusgridcell[year == years[i],])  # I() allows - to be used arithmetically, here we get similarity versus distance models
  mod_coefs[i, beta :=  -coef(mod)[2]] #change in similarity for change in distance

}

ggplot(aes(x = year, y = (beta)), data = mod_coefs) +
  geom_point() +
  labs(x = "Year", y = "Slope of Distance Decay Plot\n(ΔSimilarity/ΔDistance)") +
  geom_smooth() +
  theme_classic() +
  theme(text = element_text(size = 13))

```
Using BetaPart Package

The function returns a list with the three multiple site dissimilarity values.

For index.family="bray" the three indices are:

* beta.BRAY.BAL: value of the balanced variation component of Bray-Curtis multiple-site dissimilarity (individuals of some species in one site are substituted by the same number of individuals of different species in another site) (replacement)

* beta.BRAY.GRA: value of the abundance-gradient component of Bray-Curtis multiple-site dissimilarity (some individuals are lost from one site to the other) (nestedness)

* beta.BRAY: value of the overall dissimilarity, measured as Bray-Curtis multiple-site dissimilarity

Multi-site Dissimilarity: The number of co‐occurrences in three, four, and five (and so on) sites has an obvious impact on the total amount of dissimilarity in a set of multiple sites (or times, or, more generally, units). Therefore, averaging pairwise dissimilarity measures across multiple pairs of sites is a suboptimal approach for estimating the overall dissimilarity among more than three sites, that is, multiple‐site dissimilarity (Diserud & Ødegaard 2007; Baselga 2013a).Baselga

 2016

1982 first
```{r betapart}
# get betapart objects
reduced_year_wide.bp <- betapart.core.abund(reduced_year_wide[,2:226])
# multiple site measures
reduced_year_wide.bp.multi <- beta.multi.abund(reduced_year_wide.bp)

# sampling across equal sites
reduced_year_wide.sample <- beta.sample.abund(reduced_year_wide, sites = nrow(reduced_year_wide), samples = 100)

# plotting the distributions of components
plot(density(reduced_year_wide.sample$sampled.values$beta.BRAY),xlab='Beta
diversity', main='', lwd=4) #full Bray curtis dissimilarity
lines(density(reduced_year_wide.sample$sampled.values$beta.BRAY.BAL), lty=2, lwd=2, col = "red") #balanced
lines(density(reduced_year_wide.sample$sampled.values$beta.BRAY.GRA), lty=3, lwd=2, col = "blue") #gradient

#loop betapart


```

For index.family="jaccard" the three indices are:
beta.JTU value of the turnover component, measured as turnover fraction of Jaccard dissimilarity
beta.JNE value of the nestedness component, measured as nestedness-resultant fraction of
Jaccard dissimilarity
beta.JAC value of the overall beta diversity, measured as Jaccard dissimilarity

Overall dissimilarity using betapart Multiple Site Diversity Metrics
```{r overall jaccard and bray dissimilarity}
dissimilarities_allyears_neusgridcell <- as.data.table(matrix(nrow = length(years)))
dissimilarities_allyears_neusgridcell[,year := as.numeric(V1)][,bray_dissimilarity := as.numeric(V1)][,bray_dissimilarity_balanced := as.numeric(V1)][,bray_dissimilarity_gradient := as.numeric(V1)][,jaccard_dissimilarity := as.numeric(V1)][,jaccard_dissimilarity_turnover := as.numeric(V1)][,jaccard_dissimilarity_nestedness := as.numeric(V1)][,V1 := NULL]

#Now loop through all years starting in 1982
for (i in 1:length(years)) {
  
  dissimilarities_allyears_neusgridcell[i,year := years[i]]
  
  reduced_year <- neus_summed[year == years[i],][,presence := ifelse(wtcpue_sum >0, 1, 0)]
  
  reduced_year_wide_abundance <- dcast(reduced_year, year + cell ~ matched_name2, value.var = "wtcpue_sum", fill = 0) #long to wide data for community matrix, column names are year, cell, then scientific 
  
  reduced_year_wide_abundance.bp <- betapart.core.abund(reduced_year_wide_abundance[,3:ncol(reduced_year_wide_abundance)])
  
  # sampling across equal sites
reduced_year_wide_abundance.sample <- beta.sample.abund(reduced_year_wide_abundance[,3:ncol(reduced_year_wide_abundance)], sites = nrow(reduced_year_wide_abundance), samples = 100)

  #combined bray dissimilarity
  dissimilarities_allyears_neusgridcell[i, bray_dissimilarity := reduced_year_wide_abundance.sample$mean.values[[3]]]
  
  #balanced
  dissimilarities_allyears_neusgridcell[i, bray_dissimilarity_balanced := reduced_year_wide_abundance.sample$mean.values[[1]]]
  
  #gradients
  dissimilarities_allyears_neusgridcell[i, bray_dissimilarity_gradient := reduced_year_wide_abundance.sample$mean.values[[2]]]
  
  #AGAIN FOR OCCURRENCE
  reduced_year_wide_occurrence <- dcast(reduced_year, year + cell ~ matched_name2, value.var = "presence", fill = 0) #long to wide data for community matrix, column names are year, cell, then scientific
  
    reduced_year_wide_occurrence.bp <- betapart.core(reduced_year_wide_occurrence[,3:ncol(reduced_year_wide_occurrence)])
  
  # sampling across equal sites
reduced_year_wide_occurrence.sample <- beta.sample(reduced_year_wide_occurrence[,3:ncol(reduced_year_wide_occurrence)], sites = nrow(reduced_year_wide_occurrence), samples = 100, index.family = "jac")

  #combined jaccard dissimilarity
  dissimilarities_allyears_neusgridcell[i, jaccard_dissimilarity := reduced_year_wide_occurrence.sample$mean.values[[3]]]
  
  #balanced
  dissimilarities_allyears_neusgridcell[i, jaccard_dissimilarity_turnover := reduced_year_wide_occurrence.sample$mean.values[[1]]]
  
  #gradients
  dissimilarities_allyears_neusgridcell[i, jaccard_dissimilarity_nestedness := reduced_year_wide_occurrence.sample$mean.values[[2]]]
  
}

saveRDS(dissimilarities_allyears_neusgridcell, "dissimilarities_allyears_neusgridcell.RData")

#most of beta diversity derived from turnover
ggplot(data = dissimilarities_allyears_neusgridcell, aes(year)) +
  geom_line(aes(y = 1-bray_dissimilarity), color = "purple", linetype = 2) +
  geom_line(aes(y = 1-bray_dissimilarity_balanced), color = "darkgreen", linetype = 2) +
  geom_line(aes(y = 1-jaccard_dissimilarity), color = "purple", linetype = 1) +
  geom_line(aes(y = 1-jaccard_dissimilarity_turnover), color = "darkgreen", linetype = 1) +
  #lims(y= c(0.04,.2)) +
  labs(y = "Similarity", x = "Year") +
  theme_classic()


```
Now, same analyses but for occurrence instead of abundance


beta.SIM	
value of the turnover component, measured as Simpson dissimilarity

beta.SNE	
value of the nestedness component, measured as nestedness-resultant fraction of Sorensen dissimilarity

beta.SOR	
value of the overall beta diversity, measured as Sorensen dissimilarity

Overall dissimilarity using betapart Multiple Site Diversity Metrics
```{r overall Sorensen dissimilarity}
sorensen_dissimilarities_allyears_neusgridcell <- as.data.table(matrix(nrow = length(years)))
sorensen_dissimilarities_allyears_neusgridcell[,year := as.numeric(V1)][,sorensen_dissimilarity := as.numeric(V1)][,simpson_dissimilarity_turnover := as.numeric(V1)][,sorensen_dissimilarity_nestedness := as.numeric(V1)][,V1 := NULL]

#Now loop through all years starting in 1982
for (i in 1:length(years)) {
  
  sorensen_dissimilarities_allyears_neusgridcell[i,year := years[i]]
  
  reduced_year <- neus_summed[year == years[i],]
  reduced_year[,present := ifelse(ABUNDANCE_sum > 1, 1, 0)]
  reduced_year_wide <- dcast(reduced_year, year + cell ~ matched_name2, value.var = "present", fill = 0) #long to wide data for community matrix, column names are year, cell, then scientific 
  
  reduced_year_wide.bp <- betapart.core(reduced_year_wide[,3:ncol(reduced_year_wide)])
  
  # sampling across equal sites
reduced_year_wide.sample <- beta.sample(reduced_year_wide.bp, sites = nrow(reduced_year_wide), samples = 100)

  #combined Sorensen dissimilarity
  sorensen_dissimilarities_allyears_neusgridcell[i, sorensen_dissimilarity := reduced_year_wide.sample$mean.values[[3]]]
  
  #balanced
  sorensen_dissimilarities_allyears_neusgridcell[i, simpson_dissimilarity_turnover := reduced_year_wide.sample$mean.values[[1]]]
  
  #gradients
  sorensen_dissimilarities_allyears_neusgridcell[i, sorensen_dissimilarity_nestedness := reduced_year_wide.sample$mean.values[[2]]]
  
}

saveRDS(sorensen_dissimilarities_allyears_neusgridcell, "sorensen_dissimilarities_allyears_neusgridcell.RData")

#most of beta diversity derived from turnover
ggplot(data = sorensen_dissimilarities_allyears_neusgridcell, aes(year)) +
  geom_line(aes(y = sorensen_dissimilarity), color = "purple") +
  geom_line(aes(y = simpson_dissimilarity_turnover), color = "green") +
  labs(y = "Sorensen Dissimilarity", x = "Year") +
  theme_classic()


```

Similarity across grid cells
```{r}
#how similar are different grid cells
# pairwise
reduced_year_wide.betapair <- beta.pair.abund(reduced_year_wide)
# plotting clusters
dist.s <- reduced_year_wide.sample$sampled.values
plot(hclust(reduced_year_wide.betapair$beta.bray.bal, method="average"), hang=-1, main='', sub='', xlab='')
title(xlab=expression(beta[bray.bal]), line=0.3)
plot(hclust(reduced_year_wide.betapair$beta.bray.gra, method="average"), hang=-1, main='', sub='',, xlab='')
title(xlab=expression(beta[bray.gra]), line=0.3)
```

