---
title: "Spatial Turnover Exploratory Analysis"
output: html_notebook
---

Instructions from Malin:

Is anyone up for an exploratory analysis with the trawl data? It would be very cool to look at changes in spatial turnover, similar to [Magurran, Dornelas, Moyes, Gotelli, McGill et al. 2015](http://www.nature.com/ncomms/2015/150924/ncomms9405/full/ncomms9405.html) but across a much wider set of regions. An idea from conversations at [HIFMB](https://hifmb.de/en/) yesterday. 

From Magurran et al. *"Here, we analyse an exceptionally comprehensive 29-year time series of North Atlantic groundfish assemblages monitored over 5° latitude to the west of Scotland. These fish assemblages show no systematic change in species richness through time, but steady change in species composition, leading to an increase in spatial homogenization: the species identity of colder northern localities increasingly resembles that of warmer southern localities."*

Steps would be:

1. Calculate Jacard similarity between each haul in a single year in a single region. The vegdist() function in the vegan package would do this,OceanAdapt should have the data you’d need 
https://github.com/pinskylab/OceanAdapt/tree/master/data_raw for NEUS files

1. Calculate geographic distance between each haul in km (e.g., geosphere package in R could work).
1. Plot similarity vs. geographic distance and fit a regression line.
1. Repeat for each year.

A good place to start would just be the Northeast US. We could take a look and then think about expanding out to the other regions and to European data.

Methods from Magurran et al. 

First assign the rectangles to nine 30' latitudinal bands
Compile community time series for each latitudinal band
Sample rarefaction ensures equal sampling effort across bands and is used in the calculation of temporal $\alpha$ diversity and temporal $\beta$ diversity. 
First calculate similarity in relation to the start of the survey
Next, for each year, we compute pairwise compositional similarity of these latitudinal bands
We also construct distance-decay plots for each year

```{r setup}
library(data.table)
library(vegan)
```


```{r import all data from Ocean Adapt}
#neus_strata <- read.csv("neus_strata.csv")
#save(neus_strata, file = "neus_strata.RData")

load("neus_strata.Rdata")
load("neus_Survdat.RData")
load("neus_SVSPP.RData")

```

```{r extract columns we need from each}
neus_strata.r <- data.table(neus_strata[,c(2:3,5:6)])
neus_survdat.r <- data.table(survdat[,c(3,5,6,8,9, 10, 11)])
neus_spp.r <- data.table(spp[,c(1,2,4)])

```

```{r link data tables together}
stratamerge <- neus_survdat.r[neus_strata.r, on = "STRATUM"]
neus_full <- stratamerge[neus_spp.r, on = "SVSPP"]
neus_full

```

Make sure every observation is of an identifiable fish, aka SVSPP != 0

```{r ID fish}
neus_full.ID <- neus_full[SVSPP != 0,]
```

Group Stratums

For now, I will split into MAB, SNE, GB, and GOM

From Michelle's metadata:

Stratum group code: 
01 = Trawl, offshore north of Hatteras; 
03 = Trawl, inshore north of Hatteras; 
05 = Scotian shelf; 
07 = Trawl, inshore south of Hatteras; 
08 = Trawl, Offshore south of Hatteras;

#we only want 01 and 03's 

```{r only offshore and inshore north of hatteras}
#neus_full.ID.0103 <- neus_full.ID[STRATUM < 4000] not doing this any more, instead will reduce using below 
```

Region Stratum Key
```{r region stratum key}
region_strata_key <- data.table(read.csv("key_reg_strata.csv"))

```

```{r link key to neus trawls}
neus_full.ID.regions <- neus_full.ID[region_strata_key, on = "STRATUM_NAME"] 
neus_full.ID.regions.r <- na.omit(neus_full.ID.regions, cols = "reg") #get rid of all rows without a reg
summary(neus_full.ID.regions.r$reg)

```
Find center point of trawls in each region to find distance between regions
```{r center points of trawls }
library(geosphere)
library(rgeos)


latlon_reg <- na.omit(unique(neus_full.ID.regions.r, by = c("LAT", "LON")), c("LAT", "LON")) # single lat lon for each, to reduce bias, also, get rid of single NA row for Lat and Lon


# Define centroid function - Lon is first argument and Lat is second
# Geosphere takes a matrix with two columns: Lon|Lat, so we use cbind to coerce the data to this form
findCentroid <- function(LON, LAT, ...){
  centroid(cbind(LON, LAT), ...)
}

# Find centroid Lon and Lat by ID, as required
latlon_reg[, c("Cent_lon", "Cent_lat") := as.list(findCentroid(LON, LAT)), by = reg] #I will get an error on this when there are NA's in the LAT LON columns

reg_centroids <- unique(latlon_reg, by = c("Cent_lon", "Cent_lat"))

reg_centroids <- reg_centroids[, 13:15] #columns I need

#reorder so in order from N to S
reg_centroids[, reg := factor(reg, levels = c("GME", "GEO", "SNE", "MAB"))]

setorder(reg_centroids, reg) #set to this new factor level order 

#the location of MAB still doesn't make sense (too far south), but I'm moving forward anyway, much better than before
```



```{r reduce to simple "yes/no present in this year"}
#one PRESENT value for each YEAR, SEASON, reg, and SVSPP combination
neus_full.ID.regions.unique <- unique(neus_full.ID.regions, by = c("YEAR", "reg", "SVSPP")) #unique mini-region, year, species combinations

#reduce data table to just be svspp, year, sciname, comname, reg

neus_full.ID.regions.unique.r <- data.table(na.omit(neus_full.ID.regions.unique[, c("SVSPP", "SCINAME", "COMNAME", "YEAR", "reg")], col = "YEAR"))

#add presence column 
neus_full.ID.regions.unique.r[,"present" := 1]

#now I need to populate a comparison data table with every possible year, region, and species combination to add 0s
reg <- c("GME", "GEO", "SNE", "MAB")
spp.list <- unique(spp[SVSPP != 0,]$SVSPP) #0=no ID species
years <- unique(survdat[YEAR >1966,]$YEAR) #MAB no data pre 1967

full_set <- data.table(expand.grid("SVSPP" = spp.list, "YEAR" = years, "reg" = reg))




```

```{r bring in actual presences to expansion}
merge <- neus_full.ID.regions.unique.r[full_set, on = c("YEAR", "reg", "SVSPP")]
merge[,"present" := ifelse(is.na(SCINAME), 0,1)][,"SCINAME" := NULL][,"COMNAME" := NULL] #if scientific name and common name are empty, that is an absence, if populated, presence, also, we don't need these columns, as we'll bring them in below

#simplify spp key
spp_key <- neus_spp.r[SVSPP >= 1,]

neus_pres_abs <- merge[spp_key, on = c("SVSPP")] #also don't actually need all columns


sum(neus_pres_abs$present) #check to make sure we have the right number of presences

```
"To use the package the data have to be in the form of a community matrix.  The basic form is to have the species as columns and the plots as rows."

All presence observations corresponding to each species, reg, year under the same row
```{r long to wide}
#1967 is the first year there is data for MAB, so we will have to start there
neus_pres_abs_1year <- neus_pres_abs[YEAR == 1967,]


neus_pres_abs_1year_wide <- dcast(neus_pres_abs_1year, YEAR + reg ~ SCINAME, value.var = "present") #long to wide data for community matrix, column names are SVSPP
```


```{r jaccard dissimiliarities between regions in one year}

#to make vegdist work, each year-reg combination needs to be a "plot#"
#I'll do this super simply here
plot <- data.table("plot" = c(1:4))

neus_pres_abs_1year_wide.communitymatrix <- cbind(neus_pres_abs_1year_wide[,c(3:819)], neus_pres_abs_1year_wide[,c(1:2)])


similarities_1967 <- vegdist(neus_pres_abs_1year_wide.communitymatrix[,c(1:817)], method = "jaccard") #similarity matrix for 1967

reg_centroids_plots <- cbind(plot, reg_centroids)

reg_distances <- distm(reg_centroids_plots[,c(3, 4)]) #distance matrix, consistent across years
rownames(reg_distances) <- reg_centroids_plots$plot
colnames(reg_distances) <- reg_centroids_plots$plot

```



Turning distance and similarity matrices from long to wide form in order to do analyses
```{r matrices wide to long form}
library(reshape2)

#reshape distances
reg_distance.l <- melt(as.matrix(reg_distances), varnames = c("reg1", "reg2")) #matrix to data frame, 
reg_distance.l <- data.table(reg_distance.l) #and then to data table
reg_distance.l <- reg_distance.l[reg1 >= reg2,] # to get rid of repetitions
colnames(reg_distance.l)[colnames(reg_distance.l)=="value"] <- "distance(m)" #change column name to distance

#reshape similarities
similarities_1967.l <- melt(as.matrix(similarities_1967), varnames = c("reg1", "reg2"))
similarities_1967.l <- data.table(similarities_1967.l) #and then to data table
similarities_1967.l <- similarities_1967.l[reg1 >= reg2,] # to get rid of repetitions
colnames(similarities_1967.l)[colnames(similarities_1967.l)=="value"] <- "jaccard_similarity" #change column name to distance


similarities_1967.l[reg1 == reg2, jaccard_similarity := 1] #if they are the same value, assign jaccard_similarity of one

#add year for these values
similarities_1967.l[, "YEAR" := 1967]

#merge distance with jaccard_similarity for this year
similarities_1967.l.full <- reg_distance.l[similarities_1967.l, on = c("reg1", "reg2")]

#make plot
plot(x=similarities_1967.l.full$`distance(m)`, y=similarities_1967.l.full$jaccard_similarity, xlab = "Distance(m)", ylab="Jaccard Similarity", pch = 19) #this doesn't make sense, going back

 
```

PCA for visual
```{r PCA}
vare.pca <- rda(neus_pres_abs_1year_wide.communitymatrix[,c(1:817)])
plot(vare.pca, display = "sites")

```

```{r jaccard dissimilarities between regions in multiple years}
distances_similarities_allyears <- data.table("reg1" = integer(), "reg2" = integer(), "distance(m)" = numeric(),"jaccard_similarity" = numeric(), YEAR = integer())

#Now loop through all years starting in 1967
for (i in 1:length(years)) {
  reduced_year <- neus_pres_abs[YEAR == years[i],]
  reduced_year_wide <- dcast(reduced_year, YEAR + reg ~ SCINAME, value.var = "present") #long to wide data for community matrix, column names are SVSPP
  
  communitymatrix <- cbind(reduced_year_wide[,c(3:819)], reduced_year_wide[,c(1:2)]) #community matrix
  
  
  similarities <- vegdist(communitymatrix[,c(1:817)], method = "jaccard") #similarity matrix

  #reshape similarities
  similarities.l <- melt(as.matrix(similarities), varnames = c("reg1", "reg2"))
  similarities.l <- data.table(similarities.l) #and then to data table
  similarities.l <- similarities.l[reg1 >= reg2,] # to get rid of repetitions
  colnames(similarities.l)[colnames(similarities.l)=="value"] <- "jaccard_similarity" #change column name to distance


  similarities.l[reg1 == reg2, jaccard_similarity := 1] #if they are the same value, assign jaccard_similarity of one

#add year for these values
  similarities.l[, "YEAR" := years[i]]

#merge distance with jaccard_similarity for this year
  similarities.l.full <- reg_distance.l[similarities.l, on = c("reg1", "reg2")]

distances_similarities_allyears <- rbind(distances_similarities_allyears, similarities.l.full) #add to data table

jpeg(file=paste0("jaccardplot_", years[i], ".jpg"))
    plot(x=similarities.l.full$`distance(m)`, y=similarities.l.full$jaccard_similarity, xlab = "Distance(m)", ylab="Jaccard Similarity", ylim = c(0:1), pch = 19)
 dev.off()
 
 #PCA
vare.pca <- rda(communitymatrix[,c(1:817)])
jpeg(file=paste0("jaccardpca_", years[i], ".jpg"))
plot(vare.pca, display = "sites", xlim = c(-4,4), ylim = c(-4,4))
dev.off()
  
}

```

Plot jaccard for 1 and 2 through time
```{r 1 and 2 through time}
reduced_1_2 <- distances_similarities_allyears[reg1 == 2 & reg2 ==1,]
plot(x=reduced_1_2$YEAR, y = reduced_1_2$jaccard_similarity, pch = 19, xlab = "Year", ylab = "Jaccard Similarity", ylim= c(0.2,0.9))
abline(lm(reduced_1_2$jaccard_similarity~reduced_1_2$YEAR))

reg_1_2.lm <- lm(reduced_1_2$jaccard_similarity~reduced_1_2$YEAR)

summary(reg_1_2.lm)

jpeg(file = "GEO & GME Jaccard Through Time.jpg")
plot(x=reduced_1_2$YEAR, y = reduced_1_2$jaccard_similarity, pch = 19, xlab = "Year", ylab = "Jaccard Similarity", ylim= c(0.2,0.9))
abline(lm(reduced_1_2$jaccard_similarity~reduced_1_2$YEAR))
dev.off()
```


Plot jaccard for 1 and 3 through time
```{r 1 and 3 through time}
reduced_1_3 <- distances_similarities_allyears[reg1 == 3 & reg2 ==1,]
plot(x=reduced_1_3$YEAR, y = reduced_1_3$jaccard_similarity, pch = 19, xlab = "Year", ylab = "Jaccard Similarity", ylim= c(0.2,0.9))
abline(lm(reduced_1_3$jaccard_similarity~reduced_1_3$YEAR))

reg_1_3.lm <- lm(reduced_1_2$jaccard_similarity~reduced_1_3$YEAR)

summary(reg_1_3.lm)

jpeg(file = "GME & SNE Jaccard Through Time.jpg")
plot(x=reduced_1_3$YEAR, y = reduced_1_3$jaccard_similarity, pch = 19, xlab = "Year", ylab = "Jaccard Similarity", ylim= c(0.2,0.9))
abline(lm(reduced_1_3$jaccard_similarity~reduced_1_3$YEAR))
dev.off()
```


Plot jaccard for 1 and 4 through time
```{r 1 and 4 through time}
reduced_1_4 <- distances_similarities_allyears[reg1 == 4 & reg2 ==1,]
plot(x=reduced_1_4$YEAR, y = reduced_1_4$jaccard_similarity, pch = 19, xlab = "Year", ylab = "Jaccard Similarity", ylim= c(0.2,0.9))
abline(lm(reduced_1_4$jaccard_similarity~reduced_1_4$YEAR))

reg_1_4.lm <- lm(reduced_1_4$jaccard_similarity~reduced_1_4$YEAR)

summary(reg_1_4.lm)

jpeg(file = "MAB & GME Jaccard Through Time.jpg")
plot(x=reduced_1_4$YEAR, y = reduced_1_4$jaccard_similarity, pch = 19, xlab = "Year", ylab = "Jaccard Similarity", ylim= c(0.2,0.9))
abline(lm(reduced_1_4$jaccard_similarity~reduced_1_4$YEAR))
dev.off()
```


Plot jaccard for 2 and 3 through time
```{r 2 and 3 through time}
reduced_3_2 <- distances_similarities_allyears[reg1 == 3 & reg2 ==1,]
plot(x=reduced_1_3$YEAR, y = reduced_1_3$jaccard_similarity, pch = 19, xlab = "Year", ylab = "Jaccard Similarity", ylim= c(0.2,0.9))
abline(lm(reduced_1_3$jaccard_similarity~reduced_1_3$YEAR))

reg_1_3.lm <- lm(reduced_1_3$jaccard_similarity~reduced_1_3$YEAR)

summary(reg_1_3.lm)

jpeg(file = "GEO & SNE Jaccard Through Time.jpg")
plot(x=reduced_1_3$YEAR, y = reduced_1_3$jaccard_similarity, pch = 19, xlab = "Year", ylab = "Jaccard Similarity", ylim= c(0.2,0.9))
abline(lm(reduced_1_3$jaccard_similarity~reduced_1_3$YEAR))
dev.off()
```


Plot jaccard for 2 and 4 through time
```{r 2 and 4 through time}
reduced_4_2 <- distances_similarities_allyears[reg1 == 4 & reg2 ==2,]
plot(x=reduced_4_2$YEAR, y = reduced_4_2$jaccard_similarity, pch = 19, xlab = "Year", ylab = "Jaccard Similarity", ylim= c(0.2,0.9))
abline(lm(reduced_4_2$jaccard_similarity~reduced_4_2$YEAR))

reg_4_2.lm <- lm(reduced_4_2$jaccard_similarity~reduced_4_2$YEAR)

summary(reg_4_2.lm)

jpeg(file = "GEO & MAB Jaccard Through Time.jpg")
plot(x=reduced_4_2$YEAR, y = reduced_4_2$jaccard_similarity, pch = 19, xlab = "Year", ylab = "Jaccard Similarity", ylim= c(0.2,0.9))
abline(lm(reduced_4_2$jaccard_similarity~reduced_4_2$YEAR))
dev.off()

```


Plot jaccard for 3 and 4 through time
```{r 3 and 4 through time}
reduced_3_4 <- distances_similarities_allyears[reg1 == 4 & reg2 ==3,]
plot(x=reduced_3_4$YEAR, y = reduced_3_4$jaccard_similarity, pch = 19, xlab = "Year", ylab = "Jaccard Similarity", ylim= c(0.2,0.9))
abline(lm(reduced_3_4$jaccard_similarity~reduced_3_4$YEAR))

reg_3_4.lm <- lm(reduced_3_4$jaccard_similarity~reduced_3_4$YEAR)

summary(reg_3_4.lm)

jpeg(file = "SNE & MAB Jaccard Through Time.jpg")
plot(x=reduced_3_4$YEAR, y = reduced_3_4$jaccard_similarity, pch = 19, xlab = "Year", ylab = "Jaccard Similarity", ylim= c(0.2,0.9))
abline(lm(reduced_3_4$jaccard_similarity~reduced_3_4$YEAR))
dev.off()
```

Looking at changes in jaccard similarity across time for each region individually

```{r GME jaccard across time (between each year)}
  reduced_region <- neus_pres_abs[reg == "GME",]
  reduced_region_wide <- dcast(reduced_region, YEAR + reg ~ SCINAME, value.var = "present") #long to wide data for community matrix, column names are SVSPP
  
  communitymatrix_GME <- cbind(reduced_region_wide[,c(3:819)], reduced_region_wide[,c(1:2)]) #community matrix
  
  
  similarities_GME <- vegdist(communitymatrix_GME[,c(1:817)], method = "jaccard") #similarity matrix

  #reshape similarities
  similarities.l_GME <- melt(as.matrix(similarities_GME), varnames = c("year1", "year2"))
  similarities.l_GME <- data.table(similarities.l_GME) #and then to data table
  
  
  #I only want year 2 to be less than year 1 (don't need all comparisons)
  values_key <- c(1, rep(2:51, each=2), 52)
  year_turnover_key <- matrix(values_key, nrow = 2)
  year_turnover_key <- data.table(year_turnover_key)
  year_turnover_key.t <- transpose(year_turnover_key)
  year_turnover_key.final <- data.table("year1" = year_turnover_key.t[,V2], "year2" = year_turnover_key.t[,V1])
  
  #combine key with similarities.l_GME
similarities_final_GME <- year_turnover_key.final[similarities.l_GME, on = c("year1","year2"), nomatch = 0]
  
  
colnames(similarities_final_GME)[colnames(similarities_final_GME)=="value"] <- "jaccard_similarity"
#change column name to distance

similarities_final_GME[,"YEAR" := years[1:51]]


jpeg(file="jaccard_overtime_GME.jpg")
    plot(x=similarities_final_GME$YEAR, y=similarities_final_GME$jaccard_similarity, xlab = "Year", ylab="Jaccard Similarity", ylim = c(0:1), pch = 19)
 dev.off()


```

```{r GME jaccard across time (compared to first year)}
  reduced_region <- neus_pres_abs[reg == "GME",]
  reduced_region_wide <- dcast(reduced_region, YEAR + reg ~ SCINAME, value.var = "present") #long to wide data for community matrix, column names are SVSPP
  
  communitymatrix_GME <- cbind(reduced_region_wide[,c(3:819)], reduced_region_wide[,c(1:2)]) #community matrix
  
  
  similarities_GME <- vegdist(communitymatrix_GME[,c(1:817)], method = "jaccard") #similarity matrix

  #reshape similarities
  similarities.l_GME <- melt(as.matrix(similarities_GME), varnames = c("year1", "year2"))
  similarities.l_GME <- data.table(similarities.l_GME) #and then to data table
  
  
  #I only want to compare years with first year (don't need all comparisons)
  year_turnover_key.firstyear <- data.table("year1" = c(2:52), "year2" = 1)
  
  #combine key with similarities.l_GME
similarities_final_GME_year1 <- year_turnover_key.firstyear[similarities.l_GME, on = c("year1","year2"), nomatch = 0]
  
  
colnames(similarities_final_GME_year1)[colnames(similarities_final_GME_year1)=="value"] <- "jaccard_similarity"
#change column name to distance

similarities_final_GME_year1[,"YEAR" := years[1:51]]


jpeg(file="jaccard_overtime_GME_year1comparison.jpg")
    plot(x=similarities_final_GME_year1$YEAR, y=similarities_final_GME_year1$jaccard_similarity, xlab = "Year", ylab="Jaccard Similarity (to year 1)", ylim = c(0:1), pch = 19)
 dev.off()


```

Looking at changes in jaccard similarity across time for each region individually

```{r GEO jaccard across time}
  reduced_region <- neus_pres_abs[reg == "GEO",]
  reduced_region_wide <- dcast(reduced_region, YEAR + reg ~ SCINAME, value.var = "present") #long to wide data for community matrix, column names are SVSPP
  
  communitymatrix_GEO <- cbind(reduced_region_wide[,c(3:819)], reduced_region_wide[,c(1:2)]) #community matrix
  
  
  similarities_GEO <- vegdist(communitymatrix_GEO[,c(1:817)], method = "jaccard") #similarity matrix

  #reshape similarities
  similarities.l_GEO <- melt(as.matrix(similarities_GEO), varnames = c("year1", "year2"))
  similarities.l_GEO <- data.table(similarities.l_GEO) #and then to data table
  
  
  #I only want year 2 to be less than year 1 (don't need all comparisons)
  values_key <- c(1, rep(2:51, each=2), 52)
  year_turnover_key <- matrix(values_key, nrow = 2)
  year_turnover_key <- data.table(year_turnover_key)
  year_turnover_key.t <- transpose(year_turnover_key)
  year_turnover_key.final <- data.table("year1" = year_turnover_key.t[,V2], "year2" = year_turnover_key.t[,V1])
  
  #combine key with similarities.l_GEO
similarities_final_GEO <- year_turnover_key.final[similarities.l_GEO, on = c("year1","year2"), nomatch = 0]
  
  
colnames(similarities_final_GEO)[colnames(similarities_final_GEO)=="value"] <- "jaccard_similarity"
#change column name to distance

similarities_final_GEO[,"YEAR" := years[1:51]]


jpeg(file="jaccard_overtime_GEO.jpg")
    plot(x=similarities_final_GEO$YEAR, y=similarities_final_GEO$jaccard_similarity, xlab = "Year", ylab="Jaccard Similarity", ylim = c(0:1), pch = 19)
 dev.off()


```

```{r GEO jaccard across time (compared to first year)}
  reduced_region <- neus_pres_abs[reg == "GEO",]
  reduced_region_wide <- dcast(reduced_region, YEAR + reg ~ SCINAME, value.var = "present") #long to wide data for community matrix, column names are SVSPP
  
  communitymatrix_GEO <- cbind(reduced_region_wide[,c(3:819)], reduced_region_wide[,c(1:2)]) #community matrix
  
  
  similarities_GEO <- vegdist(communitymatrix_GEO[,c(1:817)], method = "jaccard") #similarity matrix

  #reshape similarities
  similarities.l_GEO <- melt(as.matrix(similarities_GEO), varnames = c("year1", "year2"))
  similarities.l_GEO <- data.table(similarities.l_GEO) #and then to data table
  
  
  #I only want to compare years with first year (don't need all comparisons)
  year_turnover_key.firstyear <- data.table("year1" = c(2:52), "year2" = 1)
  
  #combine key with similarities.l_GEO
similarities_final_GEO_year1 <- year_turnover_key.firstyear[similarities.l_GEO, on = c("year1","year2"), nomatch = 0]
  
  
colnames(similarities_final_GEO_year1)[colnames(similarities_final_GEO_year1)=="value"] <- "jaccard_similarity"
#change column name to distance

similarities_final_GEO_year1[,"YEAR" := years[1:51]]


jpeg(file="jaccard_overtime_GEO_year1comparison.jpg")
    plot(x=similarities_final_GEO_year1$YEAR, y=similarities_final_GEO_year1$jaccard_similarity, xlab = "Year", ylab="Jaccard Similarity (to year 1)", ylim = c(0:1), pch = 19)
 dev.off()


```


Looking at changes in jaccard similarity across time for each region individually

```{r SNE jaccard across time}
  reduced_region <- neus_pres_abs[reg == "SNE",]
  reduced_region_wide <- dcast(reduced_region, YEAR + reg ~ SCINAME, value.var = "present") #long to wide data for community matrix, column names are SVSPP
  
  communitymatrix_SNE <- cbind(reduced_region_wide[,c(3:819)], reduced_region_wide[,c(1:2)]) #community matrix
  
  
  similarities_SNE <- vegdist(communitymatrix_SNE[,c(1:817)], method = "jaccard") #similarity matrix

  #reshape similarities
  similarities.l_SNE <- melt(as.matrix(similarities_SNE), varnames = c("year1", "year2"))
  similarities.l_SNE <- data.table(similarities.l_SNE) #and then to data table
  
  
  #I only want year 2 to be less than year 1 (don't need all comparisons)
  values_key <- c(1, rep(2:51, each=2), 52)
  year_turnover_key <- matrix(values_key, nrow = 2)
  year_turnover_key <- data.table(year_turnover_key)
  year_turnover_key.t <- transpose(year_turnover_key)
  year_turnover_key.final <- data.table("year1" = year_turnover_key.t[,V2], "year2" = year_turnover_key.t[,V1])
  
  #combine key with similarities.l_SNE
similarities_final_SNE <- year_turnover_key.final[similarities.l_SNE, on = c("year1","year2"), nomatch = 0]
  
  
colnames(similarities_final_SNE)[colnames(similarities_final_SNE)=="value"] <- "jaccard_similarity"
#change column name to distance

similarities_final_SNE[,"YEAR" := years[1:51]]


jpeg(file="jaccard_overtime_SNE.jpg")
    plot(x=similarities_final_SNE$YEAR, y=similarities_final_SNE$jaccard_similarity, xlab = "Year", ylab="Jaccard Similarity", ylim = c(0:1), pch = 19)
 dev.off()


```


```{r SNE jaccard across time (compared to first year)}
  reduced_region <- neus_pres_abs[reg == "SNE",]
  reduced_region_wide <- dcast(reduced_region, YEAR + reg ~ SCINAME, value.var = "present") #long to wide data for community matrix, column names are SVSPP
  
  communitymatrix_SNE <- cbind(reduced_region_wide[,c(3:819)], reduced_region_wide[,c(1:2)]) #community matrix
  
  
  similarities_SNE <- vegdist(communitymatrix_SNE[,c(1:817)], method = "jaccard") #similarity matrix

  #reshape similarities
  similarities.l_SNE <- melt(as.matrix(similarities_SNE), varnames = c("year1", "year2"))
  similarities.l_SNE <- data.table(similarities.l_SNE) #and then to data table
  
  
  #I only want to compare years with first year (don't need all comparisons)
  year_turnover_key.firstyear <- data.table("year1" = c(2:52), "year2" = 1)
  
  #combine key with similarities.l_SNE
similarities_final_SNE_year1 <- year_turnover_key.firstyear[similarities.l_SNE, on = c("year1","year2"), nomatch = 0]
  
  
colnames(similarities_final_SNE_year1)[colnames(similarities_final_SNE_year1)=="value"] <- "jaccard_similarity"
#change column name to distance

similarities_final_SNE_year1[,"YEAR" := years[1:51]]


jpeg(file="jaccard_overtime_SNE_year1comparison.jpg")
    plot(x=similarities_final_SNE_year1$YEAR, y=similarities_final_SNE_year1$jaccard_similarity, xlab = "Year", ylab="Jaccard Similarity (to year 1)", ylim = c(0:1), pch = 19)
 dev.off()


```

Looking at changes in jaccard similarity across time for each region individually

```{r MAB jaccard across time}
  reduced_region <- neus_pres_abs[reg == "MAB",]
  reduced_region_wide <- dcast(reduced_region, YEAR + reg ~ SCINAME, value.var = "present") #long to wide data for community matrix, column names are SVSPP
  
  communitymatrix_MAB <- cbind(reduced_region_wide[,c(3:819)], reduced_region_wide[,c(1:2)]) #community matrix
  
  
  similarities_MAB <- vegdist(communitymatrix_MAB[,c(1:817)], method = "jaccard") #similarity matrix

  #reshape similarities
  similarities.l_MAB <- melt(as.matrix(similarities_MAB), varnames = c("year1", "year2"))
  similarities.l_MAB <- data.table(similarities.l_MAB) #and then to data table
  
  
  #I only want year 2 to be less than year 1 (don't need all comparisons)
  values_key <- c(1, rep(2:51, each=2), 52)
  year_turnover_key <- matrix(values_key, nrow = 2)
  year_turnover_key <- data.table(year_turnover_key)
  year_turnover_key.t <- transpose(year_turnover_key)
  year_turnover_key.final <- data.table("year1" = year_turnover_key.t[,V2], "year2" = year_turnover_key.t[,V1])
  
  #combine key with similarities.l_MAB
similarities_final_MAB <- year_turnover_key.final[similarities.l_MAB, on = c("year1","year2"), nomatch = 0]
  
  
colnames(similarities_final_MAB)[colnames(similarities_final_MAB)=="value"] <- "jaccard_similarity"
#change column name to distance

similarities_final_MAB[,"YEAR" := years[1:51]]


jpeg(file="jaccard_overtime_MAB.jpg")
    plot(x=similarities_final_MAB$YEAR, y=similarities_final_MAB$jaccard_similarity, xlab = "Year", ylab="Jaccard Similarity", ylim = c(0:1), pch = 19)
 dev.off()


```

```{r MAB jaccard across time (compared to first year)}
  reduced_region <- neus_pres_abs[reg == "MAB",]
  reduced_region_wide <- dcast(reduced_region, YEAR + reg ~ SCINAME, value.var = "present") #long to wide data for community matrix, column names are SVSPP
  
  communitymatrix_MAB <- cbind(reduced_region_wide[,c(3:819)], reduced_region_wide[,c(1:2)]) #community matrix
  
  
  similarities_MAB <- vegdist(communitymatrix_MAB[,c(1:817)], method = "jaccard") #similarity matrix

  #reshape similarities
  similarities.l_MAB <- melt(as.matrix(similarities_MAB), varnames = c("year1", "year2"))
  similarities.l_MAB <- data.table(similarities.l_MAB) #and then to data table
  
  
  #I only want to compare years with first year (don't need all comparisons)
  year_turnover_key.firstyear <- data.table("year1" = c(2:52), "year2" = 1)
  
  #combine key with similarities.l_MAB
similarities_final_MAB_year1 <- year_turnover_key.firstyear[similarities.l_MAB, on = c("year1","year2"), nomatch = 0]
  
  
colnames(similarities_final_MAB_year1)[colnames(similarities_final_MAB_year1)=="value"] <- "jaccard_similarity"
#change column name to distance

similarities_final_MAB_year1[,"YEAR" := years[1:51]]


jpeg(file="jaccard_overtime_MAB_year1comparison.jpg")
    plot(x=similarities_final_MAB_year1$YEAR, y=similarities_final_MAB_year1$jaccard_similarity, xlab = "Year", ylab="Jaccard Similarity (to year 1)", ylim = c(0:1), pch = 19)
 dev.off()


```

Simplified species richness through time
```{r spp richness through time}
library(ggplot2)

GME_spp_richness <- rowSums(communitymatrix_GME[,1:817])
GEO_spp_richness <- rowSums(communitymatrix_GEO[,1:817])
SNE_spp_richness <- rowSums(communitymatrix_SNE[,1:817])
MAB_spp_richness <- rowSums(communitymatrix_MAB[,1:817])

spp_richness <- data.table("year" = years, "GME" = GME_spp_richness, "GEO" = GEO_spp_richness, "SNE" = SNE_spp_richness, "MAB" = MAB_spp_richness)

spp_richness.l <- melt(spp_richness, id.vars = c("year"),
                measure.vars = c("GME", "GEO", "SNE", "MAB"),
                value.name = "richness")

species_richness_reg <- ggplot(data = spp_richness.l, aes(x = year, y=richness)) +
  geom_line() +
  geom_smooth(method = lm) +
  facet_wrap(~variable) +
  theme_classic()

ggsave(species_richness_reg, file = "species_richness_reg.eps")
  
```

?melt
