---
title: "Spatial Turnover Exploratory Analysis"
output: html_notebook
---

Instructions from Malin:

Is anyone up for an exploratory analysis with the trawl data? It would be very cool to look at changes in spatial turnover, similar to [Magurran, Dornelas, Moyes, Gotelli, McGill et al. 2015](http://www.nature.com/ncomms/2015/150924/ncomms9405/full/ncomms9405.html) but across a much wider set of regions. An idea from conversations at [HIFMB](https://hifmb.de/en/) yesterday. 

From Magurran et al. *"Here, we analyse an exceptionally comprehensive 29-year time series of North Atlantic groundfish assemblages monitored over 5° latitude to the west of Scotland. These fish assemblages show no systematic change in species richness through time, but steady change in species composition, leading to an increase in spatial homogenization: the species identity of colder northern localities increasingly resembles that of warmer southern localities."*

Steps would be:

1. Calculate Jacard dissimilarity between each haul in a single year in a single region. The vegdist() function in the vegan package would do this,OceanAdapt should have the data you’d need 
https://github.com/pinskylab/OceanAdapt/tree/master/data_raw for NEUS files

1. Calculate geographic distance between each haul in km (e.g., geosphere package in R could work).
1. Plot dissimilarity vs. geographic distance and fit a regression line.
1. Repeat for each year.

A good place to start would just be the Northeast US. We could take a look and then think about expanding out to the other regions and to European data.

Methods from Magurran et al. 

First assign the rectangles to nine 30' latitudinal bands
Compile community time series for each latitudinal band
Sample rarefaction ensures equal sampling effort across bands and is used in the calculation of temporal $\alpha$ diversity and temporal $\beta$ diversity. 
First calculate dissimilarity in relation to the start of the survey
Next, for each year, we compute pairwise compositional dissimilarity of these latitudinal bands
We also construct distance-decay plots for each year

```{r setup}
library(data.table)
library(vegan)
```


```{r import all data from Ocean Adapt}
#neus_strata <- read.csv("neus_strata.csv")
#save(neus_strata, file = "neus_strata.RData")

load("neus_strata.Rdata")
load("neus_Survdat.RData")
load("neus_SVSPP.RData")

```

```{r extract columns we need from each}
neus_strata.r <- data.table(neus_strata[,c(2:3,5:6)])
neus_survdat.r <- data.table(survdat[,c(3,5,6,8,9, 10, 11)])
neus_spp.r <- data.table(spp[,c(1,2,4)])

```

```{r link data tables together}
stratamerge <- neus_survdat.r[neus_strata.r, on = "STRATUM"]
neus_full <- stratamerge[neus_spp.r, on = "SVSPP"]
neus_full

```

Make sure every observation is of an identifiable fish, aka SVSPP != 0

```{r ID fish}
neus_full.ID <- neus_full[SVSPP != 0,]
```

Group Stratums

For now, I will split into MAB, SNE, GB, and GOM

From Michelle's metadata:

Stratum group code: 
01 = Trawl, offshore north of Hatteras; 
03 = Trawl, inshore north of Hatteras; 
05 = Scotian shelf; 
07 = Trawl, inshore south of Hatteras; 
08 = Trawl, Offshore south of Hatteras;

#we only want 01 and 03's 

```{r only offshore and inshore north of hatteras}
#neus_full.ID.0103 <- neus_full.ID[STRATUM < 4000] not doing this any more, instead will reduce using below 
```

Region Stratum Key
```{r region stratum key}
region_strata_key <- data.table(read.csv("key_reg_strata.csv"))

```

```{r link key to neus trawls}
neus_full.ID.regions <- neus_full.ID[region_strata_key, on = "STRATUM_NAME"] 
neus_full.ID.regions.r <- na.omit(neus_full.ID.regions, cols = "reg") #get rid of all rows without a reg
summary(neus_full.ID.regions.r$reg)

```
Find center point of trawls in each region to find distance between regions
```{r center points of trawls }
library(geosphere)
library(rgeos)


latlon_reg <- na.omit(unique(neus_full.ID.regions.r, by = c("LAT", "LON")), c("LAT", "LON")) # single lat lon for each, to reduce bias, also, get rid of single NA row for Lat and Lon


# Define centroid function - Lon is first argument and Lat is second
# Geosphere takes a matrix with two columns: Lon|Lat, so we use cbind to coerce the data to this form
findCentroid <- function(LON, LAT, ...){
  centroid(cbind(LON, LAT), ...)
}

# Find centroid Lon and Lat by ID, as required
latlon_reg[, c("Cent_lon", "Cent_lat") := as.list(findCentroid(LON, LAT)), by = reg] #I will get an error on this when there are NA's in the LAT LON columns

reg_centroids <- unique(latlon_reg, by = c("Cent_lon", "Cent_lat"))

reg_centroids <- reg_centroids[, 13:15] #columns I need

#reorder so in order from N to S
reg_centroids[, reg := factor(reg, levels = c("GME", "GEO", "SNE", "MAB"))]

setorder(reg_centroids, reg) #set to this new factor level order 

#the location of MAB still doesn't make sense (too far south), but I'm moving forward anyway, much better than before
```



```{r reduce to simple "yes/no present in this year"}
#one PRESENT value for each YEAR, SEASON, reg, and SVSPP combination
neus_full.ID.regions.unique <- unique(neus_full.ID.regions, by = c("YEAR", "reg", "SVSPP")) #unique mini-region, year, species combinations

#reduce data table to just be svspp, year, sciname, comname, reg

neus_full.ID.regions.unique.r <- data.table(na.omit(neus_full.ID.regions.unique[, c("SVSPP", "SCINAME", "COMNAME", "YEAR", "reg")], col = "YEAR"))

#add presence column 
neus_full.ID.regions.unique.r[,"present" := 1]

#now I need to populate a comparison data table with every possible year, region, and species combination to add 0s
reg <- c("GME", "GEO", "SNE", "MAB")
spp.list <- unique(spp[SVSPP != 0,]$SVSPP) #0=no ID species
years <- unique(survdat[YEAR >1966,]$YEAR) #MAB no data pre 1967

full_set <- data.table(expand.grid("SVSPP" = spp.list, "YEAR" = years, "reg" = reg))




```

```{r bring in actual presences to expansion}
merge <- neus_full.ID.regions.unique.r[full_set, on = c("YEAR", "reg", "SVSPP")]
merge[,"present" := ifelse(is.na(SCINAME), 0,1)][,"SCINAME" := NULL][,"COMNAME" := NULL] #if scientific name and common name are empty, that is an absence, if populated, presence, also, we don't need these columns, as we'll bring them in below

#simplify spp key
spp_key <- neus_spp.r[SVSPP >= 1,]

neus_pres_abs <- merge[spp_key, on = c("SVSPP")] #also don't actually need all columns


sum(neus_pres_abs$present) #check to make sure we have the right number of presences

```
"To use the package the data have to be in the form of a community matrix.  The basic form is to have the species as columns and the plots as rows."

All presence observations corresponding to each species, reg, year under the same row
```{r long to wide}
#1967 is the first year there is data for MAB, so we will have to start there
neus_pres_abs_1year <- neus_pres_abs[YEAR == 1967,]


neus_pres_abs_1year_wide <- dcast(neus_pres_abs_1year, YEAR + reg ~ SCINAME, value.var = "present") #long to wide data for community matrix, column names are SVSPP
```


```{r jaccard dissimiliarities between regions in one year}

#to make vegdist work, each year-reg combination needs to be a "plot#"
#I'll do this super simply here
plot <- data.table("plot" = c(1:4))

neus_pres_abs_1year_wide.communitymatrix <- cbind(neus_pres_abs_1year_wide[,c(3:819)], neus_pres_abs_1year_wide[,c(1:2)])


dissimilarities_1967 <- vegdist(neus_pres_abs_1year_wide.communitymatrix[,c(1:817)], method = "jaccard") #dissimilarity matrix for 1967

reg_centroids_plots <- cbind(plot, reg_centroids)

reg_distances <- distm(reg_centroids_plots[,c(3, 4)]) #distance matrix, consistent across years
rownames(reg_distances) <- reg_centroids_plots$plot
colnames(reg_distances) <- reg_centroids_plots$plot

```



Turning distance and dissimilarity matrices from long to wide form in order to do analyses
```{r matrices wide to long form}
library(reshape2)

#reshape distances
reg_distance.l <- melt(as.matrix(reg_distances), varnames = c("reg1", "reg2")) #matrix to data frame, 
reg_distance.l <- data.table(reg_distance.l) #and then to data table
reg_distance.l <- reg_distance.l[reg1 >= reg2,] # to get rid of repetitions
colnames(reg_distance.l)[colnames(reg_distance.l)=="value"] <- "distance(m)" #change column name to distance

#reshape dissimilarities
dissimilarities_1967.l <- melt(as.matrix(dissimilarities_1967), varnames = c("reg1", "reg2"))
dissimilarities_1967.l <- data.table(dissimilarities_1967.l) #and then to data table
dissimilarities_1967.l <- dissimilarities_1967.l[reg1 >= reg2,] # to get rid of repetitions
colnames(dissimilarities_1967.l)[colnames(dissimilarities_1967.l)=="value"] <- "jaccard_dissimilarity" #change column name to distance

#add year for these values
dissimilarities_1967.l[, "YEAR" := 1967]

#merge distance with jaccard_dissimilarity for this year
dissimilarities_1967.l.full <- reg_distance.l[dissimilarities_1967.l, on = c("reg1", "reg2")]

#make plot
plot(x=dissimilarities_1967.l.full$`distance(m)`, y=dissimilarities_1967.l.full$jaccard_dissimilarity, xlab = "Distance(m)", ylab="Jaccard dissimilarity", pch = 19) #this doesn't make sense, going back

 
```

PCA for visual
```{r PCA}
vare.pca <- rda(neus_pres_abs_1year_wide.communitymatrix[,c(1:817)])
plot(vare.pca, display = "sites")

```
Loop for all 
```{r jaccard disdissimilarities between regions in multiple years}
distances_dissimilarities_allyears <- data.table("reg1" = integer(), "reg2" = integer(), "distance(m)" = numeric(),"jaccard_dissimilarity" = numeric(), YEAR = integer())

#Now loop through all years starting in 1967
for (i in 1:length(years)) {
  reduced_year <- neus_pres_abs[YEAR == years[i],]
  reduced_year_wide <- dcast(reduced_year, YEAR + reg ~ SCINAME, value.var = "present") #long to wide data for community matrix, column names are SVSPP
  
  communitymatrix <- cbind(reduced_year_wide[,c(3:819)], reduced_year_wide[,c(1:2)]) #community matrix
  
  
  dissimilarities <- vegdist(communitymatrix[,c(1:817)], method = "jaccard") #dissimilarity matrix

  #reshape dissimilarities
  dissimilarities.l <- melt(as.matrix(dissimilarities), varnames = c("reg1", "reg2"))
  dissimilarities.l <- data.table(dissimilarities.l) #and then to data table
  dissimilarities.l <- dissimilarities.l[reg1 >= reg2,] # to get rid of repetitions
  colnames(dissimilarities.l)[colnames(dissimilarities.l)=="value"] <- "jaccard_dissimilarity" #change column name to distance


#add year for these values
  dissimilarities.l[, "YEAR" := years[i]]

#merge distance with jaccard_dissimilarity for this year
  dissimilarities.l.full <- reg_distance.l[dissimilarities.l, on = c("reg1", "reg2")]

distances_dissimilarities_allyears <- rbind(distances_dissimilarities_allyears, dissimilarities.l.full) #add to data table

jpeg(file=paste0("jaccardplot_", years[i], ".jpg"))
    plot(x=dissimilarities.l.full$`distance(m)`, y=dissimilarities.l.full$jaccard_dissimilarity, xlab = "Distance(m)", ylab="Jaccard dissimilarity", ylim = c(0:1), pch = 19)
 dev.off()
 
 #PCA
vare.pca <- rda(communitymatrix[,c(1:817)])
jpeg(file=paste0("jaccardpca_", years[i], ".jpg"))
plot(vare.pca, display = "sites", xlim = c(-4,4), ylim = c(-4,4))
dev.off()
  
}

```
Jaccard Dissimilarities Distance Decay all in one plot
This plot will be much more interesting with more strata, TO DO tomorrow
```{r Distance Decay one plot}
library(ggplot2)
distances_dissimilarities_allyears[,YEAR_fac := as.factor(YEAR)]
ggplot(data = distances_dissimilarities_allyears, aes(x = `distance(m)`, y = jaccard_dissimilarity, color = YEAR_fac)) +
  geom_line()
```

Plot jaccard for 1 and 2 through time
```{r 1 and 2 through time}
reduced_1_2 <- distances_dissimilarities_allyears[reg1 == 2 & reg2 ==1,]
geo_gme_overtime <- ggplot(data = reduced_1_2, aes(x = YEAR, y = jaccard_dissimilarity)) +
  geom_point() +
  geom_smooth(method = "lm") +
  ylim(0:1) +
  stat_cor(label.y = 0.1, label.x = 1990) +
  stat_regline_equation(label.y = 0.2, label.x = 1990) +
  labs(x= "Year", y = "Jaccard Dissimilarity") +
  ggtitle("GME vs. GEO") +
  theme_classic() +
  theme(text = element_text(size = 20))
  
  
ggsave(geo_gme_overtime, file ="GEO & GME Jaccard Through Time.eps")
```


Plot jaccard for 1 and 3 through time
```{r 1 and 3 through time}
reduced_1_3 <- distances_dissimilarities_allyears[reg1 == 3 & reg2 ==1,]
geo_sne_overtime <- ggplot(data = reduced_1_3, aes(x = YEAR, y = jaccard_dissimilarity)) +
  geom_point() +
  geom_smooth(method = "lm") +
  ylim(0:1) +
  stat_cor(label.y = 0.1, label.x = 1990) +
  stat_regline_equation(label.y = 0.2, label.x = 1990) +
  labs(x= "Year", y = "Jaccard Dissimilarity") +
  ggtitle("GME vs. SNE") +
  theme_classic() +
  theme(text = element_text(size = 20))
  
  
ggsave(geo_sne_overtime, file ="GEO & SNE Jaccard Through Time.eps")
```


Plot jaccard for 1 and 4 through time
```{r 1 and 4 through time}
reduced_1_4 <- distances_dissimilarities_allyears[reg1 == 4 & reg2 ==1,]
gme_mab_overtime <- ggplot(data = reduced_1_4, aes(x = YEAR, y = jaccard_dissimilarity)) +
  geom_point() +
  geom_smooth(method = "lm") +
  ylim(0:1) +
  stat_cor(label.y = 0.1, label.x = 1990) +
  stat_regline_equation(label.y = 0.2, label.x = 1990) +
  labs(x= "Year", y = "Jaccard Dissimilarity") +
  ggtitle("GME vs. MAB") +
  theme_classic() +
  theme(text = element_text(size = 20))
  
  
ggsave(gme_mab_overtime, file ="GME & MAB Jaccard Through Time.eps")
```


Plot jaccard for 2 and 3 through time
```{r 2 and 3 through time}
reduced_3_2 <- distances_dissimilarities_allyears[reg1 == 3 & reg2 ==2,]
sne_geo_overtime <- ggplot(data = reduced_3_2, aes(x = YEAR, y = jaccard_dissimilarity)) +
  geom_point() +
  geom_smooth(method = "lm") +
  ylim(0:1) +
  stat_cor(label.y = 0.1, label.x = 1990) +
  stat_regline_equation(label.y = 0.2, label.x = 1990) +
  labs(x= "Year", y = "Jaccard Dissimilarity") +
  ggtitle("SNE vs. GEO") +
  theme_classic() +
  theme(text = element_text(size = 20))
  
  
ggsave(sne_geo_overtime, file ="SNE & GEO Jaccard Through Time.eps")
```


Plot jaccard for 2 and 4 through time
```{r 2 and 4 through time}
reduced_4_2 <- distances_dissimilarities_allyears[reg1 == 4 & reg2 ==2,]
mab_geo_overtime <- ggplot(data = reduced_4_2, aes(x = YEAR, y = jaccard_dissimilarity)) +
  geom_point() +
  geom_smooth(method = "lm") +
  ylim(0:1) +
  stat_cor(label.y = 0.1, label.x = 1990) +
  stat_regline_equation(label.y = 0.2, label.x = 1990) +
  labs(x= "Year", y = "Jaccard Dissimilarity") +
  ggtitle("MAB vs. GEO") +
  theme_classic() +
  theme(text = element_text(size = 20))
  
  
ggsave(mab_geo_overtime, file ="MAB & GEO Jaccard Through Time.eps")

```


Plot jaccard for 3 and 4 through time
```{r 3 and 4 through time}
reduced_3_4 <- distances_dissimilarities_allyears[reg1 == 4 & reg2 ==3,]
sne_mab_overtime <- ggplot(data = reduced_3_4, aes(x = YEAR, y = jaccard_dissimilarity)) +
  geom_point() +
  geom_smooth(method = "lm") +
  ylim(0:1) +
  stat_cor(label.y = 0.1, label.x = 1990) +
  stat_regline_equation(label.y = 0.2, label.x = 1990) +
  labs(x= "Year", y = "Jaccard Dissimilarity") +
  ggtitle("SNE vs. MAB") +
  theme_classic() +
  theme(text = element_text(size = 20))
  
  
ggsave(sne_mab_overtime, file ="SNE & MAB Jaccard Through Time.eps")
```

Looking at changes in jaccard dissimilarity across time for each region individually

```{r GME jaccard across time (between each year)}
  reduced_region <- neus_pres_abs[reg == "GME",]
  reduced_region_wide <- dcast(reduced_region, YEAR + reg ~ SCINAME, value.var = "present") #long to wide data for community matrix, column names are SVSPP
  
  communitymatrix_GME <- cbind(reduced_region_wide[,c(3:819)], reduced_region_wide[,c(1:2)]) #community matrix
  
  
  dissimilarities_GME <- vegdist(communitymatrix_GME[,c(1:817)], method = "jaccard") #dissimilarity matrix

  #reshape dissimilarities
  dissimilarities.l_GME <- melt(as.matrix(dissimilarities_GME), varnames = c("year1", "year2"))
  dissimilarities.l_GME <- data.table(dissimilarities.l_GME) #and then to data table
  
  
  #I only want year 2 to be less than year 1 (don't need all comparisons)
  values_key <- c(1, rep(2:51, each=2), 52)
  year_turnover_key <- matrix(values_key, nrow = 2)
  year_turnover_key <- data.table(year_turnover_key)
  year_turnover_key.t <- transpose(year_turnover_key)
  year_turnover_key.final <- data.table("year1" = year_turnover_key.t[,V2], "year2" = year_turnover_key.t[,V1])
  
  #combine key with dissimilarities.l_GME
dissimilarities_final_GME <- year_turnover_key.final[dissimilarities.l_GME, on = c("year1","year2"), nomatch = 0]
  
  
colnames(dissimilarities_final_GME)[colnames(dissimilarities_final_GME)=="value"] <- "jaccard_dissimilarity"
#change column name to distance

dissimilarities_final_GME[,"YEAR" := years[1:51]]


jpeg(file="jaccard_overtime_GME.jpg")
    plot(x=dissimilarities_final_GME$YEAR, y=dissimilarities_final_GME$jaccard_dissimilarity, xlab = "Year", ylab="Jaccard dissimilarity", ylim = c(0:1), pch = 19)
 dev.off()


```

```{r GME jaccard across time (compared to first year)}
  reduced_region <- neus_pres_abs[reg == "GME",]
  reduced_region_wide <- dcast(reduced_region, YEAR + reg ~ SCINAME, value.var = "present") #long to wide data for community matrix, column names are SVSPP
  
  communitymatrix_GME <- cbind(reduced_region_wide[,c(3:819)], reduced_region_wide[,c(1:2)]) #community matrix
  
  
  dissimilarities_GME <- vegdist(communitymatrix_GME[,c(1:817)], method = "jaccard") #dissimilarity matrix

  #reshape dissimilarities
  dissimilarities.l_GME <- melt(as.matrix(dissimilarities_GME), varnames = c("year1", "year2"))
  dissimilarities.l_GME <- data.table(dissimilarities.l_GME) #and then to data table
  
  
  #I only want to compare years with first year (don't need all comparisons)
  year_turnover_key.firstyear <- data.table("year1" = c(2:52), "year2" = 1)
  
  #combine key with dissimilarities.l_GME
dissimilarities_final_GME_year1 <- year_turnover_key.firstyear[dissimilarities.l_GME, on = c("year1","year2"), nomatch = 0]
  
  
colnames(dissimilarities_final_GME_year1)[colnames(dissimilarities_final_GME_year1)=="value"] <- "jaccard_dissimilarity"
#change column name to distance

dissimilarities_final_GME_year1[,"YEAR" := years[1:51]]

jaccard_overtime_GME_year1comparison <- ggplot(data = dissimilarities_final_GME_year1, aes(x=YEAR, y = jaccard_dissimilarity)) +
  geom_point() +
  geom_smooth(method = "lm") +
  stat_cor(label.y = 0.15, label.x = 1985, size = 8) +
  stat_regline_equation(label.y =0.2, label.x = 1985, size = 8) +
  ggtitle("a. GME") +
  labs(x = "Year", y = "Jaccard dissimilarity") +
  theme_classic() +
  theme(text = element_text(size = 20))

ggsave(jaccard_overtime_GME_year1comparison, file = "jaccard_overtime_GME_year1comparison.eps" )



```

Looking at changes in jaccard dissimilarity across time for each region individually

```{r GEO jaccard across time}
  reduced_region <- neus_pres_abs[reg == "GEO",]
  reduced_region_wide <- dcast(reduced_region, YEAR + reg ~ SCINAME, value.var = "present") #long to wide data for community matrix, column names are SVSPP
  
  communitymatrix_GEO <- cbind(reduced_region_wide[,c(3:819)], reduced_region_wide[,c(1:2)]) #community matrix
  
  
  dissimilarities_GEO <- vegdist(communitymatrix_GEO[,c(1:817)], method = "jaccard") #dissimilarity matrix

  #reshape dissimilarities
  dissimilarities.l_GEO <- melt(as.matrix(dissimilarities_GEO), varnames = c("year1", "year2"))
  dissimilarities.l_GEO <- data.table(dissimilarities.l_GEO) #and then to data table
  
  
  #I only want year 2 to be less than year 1 (don't need all comparisons)
  values_key <- c(1, rep(2:51, each=2), 52)
  year_turnover_key <- matrix(values_key, nrow = 2)
  year_turnover_key <- data.table(year_turnover_key)
  year_turnover_key.t <- transpose(year_turnover_key)
  year_turnover_key.final <- data.table("year1" = year_turnover_key.t[,V2], "year2" = year_turnover_key.t[,V1])
  
  #combine key with dissimilarities.l_GEO
dissimilarities_final_GEO <- year_turnover_key.final[dissimilarities.l_GEO, on = c("year1","year2"), nomatch = 0]
  
  
colnames(dissimilarities_final_GEO)[colnames(dissimilarities_final_GEO)=="value"] <- "jaccard_dissimilarity"
#change column name to distance

dissimilarities_final_GEO[,"YEAR" := years[1:51]]


jpeg(file="jaccard_overtime_GEO.jpg")
    plot(x=dissimilarities_final_GEO$YEAR, y=dissimilarities_final_GEO$jaccard_dissimilarity, xlab = "Year", ylab="Jaccard dissimilarity", ylim = c(0:1), pch = 19)
 dev.off()


```

```{r GEO jaccard across time (compared to first year)}
  reduced_region <- neus_pres_abs[reg == "GEO",]
  reduced_region_wide <- dcast(reduced_region, YEAR + reg ~ SCINAME, value.var = "present") #long to wide data for community matrix, column names are SVSPP
  
  communitymatrix_GEO <- cbind(reduced_region_wide[,c(3:819)], reduced_region_wide[,c(1:2)]) #community matrix
  
  
  dissimilarities_GEO <- vegdist(communitymatrix_GEO[,c(1:817)], method = "jaccard") #dissimilarity matrix

  #reshape dissimilarities
  dissimilarities.l_GEO <- melt(as.matrix(dissimilarities_GEO), varnames = c("year1", "year2"))
  dissimilarities.l_GEO <- data.table(dissimilarities.l_GEO) #and then to data table
  
  
  #I only want to compare years with first year (don't need all comparisons)
  year_turnover_key.firstyear <- data.table("year1" = c(2:52), "year2" = 1)
  
  #combine key with dissimilarities.l_GEO
dissimilarities_final_GEO_year1 <- year_turnover_key.firstyear[dissimilarities.l_GEO, on = c("year1","year2"), nomatch = 0]
  
  
colnames(dissimilarities_final_GEO_year1)[colnames(dissimilarities_final_GEO_year1)=="value"] <- "jaccard_dissimilarity"
#change column name to distance

dissimilarities_final_GEO_year1[,"YEAR" := years[1:51]]


jaccard_overtime_GEO_year1comparison <- ggplot(data = dissimilarities_final_GEO_year1, aes(x=YEAR, y = jaccard_dissimilarity)) +
  geom_point() +
  geom_smooth(method = "lm") +
  stat_cor(label.y = 0.15, label.x = 1985, size = 8) +
  stat_regline_equation(label.y =0.2, label.x = 1985, size = 8) +
  ggtitle("b. GEO") +
  labs(x = "Year", y = "Jaccard dissimilarity") +
  theme_classic() +
  theme(text = element_text(size = 20))

ggsave(jaccard_overtime_GEO_year1comparison, file = "jaccard_overtime_GEO_year1comparison.eps" )


```


Looking at changes in jaccard dissimilarity across time for each region individually

```{r SNE jaccard across time}
  reduced_region <- neus_pres_abs[reg == "SNE",]
  reduced_region_wide <- dcast(reduced_region, YEAR + reg ~ SCINAME, value.var = "present") #long to wide data for community matrix, column names are SVSPP
  
  communitymatrix_SNE <- cbind(reduced_region_wide[,c(3:819)], reduced_region_wide[,c(1:2)]) #community matrix
  
  
  dissimilarities_SNE <- vegdist(communitymatrix_SNE[,c(1:817)], method = "jaccard") #dissimilarity matrix

  #reshape dissimilarities
  dissimilarities.l_SNE <- melt(as.matrix(dissimilarities_SNE), varnames = c("year1", "year2"))
  dissimilarities.l_SNE <- data.table(dissimilarities.l_SNE) #and then to data table
  
  
  #I only want year 2 to be less than year 1 (don't need all comparisons)
  values_key <- c(1, rep(2:51, each=2), 52)
  year_turnover_key <- matrix(values_key, nrow = 2)
  year_turnover_key <- data.table(year_turnover_key)
  year_turnover_key.t <- transpose(year_turnover_key)
  year_turnover_key.final <- data.table("year1" = year_turnover_key.t[,V2], "year2" = year_turnover_key.t[,V1])
  
  #combine key with dissimilarities.l_SNE
dissimilarities_final_SNE <- year_turnover_key.final[dissimilarities.l_SNE, on = c("year1","year2"), nomatch = 0]
  
  
colnames(dissimilarities_final_SNE)[colnames(dissimilarities_final_SNE)=="value"] <- "jaccard_dissimilarity"
#change column name to distance

dissimilarities_final_SNE[,"YEAR" := years[1:51]]


jpeg(file="jaccard_overtime_SNE.jpg")
    plot(x=dissimilarities_final_SNE$YEAR, y=dissimilarities_final_SNE$jaccard_dissimilarity, xlab = "Year", ylab="Jaccard dissimilarity", ylim = c(0:1), pch = 19)
 dev.off()


```


```{r SNE jaccard across time (compared to first year)}
  reduced_region <- neus_pres_abs[reg == "SNE",]
  reduced_region_wide <- dcast(reduced_region, YEAR + reg ~ SCINAME, value.var = "present") #long to wide data for community matrix, column names are SVSPP
  
  communitymatrix_SNE <- cbind(reduced_region_wide[,c(3:819)], reduced_region_wide[,c(1:2)]) #community matrix
  
  
  dissimilarities_SNE <- vegdist(communitymatrix_SNE[,c(1:817)], method = "jaccard") #dissimilarity matrix

  #reshape dissimilarities
  dissimilarities.l_SNE <- melt(as.matrix(dissimilarities_SNE), varnames = c("year1", "year2"))
  dissimilarities.l_SNE <- data.table(dissimilarities.l_SNE) #and then to data table
  
  
  #I only want to compare years with first year (don't need all comparisons)
  year_turnover_key.firstyear <- data.table("year1" = c(2:52), "year2" = 1)
  
  #combine key with dissimilarities.l_SNE
dissimilarities_final_SNE_year1 <- year_turnover_key.firstyear[dissimilarities.l_SNE, on = c("year1","year2"), nomatch = 0]
  
  
colnames(dissimilarities_final_SNE_year1)[colnames(dissimilarities_final_SNE_year1)=="value"] <- "jaccard_dissimilarity"
#change column name to distance

dissimilarities_final_SNE_year1[,"YEAR" := years[1:51]]


jaccard_overtime_SNE_year1comparison <- ggplot(data = dissimilarities_final_SNE_year1, aes(x=YEAR, y = jaccard_dissimilarity)) +
  geom_point() +
  geom_smooth(method = "lm") +
  stat_cor(label.y = 0.15, label.x = 1985, size = 8) +
  stat_regline_equation(label.y =0.2, label.x = 1985, size = 8) +
  ggtitle("c. SNE") +
  labs(x = "Year", y = "Jaccard dissimilarity") +
  theme_classic() +
  theme(text = element_text(size = 20))

ggsave(jaccard_overtime_SNE_year1comparison, file = "jaccard_overtime_SNE_year1comparison.eps" )


```

Looking at changes in jaccard dissimilarity across time for each region individually

```{r MAB jaccard across time}
  reduced_region <- neus_pres_abs[reg == "MAB",]
  reduced_region_wide <- dcast(reduced_region, YEAR + reg ~ SCINAME, value.var = "present") #long to wide data for community matrix, column names are SVSPP
  
  communitymatrix_MAB <- cbind(reduced_region_wide[,c(3:819)], reduced_region_wide[,c(1:2)]) #community matrix
  
  
  dissimilarities_MAB <- vegdist(communitymatrix_MAB[,c(1:817)], method = "jaccard") #dissimilarity matrix

  #reshape dissimilarities
  dissimilarities.l_MAB <- melt(as.matrix(dissimilarities_MAB), varnames = c("year1", "year2"))
  dissimilarities.l_MAB <- data.table(dissimilarities.l_MAB) #and then to data table
  
  
  #I only want year 2 to be less than year 1 (don't need all comparisons)
  values_key <- c(1, rep(2:51, each=2), 52)
  year_turnover_key <- matrix(values_key, nrow = 2)
  year_turnover_key <- data.table(year_turnover_key)
  year_turnover_key.t <- transpose(year_turnover_key)
  year_turnover_key.final <- data.table("year1" = year_turnover_key.t[,V2], "year2" = year_turnover_key.t[,V1])
  
  #combine key with dissimilarities.l_MAB
dissimilarities_final_MAB <- year_turnover_key.final[dissimilarities.l_MAB, on = c("year1","year2"), nomatch = 0]
  
  
colnames(dissimilarities_final_MAB)[colnames(dissimilarities_final_MAB)=="value"] <- "jaccard_dissimilarity"
#change column name to distance

dissimilarities_final_MAB[,"YEAR" := years[1:51]]


jpeg(file="jaccard_overtime_MAB.jpg")
    plot(x=dissimilarities_final_MAB$YEAR, y=dissimilarities_final_MAB$jaccard_dissimilarity, xlab = "Year", ylab="Jaccard dissimilarity", ylim = c(0:1), pch = 19)
 dev.off()


```

```{r MAB jaccard across time (compared to first year)}
  reduced_region <- neus_pres_abs[reg == "MAB",]
  reduced_region_wide <- dcast(reduced_region, YEAR + reg ~ SCINAME, value.var = "present") #long to wide data for community matrix, column names are SVSPP
  
  communitymatrix_MAB <- cbind(reduced_region_wide[,c(3:819)], reduced_region_wide[,c(1:2)]) #community matrix
  
  
  dissimilarities_MAB <- vegdist(communitymatrix_MAB[,c(1:817)], method = "jaccard") #dissimilarity matrix

  #reshape dissimilarities
  dissimilarities.l_MAB <- melt(as.matrix(dissimilarities_MAB), varnames = c("year1", "year2"))
  dissimilarities.l_MAB <- data.table(dissimilarities.l_MAB) #and then to data table
  
  
  #I only want to compare years with first year (don't need all comparisons)
  year_turnover_key.firstyear <- data.table("year1" = c(2:52), "year2" = 1)
  
  #combine key with dissimilarities.l_MAB
dissimilarities_final_MAB_year1 <- year_turnover_key.firstyear[dissimilarities.l_MAB, on = c("year1","year2"), nomatch = 0]
  
  
colnames(dissimilarities_final_MAB_year1)[colnames(dissimilarities_final_MAB_year1)=="value"] <- "jaccard_dissimilarity"
#change column name to distance

dissimilarities_final_MAB_year1[,"YEAR" := years[1:51]]

jaccard_overtime_MAB_year1comparison <- ggplot(data = dissimilarities_final_MAB_year1, aes(x=YEAR, y = jaccard_dissimilarity)) +
  geom_point() +
  geom_smooth(method = "lm") +
  stat_cor(label.y = 0.15, label.x = 1985, size = 8) +
  stat_regline_equation(label.y =0.2, label.x = 1985, size = 8) +
  ggtitle("d. MAB") +
  labs(x = "Year", y = "Jaccard dissimilarity") +
  theme_classic() +
  theme(text = element_text(size = 20))

ggsave(jaccard_overtime_MAB_year1comparison, file = "jaccard_overtime_MAB_year1comparison.eps" )

```

Simplified species richness through time
```{r spp richness through time}
library(ggplot2)
library(ggpubr)

GME_spp_richness <- rowSums(communitymatrix_GME[,1:817])
GEO_spp_richness <- rowSums(communitymatrix_GEO[,1:817])
SNE_spp_richness <- rowSums(communitymatrix_SNE[,1:817])
MAB_spp_richness <- rowSums(communitymatrix_MAB[,1:817])

spp_richness <- data.table("year" = years, "GME" = GME_spp_richness, "GEO" = GEO_spp_richness, "SNE" = SNE_spp_richness, "MAB" = MAB_spp_richness)

spp_richness.l <- melt(spp_richness, id.vars = c("year"),
                measure.vars = c("GME", "GEO", "SNE", "MAB"),
                value.name = "richness")

species_richness_reg <- ggplot(data = spp_richness.l, aes(x = year, y=richness)) +
  geom_line() +
  ylim(0,350) +
  geom_smooth(method = lm) +
  facet_wrap(~variable) +
  stat_cor(label.y = 0, label.x = 1990) +
  stat_regline_equation(label.y = 50, label.x = 1990) +
  theme_classic()




ggsave(species_richness_reg, file = "species_richness_reg.eps")
  
```

#Distance Decay with Stratum instead of region
Jaccard Dissimilarities Distance Decay all in one plot by STRATA
```{r stratum distance matrix}
neus_full.ID.uniquestrata <- unique(neus_full.ID, by = "STRATUM") #unique strata

neus_full.ID.uniquestrata.latlon <- neus_full.ID.uniquestrata[,c(1,6,7)] #only lat lon and stratum

stratum_distances <- distm(neus_full.ID.uniquestrata.latlon[,c(3, 2)]) #distance matrix, consistent across years
rownames(stratum_distances) <- neus_full.ID.uniquestrata.latlon$STRATUM
colnames(stratum_distances) <- neus_full.ID.uniquestrata.latlon$STRATUM

#reorient to long form 
stratum_distance.l <- melt(as.matrix(stratum_distances), varnames = c("strat1", "strat2")) #matrix to data frame, 
stratum_distance.l <- data.table(stratum_distance.l) #and then to data table
stratum_distance.l <- stratum_distance.l[strat1 >= strat2,] # to get rid of repetitions
colnames(stratum_distance.l)[colnames(stratum_distance.l)=="value"] <- "distance(m)" #change column name to distance

```
Above is distance info, now I need to build community matrix
```{r distance decay by strata}
neus_full.ID

#one PRESENT value for each YEAR, SEASON, STRATUM and SVSPP combination
neus_full.ID.stratum.unique <- na.omit(unique(neus_full.ID, by = c("YEAR", "STRATUM", "SVSPP")), cols = c("YEAR", "STRATUM", "SVSPP")) #unique STRATUM, year, species combinations, and get rid of any rows where any are NA

#reduce data table to just be svspp, year, sciname, comname, STRATUM

neus_full.ID.stratum.unique.r <- data.table((neus_full.ID.stratum.unique[, c("SVSPP", "SCINAME", "COMNAME", "YEAR", "STRATUM")]))

#add presence column 
neus_full.ID.stratum.unique.r[,"present" := 1]

#now I need to populate a comparison data table with every possible year, region, and species combination to add 0s
stratum <- as.integer(unique(neus_full.ID.stratum.unique.r$STRATUM)) 

table(neus_full.ID.stratum.unique.r$STRATUM, neus_full.ID.regions.unique.r$present)


full_set_stratum<- data.table(expand.grid("SVSPP" = spp.list, "YEAR" = years, "STRATUM" = stratum))

```
Bring in presences to expansion for presence by STRATUM
```{r bring in actual presences by STRATUM to expansion}
merge <- neus_full.ID.stratum.unique.r[full_set_stratum, on = c("YEAR", "STRATUM", "SVSPP")]
merge[,"present" := ifelse(is.na(SCINAME), 0,1)][,"SCINAME" := NULL][,"COMNAME" := NULL] #if scientific name and common name are empty, that is an absence, if populated, presence, also, we don't need these columns, as we'll bring them in below

#simplify spp key
#bring in spp key
spp_key

neus_pres_abs_STRATUM <- merge[spp_key, on = c("SVSPP")] #also don't actually need all columns

```
Loop for all STRATUM, YEAR, SVSPP combos
```{r jaccard disdissimilarities between STRATUM in multiple years}
distances_dissimilarities_allyears_STRATUM <- data.table("strat1" = integer(), "strat2" = integer(), "distance(m)" = numeric(),"jaccard_dissimilarity" = numeric(), YEAR = integer())

#Now loop through all years starting in 1967
for (i in 1:length(years)) {
  reduced_year <- neus_pres_abs_STRATUM[YEAR == years[i],]
  reduced_year_wide <- dcast(reduced_year, YEAR + STRATUM ~ SCINAME, value.var = "present") #long to wide data for community matrix, column names are SVSPP
  
  communitymatrix <- cbind(reduced_year_wide[,c(3:819)], reduced_year_wide[,c(1:2)]) #community matrix
  
#there are some STRATA with no presences or observations
#rowSums(communitymatrix.r[,c(1:817)]) #I guess I will get rid of those

communitymatrix.r <- communitymatrix[as.logical(rowSums(communitymatrix[,c(1:817)] != 0)), ] #reduce to only rows with atleast one species present
stratum.r <- communitymatrix.r$STRATUM

  
  dissimilarities <- vegdist(communitymatrix.r[,c(1:817)], method = "jaccard", binary = T) #dissimilarity 
  

#make into matrix
dissimilarities.m <- as.matrix(dissimilarities, labels=TRUE)
colnames(dissimilarities.m) <- rownames(dissimilarities.m) <- stratum.r

  #reshape dissimilarities
  dissimilarities.l <- melt(dissimilarities.m, varnames = c("strat1", "strat2"))
  dissimilarities.l <- data.table(dissimilarities.l) #and then to data table
  dissimilarities.l <- dissimilarities.l[strat1 >= strat2,] # to get rid of repetitions
  colnames(dissimilarities.l)[colnames(dissimilarities.l)=="value"] <- "jaccard_dissimilarity" #change column name to distance


#add year for these values
  dissimilarities.l[, "YEAR" := years[i]]

#merge distance with jaccard_dissimilarity for this year
  dissimilarities.l.full <- stratum_distance.l[dissimilarities.l, on = c("strat1", "strat2")]

distances_dissimilarities_allyears_STRATUM <- rbind(distances_dissimilarities_allyears_STRATUM, dissimilarities.l.full) #add to data table

#extract info for distance decay exponential model for each year
#reduce stratum_distances to stratum.r
stratum_distances.r <- stratum_distances[as.character(stratum.r), as.character(stratum.r)]

modname <- paste0(years[i], "_decay_mod_exponential")

assign(modname, decay.model(dissimilarities.m, stratum_distances.r, model.type = "exponential", y.type = "dissimilarities", perm = 1000))

#jpeg(file=paste0("jaccardplot_", years[i], ".jpg"))
    #plot(x=dissimilarities.l.full$`distance(m)`, y=dissimilarities.l.full$jaccard_dissimilarity, xlab = "Distance(m)", #ylab="Jaccard dissimilarity", ylim = c(0:1), pch = 19)
 #dev.off()
 
 #PCA
#vare.pca <- rda(communitymatrix[,c(1:817)])
#jpeg(file=paste0("jaccardpca_", years[i], ".jpg"))
#plot(vare.pca, display = "sites", xlim = c(-4,4), ylim = c(-4,4))
#dev.off()
  
}

summary(distances_dissimilarities_allyears_STRATUM) #here we have jaccard, distance, and 

save(distances_dissimilarities_allyears, file = "jaccard_similarities_distances_4reg.Rdata")
save(distances_dissimilarities_allyears_STRATUM, file = "jaccard_similarities_distances_stratum.Rdata")
```

Now, let's see what plots look like for distance decay over time

```{r distance decay all years}
distance_decay_stratum <- ggplot(data = distances_dissimilarities_allyears_STRATUM, aes(x = `distance(m)`, y = jaccard_dissimilarity, color = YEAR)) +
  geom_point() +
  theme_classic() +
  theme(text=element_text(size = 20))
distance_decay_stratum



ggsave(distance_decay_stratum, file = "distance_decay_stratum.eps")
```
Fitting distance decay models to pair-wise assemblage similarity: fits a negative exponential or power law function(GLM) describing decay of assemblage similarity or increase of dissimilarity with distance.


```{r distance decay model}
library(betapart)

#reduce stratum_distances to stratum.r
stratum_distances.r <- stratum_distances[as.character(stratum.r), as.character(stratum.r)]

mod_1968 <- decay.model(dissimilarities.m, stratum_distances.r, model.type = "exponential", y.type = "dissimilarities", perm = 1000)


```

