---
title: "Map of Surveys with Coefficient Values and Classifications"
author: Zoe Kitchel
date: October 19 2023
output: html_notebook
---

Figure 1 script part 2 for Kitchel et al. 2023 in prep taxonomic diversity manuscript.

Here, I plan to make a global map showing how each survey region is classified (homogenization, differentiation, no significant change.)
Run visualization_homogenization_differentiation.Rmd first!

```{r}
library(data.table)
library(ggplot2)
library(ggspatial) #basemap
#library(rgdal)
library(sp)
#library(rgeos)
library(viridis)
library(here)
library(maps)
library(mapproj)
library(concaveman)
library(dplyr)
library(purrr)
library(sf)
library(rnaturalearth)
library(rnaturalearthdata)


```

 Pull in palette and name helper
```{r}
source(here::here("analysis_code","color_links.R"))
```

Pull in data
```{r pull in data}
#trend
bray_curtis_total_coefs.r <- fread(here::here("output","bray_curtis_total_coefs.r.csv"))
bray_curtis_total_coefs.r[,`Significant Trends`:= ifelse(differentiating == 1, "differentiating",ifelse(homogenizing == 1,"homogenizing","no trend"))]

#raw data
FishGlob_clean <- readRDS(here::here("data", "cleaned", "FishGlob_clean.rds"))


```

Unique lat lon values
```{r unique lat lon}
unique_lat_lon_survey_units <- unique(FishGlob_clean[,.(survey_unit, survey, latitude, longitude)])
#shift longitude for mapping
unique_lat_lon_survey_units[,longitude_s := ifelse(longitude >160, longitude - 360, longitude)]

unique_lat_lon_survey_units.spdf <- copy(unique_lat_lon_survey_units)
#alternative to make spdf
coordinates(unique_lat_lon_survey_units.spdf) <- c(5,3)

unique_lat_lon_survey_units.spdf$long <- unique_lat_lon_survey_units$longitude_s
unique_lat_lon_survey_units.spdf$lat <- unique_lat_lon_survey_units$latitude

rm(FishGlob_clean)

```

Base map
```{r}

t <- map_data("world2",wrap=c(-200,160), ylim=c(-60,90))

ggplot(t) + geom_polygon(aes(x = long, y = lat, group = group)) + coord_map()


```

NEW basemap
```{r}
sf::sf_use_s2(FALSE) #polygons that cover more than half of the globe are inverted

#alt
world <- ne_countries(scale = "medium", returnclass = "sf")

robin_proj <-  "+proj=robin +lon_0=-45 +x_0=0 +y_0=0 +ellps=WGS84 +datum=WGS84 +units=m +no_defs"

#See here for fixing lines (https://stackoverflow.com/questions/56146735/visual-bug-when-changing-robinson-projections-central-meridian-with-ggplot2)
# define a long & slim polygon that overlaps the meridian line & set its CRS to match
# that of world
polygon <- st_polygon(x = list(rbind(c(134, 90),
                                     c(135, 90),
                                     c(135, -90),
                                     c(134, -90),
                                     c(134, 90)))) %>%
  st_sfc() %>%
  st_set_crs(4326)

# modify world dataset to remove overlapping portions with world's polygons
world2 <- world %>% st_difference(polygon)

# perform transformation on modified version of world dataset
world_robinson <- st_transform(world2, 
                               crs = '+proj=robin +lon_0=315 +x_0=0 +y_0=0 +ellps=WGS84 +datum=WGS84 +units=m +no_defs')

# plot
ggplot() +
  geom_sf(data = world_robinson)

bbox = st_bbox(world_robinson)
bbox[c(1,3)] = c(-1e-5,1e-5)
polygon2 <- st_as_sfc(bbox)

crosses = world_robinson %>%
  st_intersects(polygon2) %>%
  sapply(length) %>%
  as.logical %>%
  which

library(magrittr)
world_robinson[crosses,] %<>%
  st_buffer(0) 

ggplot(world_robinson) + geom_sf() 

```

Points and basemap
```{r}
basemap_points_color_survey_unit <- ggplot() +
  geom_polygon(data = t, aes(x = long, y = lat, group = group), fill = "lightgrey") +
  geom_point(data = data.frame(unique_lat_lon_survey_units.spdf),
             aes(x = long, y = lat, group = survey_unit, color = survey_unit), size = 0.001, alpha = 0.8) +
  scale_color_manual(values = color_link$hex) +
  coord_map() +
  theme_classic() +
  theme(legend.position = "null")

basemap_points_color_survey_unit

ggsave(basemap_points_color_survey_unit, path = here::here("figures"), filename = "basemap_points_color_survey_unit.pdf", height = 10, width = 20, unit = "in")
```

Convert points to simple feature, and then use concaveman to make corresponding polygons
```{r points to simple feature}
all_survey_units_sf <- unique_lat_lon_survey_units %>%
          st_as_sf(coords = c("longitude_s","latitude"), crs = 4326)
```


```{r concaveman to make polygons}

all_survey_units_polygons <- purrr::map(unique(all_survey_units_sf$survey_unit),
                       ~ concaveman(all_survey_units_sf[all_survey_units_sf$survey_unit %in% .,], concavity = 1, length_threshold = 2)
                       ) %>%
  map2(unique(all_survey_units_sf$survey_unit), ~ mutate(.x, survey_unit = .y)) %>%
  reduce(rbind)

#world simple feature polygon to extract land from these polygons
world <- rnaturalearth::ne_countries(type = "countries", scale="large", returnclass="sf")

#delete areas where this polygon overlaps with land
all_survey_units_polygons.clean <- rmapshaper::ms_erase(all_survey_units_polygons,world[1])

#find polygon centroid
all_survey_units_centroid_points <- sf::st_centroid(all_survey_units_polygons.clean)

#for all NZ points, convert longitudes
all_survey_units_centroid_points$geometry[all_survey_units_centroid_points$survey_unit %in% c("NZ-WCSI")]=st_point(c(171.4543-360, -41.7294))

all_survey_units_centroid_points$geometry[all_survey_units_centroid_points$survey_unit %in% c("NZ-ECSI")]=st_point(c(172.5993-360, -44.17959))

all_survey_units_centroid_points$geometry[all_survey_units_centroid_points$survey_unit %in% c("NZ-CHAT")]=st_point(c(178.9788-360, -43.58499))

all_survey_units_centroid_points$geometry[all_survey_units_centroid_points$survey_unit %in% c("NZ-SUBA")]=st_point(c(170.3432-360, -50.50603))

#for all seasonal repeats, move duplicates a bit up and down
#BITS
all_survey_units_centroid_points$geometry[all_survey_units_centroid_points$survey_unit == "BITS-4"]=st_point(c(15.98157, 57.6))

#NIGFS
all_survey_units_centroid_points$geometry[all_survey_units_centroid_points$survey_unit == "NIGFS-4"]=st_point(c(-4.910924, 55))

#NS-IBTS
all_survey_units_centroid_points$geometry[all_survey_units_centroid_points$survey_unit == "NS-IBTS-3"]=st_point(c(2.856672, 58.2))

#SWC-IBTS
all_survey_units_centroid_points$geometry[all_survey_units_centroid_points$survey_unit == "SWC-IBTS-4"]=st_point(c(-7.493093, 59.1))

#GMEX
all_survey_units_centroid_points$geometry[all_survey_units_centroid_points$survey_unit == "GMEX-Summer"]=st_point(c(-93.15577, 30))

#NEUS
all_survey_units_centroid_points$geometry[all_survey_units_centroid_points$survey_unit == "NEUS-Spring"]=st_point(c(-70.58172, 42.2))

#SEUS
all_survey_units_centroid_points$geometry[all_survey_units_centroid_points$survey_unit == "SEUS-summer"]=st_point(c(-78.27192, 32))

all_survey_units_centroid_points$geometry[all_survey_units_centroid_points$survey_unit == "SEUS-fall"]=st_point(c(-78.27192, 30.5))


#plot_polygons with points
ggplot() +
  geom_sf(data = all_survey_units_polygons.clean, aes(fill = survey_unit), color = NA) +
  geom_sf(data = all_survey_units_centroid_points, aes(fill = survey_unit)) +
  scale_fill_manual(values = color_link$hex) +
  theme_classic()
```

Link with name helper
```{r}
all_survey_units_polygons.names <- left_join(all_survey_units_polygons.clean, color_link, by = "survey_unit")
```


#Combine basemap and these regions
```{r combine basemap and survey regions}
(survey_unit_filled_polygons <- ggplot() +
  geom_sf(data = all_survey_units_polygons.names, aes(fill = reorder(Survey_Name_Season,survey_unit)), color = NA, alpha = 0.3) +
  geom_polygon(data = t, aes(x = long, y = lat, group = group), fill = "lightgrey") +
  scale_fill_manual(values = color_link$hex) +
  coord_sf() +
  lims(x = c(-200,60)) +
  theme_classic() +
  theme(axis.text = element_blank(), axis.line = element_blank(), axis.ticks = element_blank(), axis.title = element_blank()))

ggsave(survey_unit_filled_polygons, path = here::here("figures"), filename = "survey_unit_filled_polygons.pdf", height =10, width = 18, unit = "in")

#polygon outlines only
(survey_unit_outline_polygons <- ggplot() +
  geom_sf(data = all_survey_units_polygons.names, aes(color = reorder(Survey_Name_Season,survey_unit)), fill = NA, alpha = 0.5, linewidth = 0.1) +
  geom_polygon(data = t, aes(x = long, y = lat, group = group), fill = "lightgrey") +
  scale_color_manual(values = color_link$hex) +
  coord_sf() +
  lims(x = c(-200,60)) +
  theme_classic() +
  theme(axis.text = element_blank(), axis.line = element_blank(), axis.ticks = element_blank(), axis.title = element_blank()))

ggsave(survey_unit_outline_polygons, path = here::here("figures"), filename = "survey_unit_outline_polygons.pdf", height =10, width = 18, unit = "in")
```

Base theme (by Chris)
```{r}
# Base theme
base_theme <- theme(axis.text=element_text(size=7),
                    axis.text.y = element_text(angle = 90, hjust = 0.5),
                    axis.title=element_text(size=8),
                    legend.text=element_text(size=7),
                    legend.title=element_text(size=8),
                    plot.tag=element_text(size=9),
                    # Gridlines
                    panel.grid.major = element_blank(), 
                    panel.grid.minor = element_blank(),
                    panel.background = element_blank(), 
                    axis.line = element_line(colour = "black"),
                    # Legend
                    legend.background = element_rect(fill=alpha('blue', 0)))
```

Scotland shelf sea should be patterned: green and purple
Southeast US should be patterned: orange and purple
Pull each out separately
```{r}
setorder(all_survey_units_polygons.names, survey_unit)
scotland <- all_survey_units_polygons.names[38,]
southeast_us <- all_survey_units_polygons.names[35,]
#remove these from full 
all_survey_units_polygons.names <- all_survey_units_polygons.names[c(1:34,40:42),]


```


Instead color by change experienced
```{r color map by change experienced}
#merge polygons with stats
all_survey_units_polygons_stats <- left_join(all_survey_units_polygons.names, bray_curtis_total_coefs.r, by = "survey_unit")

#merge points with stats
all_survey_units_centroid_points_stats <- left_join(all_survey_units_centroid_points, bray_curtis_total_coefs.r, by = "survey_unit")

#focal regions
centroids_focal <- all_survey_units_centroid_points_stats %>%
  filter(survey_unit == "NZ-WCSI" |survey_unit == "SEUS-summer"|survey_unit == "ICE-GFS") %>%
  mutate(labels = c("d.","b.","c.")) #add labels

#edit lat lon a bit to help with visualization
centroids_focal$geometry[centroids_focal$survey_unit == "NZ-WCSI"]=st_point(c(-192, -41))
centroids_focal$geometry[centroids_focal$survey_unit == "SEUS-summer"]=st_point(c(-73, 33))
centroids_focal$geometry[centroids_focal$survey_unit == "ICE-GFS"]=st_point(c(-27, 63))

```

Change all projections to robin
```{r proj to robin}
all_survey_units_polygons_stats.robin <- st_transform(all_survey_units_polygons_stats, crs = robin_proj)
scotland.robin <- st_transform(scotland, crs = robin_proj)
southeast_us.robin <- st_transform(southeast_us, crs = robin_proj)
all_survey_units_centroid_points_stats.robin <- st_transform(all_survey_units_centroid_points_stats, crs = robin_proj)
centroids_focal.robin <- st_transform(centroids_focal, crs = robin_proj)
```





```{r figure 1 map}

map_bray_coef_trend <- ggplot() +
    #world map
  geom_sf(data = world_robinson, fill = "lightgrey", color="white", lwd=0.1) +
  #polygons
  geom_sf(data = all_survey_units_polygons_stats.robin, aes(fill = `Significant Trends`, color = `Significant Trends`), alpha = 0.7, lwd = 0.1) +
  #add scotland alone
  geom_sf(data = scotland.robin, fill = "#1E88E5", color = "#FFC109") +
  #add southeast US alone
  geom_sf(data = southeast_us.robin, fill = "#1E88E5", color = "#D81C60", lwd = 0.001) +
  #polygon outlines
#  geom_sf(data = all_survey_units_polygons_stats, aes(color = `Significant Trends`), linewidth = 0.1, alpha = 1, lwd = 0.4, fill = NA) +
  #colored points with trend assocaited with each survey
  geom_sf(all_survey_units_centroid_points_stats.robin, mapping = aes(fill = `Significant Trends`), size=1.5, pch=21, color="black", lwd = 0.05) +
 # geom_sf_text(data = all_survey_units_centroid_points_stats, aes(label = survey_unit), size = 0.5) +
  scale_fill_manual(values = c("#FFC109","#D81C60","#1E88E5"),
                    labels = c("Differentiation","Homogenization","No trend in dissimilarity")) +
  scale_color_manual(values = c("#FFC109","#D81C60","#1E88E5"),
                     labels = c("Differentiation","Homogenization","No trend in dissimilarity")) +
    # Labels
  scale_x_continuous(breaks=seq(-180, 50, 30)) +
  scale_y_continuous(breaks=seq(-90, 90, 30)) +
  labs(x = "Longitude",y = "Latitude") +
  #Annotate b,c,d from lower panel (b = W coast south island NZ, c = Southeast US Summer, d = Scotian Shelf Canada)
  geom_sf_text(centroids_focal.robin, mapping = aes(label = labels, fontface = "bold"), size = 3) +
  # Crop
  coord_sf(y=c(-6625125,8340316), x=c(-14994440,8000000), expand = F) +
 # lims(x = c(-200,60)) +
  guides(fill = guide_legend(title = "Dissimilarity trend"),
         color = guide_legend(title = "Dissimilarity trend")) +
  theme_bw() +
  base_theme +
    theme(legend.position = c(0.2, 0.4),
        legend.key.size = unit(0.6, "cm"),
        legend.text = element_text(size = 12), legend.title = element_text(size = 12),
        axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank())

map_bray_coef_trend

ggsave(map_bray_coef_trend, path = here::here("figures"), filename = "map_bray_coef_trend.pdf", height = 4, width = 8, unit = "in")
ggsave(map_bray_coef_trend, path = here::here("figures"), filename = "map_bray_coef_trend.jpg", height = 4, width = 8, unit = "in")
```

Merge completed map with visualizations of NMDS
```{r}
library(cowplot)

#load extra plots
mds_contrast <- readRDS(here::here("figures","mds_contrast.rds"))

merge_nmds_map <- plot_grid(
  map_bray_coef_trend + theme(axis.text.y = element_blank(), plot.margin = unit(c(0,0.2,0,0), "cm")),
  mds_contrast + theme(plot.margin = unit(c(0,1,0,0), "cm")), 
  ncol = 1, nrow = 2, labels = c("a.",""), label_size = 12, label_face = "bold",
  rel_heights = c(2,1)
)

merge_nmds_map <- ggdraw(xlim = c(0,10), ylim = c(0,20)) +
  draw_plot(x = 0, y = 7,map_bray_coef_trend + theme(axis.text.y = element_blank(),plot.margin = unit(c(0,0.2,0,0), "cm")), width = 10, height = 14) +
  draw_plot(x = 0, y = 0,mds_contrast + theme(plot.margin = unit(c(0,1,0,0), "cm")), width = 10, height = 8.5) +
  geom_text(x = 0.2, y = 19.5,aes(label = "a.", fontface = "bold"))


ggsave(merge_nmds_map, path = here::here("figures"), filename = "merge_nmds_map.jpg", height = 8, width = 7, unit = "in")

```
