---
title: "R Notebook"
output: html_notebook
---

Are our data autocorrelated?

This doesn't seem to be done, check manuscript for how I did this.

Maybe, avg residual per year and region?

```{r setup}
library(data.table)
library(ggplot2)
library(nlme)
library(here)
library(lme4)
```

Load Data
```{r}
distances_dissimilarities_allyears.r <- readRDS(here::here("output","distance_decay","distances_dissimilarities_allyears.r.rds"))

#survey_unit = factor
distances_dissimilarities_allyears.r[,survey_unit := factor(survey_unit)]

#add new variable for year in sequence per region
distances_dissimilarities_allyears.r[,first_year := min(year),.(survey_unit)]
distances_dissimilarities_allyears.r[,last_year := max(year),.(survey_unit)]

distances_dissimilarities_allyears.r[,year_in_seq := year-first_year+1]

distances_dissimilarities_allyears.r[,years_sampled := last_year-first_year+1]

survey_unit_sampling_years <- unique(distances_dissimilarities_allyears.r[,.(survey_unit, years_sampled)])
```

Linear Model
```{r}
bray_curtis_dissimilarity_balanced_mean_year_lm_mod <- lm(bray_curtis_dissimilarity_balanced_mean ~ year, data = distances_dissimilarities_allyears.r)
summary(bray_curtis_dissimilarity_balanced_mean_year_lm_mod)

saveRDS(bray_curtis_dissimilarity_balanced_mean_year_lm_mod, here::here("output","distance_dissimilarity","bray_curtis_dissimilarity_balanced_mean_year_mod.rds"))

```
Mod Summary

Call:
lm(formula = bray_curtis_dissimilarity_balanced_mean ~ year, 
    data = distances_dissimilarities_allyears.r)

Residuals:
     Min       1Q   Median       3Q      Max 
-0.41846 -0.09578  0.01572  0.09168  0.22807 

Coefficients:
              Estimate Std. Error t value Pr(>|t|)    
(Intercept)  4.6356456  0.8827028   5.252 1.94e-07 ***
year        -0.0020056  0.0004402  -4.556 6.04e-06 ***
---
Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1

Residual standard error: 0.1207 on 792 degrees of freedom
Multiple R-squared:  0.02554,	Adjusted R-squared:  0.02431 
F-statistic: 20.76 on 1 and 792 DF,  p-value: 6.042e-06

Now, add autocorrelation for time

```{r}
bray_curtis_dissimilarity_balanced_mean_year_ac_mod <- gls(bray_curtis_dissimilarity_balanced_mean ~ year, data=distances_dissimilarities_allyears.r, 
              correlation = corAR1(form=~year | survey_unit),
              na.action=na.omit)
summary(bray_curtis_dissimilarity_balanced_mean_year_ac_mod)

saveRDS(bray_curtis_dissimilarity_balanced_mean_year_ac_mod, here::here("output","distance_dissimilarity","bray_curtis_dissimilarity_balanced_mean_year_ac_mod.rds"))

```

Compare the two models
```{r}
AIC(bray_curtis_dissimilarity_balanced_mean_year_lm_mod, bray_curtis_dissimilarity_balanced_mean_year_ac_mod)
```

I should now include mixed effect which allows the intercept to vary across survey
```{r}
intercept_varies <- lmer(bray_curtis_dissimilarity_balanced_mean ~ year + (1|survey_unit), data = distances_dissimilarities_allyears.r)

summary(intercept_varies)

saveRDS(intercept_varies, here::here("output","distance_dissimilarity","intercept_varies.rds"))
```
Warning in checkConv(attr(opt, "derivs"), opt$par, ctrl = control$checkConv,  :
  unable to evaluate scaled gradient
Warning in checkConv(attr(opt, "derivs"), opt$par, ctrl = control$checkConv,  :
  Model failed to converge: degenerate  Hessian with 1 negative eigenvalues


Confidence intervals
```{r}
confint(intercept_varies)
```

See cofficients
```{r}
# see group coefficients
intercept_varies_model_coefs <- coef(intercept_varies)$Group %>% 
  rename(Intercept = `(Intercept)`, Slope = Days) %>% 
  rownames_to_column("Group")

# see coefficients
intercept_varies_model_coefs

#join  with OG data
sleep_groups_rani <- left_join(sleep_groups, model_coefs, by = "Group")
```

Plot
```{r}

```


Predict
```{r}
predict_with_random_intercept <- predict(intercept_varies)
```


Finally, I will allow mixed effects which allow the intercept and the slope to vary across survey
```{r}
system.time(intercept_slope_varies <- lmer(bray_curtis_dissimilarity_balanced_mean ~ year_in_seq + (1 + year_in_seq|survey_unit), data = distances_dissimilarities_allyears.r))

saveRDS(intercept_slope_varies, here::here("output","distance_dissimilarity","intercept_slope_varies.rds"))

```
NB:

Warning in checkConv(attr(opt, "derivs"), opt$par, ctrl = control$checkConv,  :
  Model failed to converge with max|grad| = 0.0367592 (tol = 0.002, component 1)
Warning in checkConv(attr(opt, "derivs"), opt$par, ctrl = control$checkConv,  :
  Model is nearly unidentifiable: very large eigenvalue
 - Rescale variables?
 
 
Plot average residuals for each year/region over time? Per Malin's recommendation: https://goodekat.github.io/redres/
```{r}
install.packages("redres")
library(redres)


```

Plot models
```{r}
library(sjPlot)
library(sjmisc)

install.packages("sjPlot")
install.packages("sjmisc")
```

Plot
```{r}
model_coefs <- coef(intercept_slope_varies)$survey_unit
model_coefs$survey_unit <- rownames(model_coefs)

model_coefs <- data.table(model_coefs)

#merge with duration of survey
model_coefs_length <- model_coefs[survey_unit_sampling_years, on = survey_unit]
```

```{r}


ggplot2(data = model_coefs, aes(x = survey_unit, y = year_in_seq, label = survey_unit)) +
  geom_point(stat = 'identity', aes(year_)) +
  scale_size(range = c(2, 8), breaks = 5)
  scale_size_manual
  
  
  # Plot
ggplot(mtcars, aes(x=`car name`, y=mpg_z, label=mpg_z)) + 
  geom_point(stat='identity', aes(col=mpg_type), size=6)  +
  scale_color_manual(name="Mileage", 
                     labels = c("Above Average", "Below Average"), 
                     values = c("above"="#00ba38", "below"="#f8766d")) + 
  geom_text(color="white", size=2) +
  labs(title="Diverging Dot Plot", 
       subtitle="Normalized mileage from 'mtcars': Dotplot") + 
  ylim(-2.5, 2.5) +
  coord_flip()
  
```

