---
title: "Calculating Dissimilarity Metrics Across Space and Time for top 85% of species"
output: html_notebook
---
Script 3b for Kitchel et al. 2023 in prep taxonomic diversity manuscript.

```{r setup}
library(data.table)
library(vegan)
library(sf)
library(concaveman) #polygon around points
library(betapart) #allows us to partition beta diversity
library(geosphere)
library(ggpubr) #stat_regline_equation
library(nlme)


FishGlob_clean <- readRDS(here::here("data", "cleaned", "FishGlob_clean.rds"))


```

Loop through all regions (takes  a few hours to run)

```{r list of all survey season combinations}
all_survey_units <- unique(FishGlob_clean.noNA[, survey_unit])
```

Delete 15% of least abundant species by biomass (across whole time series)
```{r}
#add new columns that sum biomass for each individual species per survey, and 
FishGlob_clean.noNA[survey_unit != "MEDITS",abundance_spp_sum := sum(wgt_cpue),.(accepted_name, survey_unit)][survey_unit != "MEDITS",abundance_sum := sum(wgt_cpue),survey_unit][survey_unit == "MEDITS",abundance_spp_sum := sum(num_cpue),.(accepted_name, survey_unit)][survey_unit == "MEDITS",abundance_sum := sum(num_cpue),survey_unit]



FishGlob_abundance_survey <- unique(FishGlob_clean.noNA[,.(survey_unit, accepted_name, abundance_spp_sum, abundance_sum)])

FishGlob_abundance_survey[,perc_spp_total_abundance := abundance_spp_sum/abundance_sum][,spp_count := uniqueN(accepted_name),survey_unit][,spp_count15_percent := round(0.15*spp_count,0)]

#rank 15% least abundant spp 
#rank by abundance
setkey(FishGlob_abundance_survey, perc_spp_total_abundance)

FishGlob_abundance_survey[, rank_abundance_bysurvey := frank(perc_spp_total_abundance), survey_unit]

#code each spp as 90% most common by abundance or 15% lease common by abundance in a survey
FishGlob_abundance_survey[, exclude_spp := ifelse(rank_abundance_bysurvey <= spp_count15_percent, T,F)]

#spp and survey key
spp_survey_keep_key <- unique(FishGlob_abundance_survey[,.(survey_unit, accepted_name, exclude_spp)])

FishGlob_clean.noNA_15perc <- spp_survey_keep_key[FishGlob_clean.noNA, on = .(survey_unit, accepted_name)]

FishGlob_clean.noNA_15perc_excluded <- FishGlob_clean.noNA_15perc[exclude_spp == F,]

#drops 
nrow(FishGlob_clean.noNA)-nrow(FishGlob_clean.noNA_15perc_excluded)
#4530/2759954 rows (0.2% of observations)
```

Calculate Beta Dissimilarity between all sampling sites within every survey unit
```{r loop through all regions and years}
survey_stats <- data.table(survey = character() ,  survey_unit = character(),  spp_num = numeric(),  study_period = numeric(),  study_duration = numeric(),  lat_range = numeric(),  mid_lat = numeric(),  lon_range = numeric(),  area = numeric(),  depth_range = numeric(),   mid_depth = numeric())

distances_dissimilarities_allyears_15perc_excluded_15perc_excluded <- data.table("haul_id1" = integer(), 
                                                 "haul_id2" = integer(), 
                                                 "distance" = numeric(), 
                                                 "bray_curtis_dissimilarity_balanced" = numeric(), 
                                                 year = integer(), 
                                                 abund_biomass  =  character(), 
                                                 "jaccard_dissimilarity_turnover" = numeric(), 
                                                 "jaccard_dissimilarity_nestedness" =  numeric(), 
                                                 "bray_curtis_dissimilarity_gradient" = as.numeric(), 
                                                 "jaccard_dissimilarity_total" = numeric(), 
                                                 "bray_curtis_dissimilarity_total" = numeric(), 
                                                 survey = character(), 
                                                 survey_unit = character())

for (i in 1:length(all_survey_units)) {
  
  FishGlob_clean_subset <- FishGlob_clean.noNA_15perc_excluded[survey_unit == all_survey_units[i], ]
  
  #map
  ####unique lat lon
  #add column with adjusted longitude for few surveys that cross dateline (NZ-CHAT and AI)
  FishGlob_clean_subset[,longitude_adj := ifelse((survey_unit %in% c("AI","NZ-CHAT") & longitude > 0),longitude-360,longitude)]

    lat_lon <- unique(FishGlob_clean_subset[, .(latitude,  longitude_adj)])

  pts <- st_as_sf(lat_lon,  coords=c('longitude_adj', 'latitude'),  crs=4326 )

  conc <- concaveman(pts,  concavity = 5)
  sf::sf_use_s2(FALSE) #turn off the s2 processing; GEOS treats projected coordinates as planar (i.e. two points lie on a line of infinite max lenght) while s2 is more "correct" (two points lie on a great circle of circumference of 40 075 kilometers)
  area <- st_area(conc) #m2,  check this later

  #fill row
 row <- data.table(FishGlob_clean_subset[1, survey], 
                        FishGlob_clean_subset[1, as.character(survey_unit)], 
                        as.numeric(length(unique(FishGlob_clean_subset[, accepted_name]))), 
                        as.numeric(max(FishGlob_clean_subset[, year])-min(FishGlob_clean_subset[, year])), 
                        as.numeric(length(unique(FishGlob_clean_subset[, year]))), 
                        as.numeric(max(FishGlob_clean_subset[, latitude])-min(FishGlob_clean_subset[, latitude])), 
                        as.numeric(mean(FishGlob_clean_subset[, latitude])), 
                        as.numeric(max(FishGlob_clean_subset[, longitude_adj])-min(FishGlob_clean_subset[, longitude_adj])), 
                        as.numeric(area), 
                        as.numeric(max(FishGlob_clean_subset[, as.numeric(depth)],  na.rm = T)-min(FishGlob_clean_subset[, as.numeric(depth)],  na.rm = T)), 
                   as.numeric(mean(FishGlob_clean_subset[, as.numeric(depth)],  na.rm = T)))
 
   survey_stats <- rbind(survey_stats,  row,  use.names = F)
   
   #list years
  FishGlob_clean_subset[, year:= as.numeric(year)] #make numeric
  setorder(FishGlob_clean_subset,  year)
  years <- unique(FishGlob_clean_subset[, year])
  
  #haul id keys
  haul_ids <- unique(FishGlob_clean_subset[, haul_id])
  haul_ids_key <- data.table(haul_id = haul_ids,  key_ID = seq(1, length(haul_ids),  by = 1))
  
  
  #convert haul_ids to numeric key_ids
  FishGlob_clean_subset <- FishGlob_clean_subset[haul_ids_key,  on = "haul_id"]

          for (j in 1:length(years)) {
            reduced_year <- FishGlob_clean_subset[year == years[j], ]
            
            #distances among cells
            setorder(reduced_year,  key_ID)
            
            latitude_longitude_haul_id <- unique(reduced_year[, .(latitude, longitude, key_ID)])
            distances <- distm(latitude_longitude_haul_id[, .(longitude, latitude)])
            key_IDs <- latitude_longitude_haul_id[, key_ID]
          
            colnames(distances) <- rownames(distances) <- key_IDs
          
            #wide to long
            haul_id_distances.l <- reshape2::melt(distances, varnames = (c("haul_id1",  "haul_id2")),  value.name = "distance")
            
            #make into data table
            haul_id_distances.l <- data.table(haul_id_distances.l)
          
          if(!(reduced_year[1, survey] == "MEDITS")) {
            
            #if some rows have wgt_cpue missing,  get rid of these rows
            reduced_year <- reduced_year[complete.cases(reduced_year[, wgt_cpue]), ]
            
              reduced_year_wide <- dcast(reduced_year,  key_ID + year ~ accepted_name,  value.var = "wgt_cpue",  fun.aggregate = sum) #longitude to wide data for community matrix,  column names are cell then species
              
              
              ncols <- ncol(reduced_year_wide)
              communitymatrix <- reduced_year_wide[, 3:ncols] #community matrix
              communitymatrix.occurence <- communitymatrix
              communitymatrix.occurence[communitymatrix.occurence > 0] <- 1
            
              #list of haul_id keys
              key_IDs_subset <- reduced_year_wide[,key_ID]
            
              dissimilarities_abundance <- beta.pair.abund(communitymatrix,  index.family = "bray") #dissimilarity 
              dissimilarities_occurrence <- beta.pair(communitymatrix.occurence,  index.family = "jaccard") #dissimilarity
            
              #make into matrix
              dissimilarities_abundance_balanced.m <- as.matrix(dissimilarities_abundance$beta.bray.bal,  labels=TRUE) #bal = balanced
              dissimilarities_abundance_gradient.m <- as.matrix(dissimilarities_abundance$beta.bray.gra,  labels=TRUE) #gra = gradient
              dissimilarities_abundance_total.m <- as.matrix(dissimilarities_abundance$beta.bray,  labels=TRUE) #total
              
              dissimilarities_occurrence_turnover.m <- as.matrix(dissimilarities_occurrence$beta.jtu,  labels=TRUE) #jtu = turnover
              dissimilarities_occurrence_nestedness.m <- as.matrix(dissimilarities_occurrence$beta.jne,  labels=TRUE) #jne = nestedness
              dissimilarities_occurrence_total.m <- as.matrix(dissimilarities_occurrence$beta.jac,  labels=TRUE) #total
              
              colnames(dissimilarities_abundance_balanced.m) <- rownames(dissimilarities_abundance_balanced.m) <- key_IDs_subset
              colnames(dissimilarities_abundance_gradient.m) <- rownames(dissimilarities_abundance_gradient.m) <- key_IDs_subset
              colnames(dissimilarities_abundance_total.m) <- rownames(dissimilarities_abundance_total.m) <- key_IDs_subset
              colnames(dissimilarities_occurrence_turnover.m) <- rownames(dissimilarities_occurrence_turnover.m) <- key_IDs_subset
              colnames(dissimilarities_occurrence_nestedness.m) <- rownames(dissimilarities_occurrence_nestedness.m) <- key_IDs_subset
              colnames(dissimilarities_occurrence_total.m) <- rownames(dissimilarities_occurrence_total.m) <- key_IDs_subset
            
              #reshape dissimilarities
              dissimilarities_abundance_balanced.l <- reshape2::melt(dissimilarities_abundance_balanced.m,  varnames = c("haul_id1",  "haul_id2"),  value.name = "bray_curtis_dissimilarity_balanced")
              dissimilarities_abundance_gradient.l <- reshape2::melt(dissimilarities_abundance_gradient.m,  varnames = c("haul_id1",  "haul_id2"),  value.name = "bray_curtis_dissimilarity_gradient")
              dissimilarities_abundance_total.l <- reshape2::melt(dissimilarities_abundance_total.m,  varnames = c("haul_id1",  "haul_id2"),  value.name = "bray_curtis_dissimilarity_total")
              
              dissimilarities_occurrence_turnover.l <- reshape2::melt(dissimilarities_occurrence_turnover.m,  varnames = c("haul_id1",  "haul_id2"),  value.name = "jaccard_dissimilarity_turnover")
              dissimilarities_occurrence_nestedness.l <- reshape2::melt(dissimilarities_occurrence_nestedness.m,  varnames = c("haul_id1",  "haul_id2"),  value.name = "jaccard_dissimilarity_nestedness")
              dissimilarities_occurrence_total.l <- reshape2::melt(dissimilarities_occurrence_total.m,  varnames = c("haul_id1",  "haul_id2"),  value.name = "jaccard_dissimilarity_total")
              
              #and then to data table format
              dissimilarities_abundance_balanced.l <- data.table(dissimilarities_abundance_balanced.l)
              dissimilarities_abundance_gradient.l <- data.table(dissimilarities_abundance_gradient.l)
              dissimilarities_abundance_total.l <- data.table(dissimilarities_abundance_total.l)
              dissimilarities_occurrence_turnover.l <- data.table(dissimilarities_occurrence_turnover.l)
              dissimilarities_occurrence_nestedness.l <- data.table(dissimilarities_occurrence_nestedness.l)
              dissimilarities_occurrence_total.l <- data.table(dissimilarities_occurrence_total.l)
            
              #add year for these values
              dissimilarities_abundance_balanced.l[,  "year" := years[j]]
              dissimilarities_occurrence_turnover.l[,  "year" := years[j]]
              
              
              dissimilarities_abundance_gradient.l[,  "year" := years[j]]
              dissimilarities_occurrence_nestedness.l[,  "year" := years[j]]
              
              dissimilarities_abundance_total.l[,  "year" := years[j]]
              dissimilarities_occurrence_total.l[,  "year" := years[j]]
              
              #what type of metric?
              dissimilarities_abundance_balanced.l[, "abund_biomass" := "biomass"]
              dissimilarities_occurrence_turnover.l[, "abund_biomass" := "biomass"]
              
              dissimilarities_abundance_gradient.l[, "abund_biomass" := "biomass"]
              dissimilarities_occurrence_nestedness.l[, "abund_biomass" := "biomass"]
              
              dissimilarities_abundance_total.l[, "abund_biomass" := "biomass"]
              dissimilarities_occurrence_total.l[, "abund_biomass" := "biomass"]
            
              #merge distance with dissimilarity for this year with both metrics of dissimilarity
              dissimilarities_full <- haul_id_distances.l[dissimilarities_abundance_balanced.l,  on = c("haul_id1",  "haul_id2")]
              dissimilarities_full <- dissimilarities_full[dissimilarities_occurrence_turnover.l,  on = c("haul_id1",  "haul_id2",  "year",  "abund_biomass")]
              
              dissimilarities_full <- dissimilarities_full[dissimilarities_occurrence_total.l,  on = c("haul_id1",  "haul_id2",  "year",  "abund_biomass")]
              
              dissimilarities_full <- dissimilarities_full[dissimilarities_occurrence_nestedness.l,  on = c("haul_id1",  "haul_id2",  "year",  "abund_biomass")]
              
              dissimilarities_full <- dissimilarities_full[dissimilarities_abundance_gradient.l,  on = c("haul_id1",  "haul_id2",  "year",  "abund_biomass")]
              
              dissimilarities_full <- dissimilarities_full[dissimilarities_abundance_total.l,  on = c("haul_id1",  "haul_id2",  "year",  "abund_biomass")]
              
              #add survey survey unit
              dissimilarities_full[, "survey" := FishGlob_clean_subset[1, survey]]
              dissimilarities_full[, "survey_unit" := all_survey_units[i]]
              
                            #delete if haul_id2 < haul_id1 (deletes comparisons between same site (0 inflated) and any duplicate comparisons)
              dissimilarities_full <- dissimilarities_full[haul_id2<haul_id1]
            
              #add to data table
              distances_dissimilarities_allyears_15perc_excluded <- rbind(distances_dissimilarities_allyears_15perc_excluded,  dissimilarities_full)
              
              print(paste0(j, "/", length(years)))
            
          } else { #if we do using abundance instead
            
            #if some rows have num_cpue missing,  get rid of these rows
            reduced_year <- reduced_year[complete.cases(reduced_year[, num_cpue]), ]
            
              reduced_year_wide <- dcast(reduced_year,  key_ID + year ~ accepted_name,  value.var = "num_cpue",  fun.aggregate = sum) #longitude to wide data for community matrix,  column names are cell then species
              
              
              ncols <- ncol(reduced_year_wide)
              communitymatrix <- reduced_year_wide[, 3:ncols] #community matrix
              communitymatrix.occurence <- communitymatrix
              communitymatrix.occurence[communitymatrix.occurence > 0] <- 1
            
              #list of haul_id keys
              key_IDs_subset <- reduced_year_wide[,key_ID]
            
              dissimilarities_abundance <- beta.pair.abund(communitymatrix,  index.family = "bray") #dissimilarity 
              dissimilarities_occurrence <- beta.pair(communitymatrix.occurence,  index.family = "jaccard") #dissimilarity
            
              #make into matrix
              dissimilarities_abundance_balanced.m <- as.matrix(dissimilarities_abundance$beta.bray.bal,  labels=TRUE) #bal = balanced
              dissimilarities_abundance_gradient.m <- as.matrix(dissimilarities_abundance$beta.bray.gra,  labels=TRUE) #gra = gradient
              dissimilarities_abundance_total.m <- as.matrix(dissimilarities_abundance$beta.bray,  labels=TRUE) #total
              
              dissimilarities_occurrence_turnover.m <- as.matrix(dissimilarities_occurrence$beta.jtu,  labels=TRUE) #jtu = turnover
              dissimilarities_occurrence_nestedness.m <- as.matrix(dissimilarities_occurrence$beta.jne,  labels=TRUE) #jne = nestedness
              dissimilarities_occurrence_total.m <- as.matrix(dissimilarities_occurrence$beta.jac,  labels=TRUE) #total
              
              colnames(dissimilarities_abundance_balanced.m) <- rownames(dissimilarities_abundance_balanced.m) <- key_IDs_subset
              colnames(dissimilarities_abundance_gradient.m) <- rownames(dissimilarities_abundance_gradient.m) <- key_IDs_subset
              colnames(dissimilarities_abundance_total.m) <- rownames(dissimilarities_abundance_total.m) <- key_IDs_subset
              colnames(dissimilarities_occurrence_turnover.m) <- rownames(dissimilarities_occurrence_turnover.m) <- key_IDs_subset
              colnames(dissimilarities_occurrence_nestedness.m) <- rownames(dissimilarities_occurrence_nestedness.m) <- key_IDs_subset
              colnames(dissimilarities_occurrence_total.m) <- rownames(dissimilarities_occurrence_total.m) <- key_IDs_subset
            
              #reshape dissimilarities
              dissimilarities_abundance_balanced.l <- reshape2::melt(dissimilarities_abundance_balanced.m,  varnames = c("haul_id1",  "haul_id2"),  value.name = "bray_curtis_dissimilarity_balanced")
              dissimilarities_abundance_gradient.l <- reshape2::melt(dissimilarities_abundance_gradient.m,  varnames = c("haul_id1",  "haul_id2"),  value.name = "bray_curtis_dissimilarity_gradient")
              dissimilarities_abundance_total.l <- reshape2::melt(dissimilarities_abundance_total.m,  varnames = c("haul_id1",  "haul_id2"),  value.name = "bray_curtis_dissimilarity_total")
              
              dissimilarities_occurrence_turnover.l <- reshape2::melt(dissimilarities_occurrence_turnover.m,  varnames = c("haul_id1",  "haul_id2"),  value.name = "jaccard_dissimilarity_turnover")
              dissimilarities_occurrence_nestedness.l <- reshape2::melt(dissimilarities_occurrence_nestedness.m,  varnames = c("haul_id1",  "haul_id2"),  value.name = "jaccard_dissimilarity_nestedness")
              dissimilarities_occurrence_total.l <- reshape2::melt(dissimilarities_occurrence_total.m,  varnames = c("haul_id1",  "haul_id2"),  value.name = "jaccard_dissimilarity_total")
              
              #and then to data table format
              dissimilarities_abundance_balanced.l <- data.table(dissimilarities_abundance_balanced.l)
              dissimilarities_abundance_gradient.l <- data.table(dissimilarities_abundance_gradient.l)
              dissimilarities_abundance_total.l <- data.table(dissimilarities_abundance_total.l)
              dissimilarities_occurrence_turnover.l <- data.table(dissimilarities_occurrence_turnover.l)
              dissimilarities_occurrence_nestedness.l <- data.table(dissimilarities_occurrence_nestedness.l)
              dissimilarities_occurrence_total.l <- data.table(dissimilarities_occurrence_total.l)
            
              #add year for these values
              dissimilarities_abundance_balanced.l[,  "year" := years[j]]
              dissimilarities_occurrence_turnover.l[,  "year" := years[j]]
              
              
              dissimilarities_abundance_gradient.l[,  "year" := years[j]]
              dissimilarities_occurrence_nestedness.l[,  "year" := years[j]]
              
              dissimilarities_abundance_total.l[,  "year" := years[j]]
              dissimilarities_occurrence_total.l[,  "year" := years[j]]
              
              #what type of metric?
              dissimilarities_abundance_balanced.l[, "abund_biomass" := "abund"]
              dissimilarities_occurrence_turnover.l[, "abund_biomass" := "abund"]
              
              dissimilarities_abundance_gradient.l[, "abund_biomass" := "abund"]
              dissimilarities_occurrence_nestedness.l[, "abund_biomass" := "abund"]
              
              dissimilarities_abundance_total.l[, "abund_biomass" := "abund"]
              dissimilarities_occurrence_total.l[, "abund_biomass" := "abund"]
            
              #merge distance with dissimilarity for this year with both metrics of dissimilarity
              dissimilarities_full <- haul_id_distances.l[dissimilarities_abundance_balanced.l,  on = c("haul_id1",  "haul_id2")]
              dissimilarities_full <- dissimilarities_full[dissimilarities_occurrence_turnover.l,  on = c("haul_id1",  "haul_id2",  "year",  "abund_biomass")]
              
              dissimilarities_full <- dissimilarities_full[dissimilarities_occurrence_total.l,  on = c("haul_id1",  "haul_id2",  "year",  "abund_biomass")]
              
              dissimilarities_full <- dissimilarities_full[dissimilarities_occurrence_nestedness.l,  on = c("haul_id1",  "haul_id2",  "year",  "abund_biomass")]
              
              dissimilarities_full <- dissimilarities_full[dissimilarities_abundance_gradient.l,  on = c("haul_id1",  "haul_id2",  "year",  "abund_biomass")]
              
              dissimilarities_full <- dissimilarities_full[dissimilarities_abundance_total.l,  on = c("haul_id1",  "haul_id2",  "year",  "abund_biomass")]
              
              #add survey survey unit
              dissimilarities_full[, "survey" := FishGlob_clean_subset[1, survey]]
              dissimilarities_full[, "survey_unit" := all_survey_units[i]]
            
              #add to data table
              distances_dissimilarities_allyears_15perc_excluded <- rbind(distances_dissimilarities_allyears_15perc_excluded,  dissimilarities_full)
              
                            #delete if haul_id2 < haul_id1 (deletes comparisons between same site (0 inflated) and any duplicate comparisons)
              dissimilarities_full <- dissimilarities_full[haul_id2<haul_id1]
              
              print(paste0(j, "/", length(years)))
          
              } #closes if else abundance versus biomass

  
         } #closes year
  
  print(paste0("We have cycled through ",  i, " of ", length(all_survey_units),  " total survey/season combos"))
  
  } #closes survey/region


#saveRDS(distances_dissimilarities_allyears_15perc_excluded,  here::here("output", "distance_decay", "distances_dissimilarities_allyears_15perc_excluded.rds"))

distances_dissimilarities_allyears_15perc_excluded <- readRDS(here::here("output", "distance_decay", "distances_dissimilarities_allyears_15perc_excluded.rds"))
```


Some NAs in multiple years for NEUS spring and fall, spring scotian shelf, Queen Charlotte Sound (2013 only).

```{r}
missing_values <- distances_dissimilarities_allyears[!complete.cases(distances_dissimilarities_allyears),] #rows with missing values

View(missing_values)

missing_values[,count_NA := .N,.(survey_unit, year)]

#issues in 
View(unique(missing_values[,.(survey_unit,year, count_NA)]))

#NEUS Fall
#This occurred for haul 109-110 because neither had any species (will manually change dissimilarity to 0)

#SCS Spring
#Same issue, neither observation had any species

#total percent missing?
nrow(missing_values)/nrow(distances_dissimilarities_allyears)*100

#We will keep balanced/gradient as NA, but will change totals to 0

distances_dissimilarities_allyears[,jaccard_dissimilarity_total := ifelse(is.na(jaccard_dissimilarity_total),0,jaccard_dissimilarity_total)][,bray_curtis_dissimilarity_total := ifelse(is.na(bray_curtis_dissimilarity_total),0,bray_curtis_dissimilarity_total)] #NB: does not change distribution at all

```


```{r yearly means}

distances_dissimilarities_allyears_15perc_excluded[, bray_curtis_dissimilarity_balanced_mean := mean(bray_curtis_dissimilarity_balanced, na.rm = T),.(survey_unit,year)][, bray_curtis_dissimilarity_gradient_mean := mean(bray_curtis_dissimilarity_gradient, na.rm = T),.(survey_unit,year)][, bray_curtis_dissimilarity_total_mean := mean(bray_curtis_dissimilarity_total, na.rm = T),.(survey_unit,year)][, jaccard_dissimilarity_turnover_mean := mean(jaccard_dissimilarity_turnover, na.rm = T),.(survey_unit,year)][, jaccard_dissimilarity_nestedness_mean := mean(jaccard_dissimilarity_nestedness, na.rm = T),.(survey_unit,year)][, jaccard_dissimilarity_total_mean := mean(jaccard_dissimilarity_total, na.rm = T),.(survey_unit,year)]

distances_dissimilarities_allyears_15perc_excluded.r <- unique(distances_dissimilarities_allyears_15perc_excluded[,.(survey, survey_unit, year,abund_biomass, bray_curtis_dissimilarity_balanced_mean, bray_curtis_dissimilarity_gradient_mean, bray_curtis_dissimilarity_total_mean, jaccard_dissimilarity_turnover_mean, jaccard_dissimilarity_nestedness_mean, jaccard_dissimilarity_total_mean)])

rm(distances_dissimilarities_allyears_15perc_excluded)


saveRDS(distances_dissimilarities_allyears_15perc_excluded.r,  here::here("output", "distance_decay", "distances_dissimilarities_allyears_15perc_excluded.r.rds"))
```
