 ---  "#C4451C", #G ---
title: "Bray Curtis Total Dissimilarity Surface Temperature & Fishing Interactions for All Regions"
output: html_notebook
---

Here we will link bottom temperature and fishing to dissimilarity

This code is Script 8b for Kitchel et al. Biotic homogenization, the exception and not the rule for marine fish communities manuscript.

- This project is a collaborative effort to describe changes in taxonomic composition  of fish communities around the world--as sampled by bottom tscaledl surveys.

- Code by Zoë J. Kitchel

SESSION INFO TO DO

##Here, we examine bottom temperature and fishing pressure to predict annual BC total dissimilarity for all survey units


```{r setup}

library(data.table)
library(ggplot2)
library(lme4)
library(lmerTest)
library(MuMIn)


#Pull in merged product from Regional_statistics_runbeforedrivers.Rmd

#pull in dissimilarities
distances_dissimilarities_allyears.r <- readRDS(here::here("output", "dissimilarities", "distances_dissimilarities_allyears.r.rds"))

#pull in temps
OISST_data_temp_avgs_full <- readRDS(here::here("data", "Temperature", "OISST_data_temp_avgs_full.rds"))

OISST_data_temp_avgs_full[,year:=year_for_avg]

#fishing
SAU_summed_tonnes <- readRDS(here::here("data", "sea_around_us", "SAU_summed_tonnes.rds"))

dissimilarities_temp_fishing_regstats <- distances_dissimilarities_allyears.r[OISST_data_temp_avgs_full, on = c("survey_unit","year")]

dissimilarities_temp_fishing_regstats <- dissimilarities_temp_fishing_regstats[SAU_summed_tonnes, on = c("survey_unit","year")]

#get rid of any years not between 1981-2019
dissimilarities_temp_fishing_regstats <- dissimilarities_temp_fishing_regstats[year >= 1981 & year <=2019]
```


Color palette for survey units
```{r}

survey_unit.list <- levels(factor(dissimilarities_temp_fishing_regstats[,survey_unit]))

palette_42 <- c(
  "#5A5156", #AI
  "#DF00DB", #BITS-1
  "#DB8EDA", #BITS-4
  "#F6222E", #CHL
  "#F8A19F", #DFO-NF
  "#16FF32", #DFO-QCS
  "#325A9B", #EBS
  "#3283FE", #EVHOE
  "#FEAF16", #FR-CGFS
  "#fccb6d", #GMEX-Fall
  "#1C8356", #GMEX-Summer
  "#C4451C", #GOA
  "#85660D", #GRL-DE
  "#B0009F", #GSL-N
  "#BF79B8", #GSL-S
  "#1CBE4F", #ICE-GFS
  "#782AB6", #IE-IGFS
  "#90AD1C", #MEDITS
  "#6B003A", #NAM
  "#A75B00", #NEUS-Fall
  "#E3B072", #NEUS-Spring
  "#02E8B6", #NIGFS-1
  "#97E7D5", #NIGFS-4
  "#B00068", #Nor-BTS-3
  "#00B9E3", #NS-IBTS-1
  "#95E2F4", #NS-IBTS-3
  "#B3CE73", #NZ-CHAT
  "#689500", #NZ-ECSI
  "#364d02",#NZ-SUBA
  "#AAF400", #NZ-WCSI
  "#AA0DFE", #PT-IBTS
  "#7f9eb8", #ROCKALL
  "#FA0087", #S-GEORG
  "#DEA0FD", #SCS-Summer
  "#FCEF88", #SEUS-fall
  "#A59405", #SEUS-spring
  "#FCE100", #SEUS-summer
  "#544563", #SWC-IBTS-1
  "#a37fc7", #SWC-IBTS-4
  "#C075A6", #WCANN
  "#BDCDFF", #ZAF-ATL
  "#003EFF"  #ZAF-IND
)

color_link <- data.table(survey_unit = survey_unit.list,hex = palette_42)
```

Add names for plotting
```{r add names for plotting}


name_helper <- data.table(Survey_Name_Season = c("Aleutian Islands",
                                    "Baltic Sea Q1",
                                    "Baltic Sea Q4",
                                    "Chile",
                                    "Newfoundland",
                                    "Queen Charlotte Sound",
                                    "Eastern Bering Sea",
                                    "Bay of Biscay",
                                    "English Channel",
                                    "Gulf of Mexico Summer",
                                    "Gulf of Alaska",
                                    "Greenland",
                                    "N Gulf of St. Lawrence",
                                    "S Gulf of St. Lawrence",
                                    "Iceland",
                                    "Irish Sea",
                                    "Mediterranean",
                                    "Namibia",
                                    "NE US Fall",
                                    "NE US Spring",
                                    "N Ireland Q1",
                                    "N Ireland Q4",
                                    "Barents Sea Norway Q3",
                                    "N Sea Q1",
                                    "N Sea Q3",
                                    "Chatham Rise NZ",
                                    "E Coast S Island NZ",
                                    "W Coast S Island NZ",
                                    "Portugal",
                                    "S Georgia",
                                  "Scotian Shelf Summer",
                                  "SE US Fall",
                                  "SE US Spring",
                                  "SE US Summer",
                                  "W Coast US",
                                  "Atlantic Ocean ZA",
                                  "Indian Ocean ZA",
                                   "Rockall Plateau",
                                  "Scotland Shelf Sea Q1",
                                  "Scotland Shelf Sea Q4",
                                  "Falkland Islands",
                                  "Gulf of Mexico Fall",
                                  "Barents Sea Norway Q1",
                                  "Sub-Arctic NZ",
                                  "Scotian Shelf Spring"),
                          survey_unit = c(
                                  "AI",        
                                  "BITS-1",    
                                  "BITS-4",    
                                  "CHL",       
                                  "DFO-NF",    
                                  "DFO-QCS",   
                                  "EBS",       
                                  "EVHOE",     
                                  "FR-CGFS",   
                                  "GMEX-Summer",
                                  "GOA",       
                                  "GRL-DE",    
                                  "GSL-N",     
                                  "GSL-S",     
                                  "ICE-GFS",   
                                  "IE-IGFS",   
                                  "MEDITS",    
                                  "NAM",       
                                  "NEUS-Fall", 
                                  "NEUS-Spring",
                                  "NIGFS-1",   
                                  "NIGFS-4",   
                                  "Nor-BTS-3", 
                                  "NS-IBTS-1", 
                                  "NS-IBTS-3", 
                                  "NZ-CHAT",   
                                  "NZ-ECSI",   
                                  "NZ-WCSI",   
                                  "PT-IBTS",   
                                  "S-GEORG",   
                                  "SCS-SUMMER",
                                  "SEUS-fall", 
                                  "SEUS-spring",
                                  "SEUS-summer",
                                  "WCANN",     
                                  "ZAF-ATL",   
                                  "ZAF-IND",
                                  "ROCKALL",
                                  "SWC-IBTS-1",
                                  "SWC-IBTS-4",
                                  "FALK",
                                  "GMEX-Fall",
                                  "Nor-BTS-1",
                                  "NZ-SUBA",
                                  "SCS-SPRING"
                          ))

color_link <- color_link[name_helper, on = "survey_unit"]

```

Table that will be populated with model results
```{r}
#depth, scaled, temp var, coefficient, p-value, marg_r^2, AICc
total_dissimilarity_sst_fishing_model_results <- data.table(Depth = as.character(),
                                                       Scaled = as.character(),
                                                       `Temp variable` = as.character(),
                                                       `Fishing`= as.character(),
                                                       `Temp fishing interaction` = as.character(),
                                                       `Fixed effect structure` = as.character(),
                                                       Intercept = as.numeric(),
                                                       `Temp coefficient` = as.numeric(),
                                                       `Coefficient p-value 1` = as.numeric(),
                                                       `Fishing coefficient` = as.numeric(),
                                                       `Coefficient p-value 2` = as.numeric(),
                                                       `Temp*fishing coefficient` = as.numeric(),
                                                       `Coefficient p-value 3` = as.numeric(),
                                                       `R-squared` = as.numeric(),
                                                       AICc = as.numeric())
                                                  
```

###How well does fishing alone predict mean annual BC total dissimilarity?


```{r build fishing mod}

#fishing and random slope and intercept for survey
allreg_dissimilarity_fishing_mean_mod <- lm(bray_curtis_dissimilarity_total_mean ~ summed_tonnes_scaled_byreg, data = dissimilarities_temp_fishing_regstats)

coefs <- data.frame(coef(summary(allreg_dissimilarity_fishing_mean_mod)))

total_dissimilarity_sst_fishing_model_results_row <- data.table(matrix(c(NA, #depth
                                                            NA, #scaled
                                                            NA, #temp var
                                                            T, #fishing
                                                            F, #temp fishing interaction
                                                            "summed_tonnes_scaled_byreg", #fixed effect
                                                            paste0(
                                                              round(coefs[1,1],3),
                                                              " ± ",
                                                              round(coefs[1,2],3)), #intercept
                                                            NA,#temp coef
                                                            NA,#coef p value
                                                            paste0(
                                                              round(coefs[2,1],3),
                                                              " ± ",
                                                              round(coefs[2,2],3)),#fishing coef
                                                            round(coefs[2,4],3),#coef p value
                                                            NA, #interact coef
                                                            NA, #coef p value
                                                round(summary(allreg_dissimilarity_fishing_mean_mod)$r.squared,3), #r squared
                                                            round(AICc(allreg_dissimilarity_fishing_mean_mod),3)),#AICc
                                                           nrow =1))

total_dissimilarity_sst_fishing_model_results <- rbind(total_dissimilarity_sst_fishing_model_results, total_dissimilarity_sst_fishing_model_results_row, use.names = F)

#fixed effect for survey and interaction with fishing
allreg_dissimilarity_fishing_fixed_mean_mod <- lm(bray_curtis_dissimilarity_total_mean ~ summed_tonnes_scaled_byreg*survey_unit, data = dissimilarities_temp_fishing_regstats)

coefs <- data.frame(coef(summary(allreg_dissimilarity_fishing_fixed_mean_mod)))

total_dissimilarity_sst_fishing_model_results_row <- data.table(matrix(c(NA, #depth
                                                            NA, #scaled
                                                            NA, #temp var
                                                            T, #fishing
                                                            F, #temp fishing interaction
                                                            "summed_tonnes_scaled_byreg*survey_unit", #fixed effect
                                                    
                                                            paste0(
                                                              round(coefs[1,1],3),
                                                              " ± ",
                                                              round(coefs[1,2],3)), #intercept
                                                            NA,#temp coef
                                                            NA,#coef p value
                                                            paste0(
                                                              round(coefs[2,1],3),
                                                              " ± ",
                                                              round(coefs[2,2],3)),#fishing coef
                                                            round(coefs[2,4],3),#coef p value
                                                            NA, #interact coef
                                                            NA, #coef p value
                                                round(summary(allreg_dissimilarity_fishing_fixed_mean_mod)$r.squared,3), #r squared
                                                         round(AICc(allreg_dissimilarity_fishing_fixed_mean_mod),3)),#AICc
                                                            ,nrow =1))

total_dissimilarity_sst_fishing_model_results <- rbind(total_dissimilarity_sst_fishing_model_results, total_dissimilarity_sst_fishing_model_results_row, use.names = F)

#fixed effect for survey and fishing, no interaction
allreg_dissimilarity_fishing_fixed_mean_mod_nointer <- lm(bray_curtis_dissimilarity_total_mean ~ summed_tonnes_scaled_byreg+survey_unit, data = dissimilarities_temp_fishing_regstats)

coefs <- data.frame(coef(summary(allreg_dissimilarity_fishing_fixed_mean_mod_nointer)))

total_dissimilarity_sst_fishing_model_results_row <- data.table(matrix(c(NA, #depth
                                                            NA, #scaled
                                                            NA, #temp var
                                                            T, #fishing
                                                            F, #temp fishing interaction
                                                            "summed_tonnes_scaled_byreg+survey_unit", #fixed effect
                                                         
                                                            paste0(round(coefs[1,1],3),
                                                              " ± ",
                                                              round(coefs[1,2],3)), #intercept
                                                            NA,#temp coef
                                                            NA,#coef p value
                                                            paste0(round(coefs[2,1],3),
                                                              " ± ",
                                                              round(coefs[2,2],3)),#fishing coef
                                                            round(coefs[2,4],3),#coef p value
                                                            NA, #interact coef
                                                            NA, #coef p value
                                                round(summary(allreg_dissimilarity_fishing_fixed_mean_mod_nointer)$r.squared,3), #r squared
                                                round(AICc(allreg_dissimilarity_fishing_fixed_mean_mod_nointer),3)#AICc
                                                      
                                                            ),nrow =1))

total_dissimilarity_sst_fishing_model_results <- rbind(total_dissimilarity_sst_fishing_model_results, total_dissimilarity_sst_fishing_model_results_row, use.names = F)

#fixed effect for survey only
allreg_dissimilarity_survey_fixed_mod <- lm(bray_curtis_dissimilarity_total_mean ~ survey_unit, data = dissimilarities_temp_fishing_regstats)

coefs <- data.frame(coef(summary(allreg_dissimilarity_survey_fixed_mod)))

total_dissimilarity_sst_fishing_model_results_row <- data.table(matrix(c(NA, #depth
                                                            NA, #scaled
                                                            NA, #temp var
                                                            F, #fishing
                                                            F, #temp fishing interaction
                                                            "survey_unit", #fixed effect
                                                            paste0(round(coefs[1,1],3),
                                                              " ± ",
                                                              round(coefs[1,2],3)), #intercept
                                                            NA,#temp coef
                                                            NA,#coef p value
                                                            NA,#fishing coef
                                                            NA,#coef p value
                                                            NA, #interact coef
                                                            NA, #coef p value
                                                round(summary(allreg_dissimilarity_survey_fixed_mod)$r.squared,3), #r squared
                                                            round(AICc(allreg_dissimilarity_survey_fixed_mod),3)#AICc
                                                            ),nrow =1))

total_dissimilarity_sst_fishing_model_results <- rbind(total_dissimilarity_sst_fishing_model_results, total_dissimilarity_sst_fishing_model_results_row, use.names = F)


View(total_dissimilarity_sst_fishing_model_results)

#save fishing only output
total_dissimilarity_fishing_model_results <- total_dissimilarity_sst_fishing_model_results

saveRDS(total_dissimilarity_fishing_model_results, here::here("output","total_dissimilarity_fishing_model_results.Rds"))

```
Best model includes only survey unit. Second best model includes an interaction between fishing and survey unit.

As relative fishing pressure increases, dissimilarity increases (more fished communities = more heterogenous communities). However, this varies among regions.

####Plot Fishing Only Model
```{r plot fishing vs dissimilarity}
dissimilarities_temp_fishing_regstats <- dissimilarities_temp_fishing_regstats[name_helper, on = "survey_unit"]

dissimilarities_temp_fishing_regstats <- dissimilarities_temp_fishing_regstats[!(Survey_Name_Season %in% c("Barents Sea Norway Q1","Falkland Islands","Rockall Plateau","Scotian Shelf Spring"))]

#for all regions
(allreg_BC_total_dissimilarity_fishing_mean <- ggplot(data = dissimilarities_temp_fishing_regstats) +
  labs(x = "Relative fishing pressure",  y = "BC total dissimilarity") +
  geom_point(aes(x = summed_tonnes_scaled_byreg, y = bray_curtis_dissimilarity_total_mean), alpha = 0.3) +
  geom_smooth(aes(x = summed_tonnes_scaled_byreg, y = bray_curtis_dissimilarity_total_mean), method = "lm") +
  theme_classic())

#save
ggsave(allreg_BC_total_dissimilarity_fishing_mean, path = here::here("figures","predictor_variables"), filename = "allreg_BC_total_dissimilarity_fishing_mean.jpg")

#for all regions faceted
(allreg_BC_total_dissimilarity_fishing_mean_facet <- ggplot(data = dissimilarities_temp_fishing_regstats) +
  labs(x = "Relative fishing pressure",  y = "BC total dissimilarity") +
  geom_point(aes(x = summed_tonnes_scaled_byreg, y = bray_curtis_dissimilarity_total_mean), alpha = 0.3) +
  geom_smooth(aes(x = summed_tonnes_scaled_byreg, y = bray_curtis_dissimilarity_total_mean), method = "lm") +
  facet_wrap(~Survey_Name_Season, scales = "free", ncol = 5) +
  theme_classic() +
    theme(legend.position = "null"))

#save
ggsave(allreg_BC_total_dissimilarity_fishing_mean_facet, path = here::here("figures","predictor_variables"), filename = "allreg_BC_total_dissimilarity_fishing_mean_facet.jpg", height = 13, width = 10, unit = "in")

```


###Temperature x fishing models
Best temp variables to carry forward: 

For raw temp (indicative of survey vs. survey):
-Mean temperature * survey



Mean temperature
```{r mean temp and fishing models of dissimilarity}


#fishing and temperature interaction
allreg_dissimilarity_fishing_mean_mod_sst_mean_avg <- lm(bray_curtis_dissimilarity_total_mean ~ summed_tonnes_scaled_byreg*yearly_mean_bypoint_avg, data = dissimilarities_temp_fishing_regstats)


        coefs <- data.frame(coef(summary(allreg_dissimilarity_fishing_mean_mod_sst_mean_avg)))

total_dissimilarity_sst_fishing_model_results_row <- data.table(matrix(c("sst", #depth
                                                            F, #scaled
                                                            "yearly_mean_bypoint_avg", #temp var
                                                            T, #fishing
                                                            T, #temp fishing interaction
                                                            "summed_tonnes_scaled_byreg*yearly_mean_bypoint_avg", #fixed effect
                                                            paste0(
                                                              round(coefs[1,1],3),
                                                              " ± ",
                                                              round(coefs[1,2],3)), #intercept
                                                            paste0(
                                                              round(coefs[3,1],3),
                                                              " ± ",
                                                              round(coefs[3,2],3)),#temp coef
                                                            round(coefs[3,4],3),#coef p value
                                                            paste0(
                                                              round(coefs[2,1],3),
                                                              " ± ",
                                                              round(coefs[2,2],3)),#fishing coef
                                                            round(coefs[2,4],3),#coef p value
                                                            paste0(
                                                              round(coefs[4,1],3),
                                                              " ± ",
                                                              round(coefs[4,2],3)), #interact coef
                                                            round(coefs[4,4],3), #coef p value
                                                round(summary(allreg_dissimilarity_fishing_mean_mod_sst_mean_avg)$r.squared,3), #r squared
                                                      round(AICc(allreg_dissimilarity_fishing_mean_mod_sst_mean_avg),3)#AICc
                                                            ),nrow =1))

total_dissimilarity_sst_fishing_model_results <- rbind(total_dissimilarity_sst_fishing_model_results, total_dissimilarity_sst_fishing_model_results_row, use.names = F)


#raw temp values AND interaction with both fishing and fixed survey unit

allreg_dissimilarity_fishing_mean_mod_sst_mean_avg_fixed_two_interact <- lm(bray_curtis_dissimilarity_total_mean ~ summed_tonnes_scaled_byreg*yearly_mean_bypoint_avg*survey_unit, data = dissimilarities_temp_fishing_regstats)


        coefs <- data.frame(coef(summary(allreg_dissimilarity_fishing_mean_mod_sst_mean_avg_fixed_two_interact)))

total_dissimilarity_sst_fishing_model_results_row <- data.table(matrix(c("sst", #depth
                                                            F, #scaled
                                                            "yearly_mean_bypoint_avg", #temp var
                                                            T, #fishing
                                                            T, #temp fishing interaction
                                                           "summed_tonnes_scaled_byreg*yearly_mean_bypoint_avg*survey_unit", #fixed effect
                                                            paste0(
                                                              round(coefs[1,1],3),
                                                              " ± ",
                                                              round(coefs[1,2],2)), #intercept
                                                            paste0(
                                                              round(coefs[3,1],3),
                                                              " ± ",
                                                              round(coefs[3,2],2)),#temp coef
                                                            round(coefs[3,4],3),#coef p value
                                                            paste0(
                                                              round(coefs[2,1],3),
                                                              " ± ",
                                                              round(coefs[2,2],2)),#fishing coef
                                                            round(coefs[2,4],3),#coef p value
                                                            paste0(
                                                              round(coefs[40,1],3),
                                                              " ± ",
                                                              round(coefs[40,2],2)), #interact coef (fishing and temp)
                                                            round(coefs[40,4],3), #coef p value (fishing and temp)
                                                round(r.squaredGLMM(allreg_dissimilarity_fishing_mean_mod_sst_mean_avg_fixed_two_interact)[[1]],3), #r squared
                                                            round(AICc(allreg_dissimilarity_fishing_mean_mod_sst_mean_avg_fixed_two_interact),3)#AICc
                                                            ),nrow =1))

total_dissimilarity_sst_fishing_model_results <- rbind(total_dissimilarity_sst_fishing_model_results, total_dissimilarity_sst_fishing_model_results_row, use.names = F)

#scaled temp values AND interaction with fishing but no interaction with fixed survey unit

allreg_dissimilarity_fishing_mean_mod_sst_mean_avg_fixed_temp_fish_interact <- lm(bray_curtis_dissimilarity_total_mean ~ summed_tonnes_scaled_byreg*yearly_mean_bypoint_avg+survey_unit, data = dissimilarities_temp_fishing_regstats)


        coefs <- data.frame(coef(summary(allreg_dissimilarity_fishing_mean_mod_sst_mean_avg_fixed_temp_fish_interact)))

total_dissimilarity_sst_fishing_model_results_row <- data.table(matrix(c("sst", #depth
                                                            F, #scaled
                                                            "yearly_mean_bypoint_avg", #temp var
                                                            T, #fishing
                                                            T, #temp fishing interaction
                                                            "summed_tonnes_scaled_byreg*yearly_mean_bypoint_avg+survey_unit", #fixed effect
                                                            paste0(
                                                              round(coefs[1,1],3),
                                                              " ± ",
                                                              round(coefs[1,2],3)), #intercept
                                                            paste0(
                                                              round(coefs[3,1],3),
                                                              " ± ",
                                                              round(coefs[3,2],3)),#temp coef
                                                            round(coefs[3,4],3),#coef p value
                                                            paste0(
                                                              round(coefs[2,1],3),
                                                              " ± ",
                                                              round(coefs[2,2],3)),#fishing coef
                                                            round(coefs[2,4],3),#coef p value
                                                            paste0(
                                                              round(coefs[38,1],3),
                                                              " ± ",
                                                              round(coefs[38,2],3)), #interact coef (fishing and temp)
                                                            round(coefs[38,4],3), #coef p value (fishing and temp)
                                                round(r.squaredGLMM(allreg_dissimilarity_fishing_mean_mod_sst_mean_avg_fixed_temp_fish_interact)[[1]],3), #r squared
                                                            round(AICc(allreg_dissimilarity_fishing_mean_mod_sst_mean_avg_fixed_temp_fish_interact),3)#AICc
                                                            ),nrow =1))

total_dissimilarity_sst_fishing_model_results <- rbind(total_dissimilarity_sst_fishing_model_results, total_dissimilarity_sst_fishing_model_results_row, use.names = F)

#scaled temp values AND interact with survey unit but no interact with fishing

allreg_dissimilarity_fishing_mean_mod_sst_mean_avg_fixed_temp_survey_interact <- lm(bray_curtis_dissimilarity_total_mean ~ summed_tonnes_scaled_byreg+yearly_mean_bypoint_avg*survey_unit, data = dissimilarities_temp_fishing_regstats)


        coefs <- data.frame(coef(summary(allreg_dissimilarity_fishing_mean_mod_sst_mean_avg_fixed_temp_survey_interact)))

total_dissimilarity_sst_fishing_model_results_row <- data.table(matrix(c("sst", #depth
                                                            F, #scaled
                                                            "yearly_mean_bypoint_avg", #temp var
                                                            T, #fishing
                                                            F, #temp fishing interaction
                                                          "summed_tonnes_scaled_byreg+yearly_mean_bypoint_avg*survey_unit", #fixed effect
                                                            paste0(
                                                              round(coefs[1,1],3),
                                                              " ± ",
                                                              round(coefs[1,2],3)), #intercept
                                                            paste0(
                                                              round(coefs[3,1],3),
                                                              " ± ",
                                                              round(coefs[3,2],3)),#temp coef
                                                            round(coefs[3,4],3),#coef p value
                                                            paste0(
                                                              round(coefs[2,1],3),
                                                              " ± ",
                                                              round(coefs[2,2],3)),#fishing coef
                                                            round(coefs[2,4],3),#coef p value
                                                            NA, #interact coef (fishing and temp)
                                                            NA, #coef p value (fishing and temp)
                                                round(summary(allreg_dissimilarity_fishing_mean_mod_sst_mean_avg_fixed_temp_survey_interact)$r.squared,3), #r squared
                                                
                                                            round(AICc(allreg_dissimilarity_fishing_mean_mod_sst_mean_avg_fixed_temp_survey_interact),3)#AICc
                                                            ),nrow =1))

total_dissimilarity_sst_fishing_model_results <- rbind(total_dissimilarity_sst_fishing_model_results, total_dissimilarity_sst_fishing_model_results_row, use.names = F)

#temp values AND interact with survey unit but no interact with fishing

allreg_dissimilarity_fishing_mean_mod_sst_mean_avg_fixed_fishing_survey_interact <- lm(bray_curtis_dissimilarity_total_mean ~yearly_mean_bypoint_avg + summed_tonnes_scaled_byreg*survey_unit, data = dissimilarities_temp_fishing_regstats)


        coefs <- data.frame(coef(summary(allreg_dissimilarity_fishing_mean_mod_sst_mean_avg_fixed_fishing_survey_interact)))

total_dissimilarity_sst_fishing_model_results_row <- data.table(matrix(c("sst", #depth
                                                            F, #scaled
                                                            "yearly_mean_bypoint_avg", #temp var
                                                            T, #fishing
                                                            F, #temp fishing interaction
                                                            "yearly_mean_bypoint_avg + summed_tonnes_scaled_byreg*survey_unit", #fixed effect
                                                            paste0(
                                                              round(coefs[1,1],3),
                                                              " ± ",
                                                              round(coefs[1,2],3)), #intercept
                                                            paste0(
                                                              round(coefs[3,1],3),
                                                              " ± ",
                                                              round(coefs[3,2],3)),#temp coef
                                                            round(coefs[3,4],3),#coef p value
                                                            paste0(
                                                              round(coefs[2,1],3),
                                                              " ± ",
                                                              round(coefs[2,2],3)),#fishing coef
                                                            round(coefs[2,4],3),#coef p value
                                                            NA, #interact coef (fishing and temp)
                                                            NA, #coef p value (fishing and temp)
                                                round(summary(allreg_dissimilarity_fishing_mean_mod_sst_mean_avg_fixed_fishing_survey_interact)$r.squared,3), #r squared
                                              
                                                            round(AICc(allreg_dissimilarity_fishing_mean_mod_sst_mean_avg_fixed_fishing_survey_interact),3)#AICc
                                                            ),nrow =1))

total_dissimilarity_sst_fishing_model_results <- rbind(total_dissimilarity_sst_fishing_model_results, total_dissimilarity_sst_fishing_model_results_row, use.names = F)

#scaled temp values, fixed effects for fishing and survey unit, no interact

allreg_dissimilarity_fishing_mean_mod_sst_mean_avg_fixed_fishing_no_interact <- lm(bray_curtis_dissimilarity_total_mean ~yearly_mean_bypoint_avg + summed_tonnes_scaled_byreg + survey_unit, data = dissimilarities_temp_fishing_regstats)


        coefs <- data.frame(coef(summary(allreg_dissimilarity_fishing_mean_mod_sst_mean_avg_fixed_fishing_no_interact)))

total_dissimilarity_sst_fishing_model_results_row <- data.table(matrix(c("sst", #depth
                                                            F, #scaled
                                                            "yearly_mean_bypoint_avg", #temp var
                                                            T, #fishing
                                                            F, #temp fishing interaction
                                                            "yearly_mean_bypoint_avg + summed_tonnes_scaled_byreg + survey_unit", #fixed effect
                                                            paste0(
                                                              round(coefs[1,1],3),
                                                              " ± ",
                                                              round(coefs[1,2],3)), #intercept
                                                            paste0(
                                                              round(coefs[3,1],3),
                                                              " ± ",
                                                              round(coefs[3,2],3)),#temp coef
                                                            round(coefs[3,4],3),#coef p value
                                                            paste0(
                                                              round(coefs[2,1],3),
                                                              " ± ",
                                                              round(coefs[2,2],3)),#fishing coef
                                                            round(coefs[2,4],3),#coef p value
                                                            NA, #interact coef (fishing and temp)
                                                            NA, #coef p value (fishing and temp)
                                                round(summary(allreg_dissimilarity_fishing_mean_mod_sst_mean_avg_fixed_fishing_no_interact)$r.squared,3), #r squared
                                                            round(AICc(allreg_dissimilarity_fishing_mean_mod_sst_mean_avg_fixed_fishing_no_interact),3)#AICc
                                                            ),nrow =1))

total_dissimilarity_sst_fishing_model_results <- rbind(total_dissimilarity_sst_fishing_model_results, total_dissimilarity_sst_fishing_model_results_row, use.names = F)
 


#temp only with interaction with fixed effect with survey unit

allreg_dissimilarity_sst_mean_avg_fixed <- lm(bray_curtis_dissimilarity_total_mean ~ yearly_mean_bypoint_avg*survey_unit, data = dissimilarities_temp_fishing_regstats)


        coefs <- data.frame(coef(summary(allreg_dissimilarity_sst_mean_avg_fixed)))

total_dissimilarity_sst_fishing_model_results_row <- data.table(matrix(c("sst", #depth
                                                            F, #scaled
                                                            "yearly_mean_bypoint_avg", #temp var
                                                            F, #fishing
                                                            F, #temp fishing interaction
                                                            "yearly_mean_bypoint_avg*survey_unit", #fixed effect
                                                            paste0(
                                                              round(coefs[1,1],3),
                                                              " ± ",
                                                              round(coefs[1,2],3)), #intercept
                                                            paste0(
                                                              round(coefs[2,1],3),
                                                              " ± ",
                                                              round(coefs[2,2],3)),#temp coef
                                                            round(coefs[2,4],3),#coef p value
                                                            NA,#fishing coef
                                                            NA,#coef p value
                                                            NA, #interact coef (fishing and temp)
                                                            NA, #coef p value (fishing and temp)
                                                round(summary(allreg_dissimilarity_sst_mean_avg_fixed)$r.squared,3),#r squared                                                            
                                                round(AICc(allreg_dissimilarity_sst_mean_avg_fixed),3)#AICc
                                                            ),nrow =1))

total_dissimilarity_sst_fishing_model_results <- rbind(total_dissimilarity_sst_fishing_model_results, total_dissimilarity_sst_fishing_model_results_row, use.names = F)

#temp only with fixed effect with survey unit, no interact

allreg_dissimilarity_sst_mean_avg_fixed_nointeract <- lm(bray_curtis_dissimilarity_total_mean ~ yearly_mean_bypoint_avg+survey_unit, data = dissimilarities_temp_fishing_regstats)


        coefs <- data.frame(coef(summary(allreg_dissimilarity_sst_mean_avg_fixed_nointeract)))

total_dissimilarity_sst_fishing_model_results_row <- data.table(matrix(c("sst", #depth
                                                            F, #scaled
                                                            "yearly_mean_bypoint_avg", #temp var
                                                            F, #fishing
                                                            F, #temp fishing interaction
                                                            "yearly_mean_bypoint_avg+survey_unit", #fixed effect
                                                            paste0(
                                                              round(coefs[1,1],3),
                                                              " ± ",
                                                              round(coefs[1,2],3)), #intercept
                                                            paste0(
                                                              round(coefs[2,1],3),
                                                              " ± ",
                                                              round(coefs[2,2],3)),#temp coef
                                                            round(coefs[2,4],3),#coef p value
                                                            NA,#fishing coef
                                                            NA,#coef p value
                                                            NA, #interact coef (fishing and temp)
                                                            NA, #coef p value (fishing and temp)
                                                round(r.squaredGLMM(allreg_dissimilarity_sst_mean_avg_fixed_nointeract)[[1]],3), #r squared
                                                            round(AICc(allreg_dissimilarity_sst_mean_avg_fixed_nointeract),3)),#AICc
                                                            nrow =1))

total_dissimilarity_sst_fishing_model_results <- rbind(total_dissimilarity_sst_fishing_model_results, total_dissimilarity_sst_fishing_model_results_row, use.names = F)
    

```



Show model ranking
```{r}
View(total_dissimilarity_sst_fishing_model_results)
total_dissimilarity_sst_fishing_model_results[,deltaAICc := as.numeric(AICc)-min(as.numeric(AICc))]

saveRDS(total_dissimilarity_sst_fishing_model_results, here::here("output", "total_dissimilarity_sst_fishing_model_results.Rds"))

total_dissimilarity_sst_fishing_model_results <- readRDS(here::here("output", "total_dissimilarity_sst_fishing_model_results.Rds"))

#export as CSV
fwrite(total_dissimilarity_sst_fishing_model_results, here::here("output", "total_dissimilarity_sst_fishing_model_results.csv"))
```


Best overall model (by AICc)
  - Fishing as a fixed effect without temperature or survey unit
  - Raw mean temperature with interaction with fishing 

Carry these forward
---


OA
  "#85660D", #GRL-DE
  "#B0009F", #GSL-N
  "#BF79B8", #GSL-S
  "#1CBE4F", #ICE-GFS
  "#782AB6", #IE-IGFS
  "#90AD1C", #MEDITS
  "#6B003A", #NAM
  "#A75B00", #NEUS-Fall
  "#E3B072", #NEUS-Spring
  "#02E8B6", #NIGFS-1
  "#97E7D5", #NIGFS-4
  "#B00068", #Nor-BTS-3
  "#00B9E3", #NS-IBTS-1
  "#95E2F4", #NS-IBTS-3
  "#B3CE73", #NZ-CHAT
  "#689500", #NZ-ECSI
  "#364d02",#NZ-SUBA
  "#AAF400", #NZ-WCSI
  "#AA0DFE", #PT-IBTS
  "#7f9eb8", #ROCKALL
  "#FA0087", #S-GEORG
  "#DEA0FD", #SCS-Summer
  "#FCEF88", #SEUS-fall
  "#A59405", #SEUS-spring
  "#FCE100", #SEUS-summer
  "#544563", #SWC-IBTS-1
  "#a37fc7", #SWC-IBTS-4 "#BDCDFF", #ZAF-ATL
  "#003EFF"  #ZAF-IND
)

color_link <- data.table(survey_unit = survey_unit.list,hex = palette_42)
```

Add names for plotting
```{r add names for plotting}


name_helper <- data.table(Survey_Name_Season = c("Aleutian Islands",
                                    "Baltic Sea Q1",
                                    "Baltic Sea Q4",
                                    "Chile",
                                    "Newfoundland",
                                    "Queen Charlotte Sound",
                                    "Eastern Bering Sea",
                                    "Bay of Biscay",
                                    "English Channel",
                                    "Gulf of Mexico Summer",
                                    "Gulf of Alaska",
                                    "Greenland",
                                    "N Gulf of St. Lawrence",
                                    "S Gulf of St. Lawrence",
                                    "Iceland",
                                    "Irish Sea",
                                    "Mediterranean",
                                    "Namibia",
                                    "NE US Fall",
                                    "NE US Spring",
                                    "N Ireland Q1",
                                    "N Ireland Q4",
                                    "Barents Sea Norway Q3",
                                    "N Sea Q1",
                                    "N Sea Q3",
                                    "Chatham Rise NZ",
                                    "E Coast S Island NZ",
                                    "W Coast S Island NZ",
                                    "Portugal",
                                    "S Georgia",
                                  "Scotian Shelf Summer",
                                  "SE US Fall",
                                  "SE US Spring",
                                  "SE US Summer",
                                  "W Coast US",
                                  "Atlantic Ocean ZA",
                                  "Indian Ocean ZA",
                                   "Rockall Plateau",
                                  "Scotland Shelf Sea Q1",
                                  "Scotland Shelf Sea Q4",
                                  "Falkland Islands",
                                  "Gulf of Mexico Fall",
                                  "Barents Sea Norway Q1",
                                  "Sub-Arctic NZ",
                                  "Scotian Shelf Spring"),
                          survey_unit = c(
                                  "AI",        
                                  "BITS-1",    
                                  "BITS-4",    
                                  "CHL",       
                                  "DFO-NF",    
                                  "DFO-QCS",   
                                  "EBS",       
                                  "EVHOE",     
                                  "FR-CGFS",   
                                  "GMEX-Summer",
                                  "GOA",       
                                  "GRL-DE",    
                                  "GSL-N",     
                                  "GSL-S",     
                                  "ICE-GFS",   
                                  "IE-IGFS",   
                                  "MEDITS",    
                                  "NAM",       
                                  "NEUS-Fall", 
                                  "NEUS-Spring",
                                  "NIGFS-1",   
                                  "NIGFS-4",   
                                  "Nor-BTS-3", 
                                  "NS-IBTS-1", 
                                  "NS-IBTS-3", 
                                  "NZ-CHAT",   
                                  "NZ-ECSI",   
                                  "NZ-WCSI",   
                                  "PT-IBTS",   
                                  "S-GEORG",   
                                  "SCS-SUMMER",
                                  "SEUS-fall", 
                                  "SEUS-spring",
                                  "SEUS-summer",
                                  "WCANN",     
                                  "ZAF-ATL",   
                                  "ZAF-IND",
                                  "ROCKALL",
                                  "SWC-IBTS-1",
                                  "SWC-IBTS-4",
                                  "FALK",
                                  "GMEX-Fall",
                                  "Nor-BTS-1",
                                  "NZ-SUBA",
                                  "SCS-SPRING"
                          ))

color_link <- color_link[name_helper, on = "survey_unit"]

```

Table that will be populated with model results
```{r}
#depth, scaled, temp var, coefficient, p-value, marg_r^2, AICc
total_dissimilarity_sst_fishing_model_results <- data.table(Depth = as.character(),
                                                       Scaled = as.character(),
                                                       `Temp variable` = as.character(),
                                                       `Fishing`= as.character(),
                                                       `Temp fishing interaction` = as.character(),
                                                       `Fixed effect structure` = as.character(),
                                                       Intercept = as.numeric(),
                                                       `Temp coefficient` = as.numeric(),
                                                       `Coefficient p-value 1` = as.numeric(),
                                                       `Fishing coefficient` = as.numeric(),
                                                       `Coefficient p-value 2` = as.numeric(),
                                                       `Temp*fishing coefficient` = as.numeric(),
                                                       `Coefficient p-value 3` = as.numeric(),
                                                       `R-squared` = as.numeric(),
                                                       AICc = as.numeric())
                                                  
```

###How well does fishing alone predict mean annual BC total dissimilarity?


```{r build fishing mod}

#fishing and random slope and intercept for survey
allreg_dissimilarity_fishing_mean_mod <- lm(bray_curtis_dissimilarity_total_mean ~ summed_tonnes_scaled_byreg, data = dissimilarities_temp_fishing_regstats)

coefs <- data.frame(coef(summary(allreg_dissimilarity_fishing_mean_mod)))

total_dissimilarity_sst_fishing_model_results_row <- data.table(matrix(c(NA, #depth
                                                            NA, #scaled
                                                            NA, #temp var
                                                            T, #fishing
                                                            F, #temp fishing interaction
                                                            "summed_tonnes_scaled_byreg", #fixed effect
                                                            paste0(
                                                              round(coefs[1,1],3),
                                                              " ± ",
                                                              round(coefs[1,2],3)), #intercept
                                                            NA,#temp coef
                                                            NA,#coef p value
                                                            paste0(
                                                              round(coefs[2,1],3),
                                                              " ± ",
                                                              round(coefs[2,2],3)),#fishing coef
                                                            round(coefs[2,4],3),#coef p value
                                                            NA, #interact coef
                                                            NA, #coef p value
                                                round(summary(allreg_dissimilarity_fishing_mean_mod)$r.squared,3), #r squared
                                                            round(AICc(allreg_dissimilarity_fishing_mean_mod),3)),#AICc
                                                           nrow =1))

total_dissimilarity_sst_fishing_model_results <- rbind(total_dissimilarity_sst_fishing_model_results, total_dissimilarity_sst_fishing_model_results_row, use.names = F)

#fixed effect for survey and interaction with fishing
allreg_dissimilarity_fishing_fixed_mean_mod <- lm(bray_curtis_dissimilarity_total_mean ~ summed_tonnes_scaled_byreg*survey_unit, data = dissimilarities_temp_fishing_regstats)

coefs <- data.frame(coef(summary(allreg_dissimilarity_fishing_fixed_mean_mod)))

total_dissimilarity_sst_fishing_model_results_row <- data.table(matrix(c(NA, #depth
                                                            NA, #scaled
                                                            NA, #temp var
                                                            T, #fishing
                                                            F, #temp fishing interaction
                                                            "summed_tonnes_scaled_byreg*survey_unit", #fixed effect
                                                    
                                                            paste0(
                                                              round(coefs[1,1],3),
                                                              " ± ",
                                                              round(coefs[1,2],3)), #intercept
                                                            NA,#temp coef
                                                            NA,#coef p value
                                                            paste0(
                                                              round(coefs[2,1],3),
                                                              " ± ",
                                                              round(coefs[2,2],3)),#fishing coef
                                                            round(coefs[2,4],3),#coef p value
                                                            NA, #interact coef
                                                            NA, #coef p value
                                                round(summary(allreg_dissimilarity_fishing_fixed_mean_mod)$r.squared,3), #r squared
                                                         round(AICc(allreg_dissimilarity_fishing_fixed_mean_mod),3)),#AICc
                                                            ,nrow =1))

total_dissimilarity_sst_fishing_model_results <- rbind(total_dissimilarity_sst_fishing_model_results, total_dissimilarity_sst_fishing_model_results_row, use.names = F)

#fixed effect for survey and fishing, no interaction
allreg_dissimilarity_fishing_fixed_mean_mod_nointer <- lm(bray_curtis_dissimilarity_total_mean ~ summed_tonnes_scaled_byreg+survey_unit, data = dissimilarities_temp_fishing_regstats)

coefs <- data.frame(coef(summary(allreg_dissimilarity_fishing_fixed_mean_mod_nointer)))

total_dissimilarity_sst_fishing_model_results_row <- data.table(matrix(c(NA, #depth
                                                            NA, #scaled
                                                            NA, #temp var
                                                            T, #fishing
                                                            F, #temp fishing interaction
                                                            "summed_tonnes_scaled_byreg+survey_unit", #fixed effect
                                                         
                                                            paste0(round(coefs[1,1],3),
                                                              " ± ",
                                                              round(coefs[1,2],3)), #intercept
                                                            NA,#temp coef
                                                            NA,#coef p value
                                                            paste0(round(coefs[2,1],3),
                                                              " ± ",
                                                              round(coefs[2,2],3)),#fishing coef
                                                            round(coefs[2,4],3),#coef p value
                                                            NA, #interact coef
                                                            NA, #coef p value
                                                round(summary(allreg_dissimilarity_fishing_fixed_mean_mod_nointer)$r.squared,3), #r squared
                                                round(AICc(allreg_dissimilarity_fishing_fixed_mean_mod_nointer),3)#AICc
                                                      
                                                            ),nrow =1))

total_dissimilarity_sst_fishing_model_results <- rbind(total_dissimilarity_sst_fishing_model_results, total_dissimilarity_sst_fishing_model_results_row, use.names = F)

#fixed effect for survey only
allreg_dissimilarity_survey_fixed_mod <- lm(bray_curtis_dissimilarity_total_mean ~ survey_unit, data = dissimilarities_temp_fishing_regstats)

coefs <- data.frame(coef(summary(allreg_dissimilarity_survey_fixed_mod)))

total_dissimilarity_sst_fishing_model_results_row <- data.table(matrix(c(NA, #depth
                                                            NA, #scaled
                                                            NA, #temp var
                                                            F, #fishing
                                                            F, #temp fishing interaction
                                                            "survey_unit", #fixed effect
                                                            paste0(round(coefs[1,1],3),
                                                              " ± ",
                                                              round(coefs[1,2],3)), #intercept
                                                            NA,#temp coef
                                                            NA,#coef p value
                                                            NA,#fishing coef
                                                            NA,#coef p value
                                                            NA, #interact coef
                                                            NA, #coef p value
                                                round(summary(allreg_dissimilarity_survey_fixed_mod)$r.squared,3), #r squared
                                                            round(AICc(allreg_dissimilarity_survey_fixed_mod),3)#AICc
                                                            ),nrow =1))

total_dissimilarity_sst_fishing_model_results <- rbind(total_dissimilarity_sst_fishing_model_results, total_dissimilarity_sst_fishing_model_results_row, use.names = F)


View(total_dissimilarity_sst_fishing_model_results)

#save fishing only output
total_dissimilarity_fishing_model_results <- total_dissimilarity_sst_fishing_model_results

saveRDS(total_dissimilarity_fishing_model_results, here::here("output","total_dissimilarity_fishing_model_results.Rds"))

```
Best model includes only survey unit. Second best model includes an interaction between fishing and survey unit.

As relative fishing pressure increases, dissimilarity increases (more fished communities = more heterogenous communities). However, this varies among regions.

####Plot Fishing Only Model
```{r plot fishing vs dissimilarity}
dissimilarities_temp_fishing_regstats <- dissimilarities_temp_fishing_regstats[name_helper, on = "survey_unit"]

dissimilarities_temp_fishing_regstats <- dissimilarities_temp_fishing_regstats[!(Survey_Name_Season %in% c("Barents Sea Norway Q1","Falkland Islands","Rockall Plateau","Scotian Shelf Spring"))]

#for all regions
(allreg_BC_total_dissimilarity_fishing_mean <- ggplot(data = dissimilarities_temp_fishing_regstats) +
  labs(x = "Relative fishing pressure",  y = "BC total dissimilarity") +
  geom_point(aes(x = summed_tonnes_scaled_byreg, y = bray_curtis_dissimilarity_total_mean), alpha = 0.3) +
  geom_smooth(aes(x = summed_tonnes_scaled_byreg, y = bray_curtis_dissimilarity_total_mean), method = "lm") +
  theme_classic())

#save
ggsave(allreg_BC_total_dissimilarity_fishing_mean, path = here::here("figures","predictor_variables"), filename = "allreg_BC_total_dissimilarity_fishing_mean.jpg")

#for all regions faceted
(allreg_BC_total_dissimilarity_fishing_mean_facet <- ggplot(data = dissimilarities_temp_fishing_regstats) +
  labs(x = "Relative fishing pressure",  y = "BC total dissimilarity") +
  geom_point(aes(x = summed_tonnes_scaled_byreg, y = bray_curtis_dissimilarity_total_mean), alpha = 0.3) +
  geom_smooth(aes(x = summed_tonnes_scaled_byreg, y = bray_curtis_dissimilarity_total_mean), method = "lm") +
  facet_wrap(~Survey_Name_Season, scales = "free", ncol = 5) +
  theme_classic() +
    theme(legend.position = "null"))

#save
ggsave(allreg_BC_total_dissimilarity_fishing_mean_facet, path = here::here("figures","predictor_variables"), filename = "allreg_BC_total_dissimilarity_fishing_mean_facet.jpg", height = 13, width = 10, unit = "in")

```


###Temperature x fishing models
Best temp variables to carry forward: 

For raw temp (indicative of survey vs. survey):
-Mean temperature * survey



Mean temperature
```{r mean temp and fishing models of dissimilarity}


#fishing and temperature interaction
allreg_dissimilarity_fishing_mean_mod_sst_mean_avg <- lm(bray_curtis_dissimilarity_total_mean ~ summed_tonnes_scaled_byreg*yearly_mean_bypoint_avg, data = dissimilarities_temp_fishing_regstats)


        coefs <- data.frame(coef(summary(allreg_dissimilarity_fishing_mean_mod_sst_mean_avg)))

total_dissimilarity_sst_fishing_model_results_row <- data.table(matrix(c("sst", #depth
                                                            F, #scaled
                                                            "yearly_mean_bypoint_avg", #temp var
                                                            T, #fishing
                                                            T, #temp fishing interaction
                                                            "summed_tonnes_scaled_byreg*yearly_mean_bypoint_avg", #fixed effect
                                                            paste0(
                                                              round(coefs[1,1],3),
                                                              " ± ",
                                                              round(coefs[1,2],3)), #intercept
                                                            paste0(
                                                              round(coefs[3,1],3),
                                                              " ± ",
                                                              round(coefs[3,2],3)),#temp coef
                                                            round(coefs[3,4],3),#coef p value
                                                            paste0(
                                                              round(coefs[2,1],3),
                                                              " ± ",
                                                              round(coefs[2,2],3)),#fishing coef
                                                            round(coefs[2,4],3),#coef p value
                                                            paste0(
                                                              round(coefs[4,1],3),
                                                              " ± ",
                                                              round(coefs[4,2],3)), #interact coef
                                                            round(coefs[4,4],3), #coef p value
                                                round(summary(allreg_dissimilarity_fishing_mean_mod_sst_mean_avg)$r.squared,3), #r squared
                                                      round(AICc(allreg_dissimilarity_fishing_mean_mod_sst_mean_avg),3)#AICc
                                                            ),nrow =1))

total_dissimilarity_sst_fishing_model_results <- rbind(total_dissimilarity_sst_fishing_model_results, total_dissimilarity_sst_fishing_model_results_row, use.names = F)


#raw temp values AND interaction with both fishing and fixed survey unit

allreg_dissimilarity_fishing_mean_mod_sst_mean_avg_fixed_two_interact <- lm(bray_curtis_dissimilarity_total_mean ~ summed_tonnes_scaled_byreg*yearly_mean_bypoint_avg*survey_unit, data = dissimilarities_temp_fishing_regstats)


        coefs <- data.frame(coef(summary(allreg_dissimilarity_fishing_mean_mod_sst_mean_avg_fixed_two_interact)))

total_dissimilarity_sst_fishing_model_results_row <- data.table(matrix(c("sst", #depth
                                                            F, #scaled
                                                            "yearly_mean_bypoint_avg", #temp var
                                                            T, #fishing
                                                            T, #temp fishing interaction
                                                           "summed_tonnes_scaled_byreg*yearly_mean_bypoint_avg*survey_unit", #fixed effect
                                                            paste0(
                                                              round(coefs[1,1],3),
                                                              " ± ",
                                                              round(coefs[1,2],2)), #intercept
                                                            paste0(
                                                              round(coefs[3,1],3),
                                                              " ± ",
                                                              round(coefs[3,2],2)),#temp coef
                                                            round(coefs[3,4],3),#coef p value
                                                            paste0(
                                                              round(coefs[2,1],3),
                                                              " ± ",
                                                              round(coefs[2,2],2)),#fishing coef
                                                            round(coefs[2,4],3),#coef p value
                                                            paste0(
                                                              round(coefs[40,1],3),
                                                              " ± ",
                                                              round(coefs[40,2],2)), #interact coef (fishing and temp)
                                                            round(coefs[40,4],3), #coef p value (fishing and temp)
                                                round(r.squaredGLMM(allreg_dissimilarity_fishing_mean_mod_sst_mean_avg_fixed_two_interact)[[1]],3), #r squared
                                                            round(AICc(allreg_dissimilarity_fishing_mean_mod_sst_mean_avg_fixed_two_interact),3)#AICc
                                                            ),nrow =1))

total_dissimilarity_sst_fishing_model_results <- rbind(total_dissimilarity_sst_fishing_model_results, total_dissimilarity_sst_fishing_model_results_row, use.names = F)

#scaled temp values AND interaction with fishing but no interaction with fixed survey unit

allreg_dissimilarity_fishing_mean_mod_sst_mean_avg_fixed_temp_fish_interact <- lm(bray_curtis_dissimilarity_total_mean ~ summed_tonnes_scaled_byreg*yearly_mean_bypoint_avg+survey_unit, data = dissimilarities_temp_fishing_regstats)


        coefs <- data.frame(coef(summary(allreg_dissimilarity_fishing_mean_mod_sst_mean_avg_fixed_temp_fish_interact)))

total_dissimilarity_sst_fishing_model_results_row <- data.table(matrix(c("sst", #depth
                                                            F, #scaled
                                                            "yearly_mean_bypoint_avg", #temp var
                                                            T, #fishing
                                                            T, #temp fishing interaction
                                                            "summed_tonnes_scaled_byreg*yearly_mean_bypoint_avg+survey_unit", #fixed effect
                                                            paste0(
                                                              round(coefs[1,1],3),
                                                              " ± ",
                                                              round(coefs[1,2],3)), #intercept
                                                            paste0(
                                                              round(coefs[3,1],3),
                                                              " ± ",
                                                              round(coefs[3,2],3)),#temp coef
                                                            round(coefs[3,4],3),#coef p value
                                                            paste0(
                                                              round(coefs[2,1],3),
                                                              " ± ",
                                                              round(coefs[2,2],3)),#fishing coef
                                                            round(coefs[2,4],3),#coef p value
                                                            paste0(
                                                              round(coefs[38,1],3),
                                                              " ± ",
                                                              round(coefs[38,2],3)), #interact coef (fishing and temp)
                                                            round(coefs[38,4],3), #coef p value (fishing and temp)
                                                round(r.squaredGLMM(allreg_dissimilarity_fishing_mean_mod_sst_mean_avg_fixed_temp_fish_interact)[[1]],3), #r squared
                                                            round(AICc(allreg_dissimilarity_fishing_mean_mod_sst_mean_avg_fixed_temp_fish_interact),3)#AICc
                                                            ),nrow =1))

total_dissimilarity_sst_fishing_model_results <- rbind(total_dissimilarity_sst_fishing_model_results, total_dissimilarity_sst_fishing_model_results_row, use.names = F)

#scaled temp values AND interact with survey unit but no interact with fishing

allreg_dissimilarity_fishing_mean_mod_sst_mean_avg_fixed_temp_survey_interact <- lm(bray_curtis_dissimilarity_total_mean ~ summed_tonnes_scaled_byreg+yearly_mean_bypoint_avg*survey_unit, data = dissimilarities_temp_fishing_regstats)


        coefs <- data.frame(coef(summary(allreg_dissimilarity_fishing_mean_mod_sst_mean_avg_fixed_temp_survey_interact)))

total_dissimilarity_sst_fishing_model_results_row <- data.table(matrix(c("sst", #depth
                                                            F, #scaled
                                                            "yearly_mean_bypoint_avg", #temp var
                                                            T, #fishing
                                                            F, #temp fishing interaction
                                                          "summed_tonnes_scaled_byreg+yearly_mean_bypoint_avg*survey_unit", #fixed effect
                                                            paste0(
                                                              round(coefs[1,1],3),
                                                              " ± ",
                                                              round(coefs[1,2],3)), #intercept
                                                            paste0(
                                                              round(coefs[3,1],3),
                                                              " ± ",
                                                              round(coefs[3,2],3)),#temp coef
                                                            round(coefs[3,4],3),#coef p value
                                                            paste0(
                                                              round(coefs[2,1],3),
                                                              " ± ",
                                                              round(coefs[2,2],3)),#fishing coef
                                                            round(coefs[2,4],3),#coef p value
                                                            NA, #interact coef (fishing and temp)
                                                            NA, #coef p value (fishing and temp)
                                                round(summary(allreg_dissimilarity_fishing_mean_mod_sst_mean_avg_fixed_temp_survey_interact)$r.squared,3), #r squared
                                                
                                                            round(AICc(allreg_dissimilarity_fishing_mean_mod_sst_mean_avg_fixed_temp_survey_interact),3)#AICc
                                                            ),nrow =1))

total_dissimilarity_sst_fishing_model_results <- rbind(total_dissimilarity_sst_fishing_model_results, total_dissimilarity_sst_fishing_model_results_row, use.names = F)

#temp values AND interact with survey unit but no interact with fishing

allreg_dissimilarity_fishing_mean_mod_sst_mean_avg_fixed_fishing_survey_interact <- lm(bray_curtis_dissimilarity_total_mean ~yearly_mean_bypoint_avg + summed_tonnes_scaled_byreg*survey_unit, data = dissimilarities_temp_fishing_regstats)


        coefs <- data.frame(coef(summary(allreg_dissimilarity_fishing_mean_mod_sst_mean_avg_fixed_fishing_survey_interact)))

total_dissimilarity_sst_fishing_model_results_row <- data.table(matrix(c("sst", #depth
                                                            F, #scaled
                                                            "yearly_mean_bypoint_avg", #temp var
                                                            T, #fishing
                                                            F, #temp fishing interaction
                                                            "yearly_mean_bypoint_avg + summed_tonnes_scaled_byreg*survey_unit", #fixed effect
                                                            paste0(
                                                              round(coefs[1,1],3),
                                                              " ± ",
                                                              round(coefs[1,2],3)), #intercept
                                                            paste0(
                                                              round(coefs[3,1],3),
                                                              " ± ",
                                                              round(coefs[3,2],3)),#temp coef
                                                            round(coefs[3,4],3),#coef p value
                                                            paste0(
                                                              round(coefs[2,1],3),
                                                              " ± ",
                                                              round(coefs[2,2],3)),#fishing coef
                                                            round(coefs[2,4],3),#coef p value
                                                            NA, #interact coef (fishing and temp)
                                                            NA, #coef p value (fishing and temp)
                                                round(summary(allreg_dissimilarity_fishing_mean_mod_sst_mean_avg_fixed_fishing_survey_interact)$r.squared,3), #r squared
                                              
                                                            round(AICc(allreg_dissimilarity_fishing_mean_mod_sst_mean_avg_fixed_fishing_survey_interact),3)#AICc
                                                            ),nrow =1))

total_dissimilarity_sst_fishing_model_results <- rbind(total_dissimilarity_sst_fishing_model_results, total_dissimilarity_sst_fishing_model_results_row, use.names = F)

#scaled temp values, fixed effects for fishing and survey unit, no interact

allreg_dissimilarity_fishing_mean_mod_sst_mean_avg_fixed_fishing_no_interact <- lm(bray_curtis_dissimilarity_total_mean ~yearly_mean_bypoint_avg + summed_tonnes_scaled_byreg + survey_unit, data = dissimilarities_temp_fishing_regstats)


        coefs <- data.frame(coef(summary(allreg_dissimilarity_fishing_mean_mod_sst_mean_avg_fixed_fishing_no_interact)))

total_dissimilarity_sst_fishing_model_results_row <- data.table(matrix(c("sst", #depth
                                                            F, #scaled
                                                            "yearly_mean_bypoint_avg", #temp var
                                                            T, #fishing
                                                            F, #temp fishing interaction
                                                            "yearly_mean_bypoint_avg + summed_tonnes_scaled_byreg + survey_unit", #fixed effect
                                                            paste0(
                                                              round(coefs[1,1],3),
                                                              " ± ",
                                                              round(coefs[1,2],3)), #intercept
                                                            paste0(
                                                              round(coefs[3,1],3),
                                                              " ± ",
                                                              round(coefs[3,2],3)),#temp coef
                                                            round(coefs[3,4],3),#coef p value
                                                            paste0(
                                                              round(coefs[2,1],3),
                                                              " ± ",
                                                              round(coefs[2,2],3)),#fishing coef
                                                            round(coefs[2,4],3),#coef p value
                                                            NA, #interact coef (fishing and temp)
                                                            NA, #coef p value (fishing and temp)
                                                round(summary(allreg_dissimilarity_fishing_mean_mod_sst_mean_avg_fixed_fishing_no_interact)$r.squared,3), #r squared
                                                            round(AICc(allreg_dissimilarity_fishing_mean_mod_sst_mean_avg_fixed_fishing_no_interact),3)#AICc
                                                            ),nrow =1))

total_dissimilarity_sst_fishing_model_results <- rbind(total_dissimilarity_sst_fishing_model_results, total_dissimilarity_sst_fishing_model_results_row, use.names = F)
 


#temp only with interaction with fixed effect with survey unit

allreg_dissimilarity_sst_mean_avg_fixed <- lm(bray_curtis_dissimilarity_total_mean ~ yearly_mean_bypoint_avg*survey_unit, data = dissimilarities_temp_fishing_regstats)


        coefs <- data.frame(coef(summary(allreg_dissimilarity_sst_mean_avg_fixed)))

total_dissimilarity_sst_fishing_model_results_row <- data.table(matrix(c("sst", #depth
                                                            F, #scaled
                                                            "yearly_mean_bypoint_avg", #temp var
                                                            F, #fishing
                                                            F, #temp fishing interaction
                                                            "yearly_mean_bypoint_avg*survey_unit", #fixed effect
                                                            paste0(
                                                              round(coefs[1,1],3),
                                                              " ± ",
                                                              round(coefs[1,2],3)), #intercept
                                                            paste0(
                                                              round(coefs[2,1],3),
                                                              " ± ",
                                                              round(coefs[2,2],3)),#temp coef
                                                            round(coefs[2,4],3),#coef p value
                                                            NA,#fishing coef
                                                            NA,#coef p value
                                                            NA, #interact coef (fishing and temp)
                                                            NA, #coef p value (fishing and temp)
                                                round(summary(allreg_dissimilarity_sst_mean_avg_fixed)$r.squared,3),#r squared                                                            
                                                round(AICc(allreg_dissimilarity_sst_mean_avg_fixed),3)#AICc
                                                            ),nrow =1))

total_dissimilarity_sst_fishing_model_results <- rbind(total_dissimilarity_sst_fishing_model_results, total_dissimilarity_sst_fishing_model_results_row, use.names = F)

#temp only with fixed effect with survey unit, no interact

allreg_dissimilarity_sst_mean_avg_fixed_nointeract <- lm(bray_curtis_dissimilarity_total_mean ~ yearly_mean_bypoint_avg+survey_unit, data = dissimilarities_temp_fishing_regstats)


        coefs <- data.frame(coef(summary(allreg_dissimilarity_sst_mean_avg_fixed_nointeract)))

total_dissimilarity_sst_fishing_model_results_row <- data.table(matrix(c("sst", #depth
                                                            F, #scaled
                                                            "yearly_mean_bypoint_avg", #temp var
                                                            F, #fishing
                                                            F, #temp fishing interaction
                                                            "yearly_mean_bypoint_avg+survey_unit", #fixed effect
                                                            paste0(
                                                              round(coefs[1,1],3),
                                                              " ± ",
                                                              round(coefs[1,2],3)), #intercept
                                                            paste0(
                                                              round(coefs[2,1],3),
                                                              " ± ",
                                                              round(coefs[2,2],3)),#temp coef
                                                            round(coefs[2,4],3),#coef p value
                                                            NA,#fishing coef
                                                            NA,#coef p value
                                                            NA, #interact coef (fishing and temp)
                                                            NA, #coef p value (fishing and temp)
                                                round(r.squaredGLMM(allreg_dissimilarity_sst_mean_avg_fixed_nointeract)[[1]],3), #r squared
                                                            round(AICc(allreg_dissimilarity_sst_mean_avg_fixed_nointeract),3)),#AICc
                                                            nrow =1))

total_dissimilarity_sst_fishing_model_results <- rbind(total_dissimilarity_sst_fishing_model_results, total_dissimilarity_sst_fishing_model_results_row, use.names = F)
    

```



Show model ranking
```{r}
View(total_dissimilarity_sst_fishing_model_results)
total_dissimilarity_sst_fishing_model_results[,deltaAICc := as.numeric(AICc)-min(as.numeric(AICc))]

saveRDS(total_dissimilarity_sst_fishing_model_results, here::here("output", "total_dissimilarity_sst_fishing_model_results.Rds"))
#export as CSV
fwrite(total_dissimilarity_sst_fishing_model_results, here::here("output", "total_dissimilarity_sst_fishing_model_results.csv"))
```


Best overall model (by AICc)
  - Fishing as a fixed effect without temperature or survey unit
  - Raw mean temperature with interaction with fishing 

Carry these forward
---



