---
title: "Bray Balanced Dissimilarity Bottom Temperature & Fishing Interactions for All Regions"
output: html_notebook
---

Here we will link temperature and fishing to dissimilarity

This code is Script X for Kitchel et al. TITLE manuscript.

- This project is a collaborative effort to describe changes in taxonomic composition  of fish communities around the world--as sampled by bottom trawl surveys.

- Code by Zoë J. Kitchel

SESSION INFO TO DO

##Here, we examine bottom temperature and fishing pressure to predict annual BC balanced dissimilarity for all survey units


```{r setup}

library(data.table)
library(ggplot2)
library(lme4)
library(lmerTest)
library(MuMIn)


SODA_data_temp_avgs_full <- readRDS(here::here("data","Temperature","SODA_data_temp_avgs_full.rds"))
#don't feel confident about SODA temp before 1980 right now, start then

SODA_data_temp_avgs_full <- SODA_data_temp_avgs_full[year_for_avg >= 1980,]

SODA_data_temp_avgs_full[,year := year_for_avg]


#pull in average distance decay values
distances_dissimilarities_allyears.r <- readRDS(here::here("output","distance_decay","distances_dissimilarities_allyears.r.rds"))

#pull in fishing data
SAU_summed_tonnes.r <- readRDS(here::here("data","sea_around_us","SAU_summed_tonnes.rds"))

#pull in regional statistics
region_stats <- readRDS(here::here("output","region_stats", "region_stats.rds"))


```

Put into one data table
```{r link temp and fishing to avg values}
#temp
distances_dissimilarities_allyears.r.temp <- SODA_data_temp_avgs_full[distances_dissimilarities_allyears.r, on = c("year","survey_unit")]

#fishing
dissimilarities_temp_fishing <- SAU_summed_tonnes.r[distances_dissimilarities_allyears.r.temp, on = c("year", "survey_unit")]

#region_statistics
dissimilarities_temp_fishing_regstats <- dissimilarities_temp_fishing[region_stats, on = "survey_unit"]
dissimilarities_temp_fishing_regstats[,spp_num.s := scale(spp_num)]

dissimilarities_temp_fishing_regstats[,bray_curtis_dissimilarity_balanced_mean.s := scale(bray_curtis_dissimilarity_balanced_mean, center = F),survey_unit]

#save avg values
saveRDS(dissimilarities_temp_fishing_regstats, here::here("output","distance_decay","dissimilarities_temp_fishing_regstats.rds"))
```

Color pallete for survey units
```{r}

survey_unit.list <- levels(region_stats[,factor(survey_unit)])

pal_37 <- c(
  "#5A5156", #AI
  "#F6222E", #CHL
  "#F8A19F", #DFO-NF
  "#16FF32", #DFO-QCS
  "#DF00DB", #BITS-1
  "#DB8EDA", #BITS-4
  "#325A9B", #EBS
  "#3283FE", #EVHOE
  "#FEAF16", #FR-CGFS
  "#1C8356", #GMEX-Summer
  "#C4451C", #GOA
  "#85660D", #GRL-DE
  "#B0009F", #GSL-N
  "#BF79B8", #GSL-S
  "#1CBE4F", #ICE-GFS
  "#782AB6", #IE-IGFS
  "#90AD1C", #MEDITS
  "#6B003A", #NAM
  "#A75B00", #NEUS-Fall
  "#E3B072", #NEUS-Spring
  "#02E8B6", #NIGFS-1
  "#97E7D5", #NIGFS-4
  "#B00068", #Nor-BTS-3
  "#00B9E3", #NS-IBTS-1
  "#95E2F4", #NS-IBTS-3
  "#B3CE73", #NZ-CHAT
  "#689500", #NZ-ECSI
  "#AAF400", #NZ-WCSI
  "#AA0DFE", #PT-IBTS
  "#FA0087", #S-GEORG
  "#DEA0FD", #SCS-Summer
  "#FCEF88", #SEUS-fall
  "#A59405", #SEUS-spring
  "#FCE100", #SEUS-summer
  "#C075A6", #WCANN
  "#BDCDFF", #ZAF-ATL
  "#003EFF")  #ZAF-IND
  
color_link <- data.table(survey_unit = survey_unit.list,hex = pal_37)
```

Add names for plotting
```{r add names for plotting}


name_helper <- data.table(Survey_Name_Season = c("Aleutian Islands",
                                    "Baltic Sea Q1",
                                    "Baltic Sea Q4",
                                    "Chile",
                                    "Newfoundland",
                                    "Queen Charlotte Sound",
                                    "Eastern Bering Sea",
                                    "Bay of Biscay",
                                    "English Channel",
                                    "Gulf of Mexico",
                                    "Gulf of Alaska",
                                    "Greenland",
                                    "N Gulf of St. Lawrence",
                                    "S Gulf of St. Lawrence",
                                    "Iceland",
                                    "Irish Sea",
                                    "Mediterranean",
                                    "Namibia",
                                    "NE US Fall",
                                    "NE US Spring",
                                    "N Ireland Q1",
                                    "N Ireland Q4",
                                    "Norway",
                                    "N Sea Q1",
                                    "N Sea Q3",
                                    "Chatham Rise",
                                    "E Coast S Island NZ",
                                    "W Coast S Island NZ",
                                    "Portugal",
                                    "S Georgia Straight",
                                  "Scotian Shelf",
                                  "SE US Fall",
                                  "SE US Spring",
                                  "SE US Summer",
                                  "W Coast US",
                                  "Atlantic Ocean ZA",
                                  "Indian Ocean ZA"),
                          survey_unit = c(
                                  "AI",        
                                  "BITS-1",    
                                  "BITS-4",    
                                  "CHL",       
                                  "DFO-NF",    
                                  "DFO-QCS",   
                                  "EBS",       
                                  "EVHOE",     
                                  "FR-CGFS",   
                                  "GMEX-Summer",
                                  "GOA",       
                                  "GRL-DE",    
                                  "GSL-N",     
                                  "GSL-S",     
                                  "ICE-GFS",   
                                  "IE-IGFS",   
                                  "MEDITS",    
                                  "NAM",       
                                  "NEUS-Fall", 
                                  "NEUS-Spring",
                                  "NIGFS-1",   
                                  "NIGFS-4",   
                                  "Nor-BTS-3", 
                                  "NS-IBTS-1", 
                                  "NS-IBTS-3", 
                                  "NZ-CHAT",   
                                  "NZ-ECSI",   
                                  "NZ-WCSI",   
                                  "PT-IBTS",   
                                  "S-GEORG",   
                                  "SCS-SUMMER",
                                  "SEUS-fall", 
                                  "SEUS-spring",
                                  "SEUS-summer",
                                  "WCANN",     
                                  "ZAF-ATL",   
                                  "ZAF-IND"   
                          ))

color_link <- color_link[name_helper, on = "survey_unit"]

```

Table that will be populated with model results
```{r}
#depth, scaled, temp var, coefficient, p-value, marg_r^2, cond_r^2, AICc
balanced_dissimilarity_sbt_fishing_model_results <- data.table(Depth = as.character(),
                                                       Scaled = as.character(),
                                                       `Temp variable` = as.character(),
                                                       `Fishing`= as.character(),
                                                       `Temp fishing interaction` = as.character(),
                                                       `Mixed effect` = as.character(),
                                                       Intercept = as.numeric(),
                                                       `Temp coefficient` = as.numeric(),
                                                       `Coefficient p-value 1` = as.numeric(),
                                                       `Fishing coefficient` = as.numeric(),
                                                       `Coefficient p-value 2` = as.numeric(),
                                                       `Temp*fishing coefficient` = as.numeric(),
                                                       `Coefficient p-value 3` = as.numeric(),
                                                       `Marginal R-squared` = as.numeric(),
                                                       `Conditional R-squared` = as.numeric(),
                                                       AICc = as.numeric(),
                                                       `Degrees of freedom`=as.numeric())
                                                  
```

###How well does fishing alone predict mean annual BC balanced dissimilarity?
(Note this code is identical to dissimilarity_BC_balanced_temp_fishing_link_allregs.nb_OISST.Rmd between here and ~line 249, but I've kept it in to allow easier model comparisons. Edits will need to be made in both places)
```{r build fishing mod}

#random slope and intercept for survey
allreg_dissimilarity_fishing_mean_mod <- lmer(bray_curtis_dissimilarity_balanced_mean ~ summed_tonnes_scaled_byreg + (1 + summed_tonnes_scaled_byreg|survey_unit), data = dissimilarities_temp_fishing_regstats)

coefs <- data.frame(coef(summary(allreg_dissimilarity_fishing_mean_mod)))

balanced_dissimilarity_sbt_fishing_model_results_row <- data.table(matrix(c(NA, #depth
                                                            NA, #scaled
                                                            NA, #temp var
                                                            T, #fishing
                                                            NA, #temp fishing interaction
                                                            "survey_slope_intercept", #mixed effect
                                                            round(coefs[1,1],3), #intercept
                                                            NA,#temp coef
                                                            NA,#coef p value
                                                            round(coefs[2,1],3),#fishing coef
                                                            round(coefs[2,5],3),#coef p value
                                                            NA, #interact coef
                                                            NA, #coef p value
                                                round(r.squaredGLMM(allreg_dissimilarity_fishing_mean_mod)[[1]],3), #marginal r squared
                                                round(r.squaredGLMM(allreg_dissimilarity_fishing_mean_mod)[[2]],3), #conditional r-squared
                                                            round(AICc(allreg_dissimilarity_fishing_mean_mod),3),#AICc
                                                            6 #degrees of freedom 
                                                            ),nrow =1))

balanced_dissimilarity_sbt_fishing_model_results <- rbind(balanced_dissimilarity_sbt_fishing_model_results, balanced_dissimilarity_sbt_fishing_model_results_row, use.names = F)

#random intercept only
allreg_dissimilarity_fishing_mean_mod.intercept <- lmer(bray_curtis_dissimilarity_balanced_mean ~ summed_tonnes_scaled_byreg + (1|survey_unit), data = dissimilarities_temp_fishing_regstats)

coefs <- data.frame(coef(summary(allreg_dissimilarity_fishing_mean_mod.intercept)))

balanced_dissimilarity_sbt_fishing_model_results_row <- data.table(matrix(c(NA, #depth
                                                            NA, #scaled
                                                            NA, #temp var
                                                            T, #fishing
                                                            NA, #temp fishing interaction
                                                            "survey_intercept", #mixed effect
                                                            round(coefs[1,1],3), #intercept
                                                            NA,#temp coef
                                                            NA,#coef p value
                                                            round(coefs[2,1],3),#fishing coef
                                                            round(coefs[2,5],3),#coef p value
                                                            NA, #interact coef
                                                            NA, #coef p value
                                                round(r.squaredGLMM(allreg_dissimilarity_fishing_mean_mod.intercept)[[1]],3), #marginal r squared
                                                round(r.squaredGLMM(allreg_dissimilarity_fishing_mean_mod.intercept)[[2]],3), #conditional r-squared
                                                            round(AICc(allreg_dissimilarity_fishing_mean_mod.intercept),3),#AICc
                                                            4 #degrees of freedom 
                                                            ),nrow =1))

balanced_dissimilarity_sbt_fishing_model_results <- rbind(balanced_dissimilarity_sbt_fishing_model_results, balanced_dissimilarity_sbt_fishing_model_results_row, use.names = F)

#fixed effect for survey
allreg_dissimilarity_fishing_fixed_mean_mod <- lm(bray_curtis_dissimilarity_balanced_mean ~ summed_tonnes_scaled_byreg*survey_unit, data = dissimilarities_temp_fishing_regstats)

coefs <- data.frame(coef(summary(allreg_dissimilarity_fishing_fixed_mean_mod)))

balanced_dissimilarity_sbt_fishing_model_results_row <- data.table(matrix(c(NA, #depth
                                                            NA, #scaled
                                                            NA, #temp var
                                                            T, #fishing
                                                            NA, #temp fishing interaction
                                                            "survey_fixed", #mixed effect
                                                            round(coefs[1,1],3), #intercept
                                                            NA,#temp coef
                                                            NA,#coef p value
                                                            round(coefs[2,1],3),#fishing coef
                                                            round(coefs[2,4],3),#coef p value
                                                            NA, #interact coef
                                                            NA, #coef p value
                                                round(r.squaredGLMM(allreg_dissimilarity_fishing_fixed_mean_mod)[[1]],3), #marginal r squared
                                                round(r.squaredGLMM(allreg_dissimilarity_fishing_fixed_mean_mod)[[2]],3), #conditional r-squared
                                                            round(AICc(allreg_dissimilarity_fishing_fixed_mean_mod),3),#AICc
                                                            75 #degrees of freedom 
                                                            ),nrow =1))

balanced_dissimilarity_sbt_fishing_model_results <- rbind(balanced_dissimilarity_sbt_fishing_model_results, balanced_dissimilarity_sbt_fishing_model_results_row, use.names = F)

#no effect for survey
allreg_dissimilarity_fishing_mean_mod_lm <- lm(bray_curtis_dissimilarity_balanced_mean ~ summed_tonnes_scaled_byreg, data = dissimilarities_temp_fishing_regstats)

coefs <- data.frame(coef(summary(allreg_dissimilarity_fishing_mean_mod_lm)))

balanced_dissimilarity_sbt_fishing_model_results_row <- data.table(matrix(c(NA, #depth
                                                            NA, #scaled
                                                            NA, #temp var
                                                            T, #fishing
                                                            NA, #temp fishing interaction
                                                            "exclude_survey_completely", #mixed effect
                                                            round(coefs[1,1],3), #intercept
                                                            NA,#temp coef
                                                            NA,#coef p value
                                                            round(coefs[2,1],3),#fishing coef
                                                            round(coefs[2,4],3),#coef p value
                                                            NA, #interact coef
                                                            NA, #coef p value
                                                round(r.squaredGLMM(allreg_dissimilarity_fishing_mean_mod_lm)[[1]],3), #marginal r squared
                                                round(r.squaredGLMM(allreg_dissimilarity_fishing_mean_mod_lm)[[2]],3), #conditional r-squared
                                                            round(AICc(allreg_dissimilarity_fishing_mean_mod_lm),3),#AICc
                                                            3 #degrees of freedom 
                                                            ),nrow =1))

balanced_dissimilarity_sbt_fishing_model_results <- rbind(balanced_dissimilarity_sbt_fishing_model_results, balanced_dissimilarity_sbt_fishing_model_results_row, use.names = F)

#In years of higher than average fishing pressure, dissimilarity is lower (homogenization)

AICc(allreg_dissimilarity_fishing_mean_mod, allreg_dissimilarity_fishing_mean_mod.intercept, allreg_dissimilarity_fishing_fixed_mean_mod, allreg_dissimilarity_fishing_mean_mod_lm)


View(balanced_dissimilarity_sbt_fishing_model_results)


```
Best model includes survey unit as a fixed effect (R^2 = 93% of variation), but we don't care much about survey_unit (beyond acknowledging it makes a difference), so we will move forward with random intercept and random slope for survey unit.

As relative fishing pressure increases, dissimilarity decreases (more fished communities = more homogenized communities). 

###Temperature x fishing models
Best temp variables to carry forward: 

For raw temp (indicative of survey vs. survey):
-dissimilarity_BC_balanced_sbt_mean_mod


For scaled temp (indicative of within survey: cool or warm year? heterogenous or homogenous year)
-dissimilarity_BC_balanced_sbt_mean_scaled_mod (SD a bit better, but for comparison sake, we go with 2nd best = mean)

Starting with survey unit as random slope~fishing pressure and intercept (because it performed best above)
```{r temp and fishing models of dissimilarity}
#raw temp values

allreg_dissimilarity_fishing_mean_mod_sbt_mean_avg <- lmer(bray_curtis_dissimilarity_balanced_mean ~ summed_tonnes_scaled_byreg*yearly_mean_bypoint_avg + (1 + summed_tonnes_scaled_byreg|survey_unit), data = dissimilarities_temp_fishing_regstats)


        coefs <- data.frame(coef(summary(allreg_dissimilarity_fishing_mean_mod_sbt_mean_avg)))

balanced_dissimilarity_sbt_fishing_model_results_row <- data.table(matrix(c("sbt", #depth
                                                            F, #scaled
                                                            "mean", #temp var
                                                            T, #fishing
                                                            T, #temp fishing interaction
                                                            "survey_intercept", #mixed effect
                                                            round(coefs[1,1],3), #intercept
                                                            round(coefs[3,1],3),#temp coef
                                                            round(coefs[3,5],3),#coef p value
                                                            round(coefs[2,1],3),#fishing coef
                                                            round(coefs[2,5],3),#coef p value
                                                            round(coefs[4,1],3), #interact coef
                                                            round(coefs[4,5],3), #coef p value
                                                round(r.squaredGLMM(allreg_dissimilarity_fishing_mean_mod_sbt_mean_avg)[[1]],3), #marginal r squared
                                                round(r.squaredGLMM(allreg_dissimilarity_fishing_mean_mod_sbt_mean_avg)[[2]],3), #conditional r-squared
                                                            round(AICc(allreg_dissimilarity_fishing_mean_mod_sbt_mean_avg),3),#AICc
                                                            6 #degrees of freedom 
                                                            ),nrow =1))

balanced_dissimilarity_sbt_fishing_model_results <- rbind(balanced_dissimilarity_sbt_fishing_model_results, balanced_dissimilarity_sbt_fishing_model_results_row, use.names = F)

  #raw temp, no interaction
  allreg_dissimilarity_fishing_mean_mod_sbt_mean_avg.nointeract <- lmer(bray_curtis_dissimilarity_balanced_mean ~ summed_tonnes_scaled_byreg+yearly_mean_bypoint_avg + (1 + summed_tonnes_scaled_byreg|survey_unit), data = dissimilarities_temp_fishing_regstats)

  
          coefs <- data.frame(coef(summary(allreg_dissimilarity_fishing_mean_mod_sbt_mean_avg.nointeract)))

balanced_dissimilarity_sbt_fishing_model_results_row <- data.table(matrix(c("sbt", #depth
                                                            F, #scaled
                                                            "mean", #temp var
                                                            T, #fishing
                                                            F, #temp fishing interaction
                                                            "survey_intercept", #mixed effect
                                                            round(coefs[1,1],3), #intercept
                                                            round(coefs[3,1],3),#temp coef
                                                            round(coefs[3,5],3),#coef p value
                                                            round(coefs[2,1],3),#fishing coef
                                                            round(coefs[2,5],3),#coef p value
                                                            NA, #interact coef
                                                            NA, #coef p value
                                                round(r.squaredGLMM(allreg_dissimilarity_fishing_mean_mod_sbt_mean_avg.nointeract)[[1]],3), #marginal r squared
                                                round(r.squaredGLMM(allreg_dissimilarity_fishing_mean_mod_sbt_mean_avg.nointeract)[[2]],3), #conditional r-squared
                                                            round(AICc(allreg_dissimilarity_fishing_mean_mod_sbt_mean_avg.nointeract),3),#AICc
                                                            5 #degrees of freedom 
                                                            ),nrow =1))

balanced_dissimilarity_sbt_fishing_model_results <- rbind(balanced_dissimilarity_sbt_fishing_model_results, balanced_dissimilarity_sbt_fishing_model_results_row, use.names = F)
    
  #temp only
  allreg_dissimilarity_mod_sbt_mean_avg <- lmer(bray_curtis_dissimilarity_balanced_mean ~ yearly_mean_bypoint_avg + (1 + summed_tonnes_scaled_byreg|survey_unit), data = dissimilarities_temp_fishing_regstats)

  
               coefs <- data.frame(coef(summary(allreg_dissimilarity_mod_sbt_mean_avg)))

balanced_dissimilarity_sbt_fishing_model_results_row <- data.table(matrix(c("sbt", #depth
                                                            F, #scaled
                                                            "mean", #temp var
                                                            F, #fishing
                                                            F, #temp fishing interaction
                                                            "survey_intercept", #mixed effect
                                                            round(coefs[1,1],3), #intercept
                                                            round(coefs[2,1],3),#temp coef
                                                            round(coefs[2,5],3),#coef p value
                                                            NA,#fishing coef
                                                            NA,#coef p value
                                                            NA, #interact coef
                                                            NA, #coef p value
                                                round(r.squaredGLMM(allreg_dissimilarity_mod_sbt_mean_avg)[[1]],3), #marginal r squared
                                                round(r.squaredGLMM(allreg_dissimilarity_mod_sbt_mean_avg)[[2]],3), #conditional r-squared
                                                            round(AICc(allreg_dissimilarity_mod_sbt_mean_avg),3),#AICc
                                                            4 #degrees of freedom 
                                                            ),nrow =1))

balanced_dissimilarity_sbt_fishing_model_results <- rbind(balanced_dissimilarity_sbt_fishing_model_results, balanced_dissimilarity_sbt_fishing_model_results_row, use.names = F)
    
#scaled temp values
allreg_dissimilarity_fishing_mean_mod_sbt_mean_avg.s <- lmer(bray_curtis_dissimilarity_balanced_mean ~ summed_tonnes_scaled_byreg*yearly_mean_bypoint_avg.s + (1 + summed_tonnes_scaled_byreg|survey_unit), data = dissimilarities_temp_fishing_regstats)


                  coefs <- data.frame(coef(summary(allreg_dissimilarity_fishing_mean_mod_sbt_mean_avg.s)))

balanced_dissimilarity_sbt_fishing_model_results_row <- data.table(matrix(c("sbt", #depth
                                                            T, #scaled
                                                            "mean", #temp var
                                                            T, #fishing
                                                            T, #temp fishing interaction
                                                            "survey_intercept", #mixed effect
                                                            round(coefs[1,1],3), #intercept
                                                            round(coefs[3,1],3),#temp coef
                                                            round(coefs[3,5],3),#coef p value
                                                            round(coefs[2,1],3),#fishing coef
                                                            round(coefs[2,5],3),#coef p value
                                                            round(coefs[4,1],3), #interact coef
                                                            round(coefs[4,5],3), #coef p value
                                                round(r.squaredGLMM(allreg_dissimilarity_fishing_mean_mod_sbt_mean_avg.s)[[1]],3), #marginal r squared
                                                round(r.squaredGLMM(allreg_dissimilarity_fishing_mean_mod_sbt_mean_avg.s)[[2]],3), #conditional r-squared
                                                            round(AICc(allreg_dissimilarity_fishing_mean_mod_sbt_mean_avg.s),3),#AICc
                                                            6 #degrees of freedom 
                                                            ),nrow =1))

balanced_dissimilarity_sbt_fishing_model_results <- rbind(balanced_dissimilarity_sbt_fishing_model_results, balanced_dissimilarity_sbt_fishing_model_results_row, use.names = F)

  #scaled temp, no interaction
  allreg_dissimilarity_fishing_mean_mod_sbt_mean_avg.s.nointeract <- lmer(bray_curtis_dissimilarity_balanced_mean ~ summed_tonnes_scaled_byreg+yearly_mean_bypoint_avg.s + (1 + summed_tonnes_scaled_byreg|survey_unit), data = dissimilarities_temp_fishing_regstats)

  
          coefs <- data.frame(coef(summary(allreg_dissimilarity_fishing_mean_mod_sbt_mean_avg.s.nointeract)))

balanced_dissimilarity_sbt_fishing_model_results_row <- data.table(matrix(c("sbt", #depth
                                                            T, #scaled
                                                            "mean", #temp var
                                                            T, #fishing
                                                            F, #temp fishing interaction
                                                            "survey_intercept", #mixed effect
                                                            round(coefs[1,1],3), #intercept
                                                            round(coefs[3,1],3),#temp coef
                                                            round(coefs[3,5],3),#coef p value
                                                            round(coefs[2,1],3),#fishing coef
                                                            round(coefs[2,5],3),#coef p value
                                                            NA, #interact coef
                                                            NA, #coef p value
                                                round(r.squaredGLMM(allreg_dissimilarity_fishing_mean_mod_sbt_mean_avg.s.nointeract)[[1]],3), #marginal r squared
                                                round(r.squaredGLMM(allreg_dissimilarity_fishing_mean_mod_sbt_mean_avg.s.nointeract)[[2]],3), #conditional r-squared
                                                            round(AICc(allreg_dissimilarity_fishing_mean_mod_sbt_mean_avg.s.nointeract),3),#AICc
                                                            5 #degrees of freedom 
                                                            ),nrow =1))

balanced_dissimilarity_sbt_fishing_model_results <- rbind(balanced_dissimilarity_sbt_fishing_model_results, balanced_dissimilarity_sbt_fishing_model_results_row, use.names = F)
    
  #temp only
  allreg_dissimilarity_mod_sbt_mean_avg.s <- lmer(bray_curtis_dissimilarity_balanced_mean ~ summed_tonnes_scaled_byreg+yearly_mean_bypoint_avg.s + (1 + summed_tonnes_scaled_byreg|survey_unit), data = dissimilarities_temp_fishing_regstats)

  
          coefs <- data.frame(coef(summary(allreg_dissimilarity_mod_sbt_mean_avg.s)))

balanced_dissimilarity_sbt_fishing_model_results_row <- data.table(matrix(c("sbt", #depth
                                                            T, #scaled
                                                            "mean", #temp var
                                                            T, #fishing
                                                            F, #temp fishing interaction
                                                            "survey_intercept", #mixed effect
                                                            round(coefs[1,1],3), #intercept
                                                            round(coefs[3,1],3),#temp coef
                                                            round(coefs[3,5],3),#coef p value
                                                            NA,#fishing coef
                                                            NA,#coef p value
                                                            NA, #interact coef
                                                            NA, #coef p value
                                                round(r.squaredGLMM(allreg_dissimilarity_mod_sbt_mean_avg.s)[[1]],3), #marginal r squared
                                                round(r.squaredGLMM(allreg_dissimilarity_mod_sbt_mean_avg.s)[[2]],3), #conditional r-squared
                                                            round(AICc(allreg_dissimilarity_mod_sbt_mean_avg.s),3),#AICc
                                                            5 #degrees of freedom 
                                                            ),nrow =1))

balanced_dissimilarity_sbt_fishing_model_results <- rbind(balanced_dissimilarity_sbt_fishing_model_results, balanced_dissimilarity_sbt_fishing_model_results_row, use.names = F)
  
fishing_temp_model_ranking <- AICc(allreg_dissimilarity_fishing_mean_mod_sbt_mean_avg, allreg_dissimilarity_fishing_mean_mod_sbt_mean_avg.nointeract, allreg_dissimilarity_mod_sbt_mean_avg, allreg_dissimilarity_fishing_mean_mod_sbt_mean_avg.s,  allreg_dissimilarity_fishing_mean_mod_sbt_mean_avg.s.nointeract, allreg_dissimilarity_mod_sbt_mean_avg.s, allreg_dissimilarity_fishing_mean_mod.intercept)

#Best model: best model = allreg_dissimilarity_fishing_mean_mod (no temp; different than surface temperature which was no fishing)
View(balanced_dissimilarity_sbt_fishing_model_results)

summary(allreg_dissimilarity_fishing_mean_mod)

saveRDS(balanced_dissimilarity_sbt_fishing_model_results, here::here("output","balanced_dissimilarity_sbt_fishing_model_results.rds"))

```

Significant predictors: 
summed_tonnes_scaled_byreg (as the relative fishing pressure increases, bray dissimilarity decreases)

                            Estimate Std. Error        df t value Pr(>|t|)    
(Intercept)                 0.595931   0.019402 35.941509  30.714   <2e-16 ***
summed_tonnes_scaled_byreg -0.008120   0.005989 29.195211  -1.356    0.186    

What about r^2 value
```{r}
r.squaredGLMM(allreg_dissimilarity_fishing_mean_mod)

#marginal GLMM (variance explained by fixed effects)
#conditional GLMM (variance explained by entire model)

#           R2m       R2c
#[1,]       0.002      0.92. (very little explained by fishing or temperature, most explained by survey_units)
```



---


###What if we also tie in latitude and spp num?
```{r}
allreg_dissimilarity_fishing_mean_mod_sbt_max_avg_spp_num <- lmer(bray_curtis_dissimilarity_balanced_mean ~ summed_tonnes_scaled_byreg*yearly_max_bypoint_avg*spp_num.s + (1|survey_unit), data = dissimilarities_temp_fishing_regstats)

allreg_dissimilarity_fishing_mean_mod_sbt_max_avg_mid_lat <- lmer(bray_curtis_dissimilarity_balanced_mean ~ summed_tonnes_scaled_byreg*yearly_max_bypoint_avg*mid_lat + (1|survey_unit), data = dissimilarities_temp_fishing_regstats)

allreg_dissimilarity_fishing_mean_mod_sbt_max_avg_mid_lat_spp_num <- lmer(bray_curtis_dissimilarity_balanced_mean ~ summed_tonnes_scaled_byreg*yearly_max_bypoint_avg*mid_lat*spp_num.s + (1|survey_unit), data = dissimilarities_temp_fishing_regstats)

AIC(allreg_dissimilarity_fishing_mean_mod_sbt_max_avg_spp_num, allreg_dissimilarity_fishing_mean_mod_sbt_max_avg_mid_lat,allreg_dissimilarity_fishing_mean_mod_sbt_max_avg_mid_lat_spp_num)

summary(allreg_dissimilarity_fishing_mean_mod_sbt_max_avg_spp_num)
r.squaredGLMM(allreg_dissimilarity_fishing_mean_mod_sbt_max_avg_spp_num)

#no interaction with spp num?
allreg_dissimilarity_fishing_mean_mod_sbt_max_avg_spp_num_nointeract <- lmer(bray_curtis_dissimilarity_balanced_mean ~ summed_tonnes_scaled_byreg*yearly_max_bypoint_avg+spp_num.s + (1|survey_unit), data = dissimilarities_temp_fishing_regstats)

#exclude temperature?
allreg_dissimilarity_fishing_mean_mod_spp_num <- lmer(bray_curtis_dissimilarity_balanced_mean ~ summed_tonnes_scaled_byreg*spp_num.s + (1|survey_unit), data = dissimilarities_temp_fishing_regstats)

#exclude temp no interactions
allreg_dissimilarity_fishing_mean_mod_spp_num_nointeract <- lmer(bray_curtis_dissimilarity_balanced_mean ~ summed_tonnes_scaled_byreg+spp_num.s + (1|survey_unit), data = dissimilarities_temp_fishing_regstats)

AICc(allreg_dissimilarity_fishing_mean_mod_sbt_max_avg_spp_num_nointeract,allreg_dissimilarity_fishing_mean_mod_spp_num, allreg_dissimilarity_fishing_mean_mod_sbt_max_avg_spp_num,allreg_dissimilarity_fishing_mean_mod_spp_num_nointeract)

summary(allreg_dissimilarity_fishing_mean_mod_sbt_max_avg_spp_num_nointeract)
r.squaredGLMM(allreg_dissimilarity_fishing_mean_mod_sbt_max_avg_spp_num_nointeract)

#no temp performs better
summary(allreg_dissimilarity_fishing_mean_mod_spp_num_nointeract)
r.squaredGLMM(allreg_dissimilarity_fishing_mean_mod_spp_num_nointeract)

ggplot(dissimilarities_temp_fishing_regstats) +
  geom_point(aes(x = spp_num, y = bray_curtis_dissimilarity_balanced_mean, color = yearly_max_bypoint_avg)) +
  theme_classic()

ggplot(dissimilarities_temp_fishing_regstats) +
  geom_point(aes(x = yearly_max_bypoint_avg, y = bray_curtis_dissimilarity_balanced_mean, color = yearly_max_bypoint_avg)) +
  geom_smooth(aes(x = yearly_max_bypoint_avg, y = bray_curtis_dissimilarity_balanced_mean)) +
  theme_classic()
```

Patterns appear to be very survey dependent, and there's a lot of unexplained year to year variation.
- Warmer survey regions are more highly differentiated in general (makes sense, higher diversity )
- However, when we include spp num and fishing, temp drops out as a strong predictor
- As fishing pressure increases, dissimilarity decreases (more fishing = more homogenous)
- As # species increases, dissimilarity increases (more species overall = more heterogenous)
- From before, more species, more negative dissimilarity coefficient (more likely to be differentiating over time)

#####Plot just dissimilarity ~ relative fishing pressure coefficients of each region (Like figure 1b)
```{r}
#unique survey unit year combos
survey_unit_sampling_years <- unique(region_stats[,.(survey_unit, years_sampled)]) #unique survey unit year combos
                                    

 # see group coefficients and confidence intervals
fishing_model_coefs_reduced <- data.table(transform(as.data.frame(ranef(allreg_dissimilarity_fishing_mean_mod)), lwr = condval - 1.96*condsd, upr = condval + 1.96*condsd))
#https://stackoverflow.com/questions/69805532/extract-the-confidence-intervals-of-lmer-random-effects-plotted-with-dotplotra


#ONLY SLOPES
fishing_model_coefs_reduced <- fishing_model_coefs_reduced[term == "summed_tonnes_scaled_byreg",]

fishing_model_coefs_reduced[,survey_unit := grp][,summed_tonnes_scaled_byreg := condval]

#merge with duration of survey
fishing_model_coefs_reduced_length <- fishing_model_coefs_reduced[survey_unit_sampling_years, on = "survey_unit"]


fishing_model_coefs_reduced_length[,years_sampled := as.numeric(years_sampled)][,Slope_Direction := ifelse(summed_tonnes_scaled_byreg > 0, "Positive","Negative")]

#
fishing_model_coefs_reduced_length <- fishing_model_coefs_reduced_length[color_link, on = "survey_unit"]


#does it cross zero?
fishing_model_coefs_reduced_length[,significant := ifelse(lwr >0 & upr>0,T,ifelse(lwr<0 & upr<0,T,F))]

#delete all obs that are significant
fishing_model_coefs_reduced_length.r <- fishing_model_coefs_reduced_length[significant == F,]

#order table by coefficient
setorder(fishing_model_coefs_reduced_length, summed_tonnes_scaled_byreg)

BC_balanced_fishing_model_coefs_reduced_length.unique <- unique(fishing_model_coefs_reduced_length[,.(condval,condsd, lwr, upr, survey_unit, summed_tonnes_scaled_byreg, years_sampled, Slope_Direction, hex, Survey_Name_Season, significant)]) 

#extract color hexes
#year adj coef order
color_year_adj_order <-BC_balanced_fishing_model_coefs_reduced_length.unique[,hex]

#alphabetical order
BC_balanced_fishing_model_coefs_reduced_length.unique.alpha <- setorder(BC_balanced_fishing_model_coefs_reduced_length.unique, Survey_Name_Season)
color_alpha_order <- BC_balanced_fishing_model_coefs_reduced_length.unique.alpha[,hex]

saveRDS(BC_balanced_fishing_model_coefs_reduced_length.unique, here::here("output","region_stats","BC_balanced_fishing_model_coefs_reduced_length.unique.Rds"))
```

Bar Plot Coefficient Fishing LMER
```{r bar plot of fishing coefficients}
BC_BALANCED_Dissimilarity_Coef_FISHING_errorbar_reduced <- ggplot() +
    geom_errorbar(data = fishing_model_coefs_reduced_length, aes(x = reorder(Survey_Name_Season, summed_tonnes_scaled_byreg) , y = summed_tonnes_scaled_byreg, label = Survey_Name_Season, ymin = lwr, ymax = upr), fill = "grey", width = 0) + #add confidence intervals
  geom_point(data = fishing_model_coefs_reduced_length, aes(x = reorder(Survey_Name_Season, summed_tonnes_scaled_byreg) , y = summed_tonnes_scaled_byreg, label = Survey_Name_Season, 
      #size = years_sampled,
      fill = Slope_Direction), stat = 'identity', shape = 21, color = "black") +
  scale_fill_manual(values = c("white","black"), name = "Slope Direction") +
  geom_point(data = fishing_model_coefs_reduced_length.r, aes(x = reorder(Survey_Name_Season, summed_tonnes_scaled_byreg) , y = summed_tonnes_scaled_byreg, 
         #size = years_sampled, 
          label = Survey_Name_Season), stat = 'identity', fill = "grey",color = "grey", shape = 21) +
#  scale_size_binned(range = c(1,8), name = "Survey period\nlength") +
  geom_hline(yintercept = 0) +
  xlab("Survey unit") +
  ylab("Balanced BC dissimilarity ~\nrelative fishing pressure") +
  coord_flip() +
  theme_classic()

#+
#  theme(axis.text.y = element_text(colour = color_summed_tonnes_scaled_byreg_order, face = "bold"), axis.title.y = element_blank())

BC_BALANCED_Dissimilarity_Coef_FISHING_errorbar_reduced

#save as independent grob to merge with fishing
saveRDS(BC_BALANCED_Dissimilarity_Coef_FISHING_errorbar_reduced, here::here("figures","BC_BALANCED_Dissimilarity_Coef_FISHING_errorbar_reduced.Rds"))
```

###Link temp plots with fishing pressure plot
```{r link bottom temp to lollypop plot}
#Mean temp
BC_BALANCED_Dissimilarity_Coef_BOTTOMTEMP_errorbar_reduced <- readRDS(here::here("figures","BC_BALANCED_Dissimilarity_Coef_BOTTOMTEMP_errorbar_reduced.Rds"))

#relative mean temp
BC_BALANCED_Dissimilarity_Coef_scaled_BOTTOMTEMP_errorbar_reduced <- readRDS(here::here("figures","BC_BALANCED_Dissimilarity_Coef_scaled_BOTTOMTEMP_errorbar_reduced.Rds"))

#merge
bottom_temp_fishing_coefficients_lollypop <- cowplot::plot_grid(BC_BALANCED_Dissimilarity_Coef_BOTTOMTEMP_errorbar_reduced+theme(axis.title.y = element_blank(), legend.position = "null"), BC_BALANCED_Dissimilarity_Coef_scaled_BOTTOMTEMP_errorbar_reduced+theme(axis.title.y = element_blank(), legend.position = "null"), BC_BALANCED_Dissimilarity_Coef_FISHING_errorbar_reduced+theme(axis.title.y = element_blank(), legend.position = "null"), cowplot::get_legend(BC_BALANCED_Dissimilarity_Coef_FISHING_errorbar_reduced),
          nrow = 1, ncol = 4, rel_widths = c(1,1,1,0.2), rel_height = c(1,1,1,0.2))

ggsave(bottom_temp_fishing_coefficients_lollypop, path = "figures", filename = "bottom_temp_fishing_coefficients_lollypop.jpg", height = 5, width = 17, units = "in")
```

```{r link surface temp to lollypop plot}
#Mean temp
BC_BALANCED_Dissimilarity_Coef_surfaceTEMP_errorbar_reduced <- readRDS(here::here("figures","BC_BALANCED_Dissimilarity_Coef_surfaceTEMP_errorbar_reduced.Rds"))

#relative mean temp
BC_BALANCED_Dissimilarity_Coef_scaled_surfaceTEMP_errorbar_reduced <- readRDS(here::here("figures","BC_BALANCED_Dissimilarity_Coef_scaled_surfaceTEMP_errorbar_reduced.Rds"))

#merge
surface_temp_fishing_coefficients_lollypop <- cowplot::plot_grid(BC_BALANCED_Dissimilarity_Coef_surfaceTEMP_errorbar_reduced+theme(axis.title.y = element_blank(), legend.position = "null"), BC_BALANCED_Dissimilarity_Coef_scaled_surfaceTEMP_errorbar_reduced+theme(axis.title.y = element_blank(), legend.position = "null"), BC_BALANCED_Dissimilarity_Coef_FISHING_errorbar_reduced+theme(axis.title.y = element_blank(), legend.position = "null"), cowplot::get_legend(BC_BALANCED_Dissimilarity_Coef_FISHING_errorbar_reduced),
          nrow = 1, ncol = 4, rel_widths = c(1,1,1,0.2), rel_height = c(1,1,1,0.2))

ggsave(surface_temp_fishing_coefficients_lollypop, path = "figures", filename = "surface_temp_fishing_coefficients_lollypop.jpg", height = 5, width = 17, units = "in")
```

#CODE BELOW MAY NO LONGER BE RELEVANT
Plot this model with interaction
```{r}
library(interactions)
sim_slopes(allreg_dissimilarity_fishing_temp_mod_SDonly, pred = sbt_SD_scaled, modx = summed_tonnes_scaled, johnson_neyman = FALSE)

interact_plot(allreg_dissimilarity_fishing_mean_mod_sbt_max_avg, pred = yearly_max_bypoint_avg, modx = summed_tonnes_scaled_byreg, plot.points = T, colors = c("#b3ffe4","#00d68b","#003d28"), x.label = "Maximum Sea Surface Temperature", y.label = "Mean Annual Bray Curtis Dissimilarity") + theme_classic()

ggsave(filename = "interaction_fish_temp.png" ,height = 4, width = 10)

interact_plot(allreg_dissimilarity_fishing_temp_mod_SDonly, pred = sbt_SD_scaled, modx = summed_tonnes_scaled, linearity.check = TRUE, 
              plot.points = TRUE) #check linearity 

```
Summary: Steepest slope between temp and dissimilarity when fishing is high