---
title: "Dissimilarity Temperature & Fishing Interactions for All Regions"
output: html_notebook
---

Here we will finally link temperature and fishing to dissimilarity

This code is Script X for Kitchel et al. TITLE manuscript.

- This project is a collaborative effort to describe changes in taxonomic composition  of fish communities around the world--as sampled by bottom trawl surveys.

- Code by Zoë J. Kitchel

SESSION INFO TO DO

##Here, we try to use temperature and fishing pressure to predict annual dissimilarity for all survey units


```{r setup}

library(data.table)
library(ggplot2)
library(lme4)
library(lmerTest)
library(MuMIn)


OISST_data_temp_avgs_full <- readRDS(here::here("data","Temperature","OISST_data_temp_avgs_full.rds"))

OISST_data_temp_avgs_full[,year := year_for_avg]


#pull in average distance decay values
distances_dissimilarities_allyears.r <- readRDS(here::here("output","distance_decay","distances_dissimilarities_allyears.r.rds"))

#pull in fishing data
SAU_summed_tonnes.r <- readRDS(here::here("data","sea_around_us","SAU_summed_tonnes.rds"))

```

Put into one data table
```{r link temp and fishing to avg values}
#temp
distances_dissimilarities_allyears.r.temp <- OISST_data_temp_avgs_full[distances_dissimilarities_allyears.r, on = c("year","survey_unit")]

#fishing
dissimilarities_temp_fishing <- SAU_summed_tonnes.r[distances_dissimilarities_allyears.r.temp, on = c("year", "survey_unit")]

#region_statistics
dissimilarities_temp_fishing_regstats <- dissimilarities_temp_fishing[region_stats, on = "survey_unit"]
dissimilarities_temp_fishing_regstats[,spp_num.s := scale(spp_num)]

dissimilarities_temp_fishing_regstats[,bray_curtis_dissimilarity_balanced_mean.s := scale(bray_curtis_dissimilarity_balanced_mean, center = F),survey_unit]

#save avg values
saveRDS(dissimilarities_temp_fishing_regstats, here::here("output","distance_decay","dissimilarities_temp_fishing_regstats.rds"))
```

Color pallete for survey units
```{r}
pal_37 <- c(
  "#5A5156", #AI
  "#F6222E", #CHL
  "#F8A19F", #DFO-NF
  "#16FF32", #DFO-QCS
  "#DF00DB", #BITS-1
  "#DB8EDA", #BITS-4
  "#325A9B", #EBS
  "#3283FE", #EVHOE
  "#FEAF16", #FR-CGFS
  "#1C8356", #GMEX-Summer
  "#C4451C", #GOA
  "#85660D", #GRL-DE
  "#B0009F", #GSL-N
  "#BF79B8", #GSL-S
  "#1CBE4F", #ICE-GFS
  "#782AB6", #IE-IGFS
  "#90AD1C", #MEDITS
  "#6B003A", #NAM
  "#A75B00", #NEUS-Fall
  "#E3B072", #NEUS-Spring
  "#02E8B6", #NIGFS-1
  "#97E7D5", #NIGFS-4
  "#B00068", #Nor-BTS-3
  "#00B9E3", #NS-IBTS-1
  "#95E2F4", #NS-IBTS-3
  "#B3CE73", #NZ-CHAT
  "#689500", #NZ-ECSI
  "#AAF400", #NZ-WCSI
  "#AA0DFE", #PT-IBTS
  "#FA0087", #S-GEORG
  "#DEA0FD", #SCS-Summer
  "#FCEF88", #SEUS-fall
  "#A59405", #SEUS-spring
  "#FCE100", #SEUS-summer
  "#C075A6", #WCANN
  "#BDCDFF", #ZAF-ATL
  "#003EFF")  #ZAF-IND
```


###How well does fishing predict mean annual dissimilarity?
```{r temp all regions}
#for all regions
(allreg_dissimilarity_fishing_mean <- ggplot(data = dissimilarities_temp_fishing_regstats) +
  labs(x = "Relative fishing pressure",  y = "Dissimilarity") +
  geom_point(aes(x = summed_tonnes_scaled_byreg, y = bray_curtis_dissimilarity_balanced_mean, color = survey_unit), alpha = 0.3) +
  geom_line(aes(x = summed_tonnes_scaled_byreg, y = bray_curtis_dissimilarity_balanced_mean, color = survey_unit), stat="smooth", method = "lm") +
  scale_color_manual(values =  pal_37) +
  geom_smooth(aes(x = summed_tonnes_scaled_byreg, y = bray_curtis_dissimilarity_balanced_mean), method = "lm", se = F, color  = "black", size = 1) +
  theme_classic() +
    theme(legend.position = "none"))

#all one color points
(allreg_dissimilarity_fishing_mean_SIMPLIFIED <- ggplot(data = dissimilarities_temp_fishing) +
  labs(x = "Scaled Landings (tonnes)",  y = "Average Bray Curtis Dissimilarity") +
  geom_point(aes(x = summed_tonnes_scaled_byreg, y = bray_curtis_dissimilarity_balanced_mean), alpha = 0.3) +
  geom_smooth(aes(x = summed_tonnes_scaled_byreg, y = bray_curtis_dissimilarity_balanced_mean), method = "lm", se = F, color  = "black") +
  theme_classic() +
    theme(legend.position = "none"))

#random effect for survey
allreg_dissimilarity_fishing_mean_mod <- lmer(bray_curtis_dissimilarity_balanced_mean ~ summed_tonnes_scaled_byreg + (1|survey_unit), data = dissimilarities_temp_fishing_regstats)

#fixed effect for survey
allreg_dissimilarity_fishing_fixed_mean_mod <- lm(bray_curtis_dissimilarity_balanced_mean ~ summed_tonnes_scaled_byreg*survey_unit, data = dissimilarities_temp_fishing_regstats)

#no effect for survey
allreg_dissimilarity_fishing_mean_mod_lm <- lm(bray_curtis_dissimilarity_balanced_mean ~ summed_tonnes_scaled_byreg, data = dissimilarities_temp_fishing_regstats)

#In years of higher than average fishing pressure, dissimilarity is lower (homogenization)

AICc(allreg_dissimilarity_fishing_mean_mod, allreg_dissimilarity_fishing_fixed_mean_mod, allreg_dissimilarity_fishing_mean_mod_lm)

summary(allreg_dissimilarity_fishing_mean_mod)

#
r.squaredGLMM(allreg_dissimilarity_fishing_mean_mod)
r.squaredGLMM(allreg_dissimilarity_fishing_fixed_mean_mod)
r.squaredGLMM(allreg_dissimilarity_fishing_mean_mod_lm)




```
Best model includes survey unit as a fixed effect (R^2 = 93% of variation)


###Temperature x fishing
Best temp variables to carry forward: 

For raw temp (indicative of survey vs. survey):
-dissimilarity_BC_balanced_sst_maxSD_mod (Temp heterogeneity)
-dissimilarity_BC_balanced_sst_max_mod (Maximum temperature)

For scaled temp (indicative of within survey: cool or warm year? heterogenous or homogenous year)
-dissimilarity_BC_balanced_sst_maxSD_scaled_mod
-dissimilarity_BC_balanced_sst_min_scaled_mod

Starting with survey unit as random effect again
```{r temp and fishing models of dissimilarity}
#raw temp values
allreg_dissimilarity_fishing_mean_mod_sst_maxSD <- lmer(bray_curtis_dissimilarity_balanced_mean ~ summed_tonnes_scaled_byreg*yearly_max_bypoint_SD + (1|survey_unit), data = dissimilarities_temp_fishing)

allreg_dissimilarity_fishing_mean_mod_sst_max_avg <- lmer(bray_curtis_dissimilarity_balanced_mean ~ summed_tonnes_scaled_byreg*yearly_max_bypoint_avg + (1|survey_unit), data = dissimilarities_temp_fishing)

#scaled temp values
allreg_dissimilarity_fishing_mean_mod_sst_maxSD.s <- lmer(bray_curtis_dissimilarity_balanced_mean ~ summed_tonnes_scaled_byreg*yearly_max_bypoint_SD.s + (1|survey_unit), data = dissimilarities_temp_fishing)

allreg_dissimilarity_fishing_mean_mod_sst_min_avg.s <- lmer(bray_curtis_dissimilarity_balanced_mean ~ summed_tonnes_scaled_byreg*yearly_min_bypoint_avg.s + (1|survey_unit), data = dissimilarities_temp_fishing)

AICc(allreg_dissimilarity_fishing_mean_mod_sst_maxSD, allreg_dissimilarity_fishing_mean_mod_sst_max_avg, allreg_dissimilarity_fishing_mean_mod_sst_maxSD.s, allreg_dissimilarity_fishing_mean_mod_sst_min_avg.s)

#Best model: allreg_dissimilarity_fishing_mean_mod_sst_max_avg, delta AICc = 6.1

summary(allreg_dissimilarity_fishing_mean_mod_sst_max_avg)

```

Significant predictors: 
summed_tonnes_scaled_byreg (as landings (proxy for fishing effort) increase, bray dissimilarity decreases)
sst_max_raw (as the max temp of the year before increases, bray dissimilarity increases)
summed_tonnes_scaled_byreg: sst_max_raw (this interaction is a significant predictor)

                                                    Estimate Std. Error         df t value Pr(>|t|)    
(Intercept)                                        5.376e-01  3.622e-02  1.386e+02  14.842  < 2e-16 ***
summed_tonnes_scaled_byreg                        -4.155e-02  8.180e-03  6.686e+02  -5.079 4.92e-07 ***
yearly_max_bypoint_avg                             3.191e-03  1.691e-03  2.876e+02   1.887   0.0601 .  
summed_tonnes_scaled_byreg:yearly_max_bypoint_avg  1.874e-03  4.044e-04  6.692e+02   4.635 4.30e-06 ***

What about r^2 value
```{r}
r.squaredGLMM(allreg_dissimilarity_fishing_mean_mod_sst_max_avg)

#marginal GLMM (variance explained by fixed effects)
#conditional GLMM (variance explained by entire model)

#           R2m       R2c
#[1,] 0.03197428 0.8927886 (very little explained by fishing or temperature, most explained by survey_units)
```

#Plot temp only 

#Plot fishing only

allreg_dissimilarity_fishing_temp_mod_fishonly <- lmer(bray_curtis_dissimilarity_balanced_mean ~ summed_tonnes_scaled_byreg + (1|survey_unit), data = dissimilarities_temp_fishing)

bray curtis dissimilarity = -0.005256 * tonnes + 0.6

```{r}
ggplot(data = dissimilarities_temp_fishing) +
  geom_point(aes(x=summed_tonnes_scaled_byreg, y = bray_curtis_dissimilarity_balanced_mean)) +
  geom_smooth(aes(x=summed_tonnes_scaled_byreg, y = bray_curtis_dissimilarity_balanced_mean), method = "lm") +
  theme_classic()

ggplot(data = dissimilarities_temp_fishing) +
  geom_point(aes(x=summed_tonnes_scaled_byreg, y = bray_curtis_dissimilarity_balanced_mean)) +
  geom_smooth(aes(x=summed_tonnes_scaled_byreg, y = bray_curtis_dissimilarity_balanced_mean), method = "lm") +
  facet_wrap(~survey_unit, scales = "free")
  theme_classic()
  
 # ggsave(filename = "facet_fish_model.png", height = 10, width = 10, unit = "in")
  
  
  #sst_max
  
  ggplot(data = dissimilarities_temp_fishing) +
  geom_point(aes(x=yearly_max_bypoint_avg, y = bray_curtis_dissimilarity_balanced_mean, color = survey_unit)) +
  geom_smooth(aes(x=yearly_max_bypoint_avg, y = bray_curtis_dissimilarity_balanced_mean, color = survey_unit), method = "lm", se = F) +
  theme_classic()
```




#CODE BELOW MAY NO LONGER BE RELEVANT
Plot this model with interaction
```{r}
library(interactions)
sim_slopes(allreg_dissimilarity_fishing_temp_mod_SDonly, pred = sst_SD_scaled, modx = summed_tonnes_scaled, johnson_neyman = FALSE)

interact_plot(allreg_dissimilarity_fishing_mean_mod_sst_max_avg, pred = yearly_max_bypoint_avg, modx = summed_tonnes_scaled_byreg, plot.points = T, colors = c("#b3ffe4","#00d68b","#003d28"), x.label = "Maximum Sea Surface Temperature", y.label = "Mean Annual Bray Curtis Dissimilarity") + theme_classic()

ggsave(filename = "interaction_fish_temp.png" ,height = 4, width = 10)

interact_plot(allreg_dissimilarity_fishing_temp_mod_SDonly, pred = sst_SD_scaled, modx = summed_tonnes_scaled, linearity.check = TRUE, 
              plot.points = TRUE) #check linearity 

```
Summary: Steepest slope between temp and dissimilarity when fishing is high

###What if we also tie in latitude and spp num?
```{r}
allreg_dissimilarity_fishing_mean_mod_sst_max_avg_spp_num <- lmer(bray_curtis_dissimilarity_balanced_mean ~ summed_tonnes_scaled_byreg*yearly_max_bypoint_avg*spp_num.s + (1|survey_unit), data = dissimilarities_temp_fishing_regstats)

allreg_dissimilarity_fishing_mean_mod_sst_max_avg_mid_lat <- lmer(bray_curtis_dissimilarity_balanced_mean ~ summed_tonnes_scaled_byreg*yearly_max_bypoint_avg*mid_lat + (1|survey_unit), data = dissimilarities_temp_fishing_regstats)

allreg_dissimilarity_fishing_mean_mod_sst_max_avg_mid_lat_spp_num <- lmer(bray_curtis_dissimilarity_balanced_mean ~ summed_tonnes_scaled_byreg*yearly_max_bypoint_avg*mid_lat*spp_num.s + (1|survey_unit), data = dissimilarities_temp_fishing_regstats)

AIC(allreg_dissimilarity_fishing_mean_mod_sst_max_avg_spp_num, allreg_dissimilarity_fishing_mean_mod_sst_max_avg_mid_lat,allreg_dissimilarity_fishing_mean_mod_sst_max_avg_mid_lat_spp_num)

summary(allreg_dissimilarity_fishing_mean_mod_sst_max_avg_spp_num)
r.squaredGLMM(allreg_dissimilarity_fishing_mean_mod_sst_max_avg_spp_num)

#no interaction with spp num?
allreg_dissimilarity_fishing_mean_mod_sst_max_avg_spp_num_nointeract <- lmer(bray_curtis_dissimilarity_balanced_mean ~ summed_tonnes_scaled_byreg*yearly_max_bypoint_avg+spp_num.s + (1|survey_unit), data = dissimilarities_temp_fishing_regstats)

#exclude temperature?
allreg_dissimilarity_fishing_mean_mod_spp_num <- lmer(bray_curtis_dissimilarity_balanced_mean ~ summed_tonnes_scaled_byreg*spp_num.s + (1|survey_unit), data = dissimilarities_temp_fishing_regstats)

#exclude temp no interactions
allreg_dissimilarity_fishing_mean_mod_spp_num_nointeract <- lmer(bray_curtis_dissimilarity_balanced_mean ~ summed_tonnes_scaled_byreg+spp_num.s + (1|survey_unit), data = dissimilarities_temp_fishing_regstats)

AICc(allreg_dissimilarity_fishing_mean_mod_sst_max_avg_spp_num_nointeract,allreg_dissimilarity_fishing_mean_mod_spp_num, allreg_dissimilarity_fishing_mean_mod_sst_max_avg_spp_num,allreg_dissimilarity_fishing_mean_mod_spp_num_nointeract)

summary(allreg_dissimilarity_fishing_mean_mod_sst_max_avg_spp_num_nointeract)
r.squaredGLMM(allreg_dissimilarity_fishing_mean_mod_sst_max_avg_spp_num_nointeract)

#no temp performs better
summary(allreg_dissimilarity_fishing_mean_mod_spp_num_nointeract)
r.squaredGLMM(allreg_dissimilarity_fishing_mean_mod_spp_num_nointeract)

ggplot(dissimilarities_temp_fishing_regstats) +
  geom_point(aes(x = spp_num, y = bray_curtis_dissimilarity_balanced_mean, color = yearly_max_bypoint_avg)) +
  theme_classic()

ggplot(dissimilarities_temp_fishing_regstats) +
  geom_point(aes(x = yearly_max_bypoint_avg, y = bray_curtis_dissimilarity_balanced_mean, color = yearly_max_bypoint_avg)) +
  geom_smooth(aes(x = yearly_max_bypoint_avg, y = bray_curtis_dissimilarity_balanced_mean)) +
  theme_classic()
```

Patterns appear to be very survey dependent, and there's a lot of unexplained year to year variation.
- Warmer survey regions are more highly differentiated in general (makes sense, higher diversity )
- However, when we include spp num and fishing, temp drops out as a strong predictor
- As fishing pressure increases, dissimilarity decreases (more fishing = more homogenous)
- As # species increases, dissimilarity increases (more species overall = more heterogenous)
- From before, more species, more negative dissimilarity coefficient (more likely to be differentiating over time)


