---
title: "Annual Dissimilarity vs. Temperature for all Survey Units"
output: html_notebook
---

This code is Script X for Kitchel et al. TITLE manuscript.

- This project is a collaborative effort to describe changes in taxonomic composition  of fish communities around the world--as sampled by bottom trawl surveys.

- Code by Zoë J. Kitchel

SESSION INFO TO DO

##Here, we try to use temperature to predict annual dissimilarity for all survey units

```{r setup}

library(data.table)
library(ggplot2)
library(lme4)
library(lmerTest)
library(MuMIn)


OISST_data_temp_avgs_full <- readRDS(here::here("data","Temperature","OISST_data_temp_avgs_full.rds"))

OISST_data_temp_avgs_full[,year := year_for_avg]

#pull in average distance decay values
distances_dissimilarities_allyears.r <- readRDS(here::here("output","distance_decay","distances_dissimilarities_allyears.r.rds"))

```

Put into one data table
```{r link temp to avg values}
dissimilarities_temp <- OISST_data_temp_avgs_full[distances_dissimilarities_allyears.r, on = c("year","survey_unit")]
```

Color pallete for survey units
```{r}
pal_37 <- c(
  "#5A5156", #AI
  "#F6222E", #CHL
  "#F8A19F", #DFO-NF
  "#16FF32", #DFO-QCS
  "#DF00DB", #BITS-1
  "#DB8EDA", #BITS-4
  "#325A9B", #EBS
  "#3283FE", #EVHOE
  "#FEAF16", #FR-CGFS
  "#1C8356", #GMEX-Summer
  "#C4451C", #GOA
  "#85660D", #GRL-DE
  "#B0009F", #GSL-N
  "#BF79B8", #GSL-S
  "#1CBE4F", #ICE-GFS
  "#782AB6", #IE-IGFS
  "#90AD1C", #MEDITS
  "#6B003A", #NAM
  "#A75B00", #NEUS-Fall
  "#E3B072", #NEUS-Spring
  "#02E8B6", #NIGFS-1
  "#97E7D5", #NIGFS-4
  "#B00068", #Nor-BTS-3
  "#00B9E3", #NS-IBTS-1
  "#95E2F4", #NS-IBTS-3
  "#B3CE73", #NZ-CHAT
  "#689500", #NZ-ECSI
  "#AAF400", #NZ-WCSI
  "#AA0DFE", #PT-IBTS
  "#FA0087", #S-GEORG
  "#DEA0FD", #SCS-Summer
  "#FCEF88", #SEUS-fall
  "#A59405", #SEUS-spring
  "#FCE100", #SEUS-summer
  "#C075A6", #WCANN
  "#BDCDFF", #ZAF-ATL
  "#003EFF")  #ZAF-IND
```

How does this trend hold up for all regions?
###UNSCALED
###Avg of grid cells
####Mean temperature
```{r mean sst}
#for all regions
dissimilarity_BC_balanced_sst_mean <- ggplot(data = dissimilarities_temp) +
  labs(x = "Mean surface temperature (˚C)",  y = "Bray Curtis balanced dissimilarity") +
  geom_point(aes(x = yearly_mean_bypoint_avg, y = bray_curtis_dissimilarity_balanced_mean, color = survey_unit), alpha = 0.3) +
  geom_line(aes(x = yearly_mean_bypoint_avg, y = bray_curtis_dissimilarity_balanced_mean, color = survey_unit), stat="smooth", method = "lm") +
  scale_color_manual(values = c(pal_37)) +
  geom_smooth(aes(x = yearly_mean_bypoint_avg, y = bray_curtis_dissimilarity_balanced_mean), method = "lm", se = F, color  = "black") +
  theme_classic()

ggsave(dissimilarity_BC_balanced_sst_mean, path  = here::here("figures","distance_decay"), filename = "dissimilarity_BC_balanced_sst_mean.jpg", height = 4, width = 9, unit = "in") 

dissimilarity_BC_balanced_sst_mean_facet <- ggplot(data = dissimilarities_temp) +
  labs(x = "Mean surface temperature (˚C)",  y = "Bray Curtis balanced dissimilarity") +
  geom_point(aes(x = yearly_mean_bypoint_avg, y = bray_curtis_dissimilarity_balanced_mean), alpha = 0.3) +
  geom_smooth(aes(x = yearly_mean_bypoint_avg, y = bray_curtis_dissimilarity_balanced_mean), method = "lm") +
  facet_wrap(~survey_unit, scales = "free") +
  theme_classic()

ggsave(dissimilarity_BC_balanced_sst_mean_facet, path  = here::here("figures","distance_decay"), filename = "dissimilarity_BC_balanced_sst_mean_facet.jpg", height = 10, width = 11, unit = "in")

dissimilarity_BC_balanced_sst_mean_mod <- lmer(bray_curtis_dissimilarity_balanced_mean ~ yearly_mean_bypoint_avg + (1|survey_unit), data = dissimilarities_temp)

summary(dissimilarity_BC_balanced_sst_mean_mod) #insignificant relationship
```

####Min temperature
```{r min sst}
#for all regions
dissimilarity_BC_balanced_sst_min <- ggplot(data = dissimilarities_temp) +
  labs(x = "Min surface temperature (˚C)",  y = "Bray Curtis balanced dissimilarity") +
  geom_point(aes(x = yearly_min_bypoint_avg, y = bray_curtis_dissimilarity_balanced_mean, color = survey_unit), alpha = 0.3) +
  geom_line(aes(x = yearly_min_bypoint_avg, y = bray_curtis_dissimilarity_balanced_mean, color = survey_unit), stat="smooth", method = "lm") +
  scale_color_manual(values = c(pal_37)) +
  geom_smooth(aes(x = yearly_min_bypoint_avg, y = bray_curtis_dissimilarity_balanced_mean), method = "lm", se = F, color  = "black") +
  theme_classic()

ggsave(dissimilarity_BC_balanced_sst_min, path  = here::here("figures","distance_decay"), filename = "dissimilarity_BC_balanced_sst_min.jpg", height = 4, width = 9, unit = "in") 

dissimilarity_BC_balanced_sst_min_facet <- ggplot(data = dissimilarities_temp) +
  labs(x = "Min surface temperature (˚C)",  y = "Bray Curtis balanced dissimilarity") +
  geom_point(aes(x = yearly_min_bypoint_avg, y = bray_curtis_dissimilarity_balanced_mean), alpha = 0.3) +
  geom_smooth(aes(x = yearly_min_bypoint_avg, y = bray_curtis_dissimilarity_balanced_mean), method = "lm") +
  facet_wrap(~survey_unit, scales = "free") +
  theme_classic()

ggsave(dissimilarity_BC_balanced_sst_min_facet, path  = here::here("figures","distance_decay"), filename = "dissimilarity_BC_balanced_sst_min_facet.jpg", height = 10, width = 11, unit = "in")

dissimilarity_BC_balanced_sst_min_mod <- lmer(bray_curtis_dissimilarity_balanced_mean ~ yearly_min_bypoint_avg + (1|survey_unit), data = dissimilarities_temp)

summary(dissimilarity_BC_balanced_sst_min_mod) #nearly significant relationship (0.09)
```

####Max sst
```{r max sst}
#for all regions
dissimilarity_BC_balanced_sst_max <- ggplot(data = dissimilarities_temp) +
  labs(x = "Max surface temperature (˚C)",  y = "Bray Curtis balanced dissimilarity") +
  geom_point(aes(x = yearly_max_bypoint_avg, y = bray_curtis_dissimilarity_balanced_mean, color = survey_unit), alpha = 0.3) +
  geom_line(aes(x = yearly_max_bypoint_avg, y = bray_curtis_dissimilarity_balanced_mean, color = survey_unit), stat="smooth", method = "lm") +
  scale_color_manual(values = c(pal_37)) +
  geom_smooth(aes(x = yearly_max_bypoint_avg, y = bray_curtis_dissimilarity_balanced_mean), method = "lm", se = F, color  = "black") +
  theme_classic()

ggsave(dissimilarity_BC_balanced_sst_max, path  = here::here("figures","distance_decay"), filename = "dissimilarity_BC_balanced_sst_max.jpg", height = 4, width = 9, unit = "in") 

dissimilarity_BC_balanced_sst_max_facet <- ggplot(data = dissimilarities_temp) +
  labs(x = "Max surface temperature (˚C)",  y = "Bray Curtis balanced dissimilarity") +
  geom_point(aes(x = yearly_max_bypoint_avg, y = bray_curtis_dissimilarity_balanced_mean), alpha = 0.3) +
  geom_smooth(aes(x = yearly_max_bypoint_avg, y = bray_curtis_dissimilarity_balanced_mean), method = "lm") +
  facet_wrap(~survey_unit, scales = "free") +
  theme_classic()

ggsave(dissimilarity_BC_balanced_sst_max_facet, path  = here::here("figures","distance_decay"), filename = "dissimilarity_BC_balanced_sst_max_facet.jpg", height = 10, width = 11, unit = "in")

dissimilarity_BC_balanced_sst_max_mod <- lmer(bray_curtis_dissimilarity_balanced_mean ~ yearly_max_bypoint_avg + (1|survey_unit), data = dissimilarities_temp)

summary(dissimilarity_BC_balanced_sst_max_mod) #nearly significant relationship (0.51)

```

####Seas sst
```{r seas sst}
#for all regions
dissimilarity_BC_balanced_sst_seas <- ggplot(data = dissimilarities_temp) +
  labs(x = "Seasonality of surface temperature (˚C)",  y = "Bray Curtis balanced dissimilarity") +
  geom_point(aes(x = yearly_seas_bypoint_avg, y = bray_curtis_dissimilarity_balanced_mean, color = survey_unit), alpha = 0.3) +
  geom_line(aes(x = yearly_seas_bypoint_avg, y = bray_curtis_dissimilarity_balanced_mean, color = survey_unit), stat="smooth", method = "lm") +
  scale_color_manual(values = c(pal_37)) +
  geom_smooth(aes(x = yearly_seas_bypoint_avg, y = bray_curtis_dissimilarity_balanced_mean), method = "lm", se = F, color  = "black") +
  theme_classic()

ggsave(dissimilarity_BC_balanced_sst_seas, path  = here::here("figures","distance_decay"), filename = "dissimilarity_BC_balanced_sst_seas.jpg", height = 4, width = 9, unit = "in") 

dissimilarity_BC_balanced_sst_seas_facet <- ggplot(data = dissimilarities_temp) +
  labs(x = "Seasonality of surface temperature (˚C)",  y = "Bray Curtis balanced dissimilarity") +
  geom_point(aes(x = yearly_seas_bypoint_avg, y = bray_curtis_dissimilarity_balanced_mean), alpha = 0.3) +
  geom_smooth(aes(x = yearly_seas_bypoint_avg, y = bray_curtis_dissimilarity_balanced_mean), method = "lm") +
  facet_wrap(~survey_unit, scales = "free") +
  theme_classic()

ggsave(dissimilarity_BC_balanced_sst_seas_facet, path  = here::here("figures","distance_decay"), filename = "dissimilarity_BC_balanced_sst_seas_facet.jpg", height = 10, width = 11, unit = "in")


dissimilarity_BC_balanced_sst_seas_mod <- lmer(bray_curtis_dissimilarity_balanced_mean ~ yearly_seas_bypoint_avg + (1|survey_unit), data = dissimilarities_temp)

summary(dissimilarity_BC_balanced_sst_seas_mod) #nearly significant relationship (0.08)

```

###Standard deviation of grid cells
####Mean temperature SD
```{r mean sst SD}
#for all regions
dissimilarity_BC_balanced_sst_meanSD <- ggplot(data = dissimilarities_temp) +
  labs(x = "Mean surface temperature SD (˚C)",  y = "Bray Curtis balanced dissimilarity") +
  geom_point(aes(x = yearly_mean_bypoint_SD, y = bray_curtis_dissimilarity_balanced_mean, color = survey_unit), alpha = 0.3) +
  geom_line(aes(x = yearly_mean_bypoint_SD, y = bray_curtis_dissimilarity_balanced_mean, color = survey_unit), stat="smooth", method = "lm") +
  scale_color_manual(values = c(pal_37)) +
  geom_smooth(aes(x = yearly_mean_bypoint_SD, y = bray_curtis_dissimilarity_balanced_mean), method = "lm", se = F, color  = "black") +
  theme_classic()

ggsave(dissimilarity_BC_balanced_sst_meanSD, path  = here::here("figures","distance_decay"), filename = "dissimilarity_BC_balanced_sst_meanSD.jpg", height = 4, width = 9, unit = "in") 

dissimilarity_BC_balanced_sst_meanSD_facet <- ggplot(data = dissimilarities_temp) +
  labs(x = "Mean surface temperature SD (˚C)",  y = "Bray Curtis balanced dissimilarity") +
  geom_point(aes(x = yearly_mean_bypoint_SD, y = bray_curtis_dissimilarity_balanced_mean), alpha = 0.3) +
  geom_smooth(aes(x = yearly_mean_bypoint_SD, y = bray_curtis_dissimilarity_balanced_mean), method = "lm") +
  facet_wrap(~survey_unit, scales = "free") +
  theme_classic()

ggsave(dissimilarity_BC_balanced_sst_meanSD_facet, path  = here::here("figures","distance_decay"), filename = "dissimilarity_BC_balanced_sst_meanSD_facet.jpg", height = 10, width = 11, unit = "in")

dissimilarity_BC_balanced_sst_meanSD_mod <- lmer(bray_curtis_dissimilarity_balanced_mean ~ yearly_mean_bypoint_SD + (1|survey_unit), data = dissimilarities_temp)

summary(dissimilarity_BC_balanced_sst_meanSD_mod) #insignificant relationship
```

####Min temperature SD
```{r min sst SD}
#for all regions
dissimilarity_BC_balanced_sst_minSD <- ggplot(data = dissimilarities_temp) +
  labs(x = "Min surface temperature SD (˚C)",  y = "Bray Curtis balanced dissimilarity") +
  geom_point(aes(x = yearly_min_bypoint_SD, y = bray_curtis_dissimilarity_balanced_mean, color = survey_unit), alpha = 0.3) +
  geom_line(aes(x = yearly_min_bypoint_SD, y = bray_curtis_dissimilarity_balanced_mean, color = survey_unit), stat="smooth", method = "lm") +
  scale_color_manual(values = c(pal_37)) +
  geom_smooth(aes(x = yearly_min_bypoint_SD, y = bray_curtis_dissimilarity_balanced_mean), method = "lm", se = F, color  = "black") +
  theme_classic()

ggsave(dissimilarity_BC_balanced_sst_minSD, path  = here::here("figures","distance_decay"), filename = "dissimilarity_BC_balanced_sst_minSD.jpg", height = 4, width = 9, unit = "in") 

#faceted
dissimilarity_BC_balanced_sst_minSD_facet <- ggplot(data = dissimilarities_temp) +
  labs(x = "Min surface temperature SD (˚C)",  y = "Bray Curtis balanced dissimilarity") +
  geom_point(aes(x = yearly_min_bypoint_SD, y = bray_curtis_dissimilarity_balanced_mean), alpha = 0.3) +
  geom_smooth(aes(x = yearly_min_bypoint_SD, y = bray_curtis_dissimilarity_balanced_mean), method = "lm") +
  facet_wrap(~survey_unit, scales = "free") +
  theme_classic()

ggsave(dissimilarity_BC_balanced_sst_minSD_facet, path  = here::here("figures","distance_decay"), filename = "dissimilarity_BC_balanced_sst_minSD_facet.jpg", height = 10, width = 11, unit = "in")


dissimilarity_BC_balanced_sst_minSD_mod <- lmer(bray_curtis_dissimilarity_balanced_mean ~ yearly_min_bypoint_SD + (1|survey_unit), data = dissimilarities_temp)

summary(dissimilarity_BC_balanced_sst_minSD_mod) #insignificant relationship
```

####Max sst SD
```{r max sst SD}
#for all regions
dissimilarity_BC_balanced_sst_maxSD <- ggplot(data = dissimilarities_temp) +
  labs(x = "Max surface temperature SD (˚C)",  y = "Bray Curtis balanced dissimilarity") +
  geom_point(aes(x = yearly_max_bypoint_SD, y = bray_curtis_dissimilarity_balanced_mean, color = survey_unit), alpha = 0.3) +
  geom_line(aes(x = yearly_max_bypoint_SD, y = bray_curtis_dissimilarity_balanced_mean, color = survey_unit), stat="smooth", method = "lm") +
  scale_color_manual(values = c(pal_37)) +
  geom_smooth(aes(x = yearly_max_bypoint_SD, y = bray_curtis_dissimilarity_balanced_mean), method = "lm", se = F, color  = "black") +
  theme_classic()

ggsave(dissimilarity_BC_balanced_sst_maxSD, path  = here::here("figures","distance_decay"), filename = "dissimilarity_BC_balanced_sst_maxSD.jpg", height = 4, width = 9, unit = "in") 

#faceted
dissimilarity_BC_balanced_sst_maxSD_facet <- ggplot(data = dissimilarities_temp) +
  labs(x = "Max surface temperature SD (˚C)",  y = "Bray Curtis balanced dissimilarity") +
  geom_point(aes(x = yearly_max_bypoint_SD, y = bray_curtis_dissimilarity_balanced_mean), alpha = 0.3) +
  geom_smooth(aes(x = yearly_max_bypoint_SD, y = bray_curtis_dissimilarity_balanced_mean), method = "lm") +
  facet_wrap(~survey_unit, scales = "free") +
  theme_classic()

ggsave(dissimilarity_BC_balanced_sst_maxSD_facet, path  = here::here("figures","distance_decay"), filename = "dissimilarity_BC_balanced_sst_maxSD_facet.jpg", height = 10, width = 11, unit = "in")

dissimilarity_BC_balanced_sst_maxSD_mod <- lmer(bray_curtis_dissimilarity_balanced_mean ~ yearly_max_bypoint_SD + (1|survey_unit), data = dissimilarities_temp)

summary(dissimilarity_BC_balanced_sst_maxSD_mod) #significant relationship (0.002)

```

####Seas sst SD
```{r seas sst SD}
#for all regions
dissimilarity_BC_balanced_sst_seasSD <- ggplot(data = dissimilarities_temp) +
  labs(x = "Seasonality of surface temperature SD (˚C)",  y = "Bray Curtis balanced dissimilarity") +
  geom_point(aes(x = yearly_seas_bypoint_SD, y = bray_curtis_dissimilarity_balanced_mean, color = survey_unit), alpha = 0.3) +
  geom_line(aes(x = yearly_seas_bypoint_SD, y = bray_curtis_dissimilarity_balanced_mean, color = survey_unit), stat="smooth", method = "lm") +
  scale_color_manual(values = c(pal_37)) +
  geom_smooth(aes(x = yearly_seas_bypoint_SD, y = bray_curtis_dissimilarity_balanced_mean), method = "lm", se = F, color  = "black") +
  theme_classic()

ggsave(dissimilarity_BC_balanced_sst_seasSD, path  = here::here("figures","distance_decay"), filename = "dissimilarity_BC_balanced_sst_seasSD.jpg", height = 4, width = 9, unit = "in") 

#faceted
dissimilarity_BC_balanced_sst_seasSD_facet <- ggplot(data = dissimilarities_temp) +
  labs(x = "Seasonality of surface temperature SD (˚C)",  y = "Bray Curtis balanced dissimilarity") +
  geom_point(aes(x = yearly_seas_bypoint_SD, y = bray_curtis_dissimilarity_balanced_mean), alpha = 0.3) +
  geom_smooth(aes(x = yearly_seas_bypoint_SD, y = bray_curtis_dissimilarity_balanced_mean), method = "lm") +
  facet_wrap(~survey_unit, scales = "free") +
  theme_classic()

ggsave(dissimilarity_BC_balanced_sst_seasSD_facet, path  = here::here("figures","distance_decay"), filename = "dissimilarity_BC_balanced_sst_seasSD_facet.jpg", height = 10, width = 11, unit = "in")

dissimilarity_BC_balanced_sst_seasSD_mod <- lmer(bray_curtis_dissimilarity_balanced_mean ~ yearly_seas_bypoint_SD + (1|survey_unit), data = dissimilarities_temp)

summary(dissimilarity_BC_balanced_sst_seasSD_mod) #insignificant relationship

```


###Raw temp values: Comparing temperature predictors for raw
```{r compare raw temp predictors}

raw_mods_AICc <- AICc(dissimilarity_BC_balanced_sst_maxSD_mod , dissimilarity_BC_balanced_sst_meanSD_mod,
dissimilarity_BC_balanced_sst_max_mod  ,  dissimilarity_BC_balanced_sst_minSD_mod, 
dissimilarity_BC_balanced_sst_seas_mod ,  dissimilarity_BC_balanced_sst_seasSD_mod,
dissimilarity_BC_balanced_sst_min_mod  ,  dissimilarity_BC_balanced_sst_mean_mod) #compare all_mods

#best mods: 
r.squaredGLMM(dissimilarity_BC_balanced_sst_maxSD_mod) #best (delta AICc = 6.5), p = 0.002
summary(dissimilarity_BC_balanced_sst_maxSD_mod)
r.squaredGLMM(dissimilarity_BC_balanced_sst_seasSD_mod) # p = 0.002
summary(dissimilarity_BC_balanced_sst_seasSD_mod)
r.squaredGLMM(dissimilarity_BC_balanced_sst_minSD_mod)# p = 0.04
summary(dissimilarity_BC_balanced_sst_minSD_mod)
r.squaredGLMM(dissimilarity_BC_balanced_sst_max_mod) #p = 0.04
summary(dissimilarity_BC_balanced_sst_max_mod)
#for all temp mods, almost nothing is described by temp (best model, 3% of variation), most described by survey unit (90%)
```
The raw spatial variability in max temperature in the 12 months before a survey is the best predictor.

As the raw spatial variability in max temp 12 months before a survey increases, the dissimilarity also increases. Ecosystems with higher spatial variability in temp are more heterogenous in community structure.

---
##SCALED
###Avg of grid cells
####Mean temperature scaled
```{r mean sst scaled}
#for all regions
dissimilarity_BC_balanced_sst_mean_scaled <- ggplot(data = dissimilarities_temp) +
  labs(x = "Mean surface temperature scaled (˚C)",  y = "Bray Curtis balanced dissimilarity") +
  geom_point(aes(x = yearly_mean_bypoint_avg.s, y = bray_curtis_dissimilarity_balanced_mean, color = survey_unit), alpha = 0.3) +
  geom_line(aes(x = yearly_mean_bypoint_avg.s, y = bray_curtis_dissimilarity_balanced_mean, color = survey_unit), stat="smooth", method = "lm") +
  scale_color_manual(values = c(pal_37)) +
  geom_smooth(aes(x = yearly_mean_bypoint_avg.s, y = bray_curtis_dissimilarity_balanced_mean), method = "lm", se = F, color  = "black") +
  theme_classic()

ggsave(dissimilarity_BC_balanced_sst_mean_scaled, path  = here::here("figures","distance_decay"), filename = "dissimilarity_BC_balanced_sst_mean_scaled.jpg", height = 4, width = 9, unit = "in") 

dissimilarity_BC_balanced_sst_mean_scaled_mod <- lmer(bray_curtis_dissimilarity_balanced_mean ~ yearly_mean_bypoint_avg.s + (1|survey_unit), data = dissimilarities_temp)

summary(dissimilarity_BC_balanced_sst_mean_scaled_mod) #insignificant relationship
```

####Min temperature scaled
```{r min sst scaled}
#for all regions
dissimilarity_BC_balanced_sst_min_scaled <- ggplot(data = dissimilarities_temp) +
  labs(x = "Min surface temperature scaled (˚C)",  y = "Bray Curtis balanced dissimilarity") +
  geom_point(aes(x = yearly_min_bypoint_avg.s, y = bray_curtis_dissimilarity_balanced_mean, color = survey_unit), alpha = 0.3) +
  geom_line(aes(x = yearly_min_bypoint_avg.s, y = bray_curtis_dissimilarity_balanced_mean, color = survey_unit), stat="smooth", method = "lm") +
  scale_color_manual(values = c(pal_37)) +
  geom_smooth(aes(x = yearly_min_bypoint_avg.s, y = bray_curtis_dissimilarity_balanced_mean), method = "lm", se = F, color  = "black") +
  theme_classic()

ggsave(dissimilarity_BC_balanced_sst_min_scaled, path  = here::here("figures","distance_decay"), filename = "dissimilarity_BC_balanced_sst_min_scaled.jpg", height = 4, width = 9, unit = "in") 

dissimilarity_BC_balanced_sst_min_scaled_mod <- lmer(bray_curtis_dissimilarity_balanced_mean ~ yearly_min_bypoint_avg.s + (1|survey_unit), data = dissimilarities_temp)

summary(dissimilarity_BC_balanced_sst_min_scaled_mod) #nearly significant relationship (0.09)
```

####Max sst scaled
```{r max sst scaled}
#for all regions
dissimilarity_BC_balanced_sst_max_scaled <- ggplot(data = dissimilarities_temp) +
  labs(x = "Max surface temperature scaled (˚C)",  y = "Bray Curtis balanced dissimilarity") +
  geom_point(aes(x = yearly_max_bypoint_avg.s, y = bray_curtis_dissimilarity_balanced_mean, color = survey_unit), alpha = 0.3) +
  geom_line(aes(x = yearly_max_bypoint_avg.s, y = bray_curtis_dissimilarity_balanced_mean, color = survey_unit), stat="smooth", method = "lm") +
  scale_color_manual(values = c(pal_37)) +
  geom_smooth(aes(x = yearly_max_bypoint_avg.s, y = bray_curtis_dissimilarity_balanced_mean), method = "lm", se = F, color  = "black") +
  theme_classic()

ggsave(dissimilarity_BC_balanced_sst_max_scaled, path  = here::here("figures","distance_decay"), filename = "dissimilarity_BC_balanced_sst_max_scaled.jpg", height = 4, width = 9, unit = "in") 

dissimilarity_BC_balanced_sst_max_scaled_mod <- lmer(bray_curtis_dissimilarity_balanced_mean ~ yearly_max_bypoint_avg.s + (1|survey_unit), data = dissimilarities_temp)

summary(dissimilarity_BC_balanced_sst_max_scaled_mod) #insignificant relationship

```

####Seas sst scaled
```{r seas sst scaled}
#for all regions
dissimilarity_BC_balanced_sst_seas_scaled <- ggplot(data = dissimilarities_temp) +
  labs(x = "Seasonality of surface temperature scaled (˚C)",  y = "Bray Curtis balanced dissimilarity") +
  geom_point(aes(x = yearly_seas_bypoint_avg.s, y = bray_curtis_dissimilarity_balanced_mean, color = survey_unit), alpha = 0.3) +
  geom_line(aes(x = yearly_seas_bypoint_avg.s, y = bray_curtis_dissimilarity_balanced_mean, color = survey_unit), stat="smooth", method = "lm") +
  scale_color_manual(values = c(pal_37)) +
  geom_smooth(aes(x = yearly_seas_bypoint_avg.s, y = bray_curtis_dissimilarity_balanced_mean), method = "lm", se = F, color  = "black") +
  theme_classic()

ggsave(dissimilarity_BC_balanced_sst_seas_scaled, path  = here::here("figures","distance_decay"), filename = "dissimilarity_BC_balanced_sst_seas_scaled.jpg", height = 4, width = 9, unit = "in") 

dissimilarity_BC_balanced_sst_seas_scaled_mod <- lmer(bray_curtis_dissimilarity_balanced_mean ~ yearly_seas_bypoint_avg.s + (1|survey_unit), data = dissimilarities_temp)

summary(dissimilarity_BC_balanced_sst_seas_scaled_mod) #insignificant relationship

```
###Standard deviation of grid cells
####Mean temperature SD scaled
```{r mean sst SD scaled}
#for all regions
dissimilarity_BC_balanced_sst_meanSD_scaled <- ggplot(data = dissimilarities_temp) +
  labs(x = "Mean surface temperature SD scaled (˚C)",  y = "Bray Curtis balanced dissimilarity") +
  geom_point(aes(x = yearly_mean_bypoint_SD.s, y = bray_curtis_dissimilarity_balanced_mean, color = survey_unit), alpha = 0.3) +
  geom_line(aes(x = yearly_mean_bypoint_SD.s, y = bray_curtis_dissimilarity_balanced_mean, color = survey_unit), stat="smooth", method = "lm") +
  scale_color_manual(values = c(pal_37)) +
  geom_smooth(aes(x = yearly_mean_bypoint_SD.s, y = bray_curtis_dissimilarity_balanced_mean), method = "lm", se = F, color  = "black") +
  theme_classic()

ggsave(dissimilarity_BC_balanced_sst_meanSD_scaled, path  = here::here("figures","distance_decay"), filename = "dissimilarity_BC_balanced_sst_meanSD_scaled.jpg", height = 4, width = 9, unit = "in") 

dissimilarity_BC_balanced_sst_meanSD_scaled_mod <- lmer(bray_curtis_dissimilarity_balanced_mean ~ yearly_mean_bypoint_SD.s + (1|survey_unit), data = dissimilarities_temp)

summary(dissimilarity_BC_balanced_sst_meanSD_scaled_mod) #insignificant relationship
```

####Min temperature SD scaled
```{r min sst SD scaled}
#for all regions
dissimilarity_BC_balanced_sst_minSD_scaled <- ggplot(data = dissimilarities_temp) +
  labs(x = "Min surface temperature SD scaled (˚C)",  y = "Bray Curtis balanced dissimilarity") +
  geom_point(aes(x = yearly_min_bypoint_SD.s, y = bray_curtis_dissimilarity_balanced_mean, color = survey_unit), alpha = 0.3) +
  geom_line(aes(x = yearly_min_bypoint_SD.s, y = bray_curtis_dissimilarity_balanced_mean, color = survey_unit), stat="smooth", method = "lm") +
  scale_color_manual(values = c(pal_37)) +
  geom_smooth(aes(x = yearly_min_bypoint_SD.s, y = bray_curtis_dissimilarity_balanced_mean), method = "lm", se = F, color  = "black") +
  theme_classic()

ggsave(dissimilarity_BC_balanced_sst_minSD_scaled, path  = here::here("figures","distance_decay"), filename = "dissimilarity_BC_balanced_sst_minSD_scaled.jpg", height = 4, width = 9, unit = "in") 

dissimilarity_BC_balanced_sst_minSD_scaled_mod <- lmer(bray_curtis_dissimilarity_balanced_mean ~ yearly_min_bypoint_SD.s + (1|survey_unit), data = dissimilarities_temp)

summary(dissimilarity_BC_balanced_sst_minSD_scaled_mod) #insignificant relationship
```

####Max sst SD scaled
```{r max sst SD scaled}
#for all regions
dissimilarity_BC_balanced_sst_maxSD_scaled <- ggplot(data = dissimilarities_temp) +
  labs(x = "Max surface temperature SD scaled (˚C)",  y = "Bray Curtis balanced dissimilarity") +
  geom_point(aes(x = yearly_max_bypoint_SD.s, y = bray_curtis_dissimilarity_balanced_mean, color = survey_unit), alpha = 0.3) +
  geom_line(aes(x = yearly_max_bypoint_SD.s, y = bray_curtis_dissimilarity_balanced_mean, color = survey_unit), stat="smooth", method = "lm") +
  scale_color_manual(values = c(pal_37)) +
  geom_smooth(aes(x = yearly_max_bypoint_SD.s, y = bray_curtis_dissimilarity_balanced_mean), method = "lm", se = F, color  = "black") +
  theme_classic()

ggsave(dissimilarity_BC_balanced_sst_maxSD_scaled, path  = here::here("figures","distance_decay"), filename = "dissimilarity_BC_balanced_sst_maxSD_scaled.jpg", height = 4, width = 9, unit = "in") 

dissimilarity_BC_balanced_sst_maxSD_scaled_mod <- lmer(bray_curtis_dissimilarity_balanced_mean ~ yearly_max_bypoint_SD.s + (1|survey_unit), data = dissimilarities_temp)

summary(dissimilarity_BC_balanced_sst_maxSD_scaled_mod) #nearly significant relationship (0.06)

```

####Seas sst SD scaled
```{r seas sst SD scaled}
#for all regions
dissimilarity_BC_balanced_sst_seasSD_scaled <- ggplot(data = dissimilarities_temp) +
  labs(x = "Seasonality of surface temperature SD scaled (˚C)",  y = "Bray Curtis balanced dissimilarity") +
  geom_point(aes(x = yearly_seas_bypoint_SD.s, y = bray_curtis_dissimilarity_balanced_mean, color = survey_unit), alpha = 0.3) +
  geom_line(aes(x = yearly_seas_bypoint_SD.s, y = bray_curtis_dissimilarity_balanced_mean, color = survey_unit), stat="smooth", method = "lm") +
  scale_color_manual(values = c(pal_37)) +
  geom_smooth(aes(x = yearly_seas_bypoint_SD.s, y = bray_curtis_dissimilarity_balanced_mean), method = "lm", se = F, color  = "black") +
  theme_classic()

ggsave(dissimilarity_BC_balanced_sst_seasSD_scaled, path  = here::here("figures","distance_decay"), filename = "dissimilarity_BC_balanced_sst_seasSD_scaled.jpg", height = 4, width = 9, unit = "in") 

dissimilarity_BC_balanced_sst_seasSD_scaled_mod <- lmer(bray_curtis_dissimilarity_balanced_mean ~ yearly_seas_bypoint_SD.s + (1|survey_unit), data = dissimilarities_temp)

summary(dissimilarity_BC_balanced_sst_seasSD_scaled_mod) #insignificant relationship

```



###Comparing scaled temperature predictors
```{r compare scaled temp predictors}

#list of models
list_scaled_mods <- grep("*_scaled_mod",names(.GlobalEnv),value=TRUE)

scaled_mods_AICc <- AICc(dissimilarity_BC_balanced_sst_maxSD_scaled_mod , dissimilarity_BC_balanced_sst_meanSD_scaled_mod,
dissimilarity_BC_balanced_sst_max_scaled_mod  ,  dissimilarity_BC_balanced_sst_minSD_scaled_mod, 
dissimilarity_BC_balanced_sst_seas_scaled_mod ,  dissimilarity_BC_balanced_sst_seasSD_scaled_mod,
dissimilarity_BC_balanced_sst_min_scaled_mod  ,  dissimilarity_BC_balanced_sst_mean_scaled_mod) #compare all scaled_mods

#best mods (all within 2 AICc): 
r.squaredGLMM(dissimilarity_BC_balanced_sst_maxSD_scaled_mod)
summary(dissimilarity_BC_balanced_sst_maxSD_scaled_mod) #p = 0.06
r.squaredGLMM(dissimilarity_BC_balanced_sst_min_scaled_mod)
summary(dissimilarity_BC_balanced_sst_min_scaled_mod)# 0.09
r.squaredGLMM(dissimilarity_BC_balanced_sst_seas_scaled_mod)
summary(dissimilarity_BC_balanced_sst_seas_scaled_mod) # p = 0.168
r.squaredGLMM(dissimilarity_BC_balanced_sst_mean_scaled_mod)
summary(dissimilarity_BC_balanced_sst_mean_scaled_mod) # p = 0.262



#for all temp mods, almost nothing is described by scaled temp in region (0.05% of variation), most described by survey unit (90% of variation)
```
The relative spatial variability (standard deviation) in max temperature in the 12 months before a survey is the best predictor of dissimilarity, but dissimilarity_BC_balanced_sst_min_scaled_mod, dissimilarity_BC_balanced_sst_seas_scaled_mod, and dissimilarity_BC_balanced_sst_mean_scaled_mod also perform well.

As SD of max temp increases, dissimilarity increases (more heterogeneity in temp = more heterogeneity in community structure)

As min temp increases, dissimilarity decreases. (Warmer min temps = more homogenous community structure)

As seasonality increases, dissimilarity increases. (Larger temp swings in a year = more heterogenous community structure) #this parallels the relationship between min temp and dissimilarity

####What to carry forward:

For raw temp (indicative of survey vs. survey):
-dissimilarity_BC_balanced_sst_maxSD_mod (Temp heterogeneity)
-dissimilarity_BC_balanced_sst_max_mod (Maximum temperature)

For scaled temp (indicative of within survey: cool or warm year? heterogenous or homogenous year)
-dissimilarity_BC_balanced_sst_maxSD_scaled_mod
-dissimilarity_BC_balanced_sst_min_scaled_mod