---
title: "Annual Total BC Dissimilarity vs. SODA Temperature for all Survey Units"
output: html_notebook
---

This code is Script 7a for Kitchel et al. Biotic homogenization, the exception and not the rule for marine fish communities manuscript.

- This project is a collaborative effort to describe changes in taxonomic composition  of fish communities around the world--as sampled by bottom trawl surveys.

- Code by Zoë J. Kitchel

SESSION INFO TO DO

##Here, we use bottom temperature to predict annual dissimilarity for all survey units

```{r setup}

library(data.table)
library(ggplot2)
library(lme4)
library(lmerTest)
library(MuMIn)


SODA_data_temp_avgs_full <- readRDS(here::here("data","Temperature","SODA_data_temp_avgs_full.rds"))
#don't feel confident about SODA temp before 1980 right now, start then

SODA_data_temp_avgs_full <- SODA_data_temp_avgs_full[year_for_avg >= 1980,]

SODA_data_temp_avgs_full[,year := year_for_avg]

#pull in average distance decay values
distances_dissimilarities_allyears.r <- readRDS(here::here("output","dissimilarities","distances_dissimilarities_allyears.r.rds"))

```

Model to extract p value from overall model
```{r}
#define function to extract overall p-value of model
overall_p <- function(my_model) {
    f <- summary(my_model)$fstatistic
    p <- pf(f[1],f[2],f[3],lower.tail=F)
    attributes(p) <- NULL
    return(p)
}
```


###Palette for Plotting and name helper
Palette for plotting all 42 survey units
(Prep for eventual plots)
Full Name Helper

```{r}

name_helper <- data.table(Survey_Name_Season = c("Aleutian Islands",
                                    "Baltic Sea Q1",
                                    "Baltic Sea Q4",
                                    "Chile",
                                    "Newfoundland",
                                    "Queen Charlotte Sound",
                                    "Eastern Bering Sea",
                                    "Bay of Biscay",
                                    "English Channel",
                                    "Gulf of Mexico Summer",
                                    "Gulf of Alaska",
                                    "Greenland",
                                    "N Gulf of St. Lawrence",
                                    "S Gulf of St. Lawrence",
                                    "Iceland",
                                    "Irish Sea",
                                    "Mediterranean",
                                    "Namibia",
                                    "NE US Fall",
                                    "NE US Spring",
                                    "N Ireland Q1",
                                    "N Ireland Q4",
                                    "Barents Sea Norway Q3",
                                    "N Sea Q1",
                                    "N Sea Q3",
                                    "Chatham Rise NZ",
                                    "E Coast S Island NZ",
                                    "W Coast S Island NZ",
                                    "Portugal",
                                    "S Georgia",
                                  "Scotian Shelf Summer",
                                  "SE US Fall",
                                  "SE US Spring",
                                  "SE US Summer",
                                  "W Coast US",
                                  "Atlantic Ocean ZA",
                                  "Indian Ocean ZA",
                                   "Rockall Plateau",
                                  "Scotland Shelf Sea Q1",
                                  "Scotland Shelf Sea Q4",
                                  "Falkland Islands",
                                  "Gulf of Mexico Fall",
                                  "Barents Sea Norway Q1",
                                  "Sub-Arctic NZ",
                                  "Scotian Shelf Spring"),
                          survey_unit = c(
                                  "AI",        
                                  "BITS-1",    
                                  "BITS-4",    
                                  "CHL",       
                                  "DFO-NF",    
                                  "DFO-QCS",   
                                  "EBS",       
                                  "EVHOE",     
                                  "FR-CGFS",   
                                  "GMEX-Summer",
                                  "GOA",       
                                  "GRL-DE",    
                                  "GSL-N",     
                                  "GSL-S",     
                                  "ICE-GFS",   
                                  "IE-IGFS",   
                                  "MEDITS",    
                                  "NAM",       
                                  "NEUS-Fall", 
                                  "NEUS-Spring",
                                  "NIGFS-1",   
                                  "NIGFS-4",   
                                  "Nor-BTS-3", 
                                  "NS-IBTS-1", 
                                  "NS-IBTS-3", 
                                  "NZ-CHAT",   
                                  "NZ-ECSI",   
                                  "NZ-WCSI",   
                                  "PT-IBTS",   
                                  "S-GEORG",   
                                  "SCS-SUMMER",
                                  "SEUS-fall", 
                                  "SEUS-spring",
                                  "SEUS-summer",
                                  "WCANN",     
                                  "ZAF-ATL",   
                                  "ZAF-IND",
                                  "ROCKALL",
                                  "SWC-IBTS-1",
                                  "SWC-IBTS-4",
                                  "FALK",
                                  "GMEX-Fall",
                                  "Nor-BTS-1",
                                  "NZ-SUBA",
                                  "SCS-SPRING"
                          ))



```

Color link
```{r}
survey_unit.list <- levels(factor(distances_dissimilarities_allyears.r[,survey_unit]))

palette_42 <- c(
  "#5A5156", #AI
  "#DF00DB", #BITS-1
  "#DB8EDA", #BITS-4
  "#F6222E", #CHL
  "#F8A19F", #DFO-NF
  "#16FF32", #DFO-QCS
  "#325A9B", #EBS
  "#3283FE", #EVHOE
  "#FEAF16", #FR-CGFS
  "#fccb6d", #GMEX-Fall
  "#1C8356", #GMEX-Summer
  "#C4451C", #GOA
  "#85660D", #GRL-DE
  "#B0009F", #GSL-N
  "#BF79B8", #GSL-S
  "#1CBE4F", #ICE-GFS
  "#782AB6", #IE-IGFS
  "#90AD1C", #MEDITS
  "#6B003A", #NAM
  "#A75B00", #NEUS-Fall
  "#E3B072", #NEUS-Spring
  "#02E8B6", #NIGFS-1
  "#97E7D5", #NIGFS-4
  "#B00068", #Nor-BTS-3
  "#00B9E3", #NS-IBTS-1
  "#95E2F4", #NS-IBTS-3
  "#B3CE73", #NZ-CHAT
  "#689500", #NZ-ECSI
  "#364d02",#NZ-SUBA
  "#AAF400", #NZ-WCSI
  "#AA0DFE", #PT-IBTS
  "#7f9eb8", #ROCKALL
  "#FA0087", #S-GEORG
  "#DEA0FD", #SCS-Summer
  "#FCEF88", #SEUS-fall
  "#A59405", #SEUS-spring
  "#FCE100", #SEUS-summer
  "#544563", #SWC-IBTS-1
  "#a37fc7", #SWC-IBTS-4
  "#C075A6", #WCANN
  "#BDCDFF", #ZAF-ATL
  "#003EFF"  #ZAF-IND
)

color_link <- data.table(survey_unit = survey_unit.list,hex = palette_42)

color_link <- name_helper[color_link, on = "survey_unit", nomatch = NULL]
```


Put into one data table
```{r link temp to avg values}
dissimilarities_temp <- SODA_data_temp_avgs_full[distances_dissimilarities_allyears.r, on = c("year","survey_unit")]

#few empty temp rows for early and late years (only 22, delete)
dissimilarities_temp <- na.omit(dissimilarities_temp)

#get rid of any years not between 1981-2019
dissimilarities_temp <- dissimilarities_temp[year >= 1981 & year <=2019]
```

Pull in dissimilarity ~ model coefficients
```{r}
BC_total_model_coefs_reduced_length.unique <- readRDS(here::here("output","region_stats","BC_total_model_coefs_reduced_length.unique.Rds"))
```



Table that will be populated with model results
```{r}
#depth, scaled, temp var, coefficient, p-value, r^2, AICc
total_dissimilarity_sbt_model_results <- data.table(Depth = as.character(),
                                                       Scaled = as.character(),
                                                       `Temp variable` = as.character(),
                                                       Intercept = as.numeric(),
                                                       `Temp coefficient` = as.numeric(),
                                                       `Temp p-value` = as.numeric(),
                                                       `R-squared` = as.numeric(),
                                                       AICc = as.numeric())
                                                  
```


How does this trend hold up for all regions?
###UNSCALED

####Mean temperature
```{r mean sbt}
#for all regions
(dissimilarity_BC_total_sbt_mean <- ggplot(data = dissimilarities_temp) +
  labs(x = "Mean bottom temperature (˚C)",  y = "Bray Curtis total dissimilarity") +
  geom_point(aes(x = yearly_mean_bypoint_avg, y = bray_curtis_dissimilarity_total_mean, color = survey_unit), alpha = 0.3) +
  geom_line(aes(x = yearly_mean_bypoint_avg, y = bray_curtis_dissimilarity_total_mean, color = survey_unit), stat="smooth", method = "lm") +
  scale_color_manual(values = c(palette_42)) +
  geom_smooth(aes(x = yearly_mean_bypoint_avg, y = bray_curtis_dissimilarity_total_mean), method = "lm", se = F, color  = "black") +
  theme_classic())

ggsave(dissimilarity_BC_total_sbt_mean, path  = here::here("figures","dissimilarities"), filename = "dissimilarity_BC_total_sbt_mean.jpg", height = 4, width = 9, unit = "in") 

dissimilarity_BC_total_sbt_mean_facet <- ggplot(data = dissimilarities_temp) +
  labs(x = "Mean bottom temperature (˚C)",  y = "Bray Curtis total dissimilarity") +
  geom_point(aes(x = yearly_mean_bypoint_avg, y = bray_curtis_dissimilarity_total_mean), alpha = 0.3) +
  geom_smooth(aes(x = yearly_mean_bypoint_avg, y = bray_curtis_dissimilarity_total_mean), method = "lm") +
  facet_wrap(~survey_unit, scales = "free") +
  theme_classic()

ggsave(dissimilarity_BC_total_sbt_mean_facet, path  = here::here("figures","dissimilarities"), filename = "dissimilarity_BC_total_sbt_mean_facet.jpg", height = 10, width = 11, unit = "in")

#TEMP ONLY
dissimilarity_BC_total_sbt_mean_mod <- lm(bray_curtis_dissimilarity_total_mean ~ yearly_mean_bypoint_avg, data = dissimilarities_temp) 

summary(dissimilarity_BC_total_sbt_mean_mod) #insignificant relationship

confint <- confint(dissimilarity_BC_total_sbt_mean_mod) #confidence interval

coefs <- data.frame(coef(summary(dissimilarity_BC_total_sbt_mean_mod))) #table with coefficients, #Satterthwaite's degrees of freedom, and p-value approximation for fixed effects

#input results into table
total_dissimilarity_sbt_model_results_row <- data.table(matrix(c("sbt", #depth
                                                 F, #scaled
                                                 "yearly_mean_bypoint_avg", #model structure
                                                 paste0(round(coef(dissimilarity_BC_total_sbt_mean_mod)[[1]],3),
                                                        " ± ",
                                                        round((confint[1,2]-confint[1,1])/2,3)), #intercept
                                                 paste0(round(coef(dissimilarity_BC_total_sbt_mean_mod)[[2]],3),
                                                        " ± ",
                                                        round((confint[2,2]-confint[2,1])/2,3)), #coef
                                                 round(coefs[2,4],3), #pvalue
                                                 round(summary(dissimilarity_BC_total_sbt_mean_mod)$r.squared,3),#rsquared
                                                 round(AICc(dissimilarity_BC_total_sbt_mean_mod),3)
                                                 ),nrow = 1))

#merge with larger table
total_dissimilarity_sbt_model_results <- rbind(total_dissimilarity_sbt_model_results,total_dissimilarity_sbt_model_results_row, use.names = F)

#########################
#MEAN TEMP PLUS SURVEY UNIT
#########################
dissimilarity_BC_total_sbt_mean_plus_survey_mod <- lm(bray_curtis_dissimilarity_total_mean ~ yearly_mean_bypoint_avg + survey_unit, data = dissimilarities_temp) #this converges (allows both intercept and slope to vary with survey_unit)

summary(dissimilarity_BC_total_sbt_mean_plus_survey_mod) #insignificant relationship

confint <- confint(dissimilarity_BC_total_sbt_mean_plus_survey_mod) #confidence interval

coefs <- data.frame(coef(summary(dissimilarity_BC_total_sbt_mean_plus_survey_mod))) #table with coefficients, #Satterthwaite's degrees of freedom, and p-value approximation for fixed effects

#input results into table
total_dissimilarity_sbt_model_results_row <- data.table(matrix(c("sbt", #depth
                                                 F, #scaled
                                                 "yearly_mean_bypoint_avg + survey_unit", #model structure
                                                 paste0(round(coef(dissimilarity_BC_total_sbt_mean_plus_survey_mod)[[1]],3),
                                                        " ± ",
                                                        round((confint[1,2]-confint[1,1])/2,3)), #intercept
                                                 paste0(round(coef(dissimilarity_BC_total_sbt_mean_plus_survey_mod)[[2]],3),
                                                        " ± ",
                                                        round((confint[2,2]-confint[2,1])/2,3)), #coef
                                                 signif(coefs[2,4],3), #pvalue
                                                 round(summary(dissimilarity_BC_total_sbt_mean_plus_survey_mod)$r.squared,3),#rsquared
                                                 round(AICc(dissimilarity_BC_total_sbt_mean_plus_survey_mod),3)
                                                 ),nrow = 1))

#merge with larger table
total_dissimilarity_sbt_model_results <- rbind(total_dissimilarity_sbt_model_results,total_dissimilarity_sbt_model_results_row, use.names = F)

#########################
#MEAN TEMP * SURVEY UNIT
#########################
dissimilarity_BC_total_sbt_mean_times_survey_mod <- lm(bray_curtis_dissimilarity_total_mean ~ yearly_mean_bypoint_avg * survey_unit, data = dissimilarities_temp) #this converges (allows both intercept and slope to vary with survey_unit)

summary(dissimilarity_BC_total_sbt_mean_times_survey_mod) #insignificant relationship

confint <- confint(dissimilarity_BC_total_sbt_mean_times_survey_mod) #confidence interval

coefs <- data.frame(coef(summary(dissimilarity_BC_total_sbt_mean_times_survey_mod))) #table with coefficients, #Satterthwaite's degrees of freedom, and p-value approximation for fixed effects

#input results into table
total_dissimilarity_sbt_model_results_row <- data.table(matrix(c("sbt", #depth
                                                 F, #scaled
                                                 "yearly_mean_bypoint_avg * survey_unit", #model structure
                                                 paste0(round(coef(dissimilarity_BC_total_sbt_mean_times_survey_mod)[[1]],3),
                                                        " ± ",
                                                        round((confint[1,2]-confint[1,1])/2,3)), #intercept
                                                 paste0(round(coef(dissimilarity_BC_total_sbt_mean_times_survey_mod)[[2]],3),
                                                        " ± ",
                                                        round((confint[2,2]-confint[2,1])/2,3)), #coef
                                                 signif(coefs[2,4],3), #pvalue
                                                 round(summary(dissimilarity_BC_total_sbt_mean_times_survey_mod)$r.squared,3),#rsquared
                                                 round(AICc(dissimilarity_BC_total_sbt_mean_times_survey_mod),3)
                                                 ),nrow = 1))

#merge with larger table
total_dissimilarity_sbt_model_results <- rbind(total_dissimilarity_sbt_model_results,total_dissimilarity_sbt_model_results_row, use.names = F)

#table of individual interaction coefficients
best_sbt_mod_coefs_dt <- data.table(coefs)
best_sbt_mod_coefs_dt[,predictor:=rownames(coefs)]
fwrite(best_sbt_mod_coefs_dt, here::here("output","best_sbt_mod_coefs_dt.csv"))
```

####Min temperature
```{r min sbt}
#for all regions
(dissimilarity_BC_total_sbt_min <- ggplot(data = dissimilarities_temp) +
  labs(x = "Min bottom temperature (˚C)",  y = "Bray Curtis total dissimilarity") +
  geom_point(aes(x = yearly_min_bypoint_avg, y = bray_curtis_dissimilarity_total_mean, color = survey_unit), alpha = 0.3) +
  geom_line(aes(x = yearly_min_bypoint_avg, y = bray_curtis_dissimilarity_total_mean, color = survey_unit), stat="smooth", method = "lm") +
  scale_color_manual(values = c(palette_42)) +
  geom_smooth(aes(x = yearly_min_bypoint_avg, y = bray_curtis_dissimilarity_total_mean), method = "lm", se = F, color  = "black") +
  theme_classic())

ggsave(dissimilarity_BC_total_sbt_min, path  = here::here("figures","dissimilarities"), filename = "dissimilarity_BC_total_sbt_min.jpg", height = 4, width = 9, unit = "in") 

dissimilarity_BC_total_sbt_min_facet <- ggplot(data = dissimilarities_temp) +
  labs(x = "Min bottom temperature (˚C)",  y = "Bray Curtis total dissimilarity") +
  geom_point(aes(x = yearly_min_bypoint_avg, y = bray_curtis_dissimilarity_total_mean), alpha = 0.3) +
  geom_smooth(aes(x = yearly_min_bypoint_avg, y = bray_curtis_dissimilarity_total_mean), method = "lm") +
  facet_wrap(~survey_unit, scales = "free") +
  theme_classic()

ggsave(dissimilarity_BC_total_sbt_min_facet, path  = here::here("figures","dissimilarities"), filename = "dissimilarity_BC_total_sbt_min_facet.jpg", height = 10, width = 11, unit = "in")

dissimilarity_BC_total_sbt_min_mod <- lm(bray_curtis_dissimilarity_total_mean ~ yearly_min_bypoint_avg, data = dissimilarities_temp) #this converges (allows both intercept and slope to vary with survey_unit)

summary(dissimilarity_BC_total_sbt_min_mod) #insignificant relationship

confint <- confint(dissimilarity_BC_total_sbt_min_mod) #confidence interval

coefs <- data.frame(coef(summary(dissimilarity_BC_total_sbt_min_mod))) #table with coefficients, #Satterthwaite's degrees of freedom, and p-value approximation for fixed effects

#input results into table
total_dissimilarity_sbt_model_results_row <- data.table(matrix(c("sbt", #depth
                                                 F, #scaled
                                                 "yearly_min_bypoint_avg", #model structure
                                                 paste0(round(coef(dissimilarity_BC_total_sbt_min_mod)[[1]],3),
                                                        " ± ",
                                                        round((confint[1,2]-confint[1,1])/2,3)), #intercept
                                                 paste0(round(coef(dissimilarity_BC_total_sbt_min_mod)[[2]],3),
                                                        " ± ",
                                                        round((confint[2,2]-confint[2,1])/2,3)), #coef
                                                 signif(coefs[2,4],3), #pvalue
                                                 round(summary(dissimilarity_BC_total_sbt_min_mod)$r.squared,3),#rsquared
                                                 round(AICc(dissimilarity_BC_total_sbt_min_mod),3)
                                                 ),nrow = 1))

#merge with larger table
total_dissimilarity_sbt_model_results <- rbind(total_dissimilarity_sbt_model_results,total_dissimilarity_sbt_model_results_row, use.names = F)

#########################
#min TEMP PLUS SURVEY UNIT
#########################
dissimilarity_BC_total_sbt_min_plus_survey_mod <- lm(bray_curtis_dissimilarity_total_mean ~ yearly_min_bypoint_avg + survey_unit, data = dissimilarities_temp) #this converges (allows both intercept and slope to vary with survey_unit)

summary(dissimilarity_BC_total_sbt_min_plus_survey_mod) #insignificant relationship

confint <- confint(dissimilarity_BC_total_sbt_min_plus_survey_mod) #confidence interval

coefs <- data.frame(coef(summary(dissimilarity_BC_total_sbt_min_plus_survey_mod))) #table with coefficients, #Satterthwaite's degrees of freedom, and p-value approximation for fixed effects

#input results into table
total_dissimilarity_sbt_model_results_row <- data.table(matrix(c("sbt", #depth
                                                 F, #scaled
                                                 "yearly_min_bypoint_avg + survey_unit", #model structure
                                                 paste0(round(coef(dissimilarity_BC_total_sbt_min_plus_survey_mod)[[1]],3),
                                                        " ± ",
                                                        round((confint[1,2]-confint[1,1])/2,3)), #intercept
                                                 paste0(round(coef(dissimilarity_BC_total_sbt_min_plus_survey_mod)[[2]],3),
                                                        " ± ",
                                                        round((confint[2,2]-confint[2,1])/2,3)), #coef
                                                 signif(coefs[2,4],3), #model pvalue
                                                 round(summary(dissimilarity_BC_total_sbt_min_plus_survey_mod)$r.squared,3),#rsquared
                                                 round(AICc(dissimilarity_BC_total_sbt_min_plus_survey_mod),3)
                                                 ),nrow = 1))

#merge with larger table
total_dissimilarity_sbt_model_results <- rbind(total_dissimilarity_sbt_model_results,total_dissimilarity_sbt_model_results_row, use.names = F)

#########################
#min TEMP * SURVEY UNIT
#########################
dissimilarity_BC_total_sbt_min_times_survey_mod <- lm(bray_curtis_dissimilarity_total_mean ~ yearly_min_bypoint_avg * survey_unit, data = dissimilarities_temp) #this converges (allows both intercept and slope to vary with survey_unit)

summary(dissimilarity_BC_total_sbt_min_times_survey_mod) #insignificant relationship

confint <- confint(dissimilarity_BC_total_sbt_min_times_survey_mod) #confidence interval

coefs <- data.frame(coef(summary(dissimilarity_BC_total_sbt_min_times_survey_mod))) #table with coefficients, #Satterthwaite's degrees of freedom, and p-value approximation for fixed effects

#input results into table
total_dissimilarity_sbt_model_results_row <- data.table(matrix(c("sbt", #depth
                                                 F, #scaled
                                                 "yearly_min_bypoint_avg * survey_unit", #model structure
                                                 paste0(round(coef(dissimilarity_BC_total_sbt_min_times_survey_mod)[[1]],3),
                                                        " ± ",
                                                        round((confint[1,2]-confint[1,1])/2,3)), #intercept
                                                 paste0(round(coef(dissimilarity_BC_total_sbt_min_times_survey_mod)[[2]],3),
                                                        " ± ",
                                                        round((confint[2,2]-confint[2,1])/2,3)), #coef
                                                 signif(coefs[2,4],3), #pvalue
                                                 round(summary(dissimilarity_BC_total_sbt_min_times_survey_mod)$r.squared,3),#rsquared
                                                 round(AICc(dissimilarity_BC_total_sbt_min_times_survey_mod),3)
                                                 ),nrow = 1))

#merge with larger table
total_dissimilarity_sbt_model_results <- rbind(total_dissimilarity_sbt_model_results,total_dissimilarity_sbt_model_results_row, use.names = F)
```

####Max sbt
```{r max sbt}
#for all regions
(dissimilarity_BC_total_sbt_max <- ggplot(data = dissimilarities_temp) +
  labs(x = "Max bottom temperature (˚C)",  y = "Bray Curtis total dissimilarity") +
  geom_point(aes(x = yearly_max_bypoint_avg, y = bray_curtis_dissimilarity_total_mean, color = survey_unit), alpha = 0.3) +
  geom_line(aes(x = yearly_max_bypoint_avg, y = bray_curtis_dissimilarity_total_mean, color = survey_unit), stat="smooth", method = "lm") +
  scale_color_manual(values = c(palette_42)) +
  geom_smooth(aes(x = yearly_max_bypoint_avg, y = bray_curtis_dissimilarity_total_mean), method = "lm", se = F, color  = "black") +
  theme_classic())

ggsave(dissimilarity_BC_total_sbt_max, path  = here::here("figures","dissimilarities"), filename = "dissimilarity_BC_total_sbt_max.jpg", height = 4, width = 9, unit = "in") 

dissimilarity_BC_total_sbt_max_facet <- ggplot(data = dissimilarities_temp) +
  labs(x = "Max bottom temperature (˚C)",  y = "Bray Curtis total dissimilarity") +
  geom_point(aes(x = yearly_max_bypoint_avg, y = bray_curtis_dissimilarity_total_mean), alpha = 0.3) +
  geom_smooth(aes(x = yearly_max_bypoint_avg, y = bray_curtis_dissimilarity_total_mean), method = "lm") +
  facet_wrap(~survey_unit, scales = "free") +
  theme_classic()

ggsave(dissimilarity_BC_total_sbt_max_facet, path  = here::here("figures","dissimilarities"), filename = "dissimilarity_BC_total_sbt_max_facet.jpg", height = 10, width = 11, unit = "in")

dissimilarity_BC_total_sbt_max_mod <- lm(bray_curtis_dissimilarity_total_mean ~ yearly_max_bypoint_avg, data = dissimilarities_temp) #this converges (allows both intercept and slope to vary with survey_unit)

summary(dissimilarity_BC_total_sbt_max_mod) #insignificant relationship

confint <- confint(dissimilarity_BC_total_sbt_max_mod) #confidence interval

coefs <- data.frame(coef(summary(dissimilarity_BC_total_sbt_max_mod))) #table with coefficients, #Satterthwaite's degrees of freedom, and p-value approximation for fixed effects

#input results into table
total_dissimilarity_sbt_model_results_row <- data.table(matrix(c("sbt", #depth
                                                 F, #scaled
                                                 "yearly_max_bypoint_avg", #model structure
                                                 paste0(round(coef(dissimilarity_BC_total_sbt_max_mod)[[1]],3),
                                                        " ± ",
                                                        round((confint[1,2]-confint[1,1])/2,3)), #intercept
                                                 paste0(round(coef(dissimilarity_BC_total_sbt_max_mod)[[2]],3),
                                                        " ± ",
                                                        round((confint[2,2]-confint[2,1])/2,3)), #coef
                                                 round(coefs[2,4],3), #pvalue
                                                 round(summary(dissimilarity_BC_total_sbt_max_mod)$r.squared,3),#rsquared
                                                 round(AICc(dissimilarity_BC_total_sbt_max_mod),3)
                                                 ),nrow = 1))

#merge with larger table
total_dissimilarity_sbt_model_results <- rbind(total_dissimilarity_sbt_model_results,total_dissimilarity_sbt_model_results_row, use.names = F)

#########################
#max TEMP PLUS SURVEY UNIT
#########################
dissimilarity_BC_total_sbt_max_plus_survey_mod <- lm(bray_curtis_dissimilarity_total_mean ~ yearly_max_bypoint_avg + survey_unit, data = dissimilarities_temp) #this converges (allows both intercept and slope to vary with survey_unit)

summary(dissimilarity_BC_total_sbt_max_plus_survey_mod) #insignificant relationship

confint <- confint(dissimilarity_BC_total_sbt_max_plus_survey_mod) #confidence interval

coefs <- data.frame(coef(summary(dissimilarity_BC_total_sbt_max_plus_survey_mod))) #table with coefficients, #Satterthwaite's degrees of freedom, and p-value approximation for fixed effects

#input results into table
total_dissimilarity_sbt_model_results_row <- data.table(matrix(c("sbt", #depth
                                                 F, #scaled
                                                 "yearly_max_bypoint_avg + survey_unit", #model structure
                                                 paste0(round(coef(dissimilarity_BC_total_sbt_max_plus_survey_mod)[[1]],3),
                                                        " ± ",
                                                        round((confint[1,2]-confint[1,1])/2,3)), #intercept
                                                 paste0(round(coef(dissimilarity_BC_total_sbt_max_plus_survey_mod)[[2]],3),
                                                        " ± ",
                                                        round((confint[2,2]-confint[2,1])/2,3)), #coef
                                                 signif(coefs[2,4],3), #pvalue
                                                 round(summary(dissimilarity_BC_total_sbt_max_plus_survey_mod)$r.squared,3),#rsquared
                                                 round(AICc(dissimilarity_BC_total_sbt_max_plus_survey_mod),3)
                                                 ),nrow = 1))

#merge with larger table
total_dissimilarity_sbt_model_results <- rbind(total_dissimilarity_sbt_model_results,total_dissimilarity_sbt_model_results_row, use.names = F)

#########################
#max TEMP * SURVEY UNIT
#########################
dissimilarity_BC_total_sbt_max_times_survey_mod <- lm(bray_curtis_dissimilarity_total_mean ~ yearly_max_bypoint_avg * survey_unit, data = dissimilarities_temp) #this converges (allows both intercept and slope to vary with survey_unit)

summary(dissimilarity_BC_total_sbt_max_times_survey_mod) #insignificant relationship

confint <- confint(dissimilarity_BC_total_sbt_max_times_survey_mod) #confidence interval

coefs <- data.frame(coef(summary(dissimilarity_BC_total_sbt_max_times_survey_mod))) #table with coefficients, #Satterthwaite's degrees of freedom, and p-value approximation for fixed effects

#input results into table
total_dissimilarity_sbt_model_results_row <- data.table(matrix(c("sbt", #depth
                                                 F, #scaled
                                                 "yearly_max_bypoint_avg * survey_unit", #model structure
                                                 paste0(round(coef(dissimilarity_BC_total_sbt_max_times_survey_mod)[[1]],3),
                                                        " ± ",
                                                        round((confint[1,2]-confint[1,1])/2,3)), #intercept
                                                 paste0(round(coef(dissimilarity_BC_total_sbt_max_times_survey_mod)[[2]],3),
                                                        " ± ",
                                                        round((confint[2,2]-confint[2,1])/2,3)), #coef
                                                 signif(coefs[2,4],3), #pvalue
                                                 round(summary(dissimilarity_BC_total_sbt_max_times_survey_mod)$r.squared,3),#rsquared
                                                 round(AICc(dissimilarity_BC_total_sbt_max_times_survey_mod),3)
                                                 ),nrow = 1))

#merge with larger table
total_dissimilarity_sbt_model_results <- rbind(total_dissimilarity_sbt_model_results,total_dissimilarity_sbt_model_results_row, use.names = F)

```

####Seas sbt
```{r seas sbt}
#for all regions
(dissimilarity_BC_total_sbt_seas <- ggplot(data = dissimilarities_temp) +
  labs(x = "Seasonality of bottom temperature (˚C)",  y = "Bray Curtis total dissimilarity") +
  geom_point(aes(x = yearly_seas_bypoint_avg, y = bray_curtis_dissimilarity_total_mean, color = survey_unit), alpha = 0.3) +
  geom_line(aes(x = yearly_seas_bypoint_avg, y = bray_curtis_dissimilarity_total_mean, color = survey_unit), stat="smooth", method = "lm") +
  scale_color_manual(values = c(palette_42)) +
  geom_smooth(aes(x = yearly_seas_bypoint_avg, y = bray_curtis_dissimilarity_total_mean), method = "lm", se = F, color  = "black") +
  theme_classic())

ggsave(dissimilarity_BC_total_sbt_seas, path  = here::here("figures","dissimilarities"), filename = "dissimilarity_BC_total_sbt_seas.jpg", height = 4, width = 9, unit = "in") 

dissimilarity_BC_total_sbt_seas_facet <- ggplot(data = dissimilarities_temp) +
  labs(x = "Seasonality of bottom temperature (˚C)",  y = "Bray Curtis total dissimilarity") +
  geom_point(aes(x = yearly_seas_bypoint_avg, y = bray_curtis_dissimilarity_total_mean), alpha = 0.3) +
  geom_smooth(aes(x = yearly_seas_bypoint_avg, y = bray_curtis_dissimilarity_total_mean), method = "lm") +
  facet_wrap(~survey_unit, scales = "free") +
  theme_classic()

ggsave(dissimilarity_BC_total_sbt_seas_facet, path  = here::here("figures","dissimilarities"), filename = "dissimilarity_BC_total_sbt_seas_facet.jpg", height = 10, width = 11, unit = "in")

dissimilarity_BC_total_sbt_seas_mod <- lm(bray_curtis_dissimilarity_total_mean ~ yearly_seas_bypoint_avg, data = dissimilarities_temp) #this converges (allows both intercept and slope to vary with survey_unit)

summary(dissimilarity_BC_total_sbt_seas_mod) #insignificant relationship

confint <- confint(dissimilarity_BC_total_sbt_seas_mod) #confidence interval

coefs <- data.frame(coef(summary(dissimilarity_BC_total_sbt_seas_mod))) #table with coefficients, #Satterthwaite's degrees of freedom, and p-value approximation for fixed effects

#input results into table
total_dissimilarity_sbt_model_results_row <- data.table(matrix(c("sbt", #depth
                                                 F, #scaled
                                                 "yearly_seas_bypoint_avg", #model structure
                                                 paste0(round(coef(dissimilarity_BC_total_sbt_seas_mod)[[1]],3),
                                                        " ± ",
                                                        round((confint[1,2]-confint[1,1])/2,3)), #intercept
                                                 paste0(round(coef(dissimilarity_BC_total_sbt_seas_mod)[[2]],3),
                                                        " ± ",
                                                        round((confint[2,2]-confint[2,1])/2,3)), #coef
                                                 round(coefs[2,4],3), #pvalue
                                                 round(summary(dissimilarity_BC_total_sbt_seas_mod)$r.squared,3),#rsquared
                                                 round(AICc(dissimilarity_BC_total_sbt_seas_mod),3)
                                                 ),nrow = 1))

#merge with larger table
total_dissimilarity_sbt_model_results <- rbind(total_dissimilarity_sbt_model_results,total_dissimilarity_sbt_model_results_row, use.names = F)

#########################
#seas TEMP PLUS SURVEY UNIT
#########################
dissimilarity_BC_total_sbt_seas_plus_survey_mod <- lm(bray_curtis_dissimilarity_total_mean ~ yearly_seas_bypoint_avg + survey_unit, data = dissimilarities_temp) #this converges (allows both intercept and slope to vary with survey_unit)

summary(dissimilarity_BC_total_sbt_seas_plus_survey_mod) #insignificant relationship

confint <- confint(dissimilarity_BC_total_sbt_seas_plus_survey_mod) #confidence interval

coefs <- data.frame(coef(summary(dissimilarity_BC_total_sbt_seas_plus_survey_mod))) #table with coefficients, #Satterthwaite's degrees of freedom, and p-value approximation for fixed effects

#input results into table
total_dissimilarity_sbt_model_results_row <- data.table(matrix(c("sbt", #depth
                                                 F, #scaled
                                                 "yearly_seas_bypoint_avg + survey_unit", #model structure
                                                 paste0(round(coef(dissimilarity_BC_total_sbt_seas_plus_survey_mod)[[1]],3),
                                                        " ± ",
                                                        round((confint[1,2]-confint[1,1])/2,3)), #intercept
                                                 paste0(round(coef(dissimilarity_BC_total_sbt_seas_plus_survey_mod)[[2]],3),
                                                        " ± ",
                                                        round((confint[2,2]-confint[2,1])/2,3)), #coef
                                                 signif(coefs[2,4],3), #model pvalue
                                                 round(summary(dissimilarity_BC_total_sbt_seas_plus_survey_mod)$r.squared,3),#rsquared
                                                 round(AICc(dissimilarity_BC_total_sbt_seas_plus_survey_mod),3)
                                                 ),nrow = 1))

#merge with larger table
total_dissimilarity_sbt_model_results <- rbind(total_dissimilarity_sbt_model_results,total_dissimilarity_sbt_model_results_row, use.names = F)

#########################
#seas TEMP * SURVEY UNIT
#########################
dissimilarity_BC_total_sbt_seas_times_survey_mod <- lm(bray_curtis_dissimilarity_total_mean ~ yearly_seas_bypoint_avg * survey_unit, data = dissimilarities_temp) #this converges (allows both intercept and slope to vary with survey_unit)

summary(dissimilarity_BC_total_sbt_seas_times_survey_mod) #insignificant relationship

confint <- confint(dissimilarity_BC_total_sbt_seas_times_survey_mod) #confidence interval

coefs <- data.frame(coef(summary(dissimilarity_BC_total_sbt_seas_times_survey_mod))) #table with coefficients, #Satterthwaite's degrees of freedom, and p-value approximation for fixed effects

#input results into table
total_dissimilarity_sbt_model_results_row <- data.table(matrix(c("sbt", #depth
                                                 F, #scaled
                                                 "yearly_seas_bypoint_avg * survey_unit", #model structure
                                                 paste0(round(coef(dissimilarity_BC_total_sbt_seas_times_survey_mod)[[1]],3),
                                                        " ± ",
                                                        round((confint[1,2]-confint[1,1])/2,3)), #intercept
                                                 paste0(round(coef(dissimilarity_BC_total_sbt_seas_times_survey_mod)[[2]],3),
                                                        " ± ",
                                                        round((confint[2,2]-confint[2,1])/2,3)), #coef
                                                 signif(coefs[2,4],3), #pvalue
                                                 round(summary(dissimilarity_BC_total_sbt_seas_times_survey_mod)$r.squared,3),#rsquared
                                                 round(AICc(dissimilarity_BC_total_sbt_seas_times_survey_mod),3)
                                                 ),nrow = 1))

#merge with larger table
total_dissimilarity_sbt_model_results <- rbind(total_dissimilarity_sbt_model_results,total_dissimilarity_sbt_model_results_row, use.names = F)


```
SD of mean
```{r}
#for all regions
(dissimilarity_BC_total_sbt_meanSD <- ggplot(data = dissimilarities_temp) +
  labs(x = "SD ",  y = "Bray Curtis total dissimilarity") +
  geom_point(aes(x = yearly_mean_bypoint_SD, y = bray_curtis_dissimilarity_total_mean, color = survey_unit), alpha = 0.3) +
  geom_line(aes(x = yearly_mean_bypoint_SD, y = bray_curtis_dissimilarity_total_mean, color = survey_unit), stat="smooth", method = "lm") +
  scale_color_manual(values = c(palette_42)) +
  geom_smooth(aes(x = yearly_mean_bypoint_SD, y = bray_curtis_dissimilarity_total_mean), method = "lm", se = F, color  = "black") +
  theme_classic())

ggsave(dissimilarity_BC_total_sbt_meanSD, path  = here::here("figures","dissimilarities"), filename = "dissimilarity_BC_total_sbt_meanSD.jpg", height = 4, width = 9, unit = "in") 

dissimilarity_BC_total_sbt_meanSD_facet <- ggplot(data = dissimilarities_temp) +
  labs(x = "SD of mean bottom temperature",  y = "Bray Curtis total dissimilarity") +
  geom_point(aes(x = yearly_mean_bypoint_SD, y = bray_curtis_dissimilarity_total_mean), alpha = 0.3) +
  geom_smooth(aes(x = yearly_mean_bypoint_SD, y = bray_curtis_dissimilarity_total_mean), method = "lm") +
  facet_wrap(~survey_unit, scales = "free") +
  theme_classic()

ggsave(dissimilarity_BC_total_sbt_meanSD_facet, path  = here::here("figures","dissimilarities"), filename = "dissimilarity_BC_total_sbt_meanSD_facet.jpg", height = 10, width = 11, unit = "in")

dissimilarity_BC_total_sbt_meanSD_mod <- lm(bray_curtis_dissimilarity_total_mean ~ yearly_mean_bypoint_SD, data = dissimilarities_temp) #this converges (allows both intercept and slope to vary with survey_unit)

summary(dissimilarity_BC_total_sbt_meanSD_mod) #insignificant relationship

confint <- confint(dissimilarity_BC_total_sbt_meanSD_mod) #confidence interval

coefs <- data.frame(coef(summary(dissimilarity_BC_total_sbt_meanSD_mod))) #table with coefficients, #Satterthwaite's degrees of freedom, and p-value approximation for fixed effects

#input results into table
total_dissimilarity_sbt_model_results_row <- data.table(matrix(c("sbt", #depth
                                                 F, #scaled
                                                 "yearly_mean_bypoint_SD", #model structure
                                                 paste0(round(coef(dissimilarity_BC_total_sbt_meanSD_mod)[[1]],3),
                                                        " ± ",
                                                        round((confint[1,2]-confint[1,1])/2,3)), #intercept
                                                 paste0(round(coef(dissimilarity_BC_total_sbt_meanSD_mod)[[2]],3),
                                                        " ± ",
                                                        round((confint[2,2]-confint[2,1])/2,3)), #coef
                                                 round(coefs[2,4],3), #pvalue
                                                 round(summary(dissimilarity_BC_total_sbt_meanSD_mod)$r.squared,3),#rsquared
                                                 round(AICc(dissimilarity_BC_total_sbt_meanSD_mod),3)
                                                 ),nrow = 1))

#merge with larger table
total_dissimilarity_sbt_model_results <- rbind(total_dissimilarity_sbt_model_results,total_dissimilarity_sbt_model_results_row, use.names = F)

#########################
#SD TEMP PLUS SURVEY UNIT
#########################
dissimilarity_BC_total_sbt_meanSD_plus_survey_mod <- lm(bray_curtis_dissimilarity_total_mean ~ yearly_mean_bypoint_SD + survey_unit, data = dissimilarities_temp) #this converges (allows both intercept and slope to vary with survey_unit)

summary(dissimilarity_BC_total_sbt_meanSD_plus_survey_mod) #insignificant relationship

confint <- confint(dissimilarity_BC_total_sbt_meanSD_plus_survey_mod) #confidence interval

coefs <- data.frame(coef(summary(dissimilarity_BC_total_sbt_meanSD_plus_survey_mod))) #table with coefficients, #Satterthwaite's degrees of freedom, and p-value approximation for fixed effects

#input results into table
total_dissimilarity_sbt_model_results_row <- data.table(matrix(c("sbt", #depth
                                                 F, #scaled
                                                 "yearly_mean_bypoint_SD + survey_unit", #model structure
                                                 paste0(round(coef(dissimilarity_BC_total_sbt_meanSD_plus_survey_mod)[[1]],3),
                                                        " ± ",
                                                        round((confint[1,2]-confint[1,1])/2,3)), #intercept
                                                 paste0(round(coef(dissimilarity_BC_total_sbt_meanSD_plus_survey_mod)[[2]],3),
                                                        " ± ",
                                                        round((confint[2,2]-confint[2,1])/2,3)), #coef
                                                 signif(round(coefs[2,4],3)), #pvalue
                                                 round(summary(dissimilarity_BC_total_sbt_meanSD_plus_survey_mod)$r.squared,3),#rsquared
                                                 round(AICc(dissimilarity_BC_total_sbt_meanSD_plus_survey_mod),3)
                                                 ),nrow = 1))

#merge with larger table
total_dissimilarity_sbt_model_results <- rbind(total_dissimilarity_sbt_model_results,total_dissimilarity_sbt_model_results_row, use.names = F)

#########################
#SD TEMP * SURVEY UNIT
#########################
dissimilarity_BC_total_sbt_meanSD_times_survey_mod <- lm(bray_curtis_dissimilarity_total_mean ~ yearly_mean_bypoint_SD * survey_unit, data = dissimilarities_temp) #this converges (allows both intercept and slope to vary with survey_unit)

summary(dissimilarity_BC_total_sbt_meanSD_times_survey_mod) #insignificant relationship

confint <- confint(dissimilarity_BC_total_sbt_meanSD_times_survey_mod) #confidence interval

coefs <- data.frame(coef(summary(dissimilarity_BC_total_sbt_meanSD_times_survey_mod))) #table with coefficients, #Satterthwaite's degrees of freedom, and p-value approximation for fixed effects

#input results into table
total_dissimilarity_sbt_model_results_row <- data.table(matrix(c("sbt", #depth
                                                 F, #scaled
                                                 "yearly_mean_bypoint_SD * survey_unit", #model structure
                                                 paste0(round(coef(dissimilarity_BC_total_sbt_meanSD_times_survey_mod)[[1]],3),
                                                        " ± ",
                                                        round((confint[1,2]-confint[1,1])/2,3)), #intercept
                                                 paste0(round(coef(dissimilarity_BC_total_sbt_meanSD_times_survey_mod)[[2]],3),
                                                        " ± ",
                                                        round((confint[2,2]-confint[2,1])/2,3)), #coef
                                                 signif(coefs[2,4],3), #pvalue
                                                 round(summary(dissimilarity_BC_total_sbt_meanSD_times_survey_mod)$r.squared,3),#rsquared
                                                 round(AICc(dissimilarity_BC_total_sbt_meanSD_times_survey_mod),3)
                                                 ),nrow = 1))

#merge with larger table
total_dissimilarity_sbt_model_results <- rbind(total_dissimilarity_sbt_model_results,total_dissimilarity_sbt_model_results_row, use.names = F)
```


###Raw temp values: Comparing temperature predictors for raw
```{r compare raw temp predictors}

setorder(total_dissimilarity_sbt_model_results,AICc)
View(total_dissimilarity_sbt_model_results)


```

---
##SCALED
###Avg of grid cells
####Mean temperature scaled
```{r mean sbt scaled}
#for all regions
(dissimilarity_BC_total_sbt_mean.s <- ggplot(data = dissimilarities_temp) +
    labs(x = "Mean bottom temperature scaled (˚C)",  y = "Bray Curtis total dissimilarity") +
    geom_point(aes(x = yearly_mean_bypoint_avg.s, y = bray_curtis_dissimilarity_total_mean, color = survey_unit), alpha = 0.3) +
    geom_line(aes(x = yearly_mean_bypoint_avg.s, y = bray_curtis_dissimilarity_total_mean, color = survey_unit), stat="smooth", method = "lm") +
    scale_color_manual(values = c(palette_42)) +
    geom_smooth(aes(x = yearly_mean_bypoint_avg.s, y = bray_curtis_dissimilarity_total_mean), method = "lm", se = F, color  = "black") +
    theme_classic())

ggsave(dissimilarity_BC_total_sbt_mean.s, path  = here::here("figures","dissimilarities"), filename = "dissimilarity_BC_total_sbt_mean.s.jpg", height = 4, width = 9, unit = "in") 

dissimilarity_BC_total_sbt_mean.s_facet <- ggplot(data = dissimilarities_temp) +
  labs(x = "Mean bottom temperature scaled (˚C)",  y = "Bray Curtis total dissimilarity") +
  geom_point(aes(x = yearly_mean_bypoint_avg.s, y = bray_curtis_dissimilarity_total_mean), alpha = 0.3) +
  geom_smooth(aes(x = yearly_mean_bypoint_avg.s, y = bray_curtis_dissimilarity_total_mean), method = "lm") +
  facet_wrap(~survey_unit, scales = "free") +
  theme_classic()

ggsave(dissimilarity_BC_total_sbt_mean.s_facet, path  = here::here("figures","dissimilarities"), filename = "dissimilarity_BC_total_sbt_mean.s_facet.jpg", height = 10, width = 11, unit = "in")

dissimilarity_BC_total_sbt_mean.s_mod <- lm(bray_curtis_dissimilarity_total_mean ~ yearly_mean_bypoint_avg.s, data = dissimilarities_temp) #this converges (allows both intercept and slope to vary with survey_unit)

summary(dissimilarity_BC_total_sbt_mean.s_mod) #insignificant relationship

confint <- confint(dissimilarity_BC_total_sbt_mean.s_mod) #confidence interval

coefs <- data.frame(coef(summary(dissimilarity_BC_total_sbt_mean.s_mod))) #table with coefficients, #Satterthwaite's degrees of freedom, and p-value approximation for fixed effects

#input results into table
total_dissimilarity_sbt_model_results_row <- data.table(matrix(c("sbt", #depth
                                                                 T, #scaled
                                                                 "yearly_mean_bypoint_avg.s", #model structure
                                                                 paste0(round(coef(dissimilarity_BC_total_sbt_mean.s_mod)[[1]],3),
                                                                        " ± ",
                                                                        round((confint[1,2]-confint[1,1])/2,3)), #intercept
                                                                 paste0(round(coef(dissimilarity_BC_total_sbt_mean.s_mod)[[2]],3),
                                                                        " ± ",
                                                                        round((confint[2,2]-confint[2,1])/2,3)), #coef
                                                                 round(coefs[2,4],3), #pvalue
                                                                 round(summary(dissimilarity_BC_total_sbt_mean.s_mod)$r.squared,3),#rsquared
                                                                 round(AICc(dissimilarity_BC_total_sbt_mean.s_mod),3)
),nrow = 1))

#merge with larger table
total_dissimilarity_sbt_model_results <- rbind(total_dissimilarity_sbt_model_results,total_dissimilarity_sbt_model_results_row, use.names = F)

#########################
#MEAN TEMP PLUS SURVEY UNIT
#########################
dissimilarity_BC_total_sbt_mean.s_plus_survey_mod <- lm(bray_curtis_dissimilarity_total_mean ~ yearly_mean_bypoint_avg.s + survey_unit, data = dissimilarities_temp) #this converges (allows both intercept and slope to vary with survey_unit)

summary(dissimilarity_BC_total_sbt_mean.s_plus_survey_mod) #insignificant relationship

confint <- confint(dissimilarity_BC_total_sbt_mean.s_plus_survey_mod) #confidence interval

coefs <- data.frame(coef(summary(dissimilarity_BC_total_sbt_mean.s_plus_survey_mod))) #table with coefficients, #Satterthwaite's degrees of freedom, and p-value approximation for fixed effects

#input results into table
total_dissimilarity_sbt_model_results_row <- data.table(matrix(c("sbt", #depth
                                                                 T, #scaled
                                                                 "yearly_mean_bypoint_avg.s + survey_unit", #model structure
                                                                 paste0(round(coef(dissimilarity_BC_total_sbt_mean.s_plus_survey_mod)[[1]],3),
                                                                        " ± ",
                                                                        round((confint[1,2]-confint[1,1])/2,3)), #intercept
                                                                 paste0(round(coef(dissimilarity_BC_total_sbt_mean.s_plus_survey_mod)[[2]],3),
                                                                        " ± ",
                                                                        round((confint[2,2]-confint[2,1])/2,3)), #coef
                                                                 signif(coefs[2,4],3), #pvalue
                                                                 round(summary(dissimilarity_BC_total_sbt_mean.s_plus_survey_mod)$r.squared,3),#rsquared
                                                                 round(AICc(dissimilarity_BC_total_sbt_mean.s_plus_survey_mod),3)
),nrow = 1))

#merge with larger table
total_dissimilarity_sbt_model_results <- rbind(total_dissimilarity_sbt_model_results,total_dissimilarity_sbt_model_results_row, use.names = F)

#########################
#MEAN TEMP * SURVEY UNIT
#########################
dissimilarity_BC_total_sbt_mean.s_times_survey_mod <- lm(bray_curtis_dissimilarity_total_mean ~ yearly_mean_bypoint_avg.s * survey_unit, data = dissimilarities_temp) #this converges (allows both intercept and slope to vary with survey_unit)

summary(dissimilarity_BC_total_sbt_mean.s_times_survey_mod) #insignificant relationship

confint <- confint(dissimilarity_BC_total_sbt_mean.s_times_survey_mod) #confidence interval

coefs <- data.frame(coef(summary(dissimilarity_BC_total_sbt_mean.s_times_survey_mod))) #table with coefficients, #Satterthwaite's degrees of freedom, and p-value approximation for fixed effects

#input results into table
total_dissimilarity_sbt_model_results_row <- data.table(matrix(c("sbt", #depth
                                                                 T, #scaled
                                                                 "yearly_mean_bypoint_avg.s * survey_unit", #model structure
                                                                 paste0(round(coef(dissimilarity_BC_total_sbt_mean.s_times_survey_mod)[[1]],3),
                                                                        " ± ",
                                                                        round((confint[1,2]-confint[1,1])/2,3)), #intercept
                                                                 paste0(round(coef(dissimilarity_BC_total_sbt_mean.s_times_survey_mod)[[2]],3),
                                                                        " ± ",
                                                                        round((confint[2,2]-confint[2,1])/2,3)), #coef
                                                                 signif(coefs[2,4],3), #pvalue
                                                                 round(summary(dissimilarity_BC_total_sbt_mean.s_times_survey_mod)$r.squared,3),#rsquared
                                                                 round(AICc(dissimilarity_BC_total_sbt_mean.s_times_survey_mod),3)
),nrow = 1))

#merge with larger table
total_dissimilarity_sbt_model_results <- rbind(total_dissimilarity_sbt_model_results,total_dissimilarity_sbt_model_results_row, use.names = F)
```

####Min temperature
```{r min sbt scaled}
#for all regions
(dissimilarity_BC_total_sbt_min.s <- ggplot(data = dissimilarities_temp) +
    labs(x = "Min bottom temperature scaled (˚C)",  y = "Bray Curtis total dissimilarity") +
    geom_point(aes(x = yearly_min_bypoint_avg.s, y = bray_curtis_dissimilarity_total_mean, color = survey_unit), alpha = 0.3) +
    geom_line(aes(x = yearly_min_bypoint_avg.s, y = bray_curtis_dissimilarity_total_mean, color = survey_unit), stat="smooth", method = "lm") +
    scale_color_manual(values = c(palette_42)) +
    geom_smooth(aes(x = yearly_min_bypoint_avg.s, y = bray_curtis_dissimilarity_total_mean), method = "lm", se = F, color  = "black") +
    theme_classic())

ggsave(dissimilarity_BC_total_sbt_min.s, path  = here::here("figures","dissimilarities"), filename = "dissimilarity_BC_total_sbt_min.s.jpg", height = 4, width = 9, unit = "in") 

dissimilarity_BC_total_sbt_min.s_facet <- ggplot(data = dissimilarities_temp) +
  labs(x = "Min bottom temperature scaled (˚C)",  y = "Bray Curtis total dissimilarity") +
  geom_point(aes(x = yearly_min_bypoint_avg.s, y = bray_curtis_dissimilarity_total_mean), alpha = 0.3) +
  geom_smooth(aes(x = yearly_min_bypoint_avg.s, y = bray_curtis_dissimilarity_total_mean), method = "lm") +
  facet_wrap(~survey_unit, scales = "free") +
  theme_classic()

ggsave(dissimilarity_BC_total_sbt_min.s_facet, path  = here::here("figures","dissimilarities"), filename = "dissimilarity_BC_total_sbt_min.s_facet.jpg", height = 10, width = 11, unit = "in")

dissimilarity_BC_total_sbt_min.s_mod <- lm(bray_curtis_dissimilarity_total_mean ~ yearly_min_bypoint_avg.s, data = dissimilarities_temp) #this converges (allows both intercept and slope to vary with survey_unit)

summary(dissimilarity_BC_total_sbt_min.s_mod) #insignificant relationship

confint <- confint(dissimilarity_BC_total_sbt_min.s_mod) #confidence interval

coefs <- data.frame(coef(summary(dissimilarity_BC_total_sbt_min.s_mod))) #table with coefficients, #Satterthwaite's degrees of freedom, and p-value approximation for fixed effects

#input results into table
total_dissimilarity_sbt_model_results_row <- data.table(matrix(c("sbt", #depth
                                                                 T, #scaled
                                                                 "yearly_min_bypoint_avg.s", #model structure
                                                                 paste0(round(coef(dissimilarity_BC_total_sbt_min.s_mod)[[1]],3),
                                                                        " ± ",
                                                                        round((confint[1,2]-confint[1,1])/2,3)), #intercept
                                                                 paste0(round(coef(dissimilarity_BC_total_sbt_min.s_mod)[[2]],3),
                                                                        " ± ",
                                                                        round((confint[2,2]-confint[2,1])/2,3)), #coef
                                                                 round(coefs[2,4],3), #pvalue
                                                                 round(summary(dissimilarity_BC_total_sbt_min.s_mod)$r.squared,3),#rsquared
                                                                 round(AICc(dissimilarity_BC_total_sbt_min.s_mod),3)
),nrow = 1))

#merge with larger table
total_dissimilarity_sbt_model_results <- rbind(total_dissimilarity_sbt_model_results,total_dissimilarity_sbt_model_results_row, use.names = F)

#########################
#min TEMP PLUS SURVEY UNIT
#########################
dissimilarity_BC_total_sbt_min.s_plus_survey_mod <- lm(bray_curtis_dissimilarity_total_mean ~ yearly_min_bypoint_avg.s + survey_unit, data = dissimilarities_temp) #this converges (allows both intercept and slope to vary with survey_unit)

summary(dissimilarity_BC_total_sbt_min.s_plus_survey_mod) #insignificant relationship

confint <- confint(dissimilarity_BC_total_sbt_min.s_plus_survey_mod) #confidence interval

coefs <- data.frame(coef(summary(dissimilarity_BC_total_sbt_min.s_plus_survey_mod))) #table with coefficients, #Satterthwaite's degrees of freedom, and p-value approximation for fixed effects

#input results into table
total_dissimilarity_sbt_model_results_row <- data.table(matrix(c("sbt", #depth
                                                                 T, #scaled
                                                                 "yearly_min_bypoint_avg.s + survey_unit", #model structure
                                                                 paste0(round(coef(dissimilarity_BC_total_sbt_min.s_plus_survey_mod)[[1]],3),
                                                                        " ± ",
                                                                        round((confint[1,2]-confint[1,1])/2,3)), #intercept
                                                                 paste0(round(coef(dissimilarity_BC_total_sbt_min.s_plus_survey_mod)[[2]],3),
                                                                        " ± ",
                                                                        round((confint[2,2]-confint[2,1])/2,3)), #coef
                                                                 signif(coefs[2,4],3), #model pvalue
                                                                 round(summary(dissimilarity_BC_total_sbt_min.s_plus_survey_mod)$r.squared,3),#rsquared
                                                                 round(AICc(dissimilarity_BC_total_sbt_min.s_plus_survey_mod),3)
),nrow = 1))

#merge with larger table
total_dissimilarity_sbt_model_results <- rbind(total_dissimilarity_sbt_model_results,total_dissimilarity_sbt_model_results_row, use.names = F)

#########################
#min TEMP * SURVEY UNIT
#########################
dissimilarity_BC_total_sbt_min.s_times_survey_mod <- lm(bray_curtis_dissimilarity_total_mean ~ yearly_min_bypoint_avg.s * survey_unit, data = dissimilarities_temp) #this converges (allows both intercept and slope to vary with survey_unit)

summary(dissimilarity_BC_total_sbt_min.s_times_survey_mod) #insignificant relationship

confint <- confint(dissimilarity_BC_total_sbt_min.s_times_survey_mod) #confidence interval

coefs <- data.frame(coef(summary(dissimilarity_BC_total_sbt_min.s_times_survey_mod))) #table with coefficients, #Satterthwaite's degrees of freedom, and p-value approximation for fixed effects

#input results into table
total_dissimilarity_sbt_model_results_row <- data.table(matrix(c("sbt", #depth
                                                                 T, #scaled
                                                                 "yearly_min_bypoint_avg.s * survey_unit", #model structure
                                                                 paste0(round(coef(dissimilarity_BC_total_sbt_min.s_times_survey_mod)[[1]],3),
                                                                        " ± ",
                                                                        round((confint[1,2]-confint[1,1])/2,3)), #intercept
                                                                 paste0(round(coef(dissimilarity_BC_total_sbt_min.s_times_survey_mod)[[2]],3),
                                                                        " ± ",
                                                                        round((confint[2,2]-confint[2,1])/2,3)), #coef
                                                                 signif(coefs[2,4],3), #pvalue
                                                                 round(summary(dissimilarity_BC_total_sbt_min.s_times_survey_mod)$r.squared,3),#rsquared
                                                                 round(AICc(dissimilarity_BC_total_sbt_min.s_times_survey_mod),3)
),nrow = 1))

#merge with larger table
total_dissimilarity_sbt_model_results <- rbind(total_dissimilarity_sbt_model_results,total_dissimilarity_sbt_model_results_row, use.names = F)
```

####Max sbt
```{r max sbt scaled}
#for all regions
(dissimilarity_BC_total_sbt_max.s <- ggplot(data = dissimilarities_temp) +
    labs(x = "Max bottom temperature scaled (˚C)",  y = "Bray Curtis total dissimilarity") +
    geom_point(aes(x = yearly_max_bypoint_avg.s, y = bray_curtis_dissimilarity_total_mean, color = survey_unit), alpha = 0.3) +
    geom_line(aes(x = yearly_max_bypoint_avg.s, y = bray_curtis_dissimilarity_total_mean, color = survey_unit), stat="smooth", method = "lm") +
    scale_color_manual(values = c(palette_42)) +
    geom_smooth(aes(x = yearly_max_bypoint_avg.s, y = bray_curtis_dissimilarity_total_mean), method = "lm", se = F, color  = "black") +
    theme_classic())

ggsave(dissimilarity_BC_total_sbt_max.s, path  = here::here("figures","dissimilarities"), filename = "dissimilarity_BC_total_sbt_max.s.jpg", height = 4, width = 9, unit = "in") 

dissimilarity_BC_total_sbt_max.s_facet <- ggplot(data = dissimilarities_temp) +
  labs(x = "Max bottom temperature scaled (˚C)",  y = "Bray Curtis total dissimilarity") +
  geom_point(aes(x = yearly_max_bypoint_avg.s, y = bray_curtis_dissimilarity_total_mean), alpha = 0.3) +
  geom_smooth(aes(x = yearly_max_bypoint_avg.s, y = bray_curtis_dissimilarity_total_mean), method = "lm") +
  facet_wrap(~survey_unit, scales = "free") +
  theme_classic()

ggsave(dissimilarity_BC_total_sbt_max.s_facet, path  = here::here("figures","dissimilarities"), filename = "dissimilarity_BC_total_sbt_max.s_facet.jpg", height = 10, width = 11, unit = "in")

dissimilarity_BC_total_sbt_max.s_mod <- lm(bray_curtis_dissimilarity_total_mean ~ yearly_max_bypoint_avg.s, data = dissimilarities_temp) #this converges (allows both intercept and slope to vary with survey_unit)

summary(dissimilarity_BC_total_sbt_max.s_mod) #insignificant relationship

confint <- confint(dissimilarity_BC_total_sbt_max.s_mod) #confidence interval

coefs <- data.frame(coef(summary(dissimilarity_BC_total_sbt_max.s_mod))) #table with coefficients, #Satterthwaite's degrees of freedom, and p-value approximation for fixed effects

#input results into table
total_dissimilarity_sbt_model_results_row <- data.table(matrix(c("sbt", #depth
                                                                 T, #scaled
                                                                 "yearly_max_bypoint_avg.s", #model structure
                                                                 paste0(round(coef(dissimilarity_BC_total_sbt_max.s_mod)[[1]],3),
                                                                        " ± ",
                                                                        round((confint[1,2]-confint[1,1])/2,3)), #intercept
                                                                 paste0(round(coef(dissimilarity_BC_total_sbt_max.s_mod)[[2]],3),
                                                                        " ± ",
                                                                        round((confint[2,2]-confint[2,1])/2,3)), #coef
                                                                 round(coefs[2,4],3), #pvalue
                                                                 round(summary(dissimilarity_BC_total_sbt_max.s_mod)$r.squared,3),#rsquared
                                                                 round(AICc(dissimilarity_BC_total_sbt_max.s_mod),3)
),nrow = 1))

#merge with larger table
total_dissimilarity_sbt_model_results <- rbind(total_dissimilarity_sbt_model_results,total_dissimilarity_sbt_model_results_row, use.names = F)

#########################
#max TEMP PLUS SURVEY UNIT
#########################
dissimilarity_BC_total_sbt_max.s_plus_survey_mod <- lm(bray_curtis_dissimilarity_total_mean ~ yearly_max_bypoint_avg.s + survey_unit, data = dissimilarities_temp) #this converges (allows both intercept and slope to vary with survey_unit)

summary(dissimilarity_BC_total_sbt_max.s_plus_survey_mod) #insignificant relationship

confint <- confint(dissimilarity_BC_total_sbt_max.s_plus_survey_mod) #confidence interval

coefs <- data.frame(coef(summary(dissimilarity_BC_total_sbt_max.s_plus_survey_mod))) #table with coefficients, #Satterthwaite's degrees of freedom, and p-value approximation for fixed effects

#input results into table
total_dissimilarity_sbt_model_results_row <- data.table(matrix(c("sbt", #depth
                                                                 T, #scaled
                                                                 "yearly_max_bypoint_avg.s + survey_unit", #model structure
                                                                 paste0(round(coef(dissimilarity_BC_total_sbt_max.s_plus_survey_mod)[[1]],3),
                                                                        " ± ",
                                                                        round((confint[1,2]-confint[1,1])/2,3)), #intercept
                                                                 paste0(round(coef(dissimilarity_BC_total_sbt_max.s_plus_survey_mod)[[2]],3),
                                                                        " ± ",
                                                                        round((confint[2,2]-confint[2,1])/2,3)), #coef
                                                                 signif(coefs[2,4],3), #model pvalue
                                                                 round(summary(dissimilarity_BC_total_sbt_max.s_plus_survey_mod)$r.squared,3),#rsquared
                                                                 round(AICc(dissimilarity_BC_total_sbt_max.s_plus_survey_mod),3)
),nrow = 1))

#merge with larger table
total_dissimilarity_sbt_model_results <- rbind(total_dissimilarity_sbt_model_results,total_dissimilarity_sbt_model_results_row, use.names = F)

#########################
#max TEMP * SURVEY UNIT
#########################
dissimilarity_BC_total_sbt_max.s_times_survey_mod <- lm(bray_curtis_dissimilarity_total_mean ~ yearly_max_bypoint_avg.s * survey_unit, data = dissimilarities_temp) #this converges (allows both intercept and slope to vary with survey_unit)

summary(dissimilarity_BC_total_sbt_max.s_times_survey_mod) #insignificant relationship

confint <- confint(dissimilarity_BC_total_sbt_max.s_times_survey_mod) #confidence interval

coefs <- data.frame(coef(summary(dissimilarity_BC_total_sbt_max.s_times_survey_mod))) #table with coefficients, #Satterthwaite's degrees of freedom, and p-value approximation for fixed effects

#input results into table
total_dissimilarity_sbt_model_results_row <- data.table(matrix(c("sbt", #depth
                                                                 T, #scaled
                                                                 "yearly_max_bypoint_avg.s * survey_unit", #model structure
                                                                 paste0(round(coef(dissimilarity_BC_total_sbt_max.s_times_survey_mod)[[1]],3),
                                                                        " ± ",
                                                                        round((confint[1,2]-confint[1,1])/2,3)), #intercept
                                                                 paste0(round(coef(dissimilarity_BC_total_sbt_max.s_times_survey_mod)[[2]],3),
                                                                        " ± ",
                                                                        round((confint[2,2]-confint[2,1])/2,3)), #coef
                                                                 signif(coefs[2,4],3), #model pvalue
                                                                 round(summary(dissimilarity_BC_total_sbt_max.s_times_survey_mod)$r.squared,3),#rsquared
                                                                 round(AICc(dissimilarity_BC_total_sbt_max.s_times_survey_mod),3)
),nrow = 1))

#merge with larger table
total_dissimilarity_sbt_model_results <- rbind(total_dissimilarity_sbt_model_results,total_dissimilarity_sbt_model_results_row, use.names = F)

```

####Seas sbt
```{r seas sbt scaled}
#for all regions
(dissimilarity_BC_total_sbt_seas.s <- ggplot(data = dissimilarities_temp) +
    labs(x = "Seasonality of bottom temperature scaled (˚C)",  y = "Bray Curtis total dissimilarity") +
    geom_point(aes(x = yearly_seas_bypoint_avg.s, y = bray_curtis_dissimilarity_total_mean, color = survey_unit), alpha = 0.3) +
    geom_line(aes(x = yearly_seas_bypoint_avg.s, y = bray_curtis_dissimilarity_total_mean, color = survey_unit), stat="smooth", method = "lm") +
    scale_color_manual(values = c(palette_42)) +
    geom_smooth(aes(x = yearly_seas_bypoint_avg.s, y = bray_curtis_dissimilarity_total_mean), method = "lm", se = F, color  = "black") +
    theme_classic())

ggsave(dissimilarity_BC_total_sbt_seas.s, path  = here::here("figures","dissimilarities"), filename = "dissimilarity_BC_total_sbt_seas.s.jpg", height = 4, width = 9, unit = "in") 

dissimilarity_BC_total_sbt_seas.s_facet <- ggplot(data = dissimilarities_temp) +
  labs(x = "Seasonality of bottom temperature scaled (˚C)",  y = "Bray Curtis total dissimilarity") +
  geom_point(aes(x = yearly_seas_bypoint_avg.s, y = bray_curtis_dissimilarity_total_mean), alpha = 0.3) +
  geom_smooth(aes(x = yearly_seas_bypoint_avg.s, y = bray_curtis_dissimilarity_total_mean), method = "lm") +
  facet_wrap(~survey_unit, scales = "free") +
  theme_classic()

ggsave(dissimilarity_BC_total_sbt_seas.s_facet, path  = here::here("figures","dissimilarities"), filename = "dissimilarity_BC_total_sbt_seas.s_facet.jpg", height = 10, width = 11, unit = "in")

dissimilarity_BC_total_sbt_seas.s_mod <- lm(bray_curtis_dissimilarity_total_mean ~ yearly_seas_bypoint_avg.s, data = dissimilarities_temp) #this converges (allows both intercept and slope to vary with survey_unit)

summary(dissimilarity_BC_total_sbt_seas.s_mod) #insignificant relationship

confint <- confint(dissimilarity_BC_total_sbt_seas.s_mod) #confidence interval

coefs <- data.frame(coef(summary(dissimilarity_BC_total_sbt_seas.s_mod))) #table with coefficients, #Satterthwaite's degrees of freedom, and p-value approximation for fixed effects

#input results into table
total_dissimilarity_sbt_model_results_row <- data.table(matrix(c("sbt", #depth
                                                                 T, #scaled
                                                                 "yearly_seas_bypoint_avg.s", #model structure
                                                                 paste0(round(coef(dissimilarity_BC_total_sbt_seas.s_mod)[[1]],3),
                                                                        " ± ",
                                                                        round((confint[1,2]-confint[1,1])/2,3)), #intercept
                                                                 paste0(round(coef(dissimilarity_BC_total_sbt_seas.s_mod)[[2]],3),
                                                                        " ± ",
                                                                        round((confint[2,2]-confint[2,1])/2,3)), #coef
                                                                 round(coefs[2,4],3), #pvalue
                                                                 round(summary(dissimilarity_BC_total_sbt_seas.s_mod)$r.squared,3),#rsquared
                                                                 round(AICc(dissimilarity_BC_total_sbt_seas.s_mod),3)
),nrow = 1))

#merge with larger table
total_dissimilarity_sbt_model_results <- rbind(total_dissimilarity_sbt_model_results,total_dissimilarity_sbt_model_results_row, use.names = F)

#########################
#seas TEMP PLUS SURVEY UNIT
#########################
dissimilarity_BC_total_sbt_seas.s_plus_survey_mod <- lm(bray_curtis_dissimilarity_total_mean ~ yearly_seas_bypoint_avg.s + survey_unit, data = dissimilarities_temp) #this converges (allows both intercept and slope to vary with survey_unit)

summary(dissimilarity_BC_total_sbt_seas.s_plus_survey_mod) #insignificant relationship

confint <- confint(dissimilarity_BC_total_sbt_seas.s_plus_survey_mod) #confidence interval

coefs <- data.frame(coef(summary(dissimilarity_BC_total_sbt_seas.s_plus_survey_mod))) #table with coefficients, #Satterthwaite's degrees of freedom, and p-value approximation for fixed effects

#input results into table
total_dissimilarity_sbt_model_results_row <- data.table(matrix(c("sbt", #depth
                                                                 T, #scaled
                                                                 "yearly_seas_bypoint_avg.s + survey_unit", #model structure
                                                                 paste0(round(coef(dissimilarity_BC_total_sbt_seas.s_plus_survey_mod)[[1]],3),
                                                                        " ± ",
                                                                        round((confint[1,2]-confint[1,1])/2,3)), #intercept
                                                                 paste0(round(coef(dissimilarity_BC_total_sbt_seas.s_plus_survey_mod)[[2]],3),
                                                                        " ± ",
                                                                        round((confint[2,2]-confint[2,1])/2,3)), #coef
                                                                 signif(coefs[2,4],3), #model pvalue
                                                                 round(summary(dissimilarity_BC_total_sbt_seas.s_plus_survey_mod)$r.squared,3),#rsquared
                                                                 round(AICc(dissimilarity_BC_total_sbt_seas.s_plus_survey_mod),3)
),nrow = 1))

#merge with larger table
total_dissimilarity_sbt_model_results <- rbind(total_dissimilarity_sbt_model_results,total_dissimilarity_sbt_model_results_row, use.names = F)

#########################
#seas TEMP * SURVEY UNIT
#########################
dissimilarity_BC_total_sbt_seas.s_times_survey_mod <- lm(bray_curtis_dissimilarity_total_mean ~ yearly_seas_bypoint_avg.s * survey_unit, data = dissimilarities_temp) #this converges (allows both intercept and slope to vary with survey_unit)

summary(dissimilarity_BC_total_sbt_seas.s_times_survey_mod) #insignificant relationship

confint <- confint(dissimilarity_BC_total_sbt_seas.s_times_survey_mod) #confidence interval

coefs <- data.frame(coef(summary(dissimilarity_BC_total_sbt_seas.s_times_survey_mod))) #table with coefficients, #Satterthwaite's degrees of freedom, and p-value approximation for fixed effects

#input results into table
total_dissimilarity_sbt_model_results_row <- data.table(matrix(c("sbt", #depth
                                                                 T, #scaled
                                                                 "yearly_seas_bypoint_avg.s * survey_unit", #model structure
                                                                 paste0(round(coef(dissimilarity_BC_total_sbt_seas.s_times_survey_mod)[[1]],3),
                                                                        " ± ",
                                                                        round((confint[1,2]-confint[1,1])/2,3)), #intercept
                                                                 paste0(round(coef(dissimilarity_BC_total_sbt_seas.s_times_survey_mod)[[2]],3),
                                                                        " ± ",
                                                                        round((confint[2,2]-confint[2,1])/2,3)), #coef
                                                                 signif(coefs[2,4],3), #model pvalue
                                                                 round(summary(dissimilarity_BC_total_sbt_seas.s_times_survey_mod)$r.squared,3),#rsquared
                                                                 round(AICc(dissimilarity_BC_total_sbt_seas.s_times_survey_mod),3)
),nrow = 1))

#merge with larger table
total_dissimilarity_sbt_model_results <- rbind(total_dissimilarity_sbt_model_results,total_dissimilarity_sbt_model_results_row, use.names = F)


```

SD of mean
```{r sd of mean scaled}
#for all regions
(dissimilarity_BC_total_sbt_mean_SD.s <- ggplot(data = dissimilarities_temp) +
    labs(x = "SD ",  y = "Bray Curtis total dissimilarity") +
    geom_point(aes(x = yearly_mean_bypoint_SD.s, y = bray_curtis_dissimilarity_total_mean, color = survey_unit), alpha = 0.3) +
    geom_line(aes(x = yearly_mean_bypoint_SD.s, y = bray_curtis_dissimilarity_total_mean, color = survey_unit), stat="smooth", method = "lm") +
    scale_color_manual(values = c(palette_42)) +
    geom_smooth(aes(x = yearly_mean_bypoint_SD.s, y = bray_curtis_dissimilarity_total_mean), method = "lm", se = F, color  = "black") +
    theme_classic())

ggsave(dissimilarity_BC_total_sbt_mean_SD.s, path  = here::here("figures","dissimilarities"), filename = "dissimilarity_BC_total_sbt_mean_SD.s.jpg", height = 4, width = 9, unit = "in") 

dissimilarity_BC_total_sbt_mean_SD.s_facet <- ggplot(data = dissimilarities_temp) +
  labs(x = "SD of mean bottom temperature scaled",  y = "Bray Curtis total dissimilarity") +
  geom_point(aes(x = yearly_mean_bypoint_SD.s, y = bray_curtis_dissimilarity_total_mean), alpha = 0.3) +
  geom_smooth(aes(x = yearly_mean_bypoint_SD.s, y = bray_curtis_dissimilarity_total_mean), method = "lm") +
  facet_wrap(~survey_unit, scales = "free") +
  theme_classic()

ggsave(dissimilarity_BC_total_sbt_mean_SD.s_facet, path  = here::here("figures","dissimilarities"), filename = "dissimilarity_BC_total_sbt_mean_SD.s_facet.jpg", height = 10, width = 11, unit = "in")

dissimilarity_BC_total_sbt_mean_SD.s_mod <- lm(bray_curtis_dissimilarity_total_mean ~ yearly_mean_bypoint_SD.s, data = dissimilarities_temp) #this converges (allows both intercept and slope to vary with survey_unit)

summary(dissimilarity_BC_total_sbt_mean_SD.s_mod) #insignificant relationship

confint <- confint(dissimilarity_BC_total_sbt_mean_SD.s_mod) #confidence interval

coefs <- data.frame(coef(summary(dissimilarity_BC_total_sbt_mean_SD.s_mod))) #table with coefficients, #Satterthwaite's degrees of freedom, and p-value approximation for fixed effects

#input results into table
total_dissimilarity_sbt_model_results_row <- data.table(matrix(c("sbt", #depth
                                                                 T, #scaled
                                                                 "yearly_mean_bypoint_SD.s", #model structure
                                                                 paste0(round(coef(dissimilarity_BC_total_sbt_mean_SD.s_mod)[[1]],3),
                                                                        " ± ",
                                                                        round((confint[1,2]-confint[1,1])/2,3)), #intercept
                                                                 paste0(round(coef(dissimilarity_BC_total_sbt_mean_SD.s_mod)[[2]],3),
                                                                        " ± ",
                                                                        round((confint[2,2]-confint[2,1])/2,3)), #coef
                                                                 round(coefs[2,4],3), #pvalue
                                                                 round(summary(dissimilarity_BC_total_sbt_mean_SD.s_mod)$r.squared,3),#rsquared
                                                                 round(AICc(dissimilarity_BC_total_sbt_mean_SD.s_mod),3)
),nrow = 1))

#merge with larger table
total_dissimilarity_sbt_model_results <- rbind(total_dissimilarity_sbt_model_results,total_dissimilarity_sbt_model_results_row, use.names = F)

#########################
#SD TEMP PLUS SURVEY UNIT
#########################
dissimilarity_BC_total_sbt_mean_SD.s_plus_survey_mod <- lm(bray_curtis_dissimilarity_total_mean ~ yearly_mean_bypoint_SD.s + survey_unit, data = dissimilarities_temp) #this converges (allows both intercept and slope to vary with survey_unit)

summary(dissimilarity_BC_total_sbt_mean_SD.s_plus_survey_mod) #insignificant relationship

confint <- confint(dissimilarity_BC_total_sbt_mean_SD.s_plus_survey_mod) #confidence interval

coefs <- data.frame(coef(summary(dissimilarity_BC_total_sbt_mean_SD.s_plus_survey_mod))) #table with coefficients, #Satterthwaite's degrees of freedom, and p-value approximation for fixed effects

#input results into table
total_dissimilarity_sbt_model_results_row <- data.table(matrix(c("sbt", #depth
                                                                 T, #scaled
                                                                 "yearly_mean_bypoint_SD.s + survey_unit", #model structure
                                                                 paste0(round(coef(dissimilarity_BC_total_sbt_mean_SD.s_plus_survey_mod)[[1]],3),
                                                                        " ± ",
                                                                        round((confint[1,2]-confint[1,1])/2,3)), #intercept
                                                                 paste0(round(coef(dissimilarity_BC_total_sbt_mean_SD.s_plus_survey_mod)[[2]],3),
                                                                        " ± ",
                                                                        round((confint[2,2]-confint[2,1])/2,3)), #coef
                                                                 signif(coefs[2,4],3), #model pvalue
                                                                 round(summary(dissimilarity_BC_total_sbt_mean_SD.s_plus_survey_mod)$r.squared,3),#rsquared
                                                                 round(AICc(dissimilarity_BC_total_sbt_mean_SD.s_plus_survey_mod),3)
),nrow = 1))

#merge with larger table
total_dissimilarity_sbt_model_results <- rbind(total_dissimilarity_sbt_model_results,total_dissimilarity_sbt_model_results_row, use.names = F)

#########################
#SD TEMP * SURVEY UNIT
#########################
dissimilarity_BC_total_sbt_mean_SD.s_times_survey_mod <- lm(bray_curtis_dissimilarity_total_mean ~ yearly_mean_bypoint_SD.s * survey_unit, data = dissimilarities_temp) #this converges (allows both intercept and slope to vary with survey_unit)

summary(dissimilarity_BC_total_sbt_mean_SD.s_times_survey_mod) #insignificant relationship

confint <- confint(dissimilarity_BC_total_sbt_mean_SD.s_times_survey_mod) #confidence interval

coefs <- data.frame(coef(summary(dissimilarity_BC_total_sbt_mean_SD.s_times_survey_mod))) #table with coefficients, #Satterthwaite's degrees of freedom, and p-value approximation for fixed effects

#input results into table
total_dissimilarity_sbt_model_results_row <- data.table(matrix(c("sbt", #depth
                                                                 T, #scaled
                                                                 "yearly_mean_bypoint_SD.s * survey_unit", #model structure
                                                                 paste0(round(coef(dissimilarity_BC_total_sbt_mean_SD.s_times_survey_mod)[[1]],3),
                                                                        " ± ",
                                                                        round((confint[1,2]-confint[1,1])/2,3)), #intercept
                                                                 paste0(round(coef(dissimilarity_BC_total_sbt_mean_SD.s_times_survey_mod)[[2]],3),
                                                                        " ± ",
                                                                        round((confint[2,2]-confint[2,1])/2,3)), #coef
                                                                 signif(coefs[2,4],3), #model pvalue
                                                                 round(summary(dissimilarity_BC_total_sbt_mean_SD.s_times_survey_mod)$r.squared,3),#rsquared
                                                                 round(AICc(dissimilarity_BC_total_sbt_mean_SD.s_times_survey_mod),3)
),nrow = 1))

#merge with larger table
total_dissimilarity_sbt_model_results <- rbind(total_dissimilarity_sbt_model_results,total_dissimilarity_sbt_model_results_row, use.names = F)
```

#add survey unit only model
```{r}
#########################
#SD TEMP * SURVEY UNIT
#########################
dissimilarity_BC_total_survey_mod <- lm(bray_curtis_dissimilarity_total_mean ~ survey_unit, data = dissimilarities_temp) #this converges (allows both intercept and slope to vary with survey_unit)

summary(dissimilarity_BC_total_survey_mod) #insignificant relationship

confint <- confint(dissimilarity_BC_total_survey_mod) #confidence interval

coefs <- data.frame(coef(summary(dissimilarity_BC_total_survey_mod))) #table with coefficients, #Satterthwaite's degrees of freedom, and p-value approximation for fixed effects

#input results into table
total_dissimilarity_sbt_model_results_row <- data.table(matrix(c(NA, #depth
                                                                 NA, #scaled
                                                                 "survey_unit", #model structure
                                                                 paste0(round(coef(dissimilarity_BC_total_survey_mod)[[1]],3),
                                                                        " ± ",
                                                                        round((confint[1,2]-confint[1,1])/2,3)), #intercept
                                                                 NA, #coef
                                                                 NA, #pvalue
                                                                 round(summary(dissimilarity_BC_total_survey_mod)$r.squared,3),#rsquared
                                                                 round(AICc(dissimilarity_BC_total_survey_mod),3)
),nrow = 1))

#merge with larger table
total_dissimilarity_sbt_model_results <- rbind(total_dissimilarity_sbt_model_results,total_dissimilarity_sbt_model_results_row, use.names = F)
```


Comparing temperature predictors for all (absolute/raw and scaled)
```{r compare all temp predictors}

setorder(total_dissimilarity_sbt_model_results,AICc)
total_dissimilarity_sbt_model_results[,deltaAICc := (as.numeric(AICc) - min(as.numeric(AICc)))]
View(total_dissimilarity_sbt_model_results)
```


####What to carry forward:

best model:
	
yearly_mean_bypoint_avg * survey_unit

dissimilarity_BC_total_sbt_mean_times_survey_mod <- lm(bray_curtis_dissimilarity_total_mean ~ yearly_mean_bypoint_avg * survey_unit, data = dissimilarities_temp)

Merge dissimilarities temp with names for plotting
```{r}
dissimilarities_temp <- dissimilarities_temp[color_link, on = "survey_unit"]


```



Best models
```{r}
total_dissimilarity_sbt_model_results

saveRDS(total_dissimilarity_sbt_model_results, here::here("output","total_dissimilarity_sbt_model_results.rds"))
fwrite(total_dissimilarity_sbt_model_results, here::here("output","total_dissimilarity_sbt_model_results.csv"))
```


####Plot best absolute model

dissimilarity_BC_total_sbt_mean_times_survey_mod <- lm(bray_curtis_dissimilarity_total_mean ~ yearly_mean_bypoint_avg * survey_unit, data = dissimilarities_temp)


```{r}

#for all regions merged
(preds_sbt_mean_temp_survey <- ggplot() +
  labs(x = "Mean bottom temperature (˚C)",  y = "Bray Curtis total dissimilarity") +
  geom_point(data = dissimilarities_temp, aes(x = as.numeric(yearly_mean_bypoint_avg), y = bray_curtis_dissimilarity_total_mean), alpha = 0.3) +
  geom_smooth(data = dissimilarities_temp, aes(x = as.numeric(yearly_mean_bypoint_avg), y = bray_curtis_dissimilarity_total_mean), method = "lm") +
  theme_classic())

ggsave(preds_sbt_mean_temp_survey, path  = here::here("figures","predictor_variables"), filename = "preds_sbt_mean_temp_survey.jpg", height = 4, width = 6.5, unit = "in") 

(preds_sbt_mean_temp_survey_facet <- ggplot(data = dissimilarities_temp) +
  labs(x = "Mean bottom temperature (˚C)",  y = "Bray Curtis total dissimilarity") +
  geom_point(aes(x = as.numeric(yearly_mean_bypoint_avg), y = bray_curtis_dissimilarity_total_mean), alpha = 0.3) +
  geom_smooth(aes(x = as.numeric(yearly_mean_bypoint_avg), y = bray_curtis_dissimilarity_total_mean), method = "lm") +
  facet_wrap(~Survey_Name_Season, scales= "free", ncol = 5) +
  theme_classic())


ggsave(preds_sbt_mean_temp_survey_facet, path  = here::here("figures","predictor_variables"), filename = "preds_sbt_mean_temp_survey_facet.jpg", height = 13, width = 10, unit = "in")

```

#SKIP FOR NOW
Plot just dissimilarity ~ raw temp coefficients of each region (Like figure 1b)
```{r}
#unique survey unit year combos
survey_unit_sampling_years <- unique(BC_total_model_coefs_reduced_length.unique[,.(survey_unit, years_sampled)]) #unique survey unit year combos
                                    

 # see group coefficients and confidence intervals
temp_model_coefs_reduced <- data.table(transform(as.data.frame(ranef(dissimilarity_BC_total_sbt_mean_mod)), lwr = condval - 1.96*condsd, upr = condval + 1.96*condsd))
#https://stackoverflow.com/questions/69805532/extract-the-confidence-intervals-of-lm-random-effects-plotted-with-dotplotra


#ONLY SLOPES
temp_model_coefs_reduced <- temp_model_coefs_reduced[term == "yearly_mean_bypoint_avg",]

temp_model_coefs_reduced[,survey_unit := grp][,yearly_mean_bypoint_avg := condval]

#merge with duration of survey
temp_model_coefs_reduced_length <- temp_model_coefs_reduced[survey_unit_sampling_years, on = "survey_unit"]


temp_model_coefs_reduced_length[,years_sampled := as.numeric(years_sampled)][,Slope_Direction := ifelse(yearly_mean_bypoint_avg > 0, "Positive","Negative")]

#
temp_model_coefs_reduced_length <- temp_model_coefs_reduced_length[color_link, on = "survey_unit"]


#does it cross zero?
temp_model_coefs_reduced_length[,significant := ifelse(lwr >0 & upr>0,T,ifelse(lwr<0 & upr<0,T,F))]

#delete all obs that are significant
temp_model_coefs_reduced_length.r <- temp_model_coefs_reduced_length[significant == F,]

#order table by coefficient
setorder(temp_model_coefs_reduced_length, yearly_mean_bypoint_avg)

BC_total_temp_model_coefs_reduced_length.unique <- unique(temp_model_coefs_reduced_length[,.(condval,condsd, lwr, upr, survey_unit, yearly_mean_bypoint_avg, years_sampled, Slope_Direction, hex, Survey_Name_Season, significant)]) 

#extract color hexes
#year adj coef order
color_year_adj_order <-BC_total_temp_model_coefs_reduced_length.unique[,hex]

#alphabetical order
BC_total_temp_model_coefs_reduced_length.unique.alpha <- setorder(BC_total_temp_model_coefs_reduced_length.unique, Survey_Name_Season)
color_alpha_order <- BC_total_temp_model_coefs_reduced_length.unique.alpha[,hex]

saveRDS(BC_total_temp_model_coefs_reduced_length.unique, here::here("output","region_stats","BC_total_temp_model_coefs_reduced_length.unique.Rds"))
```

Bar Plot Coefficient Bottom Temperature lm
```{r bar plot of bottom temperature coefficients}
BC_TOTAL_Dissimilarity_Coef_BOTTOMTEMP_errorbar_reduced <- ggplot() +
    geom_errorbar(data = temp_model_coefs_reduced_length, aes(x = reorder(Survey_Name_Season, yearly_mean_bypoint_avg) , y = yearly_mean_bypoint_avg, label = Survey_Name_Season, ymin = lwr, ymax = upr), fill = "grey", width = 0) + #add confidence intervals
  geom_point(data = temp_model_coefs_reduced_length, aes(x = reorder(Survey_Name_Season, yearly_mean_bypoint_avg) , y = yearly_mean_bypoint_avg, label = Survey_Name_Season, 
        #size = years_sampled,
        fill = Slope_Direction), stat = 'identity', shape = 21, color = "black") +
  scale_fill_manual(values = c("white","black"), name = "Slope Direction") +
  geom_point(data = temp_model_coefs_reduced_length.r, aes(x = reorder(Survey_Name_Season, yearly_mean_bypoint_avg) , y = yearly_mean_bypoint_avg, 
          #size = years_sampled,
          label = Survey_Name_Season), stat = 'identity', fill = "grey",color = "grey", shape = 21) +
 # scale_size_binned(range = c(1,8), name = "Survey period\nlength") +
  geom_hline(yintercept = 0) +
  xlab("Survey unit") +
  ylab("Total BC dissimilarity ~\nraw bottom temperature") +
  coord_flip() +
  theme_classic()

#+
#  theme(axis.text.y = element_text(colour = color_yearly_mean_bypoint_avg_order, face = "bold"), axis.title.y = element_blank())

BC_TOTAL_Dissimilarity_Coef_BOTTOMTEMP_errorbar_reduced

#save as independent grob to merge with fishing
saveRDS(BC_TOTAL_Dissimilarity_Coef_BOTTOMTEMP_errorbar_reduced, here::here("figures","BC_TOTAL_Dissimilarity_Coef_BOTTOMTEMP_errorbar_reduced.Rds"))
```

Scaled: dissimilarity_BC_total_sbt_maxSD_scaled_mod, dissimilarity_BC_total_sbt_minSD_scaled_mod
```{r mean scaled sbt}

#for all regions
(dissimilarity_BC_total_sbt_maxSD.s.final <- ggplot(data = dissimilarities_temp) +
  labs(x = "Relative mean bottom temperature",  y = "Bray Curtis total dissimilarity") +
  geom_point(aes(x = yearly_max_bypoint_SD.s, y = bray_curtis_dissimilarity_total_mean), alpha = 0.3) +
  geom_smooth(aes(x = yearly_max_bypoint_SD.s, y = bray_curtis_dissimilarity_total_mean), method = "lm") +
  theme_classic())

ggsave(dissimilarity_BC_total_sbt_maxSD.s.final, path  = here::here("figures","predictor_variables"), filename = "dissimilarity_BC_total_sbt_maxSD.s.final.jpg", height = 4, width = 6.5, unit = "in") 

dissimilarity_BC_total_sbt_maxSD_facet.s.final <- ggplot(data = dissimilarities_temp) +
  labs(x = "Scaled SD of maximum bottom temperature",  y = "Bray Curtis total dissimilarity") +
  geom_point(aes(x = yearly_max_bypoint_SD.s, y = bray_curtis_dissimilarity_total_mean), alpha = 0.3) +
  geom_smooth(aes(x = yearly_max_bypoint_SD.s, y = bray_curtis_dissimilarity_total_mean), method = "lm") +
  facet_wrap(~Survey_Name_Season, scales = "free", ncol = 5) +
  theme_classic()

ggsave(dissimilarity_BC_total_sbt_maxSD_facet.s.final, path  = here::here("figures","predictor_variables"), filename = "dissimilarity_BC_total_sbt_maxSD_facet.s.final.jpg", height = 13, width = 10, unit = "in")

```

```{r SD of min temp scaled}

#for all regions
(dissimilarity_BC_total_sbt_min.SD.s <- ggplot(data = dissimilarities_temp) +
  labs(x = "Relative SD of minimum temperature",  y = "Bray Curtis total dissimilarity") +
  geom_point(aes(x = yearly_min_bypoint_SD.s, y = bray_curtis_dissimilarity_total_mean), alpha = 0.3) +
  geom_smooth(aes(x = yearly_min_bypoint_SD.s, y = bray_curtis_dissimilarity_total_mean), method = "lm") +
  theme_classic())

ggsave(dissimilarity_BC_total_sbt_min.SD.s, path  = here::here("figures","predictor_variables"), filename = "dissimilarity_BC_total_sbt_min.SD.s.jpg", height = 4, width = 6.5, unit = "in") 

dissimilarity_BC_total_sbt_min.SD_facet.s.final <- ggplot(data = dissimilarities_temp) +
  labs(x = "Relative SD of minimum temperature",  y = "Bray Curtis total dissimilarity") +
  geom_point(aes(x = yearly_min_bypoint_SD.s, y = bray_curtis_dissimilarity_total_mean), alpha = 0.3) +
  geom_smooth(aes(x = yearly_min_bypoint_SD.s, y = bray_curtis_dissimilarity_total_mean), method = "lm") +
  facet_wrap(~Survey_Name_Season, scales = "free", ncol = 5) +
  theme_classic()

ggsave(dissimilarity_BC_total_sbt_min.SD_facet.s.final, path  = here::here("figures","predictor_variables"), filename = "dissimilarity_BC_total_sbt_min.SD_facet.s.final.jpg", height = 13, width = 10, unit = "in")

```
Plot mean temp for helpful visual
```{r}

#for all regions
(dissimilarity_BC_total_sbt_mean_avg <- ggplot(data = dissimilarities_temp) +
  labs(x = "Mean bottom temperature (˚C)",  y = "Bray Curtis total dissimilarity") +
  geom_point(aes(x = yearly_mean_bypoint_avg, y = bray_curtis_dissimilarity_total_mean), alpha = 0.3) +
  geom_smooth(aes(x = yearly_mean_bypoint_avg, y = bray_curtis_dissimilarity_total_mean), method = "lm") +
  theme_classic())

ggsave(dissimilarity_BC_total_sbt_mean_avg, path  = here::here("figures","predictor_variables"), filename = "dissimilarity_BC_total_sbt_mean_avg.jpg", height = 4, width = 6.5, unit = "in") 

(dissimilarity_BC_total_sbt_mean.final_facet <- ggplot(data = dissimilarities_temp) +
  labs(x = "Mean bottom temperature (˚C)",  y = "Bray Curtis total dissimilarity") +
  geom_point(aes(x = as.numeric(yearly_mean_bypoint_avg), y = bray_curtis_dissimilarity_total_mean), alpha = 0.3) +
  geom_smooth(aes(x = as.numeric(yearly_mean_bypoint_avg), y = bray_curtis_dissimilarity_total_mean), method = "lm") +
  facet_wrap(~Survey_Name_Season, scales= "free", ncol = 5) +
  theme_classic())


ggsave(dissimilarity_BC_total_sbt_mean.final_facet, path  = here::here("figures","predictor_variables"), filename = "dissimilarity_BC_total_sbt_mean.final_facet.jpg", height = 13, width = 10, unit = "in")

```

#SKIP FOR NOW
Plot just dissimilarity ~ scaled temp coefficients of each region (Like figure 1b)
```{r}

 # see group coefficients and confidence intervals
temp_scaled_model_coefs_reduced <- data.table(transform(as.data.frame(ranef(dissimilarity_BC_total_sbt_mean_scaled_mod)), lwr = condval - 1.96*condsd, upr = condval + 1.96*condsd))
#https://stackoverflow.com/questions/69805532/extract-the-confidence-intervals-of-lm-random-effects-plotted-with-dotplotra


#ONLY SLOPES
temp_scaled_model_coefs_reduced <- temp_scaled_model_coefs_reduced[term == "yearly_mean_bypoint_avg.s",]

temp_scaled_model_coefs_reduced[,survey_unit := grp][,yearly_mean_bypoint_avg.s := condval]

#merge with duration of survey
temp_scaled_model_coefs_reduced_length <- temp_scaled_model_coefs_reduced[survey_unit_sampling_years, on = "survey_unit"]


temp_scaled_model_coefs_reduced_length[,years_sampled := as.numeric(years_sampled)][,Slope_Direction := ifelse(yearly_mean_bypoint_avg.s > 0, "Positive","Negative")]

#
temp_scaled_model_coefs_reduced_length <- temp_scaled_model_coefs_reduced_length[color_link, on = "survey_unit"]


#does it cross zero?
temp_scaled_model_coefs_reduced_length[,significant := ifelse(lwr >0 & upr>0,T,ifelse(lwr<0 & upr<0,T,F))]

#delete all obs that are significant
temp_scaled_model_coefs_reduced_length.r <- temp_scaled_model_coefs_reduced_length[significant == F,]

#order table by coefficient
setorder(temp_scaled_model_coefs_reduced_length, yearly_mean_bypoint_avg.s)

BC_total_temp_scaled_model_coefs_reduced_length.unique <- unique(temp_scaled_model_coefs_reduced_length[,.(condval,condsd, lwr, upr, survey_unit, yearly_mean_bypoint_avg.s, years_sampled, Slope_Direction, hex, Survey_Name_Season, significant)]) 

#extract color hexes
#year adj coef order
color_year_adj_order <-BC_total_temp_scaled_model_coefs_reduced_length.unique[,hex]

#alphabetical order
BC_total_temp_scaled_model_coefs_reduced_length.unique.alpha <- setorder(BC_total_temp_scaled_model_coefs_reduced_length.unique, Survey_Name_Season)
color_alpha_order <- BC_total_temp_scaled_model_coefs_reduced_length.unique.alpha[,hex]

saveRDS(BC_total_temp_scaled_model_coefs_reduced_length.unique, here::here("output","region_stats","BC_total_temp_scaled_model_coefs_reduced_length.unique.Rds"))
```

Bar Plot Coefficient Scaled Bottom Temperature lm
```{r bar plot of scaled bottom temperature coefficients}
BC_TOTAL_Dissimilarity_Coef_scaled_BOTTOMTEMP_errorbar_reduced <- ggplot() +
    geom_errorbar(data = temp_scaled_model_coefs_reduced_length, aes(x = reorder(Survey_Name_Season, yearly_mean_bypoint_avg.s) , y = yearly_mean_bypoint_avg.s, label = Survey_Name_Season, ymin = lwr, ymax = upr), fill = "grey", width = 0) + #add confidence intervals
  geom_point(data = temp_scaled_model_coefs_reduced_length, aes(x = reorder(Survey_Name_Season, yearly_mean_bypoint_avg.s) , y = yearly_mean_bypoint_avg.s, label = Survey_Name_Season, 
                #size = years_sampled,
                fill = Slope_Direction), stat = 'identity', shape = 21, color = "black") +
  scale_fill_manual(values = c("white","black"), name = "Slope Direction") +
  geom_point(data = temp_scaled_model_coefs_reduced_length.r, aes(x = reorder(Survey_Name_Season, yearly_mean_bypoint_avg.s) , y = yearly_mean_bypoint_avg.s, 
                  #size = years_sampled,
                  label = Survey_Name_Season), stat = 'identity', fill = "grey",color = "grey", shape = 21) +
#  scale_size_binned(range = c(1,8), name = "Survey period\nlength") +
  geom_hline(yintercept = 0) +
  xlab("Survey unit") +
  ylab("Total BC dissimilarity ~\nrelative bottom temperature") +
  coord_flip() +
  theme_classic()

#+
#  theme(axis.text.y = element_text(colour = color_yearly_mean_bypoint_avg.s_order, face = "bold"), axis.title.y = element_blank())

BC_TOTAL_Dissimilarity_Coef_scaled_BOTTOMTEMP_errorbar_reduced

#save as independent grob to merge with fishing
saveRDS(BC_TOTAL_Dissimilarity_Coef_scaled_BOTTOMTEMP_errorbar_reduced, here::here("figures","BC_TOTAL_Dissimilarity_Coef_scaled_BOTTOMTEMP_errorbar_reduced.Rds"))
```

Merge lollypops
```{r merge lollypops}
BC_TOTAL_Dissimilarity_Coef_BOTTOMTEMP_errorbar_reduced_merge <- cowplot::plot_grid(BC_TOTAL_Dissimilarity_Coef_BOTTOMTEMP_errorbar_reduced + theme(legend.position = "null", axis.title.y = element_blank()), BC_TOTAL_Dissimilarity_Coef_scaled_BOTTOMTEMP_errorbar_reduced + theme(legend.position = "null", axis.title.y = element_blank()), cowplot::get_legend(BC_TOTAL_Dissimilarity_Coef_scaled_BOTTOMTEMP_errorbar_reduced), nrow = 1, ncol = 3, rel_widths = c(1,1,0.2), rel_heights = c(1,1,0.2))

ggsave(BC_TOTAL_Dissimilarity_Coef_BOTTOMTEMP_errorbar_reduced_merge, path = "figures", filename = "BC_TOTAL_Dissimilarity_Coef_BOTTOMTEMP_errorbar_reduced_merge.jpg", height = 5, width = 12, units = "in")
```

Merge
```{r}
library(cowplot)
(dissimilarity_BC_total_sbt_mean.final.SIMPLIFIED_merge <- plot_grid(dissimilarity_BC_total_sbt_mean.finalSIMPLIFIED, dissimilarity_BC_total_sbt_mean.s.final.SIMPLIFIED + theme(axis.title.y = element_blank(), axis.text.y = element_blank()), nrow = 1, ncol = 2, labels = c("a.","b."), hjust = -1))

ggsave(dissimilarity_BC_total_sbt_mean.final.SIMPLIFIED_merge, path  = here::here("figures","predictor_variables"), filename = "dissimilarity_BC_total_sbt_mean.final.SIMPLIFIED_merge.jpg", height = 4, width = 9, unit = "in") 
```