---
title: "Bray Balanced Dissimilarity Bottom Temperature & Fishing Interactions for All Regions"
output: html_notebook
---

Here we will link temperature and fishing to dissimilarity

This code is Script X for Kitchel et al. TITLE manuscript.

- This project is a collaborative effort to describe changes in taxonomic composition  of fish communities around the world--as sampled by bottom trawl surveys.

- Code by Zoë J. Kitchel

SESSION INFO TO DO

##Here, we examine bottom temperature and fishing pressure to predict annual BC balanced dissimilarity for all survey units


```{r setup}

library(data.table)
library(ggplot2)
library(lme4)
library(lmerTest)
library(MuMIn)


SODA_data_temp_avgs_full <- readRDS(here::here("data","Temperature","SODA_data_temp_avgs_full.rds"))
#don't feel confident about SODA temp before 1980 right now, start then

SODA_data_temp_avgs_full <- SODA_data_temp_avgs_full[year_for_avg >= 1980,]

SODA_data_temp_avgs_full[,year := year_for_avg]


#pull in average distance decay values
distances_dissimilarities_allyears.r <- readRDS(here::here("output","distance_decay","distances_dissimilarities_allyears.r.rds"))

#pull in fishing data
SAU_summed_tonnes.r <- readRDS(here::here("data","sea_around_us","SAU_summed_tonnes.rds"))

#pull in regional statistics
region_stats <- readRDS(here::here("output","region_stats", "region_stats.rds"))


```

Put into one data table
```{r link temp and fishing to avg values}
#temp
distances_dissimilarities_allyears.r.temp <- SODA_data_temp_avgs_full[distances_dissimilarities_allyears.r, on = c("year","survey_unit")]

#fishing
dissimilarities_temp_fishing <- SAU_summed_tonnes.r[distances_dissimilarities_allyears.r.temp, on = c("year", "survey_unit")]

#region_statistics
dissimilarities_temp_fishing_regstats <- dissimilarities_temp_fishing[region_stats, on = "survey_unit"]
dissimilarities_temp_fishing_regstats[,spp_num.s := scale(spp_num)]

dissimilarities_temp_fishing_regstats[,bray_curtis_dissimilarity_balanced_mean.s := scale(bray_curtis_dissimilarity_balanced_mean, center = F),survey_unit]

#save avg values
saveRDS(dissimilarities_temp_fishing_regstats, here::here("output","distance_decay","dissimilarities_temp_fishing_regstats.rds"))
```

Color pallete for survey units
```{r}
pal_37 <- c(
  "#5A5156", #AI
  "#F6222E", #CHL
  "#F8A19F", #DFO-NF
  "#16FF32", #DFO-QCS
  "#DF00DB", #BITS-1
  "#DB8EDA", #BITS-4
  "#325A9B", #EBS
  "#3283FE", #EVHOE
  "#FEAF16", #FR-CGFS
  "#1C8356", #GMEX-Summer
  "#C4451C", #GOA
  "#85660D", #GRL-DE
  "#B0009F", #GSL-N
  "#BF79B8", #GSL-S
  "#1CBE4F", #ICE-GFS
  "#782AB6", #IE-IGFS
  "#90AD1C", #MEDITS
  "#6B003A", #NAM
  "#A75B00", #NEUS-Fall
  "#E3B072", #NEUS-Spring
  "#02E8B6", #NIGFS-1
  "#97E7D5", #NIGFS-4
  "#B00068", #Nor-BTS-3
  "#00B9E3", #NS-IBTS-1
  "#95E2F4", #NS-IBTS-3
  "#B3CE73", #NZ-CHAT
  "#689500", #NZ-ECSI
  "#AAF400", #NZ-WCSI
  "#AA0DFE", #PT-IBTS
  "#FA0087", #S-GEORG
  "#DEA0FD", #SCS-Summer
  "#FCEF88", #SEUS-fall
  "#A59405", #SEUS-spring
  "#FCE100", #SEUS-summer
  "#C075A6", #WCANN
  "#BDCDFF", #ZAF-ATL
  "#003EFF")  #ZAF-IND
```


Table that will be populated with model results
```{r}
#depth, scaled, temp var, coefficient, p-value, marg_r^2, cond_r^2, AICc
balanced_dissimilarity_sbt_fishing_model_results <- data.table(Depth = as.character(),
                                                       Scaled = as.character(),
                                                       `Temp variable` = as.character(),
                                                       `Fishing`= as.character(),
                                                       `Temp fishing interaction` = as.character(),
                                                       `Mixed effect` = as.character(),
                                                       Intercept = as.numeric(),
                                                       `Temp coefficient` = as.numeric(),
                                                       `Coefficient p-value 1` = as.numeric(),
                                                       `Fishing coefficient` = as.numeric(),
                                                       `Coefficient p-value 2` = as.numeric(),
                                                       `Temp*fishing coefficient` = as.numeric(),
                                                       `Coefficient p-value 3` = as.numeric(),
                                                       `Marginal R-squared` = as.numeric(),
                                                       `Conditional R-squared` = as.numeric(),
                                                       AICc = as.numeric(),
                                                       `Degrees of freedom`=as.numeric())
                                                  
```

###How well does fishing alone predict mean annual BC balanced dissimilarity?
(Note this code is identical to dissimilarity_BC_balanced_temp_fishing_link_allregs.nb_OISST.Rmd between here and ~line 249, but I've kept it in to allow easier model comparisons. Edits will need to be made in both places)
```{r build fishing mod}

#random slope and intercept for survey
allreg_dissimilarity_fishing_mean_mod <- lmer(bray_curtis_dissimilarity_balanced_mean ~ summed_tonnes_scaled_byreg + (1 + summed_tonnes_scaled_byreg|survey_unit), data = dissimilarities_temp_fishing_regstats)

coefs <- data.frame(coef(summary(allreg_dissimilarity_fishing_mean_mod)))

balanced_dissimilarity_sbt_fishing_model_results_row <- data.table(matrix(c(NA, #depth
                                                            NA, #scaled
                                                            NA, #temp var
                                                            T, #fishing
                                                            NA, #temp fishing interaction
                                                            "survey_slope_intercept", #mixed effect
                                                            round(coefs[1,1],3), #intercept
                                                            NA,#temp coef
                                                            NA,#coef p value
                                                            round(coefs[2,1],3),#fishing coef
                                                            round(coefs[2,5],3),#coef p value
                                                            NA, #interact coef
                                                            NA, #coef p value
                                                round(r.squaredGLMM(allreg_dissimilarity_fishing_mean_mod)[[1]],3), #marginal r squared
                                                round(r.squaredGLMM(allreg_dissimilarity_fishing_mean_mod)[[2]],3), #conditional r-squared
                                                            round(AICc(allreg_dissimilarity_fishing_mean_mod),3),#AICc
                                                            6 #degrees of freedom 
                                                            ),nrow =1))

balanced_dissimilarity_sbt_fishing_model_results <- rbind(balanced_dissimilarity_sbt_fishing_model_results, balanced_dissimilarity_sbt_fishing_model_results_row, use.names = F)

#random intercept only
allreg_dissimilarity_fishing_mean_mod.intercept <- lmer(bray_curtis_dissimilarity_balanced_mean ~ summed_tonnes_scaled_byreg + (1|survey_unit), data = dissimilarities_temp_fishing_regstats)

coefs <- data.frame(coef(summary(allreg_dissimilarity_fishing_mean_mod.intercept)))

balanced_dissimilarity_sbt_fishing_model_results_row <- data.table(matrix(c(NA, #depth
                                                            NA, #scaled
                                                            NA, #temp var
                                                            T, #fishing
                                                            NA, #temp fishing interaction
                                                            "survey_intercept", #mixed effect
                                                            round(coefs[1,1],3), #intercept
                                                            NA,#temp coef
                                                            NA,#coef p value
                                                            round(coefs[2,1],3),#fishing coef
                                                            round(coefs[2,5],3),#coef p value
                                                            NA, #interact coef
                                                            NA, #coef p value
                                                round(r.squaredGLMM(allreg_dissimilarity_fishing_mean_mod.intercept)[[1]],3), #marginal r squared
                                                round(r.squaredGLMM(allreg_dissimilarity_fishing_mean_mod.intercept)[[2]],3), #conditional r-squared
                                                            round(AICc(allreg_dissimilarity_fishing_mean_mod.intercept),3),#AICc
                                                            4 #degrees of freedom 
                                                            ),nrow =1))

balanced_dissimilarity_sbt_fishing_model_results <- rbind(balanced_dissimilarity_sbt_fishing_model_results, balanced_dissimilarity_sbt_fishing_model_results_row, use.names = F)

#fixed effect for survey
allreg_dissimilarity_fishing_fixed_mean_mod <- lm(bray_curtis_dissimilarity_balanced_mean ~ summed_tonnes_scaled_byreg*survey_unit, data = dissimilarities_temp_fishing_regstats)

coefs <- data.frame(coef(summary(allreg_dissimilarity_fishing_fixed_mean_mod)))

balanced_dissimilarity_sbt_fishing_model_results_row <- data.table(matrix(c(NA, #depth
                                                            NA, #scaled
                                                            NA, #temp var
                                                            T, #fishing
                                                            NA, #temp fishing interaction
                                                            "survey_fixed", #mixed effect
                                                            round(coefs[1,1],3), #intercept
                                                            NA,#temp coef
                                                            NA,#coef p value
                                                            round(coefs[2,1],3),#fishing coef
                                                            round(coefs[2,4],3),#coef p value
                                                            NA, #interact coef
                                                            NA, #coef p value
                                                round(r.squaredGLMM(allreg_dissimilarity_fishing_fixed_mean_mod)[[1]],3), #marginal r squared
                                                round(r.squaredGLMM(allreg_dissimilarity_fishing_fixed_mean_mod)[[2]],3), #conditional r-squared
                                                            round(AICc(allreg_dissimilarity_fishing_fixed_mean_mod),3),#AICc
                                                            75 #degrees of freedom 
                                                            ),nrow =1))

balanced_dissimilarity_sbt_fishing_model_results <- rbind(balanced_dissimilarity_sbt_fishing_model_results, balanced_dissimilarity_sbt_fishing_model_results_row, use.names = F)

#no effect for survey
allreg_dissimilarity_fishing_mean_mod_lm <- lm(bray_curtis_dissimilarity_balanced_mean ~ summed_tonnes_scaled_byreg, data = dissimilarities_temp_fishing_regstats)

coefs <- data.frame(coef(summary(allreg_dissimilarity_fishing_mean_mod_lm)))

balanced_dissimilarity_sbt_fishing_model_results_row <- data.table(matrix(c(NA, #depth
                                                            NA, #scaled
                                                            NA, #temp var
                                                            T, #fishing
                                                            NA, #temp fishing interaction
                                                            "exclude_survey_completely", #mixed effect
                                                            round(coefs[1,1],3), #intercept
                                                            NA,#temp coef
                                                            NA,#coef p value
                                                            round(coefs[2,1],3),#fishing coef
                                                            round(coefs[2,4],3),#coef p value
                                                            NA, #interact coef
                                                            NA, #coef p value
                                                round(r.squaredGLMM(allreg_dissimilarity_fishing_mean_mod_lm)[[1]],3), #marginal r squared
                                                round(r.squaredGLMM(allreg_dissimilarity_fishing_mean_mod_lm)[[2]],3), #conditional r-squared
                                                            round(AICc(allreg_dissimilarity_fishing_mean_mod_lm),3),#AICc
                                                            3 #degrees of freedom 
                                                            ),nrow =1))

balanced_dissimilarity_sbt_fishing_model_results <- rbind(balanced_dissimilarity_sbt_fishing_model_results, balanced_dissimilarity_sbt_fishing_model_results_row, use.names = F)

#In years of higher than average fishing pressure, dissimilarity is lower (homogenization)

AICc(allreg_dissimilarity_fishing_mean_mod, allreg_dissimilarity_fishing_mean_mod.intercept, allreg_dissimilarity_fishing_fixed_mean_mod, allreg_dissimilarity_fishing_mean_mod_lm)


View(balanced_dissimilarity_sbt_fishing_model_results)


```
Best model includes survey unit as a fixed effect (R^2 = 93% of variation), but we don't care much about survey_unit (beyond acknowledging it makes a difference), so we will move forward with random intercept and random slope for survey unit.

As relative fishing pressure increases, dissimilarity decreases (more fished communities = more homogenized communities). 

###Temperature x fishing models
Best temp variables to carry forward: 

For raw temp (indicative of survey vs. survey):
-dissimilarity_BC_balanced_sbt_mean_mod


For scaled temp (indicative of within survey: cool or warm year? heterogenous or homogenous year)
-dissimilarity_BC_balanced_sbt_mean_scaled_mod (SD a bit better, but for comparison sake, we go with 2nd best = mean)

Starting with survey unit as random slope~fishing pressure and intercept (because it performed best above)
```{r temp and fishing models of dissimilarity}
#raw temp values

allreg_dissimilarity_fishing_mean_mod_sbt_mean_avg <- lmer(bray_curtis_dissimilarity_balanced_mean ~ summed_tonnes_scaled_byreg*yearly_mean_bypoint_avg + (1 + summed_tonnes_scaled_byreg|survey_unit), data = dissimilarities_temp_fishing_regstats)


        coefs <- data.frame(coef(summary(allreg_dissimilarity_fishing_mean_mod_sbt_mean_avg)))

balanced_dissimilarity_sbt_fishing_model_results_row <- data.table(matrix(c("sbt", #depth
                                                            F, #scaled
                                                            "mean", #temp var
                                                            T, #fishing
                                                            T, #temp fishing interaction
                                                            "survey_intercept", #mixed effect
                                                            round(coefs[1,1],3), #intercept
                                                            round(coefs[3,1],3),#temp coef
                                                            round(coefs[3,5],3),#coef p value
                                                            round(coefs[2,1],3),#fishing coef
                                                            round(coefs[2,5],3),#coef p value
                                                            round(coefs[4,1],3), #interact coef
                                                            round(coefs[4,5],3), #coef p value
                                                round(r.squaredGLMM(allreg_dissimilarity_fishing_mean_mod_sbt_mean_avg)[[1]],3), #marginal r squared
                                                round(r.squaredGLMM(allreg_dissimilarity_fishing_mean_mod_sbt_mean_avg)[[2]],3), #conditional r-squared
                                                            round(AICc(allreg_dissimilarity_fishing_mean_mod_sbt_mean_avg),3),#AICc
                                                            6 #degrees of freedom 
                                                            ),nrow =1))

balanced_dissimilarity_sbt_fishing_model_results <- rbind(balanced_dissimilarity_sbt_fishing_model_results, balanced_dissimilarity_sbt_fishing_model_results_row, use.names = F)

  #raw temp, no interaction
  allreg_dissimilarity_fishing_mean_mod_sbt_mean_avg.nointeract <- lmer(bray_curtis_dissimilarity_balanced_mean ~ summed_tonnes_scaled_byreg+yearly_mean_bypoint_avg + (1 + summed_tonnes_scaled_byreg|survey_unit), data = dissimilarities_temp_fishing_regstats)

  
          coefs <- data.frame(coef(summary(allreg_dissimilarity_fishing_mean_mod_sbt_mean_avg.nointeract)))

balanced_dissimilarity_sbt_fishing_model_results_row <- data.table(matrix(c("sbt", #depth
                                                            F, #scaled
                                                            "mean", #temp var
                                                            T, #fishing
                                                            F, #temp fishing interaction
                                                            "survey_intercept", #mixed effect
                                                            round(coefs[1,1],3), #intercept
                                                            round(coefs[3,1],3),#temp coef
                                                            round(coefs[3,5],3),#coef p value
                                                            round(coefs[2,1],3),#fishing coef
                                                            round(coefs[2,5],3),#coef p value
                                                            NA, #interact coef
                                                            NA, #coef p value
                                                round(r.squaredGLMM(allreg_dissimilarity_fishing_mean_mod_sbt_mean_avg.nointeract)[[1]],3), #marginal r squared
                                                round(r.squaredGLMM(allreg_dissimilarity_fishing_mean_mod_sbt_mean_avg.nointeract)[[2]],3), #conditional r-squared
                                                            round(AICc(allreg_dissimilarity_fishing_mean_mod_sbt_mean_avg.nointeract),3),#AICc
                                                            5 #degrees of freedom 
                                                            ),nrow =1))

balanced_dissimilarity_sbt_fishing_model_results <- rbind(balanced_dissimilarity_sbt_fishing_model_results, balanced_dissimilarity_sbt_fishing_model_results_row, use.names = F)
    
  #temp only
  allreg_dissimilarity_mod_sbt_mean_avg <- lmer(bray_curtis_dissimilarity_balanced_mean ~ yearly_mean_bypoint_avg + (1 + summed_tonnes_scaled_byreg|survey_unit), data = dissimilarities_temp_fishing_regstats)

  
               coefs <- data.frame(coef(summary(allreg_dissimilarity_mod_sbt_mean_avg)))

balanced_dissimilarity_sbt_fishing_model_results_row <- data.table(matrix(c("sbt", #depth
                                                            F, #scaled
                                                            "mean", #temp var
                                                            F, #fishing
                                                            F, #temp fishing interaction
                                                            "survey_intercept", #mixed effect
                                                            round(coefs[1,1],3), #intercept
                                                            round(coefs[2,1],3),#temp coef
                                                            round(coefs[2,5],3),#coef p value
                                                            NA,#fishing coef
                                                            NA,#coef p value
                                                            NA, #interact coef
                                                            NA, #coef p value
                                                round(r.squaredGLMM(allreg_dissimilarity_mod_sbt_mean_avg)[[1]],3), #marginal r squared
                                                round(r.squaredGLMM(allreg_dissimilarity_mod_sbt_mean_avg)[[2]],3), #conditional r-squared
                                                            round(AICc(allreg_dissimilarity_mod_sbt_mean_avg),3),#AICc
                                                            4 #degrees of freedom 
                                                            ),nrow =1))

balanced_dissimilarity_sbt_fishing_model_results <- rbind(balanced_dissimilarity_sbt_fishing_model_results, balanced_dissimilarity_sbt_fishing_model_results_row, use.names = F)
    
#scaled temp values
allreg_dissimilarity_fishing_mean_mod_sbt_mean_avg.s <- lmer(bray_curtis_dissimilarity_balanced_mean ~ summed_tonnes_scaled_byreg*yearly_mean_bypoint_avg.s + (1 + summed_tonnes_scaled_byreg|survey_unit), data = dissimilarities_temp_fishing_regstats)


                  coefs <- data.frame(coef(summary(allreg_dissimilarity_fishing_mean_mod_sbt_mean_avg.s)))

balanced_dissimilarity_sbt_fishing_model_results_row <- data.table(matrix(c("sbt", #depth
                                                            T, #scaled
                                                            "mean", #temp var
                                                            T, #fishing
                                                            T, #temp fishing interaction
                                                            "survey_intercept", #mixed effect
                                                            round(coefs[1,1],3), #intercept
                                                            round(coefs[3,1],3),#temp coef
                                                            round(coefs[3,5],3),#coef p value
                                                            round(coefs[2,1],3),#fishing coef
                                                            round(coefs[2,5],3),#coef p value
                                                            round(coefs[4,1],3), #interact coef
                                                            round(coefs[4,5],3), #coef p value
                                                round(r.squaredGLMM(allreg_dissimilarity_fishing_mean_mod_sbt_mean_avg.s)[[1]],3), #marginal r squared
                                                round(r.squaredGLMM(allreg_dissimilarity_fishing_mean_mod_sbt_mean_avg.s)[[2]],3), #conditional r-squared
                                                            round(AICc(allreg_dissimilarity_fishing_mean_mod_sbt_mean_avg.s),3),#AICc
                                                            6 #degrees of freedom 
                                                            ),nrow =1))

balanced_dissimilarity_sbt_fishing_model_results <- rbind(balanced_dissimilarity_sbt_fishing_model_results, balanced_dissimilarity_sbt_fishing_model_results_row, use.names = F)

  #scaled temp, no interaction
  allreg_dissimilarity_fishing_mean_mod_sbt_mean_avg.s.nointeract <- lmer(bray_curtis_dissimilarity_balanced_mean ~ summed_tonnes_scaled_byreg+yearly_mean_bypoint_avg.s + (1 + summed_tonnes_scaled_byreg|survey_unit), data = dissimilarities_temp_fishing_regstats)

  
          coefs <- data.frame(coef(summary(allreg_dissimilarity_fishing_mean_mod_sbt_mean_avg.s.nointeract)))

balanced_dissimilarity_sbt_fishing_model_results_row <- data.table(matrix(c("sbt", #depth
                                                            T, #scaled
                                                            "mean", #temp var
                                                            T, #fishing
                                                            F, #temp fishing interaction
                                                            "survey_intercept", #mixed effect
                                                            round(coefs[1,1],3), #intercept
                                                            round(coefs[3,1],3),#temp coef
                                                            round(coefs[3,5],3),#coef p value
                                                            round(coefs[2,1],3),#fishing coef
                                                            round(coefs[2,5],3),#coef p value
                                                            NA, #interact coef
                                                            NA, #coef p value
                                                round(r.squaredGLMM(allreg_dissimilarity_fishing_mean_mod_sbt_mean_avg.s.nointeract)[[1]],3), #marginal r squared
                                                round(r.squaredGLMM(allreg_dissimilarity_fishing_mean_mod_sbt_mean_avg.s.nointeract)[[2]],3), #conditional r-squared
                                                            round(AICc(allreg_dissimilarity_fishing_mean_mod_sbt_mean_avg.s.nointeract),3),#AICc
                                                            5 #degrees of freedom 
                                                            ),nrow =1))

balanced_dissimilarity_sbt_fishing_model_results <- rbind(balanced_dissimilarity_sbt_fishing_model_results, balanced_dissimilarity_sbt_fishing_model_results_row, use.names = F)
    
  #temp only
  allreg_dissimilarity_mod_sbt_mean_avg.s <- lmer(bray_curtis_dissimilarity_balanced_mean ~ summed_tonnes_scaled_byreg+yearly_mean_bypoint_avg.s + (1 + summed_tonnes_scaled_byreg|survey_unit), data = dissimilarities_temp_fishing_regstats)

  
          coefs <- data.frame(coef(summary(allreg_dissimilarity_mod_sbt_mean_avg.s)))

balanced_dissimilarity_sbt_fishing_model_results_row <- data.table(matrix(c("sbt", #depth
                                                            T, #scaled
                                                            "mean", #temp var
                                                            T, #fishing
                                                            F, #temp fishing interaction
                                                            "survey_intercept", #mixed effect
                                                            round(coefs[1,1],3), #intercept
                                                            round(coefs[3,1],3),#temp coef
                                                            round(coefs[3,5],3),#coef p value
                                                            NA,#fishing coef
                                                            NA,#coef p value
                                                            NA, #interact coef
                                                            NA, #coef p value
                                                round(r.squaredGLMM(allreg_dissimilarity_mod_sbt_mean_avg.s)[[1]],3), #marginal r squared
                                                round(r.squaredGLMM(allreg_dissimilarity_mod_sbt_mean_avg.s)[[2]],3), #conditional r-squared
                                                            round(AICc(allreg_dissimilarity_mod_sbt_mean_avg.s),3),#AICc
                                                            5 #degrees of freedom 
                                                            ),nrow =1))

balanced_dissimilarity_sbt_fishing_model_results <- rbind(balanced_dissimilarity_sbt_fishing_model_results, balanced_dissimilarity_sbt_fishing_model_results_row, use.names = F)
  
fishing_temp_model_ranking <- AICc(allreg_dissimilarity_fishing_mean_mod_sbt_mean_avg, allreg_dissimilarity_fishing_mean_mod_sbt_mean_avg.nointeract, allreg_dissimilarity_mod_sbt_mean_avg, allreg_dissimilarity_fishing_mean_mod_sbt_mean_avg.s,  allreg_dissimilarity_fishing_mean_mod_sbt_mean_avg.s.nointeract, allreg_dissimilarity_mod_sbt_mean_avg.s, allreg_dissimilarity_fishing_mean_mod.intercept)

#Best model: best model = allreg_dissimilarity_fishing_mean_mod (no temp; different than surface temperature which was no fishing)
View(balanced_dissimilarity_sbt_fishing_model_results)

summary(allreg_dissimilarity_fishing_mean_mod)

saveRDS(balanced_dissimilarity_sbt_fishing_model_results, here::here("output","balanced_dissimilarity_sbt_fishing_model_results.rds"))

```

Significant predictors: 
summed_tonnes_scaled_byreg (as the relative fishing pressure increases, bray dissimilarity decreases)

                            Estimate Std. Error        df t value Pr(>|t|)    
(Intercept)                 0.595931   0.019402 35.941509  30.714   <2e-16 ***
summed_tonnes_scaled_byreg -0.008120   0.005989 29.195211  -1.356    0.186    

What about r^2 value
```{r}
r.squaredGLMM(allreg_dissimilarity_fishing_mean_mod)

#marginal GLMM (variance explained by fixed effects)
#conditional GLMM (variance explained by entire model)

#           R2m       R2c
#[1,]       0.002      0.92. (very little explained by fishing or temperature, most explained by survey_units)
```



---


#CODE BELOW MAY NO LONGER BE RELEVANT
Plot this model with interaction
```{r}
library(interactions)
sim_slopes(allreg_dissimilarity_fishing_temp_mod_SDonly, pred = sbt_SD_scaled, modx = summed_tonnes_scaled, johnson_neyman = FALSE)

interact_plot(allreg_dissimilarity_fishing_mean_mod_sbt_max_avg, pred = yearly_max_bypoint_avg, modx = summed_tonnes_scaled_byreg, plot.points = T, colors = c("#b3ffe4","#00d68b","#003d28"), x.label = "Maximum Sea Surface Temperature", y.label = "Mean Annual Bray Curtis Dissimilarity") + theme_classic()

ggsave(filename = "interaction_fish_temp.png" ,height = 4, width = 10)

interact_plot(allreg_dissimilarity_fishing_temp_mod_SDonly, pred = sbt_SD_scaled, modx = summed_tonnes_scaled, linearity.check = TRUE, 
              plot.points = TRUE) #check linearity 

```
Summary: Steepest slope between temp and dissimilarity when fishing is high

###What if we also tie in latitude and spp num?
```{r}
allreg_dissimilarity_fishing_mean_mod_sbt_max_avg_spp_num <- lmer(bray_curtis_dissimilarity_balanced_mean ~ summed_tonnes_scaled_byreg*yearly_max_bypoint_avg*spp_num.s + (1|survey_unit), data = dissimilarities_temp_fishing_regstats)

allreg_dissimilarity_fishing_mean_mod_sbt_max_avg_mid_lat <- lmer(bray_curtis_dissimilarity_balanced_mean ~ summed_tonnes_scaled_byreg*yearly_max_bypoint_avg*mid_lat + (1|survey_unit), data = dissimilarities_temp_fishing_regstats)

allreg_dissimilarity_fishing_mean_mod_sbt_max_avg_mid_lat_spp_num <- lmer(bray_curtis_dissimilarity_balanced_mean ~ summed_tonnes_scaled_byreg*yearly_max_bypoint_avg*mid_lat*spp_num.s + (1|survey_unit), data = dissimilarities_temp_fishing_regstats)

AIC(allreg_dissimilarity_fishing_mean_mod_sbt_max_avg_spp_num, allreg_dissimilarity_fishing_mean_mod_sbt_max_avg_mid_lat,allreg_dissimilarity_fishing_mean_mod_sbt_max_avg_mid_lat_spp_num)

summary(allreg_dissimilarity_fishing_mean_mod_sbt_max_avg_spp_num)
r.squaredGLMM(allreg_dissimilarity_fishing_mean_mod_sbt_max_avg_spp_num)

#no interaction with spp num?
allreg_dissimilarity_fishing_mean_mod_sbt_max_avg_spp_num_nointeract <- lmer(bray_curtis_dissimilarity_balanced_mean ~ summed_tonnes_scaled_byreg*yearly_max_bypoint_avg+spp_num.s + (1|survey_unit), data = dissimilarities_temp_fishing_regstats)

#exclude temperature?
allreg_dissimilarity_fishing_mean_mod_spp_num <- lmer(bray_curtis_dissimilarity_balanced_mean ~ summed_tonnes_scaled_byreg*spp_num.s + (1|survey_unit), data = dissimilarities_temp_fishing_regstats)

#exclude temp no interactions
allreg_dissimilarity_fishing_mean_mod_spp_num_nointeract <- lmer(bray_curtis_dissimilarity_balanced_mean ~ summed_tonnes_scaled_byreg+spp_num.s + (1|survey_unit), data = dissimilarities_temp_fishing_regstats)

AICc(allreg_dissimilarity_fishing_mean_mod_sbt_max_avg_spp_num_nointeract,allreg_dissimilarity_fishing_mean_mod_spp_num, allreg_dissimilarity_fishing_mean_mod_sbt_max_avg_spp_num,allreg_dissimilarity_fishing_mean_mod_spp_num_nointeract)

summary(allreg_dissimilarity_fishing_mean_mod_sbt_max_avg_spp_num_nointeract)
r.squaredGLMM(allreg_dissimilarity_fishing_mean_mod_sbt_max_avg_spp_num_nointeract)

#no temp performs better
summary(allreg_dissimilarity_fishing_mean_mod_spp_num_nointeract)
r.squaredGLMM(allreg_dissimilarity_fishing_mean_mod_spp_num_nointeract)

ggplot(dissimilarities_temp_fishing_regstats) +
  geom_point(aes(x = spp_num, y = bray_curtis_dissimilarity_balanced_mean, color = yearly_max_bypoint_avg)) +
  theme_classic()

ggplot(dissimilarities_temp_fishing_regstats) +
  geom_point(aes(x = yearly_max_bypoint_avg, y = bray_curtis_dissimilarity_balanced_mean, color = yearly_max_bypoint_avg)) +
  geom_smooth(aes(x = yearly_max_bypoint_avg, y = bray_curtis_dissimilarity_balanced_mean)) +
  theme_classic()
```

Patterns appear to be very survey dependent, and there's a lot of unexplained year to year variation.
- Warmer survey regions are more highly differentiated in general (makes sense, higher diversity )
- However, when we include spp num and fishing, temp drops out as a strong predictor
- As fishing pressure increases, dissimilarity decreases (more fishing = more homogenous)
- As # species increases, dissimilarity increases (more species overall = more heterogenous)
- From before, more species, more negative dissimilarity coefficient (more likely to be differentiating over time)


