---
title: "Regional Statistics"
output: html_notebook
---

Here, we're looking at Balanced Bray Curtis Dissimilarity

This code is to look at regional statistics, and figures and stats will end up in supplement.

TO DO: Add bottom temperature, right now it is surface temperature

```{r setup}
library(data.table)
library(ggplot2)
library(sf)
library(rgeos)
library(concaveman)
library(raster)
library(stringr)
library(rnaturalearth)
library(rmapshaper)
library(nlme)


#raw data
FishGlob_clean <- readRDS(here::here("data", "cleaned", "FishGlob.wellsampledyearscells_complete.final.rds"))

#model coefficients from year_dissimilarity_models.Rmd
model_coefs_reduced_length.unique <- readRDS(here::here("output","region_stats","model_coefs_reduced_length.unique.Rds"))

#NB: years sampled = last_year-first_year+1

```

*Additional cleaning (first written in "dissim_metric_space_time.Rmd")*


Delete observations without CPUE except for MEDITS (we will do trends in abundance metrics for this region)
```{r delete no CPUE obs}
#delete all rows where wgt_cpue is NA,  unless it is MEDITS region
FishGlob_clean.noNA <- FishGlob_clean[(survey != "MEDITS" & !is.na(wgt_cpue)) | survey == "MEDITS"]

#check
summary(FishGlob_clean.noNA[, wgt_cpue]) #why inf? 13 observations from GSL-N,  I will delete these
FishGlob_clean.noNA <- copy(FishGlob_clean.noNA)[!is.infinite(wgt_cpue), ]

```
Full Name Helper
```{r}
name_helper <- data.table(Survey_Name_Season = c("Aleutian Islands",
                                    "Baltic Sea Q1",
                                    "Baltic Sea Q4",
                                    "Chile",
                                    "Newfoundland",
                                    "Queen Charlotte Sound",
                                    "Eastern Bering Sea",
                                    "Bay of Biscay",
                                    "English Channel",
                                    "Gulf of Mexico Summer",
                                    "Gulf of Alaska",
                                    "Greenland",
                                    "N Gulf of St. Lawrence",
                                    "S Gulf of St. Lawrence",
                                    "Iceland",
                                    "Irish Sea",
                                    "Mediterranean",
                                    "Namibia",
                                    "NE US Fall",
                                    "NE US Spring",
                                    "N Ireland Q1",
                                    "N Ireland Q4",
                                    "Barents Sea Norway Q3",
                                    "N Sea Q1",
                                    "N Sea Q3",
                                    "Chatham Rise",
                                    "E Coast S Island NZ",
                                    "W Coast S Island NZ",
                                    "Portugal",
                                    "S Georgia",
                                  "Scotian Shelf Summer",
                                  "SE US Fall",
                                  "SE US Spring",
                                  "SE US Summer",
                                  "W Coast US",
                                  "Atlantic Ocean ZA",
                                  "Indian Ocean ZA",
                                   "Rockall Plateau",
                                  "Scotland Shelf Sea Q1",
                                  "Scotland Shelf Sea Q4",
                                  "Falkland Islands",
                                  "Gulf of Mexico Fall",
                                  "Barents Sea Norway Q1",
                                  "Sub-Arctic NZ",
                                  "Scotian Shelf Spring"),
                          survey_unit = c(
                                  "AI",        
                                  "BITS-1",    
                                  "BITS-4",    
                                  "CHL",       
                                  "DFO-NF",    
                                  "DFO-QCS",   
                                  "EBS",       
                                  "EVHOE",     
                                  "FR-CGFS",   
                                  "GMEX-Summer",
                                  "GOA",       
                                  "GRL-DE",    
                                  "GSL-N",     
                                  "GSL-S",     
                                  "ICE-GFS",   
                                  "IE-IGFS",   
                                  "MEDITS",    
                                  "NAM",       
                                  "NEUS-Fall", 
                                  "NEUS-Spring",
                                  "NIGFS-1",   
                                  "NIGFS-4",   
                                  "Nor-BTS-3", 
                                  "NS-IBTS-1", 
                                  "NS-IBTS-3", 
                                  "NZ-CHAT",   
                                  "NZ-ECSI",   
                                  "NZ-WCSI",   
                                  "PT-IBTS",   
                                  "S-GEORG",   
                                  "SCS-SUMMER",
                                  "SEUS-fall", 
                                  "SEUS-spring",
                                  "SEUS-summer",
                                  "WCANN",     
                                  "ZAF-ATL",   
                                  "ZAF-IND",
                                  "ROCKALL",
                                  "SWC-IBTS-1",
                                  "SWC-IBTS-4",
                                  "FALK",
                                  "GMEX-Fall",
                                  "Nor-BTS-1",
                                  "NZ-SUBA",
                                  "SCS-SPRING"
                          ))

```


```{r list of all survey season combinations}
all_survey_units <- unique(FishGlob_clean.noNA[, survey_unit])
```

For NZ-CHAT, we want to lump December observations with next year of observations (January,  February)
```{r lump December observations with previous year}
FishGlob_clean.noNA[survey_unit == "NZ-CHAT" & month == 12,  year := year+1, ]
```

Some spp are duplicated within a haul. In all cases,  this is due to verbatim names matching to a single accepted name. Therefore,  we will sum these wgts and abundances
```{r sum duplicates}
FishGlob_clean.noNA[, wgt := sum(wgt,  na.rm = T), .(accepted_name,  haul_id)][, 
                     wgt_h := sum(wgt_h,  na.rm = T), .(accepted_name,  haul_id)][, 
                     wgt_cpue := sum(wgt_cpue,  na.rm = T), .(accepted_name,  haul_id)][, 
                     num := sum(num,  na.rm = T), .(accepted_name,  haul_id)][, 
                     num_h := sum(num_h,  na.rm = T), .(accepted_name,  haul_id)][, 
                     num_cpue := sum(num_cpue,  na.rm = T), .(accepted_name,  haul_id)]

FishGlob_clean.noNA <- unique(FishGlob_clean.noNA[,.(survey, haul_id, country, sub_area, continent, stat_rec, station, stratum, year, month, day, quarter, season, latitude, longitude, haul_dur, area_swept, gear, depth, sbt, sst, num, num_h, num_cpue, wgt, wgt_h, wgt_cpue, verbatim_aphia_id, accepted_name, aphia_id, SpecCode, kingdom, phylum, class, order, family, genus, rank, survey_unit, years_sampled, cell_ID, cell_year_count, year_cell_count, years_sampled_update)])

#make depth numeric
FishGlob_clean.noNA[,depth := as.numeric(depth)]

#make survey and survey unit factors
FishGlob_clean.noNA[,survey:=factor(survey)][,survey_unit:=factor(survey_unit)]

#adjust years
FishGlob_clean.noNA[,year_adj := year-min(year)+1]

#add column for yearly summed biomass (note that MEDITS doesn't have biomass, so it will be excluded from these plots)
FishGlob_clean.noNA[,wgt_cpue_annual := sum(wgt_cpue,na.rm = T),.(year, survey_unit)]

#add column for yearly summed # of hauls
FishGlob_clean.noNA[,haul_id_count_annual := uniqueN(haul_id),.(year, survey_unit)]

#add column for yearly summed # of spp
FishGlob_clean.noNA[,spp_count_annual := uniqueN(accepted_name),.(year_adj, survey_unit)]

FishGlob_richness_year_survey <- unique(FishGlob_clean.noNA[,.(year_adj, survey_unit, spp_count_annual,  year)])

saveRDS(FishGlob_richness_year_survey, file = here::here("output","FishGlob_richness_year_survey.Rds"))
```

Calculate coefficients from lmer for trend in species richness within a region over time
```{r}
#plot year vs richness for each reg
ggplot(data = FishGlob_richness_year_survey) +
  geom_point(aes(x = year, y = spp_count_annual), size = 2) +
  facet_wrap(~survey_unit, scales = "free") +
  theme_classic()

richness_trend_lme <- lme(spp_count_annual ~ year_adj, random = (~1 + year_adj|survey_unit),data = FishGlob_richness_year_survey)

richness_trend_coefs <- data.table(coef(richness_trend_lme))
richness_trend_coefs[,survey_unit := rownames(coef(richness_trend_lme))][,Year_Coef_richness := round(year_adj,5)][,Intercept_richness := round(`(Intercept)`,2)]
View(richness_trend_coefs)

richness_trend_coefs <- richness_trend_coefs[name_helper, on = "survey_unit"]

richness_trend_coefs.exp <- richness_trend_coefs[,.(Survey_Name_Season, Intercept_richness, Year_Coef_richness, survey_unit)]

#export this table
fwrite(richness_trend_coefs.exp, file = here::here("output","richness_trend_coefs.exp.csv"))
```

Link trend in beta diversity to trend in gamma diversity
```{r}
#merge
trend_coefs <- richness_trend_coefs.exp[model_coefs_reduced_length.unique, on = c("survey_unit")]

ggplot(data = trend_coefs, aes(x = Year_Coef_richness, y = year_adj)) +
  geom_point() +
  geom_text(aes(label = survey_unit), size = 2, nudge_y = 0.0003) +
  labs(x = "Trend in gamma diversity", y = "Trend in beta diversity") +
  geom_hline(yintercept = 0, color = "grey", linetype = "dashed") +
  geom_vline(xintercept = 0, color = "grey", linetype = "dashed") +
  theme_classic()

ggsave(path = here::here("figures"), filename = "trends_beta_gamma.jpg")

#linear model?
summary(lm(Year_Coef_richness~year_adj, data = trend_coefs))
```


Pull in avg temp values and pull in avg fishing values
```{r avg temp and fishing values}
 #open fishing and temp data
distances_dissimilarities_allyears.r.temp.fishing <- readRDS( here::here("output","distance_decay","distances_dissimilarities_allyears.r.temp.fishing.rds"))

distances_dissimilarities_allyears.r.temp.fishing[,mean_reg_temp_allyears := mean(sst_mean, na.rm = T),.(survey_unit)][,mean_reg_fishing_allyears := mean(summed_tonnes, na.rm = T),.(survey_unit)]


avg_temp_avg_fishing <- unique(distances_dissimilarities_allyears.r.temp.fishing[,.(survey_unit,mean_reg_fishing_allyears, mean_reg_temp_allyears)]) 

#link coefficients
year_survey_unit_characteristics <- model_coefs_reduced_length.unique[avg_temp_avg_fishing, on = "survey_unit"]

```

We need to look at gamma vs beta for all regions independently
```{r}
Fishglob_gamma_beta <- FishGlob_richness_year_survey[distances_dissimilarities_allyears.r.temp.fishing, on = c("survey_unit","year")]

ggplot(Fishglob_gamma_beta) +
  geom_point(aes(x = spp_count_annual, y = bray_curtis_dissimilarity_balanced_mean), color = "orange", size = 1) +
  geom_point(aes(x = spp_count_annual, y = jaccard_dissimilarity_turnover_mean), color = "slateblue", size = 1) +
  facet_wrap(~survey_unit, scales = "free") +
  labs(y = "β-diversity", x = "γ-diversity") +
  theme_classic()

ggsave(path = here::here("figures"), filename = "Fishglob_gamma_beta.jpg", width = 11, height = 8, units = "in")
```


For area, we want to delete any areas that overlap with land, therefore, we have to pull in land polygon
```{r}
#basemap
world <- rnaturalearth::ne_countries(type = "countries", scale="large", returnclass="sf")

```


Calculate stats of all regions
```{r calculate all region stats}

#survey_unit
region_stats <- as.data.table(matrix(nrow = length(all_survey_units)))

region_stats[,survey_unit := as.character(V1)][, survey := as.character(V1)][, spp_num := as.numeric(V1)][,mean_annual_biomass := as.numeric()][,mean_haul_id_count_annual := as.numeric()][, study_period := as.numeric(V1)][, years_sampled := as.numeric(V1)][, lat_range := as.numeric(V1)][, mid_lat := as.numeric(V1)][, lon_range := as.numeric(V1)][, area_km := as.numeric(V1)][, depth_range := as.numeric(V1)][,  mid_depth := as.numeric(V1)][, mean_temp := as.numeric(V1)][, mean_landings := as.numeric(V1)][,bray_coef := as.numeric()][,bray_coef_upr := as.numeric()][,bray_coef_lwr := as.numeric()][,direction_of_change := as.character()][,significant := as.logical()]

  region_stats[,V1:=NULL]

for (i in 1:length(all_survey_units)) {
  reduced_FishGlob_cleaned.10year <- FishGlob_clean.noNA[survey_unit == all_survey_units[i]]
  
    ####unique lat lon
    lat_lon <- unique(reduced_FishGlob_cleaned.10year[,.(latitude, longitude)])
    
    pts <- st_as_sf(lat_lon, coords=c('longitude','latitude'), crs=4326 )
    
    conc <- concaveman(pts, 1,2)
    
    #delete areas that overlap
    conc_nooverlap <- ms_erase(conc,world[1])
    
    sf_use_s2(FALSE) #helps with spherical geometry
    area <- st_area(conc_nooverlap) #m2

    #compile into region_Stats table    
                  region_stats[i,"survey_unit"] <- paste0(all_survey_units[i])
                  region_stats[i,"survey"] <- word(all_survey_units[i],1, sep = "_")
                  region_stats[i,"spp_num"] <- length(unique(reduced_FishGlob_cleaned.10year[,accepted_name]))
                  region_stats[i,"mean_annual_biomass"] <- mean(reduced_FishGlob_cleaned.10year[,wgt_cpue_annual])
                  region_stats[i,"mean_haul_id_count_annual"] <- mean(reduced_FishGlob_cleaned.10year[,haul_id_count_annual])
                  region_stats[i,"study_period"] <- max(reduced_FishGlob_cleaned.10year$year)-min(reduced_FishGlob_cleaned.10year$year)+1
                  region_stats[i,"years_sampled"] <- length(unique(reduced_FishGlob_cleaned.10year$year))
                  region_stats[i,"lat_range"] <- max(reduced_FishGlob_cleaned.10year$latitude)-min(reduced_FishGlob_cleaned.10year$latitude)
                  region_stats[i,"mid_lat"] <- mean(reduced_FishGlob_cleaned.10year$latitude)
                  region_stats[i,"lon_range"] <- max(reduced_FishGlob_cleaned.10year$longitude)-min(reduced_FishGlob_cleaned.10year$longitude)
                  region_stats[i,"area_km"] <- area/(1000^2)
                  region_stats[i,"depth_range"] <- max(reduced_FishGlob_cleaned.10year$depth, na.rm = T)-min(reduced_FishGlob_cleaned.10year$depth, na.rm = T)
                  region_stats[i,"mid_depth"] <- mean(reduced_FishGlob_cleaned.10year$depth, na.rm = T)
                  region_stats[i,"mean_temp"] <- year_survey_unit_characteristics[survey_unit == all_survey_units[i],mean_reg_temp_allyears]
                  region_stats[i,"mean_landings"] <- year_survey_unit_characteristics[survey_unit == all_survey_units[i],mean_reg_fishing_allyears]
                  region_stats[i,"bray_coef"] <- year_survey_unit_characteristics[survey_unit == all_survey_units[i],year_adj]
                  region_stats[i,"bray_coef_lwr"] <- year_survey_unit_characteristics[survey_unit == all_survey_units[i],lwr]
                  region_stats[i,"bray_coef_upr"] <- year_survey_unit_characteristics[survey_unit == all_survey_units[i],upr]
                  region_stats[i,"Directional_Change"] <- year_survey_unit_characteristics[survey_unit == all_survey_units[i],Directional_Change]
                  region_stats[i,"significant"] <- year_survey_unit_characteristics[survey_unit == all_survey_units[i],significant]
}
  
#if any values are -Inf, change to NA
invisible(lapply(names(region_stats),function(.name) set(region_stats, which(is.infinite(region_stats[[.name]])), j = .name,value =NA)))

#add significant directions column
region_stats[,"Significant Trends" := ifelse(significant==T, Directional_Change, "No directional change")]

#add column for avg # of hauls per area
region_stats[,avg_hauls_area := mean_haul_id_count_annual/area_km]


saveRDS(region_stats, here::here("output","region_stats","region_stats.rds"))

#or pull in
region_stats <- readRDS(here::here("output","region_stats","region_stats.rds"))


#test plot
plot(conc)
plot(conc_nooverlap)
```

Proportion test (prop homogenizing/differentiating vs. prop not)

H0: The null hypothesis is that the two hemispheres from which
   the regions were drawn have the same true proportion of homogenizers.
A:  The alternative is that this proportion is different in at
   least one of the hemispheres.
   
H0: The null hypothesis is that the two hemispheres from which
   the regions were drawn have the same true proportion of differentiators.
A:  The alternative is that this proportion is different in at
   least one of the hemispheres.
   
   Differentiation was more common in the southern hemisphere (four out of eight surveys), while homogenization was more common in the northern hemisphere (eight out of 29 surveys).

```{r prop test}
                #N, S
homogenizers <- c(8,0)
differentiators <- c(2,4)
total_surveys <- c(29,8)


prop.test(homogenizers,total_surveys) #p value << 0.05, so groups are differennt
prop.test(differentiators,total_surveys) #p value >> 0.05, so groups are different

```

Regions that are homogenizing tend to have lower overall gamma diversity than regions that are differentiating
```{r spp num}

spp_num_vs_coef <- ggplot() +
    geom_errorbar(data = region_stats, aes(x = spp_num, y = bray_coef, ymin = bray_coef_lwr, ymax = bray_coef_upr, color = `Significant Trends`), alpha = 0.8, linewidth = 0.3) + #add confidence intervals
        geom_point(data = region_stats, aes(x = spp_num , y = bray_coef, color = `Significant Trends`), stat = 'identity', size = 1) +
    scale_color_manual(values = c("#91c874", "#e7ac5b", "#cbbfde"), name = "Trend in dissimilarity") +
  geom_hline(yintercept = 0, size = 0.5) +
  geom_smooth(data = region_stats, aes(x = spp_num, y = bray_coef), method = "lm", color = "black", se = F, size = 0.5) +
  ylab("β-diversity trend") + #balanced BC dissimiliarity
  xlab("γ-diversity") +
  theme_classic()

spp_num_vs_coef

ggsave(spp_num_vs_coef, path = here::here("figures","bray_coef_vs_survey_characteristics"), filename = "spp_num_vs_coef.jpg", height =3, width = 5, unit = "in")

#linear model
summary(lm(abs(bray_coef)~spp_num, data = region_stats)) #not significant
summary(lm(bray_coef~spp_num, data = region_stats)) #significant

#low# of species, more likely to experience differentiation, high # of species, more likely to experience homogenization

saveRDS(spp_num_vs_coef, here::here("figures","bray_coef_vs_survey_characteristics","spp_num_vs_coef.rds"))

```

More likely to detect both homogenization and differentation for longer study periods
```{r study period}

study_period_vs_coef <- ggplot() +
    geom_errorbar(data = region_stats, aes(x = study_period, y = bray_coef, ymin = bray_coef_lwr, ymax = bray_coef_upr, color = `Significant Trends`), alpha = 0.8, linewidth = 0.3) + #add confidence intervals
        geom_point(data = region_stats, aes(x = study_period , y = bray_coef, color = `Significant Trends`), stat = 'identity', size = 1) +
    scale_color_manual(values = c("#91c874", "#e7ac5b", "#cbbfde"), name = "Trend in dissimilarity") +
  geom_hline(yintercept = 0, size = 0.5) +
  ylab("β-diversity trend") + #balanced BC dissimiliarity
  xlab("Survey period") + #first minus last
  theme_classic()

study_period_vs_coef

ggsave(study_period_vs_coef, path = here::here("figures","bray_coef_vs_survey_characteristics"), filename = "study_period_vs_coef.jpg", height =3, width = 5, unit = "in")

#linear model
summary(lm(abs(bray_coef)~study_period, data = region_stats)) #not significant
summary(lm(bray_coef~study_period, data = region_stats)) #not significant
```

More likely to detect differentiation when fewer years are sampled and more likely to detect homogenization when more years are sampled
```{r study duration}

years_sampled_vs_coef <- ggplot() +
    geom_errorbar(data = region_stats, aes(x = years_sampled, y = bray_coef, ymin = bray_coef_lwr, ymax = bray_coef_upr, color = `Significant Trends`), alpha = 0.8, linewidth = 0.3) + #add confidence intervals
        geom_point(data = region_stats, aes(x = years_sampled , y = bray_coef, color = `Significant Trends`), stat = 'identity', size = 1) +
    scale_color_manual(values = c("#91c874", "#e7ac5b", "#cbbfde"), name = "Trend in dissimilarity") +
  geom_hline(yintercept = 0, size = 0.5) +
  ylab("β-diversity trend") + #balanced BC dissimiliarity
  xlab("Total years sampled") + #count of years sampled
  theme_classic()

years_sampled_vs_coef

ggsave(years_sampled_vs_coef, path = here::here("figures","bray_coef_vs_survey_characteristics"), filename = "years_sampled_vs_coef.jpg", height =3, width = 5, unit = "in")

#linear model
summary(lm(abs(bray_coef)~years_sampled, data = region_stats)) #not significant
summary(lm(bray_coef~years_sampled, data = region_stats)) #not significant


```

More likely to detect differentiation when a small latitudinal range is sampled, more likely to detect homogenization when a larger latitudinal range is sampled
```{r latitude range}


lat_range_vs_coef <- ggplot() +
    geom_errorbar(data = region_stats, aes(x = lat_range, y = bray_coef, ymin = bray_coef_lwr, ymax = bray_coef_upr, color = `Significant Trends`), alpha = 0.8, linewidth = 0.3) + #add confidence intervals
        geom_point(data = region_stats, aes(x = lat_range , y = bray_coef, color = `Significant Trends`), stat = 'identity', size = 1) +
    scale_color_manual(values = c("#91c874", "#e7ac5b", "#cbbfde"), name = "Trend in dissimilarity") +
  geom_hline(yintercept = 0, size = 0.5) +
  ylab("β-diversity trend") + #balanced BC dissimiliarity
  xlab("Latitudinal range (\u00B0s)") + #count of years sampled
  theme_classic()

lat_range_vs_coef

ggsave(lat_range_vs_coef, path = here::here("figures","bray_coef_vs_survey_characteristics"), filename = "lat_range_vs_coef.jpg", height =3, width = 5, unit = "in")

#linear model
summary(lm(abs(bray_coef)~lat_range, data = region_stats)) #not significant
summary(lm(bray_coef~lat_range, data = region_stats)) #nearly but not significant
```

Differentiation occurs below 30 degrees latitude, homogenization and no change both common at higher latitudes in the northern hemisphere, NO homogenization below 30 degrees latitude
```{r middle latitude}

summary(lm(bray_coef~mid_lat, data = region_stats))


mid_lat_vs_coef <- ggplot() +
    geom_errorbar(data = region_stats, aes(x = mid_lat, y = bray_coef, ymin = bray_coef_lwr, ymax = bray_coef_upr, color = `Significant Trends`), alpha = 0.8, linewidth = 0.3) + #add confidence intervals
        geom_point(data = region_stats, aes(x = mid_lat , y = bray_coef, color = `Significant Trends`), stat = 'identity', size = 1) +
    scale_color_manual(values = c("#91c874", "#e7ac5b", "#cbbfde"), name = "Trend in dissimilarity") +
  geom_hline(yintercept = 0, size = 0.5) +
  geom_smooth(data = region_stats, aes(x = mid_lat, y = bray_coef), method = "lm", color = "black", se = F, size = 0.5) +
  ylab("β-diversity trend") + #balanced BC dissimiliarity
  xlab(paste0("Latitude (\u00B0N)")) + #middle latitude
  theme_classic()

mid_lat_vs_coef

saveRDS(mid_lat_vs_coef, here::here("figures","bray_coef_vs_survey_characteristics","mid_lat_vs_coef.rds"))

ggsave(mid_lat_vs_coef, path = here::here("figures","bray_coef_vs_survey_characteristics"), filename = "mid_lat_vs_coef.jpg", height =3, width = 5, unit = "in")

#linear model
summary(lm(abs(bray_coef)~mid_lat, data = region_stats)) #not significant
summary(lm(bray_coef~mid_lat, data = region_stats)) #significant

      
```

#NEED TO FIX FOR REGIONS THAT CROSS DATE LINE
```{r longitudinal range}

ggplot(region_stats) +
  geom_boxplot(aes(x = `Significant Trends`, y = lon_range, group = `Significant Trends`)) +
  theme_classic()

ggplot(region_stats) +
  geom_point(aes(x = lon_range, y = bray_coef)) +
  theme_classic()
```

Areas that are differentiating tend to be smaller than areas that are homogenizing
```{r area}
area_km_vs_coef <- ggplot() +
    geom_errorbar(data = region_stats, aes(x = area_km/1000, y = bray_coef, ymin = bray_coef_lwr, ymax = bray_coef_upr, color = `Significant Trends`), alpha = 0.8, linewidth = 0.3) + #add confidence intervals
        geom_point(data = region_stats, aes(x = (area_km/1000) , y = bray_coef, color = `Significant Trends`), stat = 'identity', size = 1) +
    scale_color_manual(values = c("#91c874", "#e7ac5b", "#cbbfde"), name = "Trend in dissimilarity") +
  geom_hline(yintercept = 0, size = 0.5) +
  ylab("β-diversity trend") + #balanced BC dissimiliarity
  xlab("Area (1000s of squared km)") + #area_km
  theme_classic()

area_km_vs_coef

ggsave(area_km_vs_coef, path = here::here("figures","bray_coef_vs_survey_characteristics"), filename = "area_km_vs_coef.jpg", height =3, width = 5, unit = "in")

#linear model
summary(lm(abs(bray_coef)~area_km, data = region_stats)) #not significant
summary(lm(bray_coef~area_km, data = region_stats)) #not significant

```

```{r number of hauls vs area}

avg_hauls_area_vs_coef <- ggplot() +
    geom_errorbar(data = region_stats, aes(x = avg_hauls_area, y = bray_coef, ymin = bray_coef_lwr, ymax = bray_coef_upr, color = `Significant Trends`), alpha = 0.8, linewidth = 0.3) + #add confidence intervals
        geom_point(data = region_stats, aes(x = (avg_hauls_area) , y = bray_coef, color = `Significant Trends`), stat = 'identity', size = 1) +
    scale_color_manual(values = c("#91c874", "#e7ac5b", "#cbbfde"), name = "Trend in dissimilarity") +
  geom_hline(yintercept = 0, size = 0.5) +
  ylab("β-diversity trend") + #balanced BC dissimiliarity
  xlab("Hauls per area (squared km)") + #avg_hauls_area
  theme_classic()

avg_hauls_area_vs_coef

ggsave(avg_hauls_area_vs_coef, path = here::here("figures","bray_coef_vs_survey_characteristics"), filename = "avg_hauls_area_vs_coef.jpg", height =3, width = 5, unit = "in")

#linear model
summary(lm(abs(bray_coef)~avg_hauls_area, data = region_stats)) #not significant
summary(lm(bray_coef~avg_hauls_area, data = region_stats)) #not significant

```


```{r depth_range}

depth_range_vs_coef <- ggplot() +
    geom_errorbar(data = region_stats, aes(x = depth_range, y = bray_coef, ymin = bray_coef_lwr, ymax = bray_coef_upr, color = `Significant Trends`), alpha = 0.8, linewidth = 0.3) + #add confidence intervals
        geom_point(data = region_stats, aes(x = (depth_range) , y = bray_coef, color = `Significant Trends`), stat = 'identity', size = 1) +
    scale_color_manual(values = c("#91c874", "#e7ac5b", "#cbbfde"), name = "Trend in dissimilarity") +
  geom_hline(yintercept = 0, size = 0.5) +
  ylab("β-diversity trend") + #balanced BC dissimiliarity
  xlab("Depth range (m)") + #depth_range
  theme_classic()

depth_range_vs_coef

ggsave(depth_range_vs_coef, path = here::here("figures","bray_coef_vs_survey_characteristics"), filename = "depth_range_vs_coef.jpg", height =3, width = 5, unit = "in")

#linear model
summary(lm(abs(bray_coef)~depth_range, data = region_stats)) #not significant
summary(lm(bray_coef~depth_range, data = region_stats)) #not significant
```

```{r mid_depth}

mid_depth_vs_coef <- ggplot() +
    geom_errorbar(data = region_stats, aes(x = mid_depth, y = bray_coef, ymin = bray_coef_lwr, ymax = bray_coef_upr, color = `Significant Trends`), alpha = 0.8, linewidth = 0.3) + #add confidence intervals
        geom_point(data = region_stats, aes(x = (mid_depth) , y = bray_coef, color = `Significant Trends`), stat = 'identity', size = 1) +
    scale_color_manual(values = c("#91c874", "#e7ac5b", "#cbbfde"), name = "Trend in dissimilarity") +
  geom_hline(yintercept = 0, size = 0.5) +
  ylab("β-diversity trend") + #balanced BC dissimiliarity
  xlab("Average depth (m)") + #mid_depth
  theme_classic()

mid_depth_vs_coef

ggsave(mid_depth_vs_coef, path = here::here("figures","bray_coef_vs_survey_characteristics"), filename = "mid_depth_vs_coef.jpg", height =3, width = 5, unit = "in")

#linear model
summary(lm(abs(bray_coef)~mid_depth, data = region_stats)) #not significant
summary(lm(bray_coef~mid_depth, data = region_stats)) #not significant
```

```{r mean_temp}
mean_temp_vs_coef <- ggplot() +
    geom_errorbar(data = region_stats, aes(x = mean_temp, y = bray_coef, ymin = bray_coef_lwr, ymax = bray_coef_upr, color = `Significant Trends`), alpha = 0.8, linewidth = 0.3) + #add confidence intervals
        geom_point(data = region_stats, aes(x = (mean_temp) , y = bray_coef, color = `Significant Trends`), stat = 'identity', size = 1) +
    scale_color_manual(values = c("#91c874", "#e7ac5b", "#cbbfde"), name = "Trend in dissimilarity") +
  geom_hline(yintercept = 0, size = 0.5) +
  geom_smooth(data = region_stats, aes(x = mean_temp, y = bray_coef), method = "lm", color = "black", se = F, linetype = "dashed", size = 0.5) +
  ylab("β-diversity trend") + #balanced BC dissimiliarity
  xlab("Average temperature (\u00B0C)") + #mean_temp
  theme_classic()

mean_temp_vs_coef

ggsave(mean_temp_vs_coef, path = here::here("figures","bray_coef_vs_survey_characteristics"), filename = "mean_temp_vs_coef.jpg", height =3, width = 5, unit = "in")

#linear model
summary(lm(abs(bray_coef)~mean_temp, data = region_stats)) #not significant
summary(lm(bray_coef~mean_temp, data = region_stats)) #not significant

saveRDS(mean_temp_vs_coef, here::here("figures","bray_coef_vs_survey_characteristics","mean_temp_vs_coef.rds"))
```

```{r mean_landings}
mean_landings_vs_coef <- ggplot() +
    geom_errorbar(data = region_stats, aes(x = (mean_landings/1000), y = bray_coef, ymin = bray_coef_lwr, ymax = bray_coef_upr, color = `Significant Trends`), alpha = 0.8, linewidth = 0.3) + #add confidence intervals
        geom_point(data = region_stats, aes(x = (mean_landings/1000) , y = bray_coef, color = `Significant Trends`), stat = 'identity', size = 1) +
    scale_color_manual(values = c("#91c874", "#e7ac5b", "#cbbfde"), name = "Trend in dissimilarity") +
  geom_hline(yintercept = 0, size = 0.5) +
  ylab("β-diversity trend") + #balanced BC dissimiliarity
  xlab("Mean landings (1000s of tonnes)") + #average annual landings
  theme_classic()

mean_landings_vs_coef

ggsave(mean_landings_vs_coef, path = here::here("figures","bray_coef_vs_survey_characteristics"), filename = "mean_landings_vs_coef.jpg", height =3, width = 5, unit = "in")

#linear model
summary(lm(abs(bray_coef)~mean_landings, data = region_stats)) #not significant
summary(lm(bray_coef~mean_landings, data = region_stats)) #not significant
```

#Avg number of tows per year
```{r avg number of tows per year in region}
mean_haul_id_count_annual_vs_coef <- ggplot() +
    geom_errorbar(data = region_stats, aes(x = mean_haul_id_count_annual, y = bray_coef, ymin = bray_coef_lwr, ymax = bray_coef_upr, color = `Significant Trends`), alpha = 0.8, linewidth = 0.3) + #add confidence intervals
        geom_point(data = region_stats, aes(x = mean_haul_id_count_annual , y = bray_coef, color = `Significant Trends`), stat = 'identity', size = 1) +
    scale_color_manual(values = c("#91c874", "#e7ac5b", "#cbbfde"), name = "Trend in dissimilarity") +
  geom_hline(yintercept = 0, size = 0.5) +
  ylab("β-diversity trend") + #balanced BC dissimiliarity
  xlab("Average number of hauls per year") + #average annual landings
  theme_classic()

mean_haul_id_count_annual_vs_coef

ggsave(mean_haul_id_count_annual_vs_coef, path = here::here("figures","bray_coef_vs_survey_characteristics"), filename = "mean_haul_id_count_annual_vs_coef.jpg", height =3, width = 5, unit = "in")

#linear model
summary(lm(abs(bray_coef)~mean_haul_id_count_annual, data = region_stats)) #not significant
summary(lm(bray_coef~mean_haul_id_count_annual, data = region_stats)) #not significant
```

