---
title: "Year TOTAL Dissimilarity Models"
output: html_notebook
---

This code is Script 4a  for Kitchel et al. Biotic homogenization, the exception and not the rule for marine fish communities manuscript.

COPY FROM THIS SCRIPT

(We expect this to be the main analyses in the text)

- This project is a collaborative effort to describe changes in taxonomic composition  of fish communities around the world--as sampled by bottom trawl surveys.

- Code by Zoë J. Kitchel

SESSION INFO TO DO

```{r setup}
library(data.table)
library(vegan)
library(sf)
library(concaveman) #polygon around points
library(betapart) #allows us to partition beta diversity
library(geosphere)
library(ggpubr) #stat_regline_equation
library(nlme)
library(mgcv) #to make gam
library(cowplot)
library(lme4)

#Pull Dissimilarity Means
distances_dissimilarities_allyears.r <- readRDS(here::here("output", "dissimilarities", "distances_dissimilarities_allyears.r.rds"))

#make survey and survey unit factors
distances_dissimilarities_allyears.r[,survey:=factor(survey)][,survey_unit:=factor(survey_unit)]

#adjust years
distances_dissimilarities_allyears.r[,year_adj := year-min(year)+1]

#add new variable for year in sequence per region
distances_dissimilarities_allyears.r[,first_year := min(year),.(survey_unit)]
distances_dissimilarities_allyears.r[,last_year := max(year),.(survey_unit)]

#distances_dissimilarities_allyears.r[,year_in_seq := year-first_year+1]

distances_dissimilarities_allyears.r[,years_sampled := last_year-first_year+1]


```

###Palette for Plotting
Palette for plotting all 42 survey units
```{r link colors to survey units}
survey_unit.list <- levels(factor(distances_dissimilarities_allyears.r[,survey_unit]))

palette_42 <- c(
  "#5A5156", #AI
  "#DF00DB", #BITS-1
  "#DB8EDA", #BITS-4
  "#F6222E", #CHL
  "#F8A19F", #DFO-NF
  "#16FF32", #DFO-QCS
  "#325A9B", #EBS
  "#3283FE", #EVHOE
  "#FEAF16", #FR-CGFS
  "#fccb6d", #GMEX-Fall
  "#1C8356", #GMEX-Summer
  "#C4451C", #GOA
  "#85660D", #GRL-DE
  "#B0009F", #GSL-N
  "#BF79B8", #GSL-S
  "#1CBE4F", #ICE-GFS
  "#782AB6", #IE-IGFS
  "#90AD1C", #MEDITS
  "#6B003A", #NAM
  "#A75B00", #NEUS-Fall
  "#E3B072", #NEUS-Spring
  "#02E8B6", #NIGFS-1
  "#97E7D5", #NIGFS-4
  "#B00068", #Nor-BTS-3
  "#00B9E3", #NS-IBTS-1
  "#95E2F4", #NS-IBTS-3
  "#B3CE73", #NZ-CHAT
  "#689500", #NZ-ECSI
  "#364d02",#NZ-SUBA
  "#AAF400", #NZ-WCSI
  "#AA0DFE", #PT-IBTS
  "#7f9eb8", #ROCKALL
  "#FA0087", #S-GEORG
  "#DEA0FD", #SCS-Summer
  "#FCEF88", #SEUS-fall
  "#A59405", #SEUS-spring
  "#FCE100", #SEUS-summer
  "#544563", #SWC-IBTS-1
  "#a37fc7", #SWC-IBTS-4
  "#C075A6", #WCANN
  "#BDCDFF", #ZAF-ATL
  "#003EFF"  #ZAF-IND
)

color_link <- data.table(survey_unit = survey_unit.list,hex = palette_42)
```

Add names for plotting
```{r add names for plotting}

name_helper <- data.table(Survey_Name_Season = c("Aleutian Islands",
                                    "Baltic Sea Q1",
                                    "Baltic Sea Q4",
                                    "Chile",
                                    "Newfoundland",
                                    "Queen Charlotte Sound",
                                    "Eastern Bering Sea",
                                    "Bay of Biscay",
                                    "English Channel",
                                    "Gulf of Mexico Summer",
                                    "Gulf of Alaska",
                                    "Greenland",
                                    "N Gulf of St. Lawrence",
                                    "S Gulf of St. Lawrence",
                                    "Iceland",
                                    "Irish Sea",
                                    "Mediterranean",
                                    "Namibia",
                                    "NE US Fall",
                                    "NE US Spring",
                                    "N Ireland Q1",
                                    "N Ireland Q4",
                                    "Barents Sea Norway Q3",
                                    "N Sea Q1",
                                    "N Sea Q3",
                                    "Chatham Rise NZ",
                                    "E Coast S Island NZ",
                                    "W Coast S Island NZ",
                                    "Portugal",
                                    "S Georgia",
                                  "Scotian Shelf Summer",
                                  "SE US Fall",
                                  "SE US Spring",
                                  "SE US Summer",
                                  "W Coast US",
                                  "Atlantic Ocean ZA",
                                  "Indian Ocean ZA",
                                   "Rockall Plateau",
                                  "Scotland Shelf Sea Q1",
                                  "Scotland Shelf Sea Q4",
                                  "Falkland Islands",
                                  "Gulf of Mexico Fall",
                                  "Barents Sea Norway Q1",
                                  "Sub-Arctic NZ",
                                  "Scotian Shelf Spring"),
                          survey_unit = c(
                                  "AI",        
                                  "BITS-1",    
                                  "BITS-4",    
                                  "CHL",       
                                  "DFO-NF",    
                                  "DFO-QCS",   
                                  "EBS",       
                                  "EVHOE",     
                                  "FR-CGFS",   
                                  "GMEX-Summer",
                                  "GOA",       
                                  "GRL-DE",    
                                  "GSL-N",     
                                  "GSL-S",     
                                  "ICE-GFS",   
                                  "IE-IGFS",   
                                  "MEDITS",    
                                  "NAM",       
                                  "NEUS-Fall", 
                                  "NEUS-Spring",
                                  "NIGFS-1",   
                                  "NIGFS-4",   
                                  "Nor-BTS-3", 
                                  "NS-IBTS-1", 
                                  "NS-IBTS-3", 
                                  "NZ-CHAT",   
                                  "NZ-ECSI",   
                                  "NZ-WCSI",   
                                  "PT-IBTS",   
                                  "S-GEORG",   
                                  "SCS-SUMMER",
                                  "SEUS-fall", 
                                  "SEUS-spring",
                                  "SEUS-summer",
                                  "WCANN",     
                                  "ZAF-ATL",   
                                  "ZAF-IND",
                                  "ROCKALL",
                                  "SWC-IBTS-1",
                                  "SWC-IBTS-4",
                                  "FALK",
                                  "GMEX-Fall",
                                  "Nor-BTS-1",
                                  "NZ-SUBA",
                                  "SCS-SPRING"
                          ))


color_link <- name_helper[color_link, on = "survey_unit"]

```


##Make GAMs

Bray Curtis
```{r bray curtis gams}
bray_curtis_total_gam <- gam(bray_curtis_dissimilarity_total_mean ~ year + s(survey_unit, year, bs = "fs", m = 1),#random smooth
                            data = distances_dissimilarities_allyears.r)

```


##Make LMERS

Bray
*These all converged*
```{r bray}
#running with lme instead of lmer gave same results, but allowed for calculation of p-value
bray_curtis_total_lme <- lme(bray_curtis_dissimilarity_total_mean ~ year_adj, random = (~1 + year_adj|survey_unit),data = distances_dissimilarities_allyears.r)

#but also run with lmer for confint
#total
bray_curtis_total_lmer <- lmer(bray_curtis_dissimilarity_total_mean ~ year_adj + (1 + year_adj|survey_unit),data = distances_dissimilarities_allyears.r)
#balanced (for comparison plot)
bray_curtis_balanced_lmer <- lmer(bray_curtis_dissimilarity_balanced_mean ~ year_adj + (1 + year_adj|survey_unit),data = distances_dissimilarities_allyears.r)
#gradient (for comparison plot)
bray_curtis_gradient_lmer <- lmer(bray_curtis_dissimilarity_gradient_mean ~ year_adj + (1 + year_adj|survey_unit),data = distances_dissimilarities_allyears.r)
#Jaccard total (for comparison plot)
jaccard_total_lmer <- lmer(jaccard_dissimilarity_total_mean ~ year_adj + (1 + year_adj|survey_unit),data = distances_dissimilarities_allyears.r)


summary(bray_curtis_total_lme)
anova(bray_curtis_total_lme)

bray_curtis_total_coefs <- data.table(coef(bray_curtis_total_lme))
bray_curtis_total_coefs[,survey_unit := rownames(coef(bray_curtis_total_lme))][,Year := round(year_adj,5)][,Intercept := round(`(Intercept)`,2)]
#View(bray_curtis_total_coefs)

bray_curtis_total_coefs <- bray_curtis_total_coefs[color_link, on = "survey_unit"]

bray_curtis_total_coefs.exp <- bray_curtis_total_coefs[,.(Survey_Name_Season, Intercept, Year)]

#export this table
fwrite(bray_curtis_total_coefs.exp, file = here::here("output","bray_curtis_total_coefs.exp.csv"))
```

Get LMER model as predictions
```{r TOTAL BC}

# need to sort out year in seq versus overall year models
#new data for lmer
lmer_year <- seq(min(distances_dissimilarities_allyears.r[,year]), max(distances_dissimilarities_allyears.r[,year]), by = 0.1)

lmer_year_adj <- seq(min(distances_dissimilarities_allyears.r[,year_adj]), max(distances_dissimilarities_allyears.r[,year_adj]), by = 0.1)

#predict average lmer
lmer_bray_total_predictions <- data.table(year = lmer_year, year_adj = lmer_year_adj)

#confidence intervals
bray_curtis_total_lmer_confint <- confint(bray_curtis_total_lmer)

#populate data table of lmer predictions
lmer_bray_total_predictions[,bray_curtis_lmer_preds := fixef(bray_curtis_total_lmer)[[1]] + fixef(bray_curtis_total_lmer)[[2]] * year_adj][,bray_curtis_lmer_preds_lowerCI := bray_curtis_total_lmer_confint[5] + bray_curtis_total_lmer_confint[6] * year_adj][,bray_curtis_lmer_preds_upperCI := bray_curtis_total_lmer_confint[11] + bray_curtis_total_lmer_confint[12] * year_adj]
```

```{r BALANCED BC}

# need to sort out year in seq versus overall year models
#new data for lmer


#predict average lmer
lmer_bray_balanced_predictions <- data.table(year = lmer_year, year_adj = lmer_year_adj)

#confidence intervals
bray_curtis_balanced_lmer_confint <- confint(bray_curtis_balanced_lmer)

#populate data table of lmer predictions
lmer_bray_balanced_predictions[,bray_curtis_lmer_preds := fixef(bray_curtis_balanced_lmer)[[1]] + fixef(bray_curtis_balanced_lmer)[[2]] * year_adj][,bray_curtis_lmer_preds_lowerCI := bray_curtis_balanced_lmer_confint[5] + bray_curtis_balanced_lmer_confint[6] * year_adj][,bray_curtis_lmer_preds_upperCI := bray_curtis_balanced_lmer_confint[11] + bray_curtis_balanced_lmer_confint[12] * year_adj]
```

```{r GRADIENT BC}

# need to sort out year in seq versus overall year models


#predict average lmer
lmer_bray_gradient_predictions <- data.table(year = lmer_year, year_adj = lmer_year_adj)

#confidence intervals
bray_curtis_gradient_lmer_confint <- confint(bray_curtis_gradient_lmer)

#populate data table of lmer predictions
lmer_bray_gradient_predictions[,bray_curtis_lmer_preds := fixef(bray_curtis_gradient_lmer)[[1]] + fixef(bray_curtis_gradient_lmer)[[2]] * year_adj][,bray_curtis_lmer_preds_lowerCI := bray_curtis_gradient_lmer_confint[5] + bray_curtis_gradient_lmer_confint[6] * year_adj][,bray_curtis_lmer_preds_upperCI := bray_curtis_gradient_lmer_confint[11] + bray_curtis_gradient_lmer_confint[12] * year_adj]
```

```{r total jaccard}

# need to sort out year in seq versus overall year models


#predict average lmer
lmer_jaccard_total_predictions <- data.table(year = lmer_year, year_adj = lmer_year_adj)

#confidence intervals
jaccard_total_lmer_confint <- confint(jaccard_total_lmer)

#populate data table of lmer predictions
lmer_jaccard_total_predictions[,jaccard_lmer_preds := fixef(jaccard_total_lmer)[[1]] + fixef(jaccard_total_lmer)[[2]] * year_adj][,jaccard_lmer_preds_lowerCI := jaccard_total_lmer_confint[5] + jaccard_total_lmer_confint[6] * year_adj][,jaccard_lmer_preds_upperCI := jaccard_total_lmer_confint[11] + jaccard_total_lmer_confint[12] * year_adj]
```
###Move forward with Bray Curtis total (others to supplement, although will include in plot below)


Coefficients for LMER by survey_unit
```{r}
#unique survey unit year combos
survey_unit_sampling_years <- unique(distances_dissimilarities_allyears.r[,.(survey_unit, year_adj, year, years_sampled)])

# see group coefficients
model_coefs_reduced <- data.table(transform(as.data.frame(ranef(bray_curtis_total_lmer)), lwr = condval - 1.96*condsd, upr = condval + 1.96*condsd))
#https://stackoverflow.com/questions/69805532/extract-the-confidence-intervals-of-lmer-random-effects-plotted-with-dotplotra


#ONLY SLOPES
model_coefs_reduced <- model_coefs_reduced[term == "year_adj",]

model_coefs_reduced[,survey_unit := grp][,year_adj := condval]

#merge with duration of survey
model_coefs_reduced_length <- model_coefs_reduced[survey_unit_sampling_years, on = "survey_unit"]


model_coefs_reduced_length[,years_sampled := as.numeric(years_sampled)][,Directional_Change := ifelse(year_adj > 0, "Differentiation","Homogenization")]

#does it cross zero?
model_coefs_reduced_length[,significant := ifelse(lwr >0 & upr>0,T,ifelse(lwr<0 & upr<0,T,F))]

#significant directional change
model_coefs_reduced_length[,Directional_Change_sig := ifelse(year_adj > 0 & significant == T, "Differentiation",ifelse(year_adj < 0 & significant == T, "Homogenization", "No trend in\ndissimilarity"))]


#min max distances_dissimilarities
min_bray_reduced <- min(distances_dissimilarities_allyears.r[,bray_curtis_dissimilarity_total_mean], na.rm = T)
max_bray_reduced <- max(distances_dissimilarities_allyears.r[,bray_curtis_dissimilarity_total_mean], na.rm = T)

model_coefs_reduced_length <- model_coefs_reduced_length[color_link, on = "survey_unit"]

#delete any NAs
model_coefs_reduced_length <- na.omit(model_coefs_reduced_length, cols = "significant")

#order table by coefficient
setorder(model_coefs_reduced_length, year_adj)

BC_total_model_coefs_reduced_length.unique <- unique(model_coefs_reduced_length[,.(condval,condsd, lwr, upr, survey_unit, year_adj, years_sampled, Directional_Change, hex, Survey_Name_Season, significant, Directional_Change_sig)]) 

#extract color hexes
#year adj coef order
color_year_adj_order <- BC_total_model_coefs_reduced_length.unique[,hex]

#alphabetical order
BC_total_model_coefs_reduced_length.unique.alpha <- setorder(BC_total_model_coefs_reduced_length.unique, Survey_Name_Season)

BC_total_model_coefs_reduced_length.unique.alpha[,trend_color := ifelse(Directional_Change_sig == "Homogenization", "#e7ac5b", ifelse(Directional_Change_sig == "Differentiation","#91c874","#cbbfde"))]

color_alpha_order <- BC_total_model_coefs_reduced_length.unique.alpha[,hex]
color_alpha_order_bytrend <- BC_total_model_coefs_reduced_length.unique.alpha[, trend_color]

saveRDS(BC_total_model_coefs_reduced_length.unique, here::here("output","region_stats","BC_total_model_coefs_reduced_length.unique.Rds"))
```

Total versus balanced BC plot
```{r}
ggplot(distances_dissimilarities_allyears.r) +
  geom_point(aes(bray_curtis_dissimilarity_total_mean, bray_curtis_dissimilarity_balanced_mean)) +
  geom_abline(slope = 1, intercpet = 0) +
  geom_smooth(aes(bray_curtis_dissimilarity_total_mean, bray_curtis_dissimilarity_balanced_mean)) +
  theme_classic() +
  labs(x = "Total BC dissimilarity",  y = "Balanced changes in abundance/biomass")

ggplot(distances_dissimilarities_allyears.r) +
  geom_point(aes(bray_curtis_dissimilarity_total_mean, bray_curtis_dissimilarity_gradient_mean)) +
  geom_abline(slope = 1, intercpet = 0) +
  geom_smooth(aes(bray_curtis_dissimilarity_total_mean, bray_curtis_dissimilarity_gradient_mean)) +
  theme_classic() +
  labs(x = "Total BC dissimilarity",  y = "Abundance/biomass gradient")

#all on one plot
#reduced
distances_dissimilarities_allyears.r.r <- distances_dissimilarities_allyears.r[,.(year, bray_curtis_dissimilarity_total_mean, bray_curtis_dissimilarity_gradient_mean, bray_curtis_dissimilarity_balanced_mean, jaccard_dissimilarity_total_mean)]

#wide to long
dissim_long <- melt(distances_dissimilarities_allyears.r.r, id.vars = c("year"), variable.name = "Dissimilarity metric")

dissim_long[,`Dissimilarity metric` := factor(`Dissimilarity metric`, levels = c( "bray_curtis_dissimilarity_total_mean","bray_curtis_dissimilarity_balanced_mean","bray_curtis_dissimilarity_gradient_mean", "jaccard_dissimilarity_total_mean"))]

BC_dissimilarity_components <- ggplot() +
  geom_point(data = dissim_long[`Dissimilarity metric` != "jaccard_dissimilarity_total_mean",], aes(x = year, y = value, color = `Dissimilarity metric`), alpha = 0.2, size = 1) +
  geom_ribbon(data = lmer_bray_total_predictions, aes(x = year, ymin = bray_curtis_lmer_preds_lowerCI, ymax = bray_curtis_lmer_preds_upperCI), fill = "black", alpha = 0.3) +
  geom_line(data = lmer_bray_total_predictions, aes(x = year, y = bray_curtis_lmer_preds), color = "black") +
  #balanced
    geom_ribbon(data = lmer_bray_gradient_predictions, aes(x = year, ymin = bray_curtis_lmer_preds_lowerCI, ymax = bray_curtis_lmer_preds_upperCI), fill = "cadetblue3", alpha = 0.3) +
  geom_line(data = lmer_bray_gradient_predictions, aes(x = year, y = bray_curtis_lmer_preds), color = "cadetblue3") +
  #gradient
    geom_ribbon(data = lmer_bray_balanced_predictions, aes(x = year, ymin = bray_curtis_lmer_preds_lowerCI, ymax = bray_curtis_lmer_preds_upperCI), fill = "darksalmon", alpha = 0.3) +
  geom_line(data = lmer_bray_balanced_predictions, aes(x = year, y = bray_curtis_lmer_preds), color = "darksalmon") +
  #fix. colors of points
  scale_color_manual(values = c("black","darksalmon","cadetblue3"), labels = c("Total","Balanced variation\nin abundance","Biomass gradient")) +
  labs(x = "Year", y = "Value") +
  lims(y = c(min(dissim_long$value), max(dissim_long$value))) +
  theme_classic() +
  theme(legend.position = c(0.17,0.37))

ggsave(BC_dissimilarity_components, path = here::here("figures"), filename=("BC_dissimilarity_components.jpg"), width = 6.5, height = 5, units = "in")

#Jaccard
Jaccard_dissimilarity <- ggplot() +
  geom_point(data = dissim_long[`Dissimilarity metric` == "jaccard_dissimilarity_total_mean"], aes(x = year, y = value), alpha = 0.2, size = 1) +
  geom_ribbon(data = lmer_jaccard_total_predictions, aes(x = year, ymin = jaccard_lmer_preds_lowerCI, ymax = jaccard_lmer_preds_upperCI), fill = "black", alpha = 0.3) +
  geom_line(data = lmer_jaccard_total_predictions, aes(x = year, y = jaccard_lmer_preds), color = "black") +
  labs(x = "Year", y = "Total Jaccard dissimilarity") +
  lims(y = c(min(dissim_long$value), max(dissim_long$value))) +
  theme_classic()

ggsave(Jaccard_dissimilarity, path = here::here("figures"), filename=("Jaccard_dissimilarity.jpg"), width = 6.5, height = 5, units = "in")

#merge
library(cowplot)
all_dissimilarities_over_time <- plot_grid(BC_dissimilarity_components + theme(legend.background = element_rect(fill = "transparent")),
                                           Jaccard_dissimilarity,
                                           ncol = 2, align = "hv", labels = c("a.","b."))

ggsave(all_dissimilarities_over_time, filename = "all_dissimilarities_over_time.jpg", path = here::here("figures"), width = 10, height = 4.5, units = "in")

  
```



Bar Plot Coefficient LMER
```{r bar plot of coefficients}
BC_total_Dissimilarity_Coef_errorbar_reduced <- ggplot() +
    geom_errorbar(data = model_coefs_reduced_length, aes(x = reorder(Survey_Name_Season, year_adj) , y = year_adj, label = Survey_Name_Season, ymin = lwr, ymax = upr), fill = "grey", width = 0) + #add confidence intervals
  geom_point(data = model_coefs_reduced_length, aes(x = reorder(Survey_Name_Season, year_adj) , y = year_adj, label = Survey_Name_Season, size = years_sampled, fill = Directional_Change_sig, color = Directional_Change_sig), stat = 'identity', shape = 21) +
  scale_fill_manual(values = c("white","black","grey"), name = "Dissimilarity trend", guide="none") +
  scale_color_manual(values = c("black","black","grey"), name = "Dissimilarity trend", guide="none") +
  scale_size_binned(range = c(1,8), name = "Survey period length\n(years)") +
  geom_hline(yintercept = 0) +
  scale_y_continuous(breaks = seq(-0.005, 0.0075, by = 0.0025), labels = c("-0.005","","0", "", "0.005",  "")) +
  xlab("Survey unit") +
  ylab("β-diversity trend") + #total BC dissimilarity trend
  coord_flip() +
  theme_classic() +
  theme(axis.text.y = element_text(colour = color_year_adj_order, face = "bold"), axis.title.y = element_blank(), axis.text.x = element_text(size = 15), axis.title.x = element_text(size = 15), legend.position = c(0.25,0.7), legend.direction = "vertical")

#pull legend for homogenization
directional_change_legend_plot <- BC_total_Dissimilarity_Coef_errorbar_reduced + 
  scale_fill_manual(values = c("white","black","grey"), name = "Dissimilarity trend") +
  scale_color_manual(values = c("black","black","grey"), name = "Dissimilarity trend") +
  scale_size_binned(range = c(1,8), name = "Survey period length", guide = "none") +
  theme(legend.position = "right", legend.background = element_rect(fill= "transparent")) +
  guides(colour = guide_legend(override.aes = list(size=6)))


BC_total_Dissimilarity_Coef_errorbar_reduced

#ALT grey scale
BC_total_Dissimilarity_Coef_errorbar_reduced_greyscale <- ggplot() +
    geom_errorbar(data = model_coefs_reduced_length, aes(x = reorder(Survey_Name_Season, year_adj) , y = year_adj, label = Survey_Name_Season, ymin = lwr, ymax = upr), fill = "grey", width = 0) + #add confidence intervals
  geom_point(data = model_coefs_reduced_length, aes(x = reorder(Survey_Name_Season, year_adj) , y = year_adj, label = Survey_Name_Season, size = years_sampled, fill = Directional_Change_sig, color = Directional_Change_sig), stat = 'identity', shape = 21) +
  scale_fill_manual(values = c("white","black","grey"), name = "Dissimilarity trend", guide="none") +
  scale_color_manual(values = c("black","black","grey"), name = "Dissimilarity trend", guide="none") +
  scale_size_binned(range = c(1,8), name = "Survey period length\n(years)") +
  geom_hline(yintercept = 0) +
  scale_y_continuous(breaks = seq(-0.005, 0.0075, by = 0.0025), labels = c("-0.005","","0", "", "0.005",  "")) +
  xlab("Survey unit") +
  ylab("β-diversity trend") + #total BC dissimilarity trend
  coord_flip() +
  theme_classic() +
  theme(axis.text.y = element_text(face = "bold"), axis.title.y = element_blank(), axis.text.x = element_text(size = 15), axis.title.x = element_text(size = 15), legend.position = c(0.25,0.7), legend.direction = "vertical")
```

Alternatively, we color  this plot by trend experienced

```{r}
#"#73BA4D","#E0962C","#cbbfde"

BC_total_Dissimilarity_Coef_errorbar_reduced_colorbytrend <- ggplot() +
    geom_errorbar(data = model_coefs_reduced_length, aes(x = reorder(Survey_Name_Season, year_adj) , y = year_adj, label = Survey_Name_Season, ymin = lwr, ymax = upr), fill = "grey", width = 0) + #add confidence intervals
  geom_point(data = model_coefs_reduced_length, aes(x = reorder(Survey_Name_Season, year_adj) , y = year_adj, label = Survey_Name_Season, size = years_sampled, color = Directional_Change_sig), stat = 'identity') +
  scale_color_manual(values = c("#73BA4D","#E0962C","#cbbfde"), name = "Dissimilarity trend", guide="none") +
  scale_size_binned(range = c(1,8), name = "Survey period length\n(years)") +
  geom_hline(yintercept = 0) +
  scale_y_continuous(breaks = seq(-0.005, 0.0075, by = 0.0025), labels = c("-0.005","","0", "", "0.005",  "")) +
  xlab("Survey unit") +
  ylab("β-diversity trend") + #total BC dissimilarity trend
  coord_flip() +
  theme_classic() +
  theme(axis.text.y = element_text(face = "bold"), axis.title.y = element_blank(), axis.text.x = element_text(size = 15), axis.title.x = element_text(size = 15), legend.position = c(0.3,0.8), legend.direction = "vertical", legend.text = element_text(size = 15), legend.title = element_text(size = 16))

directional_change_legend_plot_colorbytrend <- BC_total_Dissimilarity_Coef_errorbar_reduced_colorbytrend + 
  theme(legend.position = "right", legend.background = element_rect(fill= "transparent"), 
         legend.text = element_text(size = 15), legend.title = element_text(size = 16)) +
  guides(colour = guide_legend(override.aes = list(size=6)), size = "none")
```


Wavy Line Plot for GAMs

Generate predicted values
```{r generate predicted values GAM}

#add colors and names to full dissimilarity data table
distances_dissimilarities_allyears.r <- distances_dissimilarities_allyears.r[color_link, on = "survey_unit"]

#generate new data to smooth lines (need year and season survey combinations)
year_survey_unit_expand.dt <- data.table(survey_unit = as.character(NULL), year = as.numeric(NULL), year_adj = as.numeric(NULL ))

for (i in 1:length(survey_unit.list)) {
  #generate year vectors
  survey_unit_years <- unique(distances_dissimilarities_allyears.r[survey_unit == survey_unit.list[i],.(survey_unit, year, year_adj)])
  
  years <- seq(min(survey_unit_years[,year]), max(survey_unit_years[,year]), by = 0.1)
  
  year_adjust <- seq(min(survey_unit_years[,year_adj]), max(survey_unit_years[,year_adj]), by = 0.1)
  
  year_survey_unit_expand.dt_addition <- data.table(survey_unit = survey_unit.list[i], year = years, year_adj = year_adjust)
  
  year_survey_unit_expand.dt <- rbind(year_survey_unit_expand.dt, year_survey_unit_expand.dt_addition)
}

#add colors and names to full year and survey unit combination table
year_survey_unit_expand.dt <- year_survey_unit_expand.dt[color_link, on = "survey_unit"]
```



Get model as predictions
```{r}
#for plotting, get model as predictions
bray_curtis_total_gam_predictions <- predict(bray_curtis_total_gam, se.fit = TRUE, newdata = year_survey_unit_expand.dt)

#merge into table
year_survey_unit_expand.dt[,bray_glm_mod_fit := bray_curtis_total_gam_predictions$fit][,bray_glm_mod_fit_SE := bray_curtis_total_gam_predictions$se.fit]

```


Produce Plot of GAM and mean LMER line
```{r plot GAM and mean LMER lines}
points_wavylines_bray_total_year_reduced_gam_nolmer <- ggplot() +
 # geom_ribbon(data = lmer_bray_total_predictions, aes(x = year, ymin = bray_curtis_lmer_preds_lowerCI, ymax = bray_curtis_lmer_preds_upperCI), fill = "grey", alpha = 0.2) +
  geom_point(data = na.omit(distances_dissimilarities_allyears.r,cols = "year_adj"),
             aes(x = year,
                 y = bray_curtis_dissimilarity_total_mean,
                 color = Survey_Name_Season), alpha = 0.5, size = 1) +
    geom_line(data = na.omit(year_survey_unit_expand.dt,cols = "year_adj"),
             aes(x = year,
                 y = bray_glm_mod_fit,
                 color = Survey_Name_Season)) +
  geom_ribbon(data = na.omit(year_survey_unit_expand.dt,cols = "year_adj"), aes(x = year, ymin=bray_glm_mod_fit-bray_glm_mod_fit_SE, ymax=bray_glm_mod_fit+bray_glm_mod_fit_SE, fill =  Survey_Name_Season), alpha=0.1) + #add standard error
 # geom_line(data = lmer_bray_total_predictions, aes(x = year, y = bray_curtis_lmer_preds), color = "black") +
    scale_color_manual(values =  color_alpha_order, name = "Survey Unit") +
  scale_fill_manual(values =  color_alpha_order, guide = "none") +
  theme_classic() +
  lims(x = c(min(distances_dissimilarities_allyears.r[,year]),max(distances_dissimilarities_allyears.r[,year])),
       y = c(0.15,0.9)) +
  xlab("Year") +
ylab("total BC dissimilarity") +
  theme(legend.position = "null")

points_wavylines_bray_total_year_reduced_gam_nolmer

ggsave(points_wavylines_bray_total_year_reduced_gam_nolmer, path = here::here("figures"), filename ="points_wavylines_bray_total_year_reduced_gam_nolmer.jpg", height = 5, width = 5, unit = "in")

#with lmer

points_wavylines_bray_total_year_reduced_gam <- ggplot() +
  geom_ribbon(data = lmer_bray_total_predictions, aes(x = year, ymin = bray_curtis_lmer_preds_lowerCI, ymax = bray_curtis_lmer_preds_upperCI), fill = "grey", alpha = 0.3) +
  geom_point(data = na.omit(distances_dissimilarities_allyears.r, cols = "year_adj"),
             aes(x = year,
                 y = bray_curtis_dissimilarity_total_mean,
                 fill = Survey_Name_Season), alpha = 0.4, size = 1, shape = 21, color = "white") +
    geom_line(data = na.omit(year_survey_unit_expand.dt, cols = "year_adj"),
             aes(x = year,
                 y = bray_glm_mod_fit,
                 color = Survey_Name_Season), alpha = 0.6) +
  geom_ribbon(data = na.omit(year_survey_unit_expand.dt, cols = "year_adj"), aes(x = year, ymin=bray_glm_mod_fit-bray_glm_mod_fit_SE, ymax=bray_glm_mod_fit+bray_glm_mod_fit_SE, fill =  Survey_Name_Season), alpha=0.1) + #add standard error
  geom_line(data = lmer_bray_total_predictions, aes(x = year, y = bray_curtis_lmer_preds), color = "black") +
    scale_color_manual(values =  color_alpha_order, name = "Survey Unit") +
  scale_fill_manual(values =  color_alpha_order, guide = "none") +
  theme_classic() +
  lims(x = c(min(distances_dissimilarities_allyears.r[,year]),max(distances_dissimilarities_allyears.r[,year]))) +
  xlab("Year") +
ylab("β-diversity") +
  theme(legend.position = "null", axis.text = element_text(size = 15), axis.title = element_text(size = 15))

points_wavylines_bray_total_year_reduced_gam

ggsave(points_wavylines_bray_total_year_reduced_gam, path = here::here("figures"), filename ="points_wavylines_bray_total_year_reduced_gam.jpg", height = 6, width = 6, unit = "in")

#ALT
#plot all, but same color scheme (grey)
points_wavylines_bray_total_year_reduced_gam_greyscale <- ggplot() +
  geom_ribbon(data = lmer_bray_total_predictions, aes(x = year, ymin = bray_curtis_lmer_preds_lowerCI, ymax = bray_curtis_lmer_preds_upperCI), fill = "grey", alpha = 0.3) +
  geom_point(data = distances_dissimilarities_allyears.r,
             aes(x = year,
                 y = bray_curtis_dissimilarity_total_mean,
                 fill = Survey_Name_Season), alpha = 0.4, size = 1, shape = 21, color = "white") +
    geom_line(data = year_survey_unit_expand.dt,
             aes(x = year,
                 y = bray_glm_mod_fit,
                 color = Survey_Name_Season), alpha = 0.6) +
  geom_ribbon(data = year_survey_unit_expand.dt, aes(x = year, ymin=bray_glm_mod_fit-bray_glm_mod_fit_SE, ymax=bray_glm_mod_fit+bray_glm_mod_fit_SE, fill =  Survey_Name_Season), alpha=0.1) + #add standard error
  geom_line(data = lmer_bray_total_predictions, aes(x = year, y = bray_curtis_lmer_preds), color = "black") +
    scale_color_manual(values =  rep("black", times = length(unique(distances_dissimilarities_allyears.r$Survey_Name_Season))), name = "Survey Unit") +
  scale_fill_manual(values =  rep("black", times = length(unique(distances_dissimilarities_allyears.r$Survey_Name_Season))), guide = "none") +
  theme_classic() +
  lims(x = c(min(distances_dissimilarities_allyears.r[,year]),max(distances_dissimilarities_allyears.r[,year]))
       ) +
  xlab("Year") +
ylab("β-diversity") +
  theme(legend.position = "null", axis.text = element_text(size = 15), axis.title = element_text(size = 15))

points_wavylines_bray_total_year_reduced_gam_greyscale

ggsave(points_wavylines_bray_total_year_reduced_gam_greyscale, path = here::here("figures"), filename ="points_wavylines_bray_total_year_reduced_gam_greyscale.jpg", height = 6, width = 6, unit = "in")
```
Alternative, color by trend
```{r color wavy lines by trend}

points_wavylines_bray_total_year_reduced_gam_colorbytrend <- ggplot() +
  geom_ribbon(data = lmer_bray_total_predictions, aes(x = year, ymin = bray_curtis_lmer_preds_lowerCI, ymax = bray_curtis_lmer_preds_upperCI), fill = "grey", alpha = 0.3) +
  geom_point(data = na.omit(distances_dissimilarities_allyears.r, cols = "year_adj"),
             aes(x = year,
                 y = bray_curtis_dissimilarity_total_mean,
                 fill = Survey_Name_Season), alpha = 0.4, size = 1, shape = 21, color = "white") +
    geom_line(data = na.omit(year_survey_unit_expand.dt, cols = "year_adj"),
             aes(x = year,
                 y = bray_glm_mod_fit,
                 color = Survey_Name_Season), alpha = 0.6) +
  geom_ribbon(data = na.omit(year_survey_unit_expand.dt, cols = "year_adj"), aes(x = year, ymin=bray_glm_mod_fit-bray_glm_mod_fit_SE, ymax=bray_glm_mod_fit+bray_glm_mod_fit_SE, fill =  Survey_Name_Season), alpha=0.1) + #add standard error
  geom_line(data = lmer_bray_total_predictions, aes(x = year, y = bray_curtis_lmer_preds), color = "black") +
    scale_color_manual(values =  color_alpha_order_bytrend, name = "Survey Unit") +
  scale_fill_manual(values =  color_alpha_order_bytrend, guide = "none") +
  theme_classic() +
  lims(x = c(min(distances_dissimilarities_allyears.r[,year]),max(distances_dissimilarities_allyears.r[,year]))) +
  xlab("Year") +
ylab("β-diversity") +
  theme(legend.position = "null", axis.text = element_text(size = 15), axis.title = element_text(size = 15))

points_wavylines_bray_total_year_reduced_gam_colorbytrend

ggsave(points_wavylines_bray_total_year_reduced_gam_colorbytrend, path = here::here("figures"), filename ="points_wavylines_bray_total_year_reduced_gam_colorbytrend.jpg", height = 6, width = 6, unit = "in")
```

```{r plot each independently}
#plot each independently for supplement
#all survey names = 
all_survey_names <- sort(unique(color_link$Survey_Name_Season))
#list of plots
points_wavylines_bray_total_year_reduced_gam_individual <- list()
for (i in 1:length(all_survey_names)) {
points_wavylines_bray_total_year_reduced_gam_individual[[i]] <- ggplot() +
  geom_point(data = distances_dissimilarities_allyears.r[Survey_Name_Season == all_survey_names[i]],
             aes(x = year,
                 y = bray_curtis_dissimilarity_total_mean), alpha = 0.4, color = "black") +
    geom_line(data = year_survey_unit_expand.dt[Survey_Name_Season == all_survey_names[i]],
             aes(x = year,
                 y = bray_glm_mod_fit), alpha = 0.6) +
  geom_ribbon(data = year_survey_unit_expand.dt[Survey_Name_Season == all_survey_names[i]], aes(x = year, ymin=bray_glm_mod_fit-bray_glm_mod_fit_SE, ymax=bray_glm_mod_fit+bray_glm_mod_fit_SE), alpha=0.1) + #add standard error
  theme_classic() +
#  lims(x = c(min(distances_dissimilarities_allyears.r[Survey_Name_Season == all_survey_names[i],year]),max(distances_dissimilarities_allyears.r[Survey_Name_Season == all_survey_names[i],year])),
#       y = c(0.15,0.9)) +
  xlab("Year") +
ylab("beta-diversity") +
  facet_wrap(~Survey_Name_Season, ncol = 5) +
  theme(legend.position = "null", axis.text = element_text(size = 15), axis.title = element_text(size = 15))

print(points_wavylines_bray_total_year_reduced_gam_individual[[i]])

}
saveRDS(points_wavylines_bray_total_year_reduced_gam_individual, here::here("figures","points_wavylines_bray_total_year_reduced_gam_individual.Rds"))

#print to pdf
library(gridExtra)

ggsave(
   filename = here::here("figures","points_wavylines_bray_total_year_reduced_gam_individual.pdf"), 
   plot = marrangeGrob(points_wavylines_bray_total_year_reduced_gam_individual, nrow=1, ncol=1), 
   width = 8.5, height = 11
)

#Alternatively, split into 2 and use facet
#first 24
points_wavylines_bray_total_year_reduced_gam_individual_facet_1_24 <- ggplot() +
  geom_point(data = distances_dissimilarities_allyears.r[Survey_Name_Season  %in% all_survey_names[c(1:24)]],
             aes(x = year,
                 y = bray_curtis_dissimilarity_total_mean), alpha = 0.7) +
    geom_line(data = year_survey_unit_expand.dt[Survey_Name_Season   %in% all_survey_names[c(1:24)]],
             aes(x = year,
                 y = bray_glm_mod_fit)) +
  geom_ribbon(data = year_survey_unit_expand.dt[Survey_Name_Season  %in% all_survey_names[c(1:24)]],
aes(x = year, ymin=bray_glm_mod_fit-bray_glm_mod_fit_SE, ymax=bray_glm_mod_fit+bray_glm_mod_fit_SE), alpha=0.5) + #add standard error
  theme_classic() +
  xlab("Year") +
ylab("β-diversity") +
  facet_wrap(~Survey_Name_Season, ncol = 4, scales = "free") +
  theme(axis.text = element_text(size = 8), axis.title = element_text(size = 12))

ggsave(points_wavylines_bray_total_year_reduced_gam_individual_facet_1_24, path =  here::here("figures"), filename = "points_wavylines_bray_total_year_reduced_gam_individual_facet_1_24.png", height = 11, width = 9)

points_wavylines_bray_total_year_reduced_gam_individual_facet_25_42 <- ggplot() +
  geom_point(data = distances_dissimilarities_allyears.r[Survey_Name_Season  %in% all_survey_names[c(25:42)]],
             aes(x = year,
                 y = bray_curtis_dissimilarity_total_mean), alpha = 0.7) +
    geom_line(data = year_survey_unit_expand.dt[Survey_Name_Season   %in% all_survey_names[c(25:42)]],
             aes(x = year,
                 y = bray_glm_mod_fit)) +
  geom_ribbon(data = year_survey_unit_expand.dt[Survey_Name_Season  %in% all_survey_names[c(25:42)]],
aes(x = year, ymin=bray_glm_mod_fit-bray_glm_mod_fit_SE, ymax=bray_glm_mod_fit+bray_glm_mod_fit_SE), alpha=0.5) + #add standard error
  theme_classic() +
  xlab("Year") +
ylab("β-diversity") +
  facet_wrap(~Survey_Name_Season, ncol = 4, scales = "free") +
  theme(axis.text = element_text(size = 8), axis.title = element_text(size = 12))

ggsave(points_wavylines_bray_total_year_reduced_gam_individual_facet_25_42, path =  here::here("figures"), filename = "points_wavylines_bray_total_year_reduced_gam_individual_facet_25_42.png", height = 11, width = 9)


```

Region by region build up
```{r region build up}
#just EBS for personal interest
points_wavylines_bray_total_year_reduced_gam_EBS_only <- ggplot() +
  geom_point(data = distances_dissimilarities_allyears.r[survey_unit == "EBS",],
             aes(x = year,
                 y = bray_curtis_dissimilarity_total_mean), alpha = 0.5, size = 1
            # , 
            # color = BC_total_model_coefs_reduced_length.unique[Survey_Name_Season == "Eastern Bering Sea",hex]
             ) +
  #  geom_line(data = year_survey_unit_expand.dt[survey_unit == "EBS",],
   #          aes(x = year,
    #             y = bray_glm_mod_fit),
     #        color = BC_total_model_coefs_reduced_length.unique[Survey_Name_Season == "Eastern Bering Sea",hex]) +
 # geom_ribbon(data = year_survey_unit_expand.dt[survey_unit == "EBS",], aes(x = year, ymin=bray_glm_mod_fit-bray_glm_mod_fit_SE, ymax=bray_glm_mod_fit+bray_glm_mod_fit_SE), alpha=0.3, fill = BC_total_model_coefs_reduced_length.unique[Survey_Name_Season == "Eastern Bering Sea",hex]) + #add standard error
 # geom_line(data = lmer_bray_total_predictions, aes(x = year, y = bray_curtis_lmer_preds), color = "black") +
  #  scale_color_manual(name = "Survey Unit") +
 # scale_fill_manual(values =  palette_42, guide = "none") +
 #   lims(x = c(min(distances_dissimilarities_allyears.r[,year]),max(distances_dissimilarities_allyears.r[,year])),
 #      y = c(0.15,0.9)) +
  theme_classic() +
  xlab("Year") +
ylab("total BC dissimilarity")

points_wavylines_bray_total_year_reduced_gam_EBS_only

#just one region that's homogenizing (NZ WCSI)
points_wavylines_bray_total_year_reduced_gam_NZ_only <- ggplot() +
  geom_point(data = distances_dissimilarities_allyears.r[survey_unit == "NZ-WCSI",],
             aes(x = year,
                 y = bray_curtis_dissimilarity_total_mean), alpha = 0.5, size = 1, color = BC_total_model_coefs_reduced_length.unique[Survey_Name_Season == "W Coast S Island NZ",hex]) +
    geom_line(data = year_survey_unit_expand.dt[survey_unit == "NZ-WCSI",],
             aes(x = year,
                 y = bray_glm_mod_fit),
             color = BC_total_model_coefs_reduced_length.unique[Survey_Name_Season == "W Coast S Island NZ",hex]) +
  geom_ribbon(data = year_survey_unit_expand.dt[survey_unit == "NZ-WCSI",], aes(x = year, ymin=bray_glm_mod_fit-bray_glm_mod_fit_SE, ymax=bray_glm_mod_fit+bray_glm_mod_fit_SE), alpha=0.3, fill = BC_total_model_coefs_reduced_length.unique[Survey_Name_Season == "W Coast S Island NZ",hex]) + #add standard error
 # geom_line(data = lmer_bray_total_predictions, aes(x = year, y = bray_curtis_lmer_preds), color = "black") +
    scale_color_manual(name = "Survey Unit") +
 # scale_fill_manual(values =  palette_42, guide = "none") +
    lims(x = c(min(distances_dissimilarities_allyears.r[,year]),max(distances_dissimilarities_allyears.r[,year])),
       y = c(0.15,0.9)) +
  theme_classic() +
  xlab("Year") +
ylab("total BC dissimilarity")

points_wavylines_bray_total_year_reduced_gam_NZ_only

#just two regions, another that's differentiating (SEUS Summer)

points_wavylines_bray_total_year_reduced_gam_NZ_SEUS_Summer_only <- ggplot() +
  geom_point(data = distances_dissimilarities_allyears.r[survey_unit %in% c("NZ-WCSI","SEUS-summer"),],
             aes(x = year,
                 y = bray_curtis_dissimilarity_total_mean,
                 color = Survey_Name_Season), alpha = 0.5, size = 1) +
    geom_line(data = year_survey_unit_expand.dt[survey_unit %in% c("NZ-WCSI","SEUS-summer"),],
             aes(x = year,
                 y = bray_glm_mod_fit,
                 color = Survey_Name_Season)) +
  geom_ribbon(data = year_survey_unit_expand.dt[survey_unit %in% c("NZ-WCSI","SEUS-summer"),], aes(x = year, ymin=bray_glm_mod_fit-bray_glm_mod_fit_SE, ymax=bray_glm_mod_fit+bray_glm_mod_fit_SE, fill = survey_unit), alpha=0.3) + #add standard error
 # geom_line(data = lmer_bray_total_predictions, aes(x = year, y = bray_curtis_lmer_preds), color = "black") +
    scale_color_manual(name = "Survey Unit", values = c(BC_total_model_coefs_reduced_length.unique[Survey_Name_Season %in% c("W Coast S Island NZ","SE US Summer"),hex])) +
  scale_fill_manual(guide = "none", values = c(BC_total_model_coefs_reduced_length.unique[Survey_Name_Season %in% c("W Coast S Island NZ","SE US Summer"),hex])) +
    lims(x = c(min(distances_dissimilarities_allyears.r[,year]),max(distances_dissimilarities_allyears.r[,year])),
       y = c(0.15,0.9)) +
  theme_classic() +
  xlab("Year") +
ylab("total BC dissimilarity") +
  theme(legend.position = "null")

points_wavylines_bray_total_year_reduced_gam_NZ_SEUS_Summer_only

#just three regions, another one that's stable, (Canada Scotian Shelf)

points_wavylines_bray_total_year_reduced_gam_NZ_SEUS_ScoShelf_only <- ggplot() +
  geom_point(data = distances_dissimilarities_allyears.r[survey_unit %in% c("NZ-WCSI","SEUS-summer", "SCS-SUMMER"),],
             aes(x = year,
                 y = bray_curtis_dissimilarity_total_mean,
                 color = Survey_Name_Season), alpha = 0.5, size = 1) +
    geom_line(data = year_survey_unit_expand.dt[survey_unit %in% c("NZ-WCSI","SEUS-summer","SCS-SUMMER"),],
             aes(x = year,
                 y = bray_glm_mod_fit,
                 color = Survey_Name_Season)) +
  geom_ribbon(data = year_survey_unit_expand.dt[survey_unit %in% c("NZ-WCSI","SEUS-summer","SCS-SUMMER"),], aes(x = year, ymin=bray_glm_mod_fit-bray_glm_mod_fit_SE, ymax=bray_glm_mod_fit+bray_glm_mod_fit_SE, fill = Survey_Name_Season), alpha=0.3) + #add standard error
 # geom_line(data = lmer_bray_total_predictions, aes(x = year, y = bray_curtis_lmer_preds), color = "black") +
    scale_color_manual(name = "Survey Unit", values = c(BC_total_model_coefs_reduced_length.unique[Survey_Name_Season %in% c("W Coast S Island NZ","SE US Summer","Scotian Shelf"),hex])) +
  scale_fill_manual(guide = "none", values = c(BC_total_model_coefs_reduced_length.unique[Survey_Name_Season %in% c("W Coast S Island NZ","SE US Summer","Scotian Shelf"),hex])) +
    lims(x = c(min(distances_dissimilarities_allyears.r[,year]),max(distances_dissimilarities_allyears.r[,year])),
       y = c(0.15,0.9)) +
  theme_classic() +
  xlab("Year") +
ylab("total BC dissimilarity") +
  theme(legend.position = "null")

points_wavylines_bray_total_year_reduced_gam_NZ_SEUS_ScoShelf_only

ggsave(points_wavylines_bray_total_year_reduced_gam_NZ_only, path = here::here("figures"), filename ="points_wavylines_bray_total_year_reduced_gam_NZ_only.jpg", height = 5, width = 5, unit = "in")
ggsave(points_wavylines_bray_total_year_reduced_gam_EBS_only, path = here::here("figures"), filename ="points_wavylines_bray_total_year_reduced_gam_EBS_only.jpg", height = 5, width = 9, unit = "in")
ggsave(points_wavylines_bray_total_year_reduced_gam_NZ_SEUS_Summer_only, path = here::here("figures"), filename ="points_wavylines_bray_total_year_reduced_gam_NZ_SEUS_Summer_only.jpg", height = 5, width = 5, unit = "in")
ggsave(points_wavylines_bray_total_year_reduced_gam_NZ_SEUS_ScoShelf_only, path = here::here("figures"), filename ="points_wavylines_bray_total_year_reduced_gam_NZ_SEUS_ScoShelf_only.jpg", height = 5, width = 5, unit = "in")




```

Merge BC versus Year plot with GAMS and Region vs. coefficient plot for LMERs

```{r}

BC_total_GAM_LMER_merge_legend <- ggdraw(xlim = c(0, 40.5), ylim = c(0, 21)) +
    draw_plot(points_wavylines_bray_total_year_reduced_gam,
                                         x = 1, y = 1, width = 20, height = 20) +
    draw_plot(BC_total_Dissimilarity_Coef_errorbar_reduced +
        theme(legend.key.size = unit(0.5, 'cm'), #change legend key size
        legend.title = element_text(size=16), #change legend title font size
        legend.text = element_text(size=14)), #change legend text font size),
                                         x = 20, y = 1, width =19, height = 20) +
    draw_plot(get_legend(directional_change_legend_plot + 
      theme(legend.key.size = unit(0.5, 'cm'), #change legend key size
        legend.title = element_text(size=15), #change legend title font size
        legend.text = element_text(size=13))), #change legend text font size)
                                       x = 27, y = 10, width = 3, height = 2) +
  geom_text(aes(x = 2, y = 20.7), label = ("a."), size =8, fontface = "bold") +
  geom_text(aes(x =20, y = 20.7), label = ("b."), size =8, fontface = "bold")


ggsave(BC_total_GAM_LMER_merge_legend, path = here::here("figures"), filename = "BC_total_GAM_LMER_merge_legend.png", height = 8, width = 14, units = "in")

#ALT GREY SCALE
BC_total_GAM_LMER_merge_legend_greyscale <- ggdraw(xlim = c(0, 40.5), ylim = c(0, 21)) +
    draw_plot(points_wavylines_bray_total_year_reduced_gam_greyscale,
                                         x = 1, y = 1, width = 20, height = 20) +
    draw_plot(BC_total_Dissimilarity_Coef_errorbar_reduced_greyscale +
        theme(legend.key.size = unit(0.5, 'cm'), #change legend key size
        legend.title = element_text(size=16), #change legend title font size
        legend.text = element_text(size=14)), #change legend text font size),
                                         x = 20, y = 1, width = 19, height = 20) +
    draw_plot(get_legend(directional_change_legend_plot + 
      theme(legend.key.size = unit(0.5, 'cm'), #change legend key size
        legend.title = element_text(size=15), #change legend title font size
        legend.text = element_text(size=13))), #change legend text font size)
                                x = 27, y = 10, width = 3, height = 2) +
  geom_text(aes(x = 2, y = 20.7), label = ("a."), size =8, fontface = "bold") +
  geom_text(aes(x =20, y = 20.7), label = ("b."), size =8, fontface = "bold")

ggsave(BC_total_GAM_LMER_merge_legend_greyscale, path = here::here("figures"), filename = "BC_total_GAM_LMER_merge_legend_greyscale.png", height = 8, width = 14, units = "in")

#ALT COLOR BY TREND
BC_total_GAM_LMER_merge_legend_colorbytrend <- ggdraw(xlim = c(0, 40.5), ylim = c(0, 21)) +
    draw_plot(points_wavylines_bray_total_year_reduced_gam_colorbytrend,
                                         x = 1, y = 1, width = 20, height = 20) +
    draw_plot(BC_total_Dissimilarity_Coef_errorbar_reduced_colorbytrend +
        theme(legend.key.size = unit(0.5, 'cm'), #change legend key size
       # legend.title = element_text(size=16), #change legend title font size
       # legend.text = element_text(size=14)
       ), #change legend text font size),
                                         x = 20, y = 1, width = 19, height = 20) +
    draw_plot(get_legend(directional_change_legend_plot_colorbytrend + 
      theme(legend.key.size = unit(0.5, 'cm'), #change legend key size
        legend.title = element_text(size=16), #change legend title font size
        legend.text = element_text(size=15))), #change legend text font size)
                                x = 27, y = 8, width = 3, height = 2) +
  geom_text(aes(x = 2, y = 20.7), label = ("a."), size =8, fontface = "bold") +
  geom_text(aes(x =20, y = 20.7), label = ("b."), size =8, fontface = "bold")

ggsave(BC_total_GAM_LMER_merge_legend_colorbytrend, path = here::here("figures"), filename = "BC_total_GAM_LMER_merge_legend_colorbytrend.png", height = 8, width = 14, units = "in")

```

