---
title: "Dissimilarity time series analysis for reshuffled years"
author: Zoë J. Kitchel
date: October 17, 2023
output: html_notebook
---

Supplementary script 4b for Kitchel et al. 2023 in prep taxonomic diversity manuscript.

```{r setup}
library(data.table)
library(vegan)
library(sf)
library(concaveman) #polygon around points
library(betapart) #allows us to partition beta diversity
library(geosphere)
library(ggpubr) #stat_regline_equation
library(nlme)
library(mgcv) #to make gam
library(cowplot)
library(lme4)

#pull in function to calculate model estimates and standard errors
source(here::here("analysis_code","extract_coefficients_function.R"))

```

Pull dissimilarity table
```{r pull outputs from previous scripts}
#Pull Dissimilarity Means
jaccard_index_allyears <- readRDS(here::here("output","jaccard_index_VEGAN", "jaccard_index_allyears.rds"))

#make survey and survey unit factors
jaccard_index_allyears[,survey:=factor(survey)][,survey_unit:=factor(survey_unit)]

#adjust years
jaccard_index_allyears[,year_adj := year-min(year)+1]

#add new variable for year in sequence per region
jaccard_index_allyears[,first_year := min(year),.(survey_unit)]
jaccard_index_allyears[,last_year := max(year),.(survey_unit)]

#jaccard_index_allyears[,year_in_seq := year-first_year+1]

jaccard_index_allyears[,years_sampled := last_year-first_year+1]

#list of surveys
survey_unit.list <- levels(factor(jaccard_index_allyears$survey_unit))


```
 
 Pull in palette and name helper
```{r}
source(here::here("analysis_code","color_links.R"))
```

Load up 'true slopes' and CIs for comparison
```{r}

jaccard_total_coefs.r <- fread(here::here("output","jaccard_total_coefs.r.csv"))
jaccard_total_coefs.r_true <- jaccard_total_coefs.r
```


#NULL MODEL CONSTUCTION
##1) Reshuffle years (maintain correlation among species, removes temporal autocorrelation within a species)

I need to:
- reshuffle years
- reassign years
- run lmer
- pull mean and CI for each region
```{r}
#surveys
all_survey_units <- color_link$survey_unit
years <- sort(unique(jaccard_index_allyears[dissimilarity_metric == "jaccard_dissimilarity_index_binary",]$year))


for (i in 1:1000){

  jaccard_dissimilarity_time <- jaccard_index_allyears[dissimilarity_metric == "jaccard_dissimilarity_index_binary",.(survey_unit, year, year_adj, annual_dissimilarity_value)]
  
  year_run_key <- data.table(year = years, years_null = sample(years, length(years), replace = F))
  
  jaccard_dissimilarity_time  <- jaccard_dissimilarity_time[year_run_key, on = "year"]
  
  #run overall model
  jaccard_total_lm <- lm(annual_dissimilarity_value ~ years_null * survey_unit,data = jaccard_dissimilarity_time)
    
 #Use linear algebra model to calculate coefficients
  jaccard_total_coefs.r <- data.table(lm_interaction_coefficients_se(jaccard_total_lm))
    
  #homogenizing vs differentiating
  jaccard_total_coefs.r[,differentiating := ifelse((se< abs(estimate) & estimate > 0),1,0)][,homogenizing := ifelse((se< abs(estimate) & estimate < 0),1,0)]
  
    #full table with distribution of coefficients
    if(i == 1){
     jaccard_total_coefs.allruns <- jaccard_total_coefs.r }else{
    jaccard_total_coefs.allruns <- rbind(jaccard_total_coefs.allruns, jaccard_total_coefs.r)
   }
  
  
  #sum # differentiating vs homogenizing for each run
  jaccard_total_coefs_sum <- data.table(run = i, differentiating = sum(  jaccard_total_coefs.r$differentiating), homogenizing = sum(  jaccard_total_coefs.r$homogenizing))
  
  if(i == 1){
     all_null_runs_reshuffle_year <- jaccard_total_coefs_sum }else{
    all_null_runs_reshuffle_year <- rbind(all_null_runs_reshuffle_year, jaccard_total_coefs_sum)
   }
  print(i)
}

#95% quantile o(both count homogenization and count differentiation)

homogenization_95_quantile <- quantile(all_null_runs_reshuffle_year[,homogenizing], 0.95)
differentiation_95_quantile <- quantile(all_null_runs_reshuffle_year[,differentiating], 0.95)

#plot differentiation distribution
(reshuffle_differentiation_plot <- ggplot() +
  geom_histogram(data = all_null_runs_reshuffle_year, aes(differentiating), binwidth = 1) +
      geom_vline(xintercept = sum(jaccard_total_coefs.r_true$differentiating), color = "red") +
          geom_vline(xintercept = differentiation_95_quantile, linetype = "dashed") +
    scale_x_continuous(breaks = seq(0,16,by = 2)) +
  labs(x = "Number of surveys differentiating", y = "Count") +
  theme_classic())

#plot homogenization distribution
(reshuffle_homogenizing_plot <- ggplot() +
  geom_histogram(data = all_null_runs_reshuffle_year, aes(homogenizing), binwidth = 1) +
      geom_vline(xintercept = sum(jaccard_total_coefs.r_true$homogenizing), color = "red") +
          geom_vline(xintercept = homogenization_95_quantile, linetype = "dashed") +
    scale_x_continuous(breaks = seq(0,10,by = 2)) +
  labs(x = "Number of surveys homogenizing", y = "Count") +
  theme_classic())

reshuffle_null_plot <- cowplot::plot_grid(reshuffle_homogenizing_plot,reshuffle_differentiation_plot, 
                                     ncol = 2, labels = c("a.","b."))

ggsave(reshuffle_null_plot, path = here::here("figures"), filename = "reshuffle_null_plot.jpg", width = 6.5, height = 3)


#Plot the distribution of coefficients across all surveys versus the distribution of true coefficients
(reshuffle_slope_comparison_plot <- ggplot() +
  geom_histogram(data = jaccard_total_coefs.allruns, aes(estimate), fill = "#595959", binwidth = 0.001) +
  labs(x = "β diversity trend (null)", y = "Count (in 10,000s)") +
    scale_y_continuous(breaks = c(0,10000,20000,30000), labels = c("0","1","2","3")) +
    geom_vline(xintercept = median(jaccard_total_coefs.allruns$estimate)) +
   geom_vline(xintercept = quantile(jaccard_total_coefs.allruns$estimate, c(0.025,0.975)), linetype = "dashed") +
    xlim(min(c(jaccard_total_coefs.allruns$estimate,jaccard_total_coefs.r_true$estimate))-0.0035,
         max(c(jaccard_total_coefs.allruns$estimate,jaccard_total_coefs.r_true$estimate))+0.003) +
  theme_classic())

(true_slope_comparison_plot <- ggplot() +
  geom_histogram(data = jaccard_total_coefs.r_true, aes(estimate), fill = "#595959", binwidth = 0.001) +
  labs(x = "β diversity trend (observed)", y = "Count") +
    geom_vline(xintercept = median(jaccard_total_coefs.allruns$estimate)) +
    geom_vline(xintercept = quantile(jaccard_total_coefs.allruns$estimate, c(0.025,0.975)), linetype = "dashed") +
    xlim(min(c(jaccard_total_coefs.allruns$estimate,jaccard_total_coefs.r_true$estimate))-0.003,
         max(c(jaccard_total_coefs.allruns$estimate,jaccard_total_coefs.r_true$estimate))+0.003) +
  theme_classic())

#merge two plots
reshuffle_null_plot_bottom <- cowplot::plot_grid(true_slope_comparison_plot,reshuffle_slope_comparison_plot, 
                                     ncol = 1, labels = c("c.","d."), align = "h")

#merge all
reshuffle_null_plot_full <- cowplot::plot_grid(reshuffle_null_plot, reshuffle_null_plot_bottom, ncol = 1, rel_heights = c(1,2))

ggsave(reshuffle_null_plot_full, path = here::here("figures"), filename = "reshuffle_null_plot_full.jpg", width = 6.5, height = 6.5)


```
