---
title: "Dissimilarity time series analysis excluding 15% of species with lowest biomass"
author: Zoë J. Kitchel
date: May 9, 2024
output: html_notebook
---

Script 4e for Kitchel et al. 2024 in prep taxonomic diversity manuscript.

```{r setup}
library(data.table)
library(vegan)
library(sf)
library(concaveman) #polygon around points
library(betapart) #allows us to partition beta diversity
library(geosphere)
library(ggpubr) #stat_regline_equation
library(nlme)
library(mgcv) #to make gam
library(cowplot)
library(lme4)
library(stringr)
```


Pull dissimilarity table
```{r pull outputs from previous scripts}
#Pull Dissimilarity Means
distances_dissimilarities_allyears_15perc_excluded <- readRDS(here::here("output", "dissimilarities", "distances_dissimilarities_allyears_15perc_excluded.rds"))

#make survey and survey unit factors
distances_dissimilarities_allyears_15perc_excluded[,survey:=factor(survey)][,survey_unit:=factor(survey_unit)]

#adjust years
distances_dissimilarities_allyears_15perc_excluded[,year_adj := year-min(year)+1]

#add new variable for year in sequence per region
distances_dissimilarities_allyears_15perc_excluded[,first_year := min(year),.(survey_unit)]
distances_dissimilarities_allyears_15perc_excluded[,last_year := max(year),.(survey_unit)]

#distances_dissimilarities_allyears_15perc_excluded[,year_in_seq := year-first_year+1]

distances_dissimilarities_allyears_15perc_excluded[,years_sampled := last_year-first_year+1]


```

List of surveys
```{r}
#list of surveys
survey_unit.list <- levels(factor(distances_dissimilarities_allyears_15perc_excluded$survey_unit))
```

 
 Pull in palette and name helper
```{r}
source(here::here("analysis_code","color_links.R"))
```


##Make GAM (Generalized Additive Models)

Jaccard
```{r jaccard gams}
jaccard_total_gam <- gam(annual_dissimilarity_value ~ year + s(survey_unit, year, bs = "fs", m = 1),#random smooth
                            data = distances_dissimilarities_allyears_15perc_excluded[dissimilarity_metric == "jaccard_dissimilarity_total"])

```

##Make LMERS

Jaccard
*These all converged*
```{r jaccard}
#running with lme instead of lmer gave same results, but allowed for calculation of p-value

#Jaccard total with lme
jaccard_total_lme <- lme(annual_dissimilarity_value ~ year_adj, random = (~1 + year_adj|survey_unit),data = distances_dissimilarities_allyears_15perc_excluded[dissimilarity_metric == "jaccard_dissimilarity_total"])

#Jaccard total with lmer
jaccard_total_lmer <- lmer(annual_dissimilarity_value ~ year_adj + (1 + year_adj|survey_unit),data = distances_dissimilarities_allyears_15perc_excluded[dissimilarity_metric == "jaccard_dissimilarity_total"]) #FYI doesn't converge

summary(jaccard_total_lme)
anova(jaccard_total_lme)

#EXPORT JACCARD COEFFICIENTS
jaccard_total_coefs <- data.table(coef(jaccard_total_lme))
jaccard_total_coefs[,survey_unit := rownames(coef(jaccard_total_lme))][,Year := round(year_adj,5)][,Intercept := round(`(Intercept)`,2)]
#View(jaccard_total_coefs)

jaccard_total_coefs <- jaccard_total_coefs[color_link, on = "survey_unit"]

jaccard_total_coefs.exp_15perc_excluded <- jaccard_total_coefs[,.(Survey_Name_Season, Intercept, Year)]

#export this table
fwrite(jaccard_total_coefs.exp_15perc_excluded, file = here::here("output","jaccard_total_coefs.exp_15perc_excluded.csv"))
```
##Linear Model

Also, build a simple linear model with an interaction (instead of random slope and intercept for survey)
**THIS IS HOW WE CLASSIFY SURVEYS as of fall 2023 (update from before, where we classified surveys from LMER)**
(Especially good because lmer didn't converge)

Same, but for Jaccard
```{r linear model Jaccard}

jaccard_total_lm <- lm(annual_dissimilarity_value ~ year * survey_unit,data = distances_dissimilarities_allyears_15perc_excluded[dissimilarity_metric == "jaccard_dissimilarity_total"])

jaccard_total_coefs <- data.table(summary(jaccard_total_lm)$coefficients)
  jaccard_total_coefs[,var := rownames(summary(jaccard_total_lm)$coefficients)]
  
  #limit to interactions only (check this if there are any model changes!) row 2 and rows 36:68
  jaccard_total_coefs.r_15perc_excluded <- jaccard_total_coefs[c(2,36:68),]
  
  #adjust survey unit name by deleting beginning of string
  jaccard_total_coefs.r_15perc_excluded[,survey_unit := substr(var, 17, str_length(var))][var == "year",survey_unit := "AI"]
  
  #calculate interaction coefficients
  AI_estimate <- jaccard_total_coefs.r_15perc_excluded[1,Estimate]
  jaccard_total_coefs.r_15perc_excluded[1,survey_unit_coefficient := AI_estimate]
  jaccard_total_coefs.r_15perc_excluded[2:34,survey_unit_coefficient := (AI_estimate + Estimate)]
  
  #homogenizing vs differentiating
  jaccard_total_coefs.r_15perc_excluded[,differentiating := ifelse((`Std. Error`< abs(survey_unit_coefficient) & survey_unit_coefficient > 0),1,0)][,homogenizing := ifelse((`Std. Error`< abs(survey_unit_coefficient) & survey_unit_coefficient < 0),1,0)]
  
  #lower and upper CI
    jaccard_total_coefs.r_15perc_excluded[,lwr := survey_unit_coefficient-`Std. Error`][,upr:= survey_unit_coefficient+`Std. Error`]
    
    #save as output
    fwrite(jaccard_total_coefs.r_15perc_excluded, here::here("output","jaccard_total_coefs.r_15perc_excluded.csv"))
  
  jaccard_total_coefs_sum <- data.table(differentiating = sum(  jaccard_total_coefs.r_15perc_excluded$differentiating), homogenizing = sum(  jaccard_total_coefs.r_15perc_excluded$homogenizing))
  
  #hex by trend
  jaccard_total_coefs.r_15perc_excluded[,trend_color_jaccard := ifelse(homogenizing == 1,"#FFC109", ifelse(differentiating == 1,"#D81C60","#1E88E5"))]
  
  #link to color link
  jaccard_total_coefs.r_15perc_excluded <- color_link[jaccard_total_coefs.r_15perc_excluded, on = "survey_unit"]
  
  color_link_trend_jaccard <- jaccard_total_coefs.r_15perc_excluded[,.(survey_unit,Survey_Name_Season, hex,  trend_color_jaccard)]

```

##Get LMER models as predictions

```{r new data}

#new data for lmer
lmer_year <- seq(min(distances_dissimilarities_allyears_15perc_excluded[,year]), max(distances_dissimilarities_allyears_15perc_excluded[,year]), by = 0.1)

lmer_year_adj <- seq(min(distances_dissimilarities_allyears_15perc_excluded[,year_adj]), max(distances_dissimilarities_allyears_15perc_excluded[,year_adj]), by = 0.1)

```


```{r total jaccard}

#predict average lmer
lmer_jaccard_total_predictions <- data.table(year = lmer_year, year_adj = lmer_year_adj)

#confidence intervals
jaccard_total_lmer_confint <- confint(jaccard_total_lmer)

#populate data table of lmer predictions
lmer_jaccard_total_predictions[,jaccard_lmer_preds := fixef(jaccard_total_lmer)[[1]] + fixef(jaccard_total_lmer)[[2]] * year_adj][,jaccard_lmer_preds_lowerCI := jaccard_total_lmer_confint[5] + jaccard_total_lmer_confint[6] * year_adj][,jaccard_lmer_preds_upperCI := jaccard_total_lmer_confint[11] + jaccard_total_lmer_confint[12] * year_adj]
```


##Lollipop SE plot by linear trend experienced

Jaccard total dissimilarity
```{r lollipop figure jaccard total dissimilarity}


jaccard_total_coefs.r_15perc_excluded[,Directional_Change_sig := ifelse(differentiating > 0, "Differentiation",ifelse(homogenizing > 0 , "Homogenization", "No β-diversity trend"))]
years_sampled <- unique(distances_dissimilarities_allyears_15perc_excluded[,.(survey_unit, years_sampled)])
jaccard_total_coefs.r_15perc_excluded <- years_sampled[jaccard_total_coefs.r_15perc_excluded, on = "survey_unit"]

(jaccard_total_Dissimilarity_Coef_errorbar_reduced_colorbytrend_LINEAR_MODEL_15perc_excluded <- ggplot(data = jaccard_total_coefs.r_15perc_excluded, aes(x = reorder(Survey_Name_Season, survey_unit_coefficient) , y = survey_unit_coefficient), label = Survey_Name_Season) +
    geom_errorbar(aes(ymin = lwr, ymax = upr), fill = "grey", width = 0) + #add confidence intervals
  geom_point(aes(size = years_sampled, color = Directional_Change_sig), stat = 'identity') +
  scale_color_manual(values = c("#D81C60","#FFC109","#1E88E5"), name = "β-diversity trend\nTotal Jaccard dissimilarity", guide="none") +
  scale_size_binned(range = c(1,8), name = "Survey period length\n(years)") +
  geom_hline(yintercept = 0) +
  scale_y_continuous(breaks = seq(-0.0075, 0.0075, by = 0.0025), labels = c("-0.0075","","-0.0025","0", "0.0025", "",  "")) +
  xlab("Survey unit") +
  ylab("β-diversity trend\nTotal Jaccard dissimilarity") + #total jaccard dissimilarity trend
  coord_flip() +
  theme_classic() +
  theme(axis.text.y = element_text(face = "bold"), axis.title.y = element_blank(), axis.text.x = element_text(size = 15), axis.title.x = element_text(size = 15), legend.position = c(0.3,0.8), legend.direction = "vertical", legend.text = element_text(size = 15), legend.title = element_text(size = 16)))

ggsave(jaccard_total_Dissimilarity_Coef_errorbar_reduced_colorbytrend_LINEAR_MODEL_15perc_excluded, 
       path = here::here("figures"), filename = "jaccard_total_Dissimilarity_Coef_errorbar_reduced_colorbytrend_LINEAR_MODEL_15perc_excluded.jpg", height = 7, width = 7)

```

#Figure to merge three metrics for supplement



##Wavy Line Plot for GAMs (Figure 2a)

Generate predicted values (this datatable will hold both jaccard and bray curtis total dissimilarities)
```{r generate predicted values GAM}
#generate list again, as for some reason it doesn't stay in environment
survey_unit.list <- levels(factor(distances_dissimilarities_allyears_15perc_excluded$survey_unit))

#generate new data to smooth lines (need year and season survey combinations)
year_survey_unit_expand.dt <- data.table(survey_unit = as.character(NULL), year = as.numeric(NULL), year_adj = as.numeric(NULL ))

for (i in 1:length(survey_unit.list)) {
  #generate year vectors
  survey_unit_years <- unique(distances_dissimilarities_allyears_15perc_excluded[survey_unit == survey_unit.list[i],.(survey_unit, year, year_adj)])
  
  years <- seq(min(survey_unit_years[,year]), max(survey_unit_years[,year]), by = 0.1)
  
  year_adjust <- seq(min(survey_unit_years[,year_adj]), max(survey_unit_years[,year_adj]), by = 0.1)
  
  year_survey_unit_expand.dt_addition <- data.table(survey_unit = survey_unit.list[i], year = years, year_adj = year_adjust)
  
  year_survey_unit_expand.dt <- rbind(year_survey_unit_expand.dt, year_survey_unit_expand.dt_addition)
}


#add colors and names to full year and survey unit combination table
year_survey_unit_expand_jaccard.dt <- year_survey_unit_expand.dt[color_link_trend_jaccard, on = "survey_unit"]
```

Jaccard
Get model as predictions
```{r}
#for plotting, get model as predictions
jaccard_total_gam_predictions <- predict(jaccard_total_gam, se.fit = TRUE, newdata = year_survey_unit_expand_jaccard.dt)

#merge into table
year_survey_unit_expand_jaccard.dt[,jaccard_glm_mod_fit := jaccard_total_gam_predictions$fit][,jaccard_glm_mod_fit_SE := jaccard_total_gam_predictions$se.fit]

```


##Color GAMs by trend (year vs dissimilarity)
###Jaccard dissimilarity
```{r color wavy lines by trend Jaccard dissimilarity}

#link distances dissimilarities with trend color hex
distances_dissimilarities_allyears_15perc_excluded_jaccard <- distances_dissimilarities_allyears_15perc_excluded[color_link_trend_jaccard, on = "survey_unit"]
distances_dissimilarities_allyears_15perc_excluded_jaccard <- distances_dissimilarities_allyears_15perc_excluded_jaccard[dissimilarity_metric == "jaccard_dissimilarity_total_abundance"]
#and again for jaccard
distances_dissimilarities_allyears_15perc_excluded_jaccard <- distances_dissimilarities_allyears_15perc_excluded[color_link_trend_jaccard, on = "survey_unit"]
distances_dissimilarities_allyears_15perc_excluded_jaccard<- distances_dissimilarities_allyears_15perc_excluded_jaccard[dissimilarity_metric == "jaccard_dissimilarity_total"]

#change survey_unit to factor
distances_dissimilarities_allyears_15perc_excluded_jaccard[,survey_unit := factor(survey_unit)]
year_survey_unit_expand_jaccard.dt[,survey_unit := factor(survey_unit)]

#check that survey_unit in alpha order
sort(unique(distances_dissimilarities_allyears_15perc_excluded_jaccard$survey_unit)) == levels(distances_dissimilarities_allyears_15perc_excluded_jaccard$survey_unit)
sort(unique(year_survey_unit_expand_jaccard.dt$survey_unit)) == levels(year_survey_unit_expand_jaccard.dt$survey_unit)

#get list of colors in the right order
trend_colors_jaccard.dt <- unique(distances_dissimilarities_allyears_15perc_excluded_jaccard[dissimilarity_metric == "jaccard_dissimilarity_total",.(survey_unit, Survey_Name_Season, trend_color_jaccard)])

#make sure it matches year_survey_unit_expand.dt
all(trend_colors_jaccard.dt == unique(year_survey_unit_expand_jaccard.dt[,.(survey_unit, Survey_Name_Season, trend_color_jaccard)]))

setorder(trend_colors_jaccard.dt, survey_unit)


points_wavylines_jaccard_total_year_reduced_gam_colorbytrend_15perc_excluded <- ggplot() +
  #full LMER predictions (black)
  geom_ribbon(data = lmer_jaccard_total_predictions, aes(x = year, ymin = jaccard_lmer_preds_lowerCI, ymax = jaccard_lmer_preds_upperCI), fill = "grey", alpha = 0.3) +
    geom_line(data = lmer_jaccard_total_predictions, aes(x = year, y = jaccard_lmer_preds), color = "black") +
  #points by trend
  geom_point(data = distances_dissimilarities_allyears_15perc_excluded[dissimilarity_metric =="jaccard_dissimilarity_total"], cols = "year_adj",
             aes(x = year,
                 y = annual_dissimilarity_value,
                 fill = survey_unit), alpha = 0.4, size = 1, shape = 21, color = "white") +
  #squiggly lines by survey unit based GAM and colored by trend
    geom_line(data = year_survey_unit_expand_jaccard.dt, cols = "year_adj",
             aes(x = year,
                 y = jaccard_glm_mod_fit,
                 color = survey_unit), alpha = 0.6) +
  #confidence intervals for individual GAM
  geom_ribbon(data = year_survey_unit_expand_jaccard.dt, cols = "year_adj", 
              aes(x = year, ymin=jaccard_glm_mod_fit-jaccard_glm_mod_fit_SE, ymax=jaccard_glm_mod_fit+jaccard_glm_mod_fit_SE, fill =  survey_unit), alpha=0.1) + #add standard error
  #set color by survey
    scale_color_manual(values = trend_colors_jaccard.dt$trend_color, name = "Survey Unit") +
  scale_fill_manual(values =trend_colors_jaccard.dt$trend_color, guide = "none") +
  theme_classic() +
  lims(x = c(min(distances_dissimilarities_allyears_15perc_excluded[dissimilarity_metric == "jaccard_dissimilarity_total",year]),max(distances_dissimilarities_allyears_15perc_excluded[dissimilarity_metric == "jaccard_dissimilarity_total",year]))
       ) +
  xlab("Year\n") +
ylab("β-diversity\n(Jaccard dissimilarity)") +
  theme(legend.position = "null", axis.text = element_text(size = 15), axis.title = element_text(size = 15))

points_wavylines_jaccard_total_year_reduced_gam_colorbytrend_15perc_excluded

ggsave(points_wavylines_jaccard_total_year_reduced_gam_colorbytrend_15perc_excluded, path = here::here("figures"), filename ="points_wavylines_jaccard_total_year_reduced_gam_colorbytrend_15perc_excluded.jpg", height = 6, width = 11, unit = "in")

#plot for legend
directional_change_legend_plot_colorbytrend <- jaccard_total_Dissimilarity_Coef_errorbar_reduced_colorbytrend_LINEAR_MODEL_15perc_excluded + 
  theme(legend.position = "right", legend.background = element_rect(fill= "transparent"), 
         legend.text = element_text(size = 15), legend.title = element_text(size = 16)) +
  guides(colour = guide_legend(override.aes = list(size=6)), size = "none")

```

###Merge BC versus Year plot with GAMS and Region vs. coefficient plot for LMs

```{r merge lollipop and wiggly lines}

######################################################
##################JACCARDTOTAL
jaccard_total_GAM_LM_merge_legend_colorbytrend_15perc_excluded <- ggdraw(xlim = c(0, 40.5), ylim = c(0, 21)) +
    draw_plot(points_wavylines_jaccard_total_year_reduced_gam_colorbytrend_15perc_excluded + theme(axis.text = element_text(size = 14)),
                                         x = 1, y = 1, width = 20, height = 20) +
    draw_plot(jaccard_total_Dissimilarity_Coef_errorbar_reduced_colorbytrend_LINEAR_MODEL_15perc_excluded +
        theme(legend.key.size = unit(0.5, 'cm'), #change legend key size
              axis.text = element_text(size = 8)), #change legend text font size),
                                         x = 20, y = 1, width = 19, height = 20) +
    draw_plot(get_legend(directional_change_legend_plot_colorbytrend + 
      theme(legend.key.size = unit(0.5, 'cm'), #change legend key size
        legend.title = element_text(size=16), #change legend title font size
        legend.text = element_text(size=15))), #change legend text font size)
                                x = 26, y = 8, width = 3, height = 2) +
  geom_text(aes(x = 2, y = 20.7), label = ("a."), size =8, fontface = "bold") +
  geom_text(aes(x =20, y = 20.7), label = ("b."), size =8, fontface = "bold")

ggsave(jaccard_total_GAM_LM_merge_legend_colorbytrend_15perc_excluded, path = here::here("figures"), filename = "jaccard_total_GAM_LM_merge_legend_colorbytrend_15perc_excluded.png", height = 8, width = 14, units = "in")
```


###Is each survey better described by LM or by GAM?
####Also, compile predicted values from GAM
####Jaccard first
```{r compare lms to gams for each region jaccard}

#list of surveys
survey_unit.list_jaccard <- levels(factor(distances_dissimilarities_allyears_15perc_excluded_jaccard$survey_unit))

# Function to compare GAM and linear model fits
compare_models <- function(subset_data) {
  # Fit GAM model with "fs" basis and single smooth term
  gam_model <- gam(annual_dissimilarity_value ~ s(year, bs = "fs", m = 1), data = subset_data)
  
  # Fit linear model
  lm_model <- lm(annual_dissimilarity_value ~ year, data = subset_data)
  
  # Calculate AICc values
  aicc_gam <- AICc(gam_model)
  aicc_lm <- AICc(lm_model)
  
  #Calculate p-values
  gam_pvalue <- summary(gam_model)$s.table[1,4]
  lm_pvalue <- summary(lm_model)$coefficients[2,4]
  
  comparison <- ifelse((abs(aicc_gam-aicc_lm) < 2 | (aicc_gam == aicc_lm)),"Same",
                       ifelse(((abs(aicc_gam-aicc_lm) > 2 & (aicc_gam < aicc_lm))),"GAM",
                              ifelse(((abs(aicc_gam-aicc_lm) > 2 & (aicc_gam > aicc_lm))), "Linear",NA)))
  
   return(data.table(Survey = unique(subset_data$Survey_Name_Season)[1],
                     survey_unit = unique(subset_data$survey_unit),
                    comparison = comparison,
                    aicc_gam = aicc_gam,
                    gam_p = gam_pvalue,
                    aicc_lm = aicc_lm,
                    lm_p = lm_pvalue))
}


############## Compare models for each survey name and bind the results
GAM_LM_model_comparisons_jaccard_total <- data.table()

for (i in 1:length(survey_unit.list_jaccard)) {
  GAM_LM_model_comparisons_jaccard_total_single <- compare_models(distances_dissimilarities_allyears_15perc_excluded_jaccard[survey_unit == survey_unit.list_jaccard[i],])
  GAM_LM_model_comparisons_jaccard_total <- rbind(GAM_LM_model_comparisons_jaccard_total, GAM_LM_model_comparisons_jaccard_total_single)
}

GAM_LM_model_comparisons_jaccard_total

#round to 2 significant figures

#names of numeric columns
numeric_cols <- names(GAM_LM_model_comparisons_jaccard_total)[sapply(GAM_LM_model_comparisons_jaccard_total, is.numeric)]

# Apply signif() only to numeric columns
GAM_LM_model_comparisons_jaccard_total[, (numeric_cols) := lapply(.SD, function(x) if (is.numeric(x)) signif(x, digits = 2) else x), .SDcols = numeric_cols]

GAM_LM_model_comparisons_jaccard_total.f_15perc_excluded <- GAM_LM_model_comparisons_jaccard_total[,-2]

colnames(GAM_LM_model_comparisons_jaccard_total.f_15perc_excluded) <- c("Survey","Model comparison", "GAM AICc","GAM p-value","LM AICc", "LM p-value")

fwrite(GAM_LM_model_comparisons_jaccard_total.f_15perc_excluded, here::here("output","GAM_LM_model_comparisons_jaccard_total.f_15perc_excluded.csv"))


##################generate predicted GAM values
GAM_outputs <- function(subset.data) {
# Fit GAM model with "fs" basis and single smooth term
  gam_model <- gam(annual_dissimilarity_value ~ s(year, bs = "fs", m = 1), data = subset.data)
  
  pred_values <- predict(gam_model, se.fit = T)
  pred_table <- data.table(year = subset.data$year, predicted_dissimilarity = as.numeric(pred_values$fit), predicted_lwr = as.numeric(pred_values$fit)-as.numeric(pred_values$se.fit), predicted_upr = as.numeric(pred_values$fit)+as.numeric(pred_values$se.fit))
  
  st_dev_pred_dissim <- sd(pred_table$predicted_dissimilarity)
  return(signif(st_dev_pred_dissim,2))
}

predicted_values_jaccard <- data.table(survey_unit = survey_unit.list_jaccard, SD_pred_dissim = as.numeric())

for (i in 1:length(survey_unit.list_jaccard)) {
  st_dev_pred_dissim <- GAM_outputs(distances_dissimilarities_allyears_15perc_excluded_jaccard[survey_unit == survey_unit.list_jaccard[i]])
  predicted_values_jaccard[i,"SD_pred_dissim"] <- st_dev_pred_dissim
}

#merge witih GAM vs LM

GAM_lm_SD_jaccard_total <- predicted_values_jaccard[GAM_LM_model_comparisons_jaccard_total, on = "survey_unit"]

GAM_lm_SD_jaccard_total.f_15perc_excluded <- GAM_lm_SD_jaccard_total[,.(Survey, SD_pred_dissim, aicc_gam, gam_p, aicc_lm, lm_p, comparison)]

colnames(GAM_lm_SD_jaccard_total.f_15perc_excluded) <- c("Survey","SD over time of predicted GAM values", "GAM AICc","GAM p-value","LM AICc", "LM p-value","Model comparison")

View(GAM_lm_SD_jaccard_total.f_15perc_excluded)

fwrite(GAM_lm_SD_jaccard_total.f_15perc_excluded, here::here("output","GAM_lm_SD_jaccard_total.f_15perc_excluded.csv"))

```


###Model Comparison plots



Plot each jaccard total independently 

Note that this is for independent linear models and independent generalized additive models, NOT models built for all surveys at once
```{r plot each independently jaccard}
#plot each independently for supplement
#all survey names = 
all_survey_names <- color_link_trend_jaccard$survey_unit

#set order of levels of Survey Name Season by survey unit
distances_dissimilarities_allyears_15perc_excluded_jaccard[,Survey_Name_Season := factor(Survey_Name_Season, levels = trend_colors_jaccard.dt$Survey_Name_Season, ordered = T)]

#change year to integer
distances_dissimilarities_allyears_15perc_excluded_jaccard[,year:=as.integer(year)]

#split into 2 and use facet
#first 17
points_wavylines_jaccard_total_year_reduced_gam_colorbytrend_15perc_excluded_individual_facet_1_20 <- ggplot() +
  geom_point(data = distances_dissimilarities_allyears_15perc_excluded_jaccard[survey_unit  %in% all_survey_names[c(1:20)]],
             aes(x = as.integer(year),
                 y = annual_dissimilarity_value), alpha = 0.3) +

    geom_smooth(data = distances_dissimilarities_allyears_15perc_excluded_jaccard[survey_unit  %in% all_survey_names[c(1:20)]],
             aes(x = as.integer(year),
                 y = annual_dissimilarity_value,
                 color = survey_unit,
                 fill = survey_unit), method = "lm", linewidth = 0.5) +
  geom_smooth(data = distances_dissimilarities_allyears_15perc_excluded_jaccard[survey_unit %in% all_survey_names[c(1:20)]],
              aes(x = as.integer(year),
                  y = annual_dissimilarity_value), method = "gam", linewidth = 0.5, formula = y ~ s(x, bs = "fs", m = 1), color = "black") +
    scale_color_manual(values =    color_link_trend_jaccard$trend_color[1:20], name = "Survey Unit") +
    scale_fill_manual(values =    color_link_trend_jaccard$trend_color[1:20], name = "Survey Unit") +
  scale_x_continuous(breaks = ~ axisTicks(., log = FALSE)) +
  theme_classic() +
  theme(axis.text.x = element_text(size = 7)) +
  xlab("Year") +
ylab("β-diversity\n(Jaccard dissimilarity)") +
  facet_wrap(~Survey_Name_Season, ncol = 4, scales = "free") +
  theme(axis.text = element_text(size = 8), axis.title = element_text(size = 12), legend.position = "null")

ggsave(points_wavylines_jaccard_total_year_reduced_gam_colorbytrend_15perc_excluded_individual_facet_1_20, path =  here::here("figures"), filename = "points_wavylines_jaccard_total_year_reduced_gam_colorbytrend_15perc_excluded_individual_facet_1_20.png", height = 12, width = 9.5)

#second portion
points_wavylines_jaccard_total_year_reduced_gam_colorbytrend_15perc_excluded_individual_facet_21_34 <- ggplot() +
  geom_point(data = distances_dissimilarities_allyears_15perc_excluded_jaccard[survey_unit  %in% all_survey_names[c(21:34)]],
             aes(x = as.integer(year),
                 y = annual_dissimilarity_value), alpha = 0.3) +

    geom_smooth(data = distances_dissimilarities_allyears_15perc_excluded_jaccard[survey_unit  %in% all_survey_names[c(21:34)]],
             aes(x = as.integer(year),
                 y = annual_dissimilarity_value,
                 color = survey_unit,
                 fill = survey_unit), method = "lm", linewidth = 0.5) +
  geom_smooth(data = distances_dissimilarities_allyears_15perc_excluded_jaccard[survey_unit %in% all_survey_names[c(21:34)]],
              aes(x = as.integer(year),
                  y = annual_dissimilarity_value), method = "gam", linewidth = 0.5, formula = y ~ s(x, bs = "fs", m = 1), color = "black") +
    scale_color_manual(values =    color_link_trend_jaccard$trend_color[21:34], name = "Survey Unit") +
    scale_fill_manual(values =    color_link_trend_jaccard$trend_color[21:34], name = "Survey Unit") +
  scale_x_continuous(breaks = ~ axisTicks(., log = FALSE)) +
  theme_classic() +
  theme(axis.text.x = element_text(size = 7)) +
  xlab("Year") +
ylab("β-diversity\n(Jaccard dissimilarity)") +
  facet_wrap(~Survey_Name_Season, ncol = 4, scales = "free") +
  theme(axis.text = element_text(size = 8), axis.title = element_text(size = 12), legend.position = "null")

ggsave(points_wavylines_jaccard_total_year_reduced_gam_colorbytrend_15perc_excluded_individual_facet_21_34, path =  here::here("figures"), filename = "points_wavylines_jaccard_total_year_reduced_gam_colorbytrend_15perc_excluded_individual_facet_21_34.png", height = 10, width = 9.5)


```





