---
title: "Year Balanced Dissimilarity Models (replacement/turnover)"
output: html_notebook
---

This code is Script 4 v1 for Kitchel et al. TITLE manuscript.

(We expect this to be the main analyses in the text)

- This project is a collaborative effort to describe changes in taxonomic composition  of fish communities around the world--as sampled by bottom trawl surveys.

- Code by Zoë J. Kitchel

SESSION INFO TO DO

```{r setup}
library(data.table)
library(vegan)
library(sf)
library(concaveman) #polygon around points
library(betapart) #allows us to partition beta diversity
library(geosphere)
library(ggpubr) #stat_regline_equation
library(nlme)
library(mgcv) #to make gam
library(cowplot)
library(lme4)

#Pull Dissimilarity Means
distances_dissimilarities_allyears.r <- readRDS(here::here("output", "distance_decay", "distances_dissimilarities_allyears.r.rds"))

#make survey and survey unit factors
distances_dissimilarities_allyears.r[,survey:=factor(survey)][,survey_unit:=factor(survey_unit)]

#adjust years
distances_dissimilarities_allyears.r[,year_adj := year-min(year)+1]

#add new variable for year in sequence per region
distances_dissimilarities_allyears.r[,first_year := min(year),.(survey_unit)]
distances_dissimilarities_allyears.r[,last_year := max(year),.(survey_unit)]

#distances_dissimilarities_allyears.r[,year_in_seq := year-first_year+1]

distances_dissimilarities_allyears.r[,years_sampled := last_year-first_year+1]


```

###Palette for Plotting
Palette for plotting all 37 survey units
```{r link colors to survey units}
survey_unit.list <- levels(distances_dissimilarities_allyears.r[,survey_unit])

palette_37 <- c(
  "#5A5156", #AI
  "#F6222E", #CHL
  "#F8A19F", #DFO-NF
  "#16FF32", #DFO-QCS
  "#DF00DB", #BITS-1
  "#DB8EDA", #BITS-4
  "#325A9B", #EBS
  "#3283FE", #EVHOE
  "#FEAF16", #FR-CGFS
  "#1C8356", #GMEX-Summer
  "#C4451C", #GOA
  "#85660D", #GRL-DE
  "#B0009F", #GSL-N
  "#BF79B8", #GSL-S
  "#1CBE4F", #ICE-GFS
  "#782AB6", #IE-IGFS
  "#90AD1C", #MEDITS
  "#6B003A", #NAM
  "#A75B00", #NEUS-Fall
  "#E3B072", #NEUS-Spring
  "#02E8B6", #NIGFS-1
  "#97E7D5", #NIGFS-4
  "#B00068", #Nor-BTS-3
  "#00B9E3", #NS-IBTS-1
  "#95E2F4", #NS-IBTS-3
  "#B3CE73", #NZ-CHAT
  "#689500", #NZ-ECSI
  "#AAF400", #NZ-WCSI
  "#AA0DFE", #PT-IBTS
  "#FA0087", #S-GEORG
  "#DEA0FD", #SCS-Summer
  "#FCEF88", #SEUS-fall
  "#A59405", #SEUS-spring
  "#FCE100", #SEUS-summer
  "#C075A6", #WCANN
  "#BDCDFF", #ZAF-ATL
  "#003EFF"  #ZAF-IND
)

color_link <- data.table(survey_unit = survey_unit.list,hex = palette_37)
```

Add names for plotting
```{r add names for plotting}

name_helper <- data.table(Survey_Name_Season = c("Aleutian Islands",
                                    "Baltic Sea Q1",
                                    "Baltic Sea Q4",
                                    "Chile",
                                    "Newfoundland",
                                    "Queen Charlotte Sound",
                                    "Eastern Bering Sea",
                                    "Bay of Biscay",
                                    "English Channel",
                                    "Gulf of Mexico",
                                    "Gulf of Alaska",
                                    "Greenland",
                                    "N Gulf of St. Lawrence",
                                    "S Gulf of St. Lawrence",
                                    "Iceland",
                                    "Irish Sea",
                                    "Mediterranean",
                                    "Namibia",
                                    "NE US Fall",
                                    "NE US Spring",
                                    "N Ireland Q1",
                                    "N Ireland Q4",
                                    "Norway",
                                    "N Sea Q1",
                                    "N Sea Q3",
                                    "Chatham Rise",
                                    "E Coast S Island NZ",
                                    "W Coast S Island NZ",
                                    "Portugal",
                                    "S Georgia Straight",
                                  "Scotian Shelf",
                                  "SE US Fall",
                                  "SE US Spring",
                                  "SE US Summer",
                                  "W Coast US",
                                  "Atlantic Ocean ZA",
                                  "Indian Ocean ZA"),
                          survey_unit = c(
                                  "AI",        
                                  "BITS-1",    
                                  "BITS-4",    
                                  "CHL",       
                                  "DFO-NF",    
                                  "DFO-QCS",   
                                  "EBS",       
                                  "EVHOE",     
                                  "FR-CGFS",   
                                  "GMEX-Summer",
                                  "GOA",       
                                  "GRL-DE",    
                                  "GSL-N",     
                                  "GSL-S",     
                                  "ICE-GFS",   
                                  "IE-IGFS",   
                                  "MEDITS",    
                                  "NAM",       
                                  "NEUS-Fall", 
                                  "NEUS-Spring",
                                  "NIGFS-1",   
                                  "NIGFS-4",   
                                  "Nor-BTS-3", 
                                  "NS-IBTS-1", 
                                  "NS-IBTS-3", 
                                  "NZ-CHAT",   
                                  "NZ-ECSI",   
                                  "NZ-WCSI",   
                                  "PT-IBTS",   
                                  "S-GEORG",   
                                  "SCS-SUMMER",
                                  "SEUS-fall", 
                                  "SEUS-spring",
                                  "SEUS-summer",
                                  "WCANN",     
                                  "ZAF-ATL",   
                                  "ZAF-IND"   
                          ))

color_link <- color_link[name_helper, on = "survey_unit"]

```



##Make GAMs

Bray Curtis
```{r bray curtis gams}
bray_curtis_balanced_gam <- gam(bray_curtis_dissimilarity_balanced_mean ~ year + s(survey_unit, year, bs = "fs", m = 1),#random smooth
                            data = distances_dissimilarities_allyears.r)

```


##Make LMERS

Bray
*These all converged*
```{r bray}
#running with lme instead of lmer gave same results, but allowed for calculation of p-value
bray_curtis_balanced_lme <- lme(bray_curtis_dissimilarity_balanced_mean ~ year_adj, random = (~1 + year_adj|survey_unit),data = distances_dissimilarities_allyears.r)

#but also run with lmer for confint
bray_curtis_balanced_lmer <- lmer(bray_curtis_dissimilarity_balanced_mean ~ year_adj + (1 + year_adj|survey_unit),data = distances_dissimilarities_allyears.r)

summary(bray_curtis_balanced_lme)
anova(bray_curtis_balanced_lme)

bray_curtis_balanced_coefs <- data.table(coef(bray_curtis_balanced_lme))
bray_curtis_balanced_coefs[,survey_unit := rownames(coef(bray_curtis_balanced_lme))][,Year := round(year_adj,5)][,Intercept := round(`(Intercept)`,2)]
View(bray_curtis_balanced_coefs)

bray_curtis_balanced_coefs <- bray_curtis_balanced_coefs[color_link, on = "survey_unit"]

bray_curtis_balanced_coefs.exp <- bray_curtis_balanced_coefs[,.(Survey_Name_Season, Intercept, Year)]

#export this table
fwrite(bray_curtis_balanced_coefs.exp, file = here::here("output","bray_curtis_balanced_coefs.exp.csv"))
```

Get LMER model as predictions
```{r}

# need to sort out year in seq versus overall year models
#new data for lmer
lmer_year <- seq(min(distances_dissimilarities_allyears.r[,year]), max(distances_dissimilarities_allyears.r[,year]), by = 0.1)

lmer_year_adj <- seq(min(distances_dissimilarities_allyears.r[,year_adj]), max(distances_dissimilarities_allyears.r[,year_adj]), by = 0.1)

#predict average lmer
lmer_bray_balanced_predictions <- data.table(year = lmer_year, year_adj = lmer_year_adj)

#confidence intervals
bray_curtis_balanced_lmer_confint <- confint(bray_curtis_balanced_lmer)

#populate data table of lmer predictions
lmer_bray_balanced_predictions[,bray_curtis_lmer_preds := fixef(bray_curtis_balanced_lmer)[[1]] + fixef(bray_curtis_balanced_lmer)[[2]] * year_adj][,bray_curtis_lmer_preds_lowerCI := bray_curtis_balanced_lmer_confint[5] + bray_curtis_balanced_lmer_confint[6] * year_adj][,bray_curtis_lmer_preds_upperCI := bray_curtis_balanced_lmer_confint[11] + bray_curtis_balanced_lmer_confint[12] * year_adj]
```



###Move forward with Bray Curtis Balanced (others to supplement)


Coefficients for LMER by survey_unit
```{r}
#unique survey unit year combos
survey_unit_sampling_years <- unique(distances_dissimilarities_allyears.r[,.(survey_unit, year_adj, year, years_sampled)])

# see group coefficients
model_coefs_reduced <- data.table(transform(as.data.frame(ranef(bray_curtis_balanced_lmer)), lwr = condval - 1.96*condsd, upr = condval + 1.96*condsd))
#https://stackoverflow.com/questions/69805532/extract-the-confidence-intervals-of-lmer-random-effects-plotted-with-dotplotra


#ONLY SLOPES
model_coefs_reduced <- model_coefs_reduced[term == "year_adj",]

model_coefs_reduced[,survey_unit := grp][,year_adj := condval]

#merge with duration of survey
model_coefs_reduced_length <- model_coefs_reduced[survey_unit_sampling_years, on = "survey_unit"]


model_coefs_reduced_length[,years_sampled := as.numeric(years_sampled)][,Directional_Change := ifelse(year_adj > 0, "Differentiation","Homogenization")]

#min max distances_dissimilarities
min_bray_reduced <- min(distances_dissimilarities_allyears.r[,bray_curtis_dissimilarity_balanced_mean], na.rm = T)
max_bray_reduced <- max(distances_dissimilarities_allyears.r[,bray_curtis_dissimilarity_balanced_mean], na.rm = T)

model_coefs_reduced_length <- model_coefs_reduced_length[color_link, on = "survey_unit"]


#does it cross zero?
model_coefs_reduced_length[,significant := ifelse(lwr >0 & upr>0,T,ifelse(lwr<0 & upr<0,T,F))]

#delete all obs that are significant
model_coefs_reduced_length.r <- model_coefs_reduced_length[significant == F,]

#order table by coefficient
setorder(model_coefs_reduced_length, year_adj)

BC_balanced_model_coefs_reduced_length.unique <- unique(model_coefs_reduced_length[,.(condval,condsd, lwr, upr, survey_unit, year_adj, years_sampled, Directional_Change, hex, Survey_Name_Season, significant)]) 

#extract color hexes
#year adj coef order
color_year_adj_order <- BC_balanced_model_coefs_reduced_length.unique[,hex]

#alphabetical order
BC_balanced_model_coefs_reduced_length.unique.alpha <- setorder(BC_balanced_model_coefs_reduced_length.unique, Survey_Name_Season)
color_alpha_order <- BC_balanced_model_coefs_reduced_length.unique.alpha[,hex]

saveRDS(BC_balanced_model_coefs_reduced_length.unique, here::here("output","region_stats","BC_balanced_model_coefs_reduced_length.unique.Rds"))
```

Bar Plot Coefficient LMER
```{r bar plot of coefficients}
BC_BALANCED_Dissimilarity_Coef_errorbar_reduced <- ggplot() +
    geom_errorbar(data = model_coefs_reduced_length, aes(x = reorder(Survey_Name_Season, year_adj) , y = year_adj, label = Survey_Name_Season, ymin = lwr, ymax = upr), fill = "grey", width = 0) + #add confidence intervals
  geom_point(data = model_coefs_reduced_length, aes(x = reorder(Survey_Name_Season, year_adj) , y = year_adj, label = Survey_Name_Season, size = years_sampled, fill = Directional_Change), stat = 'identity', shape = 21, color = "black") +
  scale_fill_manual(values = c("white","black"), name = "Directional change", guide="none") +
  geom_point(data = model_coefs_reduced_length.r, aes(x = reorder(Survey_Name_Season, year_adj) , y = year_adj, label = Survey_Name_Season, size = years_sampled), stat = 'identity', fill = "grey",color = "grey", shape = 21) +
  scale_size_binned(range = c(1,8), name = "Survey period length") +
  geom_hline(yintercept = 0) +
  scale_y_continuous(breaks = seq(-0.005, 0.0075, by = 0.0025), labels = c("-0.005","","0", "", "0.005",  "")) +
  xlab("Survey unit") +
  ylab("β-diversity trend") + #balanced BC dissimilarity trend
  coord_flip() +
  theme_classic() +
  theme(axis.text.y = element_text(colour = color_year_adj_order, face = "bold"), axis.title.y = element_blank(), axis.text.x = element_text(size = 15), axis.title.x = element_text(size = 15), legend.position = "top")

#pull legend for homogenization
directional_change_legend_plot <- BC_BALANCED_Dissimilarity_Coef_errorbar_reduced + 
  scale_fill_manual(values = c("white","black"), name = "Directional change") +
  scale_size_binned(range = c(1,8), name = "Survey period length", guide = "none") +
  theme(legend.position = "right")


BC_BALANCED_Dissimilarity_Coef_errorbar_reduced
```

Wavy Line Plot for GAMs

Generate predicted values
```{r generate predicted values GAM}

#add colors and names to full dissimilarity data table
distances_dissimilarities_allyears.r <- distances_dissimilarities_allyears.r[color_link, on = "survey_unit"]

#generate new data to smooth lines (need year and season survey combinations)
year_survey_unit_expand.dt <- data.table(survey_unit = as.character(NULL), year = as.numeric(NULL), year_adj = as.numeric(NULL ))

for (i in 1:length(survey_unit.list)) {
  #generate year vectors
  survey_unit_years <- unique(distances_dissimilarities_allyears.r[survey_unit == survey_unit.list[i],.(survey_unit, year, year_adj)])
  
  years <- seq(min(survey_unit_years[,year]), max(survey_unit_years[,year]), by = 0.1)
  
  year_adjust <- seq(min(survey_unit_years[,year_adj]), max(survey_unit_years[,year_adj]), by = 0.1)
  
  year_survey_unit_expand.dt_addition <- data.table(survey_unit = survey_unit.list[i], year = years, year_adj = year_adjust)
  
  year_survey_unit_expand.dt <- rbind(year_survey_unit_expand.dt, year_survey_unit_expand.dt_addition)
}

#add colors and names to full year and survey unit combination table
year_survey_unit_expand.dt <- year_survey_unit_expand.dt[color_link, on = "survey_unit"]
```



Get model as predictions
```{r}
#for plotting, get model as predictions
bray_curtis_balanced_gam_predictions <- predict(bray_curtis_balanced_gam, se.fit = TRUE, newdata = year_survey_unit_expand.dt)

#merge into table
year_survey_unit_expand.dt[,bray_glm_mod_fit := bray_curtis_balanced_gam_predictions$fit][,bray_glm_mod_fit_SE := bray_curtis_balanced_gam_predictions$se.fit]

```


Produce Plot of GAM and mean LMER line
```{r plot GAM and mean LMER lines}
points_wavylines_bray_balanced_year_reduced_gam_nolmer <- ggplot() +
 # geom_ribbon(data = lmer_bray_balanced_predictions, aes(x = year, ymin = bray_curtis_lmer_preds_lowerCI, ymax = bray_curtis_lmer_preds_upperCI), fill = "grey", alpha = 0.2) +
  geom_point(data = distances_dissimilarities_allyears.r,
             aes(x = year,
                 y = bray_curtis_dissimilarity_balanced_mean,
                 color = Survey_Name_Season), alpha = 0.5, size = 1) +
    geom_line(data = year_survey_unit_expand.dt,
             aes(x = year,
                 y = bray_glm_mod_fit,
                 color = Survey_Name_Season)) +
  geom_ribbon(data = year_survey_unit_expand.dt, aes(x = year, ymin=bray_glm_mod_fit-bray_glm_mod_fit_SE, ymax=bray_glm_mod_fit+bray_glm_mod_fit_SE, fill =  Survey_Name_Season), alpha=0.1) + #add standard error
 # geom_line(data = lmer_bray_balanced_predictions, aes(x = year, y = bray_curtis_lmer_preds), color = "black") +
    scale_color_manual(values =  color_alpha_order, name = "Survey Unit") +
  scale_fill_manual(values =  color_alpha_order, guide = "none") +
  theme_classic() +
  lims(x = c(min(distances_dissimilarities_allyears.r[,year]),max(distances_dissimilarities_allyears.r[,year])),
       y = c(0.15,0.9)) +
  xlab("Year") +
ylab("Balanced BC dissimilarity") +
  theme(legend.position = "null")

points_wavylines_bray_balanced_year_reduced_gam_nolmer

ggsave(points_wavylines_bray_balanced_year_reduced_gam_nolmer, path = here::here("figures"), filename ="points_wavylines_bray_balanced_year_reduced_gam_nolmer.jpg", height = 5, width = 5, unit = "in")

#with lmer

points_wavylines_bray_balanced_year_reduced_gam <- ggplot() +
  geom_ribbon(data = lmer_bray_balanced_predictions, aes(x = year, ymin = bray_curtis_lmer_preds_lowerCI, ymax = bray_curtis_lmer_preds_upperCI), fill = "grey", alpha = 0.3) +
  geom_point(data = distances_dissimilarities_allyears.r,
             aes(x = year,
                 y = bray_curtis_dissimilarity_balanced_mean,
                 fill = Survey_Name_Season), alpha = 0.4, size = 1, shape = 21, color = "white") +
    geom_line(data = year_survey_unit_expand.dt,
             aes(x = year,
                 y = bray_glm_mod_fit,
                 color = Survey_Name_Season), alpha = 0.6) +
  geom_ribbon(data = year_survey_unit_expand.dt, aes(x = year, ymin=bray_glm_mod_fit-bray_glm_mod_fit_SE, ymax=bray_glm_mod_fit+bray_glm_mod_fit_SE, fill =  Survey_Name_Season), alpha=0.1) + #add standard error
  geom_line(data = lmer_bray_balanced_predictions, aes(x = year, y = bray_curtis_lmer_preds), color = "black") +
    scale_color_manual(values =  color_alpha_order, name = "Survey Unit") +
  scale_fill_manual(values =  color_alpha_order, guide = "none") +
  theme_classic() +
  lims(x = c(min(distances_dissimilarities_allyears.r[,year]),max(distances_dissimilarities_allyears.r[,year])),
       y = c(0.15,0.9)) +
  xlab("Year") +
ylab("β-diversity") +
  theme(legend.position = "null", axis.text = element_text(size = 15), axis.title = element_text(size = 15))

points_wavylines_bray_balanced_year_reduced_gam

ggsave(points_wavylines_bray_balanced_year_reduced_gam, path = here::here("figures"), filename ="points_wavylines_bray_balanced_year_reduced_gam.jpg", height = 6, width = 6, unit = "in")
```

Region by region build up
```{r region build up}
#just EBS for personal interest
points_wavylines_bray_balanced_year_reduced_gam_EBS_only <- ggplot() +
  geom_point(data = distances_dissimilarities_allyears.r[survey_unit == "EBS",],
             aes(x = year,
                 y = bray_curtis_dissimilarity_balanced_mean), alpha = 0.5, size = 1
            # , 
            # color = BC_balanced_model_coefs_reduced_length.unique[Survey_Name_Season == "Eastern Bering Sea",hex]
             ) +
  #  geom_line(data = year_survey_unit_expand.dt[survey_unit == "EBS",],
   #          aes(x = year,
    #             y = bray_glm_mod_fit),
     #        color = BC_balanced_model_coefs_reduced_length.unique[Survey_Name_Season == "Eastern Bering Sea",hex]) +
 # geom_ribbon(data = year_survey_unit_expand.dt[survey_unit == "EBS",], aes(x = year, ymin=bray_glm_mod_fit-bray_glm_mod_fit_SE, ymax=bray_glm_mod_fit+bray_glm_mod_fit_SE), alpha=0.3, fill = BC_balanced_model_coefs_reduced_length.unique[Survey_Name_Season == "Eastern Bering Sea",hex]) + #add standard error
 # geom_line(data = lmer_bray_balanced_predictions, aes(x = year, y = bray_curtis_lmer_preds), color = "black") +
  #  scale_color_manual(name = "Survey Unit") +
 # scale_fill_manual(values =  palette_37, guide = "none") +
 #   lims(x = c(min(distances_dissimilarities_allyears.r[,year]),max(distances_dissimilarities_allyears.r[,year])),
 #      y = c(0.15,0.9)) +
  theme_classic() +
  xlab("Year") +
ylab("Balanced BC dissimilarity")

points_wavylines_bray_balanced_year_reduced_gam_EBS_only

#just one region that's homogenizing (NZ WCSI)
points_wavylines_bray_balanced_year_reduced_gam_NZ_only <- ggplot() +
  geom_point(data = distances_dissimilarities_allyears.r[survey_unit == "NZ-WCSI",],
             aes(x = year,
                 y = bray_curtis_dissimilarity_balanced_mean), alpha = 0.5, size = 1, color = BC_balanced_model_coefs_reduced_length.unique[Survey_Name_Season == "W Coast S Island NZ",hex]) +
    geom_line(data = year_survey_unit_expand.dt[survey_unit == "NZ-WCSI",],
             aes(x = year,
                 y = bray_glm_mod_fit),
             color = BC_balanced_model_coefs_reduced_length.unique[Survey_Name_Season == "W Coast S Island NZ",hex]) +
  geom_ribbon(data = year_survey_unit_expand.dt[survey_unit == "NZ-WCSI",], aes(x = year, ymin=bray_glm_mod_fit-bray_glm_mod_fit_SE, ymax=bray_glm_mod_fit+bray_glm_mod_fit_SE), alpha=0.3, fill = BC_balanced_model_coefs_reduced_length.unique[Survey_Name_Season == "W Coast S Island NZ",hex]) + #add standard error
 # geom_line(data = lmer_bray_balanced_predictions, aes(x = year, y = bray_curtis_lmer_preds), color = "black") +
    scale_color_manual(name = "Survey Unit") +
 # scale_fill_manual(values =  palette_37, guide = "none") +
    lims(x = c(min(distances_dissimilarities_allyears.r[,year]),max(distances_dissimilarities_allyears.r[,year])),
       y = c(0.15,0.9)) +
  theme_classic() +
  xlab("Year") +
ylab("Balanced BC dissimilarity")

points_wavylines_bray_balanced_year_reduced_gam_NZ_only

#just two regions, another that's differentiating (SEUS Summer)

points_wavylines_bray_balanced_year_reduced_gam_NZ_SEUS_Summer_only <- ggplot() +
  geom_point(data = distances_dissimilarities_allyears.r[survey_unit %in% c("NZ-WCSI","SEUS-summer"),],
             aes(x = year,
                 y = bray_curtis_dissimilarity_balanced_mean,
                 color = Survey_Name_Season), alpha = 0.5, size = 1) +
    geom_line(data = year_survey_unit_expand.dt[survey_unit %in% c("NZ-WCSI","SEUS-summer"),],
             aes(x = year,
                 y = bray_glm_mod_fit,
                 color = Survey_Name_Season)) +
  geom_ribbon(data = year_survey_unit_expand.dt[survey_unit %in% c("NZ-WCSI","SEUS-summer"),], aes(x = year, ymin=bray_glm_mod_fit-bray_glm_mod_fit_SE, ymax=bray_glm_mod_fit+bray_glm_mod_fit_SE, fill = survey_unit), alpha=0.3) + #add standard error
 # geom_line(data = lmer_bray_balanced_predictions, aes(x = year, y = bray_curtis_lmer_preds), color = "black") +
    scale_color_manual(name = "Survey Unit", values = c(BC_balanced_model_coefs_reduced_length.unique[Survey_Name_Season %in% c("W Coast S Island NZ","SE US Summer"),hex])) +
  scale_fill_manual(guide = "none", values = c(BC_balanced_model_coefs_reduced_length.unique[Survey_Name_Season %in% c("W Coast S Island NZ","SE US Summer"),hex])) +
    lims(x = c(min(distances_dissimilarities_allyears.r[,year]),max(distances_dissimilarities_allyears.r[,year])),
       y = c(0.15,0.9)) +
  theme_classic() +
  xlab("Year") +
ylab("Balanced BC dissimilarity") +
  theme(legend.position = "null")

points_wavylines_bray_balanced_year_reduced_gam_NZ_SEUS_Summer_only

#just three regions, another one that's stable, (Canada Scotian Shelf)

points_wavylines_bray_balanced_year_reduced_gam_NZ_SEUS_ScoShelf_only <- ggplot() +
  geom_point(data = distances_dissimilarities_allyears.r[survey_unit %in% c("NZ-WCSI","SEUS-summer", "SCS-SUMMER"),],
             aes(x = year,
                 y = bray_curtis_dissimilarity_balanced_mean,
                 color = Survey_Name_Season), alpha = 0.5, size = 1) +
    geom_line(data = year_survey_unit_expand.dt[survey_unit %in% c("NZ-WCSI","SEUS-summer","SCS-SUMMER"),],
             aes(x = year,
                 y = bray_glm_mod_fit,
                 color = Survey_Name_Season)) +
  geom_ribbon(data = year_survey_unit_expand.dt[survey_unit %in% c("NZ-WCSI","SEUS-summer","SCS-SUMMER"),], aes(x = year, ymin=bray_glm_mod_fit-bray_glm_mod_fit_SE, ymax=bray_glm_mod_fit+bray_glm_mod_fit_SE, fill = Survey_Name_Season), alpha=0.3) + #add standard error
 # geom_line(data = lmer_bray_balanced_predictions, aes(x = year, y = bray_curtis_lmer_preds), color = "black") +
    scale_color_manual(name = "Survey Unit", values = c(BC_balanced_model_coefs_reduced_length.unique[Survey_Name_Season %in% c("W Coast S Island NZ","SE US Summer","Scotian Shelf"),hex])) +
  scale_fill_manual(guide = "none", values = c(BC_balanced_model_coefs_reduced_length.unique[Survey_Name_Season %in% c("W Coast S Island NZ","SE US Summer","Scotian Shelf"),hex])) +
    lims(x = c(min(distances_dissimilarities_allyears.r[,year]),max(distances_dissimilarities_allyears.r[,year])),
       y = c(0.15,0.9)) +
  theme_classic() +
  xlab("Year") +
ylab("Balanced BC dissimilarity") +
  theme(legend.position = "null")

points_wavylines_bray_balanced_year_reduced_gam_NZ_SEUS_ScoShelf_only

ggsave(points_wavylines_bray_balanced_year_reduced_gam_NZ_only, path = here::here("figures"), filename ="points_wavylines_bray_balanced_year_reduced_gam_NZ_only.jpg", height = 5, width = 5, unit = "in")
ggsave(points_wavylines_bray_balanced_year_reduced_gam_EBS_only, path = here::here("figures"), filename ="points_wavylines_bray_balanced_year_reduced_gam_EBS_only.jpg", height = 5, width = 9, unit = "in")
ggsave(points_wavylines_bray_balanced_year_reduced_gam_NZ_SEUS_Summer_only, path = here::here("figures"), filename ="points_wavylines_bray_balanced_year_reduced_gam_NZ_SEUS_Summer_only.jpg", height = 5, width = 5, unit = "in")
ggsave(points_wavylines_bray_balanced_year_reduced_gam_NZ_SEUS_ScoShelf_only, path = here::here("figures"), filename ="points_wavylines_bray_balanced_year_reduced_gam_NZ_SEUS_ScoShelf_only.jpg", height = 5, width = 5, unit = "in")




```

Merge BC versus Year plot with GAMS and Region vs. coefficient plot for LMERs

```{r}

BC_BALANCED_GAM_LMER_merge_legend <- ggdraw(xlim = c(0, 40), ylim = c(0, 20)) +
    draw_plot(points_wavylines_bray_balanced_year_reduced_gam,
                                         x = 1, y = 1, width = 20, height = 20) +
    draw_plot(BC_BALANCED_Dissimilarity_Coef_errorbar_reduced +
        theme(legend.key.size = unit(0.5, 'cm'), #change legend key size
        legend.title = element_text(size=16), #change legend title font size
        legend.text = element_text(size=14)), #change legend text font size),
                                         x = 20, y = 1, width = 20, height = 19) +
    draw_plot(get_legend(directional_change_legend_plot + 
      theme(legend.key.size = unit(0.5, 'cm'), #change legend key size
        legend.title = element_text(size=16), #change legend title font size
        legend.text = element_text(size=14))), #change legend text font size)
                                         x = 35, y = 4, width = 4, height = 2)

ggsave(BC_BALANCED_GAM_LMER_merge_legend, path = here::here("figures"), filename = "BC_BALANCED_GAM_LMER_merge_legend.png", height = 8, width = 14, units = "in")
```
```

