---
title: "Dissimilarity time series analysis excluding species present in less than 1/3 of years"
author: Zoë J. Kitchel
date: May 9, 2024
output: html_notebook
---

Script 4d for Kitchel et al. 2024 in prep taxonomic diversity manuscript.

```{r setup}
library(data.table)
library(vegan)
library(sf)
library(concaveman) #polygon around points
library(betapart) #allows us to partition beta diversity
library(geosphere)
library(ggpubr) #stat_regline_equation
library(nlme)
library(mgcv) #to make gam
library(cowplot)
library(lme4)
library(stringr)

#pull in function to calculate model estimates and standard errors
source(here::here("analysis_code","extract_coefficients_function.R"))
```


Pull dissimilarity table
```{r pull outputs from previous scripts}
#Pull Dissimilarity Means
jaccard_index_allyears_onethirdyears_excluded <- readRDS(here::here("output", "jaccard_index_VEGAN", "jaccard_index_allyears_onethirdyears_excluded.rds"))

#make survey and survey unit factors
jaccard_index_allyears_onethirdyears_excluded[,survey:=factor(survey)][,survey_unit:=factor(survey_unit)]

#adjust years
jaccard_index_allyears_onethirdyears_excluded[,year_adj := year-min(year)+1]

#add new variable for year in sequence per region
jaccard_index_allyears_onethirdyears_excluded[,first_year := min(year),.(survey_unit)]
jaccard_index_allyears_onethirdyears_excluded[,last_year := max(year),.(survey_unit)]

#jaccard_index_allyears_onethirdyears_excluded[,year_in_seq := year-first_year+1]

jaccard_index_allyears_onethirdyears_excluded[,years_sampled := last_year-first_year+1]


```

List of surveys
```{r}
#list of surveys
survey_unit.list <- levels(factor(jaccard_index_allyears_onethirdyears_excluded$survey_unit))
```

 
 Pull in palette and name helper
```{r}
source(here::here("analysis_code","color_links.R"))
```


##Make GAM (Generalized Additive Models)

Jaccard
```{r jaccard gams}
jaccard_total_gam <- gam(annual_dissimilarity_value ~ year + s(survey_unit, year, bs = "fs", m = 1),#random smooth
                            data = jaccard_index_allyears_onethirdyears_excluded[dissimilarity_metric == "jaccard_dissimilarity_index_binary"])

```

##Make LMERS

Jaccard
*These all converged*
```{r jaccard}
#running with lme instead of lmer gave same results, but allowed for calculation of p-value

#Jaccard total with lme
jaccard_total_lme <- lme(annual_dissimilarity_value ~ year_adj, random = (~1 + year_adj|survey_unit),data = jaccard_index_allyears_onethirdyears_excluded[dissimilarity_metric == "jaccard_dissimilarity_index_binary"])

#Jaccard total with lmer
jaccard_total_lmer <- lmer(annual_dissimilarity_value ~ year_adj + (1 + year_adj|survey_unit),data = jaccard_index_allyears_onethirdyears_excluded[dissimilarity_metric == "jaccard_dissimilarity_index_binary"]) #FYI doesn't converge

summary(jaccard_total_lme)
anova(jaccard_total_lme)

#EXPORT JACCARD COEFFICIENTS
jaccard_total_coefs <- data.table(coef(jaccard_total_lme))
jaccard_total_coefs[,survey_unit := rownames(coef(jaccard_total_lme))][,Year := round(year_adj,5)][,Intercept := round(`(Intercept)`,2)]
#View(jaccard_total_coefs)

jaccard_total_coefs <- jaccard_total_coefs[color_link, on = "survey_unit"]

jaccard_total_coefs.exp_onethirdyears_excluded <- jaccard_total_coefs[,.(Survey_Name_Season, Intercept, Year)]

#export this table
fwrite(jaccard_total_coefs.exp_onethirdyears_excluded, file = here::here("output","jaccard_total_coefs.exp_onethirdyears_excluded.csv"))
```
##Linear Model

Also, build a simple linear model with an interaction (instead of random slope and intercept for survey)
**THIS IS HOW WE CLASSIFY SURVEYS as of fall 2023 (update from before, where we classified surveys from LMER)**
(Especially good because lmer didn't converge)

Same, but for Jaccard
```{r linear model Jaccard}

jaccard_total_lm <- lm(annual_dissimilarity_value ~ year * survey_unit,data = jaccard_index_allyears_onethirdyears_excluded[dissimilarity_metric == "jaccard_dissimilarity_index_binary"])

#Use linear algebra to calculate coefficients and standard errors
jaccard_total_coefs.r_onethirdyears_excluded <- data.table(lm_interaction_coefficients_se(jaccard_total_lm))

#Add survey names
jaccard_total_coefs.r_onethirdyears_excluded[,survey_unit := sort(unique(jaccard_index_allyears_onethirdyears_excluded$survey_unit))]
  
  #homogenizing vs differentiating
  jaccard_total_coefs.r_onethirdyears_excluded[,differentiating := ifelse((se< abs(estimate) & estimate > 0),1,0)][,homogenizing := ifelse((se< abs(estimate) & estimate < 0),1,0)]
  
  #lower and upper CI
    jaccard_total_coefs.r_onethirdyears_excluded[,lwr := estimate-se][,upr:= estimate+se]
    
    #save as output
    fwrite(jaccard_total_coefs.r_onethirdyears_excluded, here::here("output","jaccard_total_coefs.r_onethirdyears_excluded.csv"))
  
  jaccard_total_coefs_sum <- data.table(differentiating = sum(  jaccard_total_coefs.r_onethirdyears_excluded$differentiating), homogenizing = sum(  jaccard_total_coefs.r_onethirdyears_excluded$homogenizing))
  
  #hex by trend
  jaccard_total_coefs.r_onethirdyears_excluded[,trend_color_jaccard := ifelse(homogenizing == 1,"#FFC109", ifelse(differentiating == 1,"#D81C60","#1E88E5"))]
  
  #link to color link
  jaccard_total_coefs.r_onethirdyears_excluded <- color_link[jaccard_total_coefs.r_onethirdyears_excluded, on = "survey_unit"]
  
  color_link_trend_jaccard <- jaccard_total_coefs.r_onethirdyears_excluded[,.(survey_unit,Survey_Name_Season, hex,  trend_color_jaccard)]

```

##Get LMER models as predictions

```{r new data}

#new data for lmer
lmer_year <- seq(min(jaccard_index_allyears_onethirdyears_excluded[,year]), max(jaccard_index_allyears_onethirdyears_excluded[,year]), by = 0.1)

lmer_year_adj <- seq(min(jaccard_index_allyears_onethirdyears_excluded[,year_adj]), max(jaccard_index_allyears_onethirdyears_excluded[,year_adj]), by = 0.1)

```


```{r total jaccard}

#predict average lmer
lmer_jaccard_total_predictions <- data.table(year = lmer_year, year_adj = lmer_year_adj)

#confidence intervals
jaccard_total_lmer_confint <- confint(jaccard_total_lmer)

#populate data table of lmer predictions
lmer_jaccard_total_predictions[,jaccard_lmer_preds := fixef(jaccard_total_lmer)[[1]] + fixef(jaccard_total_lmer)[[2]] * year_adj][,jaccard_lmer_preds_lowerCI := jaccard_total_lmer_confint[5] + jaccard_total_lmer_confint[6] * year_adj][,jaccard_lmer_preds_upperCI := jaccard_total_lmer_confint[11] + jaccard_total_lmer_confint[12] * year_adj]
```


##Lollipop SE plot by linear trend experienced

Jaccard total dissimilarity
```{r lollipop figure jaccard total dissimilarity}


jaccard_total_coefs.r_onethirdyears_excluded[,Directional_Change_sig := ifelse(differentiating > 0, "Differentiation",ifelse(homogenizing > 0 , "Homogenization", "No β-diversity trend"))]
years_sampled <- unique(jaccard_index_allyears_onethirdyears_excluded[,.(survey_unit, years_sampled)])
jaccard_total_coefs.r_onethirdyears_excluded <- years_sampled[jaccard_total_coefs.r_onethirdyears_excluded, on = "survey_unit"]

(jaccard_total_Dissimilarity_Coef_errorbar_reduced_colorbytrend_LINEAR_MODEL_onethirdyears_excluded <- ggplot(data = jaccard_total_coefs.r_onethirdyears_excluded, aes(x = reorder(Survey_Name_Season, estimate) , y = estimate), label = Survey_Name_Season) +
    geom_errorbar(aes(ymin = lwr, ymax = upr), fill = "grey", width = 0) + #add confidence intervals
  geom_point(aes(size = years_sampled, color = Directional_Change_sig), stat = 'identity') +
  scale_color_manual(values = c("#D81C60","#FFC109","#1E88E5"), name = "β-diversity trend\nTotal Jaccard dissimilarity", guide="none") +
  scale_size_binned(range = c(1,8), name = "Survey period length\n(years)") +
  geom_hline(yintercept = 0) +
 # scale_y_continuous(breaks = seq(-0.0075, 0.0075, by = 0.0025), labels = c("-0.0075","","-0.0025","0", "0.0025", "",  "")) +
  xlab("Survey unit") +
  ylab("β-diversity trend\nTotal Jaccard dissimilarity") + #total jaccard dissimilarity trend
  coord_flip() +
  theme_classic() +
  theme(axis.text.y = element_text(face = "bold"), axis.title.y = element_blank(), axis.text.x = element_text(size = 15), axis.title.x = element_text(size = 15), legend.position = c(0.3,0.8), legend.direction = "vertical", legend.text = element_text(size = 15), legend.title = element_text(size = 16)))

ggsave(jaccard_total_Dissimilarity_Coef_errorbar_reduced_colorbytrend_LINEAR_MODEL_onethirdyears_excluded, 
       path = here::here("figures"), filename = "jaccard_total_Dissimilarity_Coef_errorbar_reduced_colorbytrend_LINEAR_MODEL_onethirdyears_excluded.jpg", height = 7, width = 7)

```


#Figure to merge full dataset with two sensitivity analyses for supplement
```{r}
#load true jaccard total coefs
jaccard_total_coefs.r <- fread(here::here("output","jaccard_total_coefs.r.csv"))
jaccard_total_coefs.r[,Directional_Change_sig_fulljaccard := ifelse(differentiating > 0, "Differentiation",ifelse(homogenizing > 0 , "Homogenization", "No β-diversity trend"))]
jaccard_total_coefs.r[,trend_color_jaccard_full := ifelse(homogenizing == 1,"#FFC109", ifelse(differentiating == 1,"#D81C60","#1E88E5"))]


#load 15perc exclude jaccard total coefs
jaccard_total_coefs.r_15perc_excluded <- fread(here::here("output","jaccard_total_coefs.r_15perc_excluded.csv"))
jaccard_total_coefs.r_15perc_excluded[,Directional_Change_sig_15perc_excluded_jaccard := ifelse(differentiating > 0, "Differentiation",ifelse(homogenizing > 0 , "Homogenization", "No β-diversity trend"))]
jaccard_total_coefs.r_15perc_excluded[,trend_color_jaccard_15percexluded := ifelse(homogenizing == 1,"#FFC109", ifelse(differentiating == 1,"#D81C60","#1E88E5"))]

#full dataset coefs
jaccard_total_coefs.r[,jaccard_coefficient := estimate]
all_total_coefs <- jaccard_total_coefs.r[,.(survey_unit, jaccard_coefficient, Directional_Change_sig_fulljaccard, trend_color_jaccard_full)]

#15 percent exclude coefs
jaccard_total_coefs.r_15perc_excluded[,jaccard_coefficient_15percexclude := estimate]
jaccard_total_coefs.r_15perc_excluded.r <- jaccard_total_coefs.r_15perc_excluded[,.(survey_unit, jaccard_coefficient_15percexclude, Directional_Change_sig_15perc_excluded_jaccard, trend_color_jaccard_15percexluded)]
all_total_coefs <- all_total_coefs[jaccard_total_coefs.r_15perc_excluded.r, on = "survey_unit"]

#one third years exclude (this script)
jaccard_total_coefs.r_onethirdyears_excluded[,onethirdyears_relative_coefficient := estimate][,Directional_Change_onethirdyears := Directional_Change_sig][,trend_color_jaccard_onethirdyears_excluded := trend_color_jaccard]
jaccard_total_coefs.r_onethirdyears_excluded.r <- jaccard_total_coefs.r_onethirdyears_excluded[,.(survey_unit, Survey_Name_Season, onethirdyears_relative_coefficient, Directional_Change_onethirdyears, trend_color_jaccard_onethirdyears_excluded)]
all_total_coefs <- all_total_coefs[jaccard_total_coefs.r_onethirdyears_excluded.r, on = "survey_unit"]

#plot
(Dissimilarity_Coef_errorbar_reduced_colorbytrend_LINEAR_MODEL_spp_inclusion_sensitivity <- ggplot(data = all_total_coefs, aes(x = reorder(Survey_Name_Season, jaccard_coefficient)), label = Survey_Name_Season) +
  geom_point(aes(y = jaccard_coefficient, fill = Directional_Change_sig_fulljaccard), stat = 'identity',shape = 21, color = "black", size = 5) +
  geom_point(aes(y = jaccard_coefficient_15percexclude, color = Directional_Change_sig_15perc_excluded_jaccard), stat = 'identity', shape = 15, alpha = 0.5,size = 5) +
  geom_point(aes(y = onethirdyears_relative_coefficient, color = Directional_Change_onethirdyears), stat = 'identity', shape = 17, alpha = 0.5,size = 5) +
  scale_color_manual(values = c("#D81C60","#FFC109","#1E88E5"), name = "β-diversity trend", guide="none") +
  scale_fill_manual(values = c("#D81C60","#FFC109","#1E88E5"), name = "β-diversity trend", guide="none") +
  geom_hline(yintercept = 0) +
 # scale_y_continuous(breaks = seq(-0.0075, 0.0075, by = 0.0025), labels = c("-0.0075","","-0.0025","0", "0.0025", "",  "")) +
  xlab("Survey unit") +
  ylab("β-diversity trend") + #total jaccard dissimilarity trend
  coord_flip() +
  theme_classic() +
  theme(axis.text.y = element_text(face = "bold"), axis.title.y = element_blank(), axis.text.x = element_text(size = 15), axis.title.x = element_text(size = 15), legend.position = c(0.3,0.8), legend.direction = "vertical", legend.text = element_text(size = 15), legend.title = element_text(size = 16)))

ggsave(Dissimilarity_Coef_errorbar_reduced_colorbytrend_LINEAR_MODEL_spp_inclusion_sensitivity, 
       path = here::here("figures"), filename = "Dissimilarity_Coef_errorbar_reduced_colorbytrend_LINEAR_MODEL_spp_inclusion_sensitivity.jpg", height = 7, width = 7)

```

