---
title: "Dissimilarity time series analysis excluding 15% of species with lowest biomass"
author: Zoë J. Kitchel
date: May 9, 2024
output: html_notebook
---

Script 4c for Kitchel et al. 2024 in prep taxonomic diversity manuscript.

```{r setup}
library(data.table)
library(vegan)
library(sf)
library(concaveman) #polygon around points
library(betapart) #allows us to partition beta diversity
library(geosphere)
library(ggpubr) #stat_regline_equation
library(nlme)
library(mgcv) #to make gam
library(cowplot)
library(lme4)
library(stringr)

#pull in function to calculate model estimates and standard errors
source(here::here("analysis_code","extract_coefficients_function.R"))
```


Pull dissimilarity table
```{r pull outputs from previous scripts}
#Pull Dissimilarity Means
jaccard_index_allyears_15perc_excluded <- readRDS(here::here("output", "jaccard_index_VEGAN", "jaccard_index_allyears_15perc_excluded.rds"))

#make survey and survey unit factors
jaccard_index_allyears_15perc_excluded[,survey:=factor(survey)][,survey_unit:=factor(survey_unit)]

#adjust years
jaccard_index_allyears_15perc_excluded[,year_adj := year-min(year)+1]

#add new variable for year in sequence per region
jaccard_index_allyears_15perc_excluded[,first_year := min(year),.(survey_unit)]
jaccard_index_allyears_15perc_excluded[,last_year := max(year),.(survey_unit)]

#jaccard_index_allyears_15perc_excluded[,year_in_seq := year-first_year+1]

jaccard_index_allyears_15perc_excluded[,years_sampled := last_year-first_year+1]


```

List of surveys
```{r}
#list of surveys
survey_unit.list <- levels(factor(jaccard_index_allyears_15perc_excluded$survey_unit))
```

 
 Pull in palette and name helper
```{r}
source(here::here("analysis_code","color_links.R"))
```


##Make GAM (Generalized Additive Models)

Jaccard
```{r jaccard gams}
jaccard_total_gam <- gam(annual_dissimilarity_value ~ year + s(survey_unit, year, bs = "fs", m = 1),#random smooth
                            data = jaccard_index_allyears_15perc_excluded[dissimilarity_metric == "jaccard_dissimilarity_index_binary"])

```

##Make LMERS

Jaccard
*These all converged*
```{r jaccard}
#running with lme instead of lmer gave same results, but allowed for calculation of p-value

#Jaccard total with lme
jaccard_total_lme <- lme(annual_dissimilarity_value ~ year_adj, random = (~1 + year_adj|survey_unit),data = jaccard_index_allyears_15perc_excluded[dissimilarity_metric == "jaccard_dissimilarity_index_binary"])

#Jaccard total with lmer
jaccard_total_lmer <- lmer(annual_dissimilarity_value ~ year_adj + (1 + year_adj|survey_unit),data = jaccard_index_allyears_15perc_excluded[dissimilarity_metric == "jaccard_dissimilarity_index_binary"]) #FYI doesn't converge

summary(jaccard_total_lme)
anova(jaccard_total_lme)

#EXPORT JACCARD COEFFICIENTS
jaccard_total_coefs <- data.table(coef(jaccard_total_lme))
jaccard_total_coefs[,survey_unit := rownames(coef(jaccard_total_lme))][,Year := round(year_adj,5)][,Intercept := round(`(Intercept)`,2)]
#View(jaccard_total_coefs)

jaccard_total_coefs <- jaccard_total_coefs[color_link, on = "survey_unit"]

jaccard_total_coefs.exp_15perc_excluded <- jaccard_total_coefs[,.(Survey_Name_Season, Intercept, Year)]

#export this table
fwrite(jaccard_total_coefs.exp_15perc_excluded, file = here::here("output","jaccard_total_coefs.exp_15perc_excluded.csv"))
```
##Linear Model

Also, build a simple linear model with an interaction (instead of random slope and intercept for survey)
**THIS IS HOW WE CLASSIFY SURVEYS as of fall 2023 (update from before, where we classified surveys from LMER)**
(Especially good because lmer didn't converge)

Same, but for Jaccard
```{r linear model Jaccard}

jaccard_total_lm <- lm(annual_dissimilarity_value ~ year * survey_unit,data = jaccard_index_allyears_15perc_excluded[dissimilarity_metric == "jaccard_dissimilarity_index_binary"])

#Use linear algebra to calculate coefficients and standard errors
jaccard_total_coefs.r_15perc_excluded <- data.table(lm_interaction_coefficients_se(jaccard_total_lm))

#Add survey names
jaccard_total_coefs.r_15perc_excluded[,survey_unit := sort(unique(jaccard_index_allyears_15perc_excluded$survey_unit))]
  
  #homogenizing vs differentiating
  jaccard_total_coefs.r_15perc_excluded[,differentiating := ifelse((se< abs(estimate) & estimate > 0),1,0)][,homogenizing := ifelse((se< abs(estimate) & estimate < 0),1,0)]
  
  #lower and upper CI
    jaccard_total_coefs.r_15perc_excluded[,lwr := estimate-se][,upr:= estimate+se]
    
    #save as output
    fwrite(jaccard_total_coefs.r_15perc_excluded, here::here("output","jaccard_total_coefs.r_15perc_excluded.csv"))
  
  jaccard_total_coefs_sum <- data.table(differentiating = sum(  jaccard_total_coefs.r_15perc_excluded$differentiating), homogenizing = sum(  jaccard_total_coefs.r_15perc_excluded$homogenizing))
  
  #hex by trend
  jaccard_total_coefs.r_15perc_excluded[,trend_color_jaccard := ifelse(homogenizing == 1,"#FFC109", ifelse(differentiating == 1,"#D81C60","#1E88E5"))]
  
  #link to color link
  jaccard_total_coefs.r_15perc_excluded <- color_link[jaccard_total_coefs.r_15perc_excluded, on = "survey_unit"]
  
  color_link_trend_jaccard <- jaccard_total_coefs.r_15perc_excluded[,.(survey_unit,Survey_Name_Season, hex,  trend_color_jaccard)]

```

##Get LMER models as predictions

```{r new data}

#new data for lmer
lmer_year <- seq(min(jaccard_index_allyears_15perc_excluded[,year]), max(jaccard_index_allyears_15perc_excluded[,year]), by = 0.1)

lmer_year_adj <- seq(min(jaccard_index_allyears_15perc_excluded[,year_adj]), max(jaccard_index_allyears_15perc_excluded[,year_adj]), by = 0.1)

```


```{r total jaccard}

#predict average lmer
lmer_jaccard_total_predictions <- data.table(year = lmer_year, year_adj = lmer_year_adj)

#confidence intervals
jaccard_total_lmer_confint <- confint(jaccard_total_lmer)

#populate data table of lmer predictions
lmer_jaccard_total_predictions[,jaccard_lmer_preds := fixef(jaccard_total_lmer)[[1]] + fixef(jaccard_total_lmer)[[2]] * year_adj][,jaccard_lmer_preds_lowerCI := jaccard_total_lmer_confint[5] + jaccard_total_lmer_confint[6] * year_adj][,jaccard_lmer_preds_upperCI := jaccard_total_lmer_confint[11] + jaccard_total_lmer_confint[12] * year_adj]
```


##Lollipop SE plot by linear trend experienced

Jaccard total dissimilarity
```{r lollipop figure jaccard total dissimilarity}


jaccard_total_coefs.r_15perc_excluded[,Directional_Change_sig := ifelse(differentiating > 0, "Differentiation",ifelse(homogenizing > 0 , "Homogenization", "No β-diversity trend"))]
years_sampled <- unique(jaccard_index_allyears_15perc_excluded[,.(survey_unit, years_sampled)])
jaccard_total_coefs.r_15perc_excluded <- years_sampled[jaccard_total_coefs.r_15perc_excluded, on = "survey_unit"]

(jaccard_total_Dissimilarity_Coef_errorbar_reduced_colorbytrend_LINEAR_MODEL_15perc_excluded <- ggplot(data = jaccard_total_coefs.r_15perc_excluded, aes(x = reorder(Survey_Name_Season, estimate) , y = estimate), label = Survey_Name_Season) +
    geom_errorbar(aes(ymin = lwr, ymax = upr), fill = "grey", width = 0) + #add confidence intervals
  geom_point(aes(size = years_sampled, color = Directional_Change_sig), stat = 'identity') +
  scale_color_manual(values = c("#D81C60","#FFC109","#1E88E5"), name = "β-diversity trend\nTotal Jaccard dissimilarity", guide="none") +
  scale_size_binned(range = c(1,8), name = "Survey period length\n(years)") +
  geom_hline(yintercept = 0) +
  scale_y_continuous(breaks = seq(-0.0075, 0.0075, by = 0.0025), labels = c("-0.0075","","-0.0025","0", "0.0025", "",  "")) +
  xlab("Survey unit") +
  ylab("β-diversity trend\nTotal Jaccard dissimilarity") + #total jaccard dissimilarity trend
  coord_flip() +
  theme_classic() +
  theme(axis.text.y = element_text(face = "bold"), axis.title.y = element_blank(), axis.text.x = element_text(size = 15), axis.title.x = element_text(size = 15), legend.position = c(0.3,0.8), legend.direction = "vertical", legend.text = element_text(size = 15), legend.title = element_text(size = 16)))

ggsave(jaccard_total_Dissimilarity_Coef_errorbar_reduced_colorbytrend_LINEAR_MODEL_15perc_excluded, 
       path = here::here("figures"), filename = "jaccard_total_Dissimilarity_Coef_errorbar_reduced_colorbytrend_LINEAR_MODEL_15perc_excluded.jpg", height = 7, width = 7)

```

#Figure to merge three metrics for supplement

##Wavy Line Plot for GAMs (Figure 2a)

Generate predicted values (this datatable will hold both jaccard and bray curtis total dissimilarities)
```{r generate predicted values GAM}
#generate list again, as for some reason it doesn't stay in environment
survey_unit.list <- levels(factor(jaccard_index_allyears_15perc_excluded$survey_unit))

#generate new data to smooth lines (need year and season survey combinations)
year_survey_unit_expand.dt <- data.table(survey_unit = as.character(NULL), year = as.numeric(NULL), year_adj = as.numeric(NULL ))

for (i in 1:length(survey_unit.list)) {
  #generate year vectors
  survey_unit_years <- unique(jaccard_index_allyears_15perc_excluded[survey_unit == survey_unit.list[i],.(survey_unit, year, year_adj)])
  
  years <- seq(min(survey_unit_years[,year]), max(survey_unit_years[,year]), by = 0.1)
  
  year_adjust <- seq(min(survey_unit_years[,year_adj]), max(survey_unit_years[,year_adj]), by = 0.1)
  
  year_survey_unit_expand.dt_addition <- data.table(survey_unit = survey_unit.list[i], year = years, year_adj = year_adjust)
  
  year_survey_unit_expand.dt <- rbind(year_survey_unit_expand.dt, year_survey_unit_expand.dt_addition)
}


#add colors and names to full year and survey unit combination table
year_survey_unit_expand_jaccard.dt <- year_survey_unit_expand.dt[color_link_trend_jaccard, on = "survey_unit"]
```

Jaccard
Get model as predictions
```{r}
#for plotting, get model as predictions
jaccard_total_gam_predictions <- predict(jaccard_total_gam, se.fit = TRUE, newdata = year_survey_unit_expand_jaccard.dt)

#merge into table
year_survey_unit_expand_jaccard.dt[,jaccard_glm_mod_fit := jaccard_total_gam_predictions$fit][,jaccard_glm_mod_fit_SE := jaccard_total_gam_predictions$se.fit]

```


##Color GAMs by trend (year vs dissimilarity)
###Jaccard dissimilarity
```{r color wavy lines by trend Jaccard dissimilarity}

#link distances dissimilarities with trend color hex
jaccard_index_allyears_15perc_excluded_jaccard <- jaccard_index_allyears_15perc_excluded[color_link_trend_jaccard, on = "survey_unit"]
jaccard_index_allyears_15perc_excluded_jaccard <- jaccard_index_allyears_15perc_excluded_jaccard[dissimilarity_metric == "jaccard_dissimilarity_index_binary_abundance"]
#and again for jaccard
jaccard_index_allyears_15perc_excluded_jaccard <- jaccard_index_allyears_15perc_excluded[color_link_trend_jaccard, on = "survey_unit"]
jaccard_index_allyears_15perc_excluded_jaccard<- jaccard_index_allyears_15perc_excluded_jaccard[dissimilarity_metric == "jaccard_dissimilarity_index_binary"]

#change survey_unit to factor
jaccard_index_allyears_15perc_excluded_jaccard[,survey_unit := factor(survey_unit)]
year_survey_unit_expand_jaccard.dt[,survey_unit := factor(survey_unit)]

#check that survey_unit in alpha order
sort(unique(jaccard_index_allyears_15perc_excluded_jaccard$survey_unit)) == levels(jaccard_index_allyears_15perc_excluded_jaccard$survey_unit)
sort(unique(year_survey_unit_expand_jaccard.dt$survey_unit)) == levels(year_survey_unit_expand_jaccard.dt$survey_unit)

#get list of colors in the right order
trend_colors_jaccard.dt <- unique(jaccard_index_allyears_15perc_excluded_jaccard[dissimilarity_metric == "jaccard_dissimilarity_index_binary",.(survey_unit, Survey_Name_Season, trend_color_jaccard)])

#make sure it matches year_survey_unit_expand.dt
all(trend_colors_jaccard.dt == unique(year_survey_unit_expand_jaccard.dt[,.(survey_unit, Survey_Name_Season, trend_color_jaccard)]))

setorder(trend_colors_jaccard.dt, survey_unit)


points_wavylines_jaccard_total_year_reduced_gam_colorbytrend_15perc_excluded <- ggplot() +
  #full LMER predictions (black)
  geom_ribbon(data = lmer_jaccard_total_predictions, aes(x = year, ymin = jaccard_lmer_preds_lowerCI, ymax = jaccard_lmer_preds_upperCI), fill = "grey", alpha = 0.3) +
    geom_line(data = lmer_jaccard_total_predictions, aes(x = year, y = jaccard_lmer_preds), color = "black") +
  #points by trend
  geom_point(data = jaccard_index_allyears_15perc_excluded[dissimilarity_metric =="jaccard_dissimilarity_index_binary"], cols = "year_adj",
             aes(x = year,
                 y = annual_dissimilarity_value,
                 fill = survey_unit), alpha = 0.4, size = 1, shape = 21, color = "white") +
  #squiggly lines by survey unit based GAM and colored by trend
    geom_line(data = year_survey_unit_expand_jaccard.dt, cols = "year_adj",
             aes(x = year,
                 y = jaccard_glm_mod_fit,
                 color = survey_unit), alpha = 0.6) +
  #confidence intervals for individual GAM
  geom_ribbon(data = year_survey_unit_expand_jaccard.dt, cols = "year_adj", 
              aes(x = year, ymin=jaccard_glm_mod_fit-jaccard_glm_mod_fit_SE, ymax=jaccard_glm_mod_fit+jaccard_glm_mod_fit_SE, fill =  survey_unit), alpha=0.1) + #add standard error
  #set color by survey
    scale_color_manual(values = trend_colors_jaccard.dt$trend_color, name = "Survey Unit") +
  scale_fill_manual(values =trend_colors_jaccard.dt$trend_color, guide = "none") +
  theme_classic() +
  lims(x = c(min(jaccard_index_allyears_15perc_excluded[dissimilarity_metric == "jaccard_dissimilarity_index_binary",year]),max(jaccard_index_allyears_15perc_excluded[dissimilarity_metric == "jaccard_dissimilarity_index_binary",year]))
       ) +
  xlab("Year\n") +
ylab("β-diversity\n(Jaccard dissimilarity)") +
  theme(legend.position = "null", axis.text = element_text(size = 15), axis.title = element_text(size = 15))

points_wavylines_jaccard_total_year_reduced_gam_colorbytrend_15perc_excluded

ggsave(points_wavylines_jaccard_total_year_reduced_gam_colorbytrend_15perc_excluded, path = here::here("figures"), filename ="points_wavylines_jaccard_total_year_reduced_gam_colorbytrend_15perc_excluded.jpg", height = 6, width = 11, unit = "in")

#plot for legend
directional_change_legend_plot_colorbytrend <- jaccard_total_Dissimilarity_Coef_errorbar_reduced_colorbytrend_LINEAR_MODEL_15perc_excluded + 
  theme(legend.position = "right", legend.background = element_rect(fill= "transparent"), 
         legend.text = element_text(size = 15), legend.title = element_text(size = 16)) +
  guides(colour = guide_legend(override.aes = list(size=6)), size = "none")

```

###Merge BC versus Year plot with GAMS and Region vs. coefficient plot for LMs

```{r merge lollipop and wiggly lines}

######################################################
##################JACCARDTOTAL
jaccard_total_GAM_LM_merge_legend_colorbytrend_15perc_excluded <- ggdraw(xlim = c(0, 40.5), ylim = c(0, 21)) +
    draw_plot(points_wavylines_jaccard_total_year_reduced_gam_colorbytrend_15perc_excluded + theme(axis.text = element_text(size = 14)),
                                         x = 1, y = 1, width = 20, height = 20) +
    draw_plot(jaccard_total_Dissimilarity_Coef_errorbar_reduced_colorbytrend_LINEAR_MODEL_15perc_excluded +
        theme(legend.key.size = unit(0.5, 'cm'), #change legend key size
              axis.text = element_text(size = 8)), #change legend text font size),
                                         x = 20, y = 1, width = 19, height = 20) +
    draw_plot(get_legend(directional_change_legend_plot_colorbytrend + 
      theme(legend.key.size = unit(0.5, 'cm'), #change legend key size
        legend.title = element_text(size=16), #change legend title font size
        legend.text = element_text(size=15))), #change legend text font size)
                                x = 26, y = 8, width = 3, height = 2) +
  geom_text(aes(x = 2, y = 20.7), label = ("a."), size =8, fontface = "bold") +
  geom_text(aes(x =20, y = 20.7), label = ("b."), size =8, fontface = "bold")

ggsave(jaccard_total_GAM_LM_merge_legend_colorbytrend_15perc_excluded, path = here::here("figures"), filename = "jaccard_total_GAM_LM_merge_legend_colorbytrend_15perc_excluded.png", height = 8, width = 14, units = "in")
```
