---
title: "Dissimilarity time series analysis for reshuffled years"
author: Zoë J. Kitchel
date: October 17, 2023
output: html_notebook
---

Supplementary script 4b for Kitchel et al. 2023 in prep taxonomic diversity manuscript.

```{r setup}
library(data.table)
library(vegan)
library(sf)
library(concaveman) #polygon around points
library(betapart) #allows us to partition beta diversity
library(geosphere)
library(ggpubr) #stat_regline_equation
library(nlme)
library(mgcv) #to make gam
library(cowplot)
library(lme4)

```

Pull dissimilarity table
```{r pull outputs from previous scripts}
#Pull Dissimilarity Means
distances_dissimilarities_allyears <- readRDS(here::here("output", "dissimilarities", "distances_dissimilarities_allyears.rds"))

#make survey and survey unit factors
distances_dissimilarities_allyears[,survey:=factor(survey)][,survey_unit:=factor(survey_unit)]

#adjust years
distances_dissimilarities_allyears[,year_adj := year-min(year)+1]

#add new variable for year in sequence per region
distances_dissimilarities_allyears[,first_year := min(year),.(survey_unit)]
distances_dissimilarities_allyears[,last_year := max(year),.(survey_unit)]

#distances_dissimilarities_allyears[,year_in_seq := year-first_year+1]

distances_dissimilarities_allyears[,years_sampled := last_year-first_year+1]

#list of surveys
survey_unit.list <- levels(factor(distances_dissimilarities_allyears$survey_unit))


```
 
 Pull in palette and name helper
```{r}
source(here::here("analysis_code","color_links.R"))
```

Load up 'true slopes' and CIs for comparison
```{r}

bray_curtis_total_coefs.r <- fread(here::here("output","bray_curtis_total_coefs.r.csv"))
View(bray_curtis_total_coefs.r)
bray_curtis_total_coefs.r_true <- bray_curtis_total_coefs.r
```


#NULL MODEL CONSTUCTION
##1) Reshuffle years (maintain correlation among species, removes temporal autocorrelation within a species)

I need to:
- reshuffle years
- reassign years
- run lmer
- pull mean and CI for each region
```{r}
#surveys
all_survey_units <- color_link$survey_unit
years <- sort(unique(distances_dissimilarities_allyears$year))


for (i in 1:1000){

  BC_dissimilarity_time <- distances_dissimilarities_allyears[,.(survey_unit, year, year_adj, bray_curtis_dissimilarity_total_mean)]
  
  year_run_key <- data.table(year = years, years_null = sample(years, length(years), replace = F))
  
  BC_dissimilarity_time  <- BC_dissimilarity_time[year_run_key, on = "year"]
  
  #run overall model
  bray_curtis_total_lm <- lm(bray_curtis_dissimilarity_total_mean ~ years_null * survey_unit,data = BC_dissimilarity_time)
    
  bray_curtis_total_coefs <- data.table(summary(bray_curtis_total_lm)$coefficients)
  bray_curtis_total_coefs[,var := rownames(summary(bray_curtis_total_lm)$coefficients)]
  
  #limit to interactions only (check this if there are any model changes!) row 2 and rows 44:84
  bray_curtis_total_coefs.r <- bray_curtis_total_coefs[c(2,44:84),]
  
  #adjust survey unit name by deleting beginning of string
  bray_curtis_total_coefs.r[,survey_unit := substr(var, 23, str_length(var))][var == "years_null",survey_unit := "AI"]
  
  #calculate interaction coefficients
  AI_estimate <- bray_curtis_total_coefs.r[1,Estimate]
  bray_curtis_total_coefs.r[1,survey_unit_coefficient := AI_estimate]
  bray_curtis_total_coefs.r[2:42,survey_unit_coefficient := (AI_estimate + Estimate)]
  
  #homogenizing vs differentiating
  bray_curtis_total_coefs.r[,differentiating := ifelse((`Std. Error`< abs(survey_unit_coefficient) & survey_unit_coefficient > 0),1,0)][,homogenizing := ifelse((`Std. Error`< abs(survey_unit_coefficient) & survey_unit_coefficient < 0),1,0)]
  
    #full table with distribution of coefficients
    if(i == 1){
     bray_curtis_total_coefs.allruns <- bray_curtis_total_coefs.r }else{
    bray_curtis_total_coefs.allruns <- rbind(bray_curtis_total_coefs.allruns, bray_curtis_total_coefs.r)
   }
  
  
  #sum # differentiating vs homogenizing for each run
  bray_curtis_total_coefs_sum <- data.table(run = i, differentiating = sum(  bray_curtis_total_coefs.r$differentiating), homogenizing = sum(  bray_curtis_total_coefs.r$homogenizing))
  
  if(i == 1){
     all_null_runs_reshuffle_year <- bray_curtis_total_coefs_sum }else{
    all_null_runs_reshuffle_year <- rbind(all_null_runs_reshuffle_year, bray_curtis_total_coefs_sum)
   }
  print(i)
}

#95% quantile o(both count homogenization and count differentiation)

homogenization_95_quantile <- quantile(all_null_runs_reshuffle_year[,homogenizing], 0.95)
differentiation_95_quantile <- quantile(all_null_runs_reshuffle_year[,differentiating], 0.95)

#plot differentiation distribution
(reshuffle_differentiation_plot <- ggplot() +
  geom_histogram(data = all_null_runs_reshuffle_year, aes(differentiating), binwidth = 1) +
      geom_vline(xintercept = sum(bray_curtis_total_coefs.r_true$differentiating), color = "red") +
          geom_vline(xintercept = differentiation_95_quantile, linetype = "dashed") +
    scale_x_continuous(breaks = seq(0,16,by = 2)) +
  labs(x = "Number of surveys differentiating", y = "Count") +
  theme_classic())

#plot homogenization distribution
(reshuffle_homogenizing_plot <- ggplot() +
  geom_histogram(data = all_null_runs_reshuffle_year, aes(homogenizing), binwidth = 1) +
      geom_vline(xintercept = sum(bray_curtis_total_coefs.r_true$homogenizing), color = "red") +
          geom_vline(xintercept = homogenization_95_quantile, linetype = "dashed") +
    scale_x_continuous(breaks = seq(0,10,by = 2)) +
  labs(x = "Number of surveys homogenizing", y = "Count") +
  theme_classic())

reshuffle_null_plot <- cowplot::plot_grid(reshuffle_homogenizing_plot,reshuffle_differentiation_plot, 
                                     ncol = 2, labels = c("a.","b."))

ggsave(reshuffle_null_plot, path = here::here("figures"), filename = "reshuffle_null_plot.jpg", width = 6.5, height = 3)


#Plot the distribution of coefficients across all surveys versus the distribution of true coefficients
(reshuffle_slope_comparison_plot <- ggplot() +
  geom_histogram(data = bray_curtis_total_coefs.allruns, aes(survey_unit_coefficient), fill = "#595959", binwidth = 0.001) +
  labs(x = "Dissimilarity coefficient (null)", y = "Count (in 10,000s)") +
    scale_y_continuous(breaks = c(0,10000,20000,30000), labels = c("0","1","2","3")) +
    geom_vline(xintercept = median(bray_curtis_total_coefs.allruns$survey_unit_coefficient)) +
   geom_vline(xintercept = quantile(bray_curtis_total_coefs.allruns$survey_unit_coefficient, c(0.025,0.975)), linetype = "dashed") +
    xlim(min(c(bray_curtis_total_coefs.allruns$survey_unit_coefficient,bray_curtis_total_coefs.r_true$survey_unit_coefficient))-0.0035,
         max(c(bray_curtis_total_coefs.allruns$survey_unit_coefficient,bray_curtis_total_coefs.r_true$survey_unit_coefficient))+0.003) +
  theme_classic())

(true_slope_comparison_plot <- ggplot() +
  geom_histogram(data = bray_curtis_total_coefs.r_true, aes(survey_unit_coefficient), fill = "#595959", binwidth = 0.001) +
  labs(x = "Dissimilarity coefficient (observed)", y = "Count") +
    geom_vline(xintercept = median(bray_curtis_total_coefs.allruns$survey_unit_coefficient)) +
    geom_vline(xintercept = quantile(bray_curtis_total_coefs.allruns$survey_unit_coefficient, c(0.025,0.975)), linetype = "dashed") +
    xlim(min(c(bray_curtis_total_coefs.allruns$survey_unit_coefficient,bray_curtis_total_coefs.r_true$survey_unit_coefficient))-0.003,
         max(c(bray_curtis_total_coefs.allruns$survey_unit_coefficient,bray_curtis_total_coefs.r_true$survey_unit_coefficient))+0.003) +
  theme_classic())

#merge two plots
reshuffle_null_plot_bottom <- cowplot::plot_grid(true_slope_comparison_plot,reshuffle_slope_comparison_plot, 
                                     ncol = 1, labels = c("c.","d."), align = "h")

#merge all
reshuffle_null_plot_full <- cowplot::plot_grid(reshuffle_null_plot, reshuffle_null_plot_bottom, ncol = 1, rel_heights = c(1,2))

ggsave(reshuffle_null_plot_full, path = here::here("figures"), filename = "reshuffle_null_plot_full.jpg", width = 6.5, height = 6.5)


```














#NO LONGER DOING BELOW
##2) Vary start year (maintain correlation among species and autocorrelation within a species, but introduces weird breakpoints)

I need to:
- pick random starting years
- reassign years
- run lmer
- pull mean and CI for each region
```{r}
#surveys
all_survey_units <- levels(name_helper[,.(survey_unit)])
years_adj <- sort(unique(distances_dissimilarities_allyears$year_adj))

#only 45 possibilities, so does it make sense to sample 1000 times? probably not?

#for (i in 1:1000)){ #if we want 1000 runs
  for (i in 1:length(years_adj)){ #if we want 45 runs

  BC_dissimilarity_time <- distances_dissimilarities_allyears[,.(survey_unit, year, year_adj, bray_curtis_dissimilarity_total_mean)]
  
 # start_year <- sample(1:length(years_adj), 1) #if instead we want 1000 random runs
  start_year <- years_adj[i]
  
  random_sequence <- c(years_adj[start_year:length(years_adj)],years_adj[1:start_year-1])
  
  year_run_key <- data.table(year_adj = years_adj, years_adj_null = random_sequence)
  
  BC_dissimilarity_time  <- BC_dissimilarity_time[year_run_key, on = "year_adj"]
  
  #run overall model
  bray_curtis_total_lmer <- lmer(bray_curtis_dissimilarity_total_mean ~ years_adj_null + (1 + years_adj_null|survey_unit),data = BC_dissimilarity_time)
  
    
  bray_curtis_total_coefs <- data.table(coef(bray_curtis_total_lmer)[[1]])
  bray_curtis_total_coefs[,survey_unit := rownames(coef(bray_curtis_total_lmer)[[1]])][,Year_slope := signif(years_adj_null,3)][,Intercept := signif(`(Intercept)`,3)]
  bray_curtis_total_coefs <- bray_curtis_total_coefs[color_link, on = "survey_unit"]
  
  bray_curtis_total_coefs[,run := i]
  
  bray_curtis_total_coefs.exp <- bray_curtis_total_coefs[,.(run, Survey_Name_Season, survey_unit, Intercept, Year_slope)]
  
  if(i == 1){
     all_null_runs_cyclic_shift <- bray_curtis_total_coefs.exp }else{
    all_null_runs_cyclic_shift <- rbind(all_null_runs_cyclic_shift, bray_curtis_total_coefs.exp)
   }
  print(i)
}

#link "true" slope

true_slopes <- BC_total_model_coefs_reduced_length.unique[,.(lwr,upr, survey_unit, year_adj, Directional_Change, significant, Directional_Change_sig, trend_color)]

all_null_runs_cyclic_shift <- all_null_runs_cyclic_shift[true_slopes, on = "survey_unit"]

#take averages and quantiles across all runs
all_null_runs_cyclic_shift[,mean_slope := mean(Year_slope),Survey_Name_Season][,upper_95 := quantile(Year_slope,0.975),Survey_Name_Season][,lower_95 := quantile(Year_slope,0.025),Survey_Name_Season][,true_slope := year_adj]

saveRDS(all_null_runs_cyclic_shift, here::here("output","all_null_runs_cyclic_shift.Rds"))

#plot distribution of slopes for each survey vs. "true" slope
hist_null_slope_year_cyclic_shift <- ggplot(all_null_runs_cyclic_shift) +
  geom_histogram(aes(x = Year_slope)) +
  geom_vline(aes(xintercept = true_slope), color = "red") +
  geom_vline(aes(xintercept = upper_95), linetype = "dashed", color = "grey") +
  geom_vline(aes(xintercept = lower_95), linetype = "dashed", color = "grey") +
  facet_wrap(~Survey_Name_Season, scales = "free", ncol = 5) +
  labs(y = "Count",x = "β-diversity trend") +
  theme_classic() +
  theme(axis.text.x = element_text(size = 8))

ggsave(hist_null_slope_year_cyclic_shift, filename = "hist_null_slope_year_cyclic_shift.jpg", path = here::here("figures"), height = 10, width = 12)

#unique values per region
all_null_runs_cyclic_shift.u <- unique(all_null_runs_cyclic_shift[,.(survey_unit, Survey_Name_Season,mean_slope, upper_95, lower_95, true_slope, lwr, upr, Directional_Change, significant, Directional_Change_sig, trend_color)])

#lollipops
errorbar_slope_year_cyclic_shift <- ggplot() +
  geom_errorbar(data = all_null_runs_cyclic_shift.u, aes(x = reorder(Survey_Name_Season, true_slope) , y = mean_slope, label = Survey_Name_Season, ymin = lower_95, ymax = upper_95), color = "grey50", width = 0) + #add confidence intervals
  geom_point(data = all_null_runs_cyclic_shift.u, aes(x = reorder(Survey_Name_Season, true_slope) , y = true_slope, label = Survey_Name_Season, color = Directional_Change_sig), shape = 18, size = 3) +
  scale_color_manual(values = c("#FFC109","#D81C60","#1E88E5"), name = "Dissimilarity trend", guide="none") +
  geom_hline(yintercept = 0) +
 # scale_y_continuous(breaks = seq(-0.005, 0.0075, by = 0.0025), labels = c("-0.005","","0", "", "0.005",  "")) +
  xlab("Survey unit") +
  ylab("β-diversity trend") + #total BC dissimilarity trend
  coord_flip() +
  theme_classic() +
  theme(axis.text.y = element_text(face = "bold"), axis.title.y = element_blank(), axis.text.x = element_text(size = 15), axis.title.x = element_text(size = 15), legend.position = c(0.1,0.8), legend.direction = "vertical", legend.text = element_text(size = 15), legend.title = element_text(size = 16))

ggsave(errorbar_slope_year_cyclic_shift, filename = "errorbar_slope_year_cyclic_shift.jpg", path = here::here("figures"), height = 5, width = 5)

```
Merge these two nulls
```{r}

errorbar_null_models_merge <- plot_grid(errorbar_slope_year_reshuffle + theme(axis.title.x = element_blank(), axis.text.x =element_blank()), errorbar_slope_year_cyclic_shift, ncol = 1, labels = c("a.","b."), vjust = 1, hjust = -0.2)

ggsave(errorbar_null_models_merge, filename = "errorbar_null_models_merge.jpg", path = here::here("figures"), height = 9, width = 10)
```

