---
title: "Year Gradient Dissimilarity Models (subset/nestedness)"
output: html_notebook
---

This code is Script 4 v2 for Kitchel et al. TITLE manuscript.

(We expect this to be the main analyses in the text)

- This project is a collaborative effort to describe changes in taxonomic composition  of fish communities around the world--as sampled by bottom trawl surveys.

- Code by Zoë J. Kitchel

SESSION INFO TO DO

```{r setup}
library(data.table)
library(vegan)
library(sf)
library(concaveman) #polygon around points
library(betapart) #allows us to partition beta diversity
library(geosphere)
library(ggpubr) #stat_regline_equation
library(nlme)
library(mgcv) #to make gam
library(cowplot)

#Pull Dissimilarity Means
distances_dissimilarities_allyears.r <- readRDS(here::here("output", "distance_decay", "distances_dissimilarities_allyears.r.rds"))

#make survey and survey unit factors
distances_dissimilarities_allyears.r[,survey:=factor(survey)][,survey_unit:=factor(survey_unit)]

#adjust years
distances_dissimilarities_allyears.r[,year_adj := year-min(year)+1] #(this helps models converge)

#add new variable for year in sequence per region
distances_dissimilarities_allyears.r[,first_year := min(year),.(survey_unit)]
distances_dissimilarities_allyears.r[,last_year := max(year),.(survey_unit)]

#distances_dissimilarities_allyears.r[,year_in_seq := year-first_year+1]

distances_dissimilarities_allyears.r[,years_sampled := last_year-first_year+1]


```

##Make GAMs

Bray Curtis
```{r bray curtis gams}
bray_curtis_gradient_gam <- gam(bray_curtis_dissimilarity_gradient_mean ~ year + s(survey_unit, year, bs = "fs", m = 1),#random smooth
                            data = distances_dissimilarities_allyears.r)

```

##Make LMERS

Bray
*These all converged*
```{r bray}
bray_curtis_gradient_lmer <- lmer(bray_curtis_dissimilarity_gradient_mean ~ year_adj + (1 + year_adj|survey_unit),data = distances_dissimilarities_allyears.r)

summary(bray_curtis_gradient_lmer)

```

Get LMER model as predictions
```{r}

# need to sort out year in seq versus overall year models
#new data for lmer
lmer_year <- seq(min(distances_dissimilarities_allyears.r[,year]), max(distances_dissimilarities_allyears.r[,year]), by = 0.1)

lmer_year_adj <- seq(min(distances_dissimilarities_allyears.r[,year_adj]), max(distances_dissimilarities_allyears.r[,year_adj]), by = 0.1)

#predict average lmer
lmer_bray_gradient_predictions <- data.table(year = lmer_year, year_adj = lmer_year_adj)

#confidence intervals
bray_curtis_gradient_lmer_confint <- confint(bray_curtis_gradient_lmer)

#populate data table of lmer predictions
lmer_bray_gradient_predictions[,bray_curtis_lmer_preds := fixef(bray_curtis_gradient_lmer)[[1]] + fixef(bray_curtis_gradient_lmer)[[2]] * year_adj][,bray_curtis_lmer_preds_lowerCI := bray_curtis_gradient_lmer_confint[5] + bray_curtis_gradient_lmer_confint[6] * year_adj][,bray_curtis_lmer_preds_upperCI := bray_curtis_gradient_lmer_confint[11] + bray_curtis_gradient_lmer_confint[12] * year_adj]
```


###Palette for Plotting
Palette for plotting all 37 survey units
```{r link colors to survey units}
survey_unit.list <- levels(distances_dissimilarities_allyears.r[,survey_unit])

palette_37 <- c(
  "#5A5156", #AI
  "#F6222E", #CHL
  "#F8A19F", #DFO-NF
  "#16FF32", #DFO-QCS
  "#DF00DB", #BITS-1
  "#DB8EDA", #BITS-4
  "#325A9B", #EBS
  "#3283FE", #EVHOE
  "#FEAF16", #FR-CGFS
  "#1C8356", #GMEX-Summer
  "#C4451C", #GOA
  "#85660D", #GRL-DE
  "#B0009F", #GSL-N
  "#BF79B8", #GSL-S
  "#1CBE4F", #ICE-GFS
  "#782AB6", #IE-IGFS
  "#90AD1C", #MEDITS
  "#6B003A", #NAM
  "#A75B00", #NEUS-Fall
  "#E3B072", #NEUS-Spring
  "#02E8B6", #NIGFS-1
  "#97E7D5", #NIGFS-4
  "#B00068", #Nor-BTS-3
  "#00B9E3", #NS-IBTS-1
  "#95E2F4", #NS-IBTS-3
  "#B3CE73", #NZ-CHAT
  "#689500", #NZ-ECSI
  "#AAF400", #NZ-WCSI
  "#AA0DFE", #PT-IBTS
  "#FA0087", #S-GEORG
  "#DEA0FD", #SCS-Summer
  "#FCEF88", #SEUS-fall
  "#A59405", #SEUS-spring
  "#FCE100", #SEUS-summer
  "#C075A6", #WCANN
  "#BDCDFF", #ZAF-ATL
  "#003EFF"  #ZAF-IND
)

color_link <- data.table(survey_unit = survey_unit.list,hex = palette_37)
```

Add names for plotting
```{r add names for plotting}

color_link[,Survey_Name_Season := factor(c("Aleutian Islands",
                                    "Baltic Sea Q1",
                                    "Baltic Sea Q4",
                                    "Chile",
                                    "Newfoundland",
                                    "Queen Charlotte Sound",
                                    "Eastern Bering Sea",
                                    "Bay of Biscay",
                                    "English Channel",
                                    "Gulf of Mexico",
                                    "Gulf of Alaska",
                                    "Greenland",
                                    "N Gulf of St. Lawrence",
                                    "S Gulf of St. Lawrence",
                                    "Iceland",
                                    "Irish Sea",
                                    "Mediterranean",
                                    "Namibia",
                                    "NE US Fall",
                                    "NE US Spring",
                                    "N Ireland Q1",
                                    "N Ireland Q4",
                                    "Norway",
                                    "N Sea Q1",
                                    "N Sea Q3",
                                    "Chatham Rise",
                                    "E Coast S Island NZ",
                                    "W Coast S Island NZ",
                                    "Portugal",
                                    "S Georgia Straight",
                                  "Scotian Shelf",
                                  "SE US Fall",
                                  "SE US Spring",
                                  "SE US Summer",
                                  "W Coast US",
                                  "Atlantic Ocean ZA",
                                  "Indian Ocean ZA"))]

```


###Move forward with Bray Curtis Gradient (likelyto supplement)

Coefficients for LMER by survey_unit
```{r}
#unique survey unit year combos
survey_unit_sampling_years <- unique(distances_dissimilarities_allyears.r[,.(survey_unit, year_adj, year, years_sampled)])

# see group coefficients
model_coefs_reduced <- data.table(transform(as.data.frame(ranef(bray_curtis_gradient_lmer)), lwr = condval - 1.96*condsd, upr = condval + 1.96*condsd))
#https://stackoverflow.com/questions/69805532/extract-the-confidence-intervals-of-lmer-random-effects-plotted-with-dotplotra


#ONLY SLOPES
model_coefs_reduced <- model_coefs_reduced[term == "year_adj",]

model_coefs_reduced[,survey_unit := grp][,year_adj := condval]

#merge with duration of survey
model_coefs_reduced_length <- model_coefs_reduced[survey_unit_sampling_years, on = "survey_unit"]


model_coefs_reduced_length[,years_sampled := as.numeric(years_sampled)][,Directional_Change := ifelse(year_adj > 0, "Differentiation","Homogenization")]

#min max distances_dissimilarities
min_bray_reduced <- min(distances_dissimilarities_allyears.r[,bray_curtis_dissimilarity_gradient_mean], na.rm = T)
max_bray_reduced <- max(distances_dissimilarities_allyears.r[,bray_curtis_dissimilarity_gradient_mean], na.rm = T)

model_coefs_reduced_length <- model_coefs_reduced_length[color_link, on = "survey_unit"]


#does it cross zero?
model_coefs_reduced_length[,significant := ifelse(lwr >0 & upr>0,T,ifelse(lwr<0 & upr<0,T,F))]

#delete all obs that are significant
model_coefs_reduced_length.r <- model_coefs_reduced_length[significant == F,]

#order table by coefficient
setorder(model_coefs_reduced_length, year_adj)

BC_gradient_model_coefs_reduced_length.unique <- unique(model_coefs_reduced_length[,.(condval,condsd, lwr, upr, survey_unit, year_adj, years_sampled, Directional_Change, hex, Survey_Name_Season, significant)]) 

#extract color hexes
#year adj coef order
color_year_adj_order <- BC_gradient_model_coefs_reduced_length.unique[,hex]

#alphabetical order
BC_gradient_model_coefs_reduced_length.unique.alpha <- setorder(BC_gradient_model_coefs_reduced_length.unique, Survey_Name_Season)
color_alpha_order <- BC_gradient_model_coefs_reduced_length.unique.alpha[,hex]

saveRDS(BC_gradient_model_coefs_reduced_length.unique, here::here("output","region_stats","BC_gradient_model_coefs_reduced_length.unique.Rds"))
```

Bar Plot Coefficient LMER
```{r bar plot of coefficients}
BC_GRADIENT_Dissimilarity_Coef_errorbar_reduced <- ggplot() +
    geom_errorbar(data = model_coefs_reduced_length, aes(x = reorder(Survey_Name_Season, year_adj) , y = year_adj, label = Survey_Name_Season, ymin = lwr, ymax = upr), fill = "grey") + #add confidence intervals
  geom_point(data = model_coefs_reduced_length, aes(x = reorder(Survey_Name_Season, year_adj) , y = year_adj, label = Survey_Name_Season, size = years_sampled, fill = Directional_Change), stat = 'identity', shape = 21, color = "black") +
  scale_fill_manual(values = c("white","black"), name = "Directional Change") +
  geom_point(data = model_coefs_reduced_length.r, aes(x = reorder(Survey_Name_Season, year_adj) , y = year_adj, label = Survey_Name_Season, size = years_sampled), stat = 'identity', fill = "grey",color = "grey", shape = 21) +
  scale_size_binned(range = c(1,8), name = "Survey Period Length") +
  geom_hline(yintercept = 0) +
  xlab("Survey unit") +
  ylab("Gradient BC dissimilarity trend") +
  coord_flip() +
  theme_classic() +
  theme(axis.text.y = element_text(colour = color_year_adj_order, face = "bold"), axis.title.y = element_blank())

BC_GRADIENT_Dissimilarity_Coef_errorbar_reduced
```

Wavy Line Plot for GAMs

Generate predicted values
```{r generate predicted values GAM}

#add colors and names to full dissimilarity data table
distances_dissimilarities_allyears.r <- distances_dissimilarities_allyears.r[color_link, on = "survey_unit"]

#generate new data to smooth lines (need year and season survey combinations)
year_survey_unit_expand.dt <- data.table(survey_unit = as.character(NULL), year = as.numeric(NULL), year_adj = as.numeric(NULL ))

for (i in 1:length(survey_unit.list)) {
  #generate year vectors
  survey_unit_years <- unique(distances_dissimilarities_allyears.r[survey_unit == survey_unit.list[i],.(survey_unit, year, year_adj)])
  
  years <- seq(min(survey_unit_years[,year]), max(survey_unit_years[,year]), by = 0.1)
  
  year_adjust <- seq(min(survey_unit_years[,year_adj]), max(survey_unit_years[,year_adj]), by = 0.1)
  
  year_survey_unit_expand.dt_addition <- data.table(survey_unit = survey_unit.list[i], year = years, year_adj = year_adjust)
  
  year_survey_unit_expand.dt <- rbind(year_survey_unit_expand.dt, year_survey_unit_expand.dt_addition)
}

#add colors and names to full year and survey unit combination table
year_survey_unit_expand.dt <- year_survey_unit_expand.dt[color_link, on = "survey_unit"]
```



Get model as predictions
```{r}
#for plotting, get model as predictions
bray_curtis_gradient_gam_predictions <- predict(bray_curtis_gradient_gam, se.fit = TRUE, newdata = year_survey_unit_expand.dt)

#merge into table
year_survey_unit_expand.dt[,bray_glm_mod_fit := bray_curtis_gradient_gam_predictions$fit][,bray_glm_mod_fit_SE := bray_curtis_gradient_gam_predictions$se.fit]

```


Produce Plot of GAM and mean LMER line
```{r plot GAM and mean LMER lines}
points_wavylines_bray_gradient_year_reduced_gam_nolmer <- ggplot() +
 # geom_ribbon(data = lmer_bray_gradient_predictions, aes(x = year, ymin = bray_curtis_lmer_preds_lowerCI, ymax = bray_curtis_lmer_preds_upperCI), fill = "grey", alpha = 0.2) +
  geom_point(data = distances_dissimilarities_allyears.r,
             aes(x = year,
                 y = bray_curtis_dissimilarity_gradient_mean,
                 color = Survey_Name_Season), alpha = 0.5, size = 1) +
    geom_line(data = year_survey_unit_expand.dt,
             aes(x = year,
                 y = bray_glm_mod_fit,
                 color = Survey_Name_Season)) +
  geom_ribbon(data = year_survey_unit_expand.dt, aes(x = year, ymin=bray_glm_mod_fit-bray_glm_mod_fit_SE, ymax=bray_glm_mod_fit+bray_glm_mod_fit_SE, fill =  Survey_Name_Season), alpha=0.1) + #add standard error
 # geom_line(data = lmer_bray_gradient_predictions, aes(x = year, y = bray_curtis_lmer_preds), color = "black") +
    scale_color_manual(values =  color_alpha_order, name = "Survey Unit") +
  scale_fill_manual(values =  color_alpha_order, guide = "none") +
  theme_classic() +
  lims(x = c(min(distances_dissimilarities_allyears.r[,year]),max(distances_dissimilarities_allyears.r[,year])),
       y = c(0.05,0.45)) +
  xlab("Year") +
  ylab("Gradient BC dissimilarity") +
  theme(legend.position = "null")

points_wavylines_bray_gradient_year_reduced_gam_nolmer

#with lmer

points_wavylines_bray_gradient_year_reduced_gam <- ggplot() +
  geom_ribbon(data = lmer_bray_gradient_predictions, aes(x = year, ymin = bray_curtis_lmer_preds_lowerCI, ymax = bray_curtis_lmer_preds_upperCI), fill = "grey", alpha = 0.2) +
  geom_point(data = distances_dissimilarities_allyears.r,
             aes(x = year,
                 y = bray_curtis_dissimilarity_gradient_mean,
                 color = Survey_Name_Season), alpha = 0.5, size = 1) +
    geom_line(data = year_survey_unit_expand.dt,
             aes(x = year,
                 y = bray_glm_mod_fit,
                 color = Survey_Name_Season)) +
  geom_ribbon(data = year_survey_unit_expand.dt, aes(x = year, ymin=bray_glm_mod_fit-bray_glm_mod_fit_SE, ymax=bray_glm_mod_fit+bray_glm_mod_fit_SE, fill =  Survey_Name_Season), alpha=0.1) + #add standard error
  geom_line(data = lmer_bray_gradient_predictions, aes(x = year, y = bray_curtis_lmer_preds), color = "black") +
    scale_color_manual(values =  color_alpha_order, name = "Survey Unit") +
  scale_fill_manual(values =  color_alpha_order, guide = "none") +
  theme_classic() +
  lims(x = c(min(distances_dissimilarities_allyears.r[,year]),max(distances_dissimilarities_allyears.r[,year])),
       y = c(0.05,0.45)) +
  xlab("Year") +
  ylab("Gradient BC dissimilarity") +
  theme(legend.position = "null")

points_wavylines_bray_gradient_year_reduced_gam

#just one region that's homogenizing (NZ WCSI)
#Note, gradient is not significantly homogenizing, but it is for balanced, so we'll keep it
points_wavylines_bray_gradient_year_reduced_gam_NZ_only <- ggplot() +
  geom_point(data = distances_dissimilarities_allyears.r[survey_unit == "NZ-WCSI",],
             aes(x = year,
                 y = bray_curtis_dissimilarity_gradient_mean), alpha = 0.5, size = 1, color = BC_gradient_model_coefs_reduced_length.unique[Survey_Name_Season == "W Coast S Island NZ",hex]) +
    geom_line(data = year_survey_unit_expand.dt[survey_unit == "NZ-WCSI",],
             aes(x = year,
                 y = bray_glm_mod_fit),
             color = BC_gradient_model_coefs_reduced_length.unique[Survey_Name_Season == "W Coast S Island NZ",hex]) +
  geom_ribbon(data = year_survey_unit_expand.dt[survey_unit == "NZ-WCSI",], aes(x = year, ymin=bray_glm_mod_fit-bray_glm_mod_fit_SE, ymax=bray_glm_mod_fit+bray_glm_mod_fit_SE), alpha=0.3, fill = BC_gradient_model_coefs_reduced_length.unique[Survey_Name_Season == "W Coast S Island NZ",hex]) + #add standard error
 # geom_line(data = lmer_bray_gradient_predictions, aes(x = year, y = bray_curtis_lmer_preds), color = "black") +
    scale_color_manual(name = "Survey Unit") +
 # scale_fill_manual(values =  palette_37, guide = "none") +
    lims(x = c(min(distances_dissimilarities_allyears.r[,year]),max(distances_dissimilarities_allyears.r[,year])),
       y = c(0.05,0.45)) +
  theme_classic() +
  xlab("Year") +
  ylab("Gradient BC dissimilarity")

points_wavylines_bray_gradient_year_reduced_gam_NZ_only

#just two regions, another that's differentiating (SEUS Summer)
#Note, SEUS Summer is not differentiating for nestedness, but it is for balanced, so we'll keep anyway
points_wavylines_bray_gradient_year_reduced_gam_NZ_SEUS_Summer_only <- ggplot() +
  geom_point(data = distances_dissimilarities_allyears.r[survey_unit %in% c("NZ-WCSI","SEUS-summer"),],
             aes(x = year,
                 y = bray_curtis_dissimilarity_gradient_mean,
                 color = Survey_Name_Season), alpha = 0.5, size = 1) +
    geom_line(data = year_survey_unit_expand.dt[survey_unit %in% c("NZ-WCSI","SEUS-summer"),],
             aes(x = year,
                 y = bray_glm_mod_fit,
                 color = Survey_Name_Season)) +
  geom_ribbon(data = year_survey_unit_expand.dt[survey_unit %in% c("NZ-WCSI","SEUS-summer"),], aes(x = year, ymin=bray_glm_mod_fit-bray_glm_mod_fit_SE, ymax=bray_glm_mod_fit+bray_glm_mod_fit_SE, fill = Survey_Name_Season), alpha=0.3) + #add standard error
 # geom_line(data = lmer_bray_gradient_predictions, aes(x = year, y = bray_curtis_lmer_preds), color = "black") +
    scale_color_manual(name = "Survey Unit", values = c(BC_gradient_model_coefs_reduced_length.unique[Survey_Name_Season %in% c("W Coast S Island NZ","SE US Summer"),hex])) +
  scale_fill_manual(guide = "none", values = c(BC_gradient_model_coefs_reduced_length.unique[Survey_Name_Season %in% c("W Coast S Island NZ","SE US Summer"),hex])) +
    lims(x = c(min(distances_dissimilarities_allyears.r[,year]),max(distances_dissimilarities_allyears.r[,year])),
       y = c(0.05,0.45)) +
  theme_classic() +
  xlab("Year") +
  ylab("Gradient BC dissimilarity") +
  theme(legend.position = "null")

points_wavylines_bray_gradient_year_reduced_gam_NZ_SEUS_Summer_only

#just three regions, another one that's stable, (Canada Scotian Shelf)

points_wavylines_bray_gradient_year_reduced_gam_NZ_SEUS_ScoShelf_only <- ggplot() +
  geom_point(data = distances_dissimilarities_allyears.r[survey_unit %in% c("NZ-WCSI","SEUS-summer", "SCS-SUMMER"),],
             aes(x = year,
                 y = bray_curtis_dissimilarity_gradient_mean,
                 color = Survey_Name_Season), alpha = 0.5, size = 1) +
    geom_line(data = year_survey_unit_expand.dt[survey_unit %in% c("NZ-WCSI","SEUS-summer","SCS-SUMMER"),],
             aes(x = year,
                 y = bray_glm_mod_fit,
                 color = Survey_Name_Season)) +
  geom_ribbon(data = year_survey_unit_expand.dt[survey_unit %in% c("NZ-WCSI","SEUS-summer","SCS-SUMMER"),], aes(x = year, ymin=bray_glm_mod_fit-bray_glm_mod_fit_SE, ymax=bray_glm_mod_fit+bray_glm_mod_fit_SE, fill = Survey_Name_Season), alpha=0.3) + #add standard error
 # geom_line(data = lmer_bray_gradient_predictions, aes(x = year, y = bray_curtis_lmer_preds), color = "black") +
    scale_color_manual(name = "Survey Unit", values = c(BC_gradient_model_coefs_reduced_length.unique[Survey_Name_Season %in% c("W Coast S Island NZ","SE US Summer","Scotian Shelf"),hex])) +
  scale_fill_manual(guide = "none", values = c(BC_gradient_model_coefs_reduced_length.unique[Survey_Name_Season %in% c("W Coast S Island NZ","SE US Summer","Scotian Shelf"),hex])) +
    lims(x = c(min(distances_dissimilarities_allyears.r[,year]),max(distances_dissimilarities_allyears.r[,year])),
       y = c(0.05,0.45)) +
  theme_classic() +
  xlab("Year") +
  ylab("Gradient BC dissimilarity") +
  theme(legend.position = "null")

points_wavylines_bray_gradient_year_reduced_gam_NZ_SEUS_ScoShelf_only

ggsave(points_wavylines_bray_gradient_year_reduced_gam_NZ_only, path = here::here("figures"), filename ="points_wavylines_bray_gradient_year_reduced_gam_NZ_only.jpg", height = 5, width = 5, unit = "in")
ggsave(points_wavylines_bray_gradient_year_reduced_gam_NZ_SEUS_Summer_only, path = here::here("figures"), filename ="points_wavylines_bray_gradient_year_reduced_gam_NZ_SEUS_Summer_only.jpg", height = 5, width = 5, unit = "in")
ggsave(points_wavylines_bray_gradient_year_reduced_gam_NZ_SEUS_ScoShelf_only, path = here::here("figures"), filename ="points_wavylines_bray_gradient_year_reduced_gam_NZ_SEUS_ScoShelf_only.jpg", height = 5, width = 5, unit = "in")
points_wavylines_bray_gradient_year_reduced_gam
ggsave(points_wavylines_bray_gradient_year_reduced_gam_nolmer, path = here::here("figures"), filename ="points_wavylines_bray_gradient_year_reduced_gam_nolmer.jpg", height = 5, width = 5, unit = "in")

ggsave(points_wavylines_bray_gradient_year_reduced_gam, path = here::here("figures"), filename ="points_wavylines_bray_gradient_year_reduced_gam.jpg", height = 5, width = 5, unit = "in")

```

Merge BC versus Year plot with GAMS and Region vs. coefficient plot for LMERs

```{r}

BC_GRADIENT_GAM_LMER_merge <- plot_grid(points_wavylines_bray_gradient_year_reduced_gam + theme(legend.position = "none", plot.margin = margin(2,0,0,0, unit = "cm")), BC_GRADIENT_Dissimilarity_Coef_errorbar_reduced + theme(legend.position = "none", plot.margin = margin(2,0,0,0, unit = "cm")), labels = c("","Survey Unit  ",""), vjust = 1.5, hjust = 0)

BC_GRADIENT_GAM_LMER_merge_legend <- plot_grid(BC_GRADIENT_GAM_LMER_merge, get_legend(BC_GRADIENT_Dissimilarity_Coef_errorbar_reduced), rel_widths = c(10,2))

ggsave(BC_GRADIENT_GAM_LMER_merge_legend, path = here::here("figures"), filename = "BC_GRADIENT_GAM_LMER_merge_legend.png", height = 7, width = 14, units = "in")
```


