---
title: "Pulling SODA Bottom Temp Data"
output: html_notebook
---

This code is Script 5 v2 for Kitchel et al. TITLE manuscript.

- This project is a collaborative effort to describe changes in taxonomic composition  of fish communities around the world--as sampled by bottom trawl surveys.

- Code by Zoë J. Kitchel

SESSION INFO TO DO


SBT Temperature
1967-1979 = SODA 2.2.4
1980-2019 = SODA 3.3.2


```{r setup}
library(dplyr) # A staple for modern data management in R
library(lubridate) # Useful functions for dealing with dates
library(ggplot2) # The preferred library for data visualisation
library(tidync) # For easily dealing with NetCDF data
library(rerddap) # For easily downloading subsets of data
library(stringr)
library(data.table)

#pull in data for year and lat lon restrictions
FishGlob_clean <- readRDS(here::here("data", "cleaned", "FishGlob.wellsampledyearscells_complete.final.rds"))

#names of all survey units
all_survey_units <- unique(FishGlob_clean$survey_unit)


```

Unique Lat Lon for each Survey/season. 
```{r unique lat lon}
unique_lat_lon <- unique(FishGlob_clean[,.(latitude, longitude, survey_unit)])

#data range for SODA data is (0,360), so I will edit lat lon to match
unique_lat_lon[,longitude.shift := ifelse(longitude >= 0, longitude, 360+longitude)]

summary(unique_lat_lon[,longitude.shift])

#check to make sure shift makes sense
shift_check <- ggplot(unique_lat_lon, aes(x = longitude.shift, y = latitude)) + 
  geom_point(shape = ".") +
  theme_classic()


shift_check2 <- ggplot(unique_lat_lon, aes(x = longitude, y = latitude)) + 
  geom_point(shape = ".") +
  theme_classic()

shift_check
shift_check2

```
###Pull all date values for each Lat Lon
```{r}
extract(s,unique_lat_lon[,.(longitude.shift, latitude)])

unique_lat_lon.temp <- cbind(unique_lat_lon, extract(soda_bottom,unique_lat_lon[,.(longitude.shift, latitude)]))

#brings in dates as colnames
```

Wide to Long to make calculations
(Note, some temp values from unique_lat_lon.temp.l are NA, but only 1% so shouldn't make any difference)

```{r}
unique_lat_lon.temp.l <- melt(unique_lat_lon.temp,
            id.vars = c("latitude", "longitude","survey_unit","longitude.shift"),
            variable.name = "date", value.name = "sbt")
```
Delete X from Date column 
```{r}
unique_lat_lon.temp.l[, date.formatted := lubridate::ymd(substr(date,2,nchar(as.character(date))))]

unique_lat_lon.temp.l[, year := lubridate::year(date.formatted)][, month := lubridate::month(date.formatted)][, day := lubridate::day(date.formatted)]
```


###Average Values

First go from daily temp to average monthly temp (already the case for SODA 2.2.4 pre 1980). We need temp 12 months before haul date. We'll take yearly:
* mean
* max
* min
* seas
* SD of mean
* SD of max
* SD of min 
* SD of seas

And now all other regions
```{r loop for average values}

#to maximize memory, remove FishGlob_clean except for min_month
FishGlob_clean.months_surveys <- unique(FishGlob_clean[,.(survey_unit, month)])
FishGlob_clean.months_surveys[survey_unit == "MEDITS", month := 4]
FishGlob_clean.months_surveys[survey_unit == "GRL-DE", month := 8] #NEED TO CHECK THIS

rm(FishGlob_clean)

OISST_data_temp_avgs_full <- data.table()

for (i in 1:length(all_survey_units)) {
  min_month <- min(FishGlob_clean.months_surveys[survey_unit == all_survey_units[i]]$month)
  
  #pull in data from rds files if not already in console
  
  OISST_data <- data.table(readRDS(here::here("data","Temperature",paste0("OISST_data_",all_survey_units[i],".rds"))))
  
  #for annotate (HPC)
  OISST_data <- data.table(readRDS(here::here("data","Temperature",paste0("OISST_data_",all_survey_units[i],".rds"))))
  
  OISST_data[,month := month(t)][,year := year(t)]
  
#set year for taking temp measurements (12 months before first haul for that survey unit)

  OISST_data[, year_for_avg := ifelse(month >= min_month, year+1, year)]
  
  #next two steps take a lot of time
  
#take mean, min, max, and seas of each lat long point for each year 
  OISST_data[, yearly_mean_bypoint := mean(temp), .(year_for_avg, lat, lon)][,
              yearly_max_bypoint := max(temp), .(year_for_avg, lat, lon)][,                                                                yearly_min_bypoint := min(temp), .(year_for_avg, lat, lon)][,   
              yearly_seas_bypoint := yearly_max_bypoint-yearly_min_bypoint]
  
  #cuts down size of object from hundreds of millions to 
  OISST_data <- unique(OISST_data[,.(yearly_mean_bypoint, yearly_max_bypoint, yearly_min_bypoint, yearly_seas_bypoint, lat, lon, year_for_avg)])
  
#next, we want yearly avg of mean, min, max, and seas
    OISST_data[, yearly_mean_bypoint_avg := mean(yearly_mean_bypoint), .(year_for_avg)][,
              yearly_max_bypoint_avg := mean(yearly_max_bypoint), .(year_for_avg)][,                                                       yearly_min_bypoint_avg := mean(yearly_min_bypoint), .(year_for_avg)][,   
              yearly_seas_bypoint_avg := mean(yearly_seas_bypoint), .(year_for_avg)]
  
  #next, to calculate SD of mean, take mean of each cell in each year, and then calculate the SD of the means
  #to calculate SD of max, take max of each cell in each year, and then calculate the SD of the maxes, etc.
    OISST_data[, yearly_mean_bypoint_SD := sd(yearly_mean_bypoint), .(year_for_avg)][,
              yearly_max_bypoint_SD := sd(yearly_max_bypoint), .(year_for_avg)][,                                                       yearly_min_bypoint_SD := sd(yearly_min_bypoint), .(year_for_avg)][,   
              yearly_seas_bypoint_SD := sd(yearly_seas_bypoint), .(year_for_avg)]

   
  #finally, scaled versions by region
#next, we want yearly avg of mean, min, max, and seas
    OISST_data[, yearly_mean_bypoint_avg.s := scale(yearly_mean_bypoint_avg)][,
                 yearly_max_bypoint_avg.s := scale(yearly_max_bypoint_avg)][,                                                                        yearly_min_bypoint_avg.s := scale(yearly_min_bypoint_avg)][,   
                 yearly_seas_bypoint_avg.s := scale(yearly_seas_bypoint_avg)]
  
  #next, to calculate SD of mean, take mean of each cell in each year, and then calculate the SD of the means
  #to calculate SD of max, take max of each cell in each year, and then calculate the SD of the maxes, etc.
    OISST_data[, yearly_mean_bypoint_SD.s := scale(yearly_mean_bypoint_SD)][,
              yearly_max_bypoint_SD.s := scale(yearly_max_bypoint_SD)][,                                                                           yearly_min_bypoint_SD.s := scale(yearly_min_bypoint_SD)][,   
              yearly_seas_bypoint_SD.s := scale(yearly_seas_bypoint_SD)]
  
#uniquevalues
    OISST_data_temp_avgs <- unique(OISST_data[, .(year_for_avg, yearly_mean_bypoint_avg, yearly_max_bypoint_avg, yearly_min_bypoint_avg, yearly_seas_bypoint_avg, yearly_mean_bypoint_SD, yearly_max_bypoint_SD, yearly_min_bypoint_SD, yearly_seas_bypoint_SD, yearly_mean_bypoint_avg.s, yearly_max_bypoint_avg.s, yearly_min_bypoint_avg.s, yearly_seas_bypoint_avg.s, yearly_mean_bypoint_SD.s, yearly_max_bypoint_SD.s, yearly_min_bypoint_SD.s, yearly_seas_bypoint_SD.s)])
  
  OISST_data_temp_avgs[, survey_unit := all_survey_units[i]]
  
  OISST_data_temp_avgs_full <- rbind(OISST_data_temp_avgs_full, OISST_data_temp_avgs)

  
  print(all_survey_units[[i]]) #to help keep track


}


#save summary table!!!
saveRDS(SODA_data_temp_avgs_full, here::here("data","Temperature","SODA_data_temp_avgs_full.rds"))

#annotate (if working on HPC)
saveRDS(SODA_data_temp_avgs_full, here::here("SODA_data_temp_avgs_full.rds"))



```

NOT SURE WHY, but all data from SODA 2.2.4 is wonky, so I'm going to go ahead and only use temp data from 3.3.2, so from 1980 onward


