---
title: "Pulling OISST Temp Data"
output: html_notebook
---

Daily SST Temperature

```{r setup}
library(dplyr) # A staple for modern data management in R
library(lubridate) # Useful functions for dealing with dates
library(ggplot2) # The preferred library for data visualisation
library(tidync) # For easily dealing with NetCDF data
library(rerddap) # For easily downloading subsets of data
library(doParallel) # For parallel processing

#pull in data for year and lat lon restrictions
FishGlob_cleaned.10year.r <- readRDS(here::here("output","region_season_cleaned_data","FishGlob_cleaned.10year.r.rds"))
all_season_surveys <- unique(FishGlob_cleaned.10year.r[,survey_season])
```

Unique Lat Lon Year for each Survey/season
```{r unique lat lon}
unique_lat_lon_year <- unique(FishGlob_cleaned.10year.r[,.(latitude, longitude, year, survey, season, survey_season)])

#data range for this data is (-179.875, 179.875), so I will trim to this
unique_lat_lon_year[,longitude_trim := ifelse(longitude < -179.875, -179.875, ifelse(longitude > 179.875, 179.875, longitude))]

summary(unique_lat_lon_year[,longitude_trim])
```


```{r}
# The information for the NOAA OISST data
rerddap::info(datasetid = "ncdcOisst21Agg_LonPM180", url = "https://coastwatch.pfeg.noaa.gov/erddap/")

# Note that there is also a version with lon values from 0 yo 360
rerddap::info(datasetid = "ncdcOisst21Agg", url = "https://coastwatch.pfeg.noaa.gov/erddap/")
```

```{r}
# This function downloads and prepares data based on user provided start and end dates
OISST_sub_dl <- function(time_df, this_survey_season){
  OISST_dat <- griddap(x = "ncdcOisst21Agg_LonPM180", 
                       url = "https://coastwatch.pfeg.noaa.gov/erddap/", 
                       time = c(time_df$start, time_df$end), 
                       zlev = c(0, 0),
                       latitude = c(min(unique_lat_lon_year[survey_season == this_survey_season]$latitude),
                                    max(unique_lat_lon_year[survey_season == this_survey_season]$latitude)),
                       longitude = c(min(unique_lat_lon_year[survey_season == this_survey_season]$longitude_trim),
                                    max(unique_lat_lon_year[survey_season == this_survey_season]$longitude_trim)),
                       fields = "sst")$data %>% 
    mutate(time = as.Date(stringr::str_remove(time, "T00:00:00Z"))) %>% 
    dplyr::rename(t = time, temp = sst) %>% 
    select(lon, lat, t, temp) %>% 
    na.omit()
}
```

```{r}
    
dl_years <- data.frame(date_index = 1:5,
                           start = as.Date(c("1981-01-01", "1989-01-01", 
                                             "1997-01-01", "2005-01-01", "2013-01-01")),
                           end = as.Date(c("1988-12-31", "1996-12-31", 
                                           "2004-12-31", "2012-12-31", "2019-12-31")))


}
```

```{r}
# Download all of the data with one nested request
# The time this takes will vary greatly based on connection speed

for (i in 1:all_season_surveys)
    system.time(
      OISST_data <- dl_years %>% 
        group_by(date_index) %>% 
        group_modify(~OISST_sub_dl(.x, all_season_surveys[i])) %>% 
        ungroup() %>% 
        select(lon, lat, t, temp)
    ) # 38 seconds, ~8 seconds per batch

    assign(paste0("OISST_data_",all_season_surveys[i]), OISST_data)

OISST_data_AI <- as.data.table(OISST_data)

AI_min_positive <- min(unique_lat_lon_year[longitude > 150 & survey == "AI"]$longitude)
AI_max_negative <-max(unique_lat_lon_year[longitude < 0 & survey == "AI"]$longitude)

OISST_data_AI <- OISST_data_AI[lon >= AI_min_positive | lon <= AI_max_negative,][,lon_s := ifelse(lon > 160, lon-360, lon)]

#edit, because we don't actually need all of this data
hist(FishGlob_cleaned.10year.r[survey == "AI", longitude])
#for just AI 
#2024.317  159.159 4585.945 
```

```{r}
OISST_data_AI %>% 
  filter(t == "2018-12-01") %>% 
  ggplot(aes(x = lon_s, y = lat)) +
  geom_tile(aes(fill = temp)) +
  borders() + # Activate this line to see the global map
  scale_fill_viridis_c() +
  coord_quickmap(expand = F) +
  labs(x = NULL, y = NULL, fill = "SST (°C)") +
  theme(legend.position = "bottom")
```
Average Values

```{r}
#what time of year does AI trawl usually happen?
plot(FishGlob_cleaned.10year.r[survey == "AI"]$year, FishGlob_cleaned.10year.r[survey == "AI"]$month)

#we'll go with May
```
First go from daily temp to average monthly temp. We need mean temp 12 months before May, max temp 12 months before May, and min temp 12 months before June. At some time in the future, I may want to extract values at exact temp values, but not sure it matters much now. Instead, I'll take yearly:
* mean
* max
* min
* SD

```{r}
OISST_data_AI[,month := month(t)][,year := year(t)]

OISST_data_AI[, year_for_avg := ifelse(month >= 6, year+1, year_for_avg)][, yearly_mean_bypoint := mean(temp), .(year_for_avg, lat, lon)][, montly_mean_bypoint := mean(temp),.(year_for_avg,lat,lon,month)]

#for min, max, and mean, first temp is avg by month at each lat/lon, THEN we take mean, max, and min monthly temp per year (balances out any weird values)
#for SD, we calculate a yearly mean by lat/lon, and then look at standard deviation
OISST_data_AI[, sst_min := min(montly_mean_bypoint), year_for_avg][, sst_max := max(montly_mean_bypoint), year_for_avg][, sst_mean := mean(montly_mean_bypoint), year_for_avg][, sst_SD := sd(yearly_mean_bypoint), year_for_avg] #this doesn't take into consideration spatial component for sst min max or mean, I want to calculate mean at each point, and THEN take SD I think

OISST_data_AI_temp_avgs <- unique(OISST_data_AI[,.(year_for_avg, sst_min, sst_max, sst_mean, sst_SD)])

plot(OISST_data_AI_temp_avgs$year, OISST_data_AI_temp_avgs$sst_mean)
abline()
plot(OISST_data_AI_temp_avgs$year, OISST_data_AI_temp_avgs$sst_max)
plot(OISST_data_AI_temp_avgs$year, OISST_data_AI_temp_avgs$sst_min)
plot(OISST_data_AI_temp_avgs$year, OISST_data_AI_temp_avgs$sst_SD)

#save to play with tomorrow
saveRDS(OISST_data_AI, here::here("data","Temperature","OISST_data_AI.rds"))
saveRDS(OISST_data_AI_temp_avgs, here::here("data","Temperature","OISST_data_AI_temp_avgs.rds"))
```




I'll delete any observations from october and november sometime in the future? Not now

```{r}
system.time(
  OISST_data <- dl_years %>% 
    group_by(date_index) %>% 
    group_modify(~OISST_sub_dl(.x)) %>% 
    ungroup() %>% 
    select(lon, lat, t, temp)
) # 38 seconds, ~8 seconds per batch
```

```{r}
# Save the data as an .Rds file because it has a much better compression rate than .RData
saveRDS(OISST_data, file = "~/Desktop/OISST_vignette.Rds")
```

