---
title: "Pulling SODA Bottom Temp Data"
output: html_notebook
---

This code is Script 5 v1 step 2 for Kitchel et al. Biotic homogenization, the exception and not the rule for marine fish communities manuscript.

- This project is a collaborative effort to describe changes in taxonomic composition  of fish communities around the world--as sampled by bottom trawl surveys.

- Code by Zoë J. Kitchel

SESSION INFO TO DO

This script uses lat, long, survey, date data from FishGlob data to extract relevant bottom temperature values (SODA 3.3.2) from raster. It is best run on HPC because this raster is several GB large. 


SBT Temperature
1980-2019 = SODA 3.3.2


```{r setup}
library(dplyr) # A staple for modern data management in R
library(lubridate) # Useful functions for dealing with dates
library(ggplot2) # The preferred library for data visualisation
library(tidync) # For easily dealing with NetCDF data
library(rerddap) # For easily downloading subsets of data
library(stringr)
library(data.table)

#pull in data for year and lat lon restrictions
FishGlob_clean.singleseason <- readRDS(here::here("data", "cleaned", "FishGlob_clean.singleseason.rds"))

#names of all survey units
all_survey_units <- unique(FishGlob_clean.singleseason.singleseason.rds$survey_unit)


```

Unique Lat Lon for each Survey/season. 
```{r unique lat lon}
unique_lat_lon <- unique(FishGlob_clean.singleseason.singleseason.rds[,.(latitude, longitude, survey_unit)])

#data range for SODA data is (0,360), so I will edit lat lon to match
unique_lat_lon[,longitude.shift := ifelse(longitude >= 0, longitude, 360+longitude)]

summary(unique_lat_lon[,longitude.shift])

#check to make sure shift makes sense
shift_check <- ggplot(unique_lat_lon, aes(x = longitude.shift, y = latitude)) + 
  geom_point(shape = ".") +
  theme_classic()


shift_check2 <- ggplot(unique_lat_lon, aes(x = longitude, y = latitude)) + 
  geom_point(shape = ".") +
  theme_classic()

shift_check
shift_check2

```

###Pull all date values for each Lat Lon from SODA raster (soda_bottom_jan_2023.grd)
```{r}
#file path for annotate server
SODA_raster <- stack(file.path("OISST_temp_data_homogenization","soda_bottom_jan_2023.grd"))

SODA_values <- extract(SODA_raster,unique_lat_lon[,.(longitude.shift, latitude)])

unique_lat_lon.temp <- cbind(unique_lat_lon, SODA_values)

#brings in dates as colnames
```

Wide to Long to make calculations
(Note, some temp values from unique_lat_lon.temp.l are NA, but only 1% so shouldn't make any difference)

```{r}
unique_lat_lon.temp.l <- melt(unique_lat_lon.temp,
            id.vars = c("latitude", "longitude","survey_unit","longitude.shift"),
            variable.name = "date", value.name = "sbt")
```

Delete X from Date column 
```{r}
unique_lat_lon.temp.l[, date.formatted := lubridate::ymd(substr(date,2,nchar(as.character(date))))]

unique_lat_lon.temp.l[, year := lubridate::year(date.formatted)][, month := lubridate::month(date.formatted)][, day := lubridate::day(date.formatted)]
```





###Average Values

First go from daily temp to average monthly temp. We need temp 12 months before haul date. We'll take yearly:
* mean
* max
* min
* seas
* SD of mean
* SD of max
* SD of min 
* SD of seas

And now all other regions
```{r loop for average values}

#to maximize memory, remove FishGlob_clean.singleseason.singleseason.rds except for min_month
FishGlob_clean.singleseason.singleseason.rds.months_surveys <- unique(FishGlob_clean.singleseason.singleseason.rds[,.(survey_unit, month)])
FishGlob_clean.singleseason.singleseason.rds.months_surveys[survey_unit == "MEDITS", month := 4]
FishGlob_clean.singleseason.singleseason.rds.months_surveys[survey_unit == "GRL-DE", month := 8] #NEED TO CHECK THIS

rm(FishGlob_clean.singleseason.singleseason.rds)

SODA_data_temp_avgs_full <- data.table()

for (i in 1:length(all_survey_units)) {
  min_month <- min(FishGlob_clean.singleseason.singleseason.rds.months_surveys[survey_unit == all_survey_units[i]]$month)
  
  #pull in data from rds files if not already in console
  
  
  unique_lat_lon.temp.l[,month := month(t)][,year := year(t)]
  
#set year for taking temp measurements (12 months before first haul for that survey unit)

  unique_lat_lon.temp.l[, year_for_avg := ifelse(month >= min_month, year+1, year)]
  
    #delete any rows with NA
  unique_lat_lon.temp.l.bysurvey <- na.omit(unique_lat_lon.temp.l.bysurvey, cols = "sbt")
  
  #next two steps take a lot of time
  
#take mean, min, max, and seas of each latitude longitude.shiftg point for each year 
  unique_lat_lon.temp.l[, yearly_mean_bypoint := mean(temp), .(year_for_avg, latitude, longitude.shift)][,
              yearly_max_bypoint := max(temp), .(year_for_avg, latitude, longitude.shift)][,                                                                yearly_min_bypoint := min(temp), .(year_for_avg, latitude, longitude.shift)][,   
              yearly_seas_bypoint := yearly_max_bypoint-yearly_min_bypoint]
  
  #cuts down size of object from hundreds of millions to 
  SODA_data <- unique(unique_lat_lon.temp.l[,.(yearly_mean_bypoint, yearly_max_bypoint, yearly_min_bypoint, yearly_seas_bypoint, latitude, longitude.shift, year_for_avg)])
  
#next, we want yearly avg of mean, min, max, and seas
    unique_lat_lon.temp.l[, yearly_mean_bypoint_avg := mean(yearly_mean_bypoint), .(year_for_avg)][,
              yearly_max_bypoint_avg := mean(yearly_max_bypoint), .(year_for_avg)][,                                                       yearly_min_bypoint_avg := mean(yearly_min_bypoint), .(year_for_avg)][,   
              yearly_seas_bypoint_avg := mean(yearly_seas_bypoint), .(year_for_avg)]
  
  #next, to calculate SD of mean, take mean of each cell in each year, and then calculate the SD of the means
  #to calculate SD of max, take max of each cell in each year, and then calculate the SD of the maxes, etc.
    unique_lat_lon.temp.l[, yearly_mean_bypoint_SD := sd(yearly_mean_bypoint), .(year_for_avg)][,
              yearly_max_bypoint_SD := sd(yearly_max_bypoint), .(year_for_avg)][,                                                       yearly_min_bypoint_SD := sd(yearly_min_bypoint), .(year_for_avg)][,   
              yearly_seas_bypoint_SD := sd(yearly_seas_bypoint), .(year_for_avg)]

   
  #finally, scaled versions by region
#next, we want yearly avg of mean, min, max, and seas
    unique_lat_lon.temp.l[, yearly_mean_bypoint_avg.s := scale(yearly_mean_bypoint_avg)][,
                 yearly_max_bypoint_avg.s := scale(yearly_max_bypoint_avg)][,                                                                        yearly_min_bypoint_avg.s := scale(yearly_min_bypoint_avg)][,   
                 yearly_seas_bypoint_avg.s := scale(yearly_seas_bypoint_avg)]
  
  #to calculate SD of max, take max of each cell in each year, and then calculate the SD of the maxes, etc.
    unique_lat_lon.temp.l[, yearly_mean_bypoint_SD.s := scale(yearly_mean_bypoint_SD)][,
              yearly_max_bypoint_SD.s := scale(yearly_max_bypoint_SD)][,                                                                           yearly_min_bypoint_SD.s := scale(yearly_min_bypoint_SD)][,   
              yearly_seas_bypoint_SD.s := scale(yearly_seas_bypoint_SD)]
  
#uniquevalues
    SODA_data_temp_avgs <- unique(unique_lat_lon.temp.l[, .(year_for_avg, yearly_mean_bypoint_avg, yearly_max_bypoint_avg, yearly_min_bypoint_avg, yearly_seas_bypoint_avg, yearly_mean_bypoint_SD, yearly_max_bypoint_SD, yearly_min_bypoint_SD, yearly_seas_bypoint_SD, yearly_mean_bypoint_avg.s, yearly_max_bypoint_avg.s, yearly_min_bypoint_avg.s, yearly_seas_bypoint_avg.s, yearly_mean_bypoint_SD.s, yearly_max_bypoint_SD.s, yearly_min_bypoint_SD.s, yearly_seas_bypoint_SD.s)])
  
  SODA_data_temp_avgs[, survey_unit := all_survey_units[i]]
  
  SODA_data_temp_avgs_full <- rbind(SODA_data_temp_avgs_full, SODA_data_temp_avgs)

  
  print(all_survey_units[[i]]) #to help keep track


}


#save summary table!!!
saveRDS(SODA_data_temp_avgs_full, here::here("data","Temperature","SODA_data_temp_avgs_full.rds"))

#annotate (if working on HPC)
saveRDS(SODA_data_temp_avgs_full, here::here("SODA_data_temp_avgs_full.rds"))

summary(SODA_data_temp_avgs_full)

#check histogram

ggplot() +
  geom_histogram(data = SODA_data_temp_avgs_full, aes(x = yearly_min_bypoint_avg)) +
  facet_wrap(~survey_unit, scales = "free") +
  theme_classic()
```

NOT SURE WHY, but all data from SODA 2.2.4 is wonky, so I'm going to go ahead and only use temp data from 3.3.2, so from 1981 onward (doesn't leave out much! Only NEUS starts before 1981)


