---
title: "Sample Based Rarefaction NEUS"
output: html_notebook
---



```{r setup}
library(data.table)

#pull in NEUS data from 1982 forward (see Batt et al. for reasoning)
neus_data_gridded <- readRDS("/Users/zoekitchel/Documents/grad school/Rutgers/Repositories/trawl_spatial_turnover/data/gridded/USA-NEUS/neus_data_gridded.RData")
```

This is a medium scale, center of each hexagon = 111km away

How many tows for each year/cell combination?
```{r tow count for each year/cell combination}
unique_tows <- neus_data_gridded[,.N, .(year,cell,haulid)]
unique_year_cell <- unique_tows[,.N, .(year, cell)]

#minimum count?
min(unique_year_cell[,N]) #3
max(unique_year_cell[,N]) #31

```
The minimum tow count within a year/cell combination is 3, therefore, I will randomly select 3 tows from each year/cell

Generate 3 random haulids for each year/cell combo
*how to take seasons into account? should when in the year matter? probably yes, but will deal with later
```{r randomly select three haulids}
unique_tows

set.seed(10) #this means I'll get same random sample every time

random_tows <- unique_tows[,.SD[sample(.N, min(3,.N))],.(year,cell)]

```

Now I know which tow id's to keep

```{r narrow down original dataset to be for these randomly selected tows}
tow_ids <- unique(random_tows[,haulid])

#subset original dataset with these rows
subset_neus_full_gridded <- neus_data_gridded[haulid %in% tow_ids]
```

Save this reduced dataset
```{r save subsample}
save(subset_neus_full_gridded, file = "/Users/zoekitchel/Documents/grad school/Rutgers/Repositories/trawl_spatial_turnover/data/gridded/USA-NEUS/subset_neus_full_gridded.RData")
```


Now, I'm ready to go forward with analyses! Afterwards, I will loop back and do this process 1000 times. 