---
title: "Pull together sea around us data"
output: html_notebook
---

Pull together sea around us data from the web by closest EEZ

Script 6 for Kitchel et al. 2023 in prep taxonomic diversity manuscript.


```{r setup}
library(data.table)
library(ggplot2)
library(here)

#key
SAU_survey_key <- fread(here::here("data_prep_code","fishing","Survey_SOU_Key.csv"))

#delete rockall
SAU_survey_key <- SAU_survey_key[survey_unit != "ROCKALL",]
```

Use the key to pull in all the csvs
```{r pull in CSVs of catch data}
SAU_names <- SAU_survey_key[, survey_unit]
SAU_files <- SAU_survey_key[, FileName]

SAU_full <- data.table()

for (i in 1:length(SAU_names)) {
  SAU_data <- fread(here::here(SAU_files[i])) #pull in data for region[i]
  SAU_data.r <- SAU_data[,.(area_name, area_type, year, scientific_name, common_name, functional_group, commercial_group, fishing_entity, fishing_sector, catch_type, reporting_status, gear_type, end_use_type, tonnes, landed_value)]
  #link back to survey
  SAU_data.r[,survey_unit := SAU_names[i]]
  
  SAU_full <- rbind(SAU_full, SAU_data.r) #bind all regions together
  print(paste0(i,":",SAU_names[i],", Number of Columns: ",ncol(SAU_data)))
}



```
Cleaning SAU data

1. Limit to scientific names that are associated with fish by limiting with functional_group

```{r cleaning SAU data fish only}
#what functional groups are in this dataset
levels(as.factor(SAU_full[,functional_group]))

#get rid of 
function_groups_exclude <- c("Cephalopods","Krill","Lobsters, crabs","Shrimps","Jellyfish","Other demersal invertebrates")

SAU_full.r <- SAU_full[!(functional_group %in% function_groups_exclude)]

#check if it worked
levels(as.factor(SAU_full.r[,common_name])) #looks good

```

2. Limit to landings (exclude discards)

```{r cleaning SAU data no discards}
#what catch types
levels(as.factor(SAU_full.r[,catch_type]))

#
SAU_full.landings <- SAU_full.r[catch_type == "Landings"]

#check if it worked
levels(as.factor(SAU_full.r[,common_name])) #looks good

```

3. Limit to identified fish

```{r cleaning SAU data fish have to be IDed}
#what functional groups are in this dataset
levels(as.factor(SAU_full.landings[,scientific_name]))

#delete any observations not ID'd
SAU_full.landings.ided <- SAU_full.landings[scientific_name != "Marine fishes not identified"]


```

Now, we will sum tonnes across each year for each region
Moving forward, will use year *before* as the predictor
```{r sum tonnes}
SAU_summed_tonnes <- SAU_full.landings.ided[,summed_tonnes := sum(tonnes, na.rm = T),.(year, area_name, area_type, survey_unit)][ #key to include survey_unit because a few SAU data sets are in here multiple times
    , summed_tonnes_scaled_byreg := scale(summed_tonnes),.(survey_unit)][#scale by region (ie high relative to given region), this is what I plan to use
      , summed_tonnes_scaled := scale(summed_tonnes)] #scale across all values (ie high relative to all regions)

#reduce to unique values we need
SAU_summed_tonnes.r <- unique(SAU_summed_tonnes[,.(year, survey_unit, summed_tonnes, summed_tonnes_scaled, summed_tonnes_scaled_byreg)])


#plot to visualize
ggplot(SAU_summed_tonnes.r) +
  geom_line(aes(x = year, y = summed_tonnes_scaled_byreg)) +
  facet_wrap(~survey_unit) +
  theme_classic()

ggplot(SAU_summed_tonnes.r) +
  geom_line(aes(x = year, y = summed_tonnes_scaled_byreg, color = survey_unit)) +
  theme_classic()

#save
saveRDS(SAU_summed_tonnes.r, here::here("data","sea_around_us","SAU_summed_tonnes.rds"))
```

