---
title: "Data cleaning visualization"
author: Zoë J. Kitchel
date: "5/1/2023"
output: 
  bookdown::pdf_document2: default
header-includes: 
  - \renewcommand{\figurename}{Supplementary Figure }
  - \renewcommand{\tablename}{Supplementary Table}
  - \makeatletter
  - \def\fnum@figure{\figurename\thefigure}
  - \makeatother
editor_options: 
  markdown: 
    wrap: 72
---

This code is a supplemental script for Kitchel et al. 2023 submitted manuscript. 

Run after standardize_temp_space_2 for visual of years and cells that are excluded using 70% year 70% cell code.

Ideally, this outputs to PDF, but not currently working. Will fix later (and try on 2018 macair)


```{r setup, include=F}
library(data.table)
knitr::opts_chunk$set(echo = T)

```

Load plots
```{r}
#years
fishglob_exclude_years_plots <- readRDS(here::here("figures","standardization","fishglob_exclude_years_plots.Rds"))

#cells
fishglob_exclude_cells_plots <- readRDS(here::here("figures","standardization","fishglob_exclude_cells_plots.Rds"))
```


Plot years excluded
```{r fig-plot-years, fig.keep="all", fig.height=4, fig.cap=paste0("Number of tows per year used to standardize spatial footprint through time. Years highlighted in turquoise exluded because they sampled less than 70% of cells."), message = F, echo = F, warning = F}
for (i in c(1:44)) {
  cat('\n\n') 
print(fishglob_exclude_years_plots[[i]])
  cat('\n\n') 
}

```

Plot cells excluded

-Check EBS, GOA later which aren't printing for some reason (17 and 21)
```{r fig-plot-cells, fig.keep="all", fig.height=4, fig.cap=paste0("Tow locations and hexagonal grid cells used to standardize spatial footprint through time. Cells highlighted in turquoise exluded because they were sampled in less than 70% of years after accounting for undersampled years (See previous bar plots)."), message = F, echo = F, warning = F}
for (i in c(1:16,18:20,22:44)){
  cat('\n\n')
  print(fishglob_exclude_cells_plots[[i]])
  cat('\n\n')
}

```

