---
title: "Exploring North Sea data trimming"
output: 
    github_document:
        toc: true
        toc_depth: 3
---


```{r setup}
library(dggridR)
library(data.table)
library(rgdal)
library(raster)
library(sp)
library(here)
library(sf)
library(ggplot2) # for plotting
```


# Load data
Pull in compiled and cleaned data from fishglob February 4, 2022  
Trim to North Sea and divide into quarters 1 and 3
```{r pull in data}
FishGlob <- fread(file = here::here("temp", "FISHGLOB_v1.1_clean.csv"))
FishGlob <- FishGlob[survey == "NS-IBTS",]
FishGlob[,survey_season := as.factor(paste0(survey,"_",quarter))] # quarter indicates season
FishGlob[, summary(survey_season)]

#if positive, subtract 360
FishGlob[,longitude_s := ifelse(longitude > 150,(longitude-360),(longitude))]

#delete if NA for longitude or latitude
FishGlob <- FishGlob[complete.cases(FishGlob[,.(longitude, latitude)])]
```

# Standardize trawl footprint
## Hex size 8
```{r standardize footprint}
#if positive, subtract 360
FishGlob[,longitude_s := ifelse(longitude > 150,(longitude-360),(longitude))]

#delete if NA for longitude or latitude
FishGlob <- FishGlob[complete.cases(FishGlob[,.(longitude, latitude)])]

#set up grid
dggs <- dgconstruct(res = 8, metric = T) #with res = 8, we will need at least 3 observations per year within 7,774.2 km^2 (roughly size of some NEUS strata)

#pull out unique lat lons and get grid cells
unique_latlon <- unique(FishGlob[,.(latitude, longitude_s)])
unique_latlon[,cell := dgGEO_to_SEQNUM(dggs, longitude_s, latitude)] #get corresponding grid cells for this region/survey combo

#find cell centers
cellcenters <- dgSEQNUM_to_GEO(dggs, unique_latlon[,cell])

#linking cell centers to unique_latlon
unique_latlon[,cell_center_longitude_s := cellcenters$lon_deg][,cell_center_latitude:= cellcenters$lat_deg]

#link centers back to main data table
FishGlob.dg <- merge(FishGlob, unique_latlon, by = c("latitude", "longitude_s"), all.x = TRUE)
```

### Trim out years with <70% cells in a survey
yearkeep70 column == FALSE for years to trim out
```{r trim years}
# Find max num cells in each survey, calc 70% treshold
year_cells <- FishGlob.dg[,.(cell_count = length(unique(cell))), by = .(year, survey_season)]
year_cells[,thresh := 0.7*(max(cell_count)), by = survey_season] # of cells/ year to cut off below
  
#only keep years where over 70% of cells are sampled
year_cells[,yearkeep70 := cell_count > thresh]
  
# mark years to keep in FishGlob (yearkeep70 colum)
FishGlob.yr <- merge(FishGlob.dg, year_cells[, .(year, survey_season, yearkeep70)], all.x = TRUE, by = c('year', 'survey_season'))
```

#### Plot num. hauls by cell latitude and year. 
Second plot shows whether year is kept or not.
```{r}  
# table of years by cells, count is # hauls
yearcell_table <- FishGlob.yr[, .(nhaul = length(unique(haul_id))), by = .(cell, year, cell_center_latitude, cell_center_longitude_s, yearkeep70, survey_season)]

# color by #hauls
ggplot(yearcell_table, aes(year, cell_center_latitude, color = nhaul)) +
  geom_point() +
  facet_grid(~survey_season)


# color by keep or not at 70% threshold
ggplot(yearcell_table, aes(year, cell_center_latitude, color = yearkeep70)) +
  geom_point() +
  facet_grid(~survey_season)
```
  
### Plot num years by cell
Only count years from those remaining after the previous trim
Second plot shows whether a cell is sampled in all years
```{r}  
# table of years by cells, count is # hauls
cell_table <- FishGlob.yr[yearkeep70 == TRUE, .(nyear = length(unique(year))), by = .(cell, cell_center_latitude, cell_center_longitude_s, survey_season)]
cell_table[, allyrs := nyear == max(nyear), by = survey_season]

ggplot(cell_table, aes(cell_center_longitude_s, cell_center_latitude, color = nyear)) +
  geom_point() +
  facet_grid(~survey_season)

ggplot(cell_table, aes(cell_center_longitude_s, cell_center_latitude, color = allyrs)) +
  geom_point() +
  facet_grid(~survey_season)
  
```

### Mark cells sampled >=3x each remaining year
```{r}  
#identify num hauls per cells x year
FishGlob.yr[,year_cell_count := length(unique(haul_id)),.(year,cell, survey_season)]

#num years in which cells are sampled at least X times in any remaining year
# use X=1 as a test: should match plot of cells sampled in all years
nyrs_by_cell1dt <- FishGlob.yr[year_cell_count >= 1 & yearkeep70, .(nyrs_by_cell1 = length(unique(year))), by = .(cell, survey_season)] # number of years in which this cell has >= 1 haul
nyrs_by_cell3dt <- FishGlob.yr[year_cell_count >= 3 & yearkeep70, .(nyrs_by_cell3 = length(unique(year))), by = .(cell, survey_season)] # number of years in which this cell has >= 1 haul

FishGlob.cell <- merge(FishGlob.yr, nyrs_by_cell1dt, all.x = TRUE, by = c('survey_season', 'cell'))
FishGlob.cell[is.na(nyrs_by_cell1)==TRUE, nyrs_by_cell1 := 0] # fill missing values
FishGlob.cell <- merge(FishGlob.cell, nyrs_by_cell3dt, all.x = TRUE, by = c('survey_season', 'cell'))
FishGlob.cell[is.na(nyrs_by_cell3), nyrs_by_cell3 := 0] # fill missing values

# mark cells sampled the max years
FishGlob.cell[, cellkeep1 := (nyrs_by_cell1 == length(unique(year))) & yearkeep70, by = .(survey_season, yearkeep70)] # TRUE if all remaining years met the min num hauls threshold (1 in this case)
FishGlob.cell[, cellkeep3 := (nyrs_by_cell3 == length(unique(year))) & yearkeep70, by = .(survey_season, yearkeep70)]

#make a map of these points
ggplot(FishGlob.cell[yearkeep70 == TRUE][!duplicated(cbind(cell, survey_season)), ], 
       aes(cell_center_longitude_s, cell_center_latitude, color = cellkeep3)) +
  geom_point() +
  facet_grid(~survey_season)
```


## Hex size 7
```{r standardize footprint 7}
#set up grid
dggs7 <- dgconstruct(res = 7, metric = T) # larger than res 8

#pull out unique lat lons and get grid cells
unique_latlon7 <- unique(FishGlob[,.(latitude, longitude_s)])
unique_latlon7[,cell := dgGEO_to_SEQNUM(dggs7, longitude_s, latitude)] #get corresponding grid cells for this region/survey combo

#find cell centers
cellcenters7 <- dgSEQNUM_to_GEO(dggs7, unique_latlon7[,cell])

#linking cell centers to unique_latlon
unique_latlon7[,cell_center_longitude_s := cellcenters7$lon_deg][,cell_center_latitude:= cellcenters7$lat_deg]

#link centers back to main data table
FishGlob.dg7 <- merge(FishGlob, unique_latlon7, by = c("latitude", "longitude_s"), all.x = TRUE)
```

### Trim out years with <70% cells in a survey
yearkeep70 column == FALSE for years to trim out
```{r trim years 7}
# Find max num cells in each survey, calc 70% treshold
year_cells7 <- FishGlob.dg7[,.(cell_count = length(unique(cell))), by = .(year, survey_season)]
year_cells7[,thresh := 0.7*(max(cell_count)), by = survey_season] # of cells/ year to cut off below
  
#only keep years where over 70% of cells are sampled
year_cells7[,yearkeep70 := cell_count > thresh]
  
# mark years to keep in FishGlob (yearkeep70 colum)
FishGlob.yr7 <- merge(FishGlob.dg7, year_cells7[, .(year, survey_season, yearkeep70)], all.x = TRUE, by = c('year', 'survey_season'))
```

#### Plot num. hauls by cell latitude and year. 
Second plot shows whether year is kept or not.
```{r}  
# table of years by cells, count is # hauls
yearcell_table7 <- FishGlob.yr7[, .(nhaul = length(unique(haul_id))), by = .(cell, year, cell_center_latitude, cell_center_longitude_s, yearkeep70, survey_season)]

# color by #hauls
ggplot(yearcell_table7, aes(year, cell_center_latitude, color = nhaul)) +
  geom_point() +
  facet_grid(~survey_season)


# color by keep or not at 70% threshold
ggplot(yearcell_table7, aes(year, cell_center_latitude, color = yearkeep70)) +
  geom_point() +
  facet_grid(~survey_season)
```
  
### Plot num years by cell
Only count years from those remaining after the previous trim
Second plot shows whether a cell is sampled in all years
```{r}  
# table of years by cells, count is # hauls
cell_table7 <- FishGlob.yr7[yearkeep70 == TRUE, .(nyear = length(unique(year))), by = .(cell, cell_center_latitude, cell_center_longitude_s, survey_season)]
cell_table7[, allyrs := nyear == max(nyear), by = survey_season]

ggplot(cell_table7, aes(cell_center_longitude_s, cell_center_latitude, color = nyear)) +
  geom_point() +
  facet_grid(~survey_season)

ggplot(cell_table7, aes(cell_center_longitude_s, cell_center_latitude, color = allyrs)) +
  geom_point() +
  facet_grid(~survey_season)
  
```

### Mark cells sampled >=3x each remaining year
```{r}  
#identify num hauls per cells x year
FishGlob.yr7[,year_cell_count := length(unique(haul_id)),.(year,cell, survey_season)]

#num years in which cells are sampled at least X times in any remaining year
# use X=1 as a test: should match plot of cells sampled in all years
nyrs_by_cell1dt7 <- FishGlob.yr7[year_cell_count >= 1 & yearkeep70, .(nyrs_by_cell1 = length(unique(year))), by = .(cell, survey_season)] # number of years in which this cell has >= 1 haul
nyrs_by_cell3dt7 <- FishGlob.yr7[year_cell_count >= 3 & yearkeep70, .(nyrs_by_cell3 = length(unique(year))), by = .(cell, survey_season)] # number of years in which this cell has >= 1 haul

FishGlob.cell7 <- merge(FishGlob.yr7, nyrs_by_cell1dt7, all.x = TRUE, by = c('survey_season', 'cell'))
FishGlob.cell7[is.na(nyrs_by_cell1)==TRUE, nyrs_by_cell1 := 0] # fill missing values
FishGlob.cell7 <- merge(FishGlob.cell7, nyrs_by_cell3dt7, all.x = TRUE, by = c('survey_season', 'cell'))
FishGlob.cell7[is.na(nyrs_by_cell3), nyrs_by_cell3 := 0] # fill missing values

# mark cells sampled the max years
FishGlob.cell7[, cellkeep1 := (nyrs_by_cell1 == length(unique(year))) & yearkeep70, by = .(survey_season, yearkeep70)] # TRUE if all remaining years met the min num hauls threshold (1 in this case)
FishGlob.cell7[, cellkeep3 := (nyrs_by_cell3 == length(unique(year))) & yearkeep70, by = .(survey_season, yearkeep70)]

#make a map of these points
ggplot(FishGlob.cell7[yearkeep70 == TRUE][!duplicated(cbind(cell, survey_season)), ], 
       aes(cell_center_longitude_s, cell_center_latitude, color = cellkeep3)) +
  geom_point() +
  facet_grid(~survey_season)
```