---
title: "Regional Statistics for Jaccard Turnover Dissimilarity Metrics"
output: html_notebook
---

Here, we're looking at Jaccard Turnover Dissimilarity

This code is to look at regional statistics, and figures and stats will end up in supplement.

```{r setup}
library(data.table)
library(ggplot2)
library(sf)
library(rgeos)
library(concaveman)
library(raster)
library(stringr)
library(rnaturalearth)
library(rmapshaper)


#raw data
FishGlob_clean <- readRDS(here::here("data", "cleaned", "FishGlob.wellsampledyearscells_complete.final.rds"))

#model coefficients from year_dissimilarity_models.Rmd
jaccard_turnover_model_coefs_reduced_length.unique <- readRDS("~/Documents/grad_school/Rutgers/Repositories/trawl_spatial_turnover_git/output/region_stats_jaccard_turnover/jaccard_turnover_model_coefs_reduced_length.unique.Rds")

#NB: years sampled = last_year-first_year+1

```

*Additional cleaning (first written in "dissim_metric_space_time.Rmd")*


Delete observations without CPUE except for MEDITS (we will do trends in abundance metrics for this region)
```{r delete no CPUE obs}
#delete all rows where wgt_cpue is NA,  unless it is MEDITS region
FishGlob_clean.noNA <- FishGlob_clean[(survey != "MEDITS" & !is.na(wgt_cpue)) | survey == "MEDITS"]

#check
summary(FishGlob_clean.noNA[, wgt_cpue]) #why inf? 13 observations from GSL-N,  I will delete these
FishGlob_clean.noNA <- copy(FishGlob_clean.noNA)[!is.infinite(wgt_cpue), ]

```


```{r list of all survey season combinations}
all_survey_units <- unique(FishGlob_clean.noNA[, survey_unit])
```

For NZ-CHAT, we want to lump December observations with next year of observations (January,  February)
```{r lump December observations with previous year}
FishGlob_clean.noNA[survey_unit == "NZ-CHAT" & month == 12,  year := year+1, ]
```

Some spp are duplicated within a haul. In all cases,  this is due to verbatim names matching to a single accepted name. Therefore,  we will sum these wgts and abundances
```{r sum duplicates}
FishGlob_clean.noNA[, wgt := sum(wgt,  na.rm = T), .(accepted_name,  haul_id)][, 
                     wgt_h := sum(wgt_h,  na.rm = T), .(accepted_name,  haul_id)][, 
                     wgt_cpue := sum(wgt_cpue,  na.rm = T), .(accepted_name,  haul_id)][, 
                     num := sum(num,  na.rm = T), .(accepted_name,  haul_id)][, 
                     num_h := sum(num_h,  na.rm = T), .(accepted_name,  haul_id)][, 
                     num_cpue := sum(num_cpue,  na.rm = T), .(accepted_name,  haul_id)]

FishGlob_clean.noNA <- unique(FishGlob_clean.noNA[,.(survey, haul_id, country, sub_area, continent, stat_rec, station, stratum, year, month, day, quarter, season, latitude, longitude, haul_dur, area_swept, gear, depth, sbt, sst, num, num_h, num_cpue, wgt, wgt_h, wgt_cpue, verbatim_aphia_id, accepted_name, aphia_id, SpecCode, kingdom, phylum, class, order, family, genus, rank, survey_unit, years_sampled, cell_ID, cell_year_count, year_cell_count, years_sampled_update)])

#make depth numeric
FishGlob_clean.noNA[,depth := as.numeric(depth)]

#add column for yearly summed biomass (note that MEDITS doesn't have biomass, so it will be excluded from these plots)
FishGlob_clean.noNA[,wgt_cpue_annual := sum(wgt_cpue,na.rm = T),.(year, survey_unit)]

#add column for yearly summed # of hauls
FishGlob_clean.noNA[,haul_id_count_annual := uniqueN(haul_id),.(year, survey_unit)]
```

Pull in avg temp values and pull in avg fishing values
```{r avg temp and fishing values}
 #open fishing and temp data
distances_dissimilarities_allyears.r.temp.fishing <- readRDS( here::here("output","distance_decay","distances_dissimilarities_allyears.r.temp.fishing.rds"))

distances_dissimilarities_allyears.r.temp.fishing[,mean_reg_temp_allyears := mean(sst_mean, na.rm = T),.(survey_unit)][,mean_reg_fishing_allyears := mean(summed_tonnes, na.rm = T),.(survey_unit)]


avg_temp_avg_fishing <- unique(distances_dissimilarities_allyears.r.temp.fishing[,.(survey_unit,mean_reg_fishing_allyears, mean_reg_temp_allyears)]) 

#link coefficients
year_survey_unit_characteristics <- jaccard_turnover_model_coefs_reduced_length.unique[avg_temp_avg_fishing, on = "survey_unit"]

```

For area, we want to delete any areas that overlap with land, therefore, we have to pull in land polygon
```{r}
#basemap
world <- rnaturalearth::ne_countries(type = "countries", scale="large", returnclass="sf")

```

Map Helper
```{r}
name_helper <- data.table(Survey_Name_Season = c("Aleutian Islands",
                                    "Baltic Sea Q1",
                                    "Baltic Sea Q4",
                                    "Chile",
                                    "Newfoundland",
                                    "Queen Charlotte Sound",
                                    "Eastern Bering Sea",
                                    "Bay of Biscay",
                                    "English Channel",
                                    "Gulf of Mexico",
                                    "Gulf of Alaska",
                                    "Greenland",
                                    "N Gulf of St. Lawrence",
                                    "S Gulf of St. Lawrence",
                                    "Iceland",
                                    "Irish Sea",
                                    "Mediterranean",
                                    "Namibia",
                                    "NE US Fall",
                                    "NE US Spring",
                                    "N Ireland Q1",
                                    "N Ireland Q4",
                                    "Norway",
                                    "N Sea Q1",
                                    "N Sea Q3",
                                    "Chatham Rise",
                                    "E Coast S Island NZ",
                                    "W Coast S Island NZ",
                                    "Portugal",
                                    "S Georgia Straight",
                                  "Scotian Shelf",
                                  "SE US Fall",
                                  "SE US Spring",
                                  "SE US Summer",
                                  "W Coast US",
                                  "Atlantic Ocean ZA",
                                  "Indian Ocean ZA"),
                          survey_unit = c(
                                  "AI",        
                                  "BITS-1",    
                                  "BITS-4",    
                                  "CHL",       
                                  "DFO-NF",    
                                  "DFO-QCS",   
                                  "EBS",       
                                  "EVHOE",     
                                  "FR-CGFS",   
                                  "GMEX-Summer",
                                  "GOA",       
                                  "GRL-DE",    
                                  "GSL-N",     
                                  "GSL-S",     
                                  "ICE-GFS",   
                                  "IE-IGFS",   
                                  "MEDITS",    
                                  "NAM",       
                                  "NEUS-Fall", 
                                  "NEUS-Spring",
                                  "NIGFS-1",   
                                  "NIGFS-4",   
                                  "Nor-BTS-3", 
                                  "NS-IBTS-1", 
                                  "NS-IBTS-3", 
                                  "NZ-CHAT",   
                                  "NZ-ECSI",   
                                  "NZ-WCSI",   
                                  "PT-IBTS",   
                                  "S-GEORG",   
                                  "SCS-SUMMER",
                                  "SEUS-fall", 
                                  "SEUS-spring",
                                  "SEUS-summer",
                                  "WCANN",     
                                  "ZAF-ATL",   
                                  "ZAF-IND"   
                          ))
```


Calculate stats of all regions
```{r calculate all region stats}

#survey_unit
region_stats_jaccard_turnover <- as.data.table(matrix(nrow = length(all_survey_units)))

region_stats_jaccard_turnover[,survey_unit := as.character(V1)][, survey := as.character(V1)][, spp_num := as.numeric(V1)][,mean_annual_biomass := as.numeric()][,mean_haul_id_count_annual := as.numeric()][, study_period := as.numeric(V1)][, study_duration := as.numeric(V1)][, lat_range := as.numeric(V1)][, mid_lat := as.numeric(V1)][, lon_range := as.numeric(V1)][, area_km := as.numeric(V1)][, depth_range := as.numeric(V1)][,  mid_depth := as.numeric(V1)][, mean_temp := as.numeric(V1)][, mean_landings := as.numeric(V1)][,jaccard_coef := as.numeric()][,jaccard_coef_upr := as.numeric()][,jaccard_coef_lwr := as.numeric()][,direction_of_change := as.character()][,significant := as.logical()]

  region_stats_jaccard_turnover[,V1:=NULL]

for (i in 1:length(all_survey_units)) {
  reduced_FishGlob_cleaned.10year <- FishGlob_clean.noNA[survey_unit == all_survey_units[i]]
  
    ####unique lat lon
    lat_lon <- unique(reduced_FishGlob_cleaned.10year[,.(latitude, longitude)])
    
    pts <- st_as_sf(lat_lon, coords=c('longitude','latitude'), crs=4326 )
    
    conc <- concaveman(pts, 1,2)
    
    #delete areas that overlap
    conc_nooverlap <- ms_erase(conc,world[1])
    
    sf_use_s2(FALSE) #helps with spherical geometry
    area <- st_area(conc_nooverlap) #m2

    #compile into region_stats_jaccard_turnover table    
                  region_stats_jaccard_turnover[i,"survey_unit"] <- paste0(all_survey_units[i])
                  region_stats_jaccard_turnover[i,"survey"] <- word(all_survey_units[i],1, sep = "_")
                  region_stats_jaccard_turnover[i,"spp_num"] <- length(unique(reduced_FishGlob_cleaned.10year[,accepted_name]))
                  region_stats_jaccard_turnover[i,"mean_annual_biomass"] <- mean(reduced_FishGlob_cleaned.10year[,wgt_cpue_annual])
                  region_stats_jaccard_turnover[i,"mean_haul_id_count_annual"] <- mean(reduced_FishGlob_cleaned.10year[,haul_id_count_annual])
                  region_stats_jaccard_turnover[i,"study_period"] <- max(reduced_FishGlob_cleaned.10year$year)-min(reduced_FishGlob_cleaned.10year$year)
                  region_stats_jaccard_turnover[i,"study_duration"] <- length(unique(reduced_FishGlob_cleaned.10year$year))
                  region_stats_jaccard_turnover[i,"lat_range"] <- max(reduced_FishGlob_cleaned.10year$latitude)-min(reduced_FishGlob_cleaned.10year$latitude)
                  region_stats_jaccard_turnover[i,"mid_lat"] <- mean(reduced_FishGlob_cleaned.10year$latitude)
                  region_stats_jaccard_turnover[i,"lon_range"] <- max(reduced_FishGlob_cleaned.10year$longitude)-min(reduced_FishGlob_cleaned.10year$longitude)
                  region_stats_jaccard_turnover[i,"area_km"] <- area/(1000^2)
                  region_stats_jaccard_turnover[i,"depth_range"] <- max(reduced_FishGlob_cleaned.10year$depth, na.rm = T)-min(reduced_FishGlob_cleaned.10year$depth, na.rm = T)
                  region_stats_jaccard_turnover[i,"mid_depth"] <- mean(reduced_FishGlob_cleaned.10year$depth, na.rm = T)
                  region_stats_jaccard_turnover[i,"mean_temp"] <- year_survey_unit_characteristics[survey_unit == all_survey_units[i],mean_reg_temp_allyears]
                  region_stats_jaccard_turnover[i,"mean_landings"] <- year_survey_unit_characteristics[survey_unit == all_survey_units[i],mean_reg_fishing_allyears]
                  region_stats_jaccard_turnover[i,"jaccard_coef"] <- year_survey_unit_characteristics[survey_unit == all_survey_units[i],year_adj]
                  region_stats_jaccard_turnover[i,"jaccard_coef_lwr"] <- year_survey_unit_characteristics[survey_unit == all_survey_units[i],lwr]
                  region_stats_jaccard_turnover[i,"jaccard_coef_upr"] <- year_survey_unit_characteristics[survey_unit == all_survey_units[i],upr]
                  region_stats_jaccard_turnover[i,"Directional_Change"] <- year_survey_unit_characteristics[survey_unit == all_survey_units[i],Directional_Change]
                  region_stats_jaccard_turnover[i,"significant"] <- year_survey_unit_characteristics[survey_unit == all_survey_units[i],significant]
}
  
#if any values are -Inf, change to NA
invisible(lapply(names(region_stats_jaccard_turnover),function(.name) set(region_stats_jaccard_turnover, which(is.infinite(region_stats_jaccard_turnover[[.name]])), j = .name,value =NA)))

#add significant directions column
region_stats_jaccard_turnover[,"Significant Trends" := ifelse(significant==T, Directional_Change, "No Directional Change")]

#add column for avg # of hauls per area
region_stats_jaccard_turnover[,avg_hauls_area := mean_haul_id_count_annual/area_km]


saveRDS(region_stats_jaccard_turnover, here::here("output","region_stats_jaccard_turnover","region_stats_jaccard_turnover.rds"))

#test plot
plot(conc)
plot(conc_nooverlap)
```

Check some linear models
```{r linear models}
colnames(region_stats_jaccard_turnover)


```

Regions that are homogenizing tend to have lower overall gamma diversity than regions that are differentiating
```{r spp num}

spp_num_vs_jaccard_turnover_coef <- ggplot() +
    geom_errorbar(data = region_stats_jaccard_turnover, aes(x = spp_num, y = jaccard_coef, ymin = jaccard_coef_lwr, ymax = jaccard_coef_upr, color = `Significant Trends`), alpha = 0.8, linewidth = 0.3) + #add confidence intervals
        geom_point(data = region_stats_jaccard_turnover, aes(x = spp_num , y = jaccard_coef, color = `Significant Trends`), stat = 'identity', size = 1) +
    scale_color_manual(values = c("#91c874", "#e7ac5b", "#cbbfde"), name = "Directional Change") +
  geom_hline(yintercept = 0, size = 0.5) +
  geom_smooth(data = region_stats_jaccard_turnover, aes(x = spp_num, y = jaccard_coef), method = "lm", color = "black", se = F) +
  ylab("Dissimilarity coefficient") +
  xlab("Gamma diversity") +
  theme_classic()

spp_num_vs_jaccard_turnover_coef

ggsave(spp_num_vs_jaccard_turnover_coef, path = here::here("figures","jaccard_coef_vs_survey_characteristics"), filename = "spp_num_vs_jaccard_turnover_coef.jpg", height =3, width = 5, unit = "in")

#linear model
summary(lm(abs(jaccard_coef)~spp_num, data = region_stats_jaccard_turnover)) #not significant
summary(lm(jaccard_coef~spp_num, data = region_stats_jaccard_turnover)) #significant

#low# of species, more likely to experience differentiation, high # of species, more likely to experience homogenization

saveRDS(spp_num_vs_coef, here::here("figures","jaccard_coef_vs_survey_characteristics","spp_num_vs_coef.rds"))

```

More likely to detect both homogenization and differentation for longer study periods
```{r study period}

study_period_vs_coef <- ggplot() +
    geom_errorbar(data = region_stats_jaccard_turnover, aes(x = study_period, y = jaccard_coef, ymin = jaccard_coef_lwr, ymax = jaccard_coef_upr, color = `Significant Trends`), alpha = 0.8, linewidth = 0.3) + #add confidence intervals
        geom_point(data = region_stats_jaccard_turnover, aes(x = study_period , y = jaccard_coef, color = `Significant Trends`), stat = 'identity', size = 1) +
    scale_color_manual(values = c("#91c874", "#e7ac5b", "#cbbfde"), name = "Directional Change") +
  geom_hline(yintercept = 0, size = 0.5) +
  ylab("Dissimilarity coefficient") +
  xlab("Survey period") + #first minus last
  theme_classic()

study_period_vs_coef

ggsave(study_period_vs_coef, path = here::here("figures","jaccard_coef_vs_survey_characteristics"), filename = "study_period_vs_coef.jpg", height =3, width = 5, unit = "in")

#linear model
summary(lm(abs(jaccard_coef)~study_period, data = region_stats_jaccard_turnover)) #not significant
summary(lm(jaccard_coef~study_period, data = region_stats_jaccard_turnover)) #not significant
```

More likely to detect differentiation when fewer years are sampled and more likely to detect homogenization when more years are sampled
```{r study duration}

study_duration_vs_coef <- ggplot() +
    geom_errorbar(data = region_stats_jaccard_turnover, aes(x = study_duration, y = jaccard_coef, ymin = jaccard_coef_lwr, ymax = jaccard_coef_upr, color = `Significant Trends`), alpha = 0.8, linewidth = 0.3) + #add confidence intervals
        geom_point(data = region_stats_jaccard_turnover, aes(x = study_duration , y = jaccard_coef, color = `Significant Trends`), stat = 'identity', size = 1) +
    scale_color_manual(values = c("#91c874", "#e7ac5b", "#cbbfde"), name = "Directional Change") +
  geom_hline(yintercept = 0, size = 0.5) +
  ylab("Dissimilarity coefficient") +
  xlab("Total years sampled") + #count of years sampled
  theme_classic()

study_duration_vs_coef

ggsave(study_duration_vs_coef, path = here::here("figures","jaccard_coef_vs_survey_characteristics"), filename = "study_duration_vs_coef.jpg", height =3, width = 5, unit = "in")

#linear model
summary(lm(abs(jaccard_coef)~study_duration, data = region_stats_jaccard_turnover)) #not significant
summary(lm(jaccard_coef~study_duration, data = region_stats_jaccard_turnover)) #not significant


```

More likely to detect differentiation when a small latitudinal range is sampled, more likely to detect homogenization when a larger latitudinal range is sampled
```{r latitude range}


lat_range_vs_coef <- ggplot() +
    geom_errorbar(data = region_stats_jaccard_turnover, aes(x = lat_range, y = jaccard_coef, ymin = jaccard_coef_lwr, ymax = jaccard_coef_upr, color = `Significant Trends`), alpha = 0.8, linewidth = 0.3) + #add confidence intervals
        geom_point(data = region_stats_jaccard_turnover, aes(x = lat_range , y = jaccard_coef, color = `Significant Trends`), stat = 'identity', size = 1) +
    scale_color_manual(values = c("#91c874", "#e7ac5b", "#cbbfde"), name = "Directional Change") +
  geom_hline(yintercept = 0, size = 0.5) +
  ylab("Dissimilarity coefficient") +
  xlab("Latitudinal range (\u00B0s)") + #count of years sampled
  theme_classic()

lat_range_vs_coef

ggsave(lat_range_vs_coef, path = here::here("figures","jaccard_coef_vs_survey_characteristics"), filename = "lat_range_vs_coef.jpg", height =3, width = 5, unit = "in")

#linear model
summary(lm(abs(jaccard_coef)~lat_range, data = region_stats_jaccard_turnover)) #not significant
summary(lm(jaccard_coef~lat_range, data = region_stats_jaccard_turnover)) #nearly but not significant
```

Differentiation occurs below 30 degrees latitude, homogenization and no change both common at higher latitudes in the northern hemisphere, NO homogenization below 30 degrees latitude
```{r middle latitude}


mid_lat_vs_coef <- ggplot() +
    geom_errorbar(data = region_stats_jaccard_turnover, aes(x = mid_lat, y = jaccard_coef, ymin = jaccard_coef_lwr, ymax = jaccard_coef_upr, color = `Significant Trends`), alpha = 0.8, linewidth = 0.3) + #add confidence intervals
        geom_point(data = region_stats_jaccard_turnover, aes(x = mid_lat , y = jaccard_coef, color = `Significant Trends`), stat = 'identity', size = 1) +
    scale_color_manual(values = c("#91c874", "#e7ac5b", "#cbbfde"), name = "Directional Change") +
  geom_hline(yintercept = 0, size = 0.5) +
  geom_smooth(data = region_stats_jaccard_turnover, aes(x = mid_lat, y = jaccard_coef), method = "lm", color = "black", se = F) +
  ylab("Dissimilarity coefficient") +
  xlab(paste0("Latitude (\u00B0N)")) + #middle latitude
  theme_classic()

mid_lat_vs_coef

saveRDS(mid_lat_vs_coef, here::here("figures","jaccard_coef_vs_survey_characteristics","mid_lat_vs_coef.rds"))

ggsave(mid_lat_vs_coef, path = here::here("figures","jaccard_coef_vs_survey_characteristics"), filename = "mid_lat_vs_coef.jpg", height =3, width = 5, unit = "in")

#linear model
summary(lm(abs(jaccard_coef)~mid_lat, data = region_stats_jaccard_turnover)) #not significant
summary(lm(jaccard_coef~mid_lat, data = region_stats_jaccard_turnover)) #significant

      
```

#NEED TO FIX FOR REGIONS THAT CROSS DATE LINE
```{r longitudinal range}

ggplot(region_stats_jaccard_turnover) +
  geom_boxplot(aes(x = `Significant Trends`, y = lon_range, group = `Significant Trends`)) +
  theme_classic()

ggplot(region_stats_jaccard_turnover) +
  geom_point(aes(x = lon_range, y = jaccard_coef)) +
  theme_classic()
```

Areas that are differentiating tend to be smaller than areas that are homogenizing
```{r area}
area_km_vs_coef <- ggplot() +
    geom_errorbar(data = region_stats_jaccard_turnover, aes(x = area_km/1000, y = jaccard_coef, ymin = jaccard_coef_lwr, ymax = jaccard_coef_upr, color = `Significant Trends`), alpha = 0.8, linewidth = 0.3) + #add confidence intervals
        geom_point(data = region_stats_jaccard_turnover, aes(x = (area_km/1000) , y = jaccard_coef, color = `Significant Trends`), stat = 'identity', size = 1) +
    scale_color_manual(values = c("#91c874", "#e7ac5b", "#cbbfde"), name = "Directional Change") +
  geom_hline(yintercept = 0, size = 0.5) +
  ylab("Dissimilarity coefficient") +
  xlab("Area (1000s of squared km)") + #area_km
  theme_classic()

area_km_vs_coef

ggsave(area_km_vs_coef, path = here::here("figures","jaccard_coef_vs_survey_characteristics"), filename = "area_km_vs_coef.jpg", height =3, width = 5, unit = "in")

#linear model
summary(lm(abs(jaccard_coef)~area_km, data = region_stats_jaccard_turnover)) #not significant
summary(lm(jaccard_coef~area_km, data = region_stats_jaccard_turnover)) #not significant

```

```{r number of hauls vs area}

avg_hauls_area_vs_coef <- ggplot() +
    geom_errorbar(data = region_stats_jaccard_turnover, aes(x = avg_hauls_area, y = jaccard_coef, ymin = jaccard_coef_lwr, ymax = jaccard_coef_upr, color = `Significant Trends`), alpha = 0.8, linewidth = 0.3) + #add confidence intervals
        geom_point(data = region_stats_jaccard_turnover, aes(x = (avg_hauls_area) , y = jaccard_coef, color = `Significant Trends`), stat = 'identity', size = 1) +
    scale_color_manual(values = c("#91c874", "#e7ac5b", "#cbbfde"), name = "Directional Change") +
  geom_hline(yintercept = 0, size = 0.5) +
  ylab("Dissimilarity coefficient") +
  xlab("Hauls per area (squared km)") + #avg_hauls_area
  theme_classic()

avg_hauls_area_vs_coef

ggsave(avg_hauls_area_vs_coef, path = here::here("figures","jaccard_coef_vs_survey_characteristics"), filename = "avg_hauls_area_vs_coef.jpg", height =3, width = 5, unit = "in")

#linear model
summary(lm(abs(jaccard_coef)~avg_hauls_area, data = region_stats_jaccard_turnover)) #not significant
summary(lm(jaccard_coef~avg_hauls_area, data = region_stats_jaccard_turnover)) #not significant

```


```{r depth_range}

depth_range_vs_coef <- ggplot() +
    geom_errorbar(data = region_stats_jaccard_turnover, aes(x = depth_range, y = jaccard_coef, ymin = jaccard_coef_lwr, ymax = jaccard_coef_upr, color = `Significant Trends`), alpha = 0.8, linewidth = 0.3) + #add confidence intervals
        geom_point(data = region_stats_jaccard_turnover, aes(x = (depth_range) , y = jaccard_coef, color = `Significant Trends`), stat = 'identity', size = 1) +
    scale_color_manual(values = c("#91c874", "#e7ac5b", "#cbbfde"), name = "Directional Change") +
  geom_hline(yintercept = 0, size = 0.5) +
  ylab("Dissimilarity coefficient") +
  xlab("Depth range (m)") + #depth_range
  theme_classic()

depth_range_vs_coef

ggsave(depth_range_vs_coef, path = here::here("figures","jaccard_coef_vs_survey_characteristics"), filename = "depth_range_vs_coef.jpg", height =3, width = 5, unit = "in")

#linear model
summary(lm(abs(jaccard_coef)~depth_range, data = region_stats_jaccard_turnover)) #not significant
summary(lm(jaccard_coef~depth_range, data = region_stats_jaccard_turnover)) #not significant
```

```{r mid_depth}

mid_depth_vs_coef <- ggplot() +
    geom_errorbar(data = region_stats_jaccard_turnover, aes(x = mid_depth, y = jaccard_coef, ymin = jaccard_coef_lwr, ymax = jaccard_coef_upr, color = `Significant Trends`), alpha = 0.8, linewidth = 0.3) + #add confidence intervals
        geom_point(data = region_stats_jaccard_turnover, aes(x = (mid_depth) , y = jaccard_coef, color = `Significant Trends`), stat = 'identity', size = 1) +
    scale_color_manual(values = c("#91c874", "#e7ac5b", "#cbbfde"), name = "Directional Change") +
  geom_hline(yintercept = 0, size = 0.5) +
  ylab("Dissimilarity coefficient") +
  xlab("Average depth (m)") + #mid_depth
  theme_classic()

mid_depth_vs_coef

ggsave(mid_depth_vs_coef, path = here::here("figures","jaccard_coef_vs_survey_characteristics"), filename = "mid_depth_vs_coef.jpg", height =3, width = 5, unit = "in")

#linear model
summary(lm(abs(jaccard_coef)~mid_depth, data = region_stats_jaccard_turnover)) #not significant
summary(lm(jaccard_coef~mid_depth, data = region_stats_jaccard_turnover)) #not significant
```

```{r mean_temp}
mean_temp_vs_coef <- ggplot() +
    geom_errorbar(data = region_stats_jaccard_turnover, aes(x = mean_temp, y = jaccard_coef, ymin = jaccard_coef_lwr, ymax = jaccard_coef_upr, color = `Significant Trends`), alpha = 0.8, linewidth = 0.3) + #add confidence intervals
        geom_point(data = region_stats_jaccard_turnover, aes(x = (mean_temp) , y = jaccard_coef, color = `Significant Trends`), stat = 'identity', size = 1) +
    scale_color_manual(values = c("#91c874", "#e7ac5b", "#cbbfde"), name = "Directional Change") +
  geom_hline(yintercept = 0, size = 0.5) +
  geom_smooth(data = region_stats_jaccard_turnover, aes(x = mean_temp, y = jaccard_coef), method = "lm", color = "black", se = F) +
  ylab("Dissimilarity coefficient") +
  xlab("Average temperature (\u00B0C)") + #mean_temp
  theme_classic()

mean_temp_vs_coef

ggsave(mean_temp_vs_coef, path = here::here("figures","jaccard_coef_vs_survey_characteristics"), filename = "mean_temp_vs_coef.jpg", height =3, width = 5, unit = "in")

#linear model
summary(lm(abs(jaccard_coef)~mean_temp, data = region_stats_jaccard_turnover)) #not significant
summary(lm(jaccard_coef~mean_temp, data = region_stats_jaccard_turnover)) #not significant

saveRDS(mean_temp_vs_coef, here::here("figures","jaccard_coef_vs_survey_characteristics","mean_temp_vs_coef.rds"))
```

```{r mean_landings}
mean_landings_vs_coef <- ggplot() +
    geom_errorbar(data = region_stats_jaccard_turnover, aes(x = (mean_landings/1000), y = jaccard_coef, ymin = jaccard_coef_lwr, ymax = jaccard_coef_upr, color = `Significant Trends`), alpha = 0.8, linewidth = 0.3) + #add confidence intervals
        geom_point(data = region_stats_jaccard_turnover, aes(x = (mean_landings/1000) , y = jaccard_coef, color = `Significant Trends`), stat = 'identity', size = 1) +
    scale_color_manual(values = c("#91c874", "#e7ac5b", "#cbbfde"), name = "Directional Change") +
  geom_hline(yintercept = 0, size = 0.5) +
  ylab("Dissimilarity coefficient") +
  xlab("Mean landings (1000s of tonnes)") + #average annual landings
  theme_classic()

mean_landings_vs_coef

ggsave(mean_landings_vs_coef, path = here::here("figures","jaccard_coef_vs_survey_characteristics"), filename = "mean_landings_vs_coef.jpg", height =3, width = 5, unit = "in")

#linear model
summary(lm(abs(jaccard_coef)~mean_landings, data = region_stats_jaccard_turnover)) #not significant
summary(lm(jaccard_coef~mean_landings, data = region_stats_jaccard_turnover)) #not significant
```

