---
title: "Prepping EBS Data"
output: html_notebook
---


```{r setup}
library(dggridR)
library(data.table)
library(rgdal)
library(raster)
library(sp)
library(rnaturalearth)
library(rnaturalearthdata)
library(rgeos)
library(taxize) #standardizing names
library(geosphere)  #to calculate distance between lat lon of grid cells
library(googledrive)
```


Moving forward, I plan to use data directly from FishGlob.

November 21, 2021
```{r pull in data from fishglob for Scotian Shelf from Google Drive}
drive_download("EBS_clean.csv",
               overwrite = T)

ebs <- fread("EBS_clean.csv")

file.remove("EBS_clean.csv")

```

In the Eastern Bering Sea, sampling years prior to 1984 (data begin in 1982) were excluded from analysis due to large apparent increases in the number of species recorded in the first two years. (Ryan et al. 2017)

```{r EBS}
ebs <- ebs[year >= 1984,]

```


```{r pull in EBS}

# Get Unique lines in the data table
unique_EBS_latlon<- unique(ebs[,.(latitude, longitude)])
unique_EBS_latlon.sf <- unique_EBS_latlon
  
#make lat lon coordinates into polygon
coordinates(unique_EBS_latlon.sf) <- ~latitude + longitude
unique_EBS_latlon.sf@proj4string <- crs("+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0")

saveRDS(unique_EBS_latlon, "unique_EBS_latlon.rds")

#plot north america
world <- ne_countries(scale = "medium", returnclass = "sp")

world_ext <- extent(-180, -51.5,23, 67)

north_america_spdf <- crop(world, world_ext)

north_america_df <- fortify(north_america_spdf)

(na_outline <- ggplot(north_america_df, aes(long, lat, group = group)) +
  geom_polygon(color = "white", fill = "grey") +
  labs(x = "Longitude", y = "Latitude") +
  coord_equal() +
  theme_classic() +
  scale_x_continuous(expand = c(0,0)) +
  scale_y_continuous(expand = c(0,0)) +
  theme(panel.border = element_rect(color = "black", fill = NA, size = 1)) +
 geom_point(unique_EBS_latlon, mapping = aes(longitude, latitude), inherit.aes = FALSE, shape = 18, size = 0.001, color = "red") +
  coord_equal())

```

Playing around with dggridR (package that makes grid cells)
```{r learning dggridR}
dggs <- dgconstruct(spacing = 111, metric = T, resround = 'nearest')

unique_EBS_latlon[,cell := dgGEO_to_SEQNUM(dggs, longitude, latitude)] #get corresponding grid cells for EBS trawl

#centersof cells
EBS_cellcenters <- dgSEQNUM_to_GEO(dggs, unique_EBS_latlon[,cell])

#linking cell centers to unique_EBS_latlon
unique_EBS_latlon[,LON_CENTER := EBS_cellcenters$lon_deg][,LAT_CENTER := EBS_cellcenters$lat_deg]

#link centers back to main data table

ebs <- merge(ebs, unique_EBS_latlon, by = c("latitude", "longitude"))

#number of tows in each cell
towcount <- unique_EBS_latlon[, .N, by = cell]

#get the grid cell boundary for cells which had trawls
grid_EBS <- dgcellstogrid(dggs, unique_EBS_latlon[,cell], frame = T, wrapcells = F)

#update grid properties to include # of trawls in each cell
grid_EBS <- merge(grid_EBS, unique_EBS_latlon, by = "cell")

ggplot(data = grid_EBS, aes(x = long, y = lat, group = cell)) +
  geom_polygon() +
  coord_quickmap()

saveRDS(grid_EBS, "grid_EBS.rds")

```

Trying to plot with grid overlaying
```{r grid overlaying}

na_outline + 
   geom_polygon(grid_EBS, mapping = aes(x = long,y = lat, group = cell), inherit.aes = FALSE, color = "black", fill = NA, size = 0.1)

ggsave(filename = "grid_EBS_cells_spacing_111.pdf")
  
```

Now, we'll overlay grid cells onto this list of tows (ebs) using lat and lon and gridrr package

```{r get rid of underused grid cells}

year_grid <- data.table(table(ebs$cell, ebs$year))

#there are some grid cells with infrequent records, I will delete these
grid_cells_to_exclude <- unique(year_grid[N == 0]$V1)

#delete these grid cells from full data table
ebs.reduced <- ebs[!cell %in% grid_cells_to_exclude]

#270643 observations left, this should be great

#how many tows (TOW) per cell in a given year

haul_cell_unique <- unique(ebs.reduced[,.(haul_id, cell, year)])

ebs_samplingevents_byyear <- haul_cell_unique[,.N, .(year, cell)]

ebs_cells_lessthan3 <- ebs_samplingevents_byyear[N<3,]

cells_to_delete <- unique(ebs_cells_lessthan3[,cell])

ebs.reduced.again <- ebs.reduced[!cell %in% cells_to_delete]

#down to 114793 observations

saveRDS(ebs.reduced.again, "ebs_data_gridded.rds")

```

Delete cells that aren't sampled regularly (missing more than 2 years from 1972-2019):

32234: sampled sporadically
all others have great coverage

Infrequently sampled cells
```{r infrequently sampled cells}

dump <- 32234


#delete these grid cells from full data table
grid_EBS.reduced.again.again <- ebs.reduced.again[cell != dump]


#114262 observations (no change)

```

This is where I would use rarefaction or coverage, for now I will skip
```{r rarefaction or coverage?}
#how many tows (TOW) per cell in a given year

cell_tow_year <- unique(grid_EBS.reduced.again.again[,.(haul_id, cell, year)])

ebs_samplingevents_byyear <- cell_tow_year[,.N, .(year, cell)]

ebs_cells_lessthan3 <- ebs_samplingevents_byyear[N<3,] #need three to build rarefaction curve, so, will get rid of less than three observations

cells_to_delete <- unique(ebs_cells_lessthan3[,cell])

ebs_full_1982onward.clean.reduced.again <- grid_EBS.reduced[!cell %in% cells_to_delete]

#1299247 observations

saveRDS(ebs_full_1982onward.clean.reduced.again, "ebs_data_gridded.rds")

```


Find distance between center of each grid cell
```{r distance between grid cells}

#just need grid cell # and center lat lon for each grid cell 
grid_EBS_cells <- unique(grid_EBS.reduced.again.again[,.(cell, LON_CENTER, LAT_CENTER)])

grid_EBS_cells <- setorder(grid_EBS_cells, cell)

ebs_reg_distances <- distm(grid_EBS_cells[,.(LON_CENTER, LAT_CENTER)])

colnames(ebs_reg_distances) <- grid_EBS_cells$cell
rownames(ebs_reg_distances) <- grid_EBS_cells$cell

#reorient to long form 
ebs_reg_distances.l <- reshape2::melt(as.matrix(ebs_reg_distances), varnames = c("cell1", "cell2"), value.name = "distance(m)") #matrix to data frame
ebs_reg_distances.l <- data.table(ebs_reg_distances.l) #and then to data table
ebs_reg_distances.l <- ebs_reg_distances.l[cell1 > cell2,] # get rid of repetitions and self-comparisons

```

Check the # of tows per grid cell/year combo
```{r number of tows per grid cell/year/region combo}
#cell/year/region combos
ebs_unique_tows_cell <- grid_EBS.reduced.again.again[,.SD[1],.(year, cell, haul_id)]

#no years have way more tows than others
plot(ebs_unique_tows_cell[,.N,.(year,cell)]$year, ebs_unique_tows_cell[,.N,.(year,cell)]$N)

#therefore, I will not leave out any years, but rather just pull out cells in years they are undersampled (<N=3, which is )

summary(ebs_unique_tows_cell[,.N,.(year,cell)]$N < 3) #this occurs in 538/2081 cell/year combos  (26%) (THIS IS A LOT, return to later)

#I don't think I need to exclude in ALL years, just in a few years
#an alternative would be to assess coverage using biomass (# of singletons and doubletons)

```

Delete all observations with less than 3 tows in a given year/cell
```{r delete observations with less than 3 tows}
ebs_counts <- ebs_unique_tows_cell[,.N,.(year,cell)]
ebs_keep <- ebs_counts[N>2,][,N := NULL]

grid_EBS.reduced_3plustows <- grid_EBS.reduced.again.again[ebs_keep, on = .(cell, year)]


```


Save
```{r save}
saveRDS(grid_EBS.reduced, "/Users/zoekitchel/Documents/grad school/Rutgers/Repositories/trawl_spatial_turnover/data/gridded/USA-EBS/grid_EBS.reduced.rds")


saveRDS(ebs_reg_distances.l, "/Users/zoekitchel/Documents/grad school/Rutgers/Repositories/trawl_spatial_turnover/data/gridded/USA-EBS/ebs_reg_distances.l.rds")

saveRDS(grid_EBS.reduced_3plustows,"/Users/zoekitchel/Documents/grad school/Rutgers/Repositories/trawl_spatial_turnover/data/gridded/USA-EBS/grid_EBS.reduced_3plustows.rds")


```

