---
title: "Prepping GOA Data"
output: html_notebook
---


```{r setup}
library(dggridR)
library(data.table)
library(rgdal)
library(raster)
library(sp)
library(rnaturalearth)
library(rnaturalearthdata)
library(rgeos)
library(taxize) #standardizing names
library(geosphere)  #to calculate distance between lat lon of grid cells
library(googledrive)
```


Moving forward, I plan to use data directly from FishGlob.

November 21, 2021
```{r pull in data from fishglob for West Coast Triennial from Google Drive}
drive_download("GOA_clean.csv",
               overwrite = T)

goa <- fread("GOA_clean.csv")

file.remove("GOA_clean.csv")

```

```{r pull in GOA}

# Get Unique lines in the data table
unique_goa_latlon<- unique(goa[,.(latitude, longitude)])
unique_goa_latlon.sf <- unique_goa_latlon
  
#make lat lon coordinates into polygon
coordinates(unique_goa_latlon.sf) <- ~latitude + longitude
unique_goa_latlon.sf@proj4string <- crs("+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0")

saveRDS(unique_goa_latlon, "unique_goa_latlon.RData")

#plot north america
world <- ne_countries(scale = "medium", returnclass = "sp")

world_ext <- extent(-180, -51.5,23, 67)

north_america_spdf <- crop(world, world_ext)

north_america_df <- fortify(north_america_spdf)

(na_outline <- ggplot(north_america_df, aes(long, lat, group = group)) +
  geom_polygon(color = "white", fill = "grey") +
  labs(x = "Longitude", y = "Latitude") +
  coord_equal() +
  theme_classic() +
  scale_x_continuous(expand = c(0,0)) +
  scale_y_continuous(expand = c(0,0)) +
  theme(panel.border = element_rect(color = "black", fill = NA, size = 1)) +
 geom_point(unique_goa_latlon, mapping = aes(longitude, latitude), inherit.aes = FALSE, shape = 18, size = 0.001, color = "red") +
  coord_equal())

```

Playing around with dggridR (package that makes grid cells)
```{r learning dggridR}
dggs <- dgconstruct(spacing = 111, metric = T, resround = 'nearest')

unique_goa_latlon[,cell := dgGEO_to_SEQNUM(dggs, longitude, latitude)] #get corresponding grid cells for NEUS trawl

#centersof cells
goa_cellcenters <- dgSEQNUM_to_GEO(dggs, unique_goa_latlon[,cell])

#linking cell centers to unique_goa_latlon
unique_goa_latlon[,LON_CENTER := goa_cellcenters$lon_deg][,LAT_CENTER := goa_cellcenters$lat_deg]

#link centers back to main data table

goa <- merge(goa, unique_goa_latlon, by = c("latitude", "longitude"))

#number of tows in each cell
towcount <- unique_goa_latlon[, .N, by = cell]

#get the grid cell boundary for cells which had trawls
grid_goa <- dgcellstogrid(dggs, unique_goa_latlon[,cell], frame = T, wrapcells = F)

#update grid properties to include # of trawls in each cell
grid_goa <- merge(grid, unique_goa_latlon, by.x = "cell", by.y ="cell")

ggplot(data = grid_goa, aes(x = long, y = lat, group = cell)) +
  geom_polygon() +
  coord_quickmap()

saveRDS(grid_goa, "grid_goa.rds")

```

Trying to plot with grid overlaying
```{r grid overlaying}

na_outline + 
   geom_polygon(grid_goa, mapping = aes(x = long,y = lat, group = cell), inherit.aes = FALSE, color = "black", fill = NA, size = 0.1)

ggsave(filename = "grid_GOA_cells_spacing_111.pdf")
  
```

Cleaning spp names (must be ID'd to species!) NEED TO ADD TO OTHER REGIONS
```{r clean GOA names}
goa.clean <- goa[rank == "Species",]
#drops ~1,000
```


Now, we'll overlay grid cells onto this list of tows using LAT and LON and gridrr package
```{r get rid of underused grid cells}

year_grid <- data.table(table(goa.clean$cell, goa.clean$year))

#full matrix of year x grid x  cell
cell_numbers <- unique(goa.clean$cell)
years <- unique(goa.clean$year)
seasons <- unique(goa.clean$region)

cells_cross_years_cross_region <- data.table(expand.grid(cell = cell_numbers, year = years, region = seasons))

observations <- unique(goa.clean[,.N,by = c("year","cell")])

#merge possible cell year region combinations with actual observations
goa_cell_data_nodata <- observations[cells_cross_years_cross_region, on = .(year, cell)]

#new pres/abs column
goa_cell_data_nodata[,pres := ifelse(is.na(N), 0, 1)]

#which grid cells are sampled in which years?
ggplot(goa_cell_data_nodata, aes(as.factor(year), as.factor(cell))) + 
  geom_tile(aes(fill = factor(pres)), color = "grey", size = 0.01) +
#  facet_wrap(~region, scales="free") +
  scale_fill_manual(values = c("white", "black"), breaks = c(0,1), labels = c("No Data", "Data")) +
  labs(x = "Year", y = "Cell") +
  theme(legend.title = element_blank(),
        axis.text.y = element_text(size = 4),
        axis.text.x = element_text(size = 5, angle = 90, vjust = 0.5),
        strip.text.x = element_text(size = 10),
        axis.ticks = element_blank())

ggsave(filename = "goa_data_availability.jpg", path = "/Users/zoekitchel/Documents/grad school/Rutgers/Repositories/trawl_spatial_turnover/data/gridded/USA-GOA/")

```
Delete cells that aren't sampled regularly (missing more than 2 years from 1984-2017):

- 59451
- 59370
- 2210
- 1805
- 1642
- 1640
- 1477
- 1395
- 1233
- 1229
- 1146
- 1063
- 986
- 815
- 489



Infrequently sampled cells (cells with less than 1 survey in a region every year)
```{r infrequently sampled cells}
#dump 
dump <- c(59451, 59370, 2210, 1805, 1642, 1640, 1477, 1395, 1233, 1229, 1146, 1063, 986, 815, 489)


#delete these grid cells from full data table
goa_grid.reduced <- goa.clean[!cell %in% dump]
dim(goa_grid.reduced )
#117560 observations left

```

This is where I would use rarefaction or coverage, for now I will skip
```{r rarefaction or coverage?}
#how many tows (TOW) per cell in a given year

haul_cell_unique <- unique(goa_grid.reduced[,.(haul_id, cell, year)])

goa_samplingevents_byyear <- haul_cell_unique[,.N, .(year, cell)]

goa_cells_lessthan3 <- goa_samplingevents_byyear[N<3,]

cells_to_delete <- unique(goa_cells_lessthan3[,cell])

dat_goa.reduced.again <- goa_grid.reduced[!cell %in% cells_to_delete]

#down to 1582496   observations

saveRDS(dat_goa.reduced.again, "goa_data_gridded.rds")

```

Find distance between center of each grid cell
```{r distance between grid cells}
#just need grid cell # and center lat lon for each grid cell 
goa_grid_cells <- unique(goa_grid.reduced[,.(cell, LON_CENTER, LAT_CENTER)])

goa_grid_cells <- setorder(goa_grid_cells, cell)

goa_reg_distances <- distm(goa_grid_cells[,.(LON_CENTER, LAT_CENTER)])

colnames(goa_reg_distances) <- goa_grid_cells$cell
rownames(goa_reg_distances) <- goa_grid_cells$cell

#reorient to long form 
goa_reg_distances.l <- melt(as.matrix(goa_reg_distances), varnames = c("cell1", "cell2"), value.name = "distance(m)") #matrix to data frame
goa_reg_distances.l <- data.table(goa_reg_distances.l) #and then to data table
goa_reg_distances.l <- goa_reg_distances.l[cell1 > cell2,] # get rid of repetitions and self-comparisons
saveRDS(goa_reg_distances.l, "goa_reg_distances.l.rds")
```

Check the # of tows per grid cell/year combo
```{r number of tows per grid cell/year/region combo}
#cell/year/region combos
goa_unique_tows_cell <- goa_grid.reduced[,.SD[1],.(year, cell, haul_id)]

#no years have way more tows than others
plot(goa_unique_tows_cell[,.N,.(year,cell)]$year, goa_unique_tows_cell[,.N,.(year,cell)]$N)

#therefore, I will not leave out any years, but rather just pull out cells in years they are undersampled (<N=3, which is )

summary(goa_unique_tows_cell[,.N,.(year,cell)]$N < 3) #this occurs in 50/839 cell/year combos  (6%, no biggie)

#I don't think I need to exclude in ALL years, just in a few years
#an alternative would be to assess coverage using biomass (# of singletons and doubletons)

```

Delete all observations with less than 3 tows in a given year/cell
```{r delete observations with less than 3 tows}
goa_counts <- goa_unique_tows_cell[,.N,.(year,cell)]
goa_keep <- goa_counts[N>2,][,N := NULL]

dat_goa_grid.reduced_3plustows <- goa_grid.reduced[goa_keep, on = .(cell, year)]


```


Summary of area sampled
```{r sum area sampled}
#distance between furthest north and furthest south point
library(geosphere)
max_north_south_km <- (distm(c(mean(dat_goa_grid.reduced_3plustows$longitude), min(dat_goa_grid.reduced_3plustows$latitude)), c(mean(dat_goa_grid.reduced_3plustows$longitude), max(dat_goa_grid.reduced_3plustows$latitude)), fun = distHaversine))/1000
#distance between furthest east and furthest west point
max_west_east_km <- (distm(c(min(dat_goa_grid.reduced_3plustows$longitude), mean(dat_goa_grid.reduced_3plustows$latitude)), c(max(dat_goa_grid.reduced_3plustows$longitude), mean(dat_goa_grid.reduced_3plustows$latitude)), fun = distHaversine))/1000

max_north_south_km
max_west_east_km

summary(dat_goa_grid.reduced_3plustows$depth)

length(unique(dat_goa_grid.reduced_3plustows$accepted_name))

sort(unique(as.numeric(goa$year)))
length(unique(goa$year))
```

Save
```{r save}
saveRDS(goa_grid.reduced, "/Users/zoekitchel/Documents/grad school/Rutgers/Repositories/trawl_spatial_turnover/data/gridded/USA-GOA/GOA_grid.reduced.rds")


saveRDS(goa_reg_distances.l, "/Users/zoekitchel/Documents/grad school/Rutgers/Repositories/trawl_spatial_turnover/data/gridded/USA-GOA/GOA_reg_distances.l.rds")

saveRDS(dat_goa_grid.reduced_3plustows,"/Users/zoekitchel/Documents/grad school/Rutgers/Repositories/trawl_spatial_turnover/data/gridded/USA-GOA/GOA_grid.reduced_3plustows.rds")


```
