---
title: "Look at multiple regions: first, grid using dggridR"
output: html_notebook
---

First, I had to download dggridR. Note that the version that is installed using install.packages() was not updated to work with this version of R. Therefore, I downloaded directly from GitHub using:

library(devtools) #Use `install.packages('devtools')` if need be
install_github('r-barnes/dggridR', vignette=TRUE)

on November 23, 2020. 

```{r setup}
library(dggridR)
library(data.table)
library(rgdal)
library(raster)
library(sp)
library(rnaturalearth)
library(rnaturalearthdata)
library(rgeos)
library(taxize) #standardizing names
library(googledrive)
```


Moving forward, I plan to use data directly from FishGlob.

November 21, 2021
```{r pull in data from fishglob for wcann from Google Drive}
drive_download("NEUS_clean.csv",
               overwrite = T)

neus <- fread("NEUS_clean.csv")

file.remove("NEUS_clean.csv")

setkey(neus)

#spring only
neus.spring <- neus[season == "Spring",]

# Get unique lat lon values in the data table
unique_neus_latlon <- unique(neus.spring[,.(latitude, longitude)])
unique_neus_latlon.sf <- unique_neus_latlon
  
#make lat lon coordinates into polygon
coordinates(unique_neus_latlon.sf) <- ~latitude + longitude
unique_neus_latlon.sf@proj4string <- crs("+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0")

saveRDS(unique_neus_latlon, "unique_neus_latlon.rds") #save list of lat lon coordinates


```

Visualize NEUS sampling sites
```{r visualize sampling sites}

#open RDS if already saved
#unique_neus_latlon <- readRDS("/Users/zoekitchel/Documents/grad school/Rutgers/Repositories/trawl_spatial_turnover/unique_neus_latlon.RData")

#plot north america
world <- ne_countries(scale = "medium", returnclass = "sp")

world_ext <- extent(-80, -60,30, 50)

north_america_spdf <- crop(world, world_ext)

north_america_df <- fortify(north_america_spdf)

(na_outline <- ggplot(north_america_df, aes(long, lat, group = group)) +
  geom_polygon(color = "white", fill = "grey") +
  labs(x = "Longitude", y = "Latitude") +
  coord_equal() +
  theme_classic() +
  scale_x_continuous(expand = c(0,0)) +
  scale_y_continuous(expand = c(0,0)) +
  theme(panel.border = element_rect(color = "black", fill = NA, size = 1)) +
 geom_point(unique_neus_latlon, mapping = aes(longitude, latitude), inherit.aes = FALSE, shape = 18, size = 0.001, color = "red") +
  coord_equal())
```

Setting up hexagonal grid cells with dggridR (package that makes grid cells)
```{r learning dggridR}
dggs <- dgconstruct(spacing = 111, metric = T, resround = 'nearest') #make cells with 111km spacing for whole planet

unique_neus_latlon[,cell := dgGEO_to_SEQNUM(dggs, longitude, latitude)] #get corresponding grid cells for NEUS trawl

#centers of cells 
neus_cellcenters <- dgSEQNUM_to_GEO(dggs, unique_neus_latlon[,cell])

#linking cell centers to unique_neus_latlon
unique_neus_latlon[,LON_CENTER := neus_cellcenters$lon_deg][,LAT_CENTER := neus_cellcenters$lat_deg]

#link centers back to main data table

neus.spring_grid <- merge(neus.spring, unique_neus_latlon, by = c("latitude", "longitude"))


#number of tows in each cell by season for all years
towcount <- unique_neus_latlon[, list(Freq = .N), by = list(cell)]

#get the grid cell boundary for cells which had trawls
grid_neus <- dgcellstogrid(dggs, unique_neus_latlon[,cell], frame = T, wrapcells = F)
colnames(grid_neus) <- c("lon_grid", "lat_grid", "order", "hole", "piece", "group", "cell")

#merge polygon outlines with all haul coordinates within for each season
grid_neus <- merge(grid_neus, unique_neus_latlon, by = "cell") #this takes a while

#$region <- as.factor(grid_neus$region)


#spring
ggplot(data = grid_neus, aes(x = lon_grid, y = lat_grid, group = cell)) +
  geom_polygon() +
    lims(x = c(-77.5,-64), y = c(34.5,45.5)) +
  coord_quickmap() +
  theme_classic()

saveRDS(grid_neus, "grid_neus.RData")

```

Trying to plot with grid overlaying
```{r grid overlaying}

na_outline + 
   geom_polygon(grid_neus, mapping = aes(x = lon_grid,y = lat_grid, group = cell), inherit.aes = FALSE, color = "black", fill = NA, size = 0.1)

ggsave(filename = "neus_grid_cells_spacing_111.pdf")
  
```


Cleaning up NEUS data


```{r clean up NEUS}
neus.spring_grid.clean <- neus.spring_grid[rank == "Species"]
```

Now, we'll overlay grid cells onto this list of tows using LAT and LON and gridrr package
```{r get rid of underused grid cells}

year_grid <- data.table(table(neus.spring_grid.clean$cell, neus.spring_grid.clean$year))

#full matrix of year x grid x  cell
cell_numbers <- unique(neus.spring_grid.clean$cell)
years <- unique(neus.spring_grid.clean$year)
seasons <- unique(neus.spring_grid.clean$region)

cells_cross_years_cross_region <- data.table(expand.grid(cell = cell_numbers, year = years, region = seasons))

observations <- unique(neus.spring_grid.clean[,.N,by = c("year","cell")])

#merge possible cell year region combinations with actual observations
neus_cell_data_nodata <- observations[cells_cross_years_cross_region, on = .(year, cell)]

#new pres/abs column
neus_cell_data_nodata[,pres := ifelse(is.na(N), 0, 1)]

#which grid cells are sampled in which years?
ggplot(neus_cell_data_nodata, aes(as.factor(year), as.factor(cell))) + 
  geom_tile(aes(fill = factor(pres)), color = "grey", size = 0.01) +
#  facet_wrap(~region, scales="free") +
  scale_fill_manual(values = c("white", "black"), breaks = c(0,1), labels = c("No Data", "Data")) +
  labs(x = "Year", y = "Cell") +
  theme(legend.title = element_blank(),
        axis.text.y = element_text(size = 4),
        axis.text.x = element_text(size = 5, angle = 90, vjust = 0.5),
        strip.text.x = element_text(size = 10),
        axis.ticks = element_blank()) 

ggsave(filename = "neus_data_availability.jpg", path = "/Users/zoekitchel/Documents/grad school/Rutgers/Repositories/trawl_spatial_turnover/data/USA-NEUS/")

```
Delete cells that aren't sampled regularly (missing more than 2 years from 1972-2019):

- 6664: sampled sporadically in spring and fall
- 6583: sampled sporadically in spring and fall
- 6576: sampled sporadically in spring and fall
- 6575: sampled sporadically in spring and fall
- 6571: sampled sporadically in spring and fall
- 6076: missing a few observations in spring and fall, dump fall keep spring
- 5831: only sampled once in fall, never in spring
- 5507: sampled sporadically in fall, keep for spring only
- 5347: sampled sporadically in spring and fall, keep for spring only
- 5346: sampled sporadically in spring and fall
- 5265: missing a few observations in spring and fall, keep for spring only
- 5102: sampled sporadically in spring and fall
- 5022: missing a few observations in spring and fall, keep for spring only
- 4942:sampled sporadically in spring and fall
- 4941: sampled sporadically in spring and fall

Years with issues to look into:

- 2017 fall
- 1972, 1975, 1977, 2014 spring


Infrequently sampled cells (cells with less than 1 survey in a region every year)
```{r infrequently sampled cells}
#dump for spring
dump_spring <- c(6664, 6583, 6576, 6575, 6571, 5831, 5346, 5102, 4942, 4941)

#dump for fall
dump_fall <- c(6664, 6583, 6576, 6575, 6571, 6076, 5831, 5507, 5347, 5346, 5265, 5102, 5022, 4942, 4941)


#delete these grid cells from full data table
neus.spring_grid.clean.spring.reduced <- neus.spring_grid.clean[!cell %in% dump_spring]
#neus.spring_grid.clean.fall.reduced <- neus.spring_grid.clean[!cell %in% dump_fall][region == "Northeast US Fall",]

#1116697 fall observations left, 882424 spring observations left

```

This is where I would use rarefaction or coverage, for now I will skip
```{r rarefaction or coverage?}
#how many tows (TOW) per cell in a given year
colnames(neus.spring_grid.clean.spring.reduced)

cell_tow_year <- unique(neus.spring_grid.clean.spring.reduced[,.(haul_id, cell, year)])

neus_samplingevents_byyear <- cell_tow_year[,.N, .(year, cell)]

neus_cells_lessthan3 <- neus_samplingevents_byyear[N<3,] #need three to build rarefaction curve, so, will get rid of less than three observations

cells_to_delete <- unique(neus_cells_lessthan3[,cell])

neus.spring_grid.clean.spring.reduced.again <- neus.spring_grid.clean.spring.reduced[!cell %in% cells_to_delete]

#1375482 observations

saveRDS(neus_full_1982onward.clean.reduced.again, "neus_data_gridded.RData")
saveRDS(neus_full_1982onward.clean.reduced.again, "neus_data_gridded.RData")
```

Find distance between center of each grid cell
```{r distance between grid cells}

#spring distances
#just need grid cell # and center lat lon for each grid cell 
neus_grid_cells_spring <- unique(neus.spring_grid.clean.spring.reduced.again[,.(cell, LON_CENTER, LAT_CENTER)])

neus_grid_cells_spring <- setorder(neus_grid_cells_spring, cell)

neus_reg_distances_spring <- distm(neus_grid_cells_spring[,.(LON_CENTER, LAT_CENTER)])

colnames(neus_reg_distances_spring) <- neus_grid_cells_spring$cell
rownames(neus_reg_distances_spring) <- neus_grid_cells_spring$cell

#reorient to long form 
neus_reg_distances_spring.l <- reshape2::melt(as.matrix(neus_reg_distances_spring), varnames = c("cell1", "cell2"), value.name = "distance(m)") #matrix to data frame
neus_reg_distances_spring.l <- data.table(neus_reg_distances_spring.l) #and then to data table
neus_reg_distances_spring.l <- neus_reg_distances_spring.l[cell1 > cell2,] # get rid of repetitions and self-comparisons

##fall distances
##just need grid cell # and center lat lon for each grid cell 
#neus_grid_cells_fall <- unique(neus.spring_grid.fall.reduced[,.(cell, LON_CENTER, LAT_CENTER)])
#
#neus_grid_cells_fall <- setorder(neus_grid_cells_fall, cell)
#
#neus_reg_distances_fall <- distm(neus_grid_cells_fall[,.(LON_CENTER, LAT_CENTER)])
#
#colnames(neus_reg_distances_fall) <- neus_grid_cells_fall$cell
#rownames(neus_reg_distances_fall) <- neus_grid_cells_fall$cell
#
##reorient to long form 
#neus_reg_distances_fall.l <- reshape2::melt(as.matrix(neus_reg_distances_fall), varnames = c("cell1", "cell2"), #value.name = "distance(m)") #matrix to data frame
#neus_reg_distances_fall.l <- data.table(neus_reg_distances_fall.l) #and then to data table
#neus_reg_distances_fall.l <- neus_reg_distances_fall.l[cell1 > cell2,] # get rid of repetitions and #self-comparisons
#
#
```


Check the # of tows per grid cell/year combo
```{r number of tows per grid cell/year/region combo}
#spring cell/year/region combos
neus_spring_unique_tows_cell <- neus.spring_grid.clean.spring.reduced.again[,.SD[1],.(year, cell, haul_id)]

#no years have way more tows than others
plot(neus_spring_unique_tows_cell[,.N,.(year,cell)]$year, neus_spring_unique_tows_cell[,.N,.(year,cell)]$N)

#fall cell/year/region combos
neus_fall_unique_tows_cell <- neus.spring_grid.clean.spring.reduced.again[,.SD[1],.(year, cell, haul_id)]

#no years have way more tows than others
plot(neus_fall_unique_tows_cell[,.N,.(year,cell)]$year, neus_fall_unique_tows_cell[,.N,.(year,cell)]$N)

#therefore, I will not leave out any years, but rather just pull out cells in years they are undersampled (<N=3, which is )

summary(neus_fall_unique_tows_cell[,.N,.(year,cell)]$N < 3) #for fall, this occurs in 131/1650 (8%) year, cell combos
summary(neus_spring_unique_tows_cell[,.N,.(year,cell)]$N < 3) #for spring this occurs in 243/1638 (15%) cases, year, cell combos

#I don't think I need to exclude in ALL years, just in a few years
#an alternative would be to assess coverage using biomass (# of singletons and doubletons)

```

Delete all observations with less than 3 tows in a given year/cell
```{r delete observations with less than 3 tows}
#fall
#neus_fall_counts <- neus_fall_unique_tows_cell[,.N,.(year,cell)]
#neus_fall_keep <- neus_fall_counts[N>2,][,N := NULL]
#
#neus.spring_grid.fall.reduced_3plustows <- neus.spring_grid.fall.reduced[neus_fall_keep, on = .(cell, year)]

#spring
neus_spring_counts <- neus_spring_unique_tows_cell[,.N,.(year,cell)]
neus_spring_keep <- neus_spring_counts[N>2,][,N := NULL]

neus.spring_grid.spring.reduced_3plustows <- neus.spring_grid.clean.spring.reduced.again[neus_spring_keep, on = .(cell, year)]
```

Summary of area sampled
```{r sum area sampled}
#distance between furthest north and furthest south point
library(geosphere)
max_north_south_km <- (distm(c(mean(neus.spring_grid.spring.reduced_3plustows$longitude), min(neus.spring_grid.spring.reduced_3plustows$latitude)), c(mean(neus.spring_grid.spring.reduced_3plustows$longitude), max(neus.spring_grid.spring.reduced_3plustows$latitude)), fun = distHaversine))/1000
#distance between furthest east and furthest west point
max_west_east_km <- (distm(c(min(neus.spring_grid.spring.reduced_3plustows$longitude), mean(neus.spring_grid.spring.reduced_3plustows$latitude)), c(max(neus.spring_grid.spring.reduced_3plustows$longitude), mean(neus.spring_grid.spring.reduced_3plustows$latitude)), fun = distHaversine))/1000

max_north_south_km
max_west_east_km

summary(neus.spring_grid.spring.reduced_3plustows$depth)

length(unique(neus.spring_grid.spring.reduced_3plustows$accepted_name))

sort(unique(as.numeric(neus.spring_grid.clean$year)))
length(unique(neus.spring_grid.clean$year))
```

Make a new map but with only points we actually use spring 
```{r new map points we use spring}
(na_outline_2 <- ggplot(north_america_df, aes(long, lat, group = group)) +
  geom_polygon(color = "white", fill = "grey") +
  labs(x = "Longitude", y = "Latitude") +
  coord_equal() +
  theme_classic() +
  scale_x_continuous(expand = c(0,0)) +
  scale_y_continuous(expand = c(0,0)) +
  theme(panel.border = element_rect(color = "black", fill = NA, size = 1)) +
 geom_point(unique(neus.spring_grid.spring.reduced_3plustows[,.(latitude, longitude)]), mapping = aes(longitude, latitude), inherit.aes = FALSE, shape = 18, size = 0.001, color = "red") +
  coord_equal())
```

Make a new map but with only points we actually use fall
```{r new map points we use fall}
(na_outline_2 <- ggplot(north_america_df, aes(long, lat, group = group)) +
  geom_polygon(color = "white", fill = "grey") +
  labs(x = "Longitude", y = "Latitude") +
  coord_equal() +
  theme_classic() +
  scale_x_continuous(expand = c(0,0)) +
  scale_y_continuous(expand = c(0,0)) +
  theme(panel.border = element_rect(color = "black", fill = NA, size = 1)) +
 geom_point(unique(neus.spring_grid.spring.reduced_3plustows[,.(latitude, longitude)]), mapping = aes(longitude, latitude), inherit.aes = FALSE, shape = 18, size = 0.001, color = "red") +
  coord_equal())
```

Save
```{r save}
saveRDS(neus.spring_grid.spring.reduced, "/Users/zoekitchel/Documents/grad school/Rutgers/Repositories/trawl_spatial_turnover/data/gridded/USA-NEUS/neus.spring_grid.spring.reduced.RData")
saveRDS(neus_reg_distances_spring.l, "/Users/zoekitchel/Documents/grad school/Rutgers/Repositories/trawl_spatial_turnover/data/gridded/USA-NEUS/neus_reg_distances_spring.l.RData")

saveRDS(neus.spring_grid.fall.reduced, "/Users/zoekitchel/Documents/grad school/Rutgers/Repositories/trawl_spatial_turnover/data/gridded/USA-NEUS/neus.spring_grid.fall.reduced.RData")
saveRDS(neus_reg_distances_fall.l, "/Users/zoekitchel/Documents/grad school/Rutgers/Repositories/trawl_spatial_turnover/data/gridded/USA-NEUS/neus_reg_distances_fall.l.RData")

saveRDS(neus.spring_grid.fall.reduced_3plustows,"/Users/zoekitchel/Documents/grad school/Rutgers/Repositories/trawl_spatial_turnover/data/gridded/USA-NEUS/neus.spring_grid.fall.reduced_3plustows.RData")

saveRDS(neus.spring_grid.spring.reduced_3plustows,"/Users/zoekitchel/Documents/grad school/Rutgers/Repositories/trawl_spatial_turnover/data/gridded/USA-NEUS/neus.spring_grid.spring.reduced_3plustows.RData")


```

*************

* [bigelow fall calibration NEUS](https://github.com/pinskylab/OceanAdapt/blob/master/data_raw/bigelow-fall-calibration.csv)
* [bigelow spring calibration NEUS](https://github.com/pinskylab/OceanAdapt/blob/master/data_raw/bigelow-spring-calibration.csv)

* [AI strata](https://github.com/pinskylab/OceanAdapt/blob/master/data_raw/ai_strata.csv)
* [AI 1983-2000](https://github.com/pinskylab/OceanAdapt/blob/master/data_raw/ai1983_2000.csv)
* [AI 2002-2012](https://github.com/pinskylab/OceanAdapt/blob/master/data_raw/ai2002_2012.csv)
* [AI 2012-2018](https://github.com/pinskylab/OceanAdapt/blob/master/data_raw/ai2012_2018.csv)

* [WCTRI haul](https://github.com/pinskylab/OceanAdapt/blob/master/data_raw/wctri_haul.csv)
* [WCTRI catch](https://github.com/pinskylab/OceanAdapt/blob/master/data_raw/wctri_catch.csv)
* [WCTRI species](https://github.com/pinskylab/OceanAdapt/blob/master/data_raw/wctri_species.csv)

* [WCANN_haul](https://github.com/pinskylab/OceanAdapt/blob/master/data_raw/wcann_haul.csv)
* [WCANN_catch.csv.zip](https://github.com/pinskylab/OceanAdapt/blob/master/data_raw/wcann_catch.csv.zip)

* [SEUS catch](https://github.com/pinskylab/OceanAdapt/blob/master/data_raw/seus_catch.csv.zip)
* [SEUS haul](https://github.com/pinskylab/OceanAdapt/blob/master/data_raw/seus_haul.csv)
* [SEUS_strata](https://github.com/pinskylab/OceanAdapt/blob/master/data_raw/seus_strata.csv)

* [scotian summer](https://github.com/pinskylab/OceanAdapt/blob/master/data_raw/scot_summer.csv.zip)
* [scotian spring](https://github.com/pinskylab/OceanAdapt/blob/master/data_raw/scot_spring.csv)
* [scotian fall](https://github.com/pinskylab/OceanAdapt/blob/master/data_raw/scot_fall.csv)

* [GOA strata](https://github.com/pinskylab/OceanAdapt/blob/master/data_raw/goa_strata.csv)
* [GOA 1984-1987](https://github.com/pinskylab/OceanAdapt/blob/master/data_raw/goa2015_2017.csv)
* [GOA 1990-1999](https://github.com/pinskylab/OceanAdapt/blob/master/data_raw/goa2015_2017.csv)
* [GOA 2001-2005](https://github.com/pinskylab/OceanAdapt/blob/master/data_raw/goa2015_2017.csv)
* [GOA 2007-2013](https://github.com/pinskylab/OceanAdapt/blob/master/data_raw/goa2015_2017.csv)
* [GOA 2015-2017](https://github.com/pinskylab/OceanAdapt/blob/master/data_raw/goa2015_2017.csv)

* [GMEX starrec](https://github.com/pinskylab/OceanAdapt/blob/master/data_raw/gmex_STAREC.csv)
* [GMEX new bio codes big](https://github.com/pinskylab/OceanAdapt/blob/master/data_raw/gmex_NEWBIOCODESBIG.csv)
* [GMEX invrec](https://github.com/pinskylab/OceanAdapt/blob/master/data_raw/INVREC.csv)
* [GMEX cruises](https://github.com/pinskylab/OceanAdapt/blob/master/data_raw/gmex_CRUISES.csv)
* [GMEX BGSREC](https://github.com/pinskylab/OceanAdapt/blob/master/data_raw/gmex_BGSREC.csv.zip)

* [EBS strata](https://github.com/pinskylab/OceanAdapt/blob/master/data_raw/ebs_strata.csv)
* [EBS 1982-1984](https://github.com/pinskylab/OceanAdapt/blob/master/data_raw/ebs1982_1984.csv)
* [EBS 1985-1989](https://github.com/pinskylab/OceanAdapt/blob/master/data_raw/ebs1985_1989.csv)
* [EBS 1990-1994](https://github.com/pinskylab/OceanAdapt/blob/master/data_raw/ebs1990_1994.csv)
* [EBS 1995-1999](https://github.com/pinskylab/OceanAdapt/blob/master/data_raw/ebs1995_1999.csv)
* [EBS 2000-2004](https://github.com/pinskylab/OceanAdapt/blob/master/data_raw/ebs2000_2004.csv)
* [EBS 2005-2008](https://github.com/pinskylab/OceanAdapt/blob/master/data_raw/ebs2005_2008.csv)
* [EBS 2009-2012](https://github.com/pinskylab/OceanAdapt/blob/master/data_raw/ebs2009_2012.csv)
* [EBS 2013-2016](https://github.com/pinskylab/OceanAdapt/blob/master/data_raw/ebs2013_2016.csv)
* [EBS 2017-2018](https://github.com/pinskylab/OceanAdapt/blob/master/data_raw/ebs2017_2018.csv)

* [AI strata](https://github.com/pinskylab/OceanAdapt/blob/master/data_raw/ai_strata.csv)
* [AI 1983-2000](https://github.com/pinskylab/OceanAdapt/blob/master/data_raw/ai1983_2000.csv)
* [AI 2002-2012](https://github.com/pinskylab/OceanAdapt/blob/master/data_raw/ai2002_2012.csv)
* [AI 2014-2018](https://github.com/pinskylab/OceanAdapt/blob/master/data_raw/ai2014_2018.csv)

* [SEUS cloud](https://github.com/pinskylab/OceanAdapt/blob/master/data_raw/SVDBS_CLOUD.csv)
* [SEUS maturity codes](https://github.com/pinskylab/OceanAdapt/blob/master/data_raw/SVDBS_MATURITY_CODES.csv)
* [SEUS sex codes](https://github.com/pinskylab/OceanAdapt/blob/master/data_raw/SVDBS_SEX_CODES.csv)
* [SEUS SV gear](https://github.com/pinskylab/OceanAdapt/blob/master/data_raw/SVDBS_SVGEAR.csv)
* [SEUS SV vessel](https://github.com/pinskylab/OceanAdapt/blob/master/data_raw/SVDBS_SVVESSEL.csv)
* [SEUS weathre](https://github.com/pinskylab/OceanAdapt/blob/master/data_raw/SVDBS_WEATHER.csv)
* [SEUS xbt](https://github.com/pinskylab/OceanAdapt/blob/master/data_raw/SVDBS_XBT.csv)




```{r import data for NEUS from Ocean Adapt on GitHub}

neus_strata <- fread('https://raw.githubusercontent.com/pinskylab/OceanAdapt/master/data_raw/neus_strata.csv')
setnames(neus_strata, c('stratum', 'stratum_name'), c('STRATUM', 'STRATUM_NAME'))
load(url('https://raw.githubusercontent.com/pinskylab/OceanAdapt/master/data_raw/neus_Survdat.RData')) # loads survdat
load(url('https://raw.githubusercontent.com/pinskylab/OceanAdapt/master/data_raw/neus_SVSPP.RData')) # load spp

```

```{r }

```




Questions:
 - should I merge abundance and presence absense from all trawl surveys? or pick trawl surveys during a particular time of year?
 - abundance versus biomass in trawl data? which to use?
 - what do negative abundance values mean for US?
 - coverage versus sample based rarefaction?