---
title: "Prepping NEUS Data"
output: html_notebook
---


```{r setup}
library(dggridR)
library(data.table)
library(rgdal)
library(raster)
library(sp)
library(rnaturalearth)
library(rnaturalearthdata)
library(rgeos)
library(taxize) #standardizing names
library(geosphere)  #to calculate distance between lat lon of grid cells
library(googledrive)
```


Moving forward, I plan to use data directly from FishGlob.

November 30, 2021
```{r pull in data from fishglob for NEUS from Google Drive}
drive_download("NEUS_clean.csv",
               overwrite = T)

neus <- fread("NEUS_clean.csv")

file.remove("NEUS_clean.csv")

```
What seasons do we have? Which has most data? Eventually, look at both? (Fall and Summer have similar numbers)
```{r seasons?}
summary(as.factor(neus$season))

neus_fall <- neus[season == "Fall"]
neus_spring <- neus[season == "Spring"]
```

##FALL
```{r pull in fall NEUS}

# Get Unique lines in the data table
unique_neus_latlon_fall <- unique(neus_fall[,.(latitude, longitude)])
unique_neus_latlon.sf <- unique_neus_latlon_fall
  
#make lat lon coordinates into polygon
coordinates(unique_neus_latlon.sf) <- ~latitude + longitude
unique_neus_latlon.sf@proj4string <- crs("+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0")

saveRDS(unique_neus_latlon_fall, "unique_neus_latlon_fall.rds")

#plot north america
world <- ne_countries(scale = "medium", returnclass = "sp")

world_ext <- extent(-85, -60,30, 50)

north_america_spdf <- crop(world, world_ext)

north_america_df <- fortify(north_america_spdf)

(na_outline <- ggplot(north_america_df, aes(long, lat, group = group)) +
  geom_polygon(color = "white", fill = "grey") +
  labs(x = "Longitude", y = "Latitude") +
  coord_equal() +
  theme_classic() +
  scale_x_continuous(expand = c(0,0)) +
  scale_y_continuous(expand = c(0,0)) +
  theme(panel.border = element_rect(color = "black", fill = NA, size = 1)) +
 geom_point(unique_neus_latlon_fall, mapping = aes(longitude, latitude), inherit.aes = FALSE, shape = 18, size = 0.001, color = "red") +
  coord_equal())

```

Playing around with dggridR (package that makes grid cells)
```{r learning dggridR fall}
dggs <- dgconstruct(spacing = 111, metric = T, resround = 'nearest')

unique_neus_latlon_fall[,cell := dgGEO_to_SEQNUM(dggs, longitude, latitude)] #get corresponding grid cells for NEUS trawl

#centersof cells
neus_cellcenters <- dgSEQNUM_to_GEO(dggs, unique_neus_latlon_fall[,cell])

#linking cell centers to unique_neus_latlon
unique_neus_latlon_fall[,LON_CENTER := neus_cellcenters$lon_deg][,LAT_CENTER := neus_cellcenters$lat_deg]

#link centers back to main data table

neus_fall <- merge(neus_fall, unique_neus_latlon_fall, by = c("latitude", "longitude"))

#number of tows in each cell
towcount <- unique_neus_latlon_fall[, .N, by = cell]

#get the grid cell boundary for cells which had trawls
grid_neus_fall <- dgcellstogrid(dggs, unique_neus_latlon_fall[,cell], frame = T, wrapcells = F)

#update grid properties to include # of trawls in each cell
grid_neus_fall <- merge(grid_neus_fall, unique_neus_latlon_fall, by = "cell")

ggplot(data = grid_neus_fall, aes(x = long, y = lat, group = cell)) +
  geom_polygon() +
  coord_quickmap()

saveRDS(grid_neus_fall, "grid_neus_fall.rds")

```

Trying to plot with grid overlaying
```{r grid overlaying fall}

na_outline + 
   geom_polygon(grid_neus_fall, mapping = aes(x = long,y = lat, group = cell), inherit.aes = FALSE, color = "black", fill = NA, size = 0.1)

ggsave(filename = "grid_NEUS_cells_spacing_111_fall.pdf")
  
```
Cleaning spp names (must be ID'd to species!)
```{r clean NEUS names to species fall}
#unique species names
neus_fall.clean <- neus_fall[rank == "Species"]


```


Now, we'll overlay grid cells onto this list of tows (dat_neus) using lat and lon and gridrr package
```{r get rid of underused grid cells fall}

year_grid <- data.table(table(neus_fall.clean$cell, neus_fall.clean$year))

#there are some grid cells with infrequent records, I will delete these
grid_cells_to_exclude <- unique(year_grid[N == 0]$V1)

#delete these grid cells from full data table
neus_fall.reduced <- neus_fall.clean[!cell %in% grid_cells_to_exclude]

#115782 observations left, this should be great

#how many tows (TOW) per cell in a given year

haul_cell_unique <- unique(neus_fall.reduced[,.(haul_id, cell, year)])

neus_fall_samplingevents_byyear <- haul_cell_unique[,.N, .(year, cell)]

neus_fall_cells_lessthan3 <- neus_fall_samplingevents_byyear[N<3,]

cells_to_delete <- unique(neus_fall_cells_lessthan3[,cell])

neus_fall.reduced.again <- neus_fall.reduced[!cell %in% cells_to_delete]

#down to 85563 observations

saveRDS(neus_fall.reduced.again, "neus_fall.reduced.again.rds")

```

Find distance between center of each grid cell
```{r distance between grid cells fall}
#just need grid cell # and center lat lon for each grid cell 
neus_fall_grid_cells <- unique(neus_fall.reduced.again[,.(cell, LON_CENTER, LAT_CENTER)])

neus_fall_grid_cells <- setorder(neus_fall_grid_cells, cell)

neus_fall_reg_distances <- distm(neus_fall_grid_cells[,.(LON_CENTER, LAT_CENTER)])

colnames(neus_fall_reg_distances) <- neus_fall_grid_cells$cell
rownames(neus_fall_reg_distances) <- neus_fall_grid_cells$cell

#reorient to long form 
neus_fall_reg_distances.l <- melt(as.matrix(neus_fall_reg_distances), varnames = c("cell1", "cell2"), value.name = "distance(m)") #matrix to data frame
neus_fall_reg_distances.l <- data.table(neus_fall_reg_distances.l) #and then to data table
neus_fall_reg_distances.l <- neus_fall_reg_distances.l[cell1 > cell2,] # get rid of repetitions and self-comparisons
saveRDS(neus_fall_reg_distances.l, "neus_fall_reg_distances.l.rds")
```

CHECK CLAIM THAT #no years have way more tows than others, because this isn't true

Check the # of tows per grid cell/year combo
```{r number of tows per grid cell/year/region combo for fall}
#cell/year/region combos
neus_fall_unique_tows_cell <- neus_fall.reduced[,.SD[1],.(year, cell, haul_id)]

#no years have way more tows than others
plot(neus_fall_unique_tows_cell[,.N,.(year,cell)]$year, neus_fall_unique_tows_cell[,.N,.(year,cell)]$N)

#therefore, I will not leave out any years, but rather just pull out cells in years they are undersampled (<N=3, which is )

summary(neus_fall_unique_tows_cell[,.N,.(year,cell)]$N < 3) #this occurs in 8/520 cell/year combos  (1.5%%)

#I don't think I need to exclude in ALL years, just in a few years
#an alternative would be to assess coverage using biomass (# of singletons and doubletons)

```

Delete all observations with less than 3 tows in a given year/cell
```{r delete observations with less than 3 tows fall}
neus_fall_counts <- neus_fall_unique_tows_cell[,.N,.(year,cell)]
neus_fall_keep <- neus_fall_counts[N>2,][,N := NULL]

neus_fall_grid.reduced_3plustows <- neus_fall.reduced[neus_fall_keep, on = .(cell, year)]


```


Save
```{r save fall}
saveRDS(neus_fall.reduced, "/Users/zoekitchel/Documents/grad school/Rutgers/Repositories/trawl_spatial_turnover/data/gridded/USA-NEUS/neus_fall.reduced.rds")


saveRDS(neus_fall_reg_distances.l, "/Users/zoekitchel/Documents/grad school/Rutgers/Repositories/trawl_spatial_turnover/data/gridded/USA-NEUS/neus_fall_reg_distances.l.rds")

saveRDS(neus_fall_grid.reduced_3plustows,"/Users/zoekitchel/Documents/grad school/Rutgers/Repositories/trawl_spatial_turnover/data/gridded/USA-NEUS/neus_fall_grid.reduced_3plustows.rds")


```

##SPRING

```{r pull in spring NEUS}

# Get Unique lines in the data table
unique_neus_latlon_spring <- unique(neus_spring[,.(latitude, longitude)])
unique_neus_latlon.sf <- unique_neus_latlon_spring
  
#make lat lon coordinates into polygon
coordinates(unique_neus_latlon.sf) <- ~latitude + longitude
unique_neus_latlon.sf@proj4string <- crs("+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0")

saveRDS(unique_neus_latlon_spring, "unique_neus_latlon_spring.rds")

#plot north america
world <- ne_countries(scale = "medium", returnclass = "sp")

world_ext <- extent(-100, -80,23, 32)

north_america_spdf <- crop(world, world_ext)

north_america_df <- fortify(north_america_spdf)

(na_outline <- ggplot(north_america_df, aes(long, lat, group = group)) +
  geom_polygon(color = "white", fill = "grey") +
  labs(x = "Longitude", y = "Latitude") +
  coord_equal() +
  theme_classic() +
  scale_x_continuous(expand = c(0,0)) +
  scale_y_continuous(expand = c(0,0)) +
  theme(panel.border = element_rect(color = "black", fill = NA, size = 1)) +
 geom_point(unique_neus_latlon_spring, mapping = aes(longitude, latitude), inherit.aes = FALSE, shape = 18, size = 0.001, color = "red") +
  coord_equal())

```

Playing around with dggridR (package that makes grid cells)
```{r learning dggridR spring}
dggs <- dgconstruct(spacing = 111, metric = T, resround = 'nearest')

unique_neus_latlon_spring[,cell := dgGEO_to_SEQNUM(dggs, longitude, latitude)] #get corresponding grid cells for NEUS trawl

#centersof cells
neus_cellcenters <- dgSEQNUM_to_GEO(dggs, unique_neus_latlon_spring[,cell])

#linking cell centers to unique_neus_latlon
unique_neus_latlon_spring[,LON_CENTER := neus_cellcenters$lon_deg][,LAT_CENTER := neus_cellcenters$lat_deg]

#link centers back to main data table

neus_spring <- merge(neus_spring, unique_neus_latlon_spring, by = c("latitude", "longitude"))

#number of tows in each cell
towcount <- unique_neus_latlon_spring[, .N, by = cell]

#get the grid cell boundary for cells which had trawls
grid_neus_spring <- dgcellstogrid(dggs, unique_neus_latlon_spring[,cell], frame = T, wrapcells = F)

#update grid properties to include # of trawls in each cell
grid_neus_spring <- merge(grid_neus_spring, unique_neus_latlon_spring, by = "cell")

ggplot(data = grid_neus_spring, aes(x = long, y = lat, group = cell)) +
  geom_polygon() +
  coord_quickmap()

saveRDS(grid_neus_spring, "grid_neus_spring.rds")

```

Trying to plot with grid overlaying
```{r grid overlaying spring}

na_outline + 
   geom_polygon(grid_neus_spring, mapping = aes(x = long,y = lat, group = cell), inherit.aes = FALSE, color = "black", fill = NA, size = 0.1)

ggsave(filename = "grid_NEUS_cells_spacing_111_spring.pdf")
  
```
Cleaning spp names (must be ID'd to species!)
```{r clean NEUS names to species spring}
#unique species names
neus_spring.clean <- neus_spring[rank == "Species"]


```


Now, we'll overlay grid cells onto this list of tows (dat_neus) using lat and lon and gridrr package
```{r get rid of underused grid cells spring}

year_grid <- data.table(table(neus_spring.clean$cell, neus_spring.clean$year))

#there are some grid cells with infrequent records, I will delete these
grid_cells_to_exclude <- unique(year_grid[N == 0]$V1)

#delete these grid cells from full data table
neus_spring.reduced <- neus_spring.clean[!cell %in% grid_cells_to_exclude]

#107961 observations left, this should be great

#how many tows (TOW) per cell in a given year

haul_cell_unique <- unique(neus_spring.reduced[,.(haul_id, cell, year)])

neus_spring_samplingevents_byyear <- haul_cell_unique[,.N, .(year, cell)]

neus_spring_cells_lessthan3 <- neus_spring_samplingevents_byyear[N<3,]

cells_to_delete <- unique(neus_spring_cells_lessthan3[,cell])

neus_spring.reduced.again <- neus_spring.reduced[!cell %in% cells_to_delete]

#down to 52851 observations

saveRDS(neus_spring.reduced.again, "neus_spring.reduced.again.rds")

```

Find distance between center of each grid cell
```{r distance between grid cells spring}
#just need grid cell # and center lat lon for each grid cell 
neus_spring_grid_cells <- unique(neus_spring.reduced.again[,.(cell, LON_CENTER, LAT_CENTER)])

neus_spring_grid_cells <- setorder(neus_spring_grid_cells, cell)

neus_spring_reg_distances <- distm(neus_spring_grid_cells[,.(LON_CENTER, LAT_CENTER)])

colnames(neus_spring_reg_distances) <- neus_spring_grid_cells$cell
rownames(neus_spring_reg_distances) <- neus_spring_grid_cells$cell

#reorient to long form 
neus_spring_reg_distances.l <- melt(as.matrix(neus_spring_reg_distances), varnames = c("cell1", "cell2"), value.name = "distance(m)") #matrix to data frame
neus_spring_reg_distances.l <- data.table(neus_spring_reg_distances.l) #and then to data table
neus_spring_reg_distances.l <- neus_spring_reg_distances.l[cell1 > cell2,] # get rid of repetitions and self-comparisons
saveRDS(neus_spring_reg_distances.l, "neus_spring_reg_distances.l.rds")
```

CHECK CLAIM THAT #no years have way more tows than others, because this isn't true

Check the # of tows per grid cell/year combo
```{r number of tows per grid cell/year/region combo for spring}
#cell/year/region combos
neus_spring_unique_tows_cell <- neus_spring.reduced[,.SD[1],.(year, cell, haul_id)]

#no years have way more tows than others
plot(neus_spring_unique_tows_cell[,.N,.(year,cell)]$year, neus_spring_unique_tows_cell[,.N,.(year,cell)]$N)

#therefore, I will not leave out any years, but rather just pull out cells in years they are undersampled (<N=3, which is )

summary(neus_spring_unique_tows_cell[,.N,.(year,cell)]$N < 3) #this occurs in 17/544 cell/year combos  (3%%)

#I don't think I need to exclude in ALL years, just in a few years
#an alternative would be to assess coverage using biomass (# of singletons and doubletons)

```

Delete all observations with less than 3 tows in a given year/cell
```{r delete observations with less than 3 tows spring}
neus_spring_counts <- neus_spring_unique_tows_cell[,.N,.(year,cell)]
neus_spring_keep <- neus_spring_counts[N>2,][,N := NULL]

neus_spring_grid.reduced_3plustows <- neus_spring.reduced[neus_spring_keep, on = .(cell, year)]


```


Save
```{r save spring}
saveRDS(neus_spring.reduced, "/Users/zoekitchel/Documents/grad school/Rutgers/Repositories/trawl_spatial_turnover/data/gridded/USA-NEUS/neus_spring.reduced.rds")


saveRDS(neus_spring_reg_distances.l, "/Users/zoekitchel/Documents/grad school/Rutgers/Repositories/trawl_spatial_turnover/data/gridded/USA-NEUS/neus_spring_reg_distances.l.rds")

saveRDS(neus_spring_grid.reduced_3plustows,"/Users/zoekitchel/Documents/grad school/Rutgers/Repositories/trawl_spatial_turnover/data/gridded/USA-NEUS/neus_spring_grid.reduced_3plustows.rds")


```