---
title: "Prepping SHELF Data"
output: html_notebook
---


```{r setup}
library(dggridR)
library(data.table)
library(rgdal)
library(raster)
library(sp)
library(rnaturalearth)
library(rnaturalearthdata)
library(rgeos)
library(taxize) #standardizing names
library(geosphere)  #to calculate distance between lat lon of grid cells
library(googledrive)
```


Moving forward, I plan to use data directly from FishGlob.

November 21, 2021
```{r pull in data from fishglob for Scotian Shelf from Google Drive}
drive_download("SCS_clean.csv",
               overwrite = T)

shelf <- fread("SCS_clean.csv")

file.remove("SCS_clean.csv")

```

```{r pull in SHELF}

# Get Unique lines in the data table
unique_shelf_latlon<- unique(shelf[,.(latitude, longitude)])
unique_shelf_latlon.sf <- unique_shelf_latlon
  
#make lat lon coordinates into polygon
coordinates(unique_shelf_latlon.sf) <- ~latitude + longitude
unique_shelf_latlon.sf@proj4string <- crs("+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0")

saveRDS(unique_shelf_latlon, "unique_shelf_latlon.RData")

#plot north america
world <- ne_countries(scale = "medium", returnclass = "sp")

world_ext <- extent(-180, -51.5,23, 67)

north_america_spdf <- crop(world, world_ext)

north_america_df <- fortify(north_america_spdf)

(na_outline <- ggplot(north_america_df, aes(long, lat, group = group)) +
  geom_polygon(color = "white", fill = "grey") +
  labs(x = "Longitude", y = "Latitude") +
  coord_equal() +
  theme_classic() +
  scale_x_continuous(expand = c(0,0)) +
  scale_y_continuous(expand = c(0,0)) +
  theme(panel.border = element_rect(color = "black", fill = NA, size = 1)) +
 geom_point(unique_shelf_latlon, mapping = aes(longitude, latitude), inherit.aes = FALSE, shape = 18, size = 0.001, color = "red") +
  coord_equal())

```

Playing around with dggridR (package that makes grid cells)
```{r learning dggridR}
dggs <- dgconstruct(spacing = 111, metric = T, resround = 'nearest')

unique_shelf_latlon[,cell := dgGEO_to_SEQNUM(dggs, longitude, latitude)] #get corresponding grid cells for shelf trawl

#centersof cells
shelf_cellcenters <- dgSEQNUM_to_GEO(dggs, unique_shelf_latlon[,cell])

#linking cell centers to unique_shelf_latlon
unique_shelf_latlon[,LON_CENTER := shelf_cellcenters$lon_deg][,LAT_CENTER := shelf_cellcenters$lat_deg]

#link centers back to main data table

shelf <- merge(shelf, unique_shelf_latlon, by = c("latitude", "longitude"))

#number of tows in each cell
towcount <- unique_shelf_latlon[, .N, by = cell]

#get the grid cell boundary for cells which had trawls
grid_shelf <- dgcellstogrid(dggs, unique_shelf_latlon[,cell], frame = T, wrapcells = F)

#update grid properties to include # of trawls in each cell
grid_shelf <- merge(grid, unique_shelf_latlon, by = "cell")

ggplot(data = grid_shelf, aes(x = long, y = lat, group = cell)) +
  geom_polygon() +
  coord_quickmap()

saveRDS(grid_shelf, "grid_shelf.rds")

```

Trying to plot with grid overlaying
```{r grid overlaying}

na_outline + 
   geom_polygon(grid_shelf, mapping = aes(x = long,y = lat, group = cell), inherit.aes = FALSE, color = "black", fill = NA, size = 0.1)

ggsave(filename = "grid_SHELF_cells_spacing_111.pdf")
  
```

Now, we'll overlay grid cells onto this list of tows (shelf) using lat and lon and gridrr package
```{r get rid of underused grid cells}

year_grid <- data.table(table(shelf$cell, shelf$year))

#there are some grid cells with infrequent records, I will delete these
grid_cells_to_exclude <- unique(year_grid[N == 0]$V1)

#delete these grid cells from full data table
shelf.reduced <- shelf[!cell %in% grid_cells_to_exclude]

#270643 observations left, this should be great

#how many tows (TOW) per cell in a given year

haul_cell_unique <- unique(shelf.reduced[,.(haul_id, cell, year)])

shelf_samplingevents_byyear <- haul_cell_unique[,.N, .(year, cell)]

shelf_cells_lessthan3 <- shelf_samplingevents_byyear[N<3,]

cells_to_delete <- unique(shelf_cells_lessthan3[,cell])

shelf.reduced.again <- shelf.reduced[!cell %in% cells_to_delete]

#down to 114793 observations

saveRDS(shelf.reduced.again, "shelf_data_gridded.rds")

```

Find distance between center of each grid cell
```{r distance between grid cells}
#just need grid cell # and center lat lon for each grid cell 
shelf_grid_cells <- unique(shelf.reduced.again[,.(cell, LON_CENTER, LAT_CENTER)])

shelf_grid_cells <- setorder(shelf_grid_cells, cell)

shelf_reg_distances <- distm(shelf_grid_cells[,.(LON_CENTER, LAT_CENTER)])

colnames(shelf_reg_distances) <- shelf_grid_cells$cell
rownames(shelf_reg_distances) <- shelf_grid_cells$cell

#reorient to long form 
shelf_reg_distances.l <- melt(as.matrix(shelf_reg_distances), varnames = c("cell1", "cell2"), value.name = "distance(m)") #matrix to data frame
shelf_reg_distances.l <- data.table(shelf_reg_distances.l) #and then to data table
shelf_reg_distances.l <- shelf_reg_distances.l[cell1 > cell2,] # get rid of repetitions and self-comparisons
saveRDS(shelf_reg_distances.l, "shelf_reg_distances.l.rds")
```

Check the # of tows per grid cell/year combo
```{r number of tows per grid cell/year/region combo}
#cell/year/region combos
shelf_unique_tows_cell <- shelf.reduced[,.SD[1],.(year, cell, haul_id)]

#no years have way more tows than others
plot(shelf_unique_tows_cell[,.N,.(year,cell)]$year, shelf_unique_tows_cell[,.N,.(year,cell)]$N)

#therefore, I will not leave out any years, but rather just pull out cells in years they are undersampled (<N=3, which is )

summary(shelf_unique_tows_cell[,.N,.(year,cell)]$N < 3) #this occurs in 54/558 cell/year combos  (9.7%)

#I don't think I need to exclude in ALL years, just in a few years
#an alternative would be to assess coverage using biomass (# of singletons and doubletons)

```

Delete all observations with less than 3 tows in a given year/cell
```{r delete observations with less than 3 tows}
shelf_counts <- shelf_unique_tows_cell[,.N,.(year,cell)]
shelf_keep <- shelf_counts[N>2,][,N := NULL]

shelf_grid.reduced_3plustows <- shelf.reduced[shelf_keep, on = .(cell, year)]


```

Summary of area sampled
```{r sum area sampled}
#distance between furthest north and furthest south point
library(geosphere)
max_north_south_km <- (distm(c(mean(shelf_grid.reduced_3plustows$longitude), min(shelf_grid.reduced_3plustows$latitude)), c(mean(shelf_grid.reduced_3plustows$longitude), max(shelf_grid.reduced_3plustows$latitude)), fun = distHaversine))/1000
#distance between furthest east and furthest west point
max_west_east_km <- (distm(c(min(shelf_grid.reduced_3plustows$longitude), mean(shelf_grid.reduced_3plustows$latitude)), c(max(shelf_grid.reduced_3plustows$longitude), mean(shelf_grid.reduced_3plustows$latitude)), fun = distHaversine))/1000

max_north_south_km
max_west_east_km

summary(shelf_grid.reduced_3plustows$depth)

length(unique(shelf_grid.reduced_3plustows$accepted_name))
```

Save
```{r save}
saveRDS(shelf.reduced, "/Users/zoekitchel/Documents/grad school/Rutgers/Repositories/trawl_spatial_turnover/data/gridded/USA-SHELF/shelf_grid.reduced.rds")


saveRDS(shelf_reg_distances.l, "/Users/zoekitchel/Documents/grad school/Rutgers/Repositories/trawl_spatial_turnover/data/gridded/USA-SHELF/shelf_reg_distances.l.rds")

saveRDS(shelf_grid.reduced_3plustows,"/Users/zoekitchel/Documents/grad school/Rutgers/Repositories/trawl_spatial_turnover/data/gridded/USA-SHELF/shelf_grid.reduced_3plustows.rds")


```
