---
title: "Prepping WCANN Data"
output: html_notebook
---


```{r setup}
#if not already installed
#install.packages("/Users/zoekitchel/Downloads/dggridR", 
#                 repos = NULL, 
#                 type = "source")
library(dggridR)
library(data.table)
library(rgdal)
library(raster)
library(sp)
library(rnaturalearth)
library(rnaturalearthdata)
library(rgeos)
library(taxize) #standardizing names
library(geosphere)  #to calculate distance between lat lon of grid cells
library(googledrive)
```


Moving forward, I plan to use data directly from FishGlob.

November 21, 2021
```{r pull in data from fishglob for wcann from Google Drive}
drive_download("WCANN_clean.csv",
               overwrite = T)

wcann <- fread("WCANN_clean.csv")

file.remove("WCANN_clean.csv")

```

```{r pull in WCANN}

# Get Unique lines in the data table
unique_wcann_latlon<- unique(wcann[,.(latitude, longitude)])
unique_wcann_latlon.sf <- unique_wcann_latlon
  
#make lat lon coordinates into polygon
coordinates(unique_wcann_latlon.sf) <- ~latitude + longitude
unique_wcann_latlon.sf@proj4string <- crs("+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0")

saveRDS(unique_wcann_latlon, "unique_wcann_latlon.RData")

#plot north america
world <- ne_countries(scale = "medium", returnclass = "sp")

world_ext <- extent(-180, -51.5,23, 67)

north_america_spdf <- crop(world, world_ext)

north_america_df <- fortify(north_america_spdf)

(na_outline <- ggplot(north_america_df, aes(long, lat, group = group)) +
  geom_polygon(color = "white", fill = "grey") +
  labs(x = "Longitude", y = "Latitude") +
  coord_equal() +
  theme_classic() +
  scale_x_continuous(expand = c(0,0)) +
  scale_y_continuous(expand = c(0,0)) +
  theme(panel.border = element_rect(color = "black", fill = NA, size = 1)) +
 geom_point(unique_wcann_latlon, mapping = aes(longitude, latitude), inherit.aes = FALSE, shape = 18, size = 0.001, color = "red") +
  coord_equal())

```

Playing around with dggridR (package that makes grid cells)
```{r learning dggridR}
dggs <- dgconstruct(spacing = 111, metric = T, resround = 'nearest')

unique_wcann_latlon[,cell := dgGEO_to_SEQNUM(dggs, longitude, latitude)] #get corresponding grid cells for NEUS trawl

#centersof cells
wcann_cellcenters <- dgSEQNUM_to_GEO(dggs, unique_wcann_latlon[,cell])

#linking cell centers to unique_wcann_latlon
unique_wcann_latlon[,LON_CENTER := wcann_cellcenters$lon_deg][,LAT_CENTER := wcann_cellcenters$lat_deg]

#link centers back to main data table

wcann <- merge(wcann, unique_wcann_latlon, by = c("latitude", "longitude"))

#number of tows in each cell
towcount <- unique_wcann_latlon[, .N, by = cell]

#get the grid cell boundary for cells which had trawls
grid_wcann <- dgcellstogrid(dggs, unique_wcann_latlon[,cell], frame = T, wrapcells = F)

#update grid properties to include # of trawls in each cell
grid_wcann <- merge(grid_wcann, unique_wcann_latlon, by = "cell")

ggplot(data = grid_wcann, aes(x = long, y = lat, group = cell)) +
  geom_polygon() +
  coord_quickmap()

saveRDS(grid_wcann, "grid_wcann.rds")

```

Trying to plot with grid overlaying
```{r grid overlaying}

na_outline + 
   geom_polygon(grid_wcann, mapping = aes(x = long,y = lat, group = cell), inherit.aes = FALSE, color = "black", fill = NA, size = 0.1)

ggsave(filename = "grid_WCANN_cells_spacing_111.pdf")
  
````

Now, we'll overlay grid cells onto this list of tows (wcann) using lat and lon and gridrr package

```{r get rid of underused grid cells}

year_grid <- data.table(table(wcann$cell, wcann$year))

#there are some grid cells with infrequent records, I will delete these
grid_cells_to_exclude <- unique(year_grid[N == 0]$V1)

#delete these grid cells from full data table
wcann.reduced <- wcann[!cell %in% grid_cells_to_exclude]

#169766 observations left, this should be great

#how many tows (TOW) per cell in a given year

haul_cell_unique <- unique(wcann.reduced[,.(haul_id, cell, year)])

wcann_samplingevents_byyear <- haul_cell_unique[,.N, .(year, cell)]

wcann_cells_lessthan3 <- wcann_samplingevents_byyear[N<3,]

cells_to_delete <- unique(wcann_cells_lessthan3[,cell])

wcann.reduced.again <- wcann.reduced[!cell %in% cells_to_delete]

#down to 161508 observations

saveRDS(wcann.reduced.again, "wcann_data_gridded.rds")

```

Find distance between center of each grid cell
```{r distance between grid cells}
#just need grid cell # and center lat lon for each grid cell 
wcann_grid_cells <- unique(wcann.reduced.again[,.(cell, LON_CENTER, LAT_CENTER)])

wcann_grid_cells <- setorder(wcann_grid_cells, cell)

wcann_reg_distances <- distm(wcann_grid_cells[,.(LON_CENTER, LAT_CENTER)])

colnames(wcann_reg_distances) <- wcann_grid_cells$cell
rownames(wcann_reg_distances) <- wcann_grid_cells$cell

#reorient to long form 
wcann_reg_distances.l <- melt(as.matrix(wcann_reg_distances), varnames = c("cell1", "cell2"), value.name = "distance(m)") #matrix to data frame
wcann_reg_distances.l <- data.table(wcann_reg_distances.l) #and then to data table
wcann_reg_distances.l <- wcann_reg_distances.l[cell1 > cell2,] # get rid of repetitions and self-comparisons
saveRDS(wcann_reg_distances.l, "wcann_reg_distances.l.rds")
```

Check the # of tows per grid cell/year combo
```{r number of tows per grid cell/year/region combo}
#cell/year/region combos
wcann_unique_tows_cell <- wcann.reduced[,.SD[1],.(year, cell, haul_id)]

#no years have way more tows than others
plot(wcann_unique_tows_cell[,.N,.(year,cell)]$year, wcann_unique_tows_cell[,.N,.(year,cell)]$N)

#therefore, I will not leave out any years, but rather just pull out cells in years they are undersampled (<N=3, which is )

summary(wcann_unique_tows_cell[,.N,.(year,cell)]$N < 3) #this occurs in 14/309 cell/year combos  (4.5%, no biggie)

#I don't think I need to exclude in ALL years, just in a few years
#an alternative would be to assess coverage using biomass (# of singletons and doubletons)

```

Delete all observations with less than 3 tows in a given year/cell
```{r delete observations with less than 3 tows}
wcann_counts <- wcann_unique_tows_cell[,.N,.(year,cell)]
wcann_keep <- wcann_counts[N>2,][,N := NULL]

wcann_grid.reduced_3plustows <- wcann.reduced[wcann_keep, on = .(cell, year)]


```


Save
```{r save}
saveRDS(wcann.reduced, "/Users/zoekitchel/Documents/grad school/Rutgers/Repositories/trawl_spatial_turnover/data/gridded/USA-WCANN/dat_WCANN_grid.reduced.rds")


saveRDS(wcann_reg_distances.l, "/Users/zoekitchel/Documents/grad school/Rutgers/Repositories/trawl_spatial_turnover/data/gridded/USA-WCANN/wcann_reg_distances.l.rds")

saveRDS(wcann_grid.reduced_3plustows,"/Users/zoekitchel/Documents/grad school/Rutgers/Repositories/trawl_spatial_turnover/data/gridded/USA-WCANN/dat_WCANN_grid.reduced_3plustows.rds")


```
