---
title: "iNEXT NEUS Richness"
output: html_notebook
---

I will plot how richness changes over time but will use iNEXT to calculate richness

```{r setup}
library(data.table)
library(iNEXT)
```

Abundance Data
```{r abundance data}
neus_data <- readRDS("neus_data_gridded.RData")

head(neus_data)
```
First, we will plot richness for the entire region (ignoring cells)

Format as data frame with species as rows and years as columns , "wtcpue - numeric - the weight in kilograms captured per unit effort"

```{r sum across years}
neus_data_year <- neus_data[,wtcpue_sum := sum(wtcpue, na.rm = T), .(matched_name2, year)][,wtcpue_sum_round := round(wtcpue_sum, 0)]

neus_data_year.r <- unique(neus_data_year[,.(matched_name2, year, wtcpue_sum_round)])
```
I want to calculate richness for each individual year (site)
```{r all years x abundances}

#list of years 
neus_years <- unique(neus_data_year.r[,year])
sort(neus_years)

#long to wide
neus_wide <- dcast(neus_data_year.r, matched_name2 ~ year, value.var = "wtcpue_sum_round", fill = 0)

```

Apply iNEXT. iNEXT will literally only take data frames, so, I will make it into a data frame first.
```{r applying iNEXT to overall richness}
neus_wide.df <- data.frame(neus_wide)

neus_overall_biomass_iNEXT_richness <- iNEXT(neus_wide.df[,2:length(neus_wide.df)], q = 0, datatype = "abundance")
neus_overall_biomass_iNEXT_shannon <- iNEXT(neus_wide.df[,2:length(neus_wide.df)], q = 1, datatype = "abundance") #how to deal with the fact abundance should be a round number


neus_overall_biomass_iNEXT_simpson <- iNEXT(neus_wide.df[,2:length(neus_wide.df)], q = 2, datatype = "abundance")
saveRDS(neus_overall_biomass_iNEXT_simpson, "neus_overall_biomass_iNEXT_simpson.RData")
# to assess sample completeness (this takes a bit), iNEXT focuses on three measures of Hill numbers of order q: species richness (q = 0), Shannon diversity (q = 1, the exponential of Shannon entropy) and Simpson diversity (q = 2, the inverse of Simpson concentration), I should look at all three

neus_overall_iNEXT_min_coverage_richness <- min(neus_overall_biomass_iNEXT_richness$DataInfo$SC)
neus_overall_iNEXT_min_coverage_shannon <- min(neus_overall_biomass_iNEXT_shannon$DataInfo$SC)
neus_overall_iNEXT_min_coverage_simpson <- min(neus_overall_biomass_iNEXT_simpson$DataInfo$SC)

#minimum sample coverage across sites is 0.9998, so this will be my SC value

#this step below takes a long time as well... keep in mind

neus_overall_diversity_richness <- estimateD(neus_wide.df[,2:length(neus_wide.df)], datatype = "abundance", base = "coverage", level = neus_overall_iNEXT_min_coverage_richness) # compare point diversities for specified completeness

saveRDS(neus_overall_diversity_richness, file = "neus_overall_diversity_richness.RData")

neus_overall_diversity_shannon <- estimateD(neus_wide.df[,2:length(neus_wide.df)], datatype = "abundance", base = "coverage", level = neus_overall_iNEXT_min_coverage_shannon) # compare point diversities for specified completeness

saveRDS(neus_overall_diversity_shannon, file = "neus_overall_diversity_shannon.RData")

neus_overall_diversity_simpson <- estimateD(neus_wide.df[,2:length(neus_wide.df)], datatype = "abundance", base = "coverage", level = neus_overall_iNEXT_min_coverage_simpson) # compare point diversities for specified completeness

saveRDS(neus_overall_diversity_simpson, file = "neus_overall_diversity_simpson.RData")
#only thing that would change above is coverage estimates (q = 0-2)

#convert to data table to make easier to manipulate
neus_overall_diversity_richness.dt <- data.table(neus_overall_diversity_richness)
neus_overall_diversity_shannon.dt <- data.table(neus_overall_diversity_shannon)
neus_overall_diversity_simpson.dt <- data.table(neus_overall_diversity_simpson)

neus_overall_diversity_richness.dt[,year := as.numeric(substr(site, 2, 5))][,metric_type := "richness"]
neus_overall_diversity_shannon.dt[,year := as.numeric(substr(site, 2, 5))][,metric_type := "shannon"]
neus_overall_diversity_simpson.dt[,year := as.numeric(substr(site, 2, 5))][,metric_type := "simpson"]

#... turns out it calculates all three metrics (hill numbers) for you anyway... no need to do three times, cool bro

#rarefied richness overtime plot
ggplot(aes(x = year, y = qD), data = neus_overall_diversity_richness.dt[order == 0,]) +
  geom_smooth(aes(ymin = qD.LCL, ymax = qD.UCL), stat = "identity", color = "black") +
  labs(x= "Year", y = "Estimated Richness") +
  theme_classic() +
  theme(text = element_text(size=20))

#simpson and shannon diversity plot
ggplot(aes(x = year, y = qD, linetype = as.factor(order)), data = neus_overall_diversity_richness.dt[order != 0,]) +
  geom_smooth(aes(ymin = qD.LCL, ymax = qD.UCL), stat = "identity", color = "black", alpha = 3) +
  labs(x= "Year", y = "Diversity Metric") +
  theme_classic() +
  theme(text = element_text(size=20))


```












I need it to be a data frame formatted species (rows) by sites/years (columns)

```{r sum across grid cells and years}
neus_data_year_cell <- neus_data[,abundance_sum := sum(ABUNDANCE, na.rm = T), .(matched_name2, cell, YEAR)]

neus_data_year_cell.r <- unique(neus_data_year_cell[,.(matched_name2, cell, YEAR, abundance_sum)])
```
Let's start with one grid cell for species by year data frame for each grid cel

```{r start with one cell }

#list of grid cells 
neus_grid_cells <- unique(neus_data_year_cell.r[,cell])

i = 1

neus_subset <- neus_data_year_cell.r[cell == neus_grid_cells[i],]

#long to wide
neus_subset_wide <- dcast(neus_subset, matched_name2 ~ YEAR, value.var = "abundance_sum", fill = 0)

```

Apply iNEXT. iNEXT will literally only take data frames, so, I will make it into a data frame first.
```{r applying iNEXT}
neus_subset_wide.df <- data.frame(neus_subset_wide)

neus_abundance_iNEXT <- iNEXT(neus_subset_wide.df[,2:length(neus_subset_wide.df)], q = 0, datatype = "abundance") # to assess sample completeness (this takes a bit)

neus_abundance_iNEXT_min_coverage <- min(neus_abundance_iNEXT$DataInfo$SC)

#minimum sample coverage across sites is 0.9985, so this will be my SC value

neus_abundance_diversity_estimate <- estimateD(neus_subset_wide.df[,2:length(neus_subset_wide.df)], datatype = "abundance", base = "coverage", level = neus_abundance_iNEXT_min_coverage) # compare point diversities for specified completeness

neus_abundance_diversity_estimate.dt <- data.table(neus_abundance_diversity_estimate)

neus_abundance_diversity_estimate.dt[,YEAR := as.numeric(substr(site, 2, 5))]

ggplot(aes(x = YEAR, y = qD),data = neus_abundance_diversity_estimate.dt[order == "0",]) + #order = 0 = richness
  geom_smooth(aes(ymin = qD.LCL, ymax = qD.UCL), stat = "identity") +
  labs(x= "Year", y = "Estimated Richness", title = paste0("Cell: ", neus_grid_cells[i])) +
  geom_line() +
  theme_classic()


```
Loop to make plot for all cells using abundance
```{r}

```



Incidence Data (turn values into 1s and 0s)
```{r}
neus_subset[,abundance_incidence := if(abundance_sum >= 1, 1, 0)]

#long to wide
neus_subset_wide.incidence <- dcast(neus_subset, matched_name2 ~ cell, value.var = "abundance_incidence", fill = 0) 

```

Apply iNEXT. iNEXT will literally only take data frames, so, I will make it into a data frame first.
```{r applying iNEXT for incidence}
neus_subset_wide.incidence.df <- data.frame(neus_subset_wide.incidence)

neus_incidence_iNEXT <- iNEXT(neus_subset_wide.incidence.df[,2:length(neus_subset_wide.df)], q = 0, datatype = "abundance") # to assess sample completeness (this takes a bit)

#minimum sample coverage across sites is 0.9991, so this will be my SC value

neus_abundance_diversity_estimate <- estimateD(neus_subset_wide.df[,2:length(neus_subset_wide.df)], datatype = "abundance", base = "coverage", level = 0.9991) # compare point diversities for specified completeness
#this takes quite a while to run



```
