---
title: "Prepping Portugal (ICES) from DATRAS data"
output: html_notebook
---


```{r setup}
library(dggridR)
library(data.table)
library(rgdal)
library(raster)
library(sp)
library(rnaturalearth)
library(rnaturalearthdata)
library(rgeos)
library(taxize) #standardizing names
library(geosphere)  #to calculate distance between lat lon of grid cells
library(googledrive)
```


Moving forward, I plan to use data directly from FishGlob.

December 1, 2021
```{r pull in data from fishglob for PT from Google Drive}
drive_download("DATRAS_v2_clean.csv",
               overwrite = T)

DATRAS <- fread("DATRAS_v2_clean.csv")

file.remove("DATRAS_v2_clean.csv")

pt <- DATRAS[survey  == "PT-IBTS"]

```

What seasopt do we have? None

```{r seasopt?}
summary(as.factor(pt$season))


```

#PT
```{r pull in PT}

# Get Unique lines in the data table
unique_pt_latlon <- unique(pt[,.(latitude, longitude)])
unique_pt_latlon.sf <- unique_pt_latlon
  
#make lat lon coordinates into polygon
coordinates(unique_pt_latlon.sf) <- ~latitude + longitude
unique_pt_latlon.sf@proj4string <- crs("+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0")

saveRDS(unique_pt_latlon, "unique_pt_latlon.rds")

#plot north america
world <- ne_countries(scale = "medium", returnclass = "sp")

world_ext <- extent(-12, -5,36, 43)

europe_spdf <- crop(world, world_ext)

europe_df <- fortify(europe_spdf)

(europe_outline <- ggplot(europe_df, aes(long, lat, group = group)) +
  geom_polygon(color = "white", fill = "grey") +
  labs(x = "Longitude", y = "Latitude") +
  coord_equal() +
  theme_classic() +
  scale_x_continuous(expand = c(0,0)) +
  scale_y_continuous(expand = c(0,0)) +
  theme(panel.border = element_rect(color = "black", fill = NA, size = 1)) +
 geom_point(unique_pt_latlon, mapping = aes(longitude, latitude), inherit.aes = FALSE, shape = 18, size = 0.001, color = "red") +
  coord_equal())

```

Playing around with dggridR (package that makes grid cells)
```{r learning dggridR}
dggs <- dgconstruct(spacing = 111, metric = T, resround = 'nearest')

unique_pt_latlon[,cell := dgGEO_to_SEQNUM(dggs, longitude, latitude)] #get corresponding grid cells for NEUS trawl

#centersof cells
pt_cellcenters <- dgSEQNUM_to_GEO(dggs, unique_pt_latlon[,cell])

#linking cell centers to unique_pt_latlon
unique_pt_latlon[,LON_CENTER := pt_cellcenters$lon_deg][,LAT_CENTER := pt_cellcenters$lat_deg]

#link centers back to main data table

pt <- merge(pt, unique_pt_latlon, by = c("latitude", "longitude"))

#number of tows in each cell
towcount <- unique_pt_latlon[, .N, by = cell]

#get the grid cell boundary for cells which had trawls
grid_pt <- dgcellstogrid(dggs, unique_pt_latlon[,cell], frame = T, wrapcells = F)

#update grid properties to include # of trawls in each cell
grid_pt <- merge(grid_pt, unique_pt_latlon, by = "cell")

ggplot(data = grid_pt, aes(x = long, y = lat, group = cell)) +
  geom_polygon() +
  coord_quickmap()

saveRDS(grid_pt, "grid_pt.rds")

```

Trying to plot with grid overlaying
```{r grid overlaying}

europe_outline + 
   geom_polygon(grid_pt, mapping = aes(x = long,y = lat, group = cell), inherit.aes = FALSE, color = "black", fill = NA, size = 0.1)

ggsave(filename = "grid_PT_cells_spacing_111.pdf")
  
```
Cleaning spp names (must be ID'd to species!)
```{r clean PT names to species}
#unique species names
pt.clean <- pt[rank == "Species"]


```


Now, we'll overlay grid cells onto this list of tows (dat_pt) using lat and lon and gridrr package
```{r get rid of underused grid cells}

year_grid <- data.table(table(pt.clean$cell, pt.clean$year))

#there are some grid cells with infrequent records, I will delete these BUT IF I DELETE THESE THEY ALL DELETE, SO I WILL START TIME SERIES IN 1983 FIRST
#year_grid <- year_grid[V2 >= 1983,] #Juliette Murgier, Matthew McLean, Anthony Maire, David Mouillot, Nicolas Loiseau, François Munoz, Cyrille Violle and Arnaud Auber

grid_cells_to_exclude <- unique(year_grid[N == 0]$V1)

#delete these grid cells from full data table
pt.reduced <- pt.clean[!cell %in% grid_cells_to_exclude]

#8040 observatiopt left, this should be great

#how many tows (TOW) per cell in a given year

haul_cell_unique <- unique(pt.reduced[,.(haul_id, cell, year)])

pt_samplingevents_byyear <- haul_cell_unique[,.N, .(year, cell)]

pt_cells_lessthan3 <- pt_samplingevents_byyear[N<3,]

cells_to_delete <- unique(pt_cells_lessthan3[,cell])

pt.reduced.again <- pt.reduced[!cell %in% cells_to_delete]

#down to 7180 observatiopt

saveRDS(pt.reduced.again, "pt.reduced.again.rds")

```

Find distance between center of each grid cell
```{r distance between grid cells}
#just need grid cell # and center lat lon for each grid cell 
pt_grid_cells <- unique(pt.reduced.again[,.(cell, LON_CENTER, LAT_CENTER)])

pt_grid_cells <- setorder(pt_grid_cells, cell)

pt_reg_distances <- distm(pt_grid_cells[,.(LON_CENTER, LAT_CENTER)])

colnames(pt_reg_distances) <- pt_grid_cells$cell
rownames(pt_reg_distances) <- pt_grid_cells$cell

#reorient to long form 
pt_reg_distances.l <- melt(as.matrix(pt_reg_distances), varnames = c("cell1", "cell2"), value.name = "distance(m)") #matrix to data frame
pt_reg_distances.l <- data.table(pt_reg_distances.l) #and then to data table
pt_reg_distances.l <- pt_reg_distances.l[cell1 > cell2,] # get rid of repetitiopt and self-comparisopt
saveRDS(pt_reg_distances.l, "pt_reg_distances.l.rds")
```

CHECK CLAIM THAT #no years have way more tows than others, because this isn't true

Check the # of tows per grid cell/year combo
```{r number of tows per grid cell/year/region combo for}
#cell/year/region combos
pt_unique_tows_cell <- pt.reduced[,.SD[1],.(year, cell, haul_id)]

#no years have way more tows than others
plot(pt_unique_tows_cell[,.N,.(year,cell)]$year, pt_unique_tows_cell[,.N,.(year,cell)]$N)

#therefore, I will not leave out any years, but rather just pull out cells in years they are undersampled (<N=3, which is )

summary(pt_unique_tows_cell[,.N,.(year,cell)]$N < 3) #this occurs in 4/108 cell/year combos  (3.7%)

#I don't think I need to exclude in ALL years, just in a few years
#an alternative would be to assess coverage using biomass (# of singletopt and doubletopt)

```

Delete all observatiopt with less than 3 tows in a given year/cell AFTER 1990
```{r delete observatiopt with less than 3 tows}
pt_counts <- pt_unique_tows_cell[,.N,.(year,cell)]
pt_keep <- pt_counts[N>2,][,N := NULL]

pt_grid.reduced_3plustows <- pt.reduced[pt_keep, on = .(cell, year)]

#[year >= 1983]


```

Make a new map but with only points we actually use
```{r new map points we use}
(europe_outline_2 <- ggplot(europe_df, aes(long, lat, group = group)) +
  geom_polygon(color = "white", fill = "grey") +
  labs(x = "Longitude", y = "Latitude") +
  coord_equal() +
  theme_classic() +
  scale_x_continuous(expand = c(0,0)) +
  scale_y_continuous(expand = c(0,0)) +
  theme(panel.border = element_rect(color = "black", fill = NA, size = 1)) +
 geom_point(unique(pt_grid.reduced_3plustows[,.(latitude, longitude)]), mapping = aes(longitude, latitude), inherit.aes = FALSE, shape = 18, size = 0.001, color = "red") +
  coord_equal())
```


Save
```{r save}
saveRDS(pt.reduced, "/Users/zoekitchel/Documents/grad school/Rutgers/Repositories/trawl_spatial_turnover/data/gridded/ICES-PT/pt.reduced.rds")


saveRDS(pt_reg_distances.l, "/Users/zoekitchel/Documents/grad school/Rutgers/Repositories/trawl_spatial_turnover/data/gridded/ICES-PT/pt_reg_distances.l.rds")

saveRDS(pt_grid.reduced_3plustows,"/Users/zoekitchel/Documents/grad school/Rutgers/Repositories/trawl_spatial_turnover/data/gridded/ICES-PT/pt_grid.reduced_3plustows.rds")


```
