---
title: "3. Spatial Temporal Diversity By Strata"
output:
  html_document:
    df_print: paged
---
Here, we will do a similar analysis but by strata instead of subregion
```{r setup}
library(data.table)
library(ggplot2)
library(geosphere)
library(betapart)
library(vegan)
```

Data
```{r data}
load("data/NE_trawl_data_full.Rdata")
```

#Distance Decay with Stratum instead of region
Jaccard Dissimilarities Distance Decay all in one plot by STRATA
```{r stratum distance matrix}
neus_full.ID.uniquestrata <- unique(neus_full.ID, by = "STRATUM") #unique strata

neus_full.ID.uniquestrata.latlon <- neus_full.ID.uniquestrata[,c("LAT", "LON", "STRATUM")] #only lat lon and stratum

stratum_distances <- distm(neus_full.ID.uniquestrata.latlon[,c(2,1)]) #distance matrix, consistent across years
rownames(stratum_distances) <- neus_full.ID.uniquestrata.latlon$STRATUM
colnames(stratum_distances) <- neus_full.ID.uniquestrata.latlon$STRATUM

#reorient to long form 
stratum_distance.l <- melt(as.matrix(stratum_distances), varnames = c("strat1", "strat2"), value.name = "distance(m)") #matrix to data frame
stratum_distance.l <- data.table(stratum_distance.l) #and then to data table
stratum_distance.l <- stratum_distance.l[strat1 > strat2,] # get rid of repetitions and self-comparisons

```

Above is distance info, now I need to build community matrix
```{r list of all strata}

#one PRESENT value for each YEAR, SEASON, STRATUM and SVSPP combination
neus_full.ID.stratum.unique <- na.omit(unique(neus_full.ID, by = c("YEAR", "STRATUM", "SVSPP")), cols = c("YEAR", "STRATUM", "SVSPP")) #unique STRATUM, year, species combinations, and get rid of any rows where any are NA

#reduce data table to just be svspp, year, sciname, comname, STRATUM

neus_full.ID.stratum.unique.r <- data.table((neus_full.ID.stratum.unique[, c("SVSPP", "SCINAME", "YEAR", "STRATUM")]))

#add presence column 
neus_full.ID.stratum.unique.r[,"present" := 1]

#now I need to populate a comparison data table with every possible year, region, and species combination to add 0s
stratum <- as.integer(unique(neus_full.ID.stratum.unique.r$STRATUM)) 

# list of all species, years, strata
spp.list <- neus_full.ID.stratum.unique.r[, sort(unique(SVSPP))]
years <- neus_full.ID.stratum.unique.r[, sort(unique(YEAR))]
stratum <- neus_full.ID.stratum.unique.r[, sort(unique(STRATUM))]

full_set_stratum <- data.table(expand.grid("SVSPP" = spp.list, "YEAR" = years, "STRATUM" = stratum))

```
Bring in presences to expansion for presence by STRATUM
```{r bring in actual presences by STRATUM to expansion}
merge <- neus_full.ID.stratum.unique.r[full_set_stratum, on = c("YEAR", "STRATUM", "SVSPP")]
merge[,"present" := ifelse(is.na(SCINAME), 0,1)][, SCINAME := NULL] #if scientific name and common name are empty, that is an absence, if populated, presence, also, we don't need these columns, as we'll bring them in below

#bring in spp key
spp_key <- unique(neus_full.ID[,c("SVSPP", "SCINAME")], col = c("SVSPP", "SCINAME"))

neus_pres_abs_STRATUM <- merge[spp_key, on = c("SVSPP")] 

```

Make distance decay model for each year by loop for all STRATUM, YEAR, SVSPP combos. I've commented out lines that attempt to make a distance decay model because of error. Instead, I will just use this loop to populate a data table with distance and dissimilarities between all strata. I attempt to use decay.model() function from betapart package. 
--> I can't get this to work
Error: *cannot find valid starting values: please specify some*


```{r jaccard disdissimilarities between STRATUM in multiple years}
distances_dissimilarities_allyears_STRATUM <- data.table("strat1" = integer(), "strat2" = integer(), "distance(m)" = numeric(),"jaccard_dissimilarity" = numeric(), YEAR = integer())

#Now loop through all years starting in 1967
for (i in 1:length(years)) {
  reduced_year <- neus_pres_abs_STRATUM[YEAR == years[i],]
  reduced_year_wide <- dcast(reduced_year, YEAR + STRATUM ~ SCINAME, value.var = "present") #long to wide data for community matrix, column names are YEAR, STRATUM, then SVSPP
  
  ncols <- ncol(reduced_year_wide)
  communitymatrix <- cbind(reduced_year_wide[,3:ncols], reduced_year_wide[,1:2]) #community matrix with YEAR and STRATUM on far right
  
  #there are some STRATA with no presences or observations
  #rowSums(communitymatrix.r[,c(1:817)]) #I guess I will get rid of those

  communitymatrix.r <- communitymatrix[as.logical(rowSums(communitymatrix[,1:(ncols - 2)] != 0)), ] #reduce to only rows with atleast one species present
  stratum.r <- communitymatrix.r$STRATUM

  dissimilarities <- vegdist(communitymatrix.r[,1:(ncols-2)], method = "jaccard", binary = T) #dissimilarity 
  ?vegdist

  #make into matrix
  dissimilarities.m <- as.matrix(dissimilarities, labels=TRUE)
  colnames(dissimilarities.m) <- rownames(dissimilarities.m) <- stratum.r

  #reshape dissimilarities
  dissimilarities.l <- melt(dissimilarities.m, varnames = c("strat1", "strat2"), value.name = "jaccard_dissimilarity")
  dissimilarities.l <- data.table(dissimilarities.l) #and then to data table
  dissimilarities.l <- dissimilarities.l[strat1 > strat2,] # to get rid of repetitions and self-comparisons


  #add year for these values
  dissimilarities.l[, "YEAR" := years[i]]

  #merge distance with jaccard_dissimilarity for this year
  dissimilarities.l.full <- stratum_distance.l[dissimilarities.l, on = c("strat1", "strat2")]

  #add to data table
  distances_dissimilarities_allyears_STRATUM <- rbind(distances_dissimilarities_allyears_STRATUM, dissimilarities.l.full)

  #extract info for distance decay exponential model for each year
  #reduce stratum_distances to stratum.r
  #stratum_distances.r <- stratum_distances[as.character(stratum.r), as.character(stratum.r)]

  #modname <- paste0(years[i], "_decay_mod_exponential")

  #assign(modname, decay.model(dissimilarities.m, stratum_distances.r, model.type = "exponential", y.type = "dissimilarities", perm = 100))

  #jpeg(file=paste0("jaccardplot_", years[i], ".jpg"))
  #plot(x=dissimilarities.l.full$`distance(m)`, y=dissimilarities.l.full$jaccard_dissimilarity, xlab = "Distance(m)", #ylab="Jaccard dissimilarity", ylim = c(0:1), pch = 19)
  #dev.off()
 
  #PCA
  #vare.pca <- rda(communitymatrix[,c(1:817)])
  #jpeg(file=paste0("jaccardpca_", years[i], ".jpg"))
  #plot(vare.pca, display = "sites", xlim = c(-4,4), ylim = c(-4,4))
  #dev.off()
  
}

summary(distances_dissimilarities_allyears_STRATUM) #here we have jaccard and geographic distance

# save(distances_dissimilarities_allyears, file = "output/jaccard_similarities_distances_4reg.Rdata") # from an earlier file
save(distances_dissimilarities_allyears_STRATUM, file = "output/jaccard_similarities_distances_stratum.Rdata")

```

Because I can't get code to run for distance decay, I will look at distance decay through time visually

```{r distance decay all years}
# read in dissimilarities for all years, all strata
load('output/jaccard_similarities_distances_stratum.Rdata') # distances_dissimilarities_allyears_STRATUM

#order data table by year
setorder(distances_dissimilarities_allyears_STRATUM, by = YEAR)

distance_decay_stratum <- ggplot(data = distances_dissimilarities_allyears_STRATUM, aes(x = `distance(m)`, y = 1-jaccard_dissimilarity, color = YEAR)) +
  geom_point() +
  theme_classic() +
  labs(x = "Distance (m)", y = "Jaccard Similarity") +
  theme(text=element_text(size = 20))
distance_decay_stratum


ggsave(distance_decay_stratum, path = "plots/", file = "distance_decay_stratum.eps")
ggsave(distance_decay_stratum, path = "plots/", file = "distance_decay_stratum.jpg")


```

Extract coefficient (slope) for each year: how does it change through time? I did this because I couldn't get betapart decay.model to work.
```{r each year build linear model}
#I'm interested in the rate of decrease

mod_coefs <- data.table("year" = years, "beta" = 0)
for (i in 1:length(years)) {
  # Estimate the parameters using a linear model for this year
  mod <- lm(I(1 - jaccard_dissimilarity) ~ `distance(m)`, data = distances_dissimilarities_allyears_STRATUM[YEAR == years[i],])  # I() allows - to be used arithmetically, here we get similarity versus distance models
  mod_coefs[i, beta :=  -coef(mod)[2]] #change in similarity for change in distance

}

ggplot(aes(x = year, y = (beta)), data = mod_coefs) +
  geom_point() +
  labs(x = "Year", y = "Slope of Distance Decay Plot\n(ΔSimilarity/ΔDistance)") +
  geom_smooth() +
  theme_classic() +
  theme(text = element_text(size = 13))

ggsave(path = "plots", file = "slope_distance_decay.jpg", width = 6, height = 4, unit = "in")


#incecreasing rate of decay through time (strata become less similar through time)

```


Below, I continue to attempt to make distance decay from betapart package work, no luck yet. Some years work, others don't

Fitting distance decay models to pair-wise assemblage similarity: fits a negative exponential or power law function(GLM) describing decay of assemblage similarity or increase of dissimilarity with distance.


```{r distance decay model}

#haven't figured out error here, same error that appears above

#reduce stratum_distances to stratum.r
stratum_distances.r <- stratum_distances[as.character(stratum.r), as.character(stratum.r)]

mod_1968 <- decay.model(dissimilarities.m, stratum_distances.r, model.type = "exponential", y.type = "dissimilarities", perm = 1000)


```