---
title: "GOA Distance Decay over Time Tow x Tow"
output: html_notebook
---

```{r setup}
library(data.table)
library(vegan)
library(ggplot2)
library(broom) #augment
library(betapart)
library(geosphere)

#load data
dat_GOA_grid.reduced_3plustows <- readRDS("/Users/zoekitchel/Documents/grad school/Rutgers/Repositories/trawl_spatial_turnover/data/gridded/USA-GOA/dat_GOA_grid.reduced_3plustows.rds")


#distance among grid cells
goa_reg_distances.l <- readRDS("/Users/zoekitchel/Documents/grad school/Rutgers/Repositories/trawl_spatial_turnover/data/gridded/USA-GOA/goa_reg_distances.l.rds")



```

In this iteration of the analysis, each tow will be analyzed independently--I will not be summing or averaging over cells. However, I did use cells in order to eliminate undersampled years/tows.


Distance decay for GOA
```{r each year build exponential model}

#years
#order data table by years
dat_GOA_grid.reduced_3plustows[,year:= as.numeric(year)] #make numeric
setorder(dat_GOA_grid.reduced_3plustows, year)

goa_years <- unique(dat_GOA_grid.reduced_3plustows[,year])


goa_similarity_mod_coefs <- data.table("year" = goa_years)


for (i in 1:length(goa_years)) {
  
  reduced_year <- dat_GOA_grid.reduced_3plustows[year == goa_years[i],]
  
  #distances among cells
  setorder(reduced_year, haulid)
  
  lat_lon_haulid <- unique(reduced_year[,.(lat,lon,haulid)])
  
  nsites=nrow(lat_lon_haulid)
  
  Results_jaccardtotal=matrix(NA,nrow=nsites,ncol=2)
  Results_jaccardturnover=matrix(NA,nrow=nsites,ncol=2)
  
    
  Results_braytotal=matrix(NA,nrow=nsites,ncol=2)
  Results_braybalanced=matrix(NA,nrow=nsites,ncol=2)
  
  for (j in 1:nrow(lat_lon_haulid)){ #jackknife pulls out one haul id per run
                  
                  lat_lon_haulid.m1 <- lat_lon_haulid[-j,]
                
                distances <- distm(lat_lon_haulid.m1[,.(lon, lat)])
                
                distances.dist <- as.dist(distances)
                
                #get rid of any observations in the missing haul
                haulid.skip <- lat_lon_haulid[j,]$haulid
                reduced_year.m1 <- reduced_year[haulid != haulid.skip]
                
                #throw to wide
                
                reduced_year_wide <- dcast(reduced_year.m1, haulid ~ matched_name2, value.var = "wtcpue", fill = 0, fun.aggregate = sum) #long to wide data for community matrix, column names are cell then species
                
              #for presence absense matrix, all values !=0 are 1s
              index <- reduced_year_wide[,1]
              spp_only <- reduced_year_wide[,2:ncol(reduced_year_wide)]
              
              spp_only[spp_only > 0] <- as.numeric(1)
              
              reduced_year_wide.pres <- cbind(index, spp_only)
              
              #now, use betapart to construct objects to be used in decay models
              
              beta.pair.abund <- beta.pair.abund(reduced_year_wide[,2:ncol(reduced_year_wide)], index.family = "bray") #abundances, bray curtis
              
              beta.pair <- beta.pair(reduced_year_wide.pres[,2:ncol(reduced_year_wide.pres)], index.family = "jaccard") #occurrences, jaccard
              
              
              #ABUNDANCE
              
              #bray balanced (turnover), fitting glm because I can't get betapart to work (no initializing values)
                s = 1-as.vector(beta.pair.abund$beta.bray.bal)
                d = as.vector(distances.dist)
                g = suppressWarnings(glm(s~d,family=binomial(link='log')))
                  
                      Results_braybalanced[j,]=coef(g)
                
                
              #bray total (nestedness and turnover)  
                s = 1-as.vector(beta.pair.abund$beta.bray)
                d = as.vector(distances.dist)
                g = suppressWarnings(glm(s~d,family=binomial(link='log')))
                  
                      Results_braytotal[j,]=coef(g)
                
                
                #OCCURRENCE
                #jaccard turnover
                 s = 1-as.vector(beta.pair$beta.jtu)
                d = as.vector(distances.dist)
                g = suppressWarnings(glm(s~d,family=binomial(link='log')))
                  
                      Results_jaccardturnover[j,]=coef(g)
                
                #total jaccard from pres abs data
                s = 1-as.vector(beta.pair$beta.jac)
                d = as.vector(distances.dist)
                g = suppressWarnings(glm(s~d,family=binomial(link='log')))
                  
                      Results_jaccardtotal[j,]=coef(g)
              
                    print(paste0(j,"/",nsites))
  }
  
    Results_braytotal=cbind(Results_braytotal[,1],-Results_braytotal[,2],
                            exp(Results_braytotal[,1]),-log(2)/Results_braytotal[,2]) #check this calculation, fine for now
    colnames(Results_braytotal)=c("a_intercept","beta_slope","s0","halfd")
    Summary_braytotal=rbind(apply(Results_braytotal,2,mean),
                sqrt(diag(var(Results_braytotal)*(nsites-1)*(1-1/nsites))))
      rownames(Summary_braytotal)=c("Jnife_mean","se")
      
    Results_braybalanced=cbind(Results_braybalanced[,1],-Results_braybalanced[,2],
                            exp(Results_braybalanced[,1]),-log(2)/Results_braybalanced[,2])
    colnames(Results_braybalanced)=c("a_intercept","beta_slope","s0","halfd")
    Summary_braybalanced=rbind(apply(Results_braybalanced,2,mean),
                sqrt(diag(var(Results_braybalanced)*(nsites-1)*(1-1/nsites))))
      rownames(Summary_braybalanced)=c("Jnife_mean","se")
      
    Results_jaccardtotal=cbind(Results_jaccardtotal[,1],-Results_jaccardtotal[,2],
                            exp(Results_jaccardtotal[,1]),-log(2)/Results_jaccardtotal[,2])
    colnames(Results_jaccardtotal)=c("a_intercept","beta_slope","s0","halfd")
    Summary_jaccardtotal=rbind(apply(Results_jaccardtotal,2,mean),
                sqrt(diag(var(Results_jaccardtotal)*(nsites-1)*(1-1/nsites))))
      rownames(Summary_jaccardtotal)=c("Jnife_mean","se")
      
    Results_jaccardturnover=cbind(Results_jaccardturnover[,1],-Results_jaccardturnover[,2],
                            exp(Results_jaccardturnover[,1]),-log(2)/Results_jaccardturnover[,2])
    colnames(Results_jaccardturnover)=c("a_intercept","beta_slope","s0","halfd")
    Summary_jaccardturnover=rbind(apply(Results_jaccardturnover,2,mean),
                sqrt(diag(var(Results_jaccardturnover)*(nsites-1)*(1-1/nsites))))
      rownames(Summary_jaccardturnover)=c("Jnife_mean","se")
      
      
      #mean values of intercept and slope
      goa_similarity_mod_coefs[i,"jaccard_total_intercept"] <- Summary_jaccardtotal[1,1]
      goa_similarity_mod_coefs[i,"jaccard_total_slope"] <- Summary_jaccardtotal[1,2]
      goa_similarity_mod_coefs[i, "jaccard_turnover_intercept"] <- Summary_jaccardturnover[1,1] 
      goa_similarity_mod_coefs[i, "jaccard_turnover_slope"] <- Summary_jaccardturnover[1,2] 
      goa_similarity_mod_coefs[i, "bray_total_intercept"] <- Summary_braytotal[1,1] 
      goa_similarity_mod_coefs[i, "bray_total_slope"] <- Summary_braytotal[1,2] 
      goa_similarity_mod_coefs[i, "bray_balanced_intercept"] <- Summary_braybalanced[1,1] 
      goa_similarity_mod_coefs[i, "bray_balanced_slope"] <- Summary_braybalanced[1,2] 
      
      goa_similarity_mod_coefs[i,"jaccard_total_initial"] <- Summary_jaccardtotal[1,3]
      goa_similarity_mod_coefs[i,"jaccard_total_halving"] <- Summary_jaccardtotal[1,4]
      goa_similarity_mod_coefs[i, "jaccard_turnover_initial"] <- Summary_jaccardturnover[1,3] 
      goa_similarity_mod_coefs[i, "jaccard_turnover_halving"] <- Summary_jaccardturnover[1,4] 
      goa_similarity_mod_coefs[i, "bray_total_initial"] <- Summary_braytotal[1,3] 
      goa_similarity_mod_coefs[i, "bray_total_halving"] <- Summary_braytotal[1,4] 
      goa_similarity_mod_coefs[i, "bray_balanced_initial"] <- Summary_braybalanced[1,3] 
      goa_similarity_mod_coefs[i, "bray_balanced_halving"] <- Summary_braybalanced[1,4] 
      
      #standard error
      goa_similarity_mod_coefs[i,"jaccard_total_intercept_se"] <- Summary_jaccardtotal[2,1]
      goa_similarity_mod_coefs[i,"jaccard_total_slope_se"] <- Summary_jaccardtotal[2,2]
      goa_similarity_mod_coefs[i, "jaccard_turnover_intercept_se"] <- Summary_jaccardturnover[2,1] 
      goa_similarity_mod_coefs[i, "jaccard_turnover_slope_se"] <- Summary_jaccardturnover[2,2] 
      goa_similarity_mod_coefs[i, "bray_total_intercept_se"] <- Summary_braytotal[2,1] 
      goa_similarity_mod_coefs[i, "bray_total_slope_se"] <- Summary_braytotal[2,2] 
      goa_similarity_mod_coefs[i, "bray_balanced_intercept_se"] <- Summary_braybalanced[2,1] 
      goa_similarity_mod_coefs[i, "bray_balanced_slope_se"] <- Summary_braybalanced[2,2] 
       
      goa_similarity_mod_coefs[i,"jaccard_total_initial_se"] <- Summary_jaccardtotal[2,3]
      goa_similarity_mod_coefs[i,"jaccard_total_halving_se"] <- Summary_jaccardtotal[2,4]
      goa_similarity_mod_coefs[i, "jaccard_turnover_initial_se"] <- Summary_jaccardturnover[2,3] 
      goa_similarity_mod_coefs[i, "jaccard_turnover_halving_se"] <- Summary_jaccardturnover[2,4] 
      goa_similarity_mod_coefs[i, "bray_total_initial_se"] <- Summary_braytotal[2,3] 
      goa_similarity_mod_coefs[i, "bray_total_halving_se"] <- Summary_braytotal[2,4] 
      goa_similarity_mod_coefs[i, "bray_balanced_initial_se"] <- Summary_braybalanced[2,3] 
      goa_similarity_mod_coefs[i, "bray_balanced_halving_se"] <- Summary_braybalanced[2,4] 
      
    
      print(paste0(i,"/",length(goa_years)))
}

save(goa_similarity_mod_coefs, file = "goa_similarity_mod_coefs.RData")
load("goa_similarity_mod_coefs.RData")
# i = 7, j = 434



```

Visualizing model parameters over time
```{r visualizing model parameters HALVING DISTANCE}
#jaccard total halving distance
ggplot(aes(x = as.numeric(year)), data = goa_similarity_mod_coefs) +
  geom_errorbar(aes(ymin = jaccard_total_halving/1000-jaccard_total_halving_se/1000, ymax = jaccard_total_halving/1000+jaccard_total_halving_se/1000), alpha = 0.3, color = "black") +
  geom_point(aes(y = jaccard_total_halving/1000)) +
  geom_smooth(method = "lm", aes(y = jaccard_total_halving/1000), linetype = 1, color = "black", se = F, alpha = 0.4) +
  geom_errorbar(aes(ymin = jaccard_turnover_halving/1000-jaccard_turnover_halving_se/1000, ymax = jaccard_turnover_halving/1000+jaccard_turnover_halving_se/1000), alpha = 0.3, color = "green") +
  geom_point(aes(y = jaccard_turnover_halving/1000), color = "aquamarine4", position = position_dodge(width = 0.3)) +
  geom_smooth(method = "lm", aes(y = jaccard_turnover_halving/1000), linetype = 1, color = "aquamarine4", se = F, alpha = 0.4) +
  labs(x = "Year", y = "Jaccard Similarity Halving Distance (km)") +
  theme_classic() +
  theme(text = element_text(size = 13))

ggsave(path = here::here("figures","goa"),filename = "goa_jaccard_halving_distance.jpg")
ggsave(path = here::here("figures","goa"),filename = "goa_jaccard_halving_distance.eps")

#bray curtis total halving distance
ggplot(aes(x = as.numeric(year)), data = goa_similarity_mod_coefs) +
  geom_errorbar(aes(ymin = bray_total_halving/1000-bray_total_halving_se/1000, ymax = bray_total_halving/1000+bray_total_halving_se/1000), alpha = 0.3, color = "black") +
  geom_point(aes(y = bray_total_halving/1000)) +
  geom_smooth(method = "lm", aes(y = bray_total_halving/1000), linetype = 1, color = "black", se = F, alpha = 0.4) +
  geom_errorbar(aes(ymin = bray_balanced_halving/1000-bray_balanced_halving_se/1000, ymax = bray_balanced_halving/1000+bray_balanced_halving_se/1000), alpha = 0.3, color = "green") +
  geom_point(aes(y = bray_balanced_halving/1000), color = "aquamarine4", position = position_dodge(width = 0.3)) +
  geom_smooth(method = "lm", aes(y = bray_balanced_halving/1000), linetype = 1, color = "aquamarine4", se = F, alpha = 0.4) +
  labs(x = "Year", y = "Bray Similarity Halving Distance (km)") +
  theme_classic() +
  theme(text = element_text(size = 13))


ggsave(path = here::here("figures","goa"),filename = "goa_bray_halving_distance.jpg")

ggsave(path = here::here("figures","goa"),filename = "goa_bray_halving_distance.eps")

```

Now, same, but for  initial similarity
```{r visualizing  model parameters}
#jaccard initial similarity
ggplot(aes(x = as.numeric(year)), data = goa_similarity_mod_coefs) +
  geom_errorbar(aes(ymin = jaccard_total_initial-jaccard_total_initial_se, ymax = jaccard_total_initial+jaccard_total_initial_se), alpha = 0.3, color = "black") +
  geom_point(aes(y = jaccard_total_initial)) +
  geom_smooth(method = "lm", aes(y = jaccard_total_initial), linetype = 1, color = "black", se = F, alpha = 0.4) +
  geom_errorbar(aes(ymin = jaccard_turnover_initial-jaccard_turnover_initial_se, ymax = jaccard_turnover_initial+jaccard_turnover_initial_se), alpha = 0.3, color = "green") +
  geom_point(aes(y = jaccard_turnover_initial), color = "aquamarine4", position = position_dodge(width = 0.3)) +
  geom_smooth(method = "lm", aes(y = jaccard_turnover_initial), linetype = 1, color = "aquamarine4", se = F, alpha = 0.4) +
  labs(x = "Year", y = "Jaccard Initial Similarity") +
  theme_classic() +
  theme(text = element_text(size = 13))

ggsave(path = here::here("figures","goa"),filename = "goa_jaccard_initial_similarity.jpg")
ggsave(path = here::here("figures","goa"),filename = "goa_jaccard_initial_similarity.eps")

#bray curtis initial similarity
ggplot(aes(x = as.numeric(year)), data = goa_similarity_mod_coefs) +
  geom_errorbar(aes(ymin = bray_total_initial-bray_total_initial_se, ymax = bray_total_initial+bray_total_initial_se), alpha = 0.3, color = "black") +
  geom_point(aes(y = bray_total_initial)) +
  geom_smooth(method = "lm", aes(y = bray_total_initial), linetype = 1, color = "black", se = F, alpha = 0.4) +
  geom_errorbar(aes(ymin = bray_balanced_initial-bray_balanced_initial_se, ymax = bray_balanced_initial+bray_balanced_initial_se), alpha = 0.3, color = "green") +
  geom_point(aes(y = bray_balanced_initial), color = "aquamarine4", position = position_dodge(width = 0.3)) +
  geom_smooth(method = "lm", aes(y = bray_balanced_initial), linetype = 1, color = "aquamarine4", se = F, alpha = 0.4) +
  labs(x = "Year", y = "Bray Curtis Initial Similarity") +
  theme_classic() +
  theme(text = element_text(size = 13))


ggsave(path = here::here("figures","goa"),filename = "goa_bray_curtis_initial_similarity.jpg")
ggsave(path = here::here("figures","goa"),filename = "goa_bray_curtis_initial_similarity.eps")

```


##Plot both seasons together

Halving distance

```{r both jaccard together halving distance}
#jaccard total halving distance
ggplot() +
  #
#  geom_errorbar(data = goa_similarity_mod_coefs, aes(x = as.numeric(year), ymin = jaccard_total_halving/1000-jaccard_total_halving_se/1000, ymax = jaccard_total_halving/1000+jaccard_total_halving_se/1000), alpha = 0.3, color = "black") +
  geom_point(data = goa_similarity_mod_coefs, aes(x = as.numeric(year), y = jaccard_total_halving/1000)) +
  geom_smooth(data = goa_similarity_mod_coefs, method = "lm", aes(x = as.numeric(year), y = jaccard_total_halving/1000), linetype = 1, color = "black", se = F, alpha = 0.4) +
#  geom_errorbar(data = goa_similarity_mod_coefs, aes(x = as.numeric(year), ymin = jaccard_turnover_halving/1000-jaccard_turnover_halving_se/1000, ymax = jaccard_turnover_halving/1000+jaccard_turnover_halving_se/1000), alpha = 0.3, color = "green") +
  geom_point(data = goa_similarity_mod_coefs, aes(x = as.numeric(year), y = jaccard_turnover_halving/1000), color = "aquamarine4", position = position_dodge(width = 0.3)) +
  geom_smooth(data = goa_similarity_mod_coefs, method = "lm", aes(x = as.numeric(year), y = jaccard_turnover_halving/1000), linetype = 1, color = "aquamarine4", se = F, alpha = 0.4) +
  
    #fall
#  geom_errorbar(data = goa_similarity_mod_coefs_fall, aes(x = as.numeric(year), ymin = jaccard_total_halving/1000-jaccard_total_halving_se/1000, ymax = jaccard_total_halving/1000+jaccard_total_halving_se/1000), alpha = 0.3) +
  geom_point(data = goa_similarity_mod_coefs_fall, aes(x = as.numeric(year), y = jaccard_total_halving/1000),shape = 15, color = "grey") +
  geom_smooth(data = goa_similarity_mod_coefs_fall, method = "lm", aes(x = as.numeric(year), y = jaccard_total_halving/1000),  color = "grey", se = F, alpha = 0.4,linetype = "dashed") +
#  geom_errorbar(data = goa_similarity_mod_coefs_fall, aes(x = as.numeric(year), ymin = jaccard_turnover_halving/1000-jaccard_turnover_halving_se/1000, ymax = jaccard_turnover_halving/1000+jaccard_turnover_halving_se/1000), alpha = 0.3, color = "green") +
  geom_point(data = goa_similarity_mod_coefs_fall, aes(x = as.numeric(year), y = jaccard_turnover_halving/1000), color = "aquamarine2", position = position_dodge(width = 0.3),shape = 15) +
  geom_smooth(data = goa_similarity_mod_coefs_fall, method = "lm", aes(x = as.numeric(year), y = jaccard_turnover_halving/1000), linetype = "dashed", color = "aquamarine2", se = F, alpha = 0.4) +
  labs(x = "Year", y = "Jaccard Similarity Halving Distance (km)") +
  theme_classic() +
  theme(text = element_text(size = 13))

ggsave(path = here::here("figures","goa"), filename = "goa_bothseasons_jaccard_halvingdistance.jpg", height = 4, width = 4, unit = "in")
```
Both Bray together
```{r both bray together halving}
#jbray total halving distance
ggplot() +
  #
#  geom_errorbar(data = goa_similarity_mod_coefs, aes(x = as.numeric(year), ymin = bray_total_halving/1000-bray_total_halving_se/1000, ymax = bray_total_halving/1000+bray_total_halving_se/1000), alpha = 0.3, color = "black") +
  geom_point(data = goa_similarity_mod_coefs, aes(x = as.numeric(year), y = bray_total_halving/1000)) +
  geom_smooth(data = goa_similarity_mod_coefs, method = "lm", aes(x = as.numeric(year), y = bray_total_halving/1000), linetype = 1, color = "black", se = F, alpha = 0.4) +
#  geom_errorbar(data = goa_similarity_mod_coefs, aes(x = as.numeric(year), ymin = bray_balanced_halving/1000-bray_balanced_halving_se/1000, ymax = bray_balanced_halving/1000+bray_balanced_halving_se/1000), alpha = 0.3, color = "green") +
  geom_point(data = goa_similarity_mod_coefs, aes(x = as.numeric(year), y = bray_balanced_halving/1000), color = "aquamarine4", position = position_dodge(width = 0.3)) +
  geom_smooth(data = goa_similarity_mod_coefs, method = "lm", aes(x = as.numeric(year), y = bray_balanced_halving/1000), linetype = 1, color = "aquamarine4", se = F, alpha = 0.4) +
  
    #fall
#  geom_errorbar(data = goa_similarity_mod_coefs_fall, aes(x = as.numeric(year), ymin = bray_total_halving/1000-bray_total_halving_se/1000, ymax = bray_total_halving/1000+bray_total_halving_se/1000), alpha = 0.3) +
  geom_point(data = goa_similarity_mod_coefs_fall, aes(x = as.numeric(year), y = bray_total_halving/1000),shape = 15, color = "grey") +
  geom_smooth(data = goa_similarity_mod_coefs_fall, method = "lm", aes(x = as.numeric(year), y = bray_total_halving/1000),  color = "grey", se = F, alpha = 0.4,linetype = "dashed") +
#  geom_errorbar(data = goa_similarity_mod_coefs_fall, aes(x = as.numeric(year), ymin = bray_balanced_halving/1000-bray_balanced_halving_se/1000, ymax = bray_balanced_halving/1000+bray_balanced_halving_se/1000), alpha = 0.3, color = "green") +
  geom_point(data = goa_similarity_mod_coefs_fall, aes(x = as.numeric(year), y = bray_balanced_halving/1000), color = "aquamarine2", position = position_dodge(width = 0.3),shape = 15) +
  geom_smooth(data = goa_similarity_mod_coefs_fall, method = "lm", aes(x = as.numeric(year), y = bray_balanced_halving/1000), linetype = "dashed", color = "aquamarine2", se = F, alpha = 0.4) +
  labs(x = "Year", y = "Bray Curtis Similarity Halving Distance (km)") +
  theme_classic() +
  theme(text = element_text(size = 13))

ggsave(path = here::here("figures","goa"), filename = "goa_bothseasons_bray_halvingdistance.jpg", height = 4, width = 4, unit = "in")
```
Now, same but for initial sim
Jaccard initial similarity
```{r both seasons jaccard initial similarity}
#jaccard initial similarity
ggplot() +
  #
#  geom_errorbar(data = goa_similarity_mod_coefs, aes(x = as.numeric(year), ymin = jaccard_total_initial-jaccard_total_initial_se, ymax = jaccard_total_initial+jaccard_total_initial_se), alpha = 0.3, color = "black") +
  geom_point(data = goa_similarity_mod_coefs, aes(x = as.numeric(year), y = jaccard_total_initial)) +
  geom_smooth(data = goa_similarity_mod_coefs, method = "lm", aes(x = as.numeric(year), y = jaccard_total_initial), linetype = 1, color = "black", se = F, alpha = 0.4) +
#  geom_errorbar(data = goa_similarity_mod_coefs, aes(x = as.numeric(year), ymin = jaccard_turnover_initial-jaccard_turnover_initial_se, ymax = jaccard_turnover_initial+jaccard_turnover_initial_se), alpha = 0.3, color = "green") +
  geom_point(data = goa_similarity_mod_coefs, aes(x = as.numeric(year), y = jaccard_turnover_initial), color = "aquamarine4", position = position_dodge(width = 0.3)) +
  geom_smooth(data = goa_similarity_mod_coefs, method = "lm", aes(x = as.numeric(year), y = jaccard_turnover_initial), linetype = 1, color = "aquamarine4", se = F, alpha = 0.4) +
  
    #fall
#  geom_errorbar(data = goa_similarity_mod_coefs_fall, aes(x = as.numeric(year), ymin = jaccard_total_initial-jaccard_total_initial_se, ymax = jaccard_total_initial+jaccard_total_initial_se), alpha = 0.3) +
  geom_point(data = goa_similarity_mod_coefs_fall, aes(x = as.numeric(year), y = jaccard_total_initial),shape = 15, color = "grey") +
  geom_smooth(data = goa_similarity_mod_coefs_fall, method = "lm", aes(x = as.numeric(year), y = jaccard_total_initial),  color = "grey", se = F, alpha = 0.4,linetype = "dashed") +
#  geom_errorbar(data = goa_similarity_mod_coefs_fall, aes(x = as.numeric(year), ymin = jaccard_turnover_initial-jaccard_turnover_initial_se, ymax = jaccard_turnover_initial+jaccard_turnover_initial_se), alpha = 0.3, color = "green") +
  geom_point(data = goa_similarity_mod_coefs_fall, aes(x = as.numeric(year), y = jaccard_turnover_initial), color = "aquamarine2", position = position_dodge(width = 0.3),shape = 15) +
  geom_smooth(data = goa_similarity_mod_coefs_fall, method = "lm", aes(x = as.numeric(year), y = jaccard_turnover_initial), linetype = "dashed", color = "aquamarine2", se = F, alpha = 0.4) +
  labs(x = "Year", y = "Jaccard Initial Similarity") +
  theme_classic() +
  theme(text = element_text(size = 13))

ggsave(path = here::here("figures","goa"), filename = "goa_bothseasons_jaccard_initialdistance.jpg", height = 4, width = 4, unit = "in")
```

Both seasons of Bray together
```{r both bray together initial similarity goa}
#bray initial similarity
ggplot() +
  #
#  geom_errorbar(data = goa_similarity_mod_coefs, aes(x = as.numeric(year), ymin = bray_total_initial-bray_total_initial_se, ymax = bray_total_initial+bray_total_initial_se), alpha = 0.3, color = "black") +
  geom_point(data = goa_similarity_mod_coefs, aes(x = as.numeric(year), y = bray_total_initial)) +
  geom_smooth(data = goa_similarity_mod_coefs, method = "lm", aes(x = as.numeric(year), y = bray_total_initial), linetype = 1, color = "black", se = F, alpha = 0.4) +
#  geom_errorbar(data = goa_similarity_mod_coefs, aes(x = as.numeric(year), ymin = bray_balanced_initial-bray_balanced_initial_se, ymax = bray_balanced_initial+bray_balanced_initial_se), alpha = 0.3, color = "green") +
  geom_point(data = goa_similarity_mod_coefs, aes(x = as.numeric(year), y = bray_balanced_initial), color = "aquamarine4", position = position_dodge(width = 0.3)) +
  geom_smooth(data = goa_similarity_mod_coefs, method = "lm", aes(x = as.numeric(year), y = bray_balanced_initial), linetype = 1, color = "aquamarine4", se = F, alpha = 0.4) +
  
    #fall
#  geom_errorbar(data = goa_similarity_mod_coefs_fall, aes(x = as.numeric(year), ymin = bray_total_initial-bray_total_initial_se, ymax = bray_total_initial+bray_total_initial_se), alpha = 0.3) +
  geom_point(data = goa_similarity_mod_coefs_fall, aes(x = as.numeric(year), y = bray_total_initial),shape = 15, color = "grey") +
  geom_smooth(data = goa_similarity_mod_coefs_fall, method = "lm", aes(x = as.numeric(year), y = bray_total_initial),  color = "grey", se = F, alpha = 0.4,linetype = "dashed") +
#  geom_errorbar(data = goa_similarity_mod_coefs_fall, aes(x = as.numeric(year), ymin = bray_balanced_initial-bray_balanced_initial_se, ymax = bray_balanced_initial+bray_balanced_initial_se), alpha = 0.3, color = "green") +
  geom_point(data = goa_similarity_mod_coefs_fall, aes(x = as.numeric(year), y = bray_balanced_initial), color = "aquamarine2", position = position_dodge(width = 0.3),shape = 15) +
  geom_smooth(data = goa_similarity_mod_coefs_fall, method = "lm", aes(x = as.numeric(year), y = bray_balanced_initial), linetype = "dashed", color = "aquamarine2", se = F, alpha = 0.4) +
  labs(x = "Year", y = "Bray Curtis Initial Similarity") +
  theme_classic() +
  theme(text = element_text(size = 13))

ggsave(path = here::here("figures","goa"), filename = "goa_bothseasons_bray_initialdistance.jpg", height = 4, width = 4, unit = "in")
```