---
title: "NEUS Distance Decay over Time"
output: html_notebook
---

```{r setup}
library(data.table)
library(vegan)
library(ggplot2)
library(broom) #augment
library(betapart)

#load spring and fall data
dat_NEUS_grid.spring.reduced_3plustows <- readRDS("/Users/zoekitchel/Documents/grad school/Rutgers/Repositories/trawl_spatial_turnover/data/gridded/USA-NEUS/dat_NEUS_grid.spring.reduced_3plustows.RData")

dat_NEUS_grid.fall.reduced_3plustows <- readRDS("/Users/zoekitchel/Documents/grad school/Rutgers/Repositories/trawl_spatial_turnover/data/gridded/USA-NEUS/dat_NEUS_grid.fall.reduced_3plustows.RData")

#distance among grid cells
neus_reg_distances_fall.l <- readRDS("/Users/zoekitchel/Documents/grad school/Rutgers/Repositories/trawl_spatial_turnover/data/gridded/USA-NEUS/neus_reg_distances_fall.l.RData")

neus_reg_distances_spring.l <- readRDS("/Users/zoekitchel/Documents/grad school/Rutgers/Repositories/trawl_spatial_turnover/data/gridded/USA-NEUS/neus_reg_distances_spring.l.RData")


```

I need it to be a data frame formatted species (rows) by sites/years (columns)
TO DO -any reason I shouldn't just sum? should I sample 1 tow from each grid cell and take an average? randomly sample one tow from each cell?

-I think I should randomly sample 3 tows from one cell, calculate metrics, and then resample



```{r sum across grid cells and years}
#fall
neus_data_year_cell_fall <- dat_NEUS_grid.fall.reduced[,wtcpue_sum := sum(wtcpue, na.rm = T), .(matched_name2, cell, year)]

#spring

neus_data_year_cell.r <- unique(neus_data_year_cell[,.(matched_name2, cell, year, wtcpue_sum)])
```

I want to go year by year

Let's start with one grid cell for species by cell data frame for each year

```{r start with community matrix for one year}

#list of years 
neus_years <- unique(neus_data_year_cell.r[,year])

sort(neus_years)

i = 1

neus_subset_year <- neus_data_year_cell.r[year == neus_years[i],]

#long to wide
neus_subset_year_wide <- dcast(neus_subset_year, cell ~ matched_name2, value.var = "wtcpue_sum", fill = 0)

```


```{r bray curtis dissimilarities between cells across multiple years}
distances_dissimilarities_allyears_CELL <- data.table("cell1" = integer(), "cell2" = integer(), "distance(km)" = numeric(),"bray_curtis_dissimilarity" = numeric(), year = integer(), "jaccard_dissimilarity" = numeric())

#Now loop through all years starting in 1982
for (i in 1:length(neus_years)) {
  reduced_year <- neus_data_year_cell.r[year == neus_years[i],]
  reduced_year_wide <- dcast(reduced_year, year + cell ~ matched_name2, value.var = "wtcpue_sum", fill = 0) #long to wide data for community matrix, column names are cell then species
  
  ncols <- ncol(reduced_year_wide)
  communitymatrix <- cbind(reduced_year_wide[,3:ncols], reduced_year_wide[,1:2]) #community matrix with year and cell on far right

  cells <- communitymatrix$cell

  dissimilarities_abundance <- vegdist(communitymatrix[,1:(ncols-2)], method = "bray", binary = F) #dissimilarity 
  dissimilarities_occurrence <- vegdist(communitymatrix[,1:(ncols-2)], method = "jaccard", binary = T)

  #make into matrix
  dissimilarities_abundance.m <- as.matrix(dissimilarities_abundance, labels=TRUE)
  dissimilarities_occurrence.m <- as.matrix(dissimilarities_occurrence, labels=TRUE)
  colnames(dissimilarities_abundance.m) <- rownames(dissimilarities_abundance.m) <- cells
  colnames(dissimilarities_occurrence.m) <- rownames(dissimilarities_occurrence.m) <- cells

  #reshape dissimilarities
  dissimilarities_abundance.l <- reshape2::melt(dissimilarities_abundance.m, varnames = c("cell1", "cell2"), value.name = "bray_curtis_dissimilarity")
  dissimilarities_occurrence.l <- reshape2::melt(dissimilarities_occurrence.m, varnames = c("cell1", "cell2"), value.name = "jaccard_dissimilarity")
  dissimilarities_abundance.l <- data.table(dissimilarities_abundance.l) #and then to data table
  dissimilarities_occurrence.l <- data.table(dissimilarities_occurrence.l)
  dissimilarities_abundance.l <- dissimilarities_abundance.l[cell1 > cell2,] # to get rid of repetitions and self-comparisons
  dissimilarities_occurrence.l <- dissimilarities_occurrence.l[cell1 > cell2,]


  #add year for these values
  dissimilarities_abundance.l[, "year" := neus_years[i]]
  dissimilarities_occurrence.l[, "year" := neus_years[i]]

  #merge distance with dissimilarity for this year with both metrics of dissimilarity
  dissimilarities_full <- neus_distances.l[dissimilarities_abundance.l, on = c("cell1", "cell2")]
  dissimilarities_full <- dissimilarities_full[dissimilarities_occurrence.l, on = c("cell1", "cell2", "year")]


  #add to data table
  distances_dissimilarities_allyears_CELL <- rbind(distances_dissimilarities_allyears_CELL, dissimilarities_full)

  #extract info for distance decay exponential model for each year
  #reduce stratum_distances to stratum.r
  #stratum_distances.r <- stratum_distances[as.character(stratum.r), as.character(stratum.r)]

  #modname <- paste0(years[i], "_decay_mod_exponential")

  #assign(modname, decay.model(dissimilarities.m, stratum_distances.r, model.type = "exponential", y.type = "dissimilarities", perm = 100))

  #jpeg(file=paste0("jaccardplot_", years[i], ".jpg"))
  #plot(x=dissimilarities.l.full$`distance(m)`, y=dissimilarities.l.full$jaccard_dissimilarity, xlab = "Distance(m)", #ylab="Jaccard dissimilarity", ylim = c(0:1), pch = 19)
  #dev.off()
 
  #PCA
  #vare.pca <- rda(communitymatrix[,c(1:817)])
  #jpeg(file=paste0("jaccardpca_", years[i], ".jpg"))
  #plot(vare.pca, display = "sites", xlim = c(-4,4), ylim = c(-4,4))
  #dev.off()
  
}

summary(distances_dissimilarities_allyears_CELL) #here we have bray, jaccard and geographic distance

# save(distances_dissimilarities_allyears, file = "output/jaccard_similarities_distances_4reg.Rdata") # from an earlier file

distances_dissimilarities_allyears_CELL[,bray_curtis_similarity := (1-bray_curtis_dissimilarity)][,jaccard_similarity := (1-jaccard_dissimilarity)]

save(distances_dissimilarities_allyears_CELL, file = "neus_bray_jaccard_similarities_distances_cell.Rdata")

```

Show example of a single distance decay plot with distance on x axis and bray curtis dissimilarity on y axis
```{r distance decay}
ggplot(aes(x=`distance(km)`), 
       data = distances_dissimilarities_allyears_CELL[year == 1983]) + 
  geom_point(aes(y = 1-bray_curtis_dissimilarity), shape = 4) +
  geom_smooth(method = "lm", aes(y = 1-bray_curtis_dissimilarity), linetype = 2, color = "black") +
  geom_point(aes(y = 1-jaccard_dissimilarity), shape = 15) +
  labs(x = "Distance(km)", y = "Similarity") +
  geom_smooth(method = "lm", aes(y = 1-jaccard_dissimilarity), linetype = 1, color = "black") +
  theme_classic() +
  lims(y = c(0,1)) +
  theme(text = element_text(size = 13))
 # + facet_wrap(~year)
  
```
Instead, we will try exponential decay 
```{r exponential decay}
fit_neus1983_bray <- nls(bray_curtis_similarity ~ SSasymp(`distance(km)`, yf, y0, log_alpha), data = distances_dissimilarities_allyears_CELL[year == 1983])

fit_neus1983_jaccard <- nls(jaccard_similarity ~ SSasymp(`distance(km)`, yf, y0, log_alpha), data = distances_dissimilarities_allyears_CELL[year == 1983])


```

Plot again with these exponential fits
```{r plot with exponential}
ggplot() + 
  geom_point(data = distances_dissimilarities_allyears_CELL[year == 1983], aes(y = bray_curtis_similarity, x=`distance(km)`), shape = 4) +
  geom_line(data = augment(fit_neus1983_bray), aes(x = `distance(km)`, y = .fitted), linetype = 2, color = "black") +
  geom_point(data = distances_dissimilarities_allyears_CELL[year == 1983], aes(y = jaccard_similarity, x=`distance(km)`), shape = 15) +
  labs(x = "Distance(km)", y = "Similarity") +
  geom_line(data = augment(fit_neus1983_jaccard), aes(x = `distance(km)`, y = .fitted), linetype = 1, color = "black") +
  theme_classic() +
  lims(y = c(0,1)) +
  theme(text = element_text(size = 13))

```


Extract coefficient (slope) for each year: how does it change through time? I did this because I couldn't get betapart decay.model to work.

Setup betapart object

```{r}

#testing 1985

#for presence absense matrix, all values !=0 are 1s
index <- neus_subset_year_wide[,1]
spp_only <- neus_subset_year_wide[,2:ncol(neus_subset_year_wide)]

spp_only[spp_only > 0] <- as.numeric(1)

neus_subset_year_wide.pres <- cbind(index, spp_only)
rownames(neus_subset_year_wide.pres) <- neus_subset_year_wide.pres$cell

#now, use betapart to construct objects to be used in decay models

beta.pair <- beta.pair(neus_subset_year_wide.pres[,2:ncol(neus_subset_year_wide.pres)])
beta.pair.abund <- beta.pair.abund(neus_subset_year_wide[,2:ncol(neus_subset_year_wide)])
```

```{r each year build linear model}
#I'm interested in the rate of decrease

#distances between cells
distances <- as.dist(neus_reg_distances)

neus_similarity_mod_coefs <- data.table("year" = neus_years, "jaccard_total_initial_sim" = 0, "jaccard_total_halving" = 0, "jaccard_turnover_initial_sim" = 0, "jaccard_turnover_halving" = 0, "bray_total_initial_sim" = 0, "bray_total_halving" = 0, "bray_balanced_initial_sim" = 0, "bray_balanced_halving" = 0)


#match cells to those in this region now

for (i in 1:length(neus_years)) {
  
  reduced_year <- neus_data_year_cell.r[year == neus_years[i],]
  reduced_year_wide <- dcast(reduced_year, year + cell ~ matched_name2, value.var = "wtcpue_sum", fill = 0) #long to wide data for community matrix, column names are cell then species
  
#for presence absense matrix, all values !=0 are 1s
index <- reduced_year_wide[,1:2]
spp_only <- reduced_year_wide[,3:ncol(reduced_year_wide)]

spp_only[spp_only > 0] <- as.numeric(1)

reduced_year_wide.pres <- cbind(index, spp_only)
rownames(reduced_year_wide.pres) <- reduced_year_wide.pres$cell

#now, use betapart to construct objects to be used in decay models

beta.pair <- beta.pair(reduced_year_wide.pres[,3:ncol(reduced_year_wide.pres)], index.family = "jaccard")
beta.pair.abund <- beta.pair.abund(reduced_year_wide[,3:ncol(reduced_year_wide)], index.family = "bray")
  
#bray balanced (turnover)
  fit_bray_balanced <- decay.model(1-beta.pair.abund$beta.bray.bal, distances, model.type = "exp", y.type = "sim") # 1 minus because output of betadist object gives us dissimilarity
  
#    fit_bray$a.intercept = intercept of the model
  intercept_bray_balanced <- fit_bray_balanced$a.intercept
#    fit_bray_balanced$b.slope = rate at which similarity decreases with distance in a neg exponential model
  slope_bray_balanced <- fit_bray_balanced$b.slope
  
  halving_distance_bray_balanced <- log(2)/(-slope_bray_balanced) #halving distance
  
  neus_similarity_mod_coefs[i,"bray_balanced_halving"] <- halving_distance_bray_balanced
  
  initial_sim_bray_balanced <- fit_bray_balanced$a.intercept*1+fit_bray_balanced$b.slope # initial similarity

  
  neus_similarity_mod_coefs[i,"bray_balanced_initial_sim"] <- initial_sim_bray_balanced
  
  
#bray total (nestedness and turnover)  
  #bray balanced (turnover)
  fit_bray_total <- decay.model(1-beta.pair.abund$beta.bray, distances, model.type = "exp", y.type = "sim") # 1 minus because output of betadist object gives us dissimilarity
  
#    fit_bray$a.intercept = intercept of the model
  intercept_bray_total <- fit_bray_total$a.intercept
#    fit_bray_total$b.slope = rate at which similarity decreases with distance in a neg exponential model
  slope_bray_total <- fit_bray_total$b.slope
  
  halving_distance_bray_total <- log(2)/(-slope_bray_total) #halving distance
  
  neus_similarity_mod_coefs[i,"bray_total_halving"] <- halving_distance_bray_total
  
  initial_sim_bray_total <- fit_bray_total$a.intercept*1+fit_bray_total$b.slope # initial similarity

  
  neus_similarity_mod_coefs[i,"bray_total_initial_sim"] <- initial_sim_bray_total
  
  ########
  #jaccard turnover
  #I can't get this to run when I'm looking at presence absense
  fit_turnover_jaccard <- decay.model(1-beta.pair$beta.jtu, distances, y.type = "similarities", model.type = "exp") #beta.jtu dissimilarity matrix accounting for spatial turnover, measured as the turnoverfraction of Jaccard pair-wise dissimilarity
  
  #    fit_jaccard$a.intercept = intercept of the model
  intercept_turnover_jaccard <- fit_turnover_jaccard$a.intercept
#    fit_jaccard$b.slope = rate at which similarity decreases with distance in a neg exponential model
  slope_turnover_jaccard <- fit_turnover_jaccard$b.slope
  
  halving_distance_turnover_jaccard <- log(2)/(-slope_turnover_jaccard) #halving distance
  
  neus_similarity_mod_coefs[i,"jaccard_turnover_halving"] <- halving_distance_turnover_jaccard
  
  initial_sim_turnover_jaccard <- fit_turnover_jaccard$a.intercept*1+fit_turnover_jaccard$b.slope # initial similarity

  
  neus_similarity_mod_coefs[i,"jaccard_turnover_initial_sim"] <- initial_sim_turnover_jaccard
  
  #total jaccard from pres abs data
    fit_total_jaccard <- decay.model(1-beta.pair$beta.jac, distances, y.type = "similarities", model.type = "exp") #beta.jtu dissimilarity matrix accounting for spatial total, measured as the totalfraction of Jaccard pair-wise dissimilarity
  
  #    fit_jaccard$a.intercept = intercept of the model
  intercept_total_jaccard <- fit_total_jaccard$a.intercept
#    fit_jaccard$b.slope = rate at which similarity decreases with distance in a neg exponential model
  slope_total_jaccard <- fit_total_jaccard$b.slope
  
  halving_distance_total_jaccard <- log(2)/(-slope_total_jaccard) #halving distance
  
  neus_similarity_mod_coefs[i,"jaccard_total_halving"] <- halving_distance_total_jaccard
  
  initial_sim_total_jaccard <- fit_total_jaccard$a.intercept*1+fit_total_jaccard$b.slope # initial similarity

  
  neus_similarity_mod_coefs[i,"jaccard_total_initial_sim"] <- initial_sim_total_jaccard

}

ggplot(aes(x = as.numeric(year)), data = neus_similarity_mod_coefs) +
  geom_point(aes(y = jaccard_total_halving), shape = 15) +
  geom_smooth(aes(y = jaccard_total_halving), linetype = 1, color = "black") +
  geom_point(aes(y = jaccard_turnover_halving), shape = 15, color = "green") +
  geom_smooth(aes(y = jaccard_turnover_halving), linetype = 1, color = "green") +
  labs(x = "Year", y = "Jaccard Similarity Halving Distance (m)") +
  theme_classic() +
  theme(text = element_text(size = 13))

ggplot(aes(x = as.numeric(year)), data = neus_similarity_mod_coefs) +
  geom_point(aes(y = bray_total_halving), color = "black") +
  geom_smooth(aes(y = bray_total_halving), color = "black") +
  geom_point(aes(y = bray_balanced_halving), color = "green") +
  geom_smooth(aes(y = bray_balanced_halving), color = "green") +
  labs(x = "Year", y = "Bray Similarity Halving Distance (m)") +
  theme_classic() +
  theme(text = element_text(size = 13))

ggsave(path = "plots", file = "slopes_for_distance_decay.jpg", width = 6, height = 4, unit = "in")


#increasing rate of decay through time (strata become less similar through time)

```