---
title: "NS-IBITS_similarity_heatmap_distance"
output: html_notebook
---

```{r setup}
library(data.table)
library(viridis)
library(ggplot2)
library(reshape2)
library(geosphere)
library(vegan)
library(raster)
library(here)
library(cowplot)
library(plyr)
library(ggpubr)
library(sf)
library(concaveman)

#load data
ns_grid.reduced_3plustows <- readRDS("/Users/zoekitchel/Documents/grad school/Rutgers/Repositories/trawl_spatial_turnover/data/gridded/ICES-NS/ns_grid.reduced_3plustows.rds")

#distance among grid cells
ns_reg_distances.l <- readRDS("/Users/zoekitchel/Documents/grad school/Rutgers/Repositories/trawl_spatial_turnover/data/gridded/ICES-NS/ns_reg_distances.l.rds")

```

Region Stats

* Number of species in region
* Length of study period
* Number of years in which data was collected
* Range of latitudes sampled
* Mid latitude of region
* Range of longitudes sampled
* Overall area sampled
* Depths sampled

```{r region stats}
ns_stats <- data.table(region = character() , season = character(), spp_num = numeric(), study_period = numeric(), study_duration = numeric(), lat_range = numeric(), mid_lat = numeric(), lon_range = numeric(), area = numeric(), depth_range = numeric(),  mid_depth = numeric())


#maps
####unique lat lon
ns_lat_lon <- unique(ns_grid.reduced_3plustows[,.(latitude, longitude)])

ns_pts <- st_as_sf(ns_lat_lon, coords=c('longitude','latitude'), crs=4326 )

ns_conc <- concaveman(ns_pts)

ns_area <- st_area(ns_conc) #m2

#row
ns <- c("ns","NA",length(unique(ns_grid.reduced_3plustows$accepted_name)),
               max(ns_grid.reduced_3plustows$year)-min(ns_grid.reduced_3plustows$year),
               length(unique(ns_grid.reduced_3plustows$year)),
               max(ns_grid.reduced_3plustows$latitude)-min(ns_grid.reduced_3plustows$latitude),
               mean(ns_grid.reduced_3plustows$latitude),
               max(ns_grid.reduced_3plustows$longitude)-min(ns_grid.reduced_3plustows$longitude),
               ns_area,
               max(ns_grid.reduced_3plustows$depth, na.rm = T)-min(ns_grid.reduced_3plustows$depth, na.rm = T),
               mean(ns_grid.reduced_3plustows$depth, na.rm = T)
               )



ns_stats <- data.table(rbind(ns))

colnames(ns_stats)<-c("region","season","spp_num","study_period","study_duration","lat_range","mid_lat","lon_range","area","depth_range","mid_depth")

saveRDS(ns_stats, here::here("output","region_stats","ns_stats.rds"))
```


Lists of Years
```{r year lists}
ns_grid.reduced_3plustows[,year:= as.numeric(year)] #make numeric
setorder(ns_grid.reduced_3plustows, year)

ns_years <- unique(ns_grid.reduced_3plustows[,year])

```

Haul ID keys
```{r list of haul_ids}
ns_haul_ids <- unique(ns_grid.reduced_3plustows[,haul_id])
ns_haul_ids_key <- data.table(haul_id = ns_haul_ids, key_ID = seq(1,length(ns_haul_ids), by = 1))

```

Convert haul_ids to numeric key_IDs
```{r convert haul id to numeric key id}

ns_grid.reduced_3plustows <- ns_grid.reduced_3plustows[ns_haul_ids_key, on = "haul_id"]

```


Dissimilarities across multiple years

```{r dissimilarities between cells across multiple years for spring}
library(betapart) 
ns_distances_dissimilarities_allyears <- data.table("haul_id1" = integer(), "haul_id2" = integer(), "distance" = numeric(),"bray_curtis_dissimilarity_balanced" = numeric(), year = integer(), "jaccard_dissimilarity_turnover" = numeric())

#Now loop through all years
for (i in 1:length(ns_years)) {
  reduced_year <- ns_grid.reduced_3plustows[year == ns_years[i],]
  
#cannot have wgt or num
  reduced_year <- reduced_year[!is.na(wgt_cpue),]
  
  #distances among cells
  setorder(reduced_year, key_ID)
  
  latitude_longitude_haul_id <- unique(reduced_year[,.(latitude,longitude,key_ID)])
  ns_distances <- distm(latitude_longitude_haul_id[,.(longitude,latitude)])
  key_IDs <- latitude_longitude_haul_id[,key_ID]

  colnames(ns_distances) <- rownames(ns_distances) <- key_IDs

  #wide to longitudeg
  haul_id_distances.l <- reshape2::melt(ns_distances,varnames = (c("haul_id1", "haul_id2")), value.name = "distance")
  
  #make into data table
  haul_id_distances.l <- data.table(haul_id_distances.l)

  
  reduced_year_wide <- dcast(reduced_year, key_ID + year ~ accepted_name, value.var = "wgt_cpue", fun.aggregate = sum) #longitude to wide data for community matrix, column names are cell then species
  
  
  ncols <- ncol(reduced_year_wide)
  communitymatrix <- reduced_year_wide[,3:ncols] #community matrix with year and cell on far right
  communitymatrix.occurence <- communitymatrix
  communitymatrix.occurence[communitymatrix.occurence > 0] <- 1

  #list of haul_id keys
  key_IDs_subset <- reduced_year_wide$key_ID

  dissimilarities_abundance <- beta.pair.abund(communitymatrix, index.family = "bray") #dissimilarity 
  dissimilarities_occurrence <- beta.pair(communitymatrix.occurence, index.family = "jaccard") #dissimilarity

  #make into matrix
  dissimilarities_abundance.m <- as.matrix(dissimilarities_abundance$beta.bray.bal, labels=TRUE) #bal = balanced
  dissimilarities_occurrence.m <- as.matrix(dissimilarities_occurrence$beta.jtu, labels=TRUE) #jtu = turnover
  colnames(dissimilarities_abundance.m) <- rownames(dissimilarities_abundance.m) <- key_IDs_subset
  colnames(dissimilarities_occurrence.m) <- rownames(dissimilarities_occurrence.m) <- key_IDs_subset

  #reshape dissimilarities
  dissimilarities_abundance.l <- reshape2::melt(dissimilarities_abundance.m, varnames = c("haul_id1", "haul_id2"), value.name = "bray_curtis_dissimilarity_balanced")
  dissimilarities_occurrence.l <- reshape2::melt(dissimilarities_occurrence.m, varnames = c("haul_id1", "haul_id2"), value.name = "jaccard_dissimilarity_turnover")
  dissimilarities_abundance.l <- data.table(dissimilarities_abundance.l) #and then to data table
  dissimilarities_occurrence.l <- data.table(dissimilarities_occurrence.l)

  #add year for these values
  dissimilarities_abundance.l[, "year" := ns_years[i]]
  dissimilarities_occurrence.l[, "year" := ns_years[i]]

  #merge distance with dissimilarity for this year with both metrics of dissimilarity
  dissimilarities_full <- haul_id_distances.l[dissimilarities_abundance.l, on = c("haul_id1", "haul_id2")]
  dissimilarities_full <- dissimilarities_full[dissimilarities_occurrence.l, on = c("haul_id1", "haul_id2", "year")]


  #add to data table
  ns_distances_dissimilarities_allyears <- rbind(ns_distances_dissimilarities_allyears, dissimilarities_full)
  
  print(paste0(i,"/",length(ns_years)))
  
}

summary(ns_distances_dissimilarities_allyears) #here we have bray, jaccard and geographic distance

#delete repeats
ns_distances_dissimilarities_allyears <- ns_distances_dissimilarities_allyears[haul_id1 >= haul_id2,]


ns_distances_dissimilarities_allyears[,bray_curtis_similarity_balanced := (1-bray_curtis_dissimilarity_balanced)][,jaccard_similarity_turnover := (1-jaccard_dissimilarity_turnover)]

saveRDS(ns_distances_dissimilarities_allyears, file = "ns_distances_dissimilarities_allyears.rds")

ns_distances_dissimilarities_allyears <- readRDS("ns_distances_dissimilarities_allyears.rds")

```

###Heat map for
Average within bins for plotting

#try rounding to 10s and 1s this time!
```{r subsample}
#jaccard only
ns_distances_dissimilarities_allyears_jaccard <- ns_distances_dissimilarities_allyears[,.(year,distance,jaccard_similarity_turnover, bray_curtis_similarity_balanced)][,distance_km := distance/1000]

#new column with rounded distance
ns_distances_dissimilarities_allyears_jaccard[,distance_km_10s := round_any(distance_km,10)][,distance_km_1s := round(distance_km,0)] #round to nearest 10, 1

#new column with median jaccard similarity & bray similarity means for 1 and 10 km
ns_distances_dissimilarities_allyears_jaccard[, jaccard_similarity_mean_10s := 
                                                       mean(jaccard_similarity_turnover), .(year, distance_km_10s)][, bray_similarity_mean_10s :=
                                                      mean(bray_curtis_similarity_balanced), .(year, distance_km_1s)][, jaccard_similarity_mean_1s := 
                                                       mean(jaccard_similarity_turnover), .(year, distance_km_1s)][, bray_similarity_mean_1s :=
                                                      mean(bray_curtis_similarity_balanced), .(year, distance_km_1s)]

#reduce to unique values
ns_distances_dissimilarities_allyears_jaccard_rounded <- unique(ns_distances_dissimilarities_allyears_jaccard, by = c("year","distance_km_10s","distance_km_1s"))
```

All possible combos
```{r all combos}
#all possible 1s
seq_1 <- seq(0,max(ns_distances_dissimilarities_allyears_jaccard_rounded$distance_km_1s), by = 1)

#all possible 10s
seq_10 <- seq(0,max(ns_distances_dissimilarities_allyears_jaccard_rounded$distance_km_10s), by = 10)

#datatable with all possible combos
ns_distances_dissimilarities_allyears_jaccard_allcombos <- data.table(expand.grid(year = ns_years, distance_km_1s = seq_1))

#combine
ns_distances_dissimilarities_allyears_jaccard_rounded_allcombos <- ns_distances_dissimilarities_allyears_jaccard_allcombos[ns_distances_dissimilarities_allyears_jaccard_rounded, on = c("year","distance_km_1s")]


#CUT OUT ALL COMBOS MORE THAN 500 km CHANGE LATER
#ns_distances_dissimilarities_allyears_jaccard_rounded_allcombos <- ns_distances_dissimilarities_allyears_jaccard_rounded_allcombos[distance_km <= 500,]
```

From "Depth-Time visualization using R, the tidyverse, and ggplot2" tutorial

```{r plot ns heat maps}

#convert year to factor
setkey(ns_distances_dissimilarities_allyears_jaccard_rounded_allcombos,year)
ns_distances_dissimilarities_allyears_jaccard_rounded_allcombos[,year.f := as.factor(year)]

#save this data table
saveRDS(ns_distances_dissimilarities_allyears_jaccard_rounded_allcombos, here::here("output","distance_decay","ns_distances_dissimilarities_allyears_jaccard_rounded_allcombos.rds"))


#jaccard similarity
(ns_jaccard_heatmap <- ggplot(ns_distances_dissimilarities_allyears_jaccard_rounded_allcombos,aes(as.factor(year),distance_km_1s, color = jaccard_similarity_mean_1s)) +
  geom_point(shape = 95, size = 1) +
  scale_color_gradientn(colors =
c("purple" ,"cornflowerblue", "green" , "yellow" ,"orange", "red")
  ) +
scale_x_discrete(limits=levels(ns_distances_dissimilarities_allyears_jaccard_rounded_allcombos$year.f), breaks = c("1983","1990","1995", "2000","2005", "2010","2015","2020")) +
  labs(y = "Distance (km)", x = "Year", color = "Jaccard Similarity") +
  theme_classic())

ggsave(plot = ns_jaccard_heatmap, path = here::here("figures","ns"),file = "ns_jaccard_heatmap.jpg")
ggsave(plot = ns_jaccard_heatmap, path = here::here("figures","ns"),file = "ns_jaccard_heatmap.eps")


#bray curtis similarity
(ns_bray_heatmap <- ggplot(ns_distances_dissimilarities_allyears_jaccard_rounded_allcombos,aes(as.factor(year),distance_km_1s, color = bray_similarity_mean_1s)) +
  geom_point(shape = 95, size = 1) +
  scale_color_gradientn(colors =
c("purple" ,"cornflowerblue", "green" , "yellow" ,"orange", "red")
  ) +
  scale_x_discrete(limits=levels(ns_distances_dissimilarities_allyears_jaccard_rounded_allcombos$year.f), breaks = c("1983","1990","1995", "2000","2005", "2010","2015","2020")) +
  labs(y = "Distance (km)", x = "Year", color = "Bray Curtis Similarity") +
  theme_classic())

ggsave(plot = ns_bray_heatmap, path = here::here("figures","ns"),file = "ns_bray_heatmap.jpg")
ggsave(plot = ns_bray_heatmap, path = here::here("figures","ns"),file = "ns_bray_heatmap.eps")
```
Box plots
```{r box plots}

#jaccard similarity
(ns_jaccard_boxplot <- ggplot(ns_distances_dissimilarities_allyears_jaccard_rounded_allcombos,aes(year.f,jaccard_similarity_turnover)) +
  geom_boxplot(outlier.shape = NA, lwd = 0.2) +
  labs(x="Year", y = "Jaccard Similarity") +
  scale_x_discrete(limits=levels(ns_distances_dissimilarities_allyears_jaccard_rounded_allcombos$year.f), breaks = c("1983","1990","1995", "2000","2005", "2010","2015","2020")) +
  geom_smooth(method = "lm", se=FALSE, color="red", aes(group=1), lwd = 0.3) +
  stat_regline_equation(label.x = 3, label.y = 32) +
 # ylim(c(0,0.75)) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90)))

#bray curtis similarity
(ns_bray_boxplot <- ggplot(ns_distances_dissimilarities_allyears_jaccard_rounded_allcombos,aes(year.f,bray_curtis_similarity_balanced)) +
  geom_boxplot(outlier.shape = NA, lwd = 0.2) +
  labs(x="Year", y = "Bray Curtis Similarity") +
  scale_x_discrete(limits=levels(ns_distances_dissimilarities_allyears_jaccard_rounded_allcombos$year.f), breaks = c("1983","1990","1995", "2000","2005", "2010","2015","2020")) +
  geom_smooth(method = "lm", se=FALSE, color="red", aes(group=1), lwd = 0.3) +
  stat_regline_equation(label.x = 3, label.y = 32) +
 #   ylim(c(0,0.48)) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90)))

#corresponding models
ns_avg_jaccard_mod <- lm(data = ns_distances_dissimilarities_allyears_jaccard_rounded_allcombos, formula = jaccard_similarity_turnover ~ as.numeric(year))

ns_avg_bray_curtis_mod <- lm(data = ns_distances_dissimilarities_allyears_jaccard_rounded_allcombos, formula = bray_curtis_similarity_balanced ~ as.numeric(year))

summary(ns_avg_jaccard_mod)
summary(ns_avg_bray_curtis_mod)
```

Now, merge the previous two plots
```{r merge plots}
#pull out legend
ns_legend_heatmaps <- get_legend(ns_jaccard_heatmap + labs(color = "Similarity"))

ns_merge <- plot_grid(ns_jaccard_heatmap + theme(legend.position = "none", axis.title.x = element_blank(), axis.text.x = element_blank()) ,
                             ns_bray_heatmap + theme(legend.position = "none", axis.title.x = element_blank(), axis.text.x = element_blank()),
                             ns_jaccard_boxplot + theme(axis.text.x = element_text(size = 6)), 
                             ns_bray_boxplot + theme(axis.text.x = element_text(size = 6)),
                             nrow = 2, ncol = 2, align = "hv", axis = "lb")

ns_merge

ggsave(ns_merge, path = here::here("figures","ns"), file = "ns_heat_boxplot_merge.jpg", width = 3.5, height = 4, unit = "in")
ggsave(ns_merge, path = here::here("figures","ns"), file = "ns_heat_boxplot_merge.eps", width = 3.5, height = 4, unit = "in")

```

Add extra columns with coefficients
```{r data summary with coef}

ns_stats[,jac_coef := 
           ns_avg_jaccard_mod$coefficients[[2]]][, jac_p := 
          summary(ns_avg_jaccard_mod)$coefficients[2,4]][,bray_coef :=
          ns_avg_bray_curtis_mod$coefficients[[2]]][,bray.p := 
          summary(ns_avg_bray_curtis_mod)$coefficients[2,4]]


saveRDS(ns_stats, here::here("output","region_stats","ns_stats.rds"))
```