---
title: "SA_summer_similarity_heatmap_distance"
output: html_notebook
---

```{r setup}
library(data.table)
library(viridis)
library(ggplot2)
library(reshape2)
library(geosphere)
library(vegan)
library(raster)
library(here)
library(cowplot)
library(plyr)

#load data
dat_SA_grid.reduced_3plustows_summer <- readRDS("~/Documents/grad school/Rutgers/Repositories/trawl_spatial_turnover/data/gridded/USA-SA/seus_grid.reduced_3plustows_summer.rds")

#distance among grid cells

sa_reg_distances.l <- readRDS("/Users/zoekitchel/Documents/grad school/Rutgers/Repositories/trawl_spatial_turnover/data/gridded/USA-SA/sa_reg_distances_summer.l.rds")

```
Lists of Years
```{r year lists}
dat_SA_grid.reduced_3plustows_summer[,year:= as.numeric(year)] #make numeric
setorder(dat_SA_grid.reduced_3plustows_summer, year)

sa_years <- unique(dat_SA_grid.reduced_3plustows_summer[,year])

```

Haul ID keys
```{r list of haul_ids}
sa_haul_ids <- unique(dat_SA_grid.reduced_3plustows_summer[,haul_id])
sa_haul_ids_key <- data.table(haul_id = sa_haul_ids, key_ID = seq(1,length(sa_haul_ids), by = 1))

```

Convert haul_ids to numeric key_IDs
```{r}

dat_SA_grid.reduced_3plustows_summer <- dat_SA_grid.reduced_3plustows_summer[sa_haul_ids_key, on = "haul_id"]

```


Dissimilarities across multiple years

```{r dissimilarities between cells across multiple years for summer}
sa_distances_dissimilarities_allyears_summer <- data.table("haul_id1" = integer(), "haul_id2" = integer(), "distance" = numeric(),"bray_curtis_dissimilarity" = numeric(), year = integer(), "jaccard_dissimilarity" = numeric())

#Now loop through all years
for (i in 1:length(sa_years)) {
  reduced_year <- dat_SA_grid.reduced_3plustows_summer[year == sa_years[i],]
  
    #cannot have wgt or num = NA
  reduced_year <- reduced_year[!is.na(wgt) & !is.na(num),]
  
  #distances among cells
  setorder(reduced_year, key_ID)
  
  lat_lon_haul_id <- unique(reduced_year[,.(latitude,longitude,key_ID)])
  sa_distances <- distm(lat_lon_haul_id[,.(longitude,latitude)])
  key_IDs <- lat_lon_haul_id[,key_ID]

  colnames(sa_distances) <- rownames(sa_distances) <- key_IDs

  #wide to long
  haul_id_distances.l <- reshape2::melt(sa_distances,varnames = (c("haul_id1", "haul_id2")), value.name = "distance")
  
  #make into data table
  haul_id_distances.l <- data.table(haul_id_distances.l)

  
  reduced_year_wide <- dcast(reduced_year, key_ID + year ~ accepted_name, value.var = "wgt", fun.aggregate = sum) #long to wide data for community matrix, column names are cell then species
  
  ncols <- ncol(reduced_year_wide)
  communitymatrix <- cbind(reduced_year_wide[,3:ncols], reduced_year_wide[,1:2]) #community matrix with year and cell on far right

  #list of haul_id keys
  key_IDs_subset <- communitymatrix$key_ID

  dissimilarities_abundance <- vegdist(communitymatrix[,1:(ncols-2)], method = "bray", binary = F) #dissimilarity 
  dissimilarities_occurrence <- vegdist(communitymatrix[,1:(ncols-2)], method = "jaccard", binary = T) #T binary performs presence absence standardization before using decostand

  #make into matrix
  dissimilarities_abundance.m <- as.matrix(dissimilarities_abundance, labels=TRUE)
  dissimilarities_occurrence.m <- as.matrix(dissimilarities_occurrence, labels=TRUE)
  colnames(dissimilarities_abundance.m) <- rownames(dissimilarities_abundance.m) <- key_IDs_subset
  colnames(dissimilarities_occurrence.m) <- rownames(dissimilarities_occurrence.m) <- key_IDs_subset

  #reshape dissimilarities
  dissimilarities_abundance.l <- reshape2::melt(dissimilarities_abundance.m, varnames = c("haul_id1", "haul_id2"), value.name = "bray_curtis_dissimilarity")
  dissimilarities_occurrence.l <- reshape2::melt(dissimilarities_occurrence.m, varnames = c("haul_id1", "haul_id2"), value.name = "jaccard_dissimilarity")
  dissimilarities_abundance.l <- data.table(dissimilarities_abundance.l) #and then to data table
  dissimilarities_occurrence.l <- data.table(dissimilarities_occurrence.l)

  #add year for these values
  dissimilarities_abundance.l[, "year" := sa_years[i]]
  dissimilarities_occurrence.l[, "year" := sa_years[i]]

  #merge distance with dissimilarity for this year with both metrics of dissimilarity
  dissimilarities_full <- haul_id_distances.l[dissimilarities_abundance.l, on = c("haul_id1", "haul_id2")]
  dissimilarities_full <- dissimilarities_full[dissimilarities_occurrence.l, on = c("haul_id1", "haul_id2", "year")]


  #add to data table
  sa_distances_dissimilarities_allyears_summer <- rbind(sa_distances_dissimilarities_allyears_summer, dissimilarities_full)
  
}

summary(sa_distances_dissimilarities_allyears_summer) #here we have bray, jaccard and geographic distance

#delete repeats
sa_distances_dissimilarities_allyears_summer <- sa_distances_dissimilarities_allyears_summer[haul_id1 >= haul_id2,] #3165272 to 1588479 rows


sa_distances_dissimilarities_allyears_summer[,bray_curtis_similarity := (1-bray_curtis_dissimilarity)][,jaccard_similarity := (1-jaccard_dissimilarity)]

saveRDS(sa_distances_dissimilarities_allyears_summer, file = "sa_distances_dissimilarities_allyears_summer.rds")

sa_distances_dissimilarities_allyears_summer <- readRDS("sa_distances_dissimilarities_allyears_summer.rds")

```

#Heat map

Subsample for plotting

#try rounding to 50s this time!
```{r subsample}
#jaccard only
sa_distances_dissimilarities_allyears_summer_jaccard <- sa_distances_dissimilarities_allyears_summer[,.(year,distance,jaccard_similarity, bray_curtis_similarity)][,distance_km := distance/1000]

#new column with rounded distance
sa_distances_dissimilarities_allyears_summer_jaccard[,distance_km_50s := round_any(distance_km,50)][,distance_km_10s := round_any(distance_km,10)] #round to nearest 10

#new column with median jaccard similarity & bray similarity means
sa_distances_dissimilarities_allyears_summer_jaccard[, jaccard_similarity_mean_10s := mean(jaccard_similarity), .(year, distance_km_10s)][, bray_similarity_mean_10s := mean(bray_curtis_similarity), .(year, distance_km_10s)]

#reduce to unique values
sa_distances_dissimilarities_allyears_summer_jaccard_rounded <- unique(sa_distances_dissimilarities_allyears_summer_jaccard, by = c("year","distance_km_10s"))
```


All possible combos
```{r all combos}
#all possible 10s
seq_10 <- seq(0,max(sa_distances_dissimilarities_allyears_summer_jaccard_rounded$distance_km_10s), by = 10)

#all possiblew 50s
seq_50 <- seq(0,max(sa_distances_dissimilarities_allyears_summer_jaccard_rounded$distance_km_10), by = 50)

#datatable with all possible combos
sa_distances_dissimilarities_allyears_summer_jaccard_allcombos <- data.table(expand.grid(year = sa_years, distance_km_10s = seq_10))

#combine
sa_distances_dissimilarities_allyears_summer_jaccard_rounded_allcombos <- sa_distances_dissimilarities_allyears_summer_jaccard_allcombos[sa_distances_dissimilarities_allyears_summer_jaccard_rounded, on = c("year","distance_km_10s")]
```


Trying to use ggplot
```{r}
ggplot(sa_distances_dissimilarities_allyears_jaccard_rounded_allcombos, aes(year, distance_10s, fill = jaccard_similarity_mean)) +
  geom_raster(interpolate = T) +
  theme_classic() +
  scale_fill_viridis(discrete = F)

library(sp)

```

Convert to matrix
```{r jaccard matrix}
sa_distances_dissimilarities_allyears_jaccard_wide <-
data.table::dcast(sa_distances_dissimilarities_allyears_jaccard_rounded[,.(year,distance_10s,jaccard_similarity_mean)], distance_10s ~ year, fill = NA)

setkey(sa_distances_dissimilarities_allyears_jaccard_wide, distance_10s)

#get rid of first column
sa_distances_dissimilarities_allyears_jaccard_wide.m <- sa_distances_dissimilarities_allyears_jaccard_wide[,2:ncol(sa_distances_dissimilarities_allyears_jaccard_wide)]

sa_distances_dissimilarities_allyears_jaccard_wide.m <- as.matrix(sa_distances_dissimilarities_allyears_jaccard_wide.m)

sa_distances_dissimilarities_allyears_jaccard_wide.r <- as.raster (sa_distances_dissimilarities_allyears_jaccard_wide.m)
```

Convert comprehensive table to matrix
```{r}
sa_distances_dissimilarities_allyears_jaccard_rounded_allcombos_wide <- data.table::dcast(sa_distances_dissimilarities_allyears_jaccard_rounded_allcombos[,.(year,distance_10s,jaccard_similarity_mean)], distance_10s ~ year, fill = NA)

setkey(sa_distances_dissimilarities_allyears_jaccard_rounded_allcombos_wide, distance_10s)

#get rid of first column
sa_distances_dissimilarities_allyears_jaccard_rounded_allcombos_wide.m <- sa_distances_dissimilarities_allyears_jaccard_rounded_allcombos_wide[,2:ncol(sa_distances_dissimilarities_allyears_jaccard_rounded_allcombos_wide)]

sa_distances_dissimilarities_allyears_jaccard_rounded_allcombos_wide.m <- as.matrix(sa_distances_dissimilarities_allyears_jaccard_rounded_allcombos_wide.m)

sa_distances_dissimilarities_allyears_jaccard_rounded_allcombos_wide.r <- as.raster (sa_distances_dissimilarities_allyears_jaccard_rounded_allcombos_wide.m)
```

Plot raster
```{r}
image(sa_distances_dissimilarities_allyears_jaccard_rounded_allcombos_wide.r)
```
From "Depth-Time visualization using R, the tidyverse, and ggplot2" tutorial

```{r plot heat maps}

#convert year to factor
setkey(sa_distances_dissimilarities_allyears_summer_jaccard_rounded_allcombos,year)
sa_distances_dissimilarities_allyears_summer_jaccard_rounded_allcombos[,year.f := as.factor(year)]

sa_distances_dissimilarities_allyears_summer[,year.f := as.factor(year)]

#jaccard similarity
(sa_jaccard_heatmap <- ggplot(sa_distances_dissimilarities_allyears_summer_jaccard_rounded_allcombos,aes(as.factor(year),distance_km_10s, color = jaccard_similarity_mean_10s)) +
  geom_point(shape = "-", size = 5) +
  scale_color_gradientn(colors =
c("purple" ,"cornflowerblue", "green" , "yellow" ,"orange", "red")
  ) +
scale_x_discrete(limits=levels(sa_distances_dissimilarities_allyears_summer_jaccard_rounded_allcombos$year.f), breaks = c("1989","1995", "2000", "2005", "2010","2015")) +
  labs(y = "Distance (km)", x = "Year", color = "Jaccard Similarity") +
  theme_classic())

ggsave(plot = sa_jaccard_heatmap, path = here::here("figures","sa"),file = "sa_jaccard_heatmap.jpg")
ggsave(plot = sa_jaccard_heatmap, path = here::here("figures","sa"),file = "sa_jaccard_heatmap.eps")


#bray curtis similarity
(sa_bray_heatmap <- ggplot(sa_distances_dissimilarities_allyears_summer_jaccard_rounded_allcombos,aes(as.factor(year),distance_km_10s, color = bray_similarity_mean_10s)) +
  geom_point(shape = "-", size = 5) +
  scale_color_gradientn(colors =
c("purple" ,"cornflowerblue", "green" , "yellow" ,"orange", "red")
  ) +
  scale_x_discrete(limits=levels(sa_distances_dissimilarities_allyears_summer_jaccard_rounded_allcombos$year.f), breaks = c("1989","1995", "2000", "2005", "2010","2015")) +
  labs(y = "Distance (km)", x = "Year", color = "Bray Curtis Similarity") +
  theme_classic())

ggsave(plot = sa_bray_heatmap, path = here::here("figures","sa"),file = "sa_bray_heatmap.jpg")
ggsave(plot = sa_bray_heatmap, path = here::here("figures","sa"),file = "sa_bray_heatmap.eps")
```
Box plots
```{r summer box plots}

#jaccard similarity
(sa_jaccard_boxplot <- ggplot(sa_distances_dissimilarities_allyears_summer,aes(year.f,jaccard_similarity)) +
  geom_boxplot(outlier.shape = NA, lwd = 0.2) +
  labs(x="Year", y = "Jaccard Similarity") +
  scale_x_discrete(limits=levels(sa_distances_dissimilarities_allyears_summer$year.f), breaks = c("1989","1995", "2000", "2005", "2010","2015")) +
  geom_smooth(method = "lm", se=FALSE, color="red", aes(group=1), lwd = 0.3) +
  stat_regline_equation(label.x = 3, label.y = 32) +
  ylim(c(0,0.68)) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90)))

#bray curtis similarity
(sa_bray_boxplot <- ggplot(sa_distances_dissimilarities_allyears_summer,aes(year.f,bray_curtis_similarity)) +
  geom_boxplot(outlier.shape = NA, lwd = 0.2) +
  labs(x="Year", y = "Bray Curtis Similarity") +
  scale_x_discrete(limits=levels(sa_distances_dissimilarities_allyears_summer$year.f), breaks = c("1989","1995", "2000", "2005", "2010","2015")) +
  geom_smooth(method = "lm", se=FALSE, color="red", aes(group=1), lwd = 0.3) +
  stat_regline_equation(label.x = 3, label.y = 32) +
    ylim(c(0,0.48)) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90)))

#corresponding models
sa_avg_jaccard_mod <- lm(data = sa_distances_dissimilarities_allyears_summer, formula = jaccard_similarity ~ as.numeric(year))

sa_avg_bray_curtis_mod <- lm(data = sa_distances_dissimilarities_allyears_summer, formula = bray_curtis_similarity ~ as.numeric(year))

summary(sa_avg_jaccard_mod)
summary(sa_avg_bray_curtis_mod)
```

Now, merge the previous two plots
```{r merge plots}
#pull out legend
sa_legend_heatmaps <- get_legend(sa_jaccard_heatmap + labs(color = "Similarity"))

sa_merge_summer <- plot_grid(sa_jaccard_heatmap + theme(legend.position = "none", axis.title.x = element_blank(), axis.text.x = element_blank()) , sa_bray_heatmap + theme(legend.position = "none", axis.title.x = element_blank(), axis.text.x = element_blank(), axis.title.y = element_blank()), sa_jaccard_boxplot, sa_bray_boxplot,  nrow = 2, ncol = 2, align = "hv", axis = "lb")

sa_merge_summer

ggsave(sa_merge_summer, path = here::here("figures","sa"), file = "sa_heat_boxplot_merge_summer.jpg", width = 5, height = 5, unit = "in")
ggsave(sa_merge_summer, path = here::here("figures","sa"), file = "sa_heat_boxplot_merge_summer.eps", width = 5, height = 5, unit = "in")

```


###maybe i'll do this later? maybe not
At its heart, this graphic is just a ggplot() with a geom_raster() layer. The complication is, geom_raster() requires equally spaced points in both the x- and y-direction. The rest of this post is about how to make that happen.

Linear Interpolation

I interpolate in the distance direction first, because I think this is the better assumption: as you go down further away, a reasonable way to estimate the similarity at a distance which you did not measure is to draw a straight line between the similarity at the distance that you did measure. (Not great, but sure)
```{r}
estimate_jac_sim_by_year <- function(target_year, target_distance) {
  data_for_year <- sa_distances_dissimilarities_allyears[year == target_year]
  
  # approx() is one way to do a linear interpolation
  approx(data_for_year$distance, data_for_year$jaccard_similarity, xout = target_distance)$y
}

estimate_jac_sim_by_year(1990, c(0,3000,300000))
```
Expand data inputs
```{r}
temp_interp_depth <- data.table(expand.grid(
  # the same years as sa summer
  year = sa_years,
  # ddistances can now be any value
  distance = seq(0, max(sa_distances_dissimilarities_allyears$distance), by = 10))
)

temp_interp_depth[,jaccard_similarity_modeled := estimate_jac_sim_by_year(year,distance), year]

```
Visualize
```{r}
ggplot(temp_interp_depth,aes(year,distance, color = jaccard_similarity_modeled)) +
  geom_point() +
   scale_colour_gradient2(
    midpoint = 0.5, 
    high = scales::muted("red"), 
    low = scales::muted("blue")
  )
```

First, we write a function that estimates the temperature at any date given a depth that is in temp_interp_depth (the tibble we just calculated).

```{r}
estimate_jac_sim_by_distance <- function(target_distance, target_year) {
  data_for_distance <- temp_interp_depth[distance == target_distance,]
    setkey(data_for_distance, year)
  approx(data_for_distance$year, data_for_distance$jaccard_similarity_modeled, xout = target_year)$y
}

estimate_jac_sim_by_distance(
  target_distance = 1000000, 
  target_year = c(1990,1998,2000)
)
```
Expand year grid
```{r}
jac_sim_raster <- data.table(expand.grid(
  # dates can now be any value
  year = seq(min(sa_years), max(sa_years), by = 1),
  # depths must be the same as in temp_interp_depth
  distance = unique(temp_interp_depth$distance))
)

jac_sim_raster[,jaccard_similarity_modeled := estimate_jac_sim_by_distance(distance,year),by = distance]


```
Finally, we have equally-spaced values in both the date and depth dimensions. This can be visualized using geom_raster(), with temp mapped to the fill aesthetic. I’ve again used scale_fill_gradient2() to ensure that red values represent “hot”, blue values represent “cool”, and that there is some way to visualize the depth of the thermocline. Finally, I’ve used coord_cartesian(expand = FALSE) to eliminate the white border around the outside of the raster layer…I think it looks nicer that way.


```{r final jaccard similarity plot}
ggplot(jac_sim_raster, aes(year, distance, fill = jaccard_similarity_modeled)) +
  geom_raster() +
  scale_fill_gradientn(colors =
c("purple" ,"blue"  , "green" , "yellow" ,"orange", "red")
  ) +
  coord_cartesian(expand = FALSE)
```

                 