---
title: "Prepping North Sea (ICES) from DATRAS data"
output: html_notebook
---


```{r setup}
library(dggridR)
library(data.table)
library(rgdal)
library(raster)
library(sp)
library(rnaturalearth)
library(rnaturalearthdata)
library(rgeos)
library(taxize) #standardizing names
library(geosphere)  #to calculate distance between lat lon of grid cells
library(googledrive)
```


Moving forward, I plan to use data directly from FishGlob.

December 1, 2021
```{r pull in data from fishglob for NS from Google Drive}
drive_download("DATRAS_v2_clean.csv",
               overwrite = T)

DATRAS <- fread("DATRAS_v2_clean.csv")

file.remove("DATRAS_v2_clean.csv")

ns <- DATRAS[survey  == "NS-IBTS"]

```

What seasons do we have? None

```{r seasons?}
summary(as.factor(ns$season))


```

#NS
```{r pull in NS}

# Get Unique lines in the data table
unique_ns_latlon <- unique(ns[,.(latitude, longitude)])
unique_ns_latlon.sf <- unique_ns_latlon
  
#make lat lon coordinates into polygon
coordinates(unique_ns_latlon.sf) <- ~latitude + longitude
unique_ns_latlon.sf@proj4string <- crs("+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0")

saveRDS(unique_ns_latlon, "unique_ns_latlon.rds")

#plot north america
world <- ne_countries(scale = "medium", returnclass = "sp")

world_ext <- extent(-4, 13,49, 62)

europe_spdf <- crop(world, world_ext)

europe_df <- fortify(europe_spdf)

(europe_outline <- ggplot(europe_df, aes(long, lat, group = group)) +
  geom_polygon(color = "white", fill = "grey") +
  labs(x = "Longitude", y = "Latitude") +
  coord_equal() +
  theme_classic() +
  scale_x_continuous(expand = c(0,0)) +
  scale_y_continuous(expand = c(0,0)) +
  theme(panel.border = element_rect(color = "black", fill = NA, size = 1)) +
 geom_point(unique_ns_latlon, mapping = aes(longitude, latitude), inherit.aes = FALSE, shape = 18, size = 0.001, color = "red") +
  coord_equal())

```

Playing around with dggridR (package that makes grid cells)
```{r learning dggridR}
dggs <- dgconstruct(spacing = 111, metric = T, resround = 'nearest')

unique_ns_latlon[,cell := dgGEO_to_SEQNUM(dggs, longitude, latitude)] #get corresponding grid cells for NEUS trawl

#centersof cells
ns_cellcenters <- dgSEQNUM_to_GEO(dggs, unique_ns_latlon[,cell])

#linking cell centers to unique_ns_latlon
unique_ns_latlon[,LON_CENTER := ns_cellcenters$lon_deg][,LAT_CENTER := ns_cellcenters$lat_deg]

#link centers back to main data table

ns <- merge(ns, unique_ns_latlon, by = c("latitude", "longitude"))

#number of tows in each cell
towcount <- unique_ns_latlon[, .N, by = cell]

#get the grid cell boundary for cells which had trawls
grid_ns <- dgcellstogrid(dggs, unique_ns_latlon[,cell], frame = T, wrapcells = F)

#update grid properties to include # of trawls in each cell
grid_ns <- merge(grid_ns, unique_ns_latlon, by = "cell")

ggplot(data = grid_ns, aes(x = long, y = lat, group = cell)) +
  geom_polygon() +
  coord_quickmap()

saveRDS(grid_ns, "grid_ns.rds")

```

Trying to plot with grid overlaying
```{r grid overlaying}

europe_outline + 
   geom_polygon(grid_ns, mapping = aes(x = long,y = lat, group = cell), inherit.aes = FALSE, color = "black", fill = NA, size = 0.1)

ggsave(filename = "grid_NS_cells_spacing_111.pdf")
  
```
Cleaning spp names (must be ID'd to species!)
```{r clean NS names to species}
#unique species names
ns.clean <- ns[rank == "Species"]


```


Now, we'll overlay grid cells onto this list of tows (dat_ns) using lat and lon and gridrr package
```{r get rid of underused grid cells}

year_grid <- data.table(table(ns.clean$cell, ns.clean$year))

#there are some grid cells with infrequent records, I will delete these BUT IF I DELETE THESE THEY ALL DELETE, SO I WILL START TIME SERIES IN 1983 FIRST
year_grid <- year_grid[V2 >= 1983,] #Juliette Murgier, Matthew McLean, Anthony Maire, David Mouillot, Nicolas Loiseau, François Munoz, Cyrille Violle and Arnaud Auber

grid_cells_to_exclude <- unique(year_grid[N == 0]$V1)

#delete these grid cells from full data table
ns.reduced <- ns.clean[!cell %in% grid_cells_to_exclude]

#319715 observations left, this should be great

#how many tows (TOW) per cell in a given year

haul_cell_unique <- unique(ns.reduced[,.(haul_id, cell, year)])

ns_samplingevents_byyear <- haul_cell_unique[,.N, .(year, cell)]

ns_cells_lessthan3 <- ns_samplingevents_byyear[N<3,]

cells_to_delete <- unique(ns_cells_lessthan3[,cell])

ns.reduced.again <- ns.reduced[!cell %in% cells_to_delete]

#down to 29147 observations

saveRDS(ns.reduced.again, "ns.reduced.again.rds")

```

Find distance between center of each grid cell
```{r distance between grid cells}
#just need grid cell # and center lat lon for each grid cell 
ns_grid_cells <- unique(ns.reduced.again[,.(cell, LON_CENTER, LAT_CENTER)])

ns_grid_cells <- setorder(ns_grid_cells, cell)

ns_reg_distances <- distm(ns_grid_cells[,.(LON_CENTER, LAT_CENTER)])

colnames(ns_reg_distances) <- ns_grid_cells$cell
rownames(ns_reg_distances) <- ns_grid_cells$cell

#reorient to long form 
ns_reg_distances.l <- melt(as.matrix(ns_reg_distances), varnames = c("cell1", "cell2"), value.name = "distance(m)") #matrix to data frame
ns_reg_distances.l <- data.table(ns_reg_distances.l) #and then to data table
ns_reg_distances.l <- ns_reg_distances.l[cell1 > cell2,] # get rid of repetitions and self-comparisons
saveRDS(ns_reg_distances.l, "ns_reg_distances.l.rds")
```

CHECK CLAIM THAT #no years have way more tows than others, because this isn't true

Check the # of tows per grid cell/year combo
```{r number of tows per grid cell/year/region combo for}
#cell/year/region combos
ns_unique_tows_cell <- ns.reduced[,.SD[1],.(year, cell, haul_id)]

#no years have way more tows than others
plot(ns_unique_tows_cell[,.N,.(year,cell)]$year, ns_unique_tows_cell[,.N,.(year,cell)]$N)

#therefore, I will not leave out any years, but rather just pull out cells in years they are undersampled (<N=3, which is )

summary(ns_unique_tows_cell[,.N,.(year,cell)]$N < 3) #this occurs in 447/2811 cell/year combos  (16%)

#I don't think I need to exclude in ALL years, just in a few years
#an alternative would be to assess coverage using biomass (# of singletons and doubletons)

```

Delete all observations with less than 3 tows in a given year/cell AFTER 1990
```{r delete observations with less than 3 tows}
ns_counts <- ns_unique_tows_cell[,.N,.(year,cell)]
ns_keep <- ns_counts[N>2,][,N := NULL]

ns_grid.reduced_3plustows <- ns.reduced[ns_keep, on = .(cell, year)][year >= 1983]


```

Make a new map but with only points we actually use
```{r new map points we use}
(europe_outline_2 <- ggplot(europe_df, aes(long, lat, group = group)) +
  geom_polygon(color = "white", fill = "grey") +
  labs(x = "Longitude", y = "Latitude") +
  coord_equal() +
  theme_classic() +
  scale_x_continuous(expand = c(0,0)) +
  scale_y_continuous(expand = c(0,0)) +
  theme(panel.border = element_rect(color = "black", fill = NA, size = 1)) +
 geom_point(unique(ns_grid.reduced_3plustows[,.(latitude, longitude)]), mapping = aes(longitude, latitude), inherit.aes = FALSE, shape = 18, size = 0.001, color = "red") +
  coord_equal())
```


Save
```{r save}
saveRDS(ns.reduced, "/Users/zoekitchel/Documents/grad school/Rutgers/Repositories/trawl_spatial_turnover/data/gridded/ICES-NS/ns.reduced.rds")


saveRDS(ns_reg_distances.l, "/Users/zoekitchel/Documents/grad school/Rutgers/Repositories/trawl_spatial_turnover/data/gridded/ICES-NS/ns_reg_distances.l.rds")

saveRDS(ns_grid.reduced_3plustows,"/Users/zoekitchel/Documents/grad school/Rutgers/Repositories/trawl_spatial_turnover/data/gridded/ICES-NS/ns_grid.reduced_3plustows.rds")


```
