---
title: "Prepping Southeast US Data"
output: html_notebook
---


```{r setup}
library(dggridR)
library(data.table)
library(rgdal)
library(raster)
library(sp)
library(rnaturalearth)
library(rnaturalearthdata)
library(rgeos)
library(taxize) #standardizing names
library(geosphere)  #to calculate distance between lat lon of grid cells
library(googledrive)
```


Moving forward, I plan to use data directly from FishGlob.

November 21, 2021
```{r pull in data from fishglob for seus from Google Drive}
drive_download("SEUS_clean.csv",
               overwrite = T)

seus <- fread("SEUS_clean.csv")

file.remove("SEUS_clean.csv")

```


```{r pull in SEUS}

# Get Unique lines in the data table
unique_sa_latlon<- unique(seus[,.(latitude, longitude,season)])
unique_sa_latlon.sf <- unique_sa_latlon
  
#make lat lon coordinates into polygon
coordinates(unique_sa_latlon.sf) <- ~latitude + longitude
unique_sa_latlon.sf@proj4string <- crs("+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0")

saveRDS(unique_sa_latlon, "unique_sa_latlon.rds")

#plot north america
world <- ne_countries(scale = "medium", returnclass = "sp")

world_ext <- extent(-85, -75,23, 36)

north_america_spdf <- crop(world, world_ext)

north_america_df <- fortify(north_america_spdf)

(na_outline <- ggplot(north_america_df, aes(long, lat, group = group)) +
  geom_polygon(color = "white", fill = "grey") +
  labs(x = "Longitude", y = "Latitude") +
  coord_equal() +
  theme_classic() +
  scale_x_continuous(expand = c(0,0)) +
  scale_y_continuous(expand = c(0,0)) +
  theme(panel.border = element_rect(color = "black", fill = NA, size = 1)) +
 geom_point(unique_sa_latlon[season == "fall"], mapping = aes(longitude, latitude), inherit.aes = FALSE, size = 0.5, color = "red") +
#  facet_wrap(~season, nrow = 3, ncol = 2) +
  coord_equal())

```

Playing around with dggridR (package that makes grid cells)
```{r learning dggridR}
dggs <- dgconstruct(spacing = 111, metric = T, resround = 'nearest')

unique_sa_latlon[,cell := dgGEO_to_SEQNUM(dggs, longitude, latitude)] #get corresponding grid cells for NEUS trawl

#centersof cells
sa_cellcenters <- dgSEQNUM_to_GEO(dggs, unique_sa_latlon[,cell])

#linking cell centers to unique_sa_latlon
unique_sa_latlon[,LON_CENTER := sa_cellcenters$lon_deg][,LAT_CENTER := sa_cellcenters$lat_deg]

#link centers back to main data table

seus <- merge(seus, unique_sa_latlon, by = c("latitude", "longitude","season"))

#number of tows in each cell
towcount <- unique_sa_latlon[, .N, by = cell]

#get the grid cell boundary for cells which had trawls
grid_sa <- dgcellstogrid(dggs, unique_sa_latlon[,cell], frame = T, wrapcells = F)

#update grid properties to include # of trawls in each cell
grid_sa <- merge(grid, unique_sa_latlon, by="cell")

ggplot(data = grid_sa, aes(x = long, y = lat, group = cell)) +
  geom_polygon() +
  coord_quickmap()

saveRDS(grid_sa, "grid_sa.rds")

```

Trying to plot with grid overlaying
```{r grid overlaying}

na_outline + 
   geom_polygon(grid_sa, mapping = aes(x = long,y = lat, group = cell), inherit.aes = FALSE, color = "black", fill = NA, size = 0.1)

ggsave(filename = "grid_SEUS_cells_spacing_111.pdf")
  
```
Cleaning spp names (must be ID'd to species!)
```{r clean SEUS names}
#unique species names
seus.clean <- seus[rank == "Species",]


```

FOR NOW JUST SUMMER

Now, we'll overlay grid cells onto this list of tows (dat_sa) using lat and lon and gridrr package
```{r get rid of underused grid cells}
seus.clean.summer <- seus.clean[season == "summer"]
  
year_grid <- data.table(table(seus.clean.summer$cell, seus.clean.summer$year))

#there are some grid cells with infrequent records, I will delete these
grid_cells_to_exclude <- unique(year_grid[N == 0]$V1)

#delete these grid cells from full data table
dat_sa.reduced <- seus.clean.summer[!cell %in% grid_cells_to_exclude]

#362516 observations left, this should be great

#how many tows (TOW) per cell in a given year

haul_cell_unique <- unique(dat_sa.reduced[,.(haul_id, cell, year)])

sa_samplingevents_byyear <- haul_cell_unique[,.N, .(year, cell)]

sa_cells_lessthan3 <- sa_samplingevents_byyear[N<3,]

cells_to_delete <- unique(sa_cells_lessthan3[,cell])

dat_sa.reduced.again <- dat_sa.reduced[!cell %in% cells_to_delete]

#down to 734608 observations

saveRDS(dat_sa.reduced.again, "sa_data_gridded.rds")

```

Find distance between center of each grid cell
```{r distance between grid cells}
#just need grid cell # and center lat lon for each grid cell 
sa_grid_cells <- unique(dat_sa.reduced.again[,.(cell, LON_CENTER, LAT_CENTER)])

sa_grid_cells <- setorder(sa_grid_cells, cell)

sa_reg_distances <- distm(sa_grid_cells[,.(LON_CENTER, LAT_CENTER)])

colnames(sa_reg_distances) <- sa_grid_cells$cell
rownames(sa_reg_distances) <- sa_grid_cells$cell

#reorient to long form 
sa_reg_distances.l <- melt(as.matrix(sa_reg_distances), varnames = c("cell1", "cell2"), value.name = "distance(m)") #matrix to data frame
sa_reg_distances.l <- data.table(sa_reg_distances.l) #and then to data table
sa_reg_distances.l <- sa_reg_distances.l[cell1 > cell2,] # get rid of repetitions and self-comparisons
saveRDS(sa_reg_distances.l, "sa_reg_distances.l.rds")
```

Check the # of tows per grid cell/year combo
```{r number of tows per grid cell/year/region combo}
#cell/year/region combos
sa_unique_tows_cell <- dat_sa.reduced[,.SD[1],.(year, cell, haul_id)]

#no years have way more tows than others
plot(sa_unique_tows_cell[,.N,.(year,cell)]$year, sa_unique_tows_cell[,.N,.(year,cell)]$N)

#therefore, I will not leave out any years, but rather just pull out cells in years they are undersampled (<N=3, which is )

summary(sa_unique_tows_cell[,.N,.(year,cell)]$N < 3) #this occurs in 1/389 cell/year combos  0.2%

#I don't think I need to exclude in ALL years, just in a few years
#an alternative would be to assess coverage using biomass (# of singletons and doubletons)

```

Delete all observations with less than 3 tows in a given year/cell
```{r delete observations with less than 3 tows}
sa_counts <- sa_unique_tows_cell[,.N,.(year,cell)]
sa_keep <- sa_counts[N>2,][,N := NULL]

dat_sa_grid.reduced_3plustows <- dat_sa.reduced[sa_keep, on = .(cell, year)]


```

Summary of area sampled
```{r sum area sampled}
#distance between furthest north and furthest south point
library(geosphere)
max_north_south_km <- (distm(c(mean(dat_sa_grid.reduced_3plustows$longitude), min(dat_sa_grid.reduced_3plustows$latitude)), c(mean(dat_sa_grid.reduced_3plustows$longitude), max(dat_sa_grid.reduced_3plustows$latitude)), fun = distHaversine))/1000
#distance between furthest east and furthest west point
max_west_east_km <- (distm(c(min(dat_sa_grid.reduced_3plustows$longitude), mean(dat_sa_grid.reduced_3plustows$latitude)), c(max(dat_sa_grid.reduced_3plustows$longitude), mean(dat_sa_grid.reduced_3plustows$latitude)), fun = distHaversine))/1000

max_north_south_km
max_west_east_km

summary(dat_sa_grid.reduced_3plustows$depth)

length(unique(dat_sa_grid.reduced_3plustows$accepted_name))

sort(unique(as.numeric(seus$year)))
length(unique(seus$year))
```

Save
```{r save}
saveRDS(dat_sa.reduced, "/Users/zoekitchel/Documents/grad school/Rutgers/Repositories/trawl_spatial_turnover/data/gridded/USA-SA/seus_grid_summer.reduced.rds")


saveRDS(sa_reg_distances.l, "/Users/zoekitchel/Documents/grad school/Rutgers/Repositories/trawl_spatial_turnover/data/gridded/USA-SA/sa_reg_distances_summer.l.rds")

saveRDS(dat_sa_grid.reduced_3plustows,"/Users/zoekitchel/Documents/grad school/Rutgers/Repositories/trawl_spatial_turnover/data/gridded/USA-SA/seus_grid.reduced_3plustows_summer.rds")


```
