---
title: "All Region Similarity Trends Time"
output: html_notebook
---

Here, we will make a figure where I plot pairwise similarity trends over time for all regions we're examining right now
```{r setup}
library(data.table)
library(ggplot2)

#pull in similarity and year for each region ('REGION'_distances_dissimilarities_allyears_jaccard_rounded_allcombos)

#gmex fall

gmex_fall_distances_dissimilarities_allyears_jaccard_rounded_allcombos <- readRDS(here::here("output","distance_decay","gmex_fall_distances_dissimilarities_allyears_jaccard_rounded_allcombos.rds"))

gmex_fall_distances_dissimilarities_allyears_jaccard_rounded_allcombos[,region := "gmex"][,season := "fall"]

#gmex summer
gmex_summer_distances_dissimilarities_allyears_jaccard_rounded_allcombos <- readRDS(here::here("output","distance_decay","gmex_fall_distances_dissimilarities_allyears_jaccard_rounded_allcombos.rds"))

gmex_summer_distances_dissimilarities_allyears_jaccard_rounded_allcombos[,region := "gmex"][,season := "summer"]

#neus fall

neus_fall_distances_dissimilarities_allyears_jaccard_rounded_allcombos <- readRDS(here::here("output","distance_decay","neus_fall_distances_dissimilarities_allyears_jaccard_rounded_allcombos.rds"))

neus_fall_distances_dissimilarities_allyears_jaccard_rounded_allcombos[,year.f := as.factor(year)][,region := "neus"][,season := "fall"]

#neus spring
neus_spring_distances_dissimilarities_allyears_jaccard_rounded_allcombos <- readRDS(here::here("output","distance_decay","neus_fall_distances_dissimilarities_allyears_jaccard_rounded_allcombos.rds"))

neus_spring_distances_dissimilarities_allyears_jaccard_rounded_allcombos[,year.f := as.factor(year)][,region := "neus"][,season := "spring"]

#swc
swc_distances_dissimilarities_allyears_jaccard_rounded_allcombos <- readRDS(here::here("output","distance_decay","swc_distances_dissimilarities_allyears_jaccard_rounded_allcombos.rds"))

swc_distances_dissimilarities_allyears_jaccard_rounded_allcombos[,region := "swc"][,season := "NA"]

#ns
ns_distances_dissimilarities_allyears_jaccard_rounded_allcombos <- readRDS(here::here("output","distance_decay","ns_distances_dissimilarities_allyears_jaccard_rounded_allcombos.rds"))

ns_distances_dissimilarities_allyears_jaccard_rounded_allcombos[,region := "ns"][,season := "NA"]

#pt
pt_distances_dissimilarities_allyears_jaccard_rounded_allcombos <- readRDS(here::here("output","distance_decay","pt_distances_dissimilarities_allyears_jaccard_rounded_allcombos.rds"))

pt_distances_dissimilarities_allyears_jaccard_rounded_allcombos[,region := "pt"][,season := "NA"]

#ebs
ebs_distances_dissimilarities_allyears_jaccard_rounded_allcombos <- readRDS(here::here("output","distance_decay","ebs_distances_dissimilarities_allyears_jaccard_rounded_allcombos.rds"))
ebs_distances_dissimilarities_allyears_jaccard_rounded_allcombos[,year.f := as.factor(year)][,region := "ebs"][,season := "NA"]

#shelf
shelf_distances_dissimilarities_allyears_jaccard_rounded_allcombos <- readRDS(here::here("output","distance_decay","shelf_distances_dissimilarities_allyears_jaccard_rounded_allcombos.rds"))

shelf_distances_dissimilarities_allyears_jaccard_rounded_allcombos[,year.f := as.factor(year)][,region := "shelf"][,season := "NA"]


```


Merge 
```{r}
#any item in environment that ends with all combos
Pattern1<-grep("allcombos",names(.GlobalEnv),value=TRUE)
distances_dissimilarities_allyears_jaccard_rounded_allcombos <- do.call(rbind,mget(Pattern1[-2]))
     
#merge region season

distances_dissimilarities_allyears_jaccard_rounded_allcombos[,region_season := paste0(region,"_",season)]
```

Load region statistics
```{r load region stats}
gmex_stats <- readRDS(here::here("output","region_stats","gmex_stats.rds"))
swc_stats <- readRDS(here::here("output","region_stats","swc_stats.rds"))
ns_stats <- readRDS(here::here("output","region_stats","ns_stats.rds"))
pt_stats <- readRDS(here::here("output","region_stats","pt_stats.rds"))
ebs_stats <- readRDS(here::here("output","region_stats","ebs_stats.rds"))
ebs_stats[,region := "ebs"]
neus_stats <- readRDS(here::here("output","region_stats","neus_stats.rds"))
shelf_stats <- readRDS(here::here("output","region_stats","shelf_stats.rds"))


#any item in environment that ends with all combos

region_stats <- rbind(gmex_stats, swc_stats, ns_stats, pt_stats, ebs_stats, neus_stats,shelf_stats)

region_stats[,spp_num := as.numeric(spp_num)][,study_period := as.numeric(study_period)][,study_duration := as.numeric(study_duration)][,lat_range := as.numeric(lat_range)][,mid_lat := as.numeric(mid_lat)][,lon_range := as.numeric(lon_range)][,area := as.numeric(area)][,depth_range := as.numeric(depth_range)][,mid_depth := as.numeric(mid_depth)][,jac_change := ifelse(jac_coef > 0 & jac_p < 0.05, 1, ifelse(jac_coef < 0 & jac_p < 0.05,  -1,0))][,bray_change := ifelse(bray_coef > 0 & bray.p < 0.05, 1, ifelse(bray_coef < 0 & bray.p < 0.05,  -1,0))][,region_season := paste0(region,"_",season)]
```


```{r ggplot2}
ggplot(data = distances_dissimilarities_allyears_jaccard_rounded_allcombos,aes(x = as.numeric(year), y = jaccard_similarity_turnover, color = region_season)) +
  geom_smooth(se = F, method = "lm") +
  labs(x = "Year",y = "Jaccard Similarity") +
  theme_classic()

ggplot(data = distances_dissimilarities_allyears_jaccard_rounded_allcombos,aes(x = as.numeric(year), y = bray_curtis_similarity_balanced, color = region_season)) +
  geom_smooth(se = F, method = "lm") +
  labs(x = "Year",y = "Bray Similarity") +
  theme_classic()
```

Now, how do variables such as:
* Number of species in region
* Length of study period
* Number of years in which data was collected
* Range of latitudes sampled
* Mid latitude of region
* Range of longitudes sampled
* Overall area sampled
* Depths sampled

Come into play with the trend we identify? 

Inputs = these variables, colors = coefficient value (if significant)

```{r pca}
library(factoextra)
region_stats.descriptive <- data.frame(region_stats[,3:9])
rownames(region_stats.descriptive) <- region_stats$region_season
region_stats.pca <- prcomp(region_stats.descriptive, center = TRUE,scale. = TRUE)

fviz_pca_ind(region_stats.pca,
             repel = TRUE     # Avoid text overlapping
             )

fviz_pca_biplot(region_stats.pca, repel = TRUE,
                col.var = "#2E9FDF", # Variables color
                col.ind = "#696969"  # Individuals color
                )

summary(CaS.pca)

load.pca <- CaS.pca$loadings
print(load.pca)

fviz_pca_var(CaS.pca,
             col.var = "contrib", # Color by contributions to the PC
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
             repel = TRUE,     # Avoid text overlapping
             geom = c("arrow", "text")
             )

fviz_pca(region_stats.pca, label = c("var","ind"), repel = T, habillage = region_stats$region_season, ggtheme = theme_classic())
fviz_pca(region_stats.pca, label = c("var","ind"), repel = T, habillage = region_stats$jac_change, ggtheme = theme_classic())
fviz_pca(region_stats.pca, label = c("var","ind"), repel = T, habillage = region_stats$bray_change, ggtheme = theme_classic())


```

